### 2016-02-05 doc/toy.tex
the thing runs with reasonable-looking results.

\documentclass[preprint,3p]{elsarticle}
%\documentclass[preprint]{aastex}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\DeclareMathOperator\erf{erf}
\newcommand{\mli}[1]{\mathit{#1}}
%\usepackage{epstopdf}

\begin{document}

%\begin{frontmatter}
%
%\title{Toy Model}
%\author{A.~G. Kim\corref{cor1}}
%\ead{agkim@lbl.gov}
%\address{Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley CA, USA 94720}
%
%\begin{abstract}
%\end{abstract}
%\begin{keyword}
%\end{keyword}
%\end{frontmatter}

\section{Likelihood}
\label{likelihood:sec}

The ambition is to construct a model to determine the likelihood of DES-like
supernova data.
The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{../results//nodes_pgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   \label{pgm:fig}}
\end{figure}

In the toy model there are three types of data that
may be observed for each supernova:
\begin{itemize}
\item $c_o$: Counts at peak brightness.
\item $z_o$: Redshift.
\item $T_o$: Transient type.
\end{itemize}

There are selection criteria $S_c$ and $S_T$ for the inclusion of an object in the
sample and for a successul typing.
In what follows we assume that the typed objects are drawn from the same underlying
distribution as the sample $S_c=S_T$, and that the selection criterion is a brightness
threshold $S_c = c_o \ge c_T$.  (The spectroscopic sample $S_T$ 
eventually should depend on the
photometry and the quality of the spectroscopic data.)

For this preliminary discussion the top-level parameters are denoted by $X$,
latent parameters $L$ and $T$.

It is useful to split the data into two cases depending on whether or not
the transient has a type measurement $T_o$.

\subsection{Case Typed}
For Case Typed the likelihood is
\begin{align}
\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(T_o,z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
\end{align}
We assume that the observations give a perfect measurement of type and redshift,
so that
\begin{align}
p(T_o|T) & =\delta(T_o-T)\\
p(z_o|z) & =\delta(z_o-z);
\end{align}
practically $T$ and $z$ are no longer probabilistic but are fixed.
Then
\begin{align}
\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
&= \int dL \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
\frac{p(L, S_c, S_T | T=T_o, z=z_o, X)}{P(S_c, S_T| T=T_o,  z=z_o, X)}
\frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
&= \frac{\int dL p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}
\end{align}

The normalization term is determined by
\begin{align}
P(S_c, S_T| z=z_o, X) & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
& = \sum_i  P(T_i|z=z_o, X)\int dL p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right]
\end{align}
as $p(c_o | T=T_o, L, z)=0$ when $c_o < c_T$.

\subsection{Case Not Typed}

For Case Not Typed the likelihood is
\begin{align}
\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
\end{align}

For the redshift measurement, we take a set of potential host galaxies, where galaxy $j$
are redshift $z_j$ has probability $p_j$ of being the host
\begin{equation}
p(z_o|S_c, S_T, z, X) = \sum_j p_j\delta(z_j-z).
\end{equation}
Then
\begin{align}
\mathcal{L}(z_o,c_o | S_c, z, X) & =  \sum_j p_j \int dL \sum_{i} p(c_o | T_i, L, S_c, z=z_j, X) p(L |  T_i,  S_c,  z=z_j, X) P(T_i|S_c, z=z_j, X) \\
&=  \frac{\sum_j p_j \int dL p(c_o, S_c | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)},
\end{align}
where we use the fact that the counts do not directly depend on type.

The normalization term is
\begin{align}
P(S_c| z=z_o, X) & = \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL  p(L|T_i, z=z_j, X) 
\left[ \int_{c_T}^\infty dc_o p(c_o| T=T_i, L, z=z_j, X) \right].
\end{align}

\section{Model}
\label{model:sec}

We now go into more detail into the model and the form of the pdf's that enter
into the likelihoods described in \S\ref{likelihood:sec}.

A description of the nodes follow.  Numbers are illustrative.

\subsection{Cosmology}
The Flat $w$CDM cosmology.

\begin{align}
\Omega_M & \sim \ln{\mathcal{N}}(\ln{0.28},0.1)\\
w_0 & \sim \mathcal{N}(-1,0.05)
\end{align}

\subsection{Calibration}
Photometric zeropoints for each band.
\begin{equation}
Z \sim \mathcal{N}(0,0.02)
\end{equation}

\subsection{SNIa Rate}
Type~Ia SN rate. In the analysis only relative rates between different transients are important.
We take the rates relative to the SN~Ia rate
\begin{equation}
r_{SNIa} = 1.
\end{equation}

\subsection{SNII Rate}
Type~II SN rate, relative to the SN~Ia rate
\begin{equation}
r_{SNII} \sim \mathcal{U}(0.25, 4).
\end{equation}

\subsection{SNIa Luminosity}
The mean luminosity and intrinsic dispersion for SNe~Ia. In fact work with log-luminosity, intrinsic
dispersion in mag units.
\begin{align}
\ln{L}_{SNIa} & \sim \mathcal{N}(\ln{1}, 0.02) \\
\sigma_{SNIa} & \sim \ln{\mathcal{N}}(\ln{0.1},0.1).
\end{align}

\subsection{SNII Luminosity}
The mean luminosity and intrinsic dispersion for SNe~II. In fact work with log-luminosity, intrinsic
dispersion in mag units.
\begin{align}
\ln{L}_{SNII} & \sim \mathcal{N}(\ln{0.5}, 0.02) \\
\sigma_{SNII} & \sim \ln{\mathcal{N}}(\ln{0.4},0.1).
\end{align}

\subsection{SNIa~Probability}
Probability an object is SN~Ia
\begin{align}
p_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
p_{SNII}&=1-p_{SNIa}.
\label{prob:eqn}
\end{align}
\subsection{Type}
The type of the object $T$ is drawn from the Bernoulli distribution 
\begin{equation}
T | p_{SNIa} \sim \mathcal{B}(1-p_{SNIa}),
\end{equation}
with the convention that $T=0$ corresponds to SN~Ia.

The type probability appears explicitly in the calculation of the likelihoods.
\begin{align}
P(T_{SNIa} | z=z_j, X) &= p_{SNIa}\\
P(T_{SNII} | z=z_j, X) &= p_{SNII}
\end{align}
\subsection{Observed Type}
The observed type is assumed to be perfect.
\begin{equation}
T_o = T.
\end{equation}

\subsection{Luminosity}
For our model
\begin{equation}
\ln{L} | T, \ln{L}_{T},\sigma_{T} \sim \mathcal{N}\left( \ln{L}_{T},\frac{\ln{10}}{2.5}\sigma_{T}\right).
\end{equation}

The corresponding pdf appears in the likelihood.

\subsection{Redshift}
The redshift is drawn from a flat distribution
\begin{equation}
z\sim \mathcal{U}(0,\infty).
\end{equation}
%
%\subsection{Neighbor Galaxy Redshifts}
%Each neighbor redshift is drawn from a flat distribution.  The redshift for neighbor
%$i$ is  
%\begin{equation}
%z_{N,i} \sim \mathcal{U}(0,\infty).
%\end{equation}

\subsection{Observed Redshift}
In our model the measured redshift is a sum of delta functions
\begin{equation}
p(z_o|z) = \sum_i p_i \delta(z-z_i),
\end{equation}
where galaxy $i$ at
redshift $z_i$ has probability $p_i$ of being
the host.

\subsection{Counts}
The luminosity distance comes from the physics-theory function
\begin{equation}
d_L = d_L(z; \Omega_M, w0),
\end{equation}
the flux is
\begin{equation}
f = \frac{L}{4\pi d_L^2},
\end{equation}
and expected counts is
\begin{equation}
c_\star = 10^{Z/2.5}f.
\end{equation}
The distribution of counts for all transients is 
\begin{equation}
c_o | c_\star, \sigma_c \sim \mathcal{N}(c_\star, \sigma_c),
\end{equation}
where $\sigma_c$ is the measurement noise.

\subsection{Sample Selection Normalization}
Anticipating the use of available statistical packages, the likelihoods have been constructed
from standard distribution functions.  The truncated distributions from sample selection
are accounted by the normalization 
terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
in the PGM, they are treated as nodes who contribute 
$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
factor.

\subsection{Implementation Notes}
The following integral appears in the normalization
\begin{equation}
\int dL p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
\end{equation}
We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
\begin{equation}
n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
\end{equation}
the integral becomes
 \begin{equation}
\sigma_X \int dn p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
\end{equation}
The two probability distributions are normal so that this reduces to
 \begin{equation}
\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
\end{equation}
where
\begin{equation}
c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
\end{equation}
\end{document}

### 2016-02-05 src/Nemcee.py

@@ -13,13 +13,18 @@ import theano.tensor as T
 
 import matplotlib.pyplot as plt
 
-magtoflux  = numpy.exp(numpy.log(10)/2.5)
+ln10_2p5 = numpy.log(10)/2.5
+magtoflux  = numpy.exp(ln10_2p5)
+
+params = [0.28,-1, 0.5,numpy.log(1.), 0.1, numpy.log(0.5), 0.4,0]
+
+fluxthreshold = 0.4e-8
 
 def luminosity_distance(z, Om0, w0):
     cosmo = FlatwCDM(H0=72, Om0=Om0, w0=w0)
     return cosmo.luminosity_distance(z).value
 
-def normalization_integrand(n, lnL0, sigma_lnL, Z, threshold, ld):
+def normalization_integrand(n, lnL0, sigma_lnL, Z, threshold, ld, dcounts):
     #   The L value being sampled:
     #   lnL = n*sigma_lnL+lnL0
     #   exp(lnL) = exp(n)**sigma_lnL * exp(lnL0)
@@ -35,15 +40,14 @@ def normalization_integrand(n, lnL0, sigma_lnL, Z, threshold, ld):
     #   counts = 1/4/pi exp(lnL)/ld**2 [exp(ln10/2.5)]**Z
     #   counts = 1/4/pi exp(n)**sigma_lnL * exp(lnL0)/ld**2 [exp(ln10/2.5)]**Z
 
-    expn = numpy.exp(n)
-    counts = (1./4/numpy.pi)* numpy.power(expn, sigma_lnL) * numpy.exp(lnL0) / ld**2 * numpy.power(magtoflux,Z)
-#    ccdf  = 0.5 - 0.5*T.erf((threshold-counts)/ (numpy.sqrt(2.)*1e-9))
-#    return numpy.exp(-n**2/2.)*ccdf/sigma_lnL
+    # the gaussian on logL (leaving out the sqrt 2pi)
 
-    factor =  numpy.exp(-n**2/2.)*0.5 / sigma_lnL
-    divisor = numpy.sqrt(2.)*1e-9
-    return factor - factor*scipy.special.erf(threshold/divisor-counts/divisor) # return the incorrect answer since the integral multiplies by sigma_lnL
-
+    firstterm  = numpy.exp(-n**2/2) #the sigma_lnL here cancels with the normalization factor
+    L = numpy.exp(lnL0+n*sigma_lnL)
[…]

### 2016-02-17 doc/toy.tex
add multiple photometric points and covariance. Implement "N choose k" selection criteria.  Runs and OK-looking results on laptop

@@
 \begin{align}
-\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
-&= \int dL \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
+\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL\, p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
+&= \int dL\, \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
 \frac{p(L, S_c, S_T | T=T_o, z=z_o, X)}{P(S_c, S_T| T=T_o,  z=z_o, X)}
 \frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
-&= \frac{\int dL p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}
+&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}\\
+&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
 \end{align}
 
-The normalization term is determined by
+The  normalization term sums/integrates over all possible observed types/counts 
 \begin{align}
-P(S_c, S_T| z=z_o, X) & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
-& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& = \sum_i  P(T_i|z=z_o, X)\int dL p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right]
+P(S_c, S_T| z=z_o, X) & =\sum_i \int dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
+% & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
+%& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
+& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right].
 \end{align}
-as $p(c_o | T=T_o, L, z)=0$ when $c_o < c_T$.
+
+In the case of a single count measurement per transient and neglecting the typing selection criteria,
+\begin{align}
+ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
+ &= \begin{cases}
+   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
+   0 & \text{if }  c_o < c_T
+ \end{cases}
+\end{align}
+so that the integral in brackets is
+\begin{equation}
+\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
+\end{equation}
+
+For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
+noise of a single observation is expected to be dominated by its own photon noise making this
+supposition a good approximation.} and $k$ of those must be above threshold in order to be
+included in the sample.
+Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
+of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
+Then
+\begin{align}
+p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
+ &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
+   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } c_{o,\alpha} \ge c_T == g_\alpha \\
+   0 & \text{if }  c_{o,\alpha} \ge c_T\  != g_\alpha
+ \end{cases}
+\end{align}
+and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
+of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
 
@@
 \begin{align}
-\mathcal{L}(z_o,c_o | S_c, z, X) & =  \sum_j p_j \int dL \sum_{i} p(c_o | T_i, L, S_c, z=z_j, X) p(L |  T_i,  S_c,  z=z_j, X) P(T_i|S_c, z=z_j, X) \\
-&=  \frac{\sum_j p_j \int dL p(c_o, S_c | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)},
+\mathcal{L}(z_o,c_o | S_c, z, X) & =  \sum_j p_j \int dL \,\sum_{i} p(c_o | T_i, L, S_c, z=z_j, X) p(L |  T_i,  S_c,  z=z_j, X) P(T_i|S_c, z=z_j, X) \\
+&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)}\\
+&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)},
 \end{align}
@@
 \begin{align}
-P(S_c| z=z_o, X) & = \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL  p(L|T_i, z=z_j, X) 
-\left[ \int_{c_T}^\infty dc_o p(c_o| T=T_i, L, z=z_j, X) \right].
+P(S_c| z=z_o, X) & = \int dc_o \, \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
+& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL \, p(L|T_i, z=z_j, X) 
+\left[ \int dc_o \, p(c_o, S_c| T=T_i, L, z=z_j, X) \right].
 \end{align}

### 2016-02-17 src/Nemcee.py

@@ -65,6 +65,7 @@ def lognormal_logp(value, mu=None, tau=None):
 def normal_logp(value, mu=None, tau=None):
     return (-tau * (value - mu)**2 + numpy.log(tau / numpy.pi / 2.)) / 2
 
+
 def LogLuminosityMarginalizedOverType_logp(value, mus=None, taus=None, p=None):
     r"""The distribution for the luminosity marginalized over two kinds
     of astronomical sources:
@@ -127,7 +128,7 @@ def LogLuminosityGivenSpectype_logp(value, mu=None, tau=None, p=None):
     # print 'LLGST', numpy.log(p) + normal_logp(value, mu, tau)
     return numpy.log(p) + normal_logp(value, mu, tau)
 
-def Counts_logp(value, fluxes=None, pzs = None, Z=None, csigma=None):
+def Counts_logp(value, fluxes=None, pzs = None, Z=None, counts_invcov=None):
     r"""The distribution for the joint spectype and log-luminosity
 
     pdf of counts given a threshold
@@ -142,27 +143,49 @@ def Counts_logp(value, fluxes=None, pzs = None, Z=None, csigma=None):
         minimum counts in order to be discovered
 
     """
-    ans=0.
     logterms = []
     for index in xrange(len(pzs)):
         flux = fluxes[index]
         counts  = flux*magtoflux**Z
-        tau = csigma**(-2)
-        logterms.append(normal_logp(value, counts, tau))
+
+        # implement diagonal covariance matrix for now
+        ans = 0.
+        for i in xrange(len(value)):
+            tau = counts_invcov[i,i]
+            ans += normal_logp(value[i], counts, tau)
+        logterms.append(ans)
     logterms = numpy.array(logterms)
     logterm_max = numpy.max(logterms)
     logterms = pzs * numpy.exp(logterms-logterm_max)
     # print 'C',logterm_max + numpy.log(logterms.sum())
     return logterm_max + numpy.log(logterms.sum())
 
-    #ans = ans + pzs[index] *numpy.exp(normal_logp(value, counts, tau))
 
-    #need to add threshold
+"""
+A generator of all combinations of observation success given a minimum number of successes. 
+"""     
+import itertools
+# def sampleCombinations(obs_list, ntrue):
+#   num = ntrue
+#   while num <= len(obs_list):
+#       for a in itertools.combinations(obs_list,num):
[…]

### 2016-02-23 doc/article.tex
rewrite of toy note and cleaning up of some hanging changes

\documentclass[preprint,3p]{elsarticle}
%\documentclass[preprint]{aastex}
\usepackage{aas_macros}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{hyperref}

\newcommand{\mli}[1]{\mathit{#1}}
%\usepackage{epstopdf}

\begin{document}

\begin{frontmatter}

\title{Standard Candle Cosmology With Incomplete Spectroscopic Classification}
\author{A.~G. Kim\corref{cor1}}
\ead{agkim@lbl.gov}
\address{Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley CA, USA 94720}

\begin{abstract}
\end{abstract}
\begin{keyword}
\end{keyword}
\end{frontmatter}

\section{Introduction}
A class of astronomical objects who share a common luminosity are called a standard
candle.  The relative fluxes (or magnitudes) for a set of these objects distributed in space
provide a measurement of their relative luminosity distances. These data, combined with
measurements of the standard candle redshifts, constitute a Hubble
diagram which is used to map the expansion history of the Universe.

Type~Ia supernovae (SNe~Ia) are an example of a standard candle (more precisely a standardizable
candle), whose redshift-magnitude relation provides an accurate expansion history
used to measure the Hubble Constant
\citep{2001ApJ...553...47F} and detect the accelerated expansion of the Universe
\citep{1998AJ....116.1009R, 1999ApJ...517..565P}.  SN~Ia measurements continue
to improve 
\citep{2014A&A...568A..22B} and play a critical role in probing the physics
responsible for the acceleration \citep{2013PhR...530...87W}.
Looking at the present and toward the future, SN~Ia cosmology is
a component of the Dark Energy Survey\footnote{\url{http://www.darkenergysurvey.org}} (DES),
and a science driver in the design of the
Large Synoptic Survey Telescope\footnote{\url{http://www.lsst.org}} (LSST)
and the Wide-Field InfraRed Survey Telescope
\citep{2015arXiv150303757S}.

The information necessary to construct a Hubble diagram consists of, per object,
its classification as a
member of the standard candle class, its redshift, and its flux.  For SNe~Ia, the flux at
peak brightness is typically determined from multi-band light curves obtained from repeated
photometric observations. Traditionally, spectroscopic data is used to classify the transient
as SN~Ia through the lack of hydrogen and the presence of SiII in its spectrum,
and determine its redshift either directly through the transient spectrum or that of its host
galaxy. While the classification and redshift could be determined from the multi-band
light curves and colors of the host galaxy, the accuracy is insufficient for precision
cosmology \citep{2011ApJ...738..162S}.  Consequently spectroscopic follow-up
has played a critical role in cosmological supernova surveys.

While the large focal planes of the Dark Energy Camera (3 square degrees)
and the LSST (9.6 square degrees) open the
possibility for the photometric discovery and observation of thousands to millions
of SNe~Ia out to $z\sim1$, a corresponding multiplex advantage has not been
achieved for the 10-m
class telescopes used for the spectroscopic classification of faint discoveries:
it is unfeasible to obtain a spectrum for all SN~Ia candidates with well-measured
photometry.

Photometric supernova analysis has been proposed to  take advantage of the statistical weight imported by large numbers of supernova light curves, despite the lack of spectroscopy.
An example of such a program would proceed as follows: A rolling wide-field imaging
survey generates light curves for all transients in the survey area.  In near real-time,
a subset 
of transients from those images are selected for triggered spectroscopic observations
to obtain classifications and redshifts.  At the end of the survey, the
analysis sample is the subset of likely SN~Ia
candidates based on photometric data;  the analysis sample and its subset
with triggered spectroscopy are constructed so as to be drawn from the same parent distribution.
Spectroscopic redshifts are obtained for likely host galaxies.  The Hubble diagram of the
analysis sample is analyzed with the understanding that it may contain multiple populations.

The Sloan Digital Sky Survey-II Supernova Survey is an example of a program
that produced significant numbers of transients
with light curves but without spectroscopic typing \citep{2014arXiv1401.3317S}:
photometric SN analysis has been developed applied to these data \citep{2012ApJ...752...79H,
2013ApJ...763...88C}.   Such analyses are anticipated in the projected
supernova programs of 
DES \citep{2012ApJ...753..152B} and LSST \citep{2012arXiv1211.0310L}.

Current models used for supernova cosmology analysis account for unobserved quantities.
SNe~Ia are not standard but are ``standardizable'' candles, implying that the class may be composed of distinct subpopulations each with its own redshift-dependent rates, magnitude distributions, and secondary
observable (e.g.\ light-curve shape, color) distributions.  Bayesian hierarchical modeling is 
used to account for this extra layer of complexity  to determine characteristics of the SN~Ia population simultaneously
with the cosmology \citep{2011MNRAS.418.2308M,
2015arXiv150701602R}.  Consideration of missing classification is addressed
by
BEAMS \citep{2007PhRvD..75j3508K}, which  includes a non-Ia population in sample, whose luminosity distribution
is inferred along with the cosmology:
BEAMS was used in the analysis of SDSS-II SN data \citep{2012ApJ...752...79H}.

This article addresses the affect of missing spectroscopic typing in
a transient standard candle analysis, with the eventual goal of quantifying
requirements for the spectroscopic subsample in a broader photometric analysis.
The lack of spectroscopy leads to ambiguity in both the transient class and
in the host galaxy due to projected galaxies that appear close to the transient or
the relative low-surface brightness of the host.  The subsample with
spectroscopic typing informs the interpretation of objects lacking spectroscopy.
Differences between observed and
intrinsic distributions in a flux-limited sample can have important effects on the Hubble
diagram  (e.g.\ Malmquist bias) so are accounted for explicitly.

\section{DES SN Survey Model}
\label{snIamodel:sec}
\subsection{Model}
\label{model:sec}
This Section presents a candidate model
for supernova surveys that has the fidelity needed for the photometric
supernova analysis of the Dark Energy Survey and the Large Synoptic Survey Telescope.
This model captures the current treatment in current analyses and is expanded to include
photometry-only transients.  This may not be the ultimate model used, indeed many models
need to be considered for model testing.

The Probabilistic Graphical Model is shown in Figure~\ref{pgm:fig}. The underlying behavior of the universe is described by the following elements:
\begin{description}
\item [Cosmology] Luminosity distance for the cosmological model being considered,
e.g.\ a $wCDM$ universe parameterized by $\Omega_M$, $\Omega_w$, $w_0$, $w_a$.
\item[SNe~Ia, Non-Ia Populations] Time-dependent luminosity for all transient populations
that may enter
the analysis.  The SALT2 model is a candidate for SNe~Ia, with parameters $\alpha$, $\beta$.  The PSNID treatment of non-Ia's can be made into a model for the other
transients or Natahsa Karpenka is working on a model of non-Type~II core collapse SNe.
\item [Relative Rates] Relative rates of all the transients populations and
subpopulations (e.g. SALT2 $x_1$, $c$) under consideration.  The
model could of been extended to measure absolute rates, but that is unnecessary if
the cosmological parameters are the goal.
\end{description}
Each transient is described by the following elements:
\begin{description}
\item[Host Galaxy] The host galaxy, or more generally for the case of hostless supernovae,
the environment from which the transient arises.  The host is described by its redshift
and other parameters that may be of interest such as mass,
specific star formation rate, metallicity.
\item[Type Subtype] Specification of the the transient: the population to which it belongs
and its internal parameters.  For example, an object could be SN~Ia with specific
values for its SALT2 parameters $c$, $x_1$, $t_0$ and its random dispersion
realization (both absolute magnitude and from the error snake). 
This is dependent on the relative rates and
its host galaxy including its redshift.
\item[Distance] Distance modulus or equivalently the luminosity distance of the host galaxy,
which in our model corresponds to the cosmological redshift of the transient.  The value
is fixed based on the cosmology and redshift.
\end{description}
The individual photometric measurements of a transient are described by the following
elements:
\begin{description}
\item[Luminosity]  Luminosity as a function of wavelength
corresponding to the date of observation.   This depends on the population, type,
and subtype.  Given that searching for correlations between  host properties and Hubble
residuals is a good probe of systematics,
a direct connection between host galaxy and luminosity is made.
\item[Flux] Flux of the transient incident to earth at the date of observation.  It is 
fixed based on the luminosity, redshift, and distance.
\end{description}
The observatory is described by the following elements:
\begin{description}
\item[Throughput] The transmission function experienced by the transient during
the observation.  It depends on band, position on the focal plane, and date of observation.
\end{description}
The data (subscripted by ``o'' for observation)  are:
\begin{description}
\item[$\text{Photometry}_o$] Photometric flux measurement of the transient.
The expected measurement
is the incident flux integrated over the throughput.
\item[$\text{Type Subtype}_o$] Spectroscopic
measurements of the intrinsic transient type.  It is dependent on the underlying
type.
\item[$\text{Redshift}_o$] Measurement of the host redshift.  The redshift can be
spectroscopic or photometric.  It is dependent on the host redshift, and
redshifts of any other galaxies that may be confused as the host (not shown explicitly
in the PGM).
\item[$\text{Host Properties}_o$] Measurement of the host properties such
as mass, sSFR, metallicity. It is dependent on the host galaxy, and
any other galaxies that may be confused as the host.
\end{description}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=6.5in]{../results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The description of the elements and dependencies of the model are given in
   \S\ref{model:sec}.
   \label{pgm:fig}}
\end{figure}

This hierarchical model has several features worth noting.  Transients with and without
spectroscopic classification are treated together: those with classification help constrain
the properties of the underlying population to inform the treatment of those without.
Photometric classification is treated internally, through the
latent Type.  The system throughput is treated as a probabilistic variable
allowing for self-calibration; this could be omitted from the model with calibration
and its uncertainties applied
to the Photometry measurements.  Host-galaxy misclassification enters as measurement
uncertainties.  Malmquist bias and other selection effects do not appear explicitly, but are accounted
for in the likelihoods.  The model could be extended to include summary statistics,
for example the results of an external supernova classifier; the use
of summary statistics would affect the manner in which the model is evaluated
due to our inability to write their likelihoods.

\subsection{Sample Selection}
The transients that enter the analysis satisfy some selection criteria.  Trivially, they
have to be discovered in the first place and often more stringent photometric signal-to-noise
requirements are applied.  Additionally, transients that have
successful spectroscopy may be drawn from an underlying population that differs from the
underlying or photometrically-selected samples.

Sample selection criteria for DES have not yet been finalized
(Argonne Eve for photometry, OzDES/Ryan for spectroscopy are candidates to do this);
presented here is a sketch of the likelihood calculation for a reasonable
photometric selection criterion:
for each transient $\nu$, the criterion is to have $\ge N$ photometric points
above a signal-to-noise threshold given $M_\nu$ available measurements.
Each photometric point is associated with a flux threshold that it either passes
or fails.  Given a total $M=\sum_\nu M_\nu$ data points, there are
$2^M$ possible configurations
of the individual points passing or failing theshold, of which only
$\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}$  configurations lead to
the inclusion of the transients in the sample.

The likelihood for the analysis
sample is the conditional distribution obtained by applying the selection criteria
on the domain of the likelihood given no selection criteria.  This latter likelihood
can be written
as the sum of the contributions from all possible threshold-passing configurations
\begin{equation}
p(data|model) = \sum_{k=1}^{2^M} p(data, \hat{b}_k| model),
\label{likelihood_u:eqn}
\end{equation}
where $\hat{b}$ is the set containing all possible configurations.
Let $b$ be the subset of $\hat{b}$ 
that satisfies sample selection. Then the selection criteria is $\mathcal{S} = \hat{b}_k
\in b$.
The likelihood of the analysis sample is proportional to Eqn.~\ref{likelihood_u:eqn}
with failing configurations removed
\begin{equation}
\mathcal{L}(data|\mathcal{S}, model) =
\sum_{k=1}^{2^M} p(data, \hat{b}_k| \hat{b}_k
\in b, model) \propto \sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} p(data, b_k| model),
\end{equation}

The probability for the data to be in a specific configuration $b_k$ of
being below or above flux threshold is expressed
as
\begin{equation}
p(data, b_k| model)  = \epsilon(Photometry_o, b_k)p(data| model)\\
\end{equation}
where
\begin{equation}
\epsilon(Photometry_o; b_{k})= \prod_{\nu i}\mathcal{H}\left((-1)^{b_{k\nu i}+1}(Photometry_{o,k\nu i}-Threshold_{\nu i})\right),
\label{eff:eqn}
\end{equation}
where $\mathcal{H}$ is the Heaviside step function; when $b_{k\nu i}=1/0$ the step in the efficiency
is up/down.  The indices correspond to configuration $k$,
transient $\nu$, and photometric point $i$.

Assembling this together, the likelihood is
\begin{align}
\mathcal{L}(data|\mathcal{S}, model)
% & = \frac{\sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} \epsilon(Photometry_o, b_k)p(data| model)}{\int ddata'\,
%\sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} \epsilon(Photometry_o', b_k)p(data'| model)
%}\\
& = \frac{p(data| model)}{\int ddata'\,
\sum_{k=1}^{\prod_\nu \sum_{k=N}^{M_{\nu}} \binom{M_{\nu}}{k}} \epsilon(Photometry_o', b_k)p(data'| model)
},
\end{align}
where the normalizing factor is included explicitly as it has 
parameter dependence and must not be neglected in calculations.
The numerator is $\le 1$, where the equality occurs when there are no selection criteria.
If $p(data|model)$ is a multi-dimensional normal distribution, the numerator is
the sum of error functions.

Consider the traditional case of Malmquist bias of a single object,
where a single measurement has
to be above flux threshold.  For the underlying distribution with no flux threshold there
are $2^1=2$ possible configurations, the flux is either above or below threshold.  For
the sample distribution, there is $\binom{1}{1}=1$ possible configuration: the flux
is above threshold.   The likelihood for the measured flux is proportional to the original
distribution truncated so that fluxes below threshold are assigned zero probability.

For the case of a single object with two measurements of which only one is needed to
pass threshold,
there are $2^2=4$ possible configurations of which $\sum_{k=1}^{2} \binom{2}{k}=3$
allow sample selection, the exception being when both measurements fall below threshold.
The likelihood for each flux is composed from
the sum of three truncated distributions of which one accepts faint fluxes,
resulting in likelihoods that are not zero below detection threshold.

\subsection{Underlying Likelihood}
The likelihood of the analysis sample is calculated from the underlying pdf of data.  Based on Figure~\ref{pgm:fig},
\begin{align}
p(data|model) &=\int dType\, dLuminosity\, p(Photometry_o |  Luminosity, Cosmology,Host,Throughput)  \nonumber\\
& p(Luminosity| Type
, Host, Population) p(Type | Rates, Host)  \nonumber\\
& \times \int dType\, p(Type_o|Type) p(Type|Rates, Host) \nonumber \\
& \times p(Redshift_o, Host_o| Host),
\label{like:eqn}
\end{align}
where $data=\{Photometry_o, Type_o, Redshift_o, Host_o\}$,
$model=\{Cosmology, Populations, Rates, Throughput, Host\}$, and the independence
of photometric and spectroscopic typing measurements is anticipated.

For specificity candidate models for the pdf's that appear in Eqn.~\ref{like:eqn}
are now presented.
\begin{itemize}
\item The type is drawn from the host-dependent relative rate between SNe~Ia and non-Ia's 
\begin{equation}
Type | Rates, Host \sim {\rm Bernoulli}(\pi(z;r)),
\end{equation}
where $\pi$ is a model for the relative rate of SNe~Ia parameterized by $r$ as a
function of galaxy redshift $z$.  {\bf A reasonable choice for $\pi$ needs to be
selected.}

Type~Ia supernovae have values of SALT2 parameters
$x_1$ and $c$ drawn from Normal distributions
with parameters for the variance
\begin{equation}
x_1, c \sim \mathcal{N}(0, (\sigma_{x_1}, \sigma_c)).
\end{equation}

Models for non-Ia supernovae need to be selected.

Note that the rates may not have been important in spectroscopic analyses, but are important here to help infer the types/subtypes of transients without
a spectrum.
\item For the underlying SED of transients for SNe~Ia the SALT2 model
is an option, although new models are expected to appear shortly.  Absolute magnitudes
that depend on host-galaxy properties also need to be considered. 
In the SALT2 context,
\begin{equation}
Luminosity(Type
, Host, Population)= L_{SALT2}\left(\frac{t-t_0}{1+z}; \alpha, \beta, x_1, c\right) \exp{\left(\delta_G\mathcal{H}(g)\right)},
\end{equation}
where the first  term is the SALT2 light-curve and absolute magnitude
model with error snake.  Possible host-galaxy effects like the ``mass step'' are
accounted in $\mathcal{H}(g)$ where $g$ is the corresponding host parameter
and $\delta_G$ is the size of the offset.  The host redshift is $z$.
Note that for this model the parameter is fixed, though
generally this may not be the case.

{\bf The non-Ia model needs to be identified.}
\item The theoretical distance modulus is
\begin{equation}
\mu(Host, Cosmology) = \mu_{wCDM}(z; \Omega_M, w)
\end{equation}
is the predicted distance modulus of a flat dark energy cosmology
with parameters $\Omega_M$ and $w$, and host redshift $z$.

The flux incident on Earth is the redshifted and time-dilated SED scaled
by the distance modulus
\begin{equation}
f(Luminosity, Host, \mu) = 10^{-\mu/2.5} Luminosity(\lambda/(1+z)).
\end{equation}

The expected photometric signal in counts depends on the flux and throughput of the
optical chain.  A simple model encapsulates flux calibration through a zeropoint $Z$
and $T_0$  a fiducial transmission function for the appropriate filter.
More
sophisticated descriptions of calibration can vary the shape, not only the normalization, of the
transmission.  For the simple model,
\begin{equation}
counts(Throughput, f) = 10^{Z/2.5} \int  f(\lambda) T_0(\lambda)d\lambda.
\end{equation}

Practically the photometric extraction is handled by independent software, which calculates
the integrated counts with uncertainties $C_P$
by modeling and separating out polluting background light from the transient
signal
\begin{equation}
Photometry_o |  Luminosity, Cosmology,Host,Throughput \sim \mathcal{N}(counts,C_P).
\end{equation}
Dillon is providing this $Photometry_o$ and $C_P$.

An alternative model would be to remove Throughput
and include calibration uncertainties in $C_P$.  This would be a huge covariance
matrix!
\item Spectroscopic measurements are assumed
to provide prefect determinations of transient type
\begin{equation}
Type_o(Type)  = Type.
\end{equation}
\item There are measurements of the host galaxy properties.
For transients that have spectroscopic
confirmation, an unambiguous host determination is assumed such that
\begin{align}
Redshift_o(Host) & = z\\
Host~properties_o  | Host & \sim \mathcal{N}(p, C),
\end{align}
where $p$ are the Host properties being measured
and
$C$ their measurement uncertainty.

For the photometric sample, there is uncertainty in the identity of the host galaxy.
Suppose host-galaxy identification gives
a set of potential hosts $\hat{G}$ each with probability $\pi$ of being the correct host,
what Ravi is doing.
Then for the photometric sample
\begin{equation}
p(Redshift_o, Host~properties_o   | Host) = \sum_i \pi_i \delta(Redshift_o- z_i)\mathcal{N}(Host~properties_o-p_i, C)).
\end{equation}
Recall that DES plans to get spectroscopic redshifts with negligible measurement uncertainties on all
possible hosts.
\end{itemize}

This accounts for all the terms needed to assemble to pdf of the underlying distribution,
from which the likelihood of the analysis sample is assembled.

\section{Design}
\subsection{Data}
There is photometry, redshift, host properties, type/subtype.  The pdf's of redshift and
host-properties may be more complicated than a normal distribution due to uncertain host
determination and photometric redshifts.  The photometry is expected to be a Gaussian
with a full covariance matrix due to the host-galaxy subtraction.

\section{Tasks}
Here are a list of initial tasks and ingredients needed to proceed.
\subsection{Photometric Selection}
Here are desired properties of the set (or sets) of objects obtained from photometric selection.
\begin{itemize}
\item Sufficient data-quality to have a chance at photometric classification.
SNe~Ia do look different from other transients: the principal source of type  ambiguity
comes from poor signal-to-noise and poor temporal/wavelength sampling.
Having enough points with sufficient signal-to-noise over three adjacent bands over a 
sufficiently large time window.
\item A large set of likely SNe~Ia.
\item All objects with photometry-only must have spectroscopically-typed analogs, preferably though not necessarily at the same redshift.  SNe~Ia subpopulations whose characteristics
are unconstrained are not useful for cosmology.
\item Obvious non-Ia's from populations that could be misconstrued as SN~Ia.  For example,
a non-Ia at low redshift could be
mistaken to be SN~Ia at high redshift due to degraded signal.  
\item Exclusion of objects that are not part of populations that could be misconstrued as SN~Ia.
They would pollute inference of the non-Ia population.
\end{itemize}

\subsection{Detection Criteria (Malmquist Bias)}
The photometric selection criteria are expected to be significantly more rigorous than
the detection criteria.  Nevertheless, the detection efficiencies need to be characterized.
Fakes already provide a good measure of this.  To first order, the efficiency depends
on $S/N$, but a secondary dependence on the structure of background galaxies.
The detection efficiency is more complicated than the step function
shown in Eqn.~\ref{eff:eqn}.

Deliverable is the detection efficiency to take the place of Eqn.~\ref{eff:eqn}.

\subsection{Per-Exposure Thresholds}
To understand the effect of sample-selection criteria and detection efficiency
on a specific object, the
properties of each photometric point are needed.  Although the photometric criteria
have not been formalized, a $S/N$ threshold is a possible statistic to be used.
As such, a clean deliverable would be:
\begin{itemize}
\item Determination of flux thresholds for each photometric point of each candidate
given
a $S/N$
\begin{equation}
threshold(point, S/N).
\end{equation}
If we work in the sky-noise dominated regime, I believe this information is already
calculated by DM.
\end{itemize}

\subsection{Spectroscopic Selection}
Identification of selection criteria for spectroscopic success, as a function
of data.

\subsection{Relative Rates}
\label{rates:sec}
A rates model for transient types and subtypes
\begin{equation}
\pi(z;r).
\end{equation}
\subsection{SALT2}
Decision on how to model the distribution of $x_1$ and $c$.  For a first try a
single normal distribution seems sufficient but do we want to be more ambitious?
\subsection{Non-Ia Transient Model}
Needed is a model(s) for non-transients that could be mistaken as SN~Ia:
both a description of the time-evolving SED and the relative rates
with respect to Type~Ia's (see \S\ref{rates:sec}).

The SED model is of the time evolving luminosity (or more precisely $\mathcal{M}$)
for the range of possible non-Ia's.  It describes not only colors and light curve shapes,
but also absolute magnitude.  Its parameters can be constrained part of the
global analysis, indeed the pdf of the absolute magnitude is candidate.

As much of its usage is in classification, PSNID is a place to start.
Natasha Karpenka has been compiling a descriptions of non-Ia-Type~I's.  Can these
be packaged into the following deliverable?

The deliverable is a parameterized SED model that can be accessed through
an interface such as $L(t,\lambda; \text{Non-Ia parameters})$.

\subsection{Host-Galaxy Parameter Determination}
Although a different group within DES is charged with deriving properties
of galaxies (including our hosts), results do not seem to be forthcoming.
Mat has said that there is a group within the SNWG interested in
taking over this task.  Chris has discussed getting deep
co-adds of galaxies uncontaminated by supernova light.
Can we confirm who is part of this effort, if there is forward momentum,
and get a single name that used as the point of contact?

The deliverable are galaxy properties and a covariance matrix of their uncertainties,
which are treated in the analysis as data.

\subsection{NIR Data}
VISTA NIR data have important probative value that can enhance about all aspects
of our analysis even if available for a subset of objects.
We should decide our intents on the timing of the use of this data.
For example, on whether we plan on a DES-only paper and DES+NIR, and if
we want to focus efforts on the former in the short term.

\subsection{Photometric Data}

Deliverable:
transient counts and associated covariance matrix.  For flexibility, the counts and the
flux calibration are delivered separately.

\subsection{Calibration model and uncertainties}
For the first analysis, it is presumably sufficient to describe calibration uncertainties
as global zeropoints.  In the long term, it would be great to have a transmission
function for every photometric point.

It has to be decided whether calibration should be included
as part of the photometric uncertainties, or as model parameters.
The former requires the inversion of a full covariance matrix for all photometric
points.  The latter may have a per-SN solution through importance sampling.

Deliverable of $p(Z)$.

\subsection{Others}

\begin{itemize}
\item Data! Spectroscopic Type and uncertainties, Spectroscopic redshifts and
uncertainties, galaxy properties and uncertainties, host-galaxy associations
and uncertainties, photometry and uncertainties.
\item Determine the mechanics of doing the analysis.  A Bayesian MCMC could produce posteriors, which sampler that best suits our needs? Is a frequentist
analysis technically unfeasible? If redshifts were free to wander I would think so, but maybe
it is possible in this model.
\end{itemize}

Danny writes everything!
%\section{Misc}
%The simulation of data from the model presented in
%and their analysis is implemented in Python and are available
%on GitHub\footnote{{git@github.com:AlexGKim/abc.git}}.
%
%
%\section{Standard Candle}
%\label{toy:sec}
%To introduce basic model elements, this Section considers a
%simplified view of SNe~Ia
%as transients that are bolometric standard candles with an intrinsic random
%dispersion
%at peak brightness, and assumes data with negligible measurement uncertainties.
%For convenience,
%the transients
%in this Section are  referred to as SNe~Ia.
%The added complexity of correlations between absolute magnitude
%and multi-band light curves, and measurement uncertainties
%are discussed in
%\S\ref{snIamodel:sec}. 
%
%\subsection{Base Model}
%\label{model:sec}
%This Section presents a model that is used both for generating simulated data, and for analysis.
%The data used to construct a Hubble diagram and the fundamental parameters of the
%model used to describe those data are succinctly summarized in the notation
%for the likelihood 
%\begin{equation}
%p(f, {{T}}_S,{{z}}_S, G|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}}),
%\label{likelihood:eqn}
%\end{equation}
%whose relations are sketched in the probabilistic graphical model in Figure~\ref{toypgm:fig}.
%The measurements are
%\begin{itemize}
%\item $f$: the flux at peak brightness, taken 
%to be bolometric to facilitate the treatment of
%standard candles at different redshift;
%\item ${{T}}_S$, ${{z}}_S$: the spectroscopic type and redshift available for a subset of transients;
%\item $\theta_{G1}$, $\theta_{G2}$: the spectroscopic redshifts of the galaxies identified
%as host and neighbor from imaging.
%\end{itemize}
%The model parameters are
%\begin{itemize}
%\item $\Omega_M$, $w$: the mass density and dark-energy equation of state parameters of a flat,
%constant-$w$ dark-energy cosmology;
%\item $\theta_{r1}$, $\theta_{r2}$:  the parameters
%that describe the redshift-dependent relative fraction of SNe~Ia and non-Ia's;
%\item $\alpha_{Ia}$, $\sigma_{Ia}$, $\alpha_{\mathit{!Ia}}$, $\sigma_{\mathit{!Ia}}$:
%the luminosity (times $4\pi$) and intrinsic dispersion for SNe~Ia and non-Ia at peak brightness.
%\end{itemize}
%Not listed explicitly in Eqn.~\ref{likelihood:eqn} but shown in Figure~\ref{toypgm:fig} are latent model parameters for each transient:
%\begin{itemize}
%\item $z$: the cosmological redshifts of the host and neighboring galaxies;
%\item $T$: the class of the objects, SN~Ia or non-Ia;
%\item $L$: the luminosity of the transient.
%\end{itemize}
%There is a  fixed parameter
%\begin{itemize}
%\item $\mu$: the distance modulus given the cosmological parameters and the redshift.
%\end{itemize}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=4.5in]{/Users/akim/project/abc/results/toy_pgm.pdf} 
%   \caption{Probabilistic Graphical Model for the standard candle analysis. The
%   parameters include the cosmological parameters $\Omega_M$ and $w$,
%   the luminosities $\alpha$ and intrinsic dispersions $\sigma$ of SNe~Ia and non-Ia's,
%   and the relative fraction parameters $\theta_{r1}$ and $\theta_{r2}$.  Latent parameters
%   include 
%   for each transient the type $T$ and redshifts of the host and neighboring galaxies $z$ and $z_N$.  The  distance modulus $\mu$ and
%   luminosity $L$ are fixed by the other parameters.  The observables include the measured
%   flux $f$, spectroscopic type and redshift $T_S$ and $z_S$, and
%   the inferred host and neighbor redshifts $\theta_{G1}$ and $\theta_{G2}$.
%   The host and neighbor galaxies of all transients are inferred based a common galaxy catalog, 
%   and so strictly are not independent from supernova to supernova.
%   \label{toypgm:fig}}
%\end{figure}
%
%
%The transients being analyzed have passed some detection threshold
%and thus are drawn from
%distributions that have different shapes than the underlying population of both discovered and
%non-discovered objects. A  threshold flux of $f_0$ is adopted for the detection criterion.
%Otherwise, no other sample-selection criteria are assumed.
%
%\subsubsection{Likelihood for the {\it obs} set data}
%\label{obs:sec}
%The likelihood of data from the {\it obs} set, which has
%spectroscopic observations $T_S$ and redshift
%$z_S$, is described as follow.  The likelihood is
%\begin{align}
%p(f, {{T}}_S,{{z}}_S|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}}, z)  &=\sum_T
%p(f, {{T}}_S,{{z}}_S, T| \ldots)\\
%&= \sum_T 
%p(f, {{T}}_S,{{z}}_S| T,\dots) p(T | \ldots).
%\end{align}
%
%Direct associations between model and spectroscopic redshift and type are made
%\begin{align}
%p(z_S|z) &= \delta(z_S-z)\\
%p(T_S|T) &= \delta(T_S-T),
%\label{specz:eqn}
%\end{align}
%so that
%\begin{align}
%p(f, {{T}}_S,{{z}}_S|  \ldots) &= 
%p(f| T=T_S, z=z_s,\dots) p(T_S|z=z_s, \ldots).
%\label{obs:eqn}
%\end{align}
%
%
%The first term describes the expected counts. Given the redshift
%and transient type, the brightness of the underlying population
%is modeled as being drawn from a standard candle
%with mean luminosity $4\pi\alpha_X$ and a normal distribution in magnitude
%space with intrinsic dispersion  $\sigma_X$:
%\begin{equation}
%f| T=T_S, z=z_S, \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}} \sim \mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{T_S}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{T_S}\right),
%\label{flux:eqn}
%\end{equation}
%where $\mathcal{N}_{\ln}$ is the log-normal distribution and $d_L$ is the luminosity distance.
%Measurement uncertainty is taken to be negligible.
%In a flux-limited survey, the pdf's for the data are truncated versions of the underlying
%pdf's.
%\begin{itemize}
%\item For SNe~Ia  $T_S=1$:
%\begin{equation}
%f | T=1, z=z_S, \Omega_M, w, \alpha_{Ia},\sigma_{Ia} \sim
%\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{Ia}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{Ia}\right)}{P_{Ia}(f > f_0)}.
%\label{adusnIa:eqn}
%\end{equation}
%\item For non-Ia's in set $T_S=0$:
%\begin{equation}
%f | T=0, z=z_S, \Omega_M, w, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}}\sim 
%\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{\mathit{!Ia}}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{\mathit{!Ia}}\right)}{P_{!Ia}(f > f_0)},
%\label{adunonIa:eqn}
%\end{equation}
%\end{itemize}
%where the normalization factor is the complementary cumulative distribution function
%for the corresponding log-normal distribution that accounts for the truncated
%distribution of the detected sample.
%It is understood implicitly that as all objects must pass the detection threshold,
%the case of $f<f_0$ is never confronted in the analysis.
%
%The second term gives the type probability. In our model,
%the underlying relative rates between SNe~Ia and non-Ia are linear in redshift,
% parameterized by the
%its values at $z=0$ and $1.1 \times z_{max}$, 
%\begin{equation}
%\hat{\theta}_r=\theta_{r0}+z\left(\frac{\theta_{r1}-\theta_{r0}}{1.1 \times z_{max}}\right).
%\label{rate:eqn}
%\end{equation}
%The relative rate for SN~Ia discovery for the discovered (truncated) distribution is
%\begin{equation}
%\theta_r=\frac{\hat{\theta}_rP_{Ia}(f > f_0)}{\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{!Ia}(f > f_0)}.
%\end{equation}
%The spectroscopic classifications are drawn from
%\begin{equation}
%T_S | z=z_S, \Omega_M, w, \theta_{r0}, \theta_{r1} , \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}} \sim \text{Bernoulli}(\theta_r).
%\end{equation}
%Although the underlying type distribution depends only on $ \theta_{r0}$ and $ \theta_{r1}$,  
%the observed distribution does depend on the other parameters through $\theta_r$'s dependence on
%the two normalization factors
%$P_{X}(f > f_0)$.
%
%%The third term is the redshift distribution of transient discoveries.
%%The model takes an underlying redshift distribution of transients in the universe
%%\begin{equation}
%%p(z) \propto z^2.
%%\end{equation}
%%The redshift distribution of discovered transients is then
%%\begin{equation}
%%p(z) = \frac{z^2\left(\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{non-Ia}(f > f_0)
%%\right)}{\int_{z_{min}}^{z_{max}} dz'\, z'^2\left( 
%%\hat{\theta}_{r}P_{Ia}(f > f_0) + (1-\hat{\theta}_{r})P_{non-Ia}(f > f_0)
%%\right)}.
%%\end{equation}
%
%
%\subsubsection{Likelihood for the {\it mis} set data}
%
%Transients in the {\it mis}
%set have no spectroscopic confirmation, and their data are treated as follows:
%With no redshift from the transient itself, 
%the redshift measurement is inferred from potential host galaxies.
%In this Section, each transient is associated with two galaxies, one 
%with redshift $\theta_{G1}$
%that has high probability $\pi_{host}$
%of being the host and the other
%with redshift $\theta_{G2}$ that has probability $(1-\pi_{host})$.
%In the model, the
%true host redshift is $z$ and that of the interloping galaxy is $z_N$.
%The likelihood for the galaxy redshifts is
%\begin{equation}
%P(\theta_{G1},\theta_{G2}|z, z_N) =
%	\pi_{host}\delta(z-\theta_{G1})\delta(z_N-\theta_{G2}) +
%	(1-\pi_{host}) \delta(z-\theta_{G2})\delta(z_N-\theta_{G1}).
%\end{equation}
%As with the transients, the redshifts of the host
% and neighbor galaxies are drawn from underlying distributions $p(z) \propto z^2$, $p(z_N) \propto z^2_N$.
%
%
%The distribution of the underlying distribution is 
%\begin{multline}
%p(f, \theta_{G1}, \theta_{G2} | \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{!Ia}},\sigma_{\mathit{!Ia}}, z, z_N)  \\
%= \sum_{T} p(f, \theta_{G1}, \theta_{G2}| T, \ldots)P(T| \ldots)\\
% =\sum_{T} p(f| T, \ldots)p(\theta_{G1}, \theta_{G2}| T, \ldots) P(T| \ldots) \\
%= \sum_{T} \pi_{host} p(f| T, z=\theta_{G1} ,z_N=\theta_{G2}, \ldots) P(T| z=\theta_{G1}, ,z_N=\theta_{G2}, \ldots)  \\
% +  (1-\pi_{host}) p(f| T, z=\theta_{G2}, z_N=\theta_{G1}, \ldots) P(T| z=\theta_{G2}, z_N=\theta_{G1}, \ldots).
%\label{adumis:eqn}
%\end{multline}
%The probabilities  for the underlying flux and
%relative rate distributions
% Eqns.~\ref{flux:eqn} and \ref{rate:eqn} introduced in \S\ref{obs:sec} apply.
%
%The data likelihood is a truncated version of the
%underlying distribution: the distribution in  Eqn.~\ref{adumis:eqn} normalized by
%the complementary cumulative distribution function
%\begin{equation}
%p(f>f_0, \theta_{G1}, \theta_{G2}|\ldots) = \int_{f_0}^{\infty} p(f, \theta_{G1}, \theta_{G2}|\ldots)df
%\end{equation}
%for a detection threshold of $f_0$.
%
%\subsection{Instantiation of Simulated Data}
%Two simulated data sets are realized.
%The generative model of the first data set (Set 1) is
% identical to the model described in \S\ref{model:sec} with the following parameter choices:
%\begin{itemize}
%\item Cosmological parameters $\Omega_M=0.28$, $w=-1$.
%\item Type~Ia supernovae with intrinsic dispersion $\sigma_{SNIa}=0.1$ mag.
%\item Non-Ia supernovae 2 magnitudes fainter than SNe~Ia with intrinsic
%dispersion $\sigma_{!Ia}=1$ mag.
%\item $z_{max}=1.4$.
%\item $\theta_{r0}=0.95$, $\theta_{r1}=0.2$ for the relative fraction of SNe~Ia
%at $z=0$ and $z=1.1 \times z_{max}$ respectively.
%\item 2000 transients total, including discovered and non-discovered,
%in the range $0.1<z<z_{max}$.
%\item Detection threshold $f_0$ at approximately the mean brightness SN~Ia at $z=1$.
%\item Limits of neighbor-galaxy redshifts of $0.1/1.1$ and $1.1\times z_{max}$.
%\item $\pi_{host}=0.98$ for the probability of correct host-galaxy assignment.
%\item Accumulated spectra for 0\%, 20\%, 60\%, and 100\% of discovered transients.
%\end{itemize}
%
%Figures~\ref{hd:fig} and \ref{hist:fig} show
%the Hubble diagram 
%and redshift histogram for discovered SNe~Ia and non-Ia respectively
%for a realization drawn from this model.
%SNe~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
%are transparent.  The solid points
%represent those objects spectroscopically typed for the case of 20\% follow-up.
%The green X's represent the 2\% among those without spectroscopy that
%are assigned an incorrect redshift.
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2.pdf}
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2_pop2.pdf}  
%\caption{Hubble diagrams for a realization drawn from Set 1 (left)
%and Set 2 that contains  a second non-Ia population (right).
%Type~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
%are transparent.  The green X's represent the 2\% among those without spectroscopy that
%are assigned an incorrect redshift.
%   \label{hd:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2_hist.pdf}
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2_pop2_hist.pdf}  
%\caption{Histogram of redshifts of discovered SNe~Ia and non-Ia from Set 1 (left)
%and Set 2  (right).
%   \label{hist:fig}}
%\end{figure}
%
%The generative model for the second realized data set (Set 2)
%is a perturbed version of that used to make Set 1.
%A second non-Ia population is introduced with average luminosity 0.5 mag fainter
%than the Ia's and with 0.25 mag dispersion.
%The relative rate of the original non-Ia population relative to the second is quadratic
%in redshift:
%\begin{equation}
%1 -\frac{0.8}{\left(1.1\times z_{max}\right)^2}z^2.
%\end{equation}
%A Hubble diagram realization for Set 2 is shown in Figure~\ref{hd:fig} and its redshift histogram
%is in Figure~\ref{hist:fig}.  Sets 1 and 2 are realized using the same seeds:
%The salient differences
%between the two realizations are that a fraction of non-Ia's have absolute magnitude
%based on the second population's average luminosity and intrinsic dispersion; from
%that population different transients satisfy the detection threshold; spectroscopic follow-up is redistributed
%for the new set that is detected.
%
%Sets 1 and 2 share identical discovered SNe~Ia.  For reference, this pure SN~Ia
%set is analyzed with 100\% spectroscopic typing. 
%
%
%\subsection{Analysis Results}
%The simulated data Sets 1 and 2 are analyzed using the model described
%in \S\ref{model:sec}.  The generative model for Set 1 and the analysis model are 
%identical, but for Set 2 they differ.
%The analysis is performed using STAN 
%\citep{stan-software:2015},
%a probabilistic programming language for
%inference of Bayesian models with Hamiltonian Monte Carlo
%using the No-U-Turn sampler \citep{Homan:2014:NSA:2627435.2638586}.  Each run of the analysis
%uses 24 chains each with 1000 links.  In warmup, the first half of the links are used to
%determine parameters for the integration step size and the ``mass''
%parameters of the Hamiltonian. The complementary
%12,000 links represent draws from the model-parameter posterior.
%
%In lieu of including data of
%a pure large SN~Ia sample at low-redshift,
%priors are applied to the SN~Ia absolute magnitude and intrinsic dispersion
%\begin{align}
%\alpha_{SNIa} & \sim \mathcal{N}_{\ln}\left(\ln\left(\alpha^0_{SNIa}\right),\frac{\ln{10}}{2.5}0.02\right)\\
%\sigma_{SNIa} & \sim \mathcal{N}_{\ln}\left(\ln\left(0.1\right),0.1\right),
%\end{align}
%where $\alpha^0_{SNIa}$ is the true input luminosity.
%
%The output of the analyses are presented as contours of the model
%parameter pdf's for the case of 100\% spectroscopic
%typing in Figure~\ref{hd:fig}, and as the mean, median, and
%credible intervals of $w$ for different fractions of
%spectroscopic completeness in Table~\ref{seed2:tab}.
%For Set 1  the input parameters lie 
%within the 68\% credible intervals of the resultant posterior despite the non-trivial appearance of
%the Hubble diagram in Figure~\ref{hd:fig}.
%The decrease in spectroscopic coverage leads to shifts in $w$ of 0.04 relative to the 100\%-coverage case, which is small
%relative to the credible interval range of 0.281.  However, there is an
%increase in parameter uncertainties, going from $\Delta w= 0.281$ to 0.309
%when there is no spectroscopy; the latter is a constraining result considering that none of the
%objects are directly typed and redshifted.
%Relative to Set 1 with 100\% spectrosocopy,
%the subset of SNe~Ia-only has a shift in the mean $w$ of 0.12
%and no appreciable increase in the uncertainty of $w$.
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2/{contour.2000.1.0}.pdf}
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2_pop2/{contour.2000.1.0}.pdf} 
%\caption{Probability distribution functions of the model parameters for the case of
% 100\% spectroscopy for Model 1 (left) and Model 2 (right).
%    Solid lines represent the input parameters.  Dashed lines in the
%   histograms are the 0.16 and 0.84 quantiles. 
%   \label{contour1:fig}}
%\end{figure}
%
%\begin{table}
%\centering
%\begin{tabular}{|c|cc|ccc|}
%\hline
%Fraction Spec&$<w>$ & $w$ median &\%-ile  &$w$ range & $\Delta w$\\ \hline
%\multicolumn{6}{|c|}{SN Ia only}\\ \hline
%$1.00$& $-0.94 \pm 0.09$ &$-0.94$ &$0.68$ & $[-1.076, -0.801]$ & $0.275$ \\
%&  &  & $0.90$ & $[-1.166, -0.718]$ & $0.448$ \\
%&  &  & $0.95$ & $[-1.208, -0.682]$ & $0.525$ \\
% \hline
%\multicolumn{6}{|c|}{Set 1}\\ \hline
%$1.00$ &$-1.06 \pm  0.09$ &$-1.05$ &$0.68$ & $[-1.196, -0.914]$ & $0.281$ \\
%&&&$0.90$ & $[-1.292, -0.826]$ & $0.466$ \\
%&&&$0.95$ & $[-1.340, -0.784]$ & $0.556$ \\
%\hline
%$0.60$ &$-1.08 \pm  0.10$ & $-1.08$ &$0.68$ & $[-1.227, -0.933]$ & $0.293$ \\
%& & &$0.90$ & $[-1.331, -0.836]$ & $0.495$ \\
%& & &$0.95$ & $[-1.385, -0.787]$ & $0.598$ \\
%\hline
%$0.20$ &$-1.10 \pm  0.10$ & $-1.09$ &$0.68$ & $[-1.249, -0.943]$ & $0.305$ \\
%& & &$0.90$ & $[-1.354, -0.851]$ & $0.502$ \\
%& & &$0.95$ & $[-1.408, -0.810]$ & $0.597$ \\
%\hline
%$0.00$ &$-1.07 \pm  0.10$ & $-1.06$ &$0.68$ & $[-1.218, -0.909]$ & $0.309$ \\
%& & &$0.90$ & $[-1.320, -0.812]$ & $0.508$ \\
%& & &$0.95$ & $[-1.372, -0.770]$ & $0.602$ \\
%\hline
%\multicolumn{6}{|c|}{Set 2}\\ \hline
%$1.00$ &$-0.94 \pm  0.09$ &$-0.94$ &$0.68$ & $[-1.078, -0.804]$ & $0.274$ \\
%&&&$0.90$ & $[-1.168, -0.720]$ & $0.447$ \\
%&&&$0.95$ & $[-1.214, -0.680]$ & $0.534$ \\
%\hline
%$0.60$ &$-1.04 \pm  0.10$ & $-1.03$ &$0.68$ & $[-1.182, -0.883]$ & $0.298$ \\
%& & &$0.90$ & $[-1.285, -0.791]$ & $0.494$ \\
%& & &$0.95$ & $[-1.332, -0.749]$ & $0.583$ \\
%\hline
%$0.20$ &$-1.10 \pm  0.10$ & $-1.09$ &$0.68$ & $[-1.247, -0.936]$ & $0.311$ \\
%& & &$0.90$ & $[-1.359, -0.840]$ & $0.518$ \\
%& & &$0.95$ & $[-1.413, -0.798]$ & $0.614$ \\
%\hline
%$0.00$ &$-1.11 \pm  0.10$ & $-1.10$ &$0.68$ & $[-1.266, -0.948]$ & $0.317$ \\
%& & &$0.90$ & $[-1.375, -0.850]$ & $0.525$ \\
%& & &$0.95$ & $[-1.428, -0.808]$ & $0.621$ \\
%\hline
%\end{tabular}
%\caption{Mean, median, and credible intervals of $w$ for the SNe~Ia only,
%Model 1, and Model 2 for varying
%fractions of spectroscopic coverage. \label{seed2:tab}}
%\end{table}
%
%The generative model for Set 2 differs from the analysis model.
%While it is expected for the input and resultant non-Ia parameters not to match, 
%when there is 100\% spectroscopic coverage the posteriors of the other
%parameters accommodate the input with the $w$ posterior matching
%closely that of the SN~Ia-only analysis.  Unlike Model~1, which exhibits no bias
%in $w$ with decreasing spectroscopy, Model~2 does have a shift in the $w$
%credible region.  
%The values of the mean and median values of
%$w$ for both 20\% and 0\% coverage are outside the 68\%
%credible region of the ideal 100\% coverage.
%The Set 2 $\Delta w$ increases quicker with decreasing spectroscopy
%relative to Set 1.
%
%These results are consistent with the trivial expectation that when generative
%and analysis models are identical statistically correct parameter determinations
%occur, and that analysis with an incorrect non-Ia model can lead to bias.
%The latter case is mitigated with complete spectroscopic typing, though then
%the non-Ia's contribute little statistically to the $w$ constraints. 
%

\bibliographystyle{elsarticle-num} 
\bibliography{alex}

\end{document}

%
%\subsubsection{SN~Ia Only}
%A classical analysis would only include confirmed SNe~Ia with a definitive
%redshift. The realization of simulated data has 415 Type~Ia supernovae.
%Results of the analysis of a pure SN~Ia sample with no
%host misclassification are presented:  Figure~\ref{contour.ia_only:fig}
%shows the posteriors for  $\Omega_M$--$w$ and Table~\ref{ia_only:tab} the
%credible equal-tailed intervals for $w$.  The true $w=-1$ is within
%the 0.58 credible interval. (Note that $w$ is the parameter
%of most interest for the experiment.)
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.55]{/Users/akim/project/abc/results/{contour..ia_only.500}.pdf} \newline
%\caption{Probability distribution function for $\Omega_M$--$w$ for only SNe~Ia
%   all with spectra.
%   Solid lines represent the input cosmology.  Dashed lines in the
%   histograms are the 0.16 and 0.86 quantiles. 
%   \label{contour.ia_only:fig}}
%\end{figure}
%
%\begin{table}
%\centering
%\begin{tabular}{|c|cc|}
%\hline
%\%-ile & $w$ range & $\Delta w$\\ \hline
%$0.68$ & $[-1.254, -0.975]$ & $0.279$ \\
%$0.90$ & $[-1.347, -0.889]$ & $0.458$ \\
%$0.95$ & $[-1.396, -0.846]$ & $0.550$ \\
%\hline
%\end{tabular}
%\caption{Credible intervals of $w$ using the SNe~Ia only with spectroscopic
%observations.\label{ia_only:tab}}
%\end{table}
%
%\subsubsection{Full Sample: Varying Spectroscopic Completeness}
%We now turn to the full sample of SNe~Ia and non-Ia's.  Recall that our simple
%model treats the non-Ia's in the sample
%as standard candles with a large (1 mag) intrinsic dispersion.
%
%Results for the $\Omega_M$--$w$ posterior for the
%case of 40\%, 70\%, and 100\% spectroscopic completeness
%are shown in Figures~\ref{contour.200:fig}, \ref{contour.350:fig}, and
%\ref{contour.500:fig} respectively.
%The credible intervals
%of $w$ for are given in Table~\ref{compare:tab}, with
%the input $w=-1$ within the 0.65, 0.49, and 0.47  intervals
%respectively.
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.200}.pdf} 
%   \caption{Probability distribution function for $\Omega_M$--$w$ for:
%    SNe~Ia and non-Ia with 40\%  spectroscopic completeness.
%   \label{contour.200:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.350}.pdf} 
%   \caption{Probability distribution function for $\Omega_M$--$w$ for
%    SNe~Ia and non-Ia with 70\%  spectroscopic completeness.
%   \label{contour.350:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.500}.pdf} 
%   %\includegraphics[scale=0.55]{/Users/akim/project/abc/results/{contour.200}.pdf}    
%   \caption{Probability distribution function for $\Omega_M$--$w$ for
%    SNe~Ia and non-Ia with 100\%  spectroscopic completeness.
%   \label{contour.500:fig}}
%\end{figure}
%
%\begin{table}[htbp] 
%\centering
%\begin{tabular}{|cc|cc|}
%\hline
%\# Spectra & \%-ile & $w$ range & $\Delta$\\ \hline
%200& $0.68$ & $[-1.284, -0.991]$ & $0.293$ \\
%200& $0.90$ & $[-1.385, -0.900]$ & $0.485$ \\
%200& $0.95$ & $[-1.435, -0.860]$ & $0.575$ \\
%350& $0.68$ & $[-1.243, -0.952]$ & $0.291$ \\
%350& $0.90$ & $[-1.339, -0.858]$ & $0.481$ \\
%350& $0.95$ & $[-1.390, -0.817]$ & $0.573$ \\
%500& $0.68$ & $[-1.236, -0.948]$ & $0.289$ \\
%500& $0.90$ & $[-1.336, -0.861]$ & $0.475$ \\
%500& $0.95$ & $[-1.385, -0.815]$ & $0.570$ \\
%\hline
%\end{tabular}
%\caption{Credible intervals of $w$ given different levels of spectroscopy.  \label{compare:tab}}
%\end{table}
%
%As would be expected, the better the spectroscopic completeness, the less bias
%and smaller uncertainties that are obtained.
%
%
%\section{Subsections of the Model}
%
%\subsection{Type}
%\label{type:sec}
%Since our transient sample may not consist purely of SNe~Ia, the model
%has a term of the form $P(T, G | z,\mu)$, which should account for all types of
%objects that could potentially be mistaken for Type~Ia:
%the rates and association with host galaxies are described for
%the full population of potential interlopers.
%This term is not currently well known, particularly at the highest redshifts probed by DES.
%
%An idea is to consider two types: SN~Ia and non-SN~Ia, where the latter's
%intrinsic distributions have loose priors.  Transients that have been spectroscopically
%classified as non-Ia will provide the strongest leverage in constraining the
%contamination: this subset must be unbiased or properly weighted
%relative to the underlying population that constitutes the Hubble diagram.
%Running the model on a pure sample of SNe~Ia can have benefits.
%As discussed in \S\ref{systematics:sec}, comparing these results with
%those from the full sample provides
%a test of systematics.  Alternatively, the non-Ia model may be better constrained
%with the pure sample.
%
%
%To minimize the contribution of non-Ia contamination we strive for
%a pure sample selection so that $P(T=\text{non-Ia}| \mu) \rightarrow 0$.
%
%
%\subsection{Host Matching}
%\sloppy
%The host-galaxy properties depend on the GC and the transient
%coordinates $P({z}_H, {\theta}_H | {\mathit{Gals.}}, \text{RA/Dec})$.  With a
%spectrum of the transient and the underlying background, identification of the host galaxy
%is usually straightforward.
%Otherwise, projections or ambiguous cases can result in the misidentification of
%the host.
%Other supernova surveys (e.g.\ SNLS) 
%with spectroscopically confirmed host associations (e.g.\ with
%matched transient and galaxy redshifts) that replicate the DES population
%can be used to put a prior on this probability.
%
%%An effective way of addressing this term is to
%%analyze the supernovae assuming a correct host match, and
%%experimentally determine $P(\text{mismatch} )$ and
%%$\left. \langle {\mu}-\mu(z)\rangle \right|_{\text{mismatch}}$, $\left. \langle {z}-z \rangle \right|_{\text{mismatch}}$ and cross-terms to correct bias in the Hubble diagram. 
%
%
%% only model $T=\text{SN~Ia}$,
%%and experimentally determine $P(\text{non-Ia})$, 
%%$\left. \langle \mu-\mu_{model} \rangle \right|_{non-Ia}$, and its uncertainty for the
%%sample of non-Ia's that are discovered and misclassified as SN~Ia.
%%This measurement is achieved through spectroscopic
%%typing of a subset drawn from the populations used in the Hubble diagram.
%%The weighted bias would be added as a correction to the Hubble diagram under the assumption that all objects are SNe~Ia.
%
%
%\section{Systematic Tests}
%\label{systematics:sec}
%The Hubble diagrams of subsets of data quantify systematic uncertainty.  
%Subsets that may be expected to show evidence for systematics include:
%spectroscopically typed SNe~Ia; transients in deep (3) versus shallow (1/2) DES fields;
%transients with NIR data; splits based on host properties.
%
%A potentially more probative test is to compare the distance moduii inferred for the same
%supernovae, but reducing the amount of data considered.  A subset of DES
%transients have higher signal-to-noise, spectroscopic confirmation, and NIR coverage than
%the average transient, we can take away that extra information in calculating distance modulus
%to get $P(\mu | \mathit{data} - \mu | \mathit{data}^-, z | \mathit{data} - z | \mathit{data}^-)$.
%It is possible that differences in $\mu$s and $z$s have smaller uncertainties that the
%$\mu$ and $z$ alone due to common measurement uncertainties. 
%
%
%
%Generically ${{T}}$ can describe sample selection.  I choose to focus
%on objects typed as SN~Ia, rather than  all discovered transients, not just those typed as SN~Ia. 
%The non-Ia transient population has significant model uncertainty and the information they impart on the expansion history is weak:  I anticipate it is cleaner to limit ourselves to
%modeling discoveries we think are SN~Ia. This restriction does not obviate the need
%to consider non-Ia contamination in the sample, rather
%we only need to model non-Ia's
%that are mistaken for Ia and not all discovered transients.

### 2016-02-23 doc/toy.tex
rewrite of toy note and cleaning up of some hanging changes

@@
 The ambition is to construct a model to determine the likelihood of DES-like
-supernova data.
+supernova data.  We start with a toy model and plan to evolve its complexity to
+have the fidelity necessary for the DES analysis.
 The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
@@
 \begin{itemize}
-\item $c_o$: Counts at peak brightness.
+\item $c_o$: Counts at peak brightness, one or more independent measurements.
 \item $z_o$: Redshift.
@@
 In what follows we assume that the typed objects are drawn from the same underlying
-distribution as the sample $S_c=S_T$, and that the selection criterion is a brightness
+distribution as the sample $S_c=S_T$, and that the selection criterion depends on a  brightness
 threshold $S_c = c_o \ge c_T$.  (The spectroscopic sample $S_T$ 
@@
 For this preliminary discussion the top-level parameters are denoted by $X$,
-latent parameters $L$ and $T$.
+the latent parameters are $L$ and $T$ (for luminosity and type).
 
@@
 \end{align}
-practically $T$ and $z$ are no longer probabilistic but are fixed.
+$T$ and $z$ are no longer probabilistic but are fixed.
 Then
@@
 
-In the case of a single count measurement per transient and neglecting the typing selection criteria,
-\begin{align}
- p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
- &= \begin{cases}
-   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
-   0 & \text{if }  c_o < c_T
- \end{cases}
-\end{align}
-so that the integral in brackets is
-\begin{equation}
-\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
-\end{equation}
-
-For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
-noise of a single observation is expected to be dominated by its own photon noise making this
-supposition a good approximation.} and $k$ of those must be above threshold in order to be
-included in the sample.
-Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
-of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
-Then
-\begin{align}
-p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
- &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
-   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } c_{o,\alpha} \ge c_T == g_\alpha \\
-   0 & \text{if }  c_{o,\alpha} \ge c_T\  != g_\alpha
- \end{cases}
-\end{align}
-and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
-of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
-
 \subsection{Case Not Typed}
@@
 \begin{align}
-\mathcal{L}(z_o,c_o | S_c, z, X) & =  \sum_j p_j \int dL \,\sum_{i} p(c_o | T_i, L, S_c, z=z_j, X) p(L |  T_i,  S_c,  z=z_j, X) P(T_i|S_c, z=z_j, X) \\
-&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)}\\
-&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c| z=z_j, X)},
+\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
+&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c, S_T | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)}\\
+&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)},
 \end{align}
@@
 
+\subsection{Sample Selection}
+Both the Typed and Not Typed cases have in their normalizations the term
+\begin{equation}
+\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X),
+\end{equation}
+which is the fraction of objects whose counts satisfy the sample selection criteria.
+
+In the case of a single count measurement per transient and neglecting the typing selection criteria,
+\begin{align}
+ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
+ &= \begin{cases}
+   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
+   0 & \text{if }  c_o < c_T
+ \end{cases}
+\end{align}
+so that the integral is
+\begin{equation}
+\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
+\end{equation}
+Our count measurements come from a Normal distribution, so that integral is simply related to the error function.
+
+For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
+noise of a single observation is expected to be dominated by its own photon noise making this
+supposition a good approximation.} and $k$ of those must be above threshold in order to be
+included in the sample.
+Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
+of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
+Then
+\begin{align}
+p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
+ &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
+   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } (c_{o,\alpha} \ge c_T) = g_\alpha\\
+   0 & \text{if }  (c_{o,\alpha} \ge c_T)  \ne g_\alpha
+ \end{cases}
+\end{align}
+and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
+of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
+If the count measurement uncertainties are independent, which may not be exactly but is approximately the case when we have high
+signal-to-noise references, the integral because a sum of products of normalized error functions.
+
 \section{Model}
@@
 \subsection{Cosmology}
-The Flat $w$CDM cosmology.
-
+The Flat $w$CDM cosmological parameters are drawn from top-hat distributions
 \begin{align}
-\Omega_M & \sim \ln{\mathcal{N}}(\ln{0.28},0.1)\\
-w_0 & \sim \mathcal{N}(-1,0.05)
+\Omega_M & \sim  {\mathcal{U}}(0,1)\\
+w_0 & \sim \mathcal{U}(-1.5, -0.1).
 \end{align}
@@
 \begin{equation}
-Z \sim \mathcal{N}(0,0.02)
+Z \sim \mathcal{N}(0,\frac{\ln{10}}{2.5}0.01)
 \end{equation}
@@
 
-\subsection{Sample Selection Normalization}
+\subsection{Implementation Notes}
 Anticipating the use of available statistical packages, the likelihoods have been constructed
@@
 
-\subsection{Implementation Notes}
 The following integral appears in the normalization
@@
 \end{equation}
+
 \end{document}

### 2016-03-01 tex/toy.tex
Initial refactor. Nothing works at the moment, but want to get something committed.

--- 2016-02-23 doc/toy.tex
@@
 
+\section{Adding Complexity}
+\subsection{SALT2 Model for Transient Luminosity}
+The SALT2 model for supernova luminosity as a function of SN-frame time $t$  and wavelength $\lambda$ is
+\begin{equation}
+L(t-t_0, \lambda; M-x_0-\alpha x_1 + \beta c, x_1, c),
+\label{SALTL:eqn}
+\end{equation}
+where $\alpha$ and $\beta$ are global parameters and $t_0$, $x_0$, $x_1$ and $c$ are per-supernova parameters.
+
+Incorporating the SALT2 model leads to changes to the PGM in Figure~\ref{pgm:fig}, specifically in the following Nodes.
+UNITY notation and priors are taken.
+\subsubsection{SNX Luminosity}
+There are parameters for the global supernova parameters.  For SNe~Ia
+\begin{align}
+M_{SNIa} & \sim \mathcal{U}(-20, -18) \\
+\arctan{\alpha_{SNIa}} & \sim \mathcal{U}(-0.2, 0.3) \\
+\arctan{\beta_{SNIa}} & \sim \mathcal{U}(-1.4, 1.4)
+\end{align}
+and for ``hyperparameters'' that describe the distributions of the per-supernova parameters
+\begin{align}
+\log_{10}{R^{x_0}_{SNIa}} & \sim \mathcal{U}({-2}, {-0.5})\\
+\log_{10}{R^{x_1}_{SNIa}} & \sim \mathcal{U}({-0.5}, {0.5})\\
+\log_{10}{R^{c}_{SNIa}} & \sim \mathcal{U}({-1.5}, {-0.5})\\
+x_{1,SNIa}^\star& \sim \text{Cauchy}(0,1)\\
+c^\star_{SNIa} & \sim \text{Cauchy}(0,0.3).
+\end{align}
+
+\subsubsection{Type}
+The type of the object is still obtained as before.  Now we have the parameters that specify the sub-type.
+\begin{align}
+t_0 & \sim \mathcal{U}\\
+x_{0, SNIa} & \sim \mathcal{N}(0,R^{x_0}_{SNX})\\
+x_{1,SNIa} & \sim \mathcal{N}(x_{1,SNIa}^\star,R^{x_1}_{SNIa})\\
+c_{SNIa} & \sim \text{SkewNormal}(c^\star_{SNIa},R^{c}_{SNIa} )
+\end{align}
+
+\subsubsection{Luminosity}
+The luminosity is now fixed by Eqn.\ \ref{SALTL:eqn}.  It may be made probabilistic if an error snake is considered.
+
+\subsubsection{Likelihood}
+The likelihood for typed objects is
+\begin{multline}
+\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X)
+ \propto \int dt_0 \, dx_{0,T_o} \, dx_{1,T_o} dc_{T_o} 
+ p(t_0, x_{0,T_o}, x_{1,T_o}, c_{T_o} |T_o, z=z_o, X) \\ P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L_{T_o}(t_0, x_{0,T_o}, x_{1,T_o}, c_{T_o}), z=z_o, X)
+\end{multline}
+and for non-typed
+\begin{multline}
+\mathcal{L}(z_o,c_o | S_c, S_T, z, X)
+ \propto \sum_j p_j \sum_i P(T_i|z=z_j, X)   \int dt_0 \, dx_{0,T_i} \, dx_{1,T_i} dc_{T_i} 
+ p(t_0, x_{0,T_i}, x_{1,T_i}, c_{T_i} |T_i, z=z_j, X) \\  p(c_o, S_c, S_T | T=T_o, L_{T_i}(t_0, x_{0,T_i}, x_{1,T_i}, c_{T_i}), z=z_j, X),
+ \end{multline}
+where both likelihoods must be normalized.
+
 \end{document}

### 2016-03-01 dessn/model/model.py

@@ -1,0 +1,12 @@
+import numpy as np
+
+from dessn.model.nodes.cosmology import Cosmology, FlatWCDM
+from dessn.model.nodes.supernova import Supernova, SupernovaIa
+
+class Model(object):
+
+    def __init__(self):
+        self.nodes = []
+
+
+

### 2016-03-01 dessn/model/node.py

@@ -1,0 +1,22 @@
+import abc
+
+
+class Node(object):
+    __metaclass__ = abc.ABCMeta
+
+    def __init__(self):
+        self.dependencies = []
+
+    def add_dependency(self, dependency_class):
+        self.dependencies.append(dependency_class)
+
+    def get_dependencies(self):
+        return self.dependencies
+
+    @abc.abstractmethod
+    def get_name(self):
+        return
+
+    @abc.abstractmethod
+    def __call__(self):
+        return

### 2016-03-01 dessn/model/nodes/cosmology.py

@@ -1,0 +1,14 @@
+import abc
+from dessn.model.node import Node
+
+
+class Cosmology(Node):
+    __metaclass__ = abc.ABCMeta
+
+    def get_name(self):
+        return "Cosmology"
+
+
+class FlatWCDM(Cosmology):
+    def __call__(self):
+        pass

### 2016-03-01 dessn/model/nodes/supernova.py

@@ -1,0 +1,42 @@
+import abc
+from dessn.model.node import Node
+from dessn.model.nodes.typeProb import Types
+
+
+class Supernova(Node):
+    """ Abstract supernova which others must implement
+    """
+    __metaclass__ = abc.ABCMeta
+
+    @abc.abstractproperty
+    def type(self):
+        return
+
+
+class SupernovaIa(object):
+    """Models a type Ia supernova
+
+    Models the luminosity distribution of type Ia supernovas statically (without
+    internal parameters).
+
+    .. math::
+        P(L) \sim N(
+
+    Parameters
+    ----------
+    log_luminosity : float
+        Represented by the variable :math:`\mu`
+    sigma_luminosity : float
+        Represented by the variable :math:`\sigma`
+    """
+    def __init__(self, log_luminosity, sigma_luminosity):
+        self.log_luminosity = log_luminosity
+        self.sigma_luminosity = sigma_luminosity
+
+    def get_luminosity_prob(self, log_luminosity):
+        return
+
+    def type(self):
+        return Types.snIa
+
+

### 2016-03-01 dessn/model/nodes/typeProb.py

@@ -1,0 +1,43 @@
+from dessn.model.node import Node
+from enum import Enum
+import abc
+
+
+class Types(Enum):
+    """Possible target types
+    """
+    snIa = 1
+    snII = 2
+
+
+# todo: Consider if we should have a discrete and continuous subclasses of Node
+
+class TypeProbability(Node):
+    """Abstract type probability node, from which all implementations should inherit
+    """
+    __metaclass__ = abc.ABCMeta
+
+    def get_name(self):
+        return "Type Probability"
+
+
+class TypeProbabilitySimple(TypeProbability):
+    """The Type probability node.
+
+    Takes *some information* to determine the probability of
+    of the object in question being a type of supernova.
+
+    Parameters
+    ----------
+    relative_rate : Optional[str]
+        Relative rate of SnIa / SnII
+
+
+    """
+
+    def __init__(self, relative_rate=0.333):
+        super(TypeProbabilitySimple, self).__init__()
+        self.relative_rate = relative_rate
+
+    def __call__(self):
+        return {Types.snIa: self.relative_rate, Types.snII: 1 - self.relative_rate}

### 2016-03-01 dessn/simple/exampleIntegral.py

@@ -1,0 +1,160 @@
+import numpy as np
+import emcee
+from scipy.integrate import quad
+from scipy import stats
+import matplotlib, matplotlib.pyplot as plt
+from progressbar import Bar, Counter, ETA, Percentage, ProgressBar, Timer
+import corner
+
+class ExampleIntegral(object):
+    r"""An example implementation using integration over a latent parameter.
+
+    Let us assume that we are observing supernova that a drawn from an underlying
+    supernova distribution parameterised by :math:`\theta`,
+    where the supernova itself simply a luminosity :math:`L`. We measure the luminosity
+    of multiple supernovas, giving us an array of measurements :math:`D`. If we wish to recover
+    the underlying distribution of supernovas from our measurements, we wish to find :math:`P(\theta|D)`,
+    which is given by
+
+    .. math::
+        P(\theta|D) \propto P(D|\theta)P(\theta)
+
+
+    Note that in the above equation, we realise that :math:`P(D|L) = \prod_{i=1}^N P(D_i|L_i)` as our measurements are
+    independent. The likelihood :math:`P(D|\theta)` is given by
+
+    .. math::
+        P(D|\theta) =  \prod_{i=1}^N  \int_{-\infty}^\infty P(D_i|L_i) P(L_i|\theta) dL_i
+
+
+
+    We now have two distributions to characterise. Let us assume both are gaussian, that is
+    our observed luminosity :math:`x_i` has gaussian error :math:`\sigma_i` from the actual supernova
+    luminosity, and the supernova luminosity is drawn from an underlying gaussian distribution
+    parameterised by :math:`\theta`.
+
+     .. math::
+        P(D_i|L_i) = \frac{1}{\sqrt{2\pi}\sigma_i}\exp\left(-\frac{(x_i-L_i)^2}{\sigma_i^2}\right)
+
+        P(L_i|\theta) = \frac{1}{\sqrt{2\pi}\theta_2}\exp\left(-\frac{(L_i-\theta_1)^2}{\theta_2^2}\right)
+
+
+
+    This gives us a likelihood of
+
+    .. math::
+
+        P(D|\theta) = \prod_{i=1}^N  \frac{1}{2\pi \theta_2 \sigma_i}  \int_{-\infty}^\infty
+        \exp\left(-\frac{(x_i-L_i)^2}{\sigma_i^2} -\frac{(L_i-\theta_1)^2}{\theta_2^2} \right) dL_i
+
+
[…]

### 2016-03-01 dessn/simulation/observationFactory.py

@@ -1,0 +1,82 @@
+import logging
+import numpy as np
+
+
+class ObservationFactory(object):
+    def __init__(self, **kwargs):
+        self.kwargs = kwargs
+        self.logger = logging.getLogger(__name__)
+        self.check_kwargs()
+
+    def check_kwargs(self):
+        reqs = ['rate_II_r', 'logL_snIa', 'logL_snII', 'sigma_snIa', 'sigma_snII', 'cosmology']
+        keys = self.kwargs.keys()
+        fail = []
+        for req in reqs:
+            if req not in keys:
+                fail.append(req)
+        if len(fail) > 0:
+            message = "Required keys %s not found" % fail
+            self.logger.error(message, exc_info=True)
+            raise ValueError("Required keys %s not found" % fail)
+
+    def get_observations(self, num):
+        """ Still needs massive refactoring """
+        specz = np.random.uniform(low=0.1, high=0.8, size=num)
+        zprob = np.zeros(num) + 1.0
+
+        snIa_rate = 1.0 / (1 + self.kwargs['rate_II_r'])
+
+        spectype = (np.random.uniform(low=0, high=1, size=num) > snIa_rate) * 1
+        # 0 represents snIa, 1 represents snII
+
+        snIa_luminosity = np.exp(self.kwargs['logL_snIa'] * np.power(10, np.random.normal(0, self.kwargs['sigma_snIa']/2.5), size=num))
+        snII_luminosity = np.exp(self.kwargs['logL_snII'] * np.power(10, np.random.normal(0, self.kwargs['sigma_snII']/2.5), size=num))
+
+        luminosity = (1 - spectype) * snIa_luminosity + spectype * snII_luminosity
+        luminosity_distance = self.kwargs['cosmology'].luminosity_distance(specz).value
+
+        npts = 2
+        cov = np.zeros((2,2))
+        cov[0,0] = 1e-20
+        cov[1,1] = 1e-20
+        cov[0,1] = 0     #for now uncorrelated as algorithm handles that
+        cov[1,0] = 0     #for now uncorrelated as algorithm handles that
+        invcov = np.linalg.inv(cov)
+
+        counts = []
+        counts_invcov = []
+        for i in xrange(num):
+            ans = np.random.multivariate_normal(np.zeros(npts)+ luminosity[i] / (4 * np.pi * luminosity_distance[i] * luminosity_distance[i] * 10**(self.kwargs['Z']/2.5)), cov)
+            counts.append(ans)
[…]

### 2016-03-01 doc/conf.py

@@ -1,0 +1,288 @@
+# -*- coding: utf-8 -*-
+#
+# dessn documentation build configuration file, created by
+# sphinx-quickstart on Mon Feb 29 10:30:30 2016.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+import sys
+import os
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+sys.path.insert(0, os.path.abspath('..'))
+
+# -- General configuration ------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+    'sphinx.ext.autodoc',
+    'sphinx.ext.mathjax',
+    'sphinx.ext.viewcode',
+    'sphinx.ext.autosummary',
+    'numpydoc']
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The encoding of source files.
+#source_encoding = 'utf-8-sig'
+
+# The master toctree document.
+master_doc = 'modules'
+
+# General information about the project.
+project = u'dessn'
+copyright = u'2016, dessn'
+author = u'dessn'
+
+# The version info for the project you're documenting, acts as replacement for
+# |version| and |release|, also used in various other places throughout the
+# built documents.
+#
+# The short X.Y version.
+version = u'0.0.1'
+# The full version, including alpha/beta/rc tags.
+release = u'0.0.1'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
+# This is also used if you do content translation via gettext catalogs.
+# Usually you set "language" from the command line for these cases.
+language = None
+
+# There are two options for replacing |today|: either, you set today to some
+# non-false value, then it is used:
+#today = ''
[…]

### 2016-03-01 old/Nemcee.py

@@ -366,7 +366,7 @@ def simulateData():
     observation['counts'] = []
     observation['counts_invcov']=[]
     for i in xrange(nTrans):
-        ans =numpy.random.multivariate_normal(numpy.zeros(npts)+ luminosity[i] / 4/numpy.pi/ld[i]/ld[i]*10**(inputs.Z/2.5), cov)
+        ans = numpy.random.multivariate_normal(numpy.zeros(npts)+ luminosity[i] / 4/numpy.pi/ld[i]/ld[i]*10**(inputs.Z/2.5), cov)
         observation['counts'].append(ans)
         observation['counts_invcov'].append(invcov)
         #luminosity / 4/numpy.pi/ld/ld*10**(inputs.Z/2.5)
@@ -386,7 +386,7 @@ def simulateData():
     nTrans =  found.sum()
     observation['specz'] = [numpy.array([dum]) for dum in observation['specz'][found]]
     observation['zprob'] = [numpy.array([dum]) for dum in observation['zprob'][found]]
-    observation['spectype'] =observation['spectype'][found]
+    observation['spectype'] = observation['spectype'][found]
 
     observation['spectype'][0] = -1   # case of no spectral type
     observation['specz'][0] = numpy.array([observation['specz'][0][0], 0.2])
@@ -723,17 +723,17 @@ import matplotlib.pyplot as plt
 
 def runModel():
 
-    pool = MPIPool()
-    if not pool.is_master():
-        pool.wait()
-        sys.exit(0)
-    # pool=None
+    #pool = MPIPool()
+    #if not pool.is_master():
+    #    pool.wait()
+    #    sys.exit(0)
+    #pool=None
 
     observation = simulateData()
     nTrans = len(observation['spectype'])
 
-    ndim, nwalkers = 8+ nTrans, 1000
-
+    ndim, nwalkers = 8+ nTrans, 70
+    print("1")
     # mns = numpy.concatenate(([inputs.Om0, inputs.w0, inputs.rate_II_r, inputs.logL_snIa, inputs.sigma_snIa, \
     #             inputs.logL_snII,inputs.sigma_snII,inputs.Z], -.35*numpy.zeros(nTrans)))
     sigs = numpy.concatenate(([.1,.2,.1, uncertainties.logL_snIa, uncertainties.sigma_snIa, uncertainties.logL_snII, uncertainties.sigma_snII, uncertainties.Z], .05+numpy.zeros(nTrans)))
@@ -755,8 +755,9 @@ def runModel():
 
 
[…]

### 2016-03-01 old/data.py

@@ -2,17 +2,6 @@
 
 #import matplotlib
 #matplotlib.use('Agg')
-import matplotlib.pyplot as plt
-from matplotlib.backends.backend_pdf import PdfPages
-import pystan
-import numpy.random
-from astropy.cosmology import FlatwCDM
-import scipy
-import sncosmo
-import pickle
-import astropy.units as u
-import abc
-import copy
 import Sim
 		
 

### 2016-03-01 doc/conf.py
Reducing warnings from nupmydoc

@@ -48,7 +48,7 @@ source_suffix = '.rst'
 
 # The master toctree document.
 master_doc = 'modules'
-
+numpydoc_show_class_members = False
 # General information about the project.
 project = u'dessn'
 copyright = u'2016, dessn'

### 2016-03-01 doc/_build/latex/dessn.tex
Adding built documentation for github pages

% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{00A0}{\nobreakspace}
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{longtable}
\usepackage{sphinx}
\usepackage{multirow}
\usepackage{eqparbox}
\usepackage{amsfonts}

\addto\captionsenglish{\renewcommand{\figurename}{Fig. }}
\addto\captionsenglish{\renewcommand{\tablename}{Table }}
\SetupFloatingEnvironment{literal-block}{name=Listing }

\title{dessn Documentation}
\date{March 01, 2016}
\release{0.0.1}
\author{dessn}
\newcommand{\sphinxlogo}{}
\renewcommand{\releasename}{Release}
\setcounter{tocdepth}{3}
\makeindex

\makeatletter
\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
    \let\PYG@ul=\relax \let\PYG@tc=\relax%
    \let\PYG@bc=\relax \let\PYG@ff=\relax}
\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
\def\PYG@toks#1+{\ifx\relax#1\empty\else%
    \PYG@tok{#1}\expandafter\PYG@toks\fi}
\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}

\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@cs\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{1.00,0.94,0.94}{\strut ##1}}}
\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.20,0.20,0.20}{##1}}}
\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.84,0.33,0.22}{##1}}}
\expandafter\def\csname PYG@tok@nl\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.13,0.44}{##1}}}
\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.38,0.68,0.84}{##1}}}
\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@nd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
\expandafter\def\csname PYG@tok@ne\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.49}{##1}}}
\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.44,0.63,0.82}{##1}}}
\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.45}{##1}}}
\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@ch\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PYG@tok@mb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.32,0.47,0.09}{##1}}}
\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.14,0.33,0.53}{##1}}}
\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@cpf\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.56,0.13,0.00}{##1}}}
\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}

\def\PYGZbs{\char`\\}
\def\PYGZus{\char`\_}
\def\PYGZob{\char`\{}
\def\PYGZcb{\char`\}}
\def\PYGZca{\char`\^}
\def\PYGZam{\char`\&}
\def\PYGZlt{\char`\<}
\def\PYGZgt{\char`\>}
\def\PYGZsh{\char`\#}
\def\PYGZpc{\char`\%}
\def\PYGZdl{\char`\$}
\def\PYGZhy{\char`\-}
\def\PYGZsq{\char`\'}
\def\PYGZdq{\char`\"}
\def\PYGZti{\char`\~}
% for compatibility with earlier versions
\def\PYGZat{@}
\def\PYGZlb{[}
\def\PYGZrb{]}
\makeatother

\renewcommand\PYGZsq{\textquotesingle}

\begin{document}

\maketitle
\tableofcontents
\phantomsection\label{modules::doc}

\chapter{dessn package}
\label{dessn:dessn}\label{dessn::doc}\label{dessn:dessn-package}

\section{Subpackages}
\label{dessn:subpackages}

\subsection{dessn.entry package}
\label{dessn.entry:dessn-entry-package}\label{dessn.entry::doc}

\subsubsection{Submodules}
\label{dessn.entry:submodules}

\subsubsection{dessn.entry.sim module}
\label{dessn.entry:dessn-entry-sim-module}\label{dessn.entry:module-dessn.entry.sim}\index{dessn.entry.sim (module)}

\subsubsection{Module contents}
\label{dessn.entry:module-dessn.entry}\label{dessn.entry:module-contents}\index{dessn.entry (module)}

\subsection{dessn.model package}
\label{dessn.model:dessn-model-package}\label{dessn.model::doc}

\subsubsection{Subpackages}
\label{dessn.model:subpackages}

\paragraph{dessn.model.nodes package}
\label{dessn.model.nodes:dessn-model-nodes-package}\label{dessn.model.nodes::doc}

\subparagraph{Submodules}
\label{dessn.model.nodes:submodules}

\subparagraph{dessn.model.nodes.cosmology module}
\label{dessn.model.nodes:dessn-model-nodes-cosmology-module}\label{dessn.model.nodes:module-dessn.model.nodes.cosmology}\index{dessn.model.nodes.cosmology (module)}\index{Cosmology (class in dessn.model.nodes.cosmology)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.cosmology.Cosmology}\pysigline{\strong{class }\code{dessn.model.nodes.cosmology.}\bfcode{Cosmology}}
Bases: {\hyperref[dessn.model:dessn.model.node.Node]{\emph{\code{dessn.model.node.Node}}}}
\index{get\_name() (dessn.model.nodes.cosmology.Cosmology method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.cosmology.Cosmology.get_name}\pysiglinewithargsret{\bfcode{get\_name}}{}{}
\end{fulllineitems}

\end{fulllineitems}

\index{FlatWCDM (class in dessn.model.nodes.cosmology)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.cosmology.FlatWCDM}\pysigline{\strong{class }\code{dessn.model.nodes.cosmology.}\bfcode{FlatWCDM}}
Bases: {\hyperref[dessn.model.nodes:dessn.model.nodes.cosmology.Cosmology]{\emph{\code{dessn.model.nodes.cosmology.Cosmology}}}}

\end{fulllineitems}

\subparagraph{dessn.model.nodes.supernova module}
\label{dessn.model.nodes:dessn-model-nodes-supernova-module}\label{dessn.model.nodes:module-dessn.model.nodes.supernova}\index{dessn.model.nodes.supernova (module)}\index{Supernova (class in dessn.model.nodes.supernova)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.supernova.Supernova}\pysigline{\strong{class }\code{dessn.model.nodes.supernova.}\bfcode{Supernova}}
Bases: {\hyperref[dessn.model:dessn.model.node.Node]{\emph{\code{dessn.model.node.Node}}}}

Abstract supernova which others must implement
\index{type (dessn.model.nodes.supernova.Supernova attribute)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.supernova.Supernova.type}\pysigline{\bfcode{type}}
\end{fulllineitems}

\end{fulllineitems}

\index{SupernovaIa (class in dessn.model.nodes.supernova)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.supernova.SupernovaIa}\pysiglinewithargsret{\strong{class }\code{dessn.model.nodes.supernova.}\bfcode{SupernovaIa}}{\emph{log\_luminosity}, \emph{sigma\_luminosity}}{}
Bases: \code{object}

Models a type Ia supernova

Models the luminosity distribution of type Ia supernovas statically (without
internal parameters).
\begin{gather}
\begin{split}P(L) \sim N(\end{split}\notag
\end{gather}\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{log\_luminosity} : float
\begin{quote}

Represented by the variable \(\mu\)
\end{quote}

\textbf{sigma\_luminosity} : float
\begin{quote}

Represented by the variable \(\sigma\)
\end{quote}

\end{description}\end{quote}
\index{get\_luminosity\_prob() (dessn.model.nodes.supernova.SupernovaIa method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.supernova.SupernovaIa.get_luminosity_prob}\pysiglinewithargsret{\bfcode{get\_luminosity\_prob}}{\emph{log\_luminosity}}{}
\end{fulllineitems}

\index{type() (dessn.model.nodes.supernova.SupernovaIa method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.supernova.SupernovaIa.type}\pysiglinewithargsret{\bfcode{type}}{}{}
\end{fulllineitems}

\end{fulllineitems}

\subparagraph{dessn.model.nodes.typeProb module}
\label{dessn.model.nodes:dessn-model-nodes-typeprob-module}\label{dessn.model.nodes:module-dessn.model.nodes.typeProb}\index{dessn.model.nodes.typeProb (module)}\index{TypeProbability (class in dessn.model.nodes.typeProb)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.TypeProbability}\pysigline{\strong{class }\code{dessn.model.nodes.typeProb.}\bfcode{TypeProbability}}
Bases: {\hyperref[dessn.model:dessn.model.node.Node]{\emph{\code{dessn.model.node.Node}}}}

Abstract type probability node, from which all implementations should inherit
\index{get\_name() (dessn.model.nodes.typeProb.TypeProbability method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.TypeProbability.get_name}\pysiglinewithargsret{\bfcode{get\_name}}{}{}
\end{fulllineitems}

\end{fulllineitems}

\index{TypeProbabilitySimple (class in dessn.model.nodes.typeProb)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.TypeProbabilitySimple}\pysiglinewithargsret{\strong{class }\code{dessn.model.nodes.typeProb.}\bfcode{TypeProbabilitySimple}}{\emph{relative\_rate=0.333}}{}
Bases: {\hyperref[dessn.model.nodes:dessn.model.nodes.typeProb.TypeProbability]{\emph{\code{dessn.model.nodes.typeProb.TypeProbability}}}}

The Type probability node.

Takes \emph{some information} to determine the probability of
of the object in question being a type of supernova.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{relative\_rate} : Optional{[}str{]}
\begin{quote}

Relative rate of SnIa / SnII
\end{quote}

\end{description}\end{quote}

\end{fulllineitems}

\index{Types (class in dessn.model.nodes.typeProb)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.Types}\pysigline{\strong{class }\code{dessn.model.nodes.typeProb.}\bfcode{Types}}
Bases: \code{enum.Enum}

Possible target types
\index{snII (dessn.model.nodes.typeProb.Types attribute)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.Types.snII}\pysigline{\bfcode{snII}\strong{ = \textless{}Types.snII: 2\textgreater{}}}
\end{fulllineitems}

\index{snIa (dessn.model.nodes.typeProb.Types attribute)}

\begin{fulllineitems}
\phantomsection\label{dessn.model.nodes:dessn.model.nodes.typeProb.Types.snIa}\pysigline{\bfcode{snIa}\strong{ = \textless{}Types.snIa: 1\textgreater{}}}
\end{fulllineitems}

\end{fulllineitems}

\subparagraph{Module contents}
\label{dessn.model.nodes:module-contents}\label{dessn.model.nodes:module-dessn.model.nodes}\index{dessn.model.nodes (module)}

\subsubsection{Submodules}
\label{dessn.model:submodules}

\subsubsection{dessn.model.abstracts module}
\label{dessn.model:module-dessn.model.abstracts}\label{dessn.model:dessn-model-abstracts-module}\index{dessn.model.abstracts (module)}

\subsubsection{dessn.model.model module}
\label{dessn.model:dessn-model-model-module}\label{dessn.model:module-dessn.model.model}\index{dessn.model.model (module)}\index{Model (class in dessn.model.model)}

\begin{fulllineitems}
\phantomsection\label{dessn.model:dessn.model.model.Model}\pysigline{\strong{class }\code{dessn.model.model.}\bfcode{Model}}
Bases: \code{object}

\end{fulllineitems}

\subsubsection{dessn.model.node module}
\label{dessn.model:module-dessn.model.node}\label{dessn.model:dessn-model-node-module}\index{dessn.model.node (module)}\index{Node (class in dessn.model.node)}

\begin{fulllineitems}
\phantomsection\label{dessn.model:dessn.model.node.Node}\pysigline{\strong{class }\code{dessn.model.node.}\bfcode{Node}}
Bases: \code{object}
\index{add\_dependency() (dessn.model.node.Node method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model:dessn.model.node.Node.add_dependency}\pysiglinewithargsret{\bfcode{add\_dependency}}{\emph{dependency\_class}}{}
\end{fulllineitems}

\index{get\_dependencies() (dessn.model.node.Node method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model:dessn.model.node.Node.get_dependencies}\pysiglinewithargsret{\bfcode{get\_dependencies}}{}{}
\end{fulllineitems}

\index{get\_name() (dessn.model.node.Node method)}

\begin{fulllineitems}
\phantomsection\label{dessn.model:dessn.model.node.Node.get_name}\pysiglinewithargsret{\bfcode{get\_name}}{}{}
\end{fulllineitems}

\end{fulllineitems}

\subsubsection{Module contents}
\label{dessn.model:module-dessn.model}\label{dessn.model:module-contents}\index{dessn.model (module)}

\subsection{dessn.simple package}
\label{dessn.simple::doc}\label{dessn.simple:dessn-simple-package}

\subsubsection{Submodules}
\label{dessn.simple:submodules}

\subsubsection{dessn.simple.exampleIntegral module}
\label{dessn.simple:dessn-simple-exampleintegral-module}\label{dessn.simple:module-dessn.simple.exampleIntegral}\index{dessn.simple.exampleIntegral (module)}\index{ExampleIntegral (class in dessn.simple.exampleIntegral)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral}\pysiglinewithargsret{\strong{class }\code{dessn.simple.exampleIntegral.}\bfcode{ExampleIntegral}}{\emph{n=900}, \emph{theta\_1=100.0}, \emph{theta\_2=30.0}}{}
Bases: \code{object}

An example implementation using integration over a latent parameter.

Let us assume that we are observing supernova that a drawn from an underlying
supernova distribution parameterised by \(\theta\),
where the supernova itself simply a luminosity \(L\). We measure the luminosity
of multiple supernovas, giving us an array of measurements \(D\). If we wish to recover
the underlying distribution of supernovas from our measurements, we wish to find \(P(\theta|D)\),
which is given by
\begin{gather}
\begin{split}P(\theta|D) \propto P(D|\theta)P(\theta)\end{split}\notag
\end{gather}
Note that in the above equation, we realise that \(P(D|L) = \prod_{i=1}^N P(D_i|L_i)\) as our measurements are
independent. The likelihood \(P(D|\theta)\) is given by
\begin{gather}
\begin{split}P(D|\theta) =  \prod_{i=1}^N  \int_{-\infty}^\infty P(D_i|L_i) P(L_i|\theta) dL_i\end{split}\notag
\end{gather}
We now have two distributions to characterise. Let us assume both are gaussian, that is
our observed luminosity \(x_i\) has gaussian error \(\sigma_i\) from the actual supernova
luminosity, and the supernova luminosity is drawn from an underlying gaussian distribution
parameterised by \(\theta\).
\begin{quote}
\begin{gather}
\begin{split}P(D_i|L_i) = \frac{1}{\sqrt{2\pi}\sigma_i}\exp\left(-\frac{(x_i-L_i)^2}{\sigma_i^2}\right)\end{split}\notag\\\begin{split}P(L_i|\theta) = \frac{1}{\sqrt{2\pi}\theta_2}\exp\left(-\frac{(L_i-\theta_1)^2}{\theta_2^2}\right)\end{split}\notag
\end{gather}\end{quote}

This gives us a likelihood of
\begin{gather}
\begin{split}P(D|\theta) = \prod_{i=1}^N  \frac{1}{2\pi \theta_2 \sigma_i}  \int_{-\infty}^\infty
\exp\left(-\frac{(x_i-L_i)^2}{\sigma_i^2} -\frac{(L_i-\theta_1)^2}{\theta_2^2} \right) dL_i\end{split}\notag
\end{gather}
Working in log space for as much as possible will assist in numerical precision, so we can rewrite this as
\begin{gather}
\begin{split}\log\left(P(D|\theta)\right) =  \sum_{i=1}^N  \left[
        \log\left( \int_{-\infty}^\infty \exp\left(-\frac{(x_i-L_i)^2}{\sigma_i^2} -
\frac{(L_i-\theta_1)^2}{\theta_2^2} \right) dL_i \right) -\log(2\pi\theta_2\sigma_i) \right]\end{split}\notag
\end{gather}
Creating this class will set up observations from an underlying distribution.
Invoke \code{emcee} by calling the object. Notice that performing the marginalisation over
\(dL_i\) requires computing \(n\) integrals for each step in the MCMC.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{n} : int, optional
\begin{quote}

The number of supernova to `observe'
\end{quote}

\textbf{theta\_1} : float, optional
\begin{quote}

The mean of the underlying supernova luminosity distribution
\end{quote}

\textbf{theta\_2} : float, optional
\begin{quote}

The standard deviation of the underlying supernova luminosity distribution
\end{quote}

\end{description}\end{quote}
\index{do\_emcee() (dessn.simple.exampleIntegral.ExampleIntegral method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral.do_emcee}\pysiglinewithargsret{\bfcode{do\_emcee}}{\emph{nwalkers=20}, \emph{nburn=500}, \emph{nsteps=3000}}{}
Run the \emph{emcee} chain and produce a corner plot.

Saves a png image of the corner plot to plots/exampleIntegration.png.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{nwalkers} : int, optional
\begin{quote}

The number of walkers to use. Minimum of four.
\end{quote}

\textbf{nburn} : int, optional
\begin{quote}

The burn in period of the chains.
\end{quote}

\textbf{nsteps} : int, optional
\begin{quote}

The number of steps to run
\end{quote}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_likelihood() (dessn.simple.exampleIntegral.ExampleIntegral method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral.get_likelihood}\pysiglinewithargsret{\bfcode{get\_likelihood}}{\emph{theta}, \emph{data}, \emph{error}}{}
Gets the log likelihood given the supplied input parameters.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{theta} : array of size 2
\begin{quote}

An array representing \([\theta_1,\theta_2]\)
\end{quote}

\textbf{data} : array of length \emph{n}
\begin{quote}

An array of observed luminosities
\end{quote}

\textbf{error} : array of length \emph{n}
\begin{quote}

An array of observed luminosity errors
\end{quote}

\item[{Returns}] \leavevmode
float
\begin{quote}

the log likelihood probability
\end{quote}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_posterior() (dessn.simple.exampleIntegral.ExampleIntegral method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral.get_posterior}\pysiglinewithargsret{\bfcode{get\_posterior}}{\emph{theta}, \emph{data}, \emph{error}}{}
Gives the log posterior probability given the supplied input parameters.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{theta} : array of size 2
\begin{quote}

An array representing \([\theta_1,\theta_2]\)
\end{quote}

\textbf{data} : array of length \emph{n}
\begin{quote}

An array of observed luminosities
\end{quote}

\textbf{error} : array of length \emph{n}
\begin{quote}

An array of observed luminosity errors
\end{quote}

\item[{Returns}] \leavevmode
float
\begin{quote}

the log posterior probability
\end{quote}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_prior() (dessn.simple.exampleIntegral.ExampleIntegral method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral.get_prior}\pysiglinewithargsret{\bfcode{get\_prior}}{\emph{theta}}{}
Get the log prior probability given the input.

The prior distribution is currently implemented as flat prior.
\begin{quote}\begin{description}
\item[{Parameters}] \leavevmode
\textbf{theta} : array of size 2
\begin{quote}

An array representing \([\theta_1,\theta_2]\)
\end{quote}

\item[{Returns}] \leavevmode
float
\begin{quote}

the log prior probability
\end{quote}

\end{description}\end{quote}

\end{fulllineitems}

\index{plot\_observations() (dessn.simple.exampleIntegral.ExampleIntegral method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simple:dessn.simple.exampleIntegral.ExampleIntegral.plot_observations}\pysiglinewithargsret{\bfcode{plot\_observations}}{}{}
Plot the observations and observation distribution.

\end{fulllineitems}

\end{fulllineitems}

\subsubsection{Module contents}
\label{dessn.simple:module-contents}\label{dessn.simple:module-dessn.simple}\index{dessn.simple (module)}

\subsection{dessn.simulation package}
\label{dessn.simulation:dessn-simulation-package}\label{dessn.simulation::doc}

\subsubsection{Submodules}
\label{dessn.simulation:submodules}

\subsubsection{dessn.simulation.observationFactory module}
\label{dessn.simulation:dessn-simulation-observationfactory-module}\label{dessn.simulation:module-dessn.simulation.observationFactory}\index{dessn.simulation.observationFactory (module)}\index{ObservationFactory (class in dessn.simulation.observationFactory)}

\begin{fulllineitems}
\phantomsection\label{dessn.simulation:dessn.simulation.observationFactory.ObservationFactory}\pysiglinewithargsret{\strong{class }\code{dessn.simulation.observationFactory.}\bfcode{ObservationFactory}}{\emph{**kwargs}}{}
Bases: \code{object}
\index{check\_kwargs() (dessn.simulation.observationFactory.ObservationFactory method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simulation:dessn.simulation.observationFactory.ObservationFactory.check_kwargs}\pysiglinewithargsret{\bfcode{check\_kwargs}}{}{}
\end{fulllineitems}

\index{get\_observations() (dessn.simulation.observationFactory.ObservationFactory method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simulation:dessn.simulation.observationFactory.ObservationFactory.get_observations}\pysiglinewithargsret{\bfcode{get\_observations}}{\emph{num}}{}
Still needs massive refactoring

\end{fulllineitems}

\end{fulllineitems}

\subsubsection{dessn.simulation.simulation module}
\label{dessn.simulation:module-dessn.simulation.simulation}\label{dessn.simulation:dessn-simulation-simulation-module}\index{dessn.simulation.simulation (module)}\index{Simulation (class in dessn.simulation.simulation)}

\begin{fulllineitems}
\phantomsection\label{dessn.simulation:dessn.simulation.simulation.Simulation}\pysigline{\strong{class }\code{dessn.simulation.simulation.}\bfcode{Simulation}}
Bases: \code{object}
\index{get\_simulation() (dessn.simulation.simulation.Simulation method)}

\begin{fulllineitems}
\phantomsection\label{dessn.simulation:dessn.simulation.simulation.Simulation.get_simulation}\pysiglinewithargsret{\bfcode{get\_simulation}}{\emph{num\_trans=30}}{}
\end{fulllineitems}

\end{fulllineitems}

\subsubsection{Module contents}
\label{dessn.simulation:module-contents}\label{dessn.simulation:module-dessn.simulation}\index{dessn.simulation (module)}

\section{Module contents}
\label{dessn:module-dessn}\label{dessn:module-contents}\index{dessn (module)}

\renewcommand{\indexname}{Python Module Index}
\begin{theindex}
\def\bigletter#1{{\Large\sffamily#1}\nopagebreak\vspace{1mm}}
\bigletter{d}
\item {\texttt{dessn}}, \pageref{dessn:module-dessn}
\item {\texttt{dessn.entry}}, \pageref{dessn.entry:module-dessn.entry}
\item {\texttt{dessn.entry.sim}}, \pageref{dessn.entry:module-dessn.entry.sim}
\item {\texttt{dessn.model}}, \pageref{dessn.model:module-dessn.model}
\item {\texttt{dessn.model.abstracts}}, \pageref{dessn.model:module-dessn.model.abstracts}
\item {\texttt{dessn.model.model}}, \pageref{dessn.model:module-dessn.model.model}
\item {\texttt{dessn.model.node}}, \pageref{dessn.model:module-dessn.model.node}
\item {\texttt{dessn.model.nodes}}, \pageref{dessn.model.nodes:module-dessn.model.nodes}
\item {\texttt{dessn.model.nodes.cosmology}}, \pageref{dessn.model.nodes:module-dessn.model.nodes.cosmology}
\item {\texttt{dessn.model.nodes.supernova}}, \pageref{dessn.model.nodes:module-dessn.model.nodes.supernova}
\item {\texttt{dessn.model.nodes.typeProb}}, \pageref{dessn.model.nodes:module-dessn.model.nodes.typeProb}
\item {\texttt{dessn.simple}}, \pageref{dessn.simple:module-dessn.simple}
\item {\texttt{dessn.simple.exampleIntegral}}, \pageref{dessn.simple:module-dessn.simple.exampleIntegral}
\item {\texttt{dessn.simulation}}, \pageref{dessn.simulation:module-dessn.simulation}
\item {\texttt{dessn.simulation.observationFactory}}, \pageref{dessn.simulation:module-dessn.simulation.observationFactory}
\item {\texttt{dessn.simulation.simulation}}, \pageref{dessn.simulation:module-dessn.simulation.simulation}
\end{theindex}

\renewcommand{\indexname}{Index}
\printindex
\end{document}

### 2016-03-31 doc/toy.tex
add spectroscopic efficiency

@@
 \frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
-&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}\\
-&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
+&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.%\\
+%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
 \end{align}
+
+The photometric and spectroscopic sample selections can be treated as efficiencies $\epsilon_c$, $\epsilon_T$ as
+\begin{align}
+p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) & = \int ds\, p(c_o, s, S_c, S_T | T=T_o, L, z=z_o, X) \\
+ & =p(c_o, S_c | T=T_o, L, z=z_o, X)  \int ds\,  p(s, S_T | T=T_o, L, z=z_o, X) \\
+ & = \epsilon_c(c_o) p(c_o| T=T_o, L, z=z_o, X)  \int ds\, \epsilon_T(s)  p(s | T=T_o, L, z=z_o, X) \\
+ & \approx \epsilon_c(c_o)   \epsilon_T(s_\star) p(c_o| T=T_o, L, z=z_o, X).
+\end{align}
+where the supernova signal in the spectrum $s$ is not extracted and we make the approximation $ p(s | T=T_o, L, z=z_o, X)  = \delta(s-s_\star)$
+and $s_\star(t_T; T=T_o, L, z=z_o, X)$ is the expected signal.
 
@@
 %& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right].
+& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right] \\
+& \approx  \sum_i   P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[  \epsilon_T(s_\star)   \int dc_o\, \epsilon_c(c_o)  p(c_o| T=T_o, L, z=z_o, X)  \right] .
 \end{align}

### 2016-05-08 tex/model.tex

\documentclass[preprint]{aastex}
\usepackage{amsmath,amssymb}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{bm}
\bibliographystyle{apj}

\newcommand{\mli}[1]{\mathit{#1}}
%\usepackage{epstopdf}

\begin{document}

\title{Model for DES Type Ia Supernova Cosmology Analysis}
\author{Alex Kim}

\section{Model}
We infer a Hubble diagram based on measurements of discovered transients
that have passed some selection criteria (e.g.\ classified as Type~Ia supernovae).  The posterior
of the Hubble diagram is denoted as
\begin{equation}
p({\mu},{z} |  {{ADU}}, {{T}}_S,{{z}}_S,
{{\theta}}_S, \theta_G, \tau).
\label{hd:eqn}
\end{equation}
The set of distance moduli
and redshifts $(\mu, z)$ contains the points that constitute the Hubble diagram.  The
data
from which the Hubble diagram is inferred include:
the transient photometry ${ADU}$; spectral measurements of
transient
type, redshift and SN-parameters ${T}_S$, ${z}_S$, ${\theta}_S$;
host-galaxy parameter measurements $\theta_H$ (including redshift,
mass, sSFR, etc.).  
The subset transients that enter the cosmology
analysis is based on the sample-selection criteria $\tau$: for example
the classification score from PSNID (a classifier based on multi-band
light curves).

The posterior of $\mu$ and $z$ is an efficient way
to distribute supernova results in is a way that is independent of cosmological model,
facilitating joint-probe analysis or consideration of non-standard models.  Choosing
a specific cosmological model, the posterior of the model parameters $\theta_\mu$
is of interest.  The model presented here accommodates both options.

A sketch of a more extensive model that includes the
model of Equation~\ref{hd:eqn} is depicted in the Probabilistic Graphical Model
(PGM)
shown in Figure~\ref{pgm:fig}.
The fundamental parameters are those with no incoming arrows;  most are
used to model the universe: 
the background cosmology, transients that could enter the sample, and their host galaxies.
\begin{itemize}
\item  $g$ labels the  galaxy that hosts the transient.
\item $\theta_\mu$ are the parameters that specify that distance modulus.  They could
be the parameters of a specific cosmological
parameters, or an empirical model independent of physics.
\item $\theta_\tau$ govern the relative rates of supernovae of different types in galaxies.
\item $\theta_T^{Ia}$, $\theta_{Ti}^{Ia}$.  The model parameters for SNe~Ia, including
parameters that describe distributions.
Different SN~Ia models could be considered.  The treatment of SALT2
has been presented in \citet{2011MNRAS.418.2308M}:
in this model each supernova
has individual $c$ and $x_1$ parameters that constitute $\theta_{Ti}^{Ia}$;
global SN~Ia properties
$\alpha$, $\beta$, $\sigma_{int}$ 
 and  underlying  $c$ and $x_1$ distributions   constitute $\theta_T^{Ia}$.
\item $\theta_T^{\mathit{non-Ia}}$, $\theta_{Ti}^{\mathit{non-Ia}}$.  False-positive non-Ia transients
that enter the analysis are described by a parameterized model.
\item The transmission functions $\phi(\lambda)$ describe the optical path from the
atmosphere to counts.  Calibration solutions are generally provided
but are included in Figure~\ref{pgm:fig} to encompass the case of self-calibration.
\end{itemize}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=7in]{../results/hdpgm.pdf} 
   \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
   The forward modeling
   flow goes from left to right, starting with the description of the universe, observatory,
   and data.    A transient has host galaxy $g_i$, which determines its redshift $z_i$
   and galaxy parameters $\theta_{gi}$.
   A model with parameters $\theta_\mu$ fix the distance modulus $\mu_i$.
   The transient type $T$ depends on the host-galaxy parameters  $\theta_{gi}$
   and rates $\theta_r$.   The transient
   parameters $\theta_T^X$, $\theta_{Ti}^X$ (and perhaps the host galaxy) determine the luminosity $L$.       The 
   incoming photon flux $n$, $n_g$  are then fixed
   with redshift and distance modulus.
   The instrumental transmission function $\phi$ is calibrated with data ${Z}$ and
   gives the expected
   counts $\overline{\mathit{ADU}}$, $\overline{\mathit{ADU}}_g$. 
   The transient has realized light curves (${ADU}$) and passes selection criteria
   through $\tau_i$ to be included in the sample.  Some transients have spectral data
   ${X}_S$.  The properties of the galaxy identified as the host galaxy is $\theta_G$. 
   \label{pgm:fig}}
\end{figure}

From these fundamental parameters we can determine the expected
signals
through a series of intermediary parameters,
some of which are deterministic (points in Figure~\ref{pgm:fig}) and others
probabilistic (ovals).
\begin{itemize}
\item $z_i$, $\theta_{gi}$ are the galaxy redshift and parameters (e.g.\ mass, sSFR, metallicity).
Each galaxy $g_i$ has its values for these quantities.
\item $\mu_i=\mu(z_i; \theta_\mu)$.  The distance modulus is a function of redshift,
Unspecified at the moment,  it could
be a $\Lambda$CDM prediction with cosmological parameters, a spline with knot values
as parameters.
The redshift of the galaxy is used in the Hubble diagram, peculiar velocities are not considered.
\item $P(T_i | \theta_r, \theta_{gi})$.  The transient type $T_i$, SN~Ia or non-Ia, depends
on the relative rates given the host properties.
\item $P(L_i(t,\lambda)| T_i, \theta_T^{Ia}, \theta_{Ti}^{Ia}, \theta_T^{\mathit{non-Ia}}, \theta_{Ti}^{\mathit{non-Ia}},
\theta_{gi})$.  The source model determines
the luminosity; the model may include intrinsic dispersion to make the luminosity
probabilistic. The  model includes  information on the
source and line-of-sight effects, which specify the effective SED.   Only the
parameters of the model of the appropriate type $T$ are relevant.  The galaxy parameters
$\theta_{gi}$ are included if correlations with host-galaxy properties are to be modeled.
\item $n_i(t,\lambda; L_i, \mu_i, z_i)$, $n_{gi}(t; g_i, \mu_i, z_i)$.  The  fluxes of
the transient and host that are incident at Earth
are fixed by the luminosity, distance modulus, and redshift.
The supernova redshift is required for this calculation,
for simplicity the supernova and host-galaxy redshift are assumed to be equal.
\item $\overline{\mathit{ADU}}$ and
$\overline{\mathit{ADU}}_g$ are the expected counts of the transient and galaxy respectively,
coming from $\int n_i \phi d\lambda$.  
\end{itemize}

The observations are as follows:
\begin{itemize}
\item ${\mathit{ADU}}$ is the realized flux.  It depends on the underlying flux and
the photometric noise.
\item $\theta_G$ represents the galaxy parameters of the
galaxy catalog used to identify prospective host galaxies.
\item ${T}_S$, ${z}_S$, ${\theta}_S$ are the transient type, redshift and properties from
spectroscopic data. Details of the actual measurement are ignored and a direct association
is made with the fluxes of the transient and underlying galaxy $n$, $n_g$.
\item $\tau$ is the score used to decide if the transient is used in the analysis.  It is
determined based on other measurements $\tau(\mathit{ADU},T_S)$.
\item ${Z}$ represents the calibration measurements.   Although this may be given
by the Calibration Task Force, the PGM is written for the more general case where the calibration
is modeled within our analysis.
\end{itemize}

%The likelihood for the model is
%\begin{multline}
%p(\mathit{ADU}, {{T}}_S,{{z}}_S,
%{{\theta}}_S, \theta_G, \tau, Z | g, \theta_\mu, \theta_r, \theta_T^{Ia}, \theta_{Ti}^{Ia}, \theta_T^{non-Ia}, \theta_{Ti}^{non-Ia}, \phi) = \\
%p(\theta_G|g)p(Z|\phi)p(\mathit{ADU}, {{T}}_S,{{z}}_S,
%{{\theta}}_S, \tau | g, \theta_\mu, \theta_r, \theta_T^{Ia}, \theta_{Ti}^{Ia}, \theta_T^{non-Ia}, \theta_{Ti}^{non-Ia}, \phi).
%\label{likelihood:eqn}
%\end{multline}

\section{Simple Example}
\subsection{Model}
\label{example:sec}
A simplified version of the above model,
\begin{equation}
p(\mathit{ADU}, {{T}}_S,{{z}}_S, \theta_G|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}),
\end{equation}
is examined in this section and is sketched in Figure~\ref{toypgm:fig}.
The measurements are
\begin{itemize}
\item $\mathit{ADU}$: the measured counts at peak brightness;
\item ${{T}}_S$, ${{z}}_S$: the spectroscopic type and redshift for a subset of transients;
\item $\theta_{G1}$, $\theta_{G2}$: the redshifts of the identified host and neighbor
galaxies.
\end{itemize}
The model parameters are
\begin{itemize}
\item $\Omega_M$, $w$: the parameters of a flat dark-energy cosmology;
\item $\theta_{r1}$, $\theta_{r2}$:  for two redshifts the fraction of transients that are SN~Ia;
\item $\alpha_{Ia}$, $\sigma_{Ia}$, $\alpha_{\mathit{non-Ia}}$, $\sigma_{\mathit{non-Ia}}$:
the luminosity (times $4\pi$) and intrinsic dispersion for SNe~Ia and non-Ia at peak.
Light-curve shape and color are not considered in this simple model.
\end{itemize}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5in]{../results/toy_pgm.pdf} 
   \caption{Probabilistic Graphical Model for the simple SN~Ia analysis. The
   parameters include the cosmological parameters $\Omega_M$ and $w$,
   the luminosities $\alpha$ and intrinsic dispersions $\sigma$ of SNe~Ia and non-Ia's,
   and the relative rate parameters $\theta_{r1}$ and $\theta_{r2}$.  Latent parameters
   include the type $T$ and redshifts of the host and neighboring galaxies $z$ and $z_N$
   for each transient.  The  parameters for distance modulus $\mu$ and
   luminosity $L$ are fixed by the other parameters.  The observables include the measured
   flux $\mathit{ADU}$, spectroscopic type and redshift $T_S$ and $z_S$, and
   the inferred host and neighbor redshifts $\theta_{G1}$ and $\theta_{G2}$.
   \label{toypgm:fig}}
\end{figure}

There are latent model parameters for each transient used internally
by the model:
\begin{itemize}
\item $z$: the cosmological redshifts.
\item $T$: the type, SN~Ia or non-Ia.
\end{itemize}

The transients are divided into samples defined by available data: {\it obs} are transients observed spectroscopically while active, from which $T_S=1$ are those typed
as SN~Ia and $T_S=0$ as non-Ia.  {\it mis} transients are not observed
spectroscopically when active.  The  {\it obs} and {\it mis} samples
are assumed to be drawn from the same underlying distribution, so that 
same SN~Ia and non-Ia parameters apply for both sets.

\subsubsection{Model for the {\it obs} set data}
The modeling of data from the {\it obs} set, which has
spectroscopic observations $T_S$ and redshift
$z_S$, is described as follows.  We are interested in
\begin{align}
p(\mathit{ADU}, {{T}}_S,{{z}}_S|  \Omega_M, w, \theta_r,\alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  &=\sum_T \int_z dz\,
p(\mathit{ADU}, {{T}}_S,{{z}}_S, T, z| \ldots)\\
&= \sum_T \int_z dz\,
p(\mathit{ADU}, {{T}}_S,{{z}}_S| T, z,\dots) p(T,z | \ldots)
\end{align}

A direct association of spectroscopic redshift and type are made so that
\begin{align}
p(z_S|z) &= \delta(z_S-z)\\
p(T_S|T) &= \delta(T_S-T),
\label{specz:eqn}
\end{align}
so that
\begin{align}
p(\mathit{ADU}, {{T}}_S,{{z}}_S|  \ldots) &= 
p(\mathit{ADU}| T=T_S, z=z_s,\dots) P(T_S| z= z_S , \ldots) p(z_S|\ldots).
\label{obs:eqn}
\end{align}
We remark that in a flux limited survey, the pdf's for the data are truncated versions of the universal
pdf's of the observables.

The first term describes the expected counts. Given the redshift
and transient type, the brightness of the underlying population
is modeled as being drawn from a standard candle
with luminosity $4\pi\alpha_X$ and intrinsic magnitude dispersion  $\sigma_X$:
\begin{equation}
\mathit{ADU}| T_S, z_S, \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{T_S}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{T_S}\right),
\end{equation}
where $\mathcal{N}_{\ln}$ is the log-normal distribution and $d_L$ is the luminosity distance.

The subset of objects that enters our analysis comes from a truncated distribution:
\begin{itemize}
\item For transients in set $T_S=1$:
\begin{equation}
\mathit{ADU} | T_S=1, z_S, \Omega_M, w, \alpha_{Ia},\sigma_{Ia} \sim
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{Ia}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{Ia}\right)}{P_{Ia}(\mathit{ADU} > \mathit{ADU}_0)}.
\label{adusnIa:eqn}
\end{equation}
\item For transients in set $T_S=0$:
\begin{equation}
\mathit{ADU} | T_S=0, z_S, \Omega_M, w, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}}\sim 
\frac{\mathcal{N}_{\ln}\left(\ln{\left(\frac{\alpha_{\mathit{non-Ia}}}{d_L^2(z_S;\Omega_M, w)}\right)}, \frac{\ln{10}}{2.5}\sigma_{\mathit{non-Ia}}\right)}{P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)},
\label{adunonIa:eqn}
\end{equation}
\end{itemize}
where the normalization factor is the complementary cumulative distribution function
for the corresponding log-normal distribution.
It is understood that $p(\mathit{ADU}<\mathit{ADU}_0)=0$ is never evaluated.

The second term gives the type probability. In our model,
the underlying relative rates between SNe~Ia and non-Ia is linear in redshift,
 parameterized by the
its value at $z=0$ and $1.1 \times z_{max}$, 
\begin{equation}
\hat{\theta}_{r}=\theta_{r0}+z\left(\frac{\theta_{r1}-\theta_{r0}}{1.1 \times z_{max}}\right).
\end{equation}
The relative rate for SN~Ia discovery for the truncated distribution is
\begin{equation}
\theta_{r}=\frac{\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0)}{\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0) + (1-\hat{\theta}_{r})P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)}.
\end{equation}
The spectroscopic classifications are drawn from
\begin{equation}
T_S | \Omega_M, w, \theta_{r0}, \theta_{r1} , \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}} \sim \text{Bernoulli}(\theta_r).
\end{equation}
Although the underlying type distribution depends only on $ \theta_{r0}$ and $ \theta_{r1}$,  
the observed truncated distribution does depend on the other parameters through 
$P_{X}(\mathit{ADU} > \mathit{ADU}_0)$.

The third term is the redshift distribution of transient discoveries.
In our model we adopt
a redshift distribution of transients in the universe
\begin{equation}
P(z) \propto z^2.
\end{equation}
The redshift distribution of discovered transients is then
\begin{equation}
P(z) = \left(\int_{z_{min}}^{z_{max}} dz'\, z'^2\left( 
\hat{\theta}_{r}P_{Ia}(\mathit{ADU} > \mathit{ADU}_0) + (1-\hat{\theta}_{r})P_{non-Ia}(\mathit{ADU} > \mathit{ADU}_0)
\right)\right)^{-1}z^2.
\end{equation}

In the presented model, the data from different transients are independent.

\subsubsection{Model for the {\it mis} set data}

We now consider the modeling of data from the  {\it mis}
set that has no spectroscopic confirmation.
The redshift measurement is treated as follows:
For each transient there are two galaxies identified as
the potential host, one ($\theta_{G1}$) with high probability $P_{host}$
of being the host and the other
($\theta_{G2}$) at lower $(1-P_{host})$ probability.
In the model, the
true host redshift is $z$ and that of the interloping galaxy is $z_N$.
\begin{equation}
P(\theta_{G1},\theta_{G2}|z, z_N) =
	P_{host}\delta(z-\theta_{G1})\delta(z_N-\theta_{G2}) +
	(1-P_{host}) \delta(z-\theta_{G2})\delta(z_N-\theta_{G1}).
\end{equation}
As for the {\it obs} set, the redshifts of the supernova
 and neighbor are drawn from distributions $P(z) \propto z^2$, $P(z_N) \propto z^2_N$.

The data likelihood is 
\begin{multline}
p(\mathit{ADU}, \theta_{G1}, \theta_{G2} | \Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})  \\
= \sum_{T}\int_{z,z_N} dz\,dz_N \, p(\mathit{ADU, \theta_{G1}, \theta_{G2}}| z , z_N, T, \ldots)P(T| z, z_N,\ldots) p(z, z_N|\ldots)\\
 =\sum_{T} \int_{z,z_N} dz\,dz_N\, p(\mathit{ADU}| z ,T, \ldots)p(\theta_{G1}, \theta_{G2}| z , z_N) P(T| z, \ldots) p(z, z_N|\ldots)\\
= \sum_{T} P_{host} p(\mathit{ADU}| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) p(\theta_{G1},\theta_{G2}|\ldots) \\
 +  (1-P_{host}) p(\mathit{ADU}| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots) p(\theta_{G2},
 \theta_{G1}|\ldots)\\
\propto \sum_{T} P_{host} p(\mathit{ADU}| z=\theta_{G1} ,T, \ldots) P(T| z=\theta_{G1}, \ldots) \\
 +  (1-P_{host}) p(\mathit{ADU}| z=\theta_{G2} ,T, \ldots) P(T| z=\theta_{G2}, \ldots),
\label{adumis:eqn}
\end{multline}
where we take advantage of the fact that
$p(\theta_{G1}, \theta_{G2} | \ldots) = p(\theta_{G2}, \theta_{G1} | \ldots)$ does
not have parameter dependence.

The distribution of $\mathit{ADU}$ for our sample is a truncated version of the
underlying distribution.  
For our analysis, we use the distribution in  Eqn.~\ref{adumis:eqn} normalized by
the complementary cumulative distribution function
\begin{equation}
P(\mathit{ADU}>\mathit{ADU}_0, \theta_{G1}, \theta_{G2}|\ldots) = \int_{\mathit{ADU}_0}^{\infty} p(\mathit{ADU}, \theta_{G1}, \theta_{G2}|\Omega_M, w, \theta_r, \alpha_{Ia},\sigma_{Ia}, \alpha_{\mathit{non-Ia}},\sigma_{\mathit{non-Ia}})d\mathit{ADU}
\end{equation}
for a detection threshold of $\mathit{ADU}_0$.

%I highlight the treatment of redshift in this model; the cosmological redshift
%$z$ is set to the values of the
%observed redshifts through $\delta$-functions.  This choice was made specifically to aid
%in the convergence of the Monte Carlo: the discrete possibilities for redshift allow
%for their marginalized likelihood to be used in the calculation of the posterior.
%A more general model would replace the $\delta$-functions
%with coming from a continuous PDF's.
%Peculiar velocities, redshift measurement errors (particularly from
%photometric redshifts) justify this generalization.
%A transient
%without a spectrum has uncertain type and redshift: the redshift
%of the associated host galaxy, and redshifts expected based on the
%flux of the SN~Ia and non-Ia standard candles,
%are viable but disjoint possibilities for the posterior of
%the true underlying redshift. My runs of this scenario produce chains
%that do not mix so I resorted to the simplified treatment presented here.

\subsection{Analysis of Simulated Data}
The simulation of data from the model presented in \S\ref{example:sec}
and their analysis is implemented in Python and are available
on GitHub\footnote{\url{git@github.com:AlexGKim/abc.git}}.

The results that follow are based on the parameter
choices:
\begin{itemize}
\item Cosmological parameters $\Omega_M=0.28$, $w=-1$.
\item Type~Ia supernovae with intrinsic dispersion $\sigma_{SNIa}=0.1$.
\item Non-Ia supernovae 2 magnitudes fainter than SNe~Ia with intrinsic
dispersion $\sigma_{non-Ia}=1$.
\item $\theta_r=r_0 + (r_1-r_0)z/1.54$ for a constant relative fraction of SNe~Ia where $r_1=0.2$,
$r_0=0.95$.
\item 2000 transients distributed by co-moving volume in the range $0.1<z<1.4$.
\item Detection threshold at approximately the mean brightness SN~Ia at $z=1$.
\item $z_{min}=0.1/1.1$ and $z_{max}=1.4\times 1.1$ is the range of possible misassigned
host redshift.
\item $P_{host}=0.98$ for the probability of correct host-galaxy assignment.
\item Accumulated spectra for 0\%, 20\%, 60\%, and 100\% of discovered transients.
\end{itemize}
This setup is referred to as Model 1.

The Hubble diagram for a realization drawn from this model is shown in Figure~\ref{hd:fig}.
Type~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
are faded.  The solid points
represent those objects spectroscopically typed for the case of 20\% follow-up.
The green X's represent the 2\% among those without spectroscopy that
are assigned an incorrect redshift.
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2.pdf}
%   \includegraphics[scale=0.4]{/Users/akim/project/abc/doc/seed2_pop2.pdf}  
%\caption{Hubble diagrams for a realization drawn from the baseline Model 1 (left)
%and Model 2 with a second non-Ia population (right).
%Type~Ia's are shown in blue and non-Ia's in red, those that fall below detection threshold
%are faded.  The green X's represent the 2\% among those without spectroscopy that
%are assigned an incorrect redshift.
%   \label{hd:fig}}
%\end{figure}

The simulated data are analyzed using 
\citet{stan-software:2015},
a probabilistic programming language for
inference of Bayesian models with Hamiltonian Monte Carlo.
In addition to the simulated data,
supplemental priors are applied to the SN~Ia absolute magnitude and intrinsic dispersion
\begin{align}
\alpha_{SNIa} & \sim \mathcal{N}_{\ln}\left(\ln\left(\alpha^0_{SNIa}\right),\frac{\ln{10}}{2.5}0.02\right)\\
\sigma_{SNIa} & \sim \mathcal{N}_{\ln}\left(\ln\left(0.1\right),0.1\right),
\end{align}
where $\alpha^0_{SNIa}$ is the true input luminosity.  This prior is added in lieu of
a pure large SN~Ia sample at low-redshift.

The results of the analyses meet expectations.  As the input data and analyses are
based on the same model, the input parameters lie in reasonable position
within the resultant posterior despite the non-trivial appearance of
the Hubble diagram in Figure~\ref{hd:fig}.
The parameter pdf's are shown in Figure~\ref{contour1:fig} for the ideal case of 100\% spectroscopic classification.  The input values for all 7 parameters fall within the 68\%
credible interval.  The same is true for the unideal case of no
spectroscopic classification.  Table~\ref{seed2:tab} lists the modes
of the $w$ distribution for varying spectroscopic coverage: there is no appreciable
bias relative to the statistical uncertainties.

%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2/{contour.2000.1.0}.pdf}
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2_pop2/{contour.2000.1.0}.pdf} 
%\caption{Probability distribution functions of the model parameters for the case of
% 100\% spectroscopy for Model 1 (left) and Model 2 (right).
%    Solid lines represent the input parameters.  Dashed lines in the
%   histograms are the 0.16 and 0.84 quantiles. 
%   \label{contour1:fig}}
%\end{figure}

\begin{table}
\centering
\begin{tabular}{|c|c|ccc|}
\hline
Fraction Spec&\%-ile & $w$ &$w$ range & $\Delta w$\\ \hline
SN Ia only&&&&\\
$1.00$ & $0.68$ &$-0.97$ & $[-1.076, -0.801]$ & $0.275$ \\
$1.00$& $0.90$ &$-0.97$ & $[-1.166, -0.718]$ & $0.448$ \\
$1.00$ & $0.95$ &$-0.97$ & $[-1.208, -0.682]$ & $0.525$ \\ \hline
Model 1&&&&\\
$1.00$ &$0.68$ &$-1.08$ & $[-1.196, -0.914]$ & $0.281$ \\
$1.00$ &$0.90$ &$-1.08$ & $[-1.292, -0.826]$ & $0.466$ \\
$1.00$ &$0.95$ &$-1.08$ & $[-1.340, -0.784]$ & $0.556$ \\
$0.60$ &$0.68$ &$-1.08$ & $[-1.227, -0.933]$ & $0.293$ \\
$0.60$ &$0.90$ &$-1.08$ & $[-1.331, -0.836]$ & $0.495$ \\
$0.60$ &$0.95$ &$-1.08$ & $[-1.385, -0.787]$ & $0.598$ \\
$0.20$ &$0.68$ &$-1.15$ & $[-1.249, -0.943]$ & $0.305$ \\
$0.20$ &$0.90$ &$-1.15$ & $[-1.354, -0.851]$ & $0.502$ \\
$0.20$ &$0.95$ &$-1.15$ & $[-1.408, -0.810]$ & $0.597$ \\
$0.00$ &$0.68$ &$-1.09$ & $[-1.218, -0.909]$ & $0.309$ \\
$0.00$ &$0.90$ &$-1.09$ & $[-1.320, -0.812]$ & $0.508$ \\
$0.00$ &$0.95$ &$-1.09$ & $[-1.372, -0.770]$ & $0.602$ \\ \hline
Model 2&&&&\\
$1.00$ &$0.68$ &$-0.94$ & $[-1.078, -0.804]$ & $0.274$ \\
$1.00$ &$0.90$ &$-0.94$ & $[-1.168, -0.720]$ & $0.447$ \\
$1.00$ &$0.95$ &$-0.94$ & $[-1.214, -0.680]$ & $0.534$ \\
$0.60$ &$0.68$ &$-1.09$ & $[-1.182, -0.883]$ & $0.298$ \\
$0.60$ &$0.90$ &$-1.09$ & $[-1.285, -0.791]$ & $0.494$ \\
$0.60$ &$0.95$ &$-1.09$ & $[-1.332, -0.749]$ & $0.583$ \\
$0.20$ &$0.68$ &$-1.09$ & $[-1.247, -0.936]$ & $0.311$ \\
$0.20$ &$0.90$ &$-1.09$ & $[-1.359, -0.840]$ & $0.518$ \\
$0.20$ &$0.95$ &$-1.09$ & $[-1.413, -0.798]$ & $0.614$ \\
$0.00$ &$0.68$ &$-1.11$ & $[-1.266, -0.948]$ & $0.317$ \\
$0.00$ &$0.90$ &$-1.11$ & $[-1.375, -0.850]$ & $0.525$ \\
$0.00$ &$0.95$ &$-1.11$ & $[-1.428, -0.808]$ & $0.621$ \\
\hline
\end{tabular}
\caption{Mode and credible intervals of $w$ for only the SNe~Ia, Model 1, and Model 2 for varying
fractions of spectroscopic coverage. \label{seed2:tab}}
\end{table}

%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2/{contour.2000.0.0}.pdf}
%   \includegraphics[scale=0.17]{/Users/akim/project/abc/results/seed2_pop2/{contour.2000.0.0}.pdf} 
%\caption{Probability distribution functions for the difference in parameters
%with 100\% and 0\% spectroscopic folowup for Model 1 (left) and Model 2 (right).  \label{contour2:fig}}
%\end{figure}

Though the decrease in spectroscopic coverage produces low bias, it does lead to
increased parameter uncertainties.
The credible intervals for $w$ are shown in Table~\ref{seed2:tab}:
for the case of 100\% follow-up the range of the 68\% $w$ interval is $\Delta w = 0.281$. 
At the opposite extreme of no spectroscopic classification the $w$ interval inflates to
$\Delta w = 0.309$, a constraining result nevertheless considering that none of the
objects are directly typed and redshifted.

As a stress test, we generate a simulated data based on a different model
referred to as Model 2, which
has a second non-Ia population with average luminosity 0.5 mag fainter
than the Ia's and 0.25 mag dispersion.
The relative rates of the original non-Ia population relative to the second is quadratic:
\begin{equation}
1 -\frac{0.8}{1.54}z^2.
\end{equation}
A realization of transients from this model, using the same random seeds
as Model 1, is shown in Figure~\ref{hd:fig}.

When there is 100\% spectroscopic coverage the   $\Omega_M$--$w$  posterior accommodates the input 
parameters, as shown  in
Figure~\ref{contour1:fig}.
A salient feature that contrasts with Model 1 is 
seen in the modes for $w$ listed in Table~\ref{seed2:tab}: biases
larger than the statistical uncertainties appear with decreasing
spectroscopic coverage.  Note that the Hubble diagrams
in these analyses are almost identical: there are redshift misassignments
for 2\% of the transients that lack spectroscopy.

%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.35]{/Users/akim/project/abc/results/seed2_pop2/{contour.2000.1.0}.pdf} \newline
%\caption{Probability distribution function for $\Omega_M$--$w$,
% 100\% spectroscopy, biased.
%    Solid lines represent the input cosmology.  Dashed lines in the
%   histograms are the 0.16 and 0.84 quantiles. 
%   \label{contour3:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.35]{/Users/akim/project/abc/results/seed2_pop2/{contour.2000.0.0}.pdf} \newline
%\caption{Probability distribution function for $\Omega_M$--$w$,
% no spectroscopy, biased.
%   \label{contour4:fig}}
%\end{figure}

Some initial conclusions.
\begin{itemize}
\item Knowing the underlying model that produces data, you should
get results statistically consistent with the truth.  A tautology?
\item Typing is important when the non-Ia model is not accurate.
\end{itemize}

%
%\subsubsection{SN~Ia Only}
%A classical analysis would only include confirmed SNe~Ia with a definitive
%redshift. The realization of simulated data has 415 Type~Ia supernovae.
%Results of the analysis of a pure SN~Ia sample with no
%host misclassification are presented:  Figure~\ref{contour.ia_only:fig}
%shows the posteriors for  $\Omega_M$--$w$ and Table~\ref{ia_only:tab} the
%credible equal-tailed intervals for $w$.  The true $w=-1$ is within
%the 0.58 credible interval. (Note that $w$ is the parameter
%of most interest for the experiment.)
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.55]{/Users/akim/project/abc/results/{contour..ia_only.500}.pdf} \newline
%\caption{Probability distribution function for $\Omega_M$--$w$ for only SNe~Ia
%   all with spectra.
%   Solid lines represent the input cosmology.  Dashed lines in the
%   histograms are the 0.16 and 0.86 quantiles. 
%   \label{contour.ia_only:fig}}
%\end{figure}
%
%\begin{table}
%\centering
%\begin{tabular}{|c|cc|}
%\hline
%\%-ile & $w$ range & $\Delta w$\\ \hline
%$0.68$ & $[-1.254, -0.975]$ & $0.279$ \\
%$0.90$ & $[-1.347, -0.889]$ & $0.458$ \\
%$0.95$ & $[-1.396, -0.846]$ & $0.550$ \\
%\hline
%\end{tabular}
%\caption{Credible intervals of $w$ using the SNe~Ia only with spectroscopic
%observations.\label{ia_only:tab}}
%\end{table}
%
%\subsubsection{Full Sample: Varying Spectroscopic Completeness}
%We now turn to the full sample of SNe~Ia and non-Ia's.  Recall that our simple
%model treats the non-Ia's in the sample
%as standard candles with a large (1 mag) intrinsic dispersion.
%
%Results for the $\Omega_M$--$w$ posterior for the
%case of 40\%, 70\%, and 100\% spectroscopic completeness
%are shown in Figures~\ref{contour.200:fig}, \ref{contour.350:fig}, and
%\ref{contour.500:fig} respectively.
%The credible intervals
%of $w$ for are given in Table~\ref{compare:tab}, with
%the input $w=-1$ within the 0.65, 0.49, and 0.47  intervals
%respectively.
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.200}.pdf} 
%   \caption{Probability distribution function for $\Omega_M$--$w$ for:
%    SNe~Ia and non-Ia with 40\%  spectroscopic completeness.
%   \label{contour.200:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.350}.pdf} 
%   \caption{Probability distribution function for $\Omega_M$--$w$ for
%    SNe~Ia and non-Ia with 70\%  spectroscopic completeness.
%   \label{contour.350:fig}}
%\end{figure}
%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[scale=0.40]{/Users/akim/project/abc/results/{contour.500}.pdf} 
%   %\includegraphics[scale=0.55]{/Users/akim/project/abc/results/{contour.200}.pdf}    
%   \caption{Probability distribution function for $\Omega_M$--$w$ for
%    SNe~Ia and non-Ia with 100\%  spectroscopic completeness.
%   \label{contour.500:fig}}
%\end{figure}
%
%\begin{table}[htbp] 
%\centering
%\begin{tabular}{|cc|cc|}
%\hline
%\# Spectra & \%-ile & $w$ range & $\Delta$\\ \hline
%200& $0.68$ & $[-1.284, -0.991]$ & $0.293$ \\
%200& $0.90$ & $[-1.385, -0.900]$ & $0.485$ \\
%200& $0.95$ & $[-1.435, -0.860]$ & $0.575$ \\
%350& $0.68$ & $[-1.243, -0.952]$ & $0.291$ \\
%350& $0.90$ & $[-1.339, -0.858]$ & $0.481$ \\
%350& $0.95$ & $[-1.390, -0.817]$ & $0.573$ \\
%500& $0.68$ & $[-1.236, -0.948]$ & $0.289$ \\
%500& $0.90$ & $[-1.336, -0.861]$ & $0.475$ \\
%500& $0.95$ & $[-1.385, -0.815]$ & $0.570$ \\
%\hline
%\end{tabular}
%\caption{Credible intervals of $w$ given different levels of spectroscopy.  \label{compare:tab}}
%\end{table}
%
%As would be expected, the better the spectroscopic completeness, the less bias
%and smaller uncertainties that are obtained.
%
%
%\section{Subsections of the Model}
%
%\subsection{Type}
%\label{type:sec}
%Since our transient sample may not consist purely of SNe~Ia, the model
%has a term of the form $P(T, G | z,\mu)$, which should account for all types of
%objects that could potentially be mistaken for Type~Ia:
%the rates and association with host galaxies are described for
%the full population of potential interlopers.
%This term is not currently well known, particularly at the highest redshifts probed by DES.
%
%An idea is to consider two types: SN~Ia and non-SN~Ia, where the latter's
%intrinsic distributions have loose priors.  Transients that have been spectroscopically
%classified as non-Ia will provide the strongest leverage in constraining the
%contamination: this subset must be unbiased or properly weighted
%relative to the underlying population that constitutes the Hubble diagram.
%Running the model on a pure sample of SNe~Ia can have benefits.
%As discussed in \S\ref{systematics:sec}, comparing these results with
%those from the full sample provides
%a test of systematics.  Alternatively, the non-Ia model may be better constrained
%with the pure sample.
%
%
%To minimize the contribution of non-Ia contamination we strive for
%a pure sample selection so that $P(T=\text{non-Ia}| \mu) \rightarrow 0$.
%
%
%\subsection{Host Matching}
%\sloppy
%The host-galaxy properties depend on the GC and the transient
%coordinates $P({z}_H, {\theta}_H | {\mathit{Gals.}}, \text{RA/Dec})$.  With a
%spectrum of the transient and the underlying background, identification of the host galaxy
%is usually straightforward.
%Otherwise, projections or ambiguous cases can result in the misidentification of
%the host.
%Other supernova surveys (e.g.\ SNLS) 
%with spectroscopically confirmed host associations (e.g.\ with
%matched transient and galaxy redshifts) that replicate the DES population
%can be used to put a prior on this probability.
%
%%An effective way of addressing this term is to
%%analyze the supernovae assuming a correct host match, and
%%experimentally determine $P(\text{mismatch} )$ and
%%$\left. \langle {\mu}-\mu(z)\rangle \right|_{\text{mismatch}}$, $\left. \langle {z}-z \rangle \right|_{\text{mismatch}}$ and cross-terms to correct bias in the Hubble diagram. 
%
%
%% only model $T=\text{SN~Ia}$,
%%and experimentally determine $P(\text{non-Ia})$, 
%%$\left. \langle \mu-\mu_{model} \rangle \right|_{non-Ia}$, and its uncertainty for the
%%sample of non-Ia's that are discovered and misclassified as SN~Ia.
%%This measurement is achieved through spectroscopic
%%typing of a subset drawn from the populations used in the Hubble diagram.
%%The weighted bias would be added as a correction to the Hubble diagram under the assumption that all objects are SNe~Ia.
%
%
%\section{Systematic Tests}
%\label{systematics:sec}
%The Hubble diagrams of subsets of data quantify systematic uncertainty.  
%Subsets that may be expected to show evidence for systematics include:
%spectroscopically typed SNe~Ia; transients in deep (3) versus shallow (1/2) DES fields;
%transients with NIR data; splits based on host properties.
%
%A potentially more probative test is to compare the distance moduii inferred for the same
%supernovae, but reducing the amount of data considered.  A subset of DES
%transients have higher signal-to-noise, spectroscopic confirmation, and NIR coverage than
%the average transient, we can take away that extra information in calculating distance modulus
%to get $P(\mu | \mathit{data} - \mu | \mathit{data}^-, z | \mathit{data} - z | \mathit{data}^-)$.
%It is possible that differences in $\mu$s and $z$s have smaller uncertainties that the
%$\mu$ and $z$ alone due to common measurement uncertainties. 
%
%
%
%Generically ${{T}}$ can describe sample selection.  I choose to focus
%on objects typed as SN~Ia, rather than  all discovered transients, not just those typed as SN~Ia. 
%The non-Ia transient population has significant model uncertainty and the information they impart on the expansion history is weak:  I anticipate it is cleaner to limit ourselves to
%modeling discoveries we think are SN~Ia. This restriction does not obviate the need
%to consider non-Ia contamination in the sample, rather
%we only need to model non-Ia's
%that are mistaken for Ia and not all discovered transients.

\bibliography{/Users/akim/Documents/alex}

\end{document}

### 2016-05-08 tex/toy.tex

@@
 \frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
-&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}\\
-&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
+&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.%\\
+%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
 \end{align}
+
+The photometric and spectroscopic sample selections can be treated as efficiencies $\epsilon_c$, $\epsilon_T$ as
+\begin{align}
+p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) & = \int ds\, p(c_o, s, S_c, S_T | T=T_o, L, z=z_o, X) \\
+ & =p(c_o, S_c | T=T_o, L, z=z_o, X)  \int ds\,  p(s, S_T | T=T_o, L, z=z_o, X) \\
+ & = \epsilon_c(c_o) p(c_o| T=T_o, L, z=z_o, X)  \int ds\, \epsilon_T(c_o, s)  p(s | T=T_o, L, z=z_o, X) \\
+ & \approx \epsilon_c(c_o)   \epsilon_T(c_o, s_\star) p(c_o| T=T_o, L, z=z_o, X).
+\end{align}
+where the supernova signal in the spectrum $s$ is not extracted and we make the approximation $ p(s | T=T_o, L, z=z_o, X)  = \delta(s-s_\star)$
+and $s_\star(t_T; T=T_o, L, z=z_o, X)$ is the expected signal.
 
@@
 %& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right].
+& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right] \\
+& \approx  \sum_i   P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[ \int dc_o\, \epsilon_c(c_o)  \epsilon_T(c_o, s_\star)  p(c_o| T=T_o, L, z=z_o, X)  \right] .
 \end{align}

### 2016-05-08 tex/toy.tex
slight modification of toy model

@@
 \begin{align}
-\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
+\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int ds \,\mathcal{L}(z_o,c_o, s | S_c, S_T, z, X) \\
+& =  \int ds \, \int dL \, \sum_i p(z_o,c_o, s | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
 \end{align}
@@
 \begin{align}
-\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
+\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(s | T_i, L, S_c, S_T, z=z_j, X) \\
+& \times p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
+& \propto  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o, S_c, S_T | T_i, L, z=z_j, X) p(s, S_c, S_T | T_i, L, S_c, S_T, z=z_j, X) \\
+& \times p(L,, S_c, S_T |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
 &=  \frac{\sum_j p_j \int dL \, p(c_o, S_c, S_T | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)}\\

### 2016-05-11 tex/toy.tex
fix likelihood discussion.

@@
 The ambition is to construct a model to determine the likelihood of DES-like
-supernova data.  We start with a toy model and plan to evolve its complexity to
-have the fidelity necessary for the DES analysis.
+supernova data. 
 The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
@@
    \centering
-   \includegraphics[width=6.5in]{../results//nodes_pgm.pdf} 
+   \includegraphics[width=6.5in]{../results//hdpgm.pdf} 
    \caption{Probabilistic Graphical Model for the SN~Ia analysis.  
@@
 
-In the toy model there are three types of data that
+The model has three types of observables $O$ that
 may be observed for each supernova:
@@
 \end{itemize}
+
+There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
+host-galaxy properties, and transient properties.
+There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $T$.
+
+The likelihood is expressed as
+\begin{align*}
+\mathcal{L}(O| M, M_I) & = \frac{\epsilon(O, M, M_I) p(O|M, M_I)}{\int dO\, \epsilon(O, M, M_I) p(O|M, M_I)},
+\end{align*}
+where $p(O|M, M_I)$ is the pdf of all possible observables, and $\epsilon(O, M, M_I)$
+is the probability those observables will be included in the experimental analysis.
+
+The dependencies of the parameters are charted through the PGM to give for transients with a type measurement
+\begin{align*}
+p(O| M, M_I) & = P(T_o | T) p(z | M) p(c_o | L, M)
+\end{align*}
+and for transients without a type measurement
+\begin{align*}
+p(O| M, M_I) & = p(z | M) p(c_o | L, M),
+\end{align*}
+where the observable measurements are taken to be independent.
+
+The efficiency $\epsilon$ depends on the properties of the experiment and the sample selection criteria:
+the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, the typing
+is successful, and that the type is SN~Ia:
+\begin{align*}
+\epsilon(O, M, M_I) & = p(SNIa, typed, sample, discovered |O, M, L, T)\\
+& = p(SNIa|  typed, T_o)p(typed | sample, c_o)p(sample| discovered, c_o)p(discovered |c_o),
+\end{align*}
+where transient discovery, sample selection, and typing is based on observed photometry and
+identification as a SN~Ia is based on observed type.
 

### 2016-05-11 tex/toy.tex
a version for sam to look at.

--- 2016-05-08 tex/toy.tex
@@
 
-\section{Likelihood}
+\section{Likelihood and Posterior}
 \label{likelihood:sec}
@@
 
-In the toy model there are three types of data that
+The model has three types of observables $O$ that
 may be observed for each supernova:
 \begin{itemize}
-\item $c_o$: Counts at peak brightness, one or more independent measurements.
+\item $c_o$: Counts of the photometric measurements (light curves).
+\item $s_o$: Counts of the spectroscopic measurements.
 \item $z_o$: Redshift.
@@
 \end{itemize}
+Measurements of an observable are specified with a subscript ``o''.
+
+There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
+host-galaxy properties, and transient properties.
+There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $\tau$.
+
+The likelihood for observations $O_o$ is expressed as
+\begin{align*}
+\mathcal{L}(O_o, sample| M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
+\end{align*}
+where $p(O|M, M_I)$ is the pdf of all possible observables, and $\epsilon(O, M, M_I) $
+is the probability those observables will be included in the experimental analysis.
+
+The dependencies of the parameters are charted through the PGM to give for transients with a type measurement
+\begin{align*}
+p(O| M, M_I) & = P(T | \tau) p(z | M) p(c | L, M)
+\end{align*}
+and for transients without a type measurement (not relevant for the SN~Ia-only analysis but included for completeness)
+\begin{align*}
+p(O| M, M_I) & = p(z | M) p(c | L, M),
+\end{align*}
+where the observable measurements are taken to be independent.
+
+The efficiency $\epsilon(O, M, M_I) $ depends on the properties of the experiment and the sample selection criteria:
+the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, the typing
+is successful, and that the type is SN~Ia:
+\begin{align*}
+\epsilon(O, M, M_I)  & = P(SNIa, typed, spec, sample, discovered |O, M, L, \tau)\\
+& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | sample, c) \\
+& \times P(sample| discovered, c)P(discovered |c),
+\end{align*}
+where transient discovery, sample selection, and spectroscopic follow-up are based on observed photometry.
+
+The posterior is given by Bayes theorem
+\begin{align*}
+p(M, M_i | O_o) & \propto \mathcal{L}(O_o| M, M_I) p(M_I|M) p(M)\\
+ & \propto \mathcal{L}(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
+\end{align*}
+
+\section{Probabilities}
+Section~\ref{likelihood:sec} introduced a series of probabilities based on the structure of our model, which are now described
+in further detail.
+
+\subsection{$p(M)$: Priors on the Top-Level Parameters}
+\begin{itemize}
+\item
+The Flat $w$CDM cosmological parameters are drawn from top-hat distributions
+\begin{align}
+\Omega_M & \sim  {\mathcal{U}}(0,1)\\
+w_0 & \sim \mathcal{U}(-1.5, -0.1).
+\end{align}
+
+\item
+Photometric zeropoints for each band.
+\begin{equation}
+Z \sim \mathcal{N}\left(0,\frac{\ln{10}}{2.5}0.01\right)
+\end{equation}
+
+\item
+There are a number of parameters that describe the supernova populations.
+
+The luminosity of SNe~Ia as a function of time and wavelength are described by the SALT2 model and
+its relation to absolute magnitude.
+The mean luminosity and intrinsic magnitude dispersion for SNe~Ia are
+\begin{align}
+\ln{L}_{SNIa} & \sim \mathcal{N}(\ln{1}, 0.02) \\
+\sigma_{SNIa} & \sim \ln{\mathcal{N}}(\ln{0.1},0.1).
+\end{align}
+and for the moment we don't worry about priors on the distribution of $\alpha$ and $\beta$.
+
+For the cosmology analysis only the relative SN~Ia rate between different transients are important.
+We take the rates relative to the SN~Ia rate
+\begin{equation}
+r_{SNIa} = 1.
+\end{equation}
+This is sufficient for the SN~Ia sample analysis.
+
+When considering the photometric sample the model needs to consider non-Ia populations.
+The mean luminosity and intrinsic dispersion for non-Ia in a simplified toy model
+\begin{align}
+\ln{L}_{SNII} & \sim \mathcal{N}(\ln{0.5}, 0.02) \\
+\sigma_{SNII} & \sim \ln{\mathcal{N}}(\ln{0.4},0.1),
+\end{align}
+and make the non-Ia rate a stochastic parameter
+\begin{equation}
+r_{SNII} \sim \mathcal{U}(0.25, 4).
+\end{equation}
+
+\end{itemize}
+
+\subsection{$p(M)$: Priors on the Top-Level Parameters}
 
@@
 A description of the nodes follow.  Numbers are illustrative.
-
-\subsection{Cosmology}
-The Flat $w$CDM cosmological parameters are drawn from top-hat distributions
-\begin{align}
-\Omega_M & \sim  {\mathcal{U}}(0,1)\\
-w_0 & \sim \mathcal{U}(-1.5, -0.1).
-\end{align}
-
-\subsection{Calibration}
-Photometric zeropoints for each band.
-\begin{equation}
-Z \sim \mathcal{N}(0,\frac{\ln{10}}{2.5}0.01)
-\end{equation}
-
-\subsection{SNIa Rate}
-Type~Ia SN rate. In the analysis only relative rates between different transients are important.
-We take the rates relative to the SN~Ia rate
-\begin{equation}
-r_{SNIa} = 1.
-\end{equation}
-
-\subsection{SNII Rate}
-Type~II SN rate, relative to the SN~Ia rate
-\begin{equation}
-r_{SNII} \sim \mathcal{U}(0.25, 4).
-\end{equation}
-
-\subsection{SNIa Luminosity}
-The mean luminosity and intrinsic dispersion for SNe~Ia. In fact work with log-luminosity, intrinsic
-dispersion in mag units.
-\begin{align}
-\ln{L}_{SNIa} & \sim \mathcal{N}(\ln{1}, 0.02) \\
-\sigma_{SNIa} & \sim \ln{\mathcal{N}}(\ln{0.1},0.1).
-\end{align}
-
-\subsection{SNII Luminosity}
-The mean luminosity and intrinsic dispersion for SNe~II. In fact work with log-luminosity, intrinsic
-dispersion in mag units.
-\begin{align}
-\ln{L}_{SNII} & \sim \mathcal{N}(\ln{0.5}, 0.02) \\
-\sigma_{SNII} & \sim \ln{\mathcal{N}}(\ln{0.4},0.1).
-\end{align}
 

### 2016-05-11 tex/toy.tex
Pause for a moment

@@
 \end{align}
-and for the moment we don't worry about priors on the distribution of $\alpha$ and $\beta$.
+and for the moment we don't worry about priors on $\alpha$ and $\beta$ or the distributions of $x_1$ and $c$.
 
@@
 
-\subsection{$p(M)$: Priors on the Top-Level Parameters}
+\subsection{$p(M_I|M)$: Probabilities of the Intermediary Parameters.}
+\begin{itemize}
+\item We only have to worry about the type probabilities in the photometric-typing analysis.
+The probability an object is SN~Ia is
+\begin{align}
+p_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
+p_{SNII}&=1-p_{SNIa}.
+\label{prob:eqn}
+\end{align}
+
+The type of the object $\tau$ is drawn from the Bernoulli distribution 
+\begin{equation}
+\tau | p_{SNIa} \sim \mathcal{B}(1-p_{SNIa}),
+\end{equation}
+with the convention that $T=0$ corresponds to SN~Ia.
+
+For each SN~Ia the SALT2 parameters are drawn
+\begin{equation}
+x_1, c | \text{distribution hyperparameters }  \sim TBD.
+\end{equation}
+
+\item
+The peak luminosity is drawn from
+\begin{equation}
+\ln{L} | \tau, \ln{L}_{\tau},\sigma_{\tau} \sim \mathcal{N}\left( \ln{L}_{\tau},\frac{\ln{10}}{2.5}\sigma_{\tau}\right).
+\end{equation}
+
+\end{itemize}
+
+\subsection{$p(O| M, M_I)$: Probability of All Possible Observables}
+\begin{itemize}
+\item The spectroscopic typing is assumed to be perfect
+\begin{equation}
+T | \tau ~ \delta (T-\tau).
+\end{equation}
+
+\item The spectroscopic redshift is also assumed to be perfect
+\begin{equation}
+z | z_{host} ~ \delta (z-z_{host}).
+\end{equation}
+\item  The luminosity distance comes from the physics-theory function
+\begin{equation}
+d_L = d_L(z; \Omega_M, w0),
+\end{equation}
+the flux is something like
+\begin{equation}
+f = 10^{-\frac{\alpha x_1 + \beta c}{2.5}}\frac{L}{4\pi d_L^2} SALT2(t-t_0; x_1, c),
+\end{equation}
+and expected counts is
+\begin{equation}
+c_\star = 10^{Z/2.5}f.
+\end{equation}
+The distribution of counts for all transients is 
+\begin{equation}
+c | c_\star, \sigma_c \sim \mathcal{N}(c_\star, \sigma_c),
+\end{equation}
+where $\sigma_c$ is the measurement noise.
+
+\end{itemize}
+
+\subsection{$\epsilon(O, M, M_I)$: Sample Inclusion Efficiency}
 
@@
 \subsection{SNIa~Probability}
-Probability an object is SN~Ia
-\begin{align}
-p_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
-p_{SNII}&=1-p_{SNIa}.
-\label{prob:eqn}
-\end{align}
-\subsection{Type}
-The type of the object $T$ is drawn from the Bernoulli distribution 
-\begin{equation}
-T | p_{SNIa} \sim \mathcal{B}(1-p_{SNIa}),
-\end{equation}
-with the convention that $T=0$ corresponds to SN~Ia.
-
-The type probability appears explicitly in the calculation of the likelihoods.
-\begin{align}
-P(T_{SNIa} | z=z_j, X) &= p_{SNIa}\\
-P(T_{SNII} | z=z_j, X) &= p_{SNII}
-\end{align}
 \subsection{Observed Type}
@@
 T_o = T.
-\end{equation}
-
-\subsection{Luminosity}
-For our model
-\begin{equation}
-\ln{L} | T, \ln{L}_{T},\sigma_{T} \sim \mathcal{N}\left( \ln{L}_{T},\frac{\ln{10}}{2.5}\sigma_{T}\right).
 \end{equation}
@@
 \subsection{Counts}
-The luminosity distance comes from the physics-theory function
-\begin{equation}
-d_L = d_L(z; \Omega_M, w0),
-\end{equation}
-the flux is
-\begin{equation}
-f = \frac{L}{4\pi d_L^2},
-\end{equation}
-and expected counts is
-\begin{equation}
-c_\star = 10^{Z/2.5}f.
-\end{equation}
-The distribution of counts for all transients is 
-\begin{equation}
-c_o | c_\star, \sigma_c \sim \mathcal{N}(c_\star, \sigma_c),
-\end{equation}
-where $\sigma_c$ is the measurement noise.
 

### 2016-05-17 tex/toy.tex
and notation to specficy different samples.
Almost done with draft that is presentable.

@@
 \begin{itemize}
-\item $c_o$: Counts of the photometric measurements (light curves).
-\item $s_o$: Counts of the spectroscopic measurements.
-\item $z_o$: Redshift.
-\item $T_o$: Transient type.
+\item $c$: Counts of the photometric measurements (light curves).
+\item $s$: Counts of the spectroscopic measurements.
+\item $z$: Redshift.
+\item $T$: Transient type.
 \end{itemize}
-Measurements of an observable are specified with a subscript ``o''.
+The measurements made of the observables are labeled  with a subscript ``o''.
 
@@
 
+Not all transients that occur in nature will be detected or pass our sample cuts, and the parameter pdf's
+depend on the sample under consideration: the $S$ subscript specifies a pdf for the analysis sample,
+whereas the $p$'s with no subscript specifies the pdf of the underlying sample.
 The likelihood for observations $O_o$ is expressed as
 \begin{align*}
-\mathcal{L}(O_o, sample| M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
+\mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
 \end{align*}
-where $p(O|M, M_I)$ is the pdf of all possible observables, and $\epsilon(O, M, M_I) $
-is the probability those observables will be included in the experimental analysis.
+$p(O|M, M_I)$ is the pdf of the underlying sample, and $\epsilon(O, M, M_I) $
+is the probability those observables will be included in the analysis sample.
+As the observable uncertainties
+are expected to be independent between different transients we can decompose the likelihood into terms
+for each.
+\begin{align*}
+\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients} \epsilon(O_o^\alpha, M, M_I) p(O_o^\alpha|M, M_I)}{\int \prod_{\alpha \in transients}  \epsilon(O^\alpha, M, M_I) p(O^\alpha|M, M_I)dO^\alpha}.
+\end{align*}
+Note that we work with counts: for fluxes this splitting would not be possible due to correlated calibration uncertainties.
 
+We now consider the terms of the likelihood contributed by each transient and omit the $\alpha$ superscript. 
 The dependencies of the parameters are charted through the PGM to give for transients with a type measurement
 \begin{align*}
-p(O| M, M_I) & = P(T | \tau) p(z | M) p(c | L, M)
+p(O| M, M_I) & = P(T | \tau) p(z | M) p(c | L, M) p(s|L, M)
 \end{align*}
@@
 \end{align*}
-where the observable measurements are taken to be independent.
+where the observations are taken to have measurement uncertainties that are independent.
 
@@
 \epsilon(O, M, M_I)  & = P(SNIa, typed, spec, sample, discovered |O, M, L, \tau)\\
-& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | sample, c) \\
-& \times P(sample| discovered, c)P(discovered |c),
+& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | discovered, c) \\
+& \times P(sample| discovered, c)P(discovered |c).
 \end{align*}
-where transient discovery, sample selection, and spectroscopic follow-up are based on observed photometry.
+Transient discovery and sample selection are purely based on observed photometry $c$.
+The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
+of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
+on the signal $s$, but may also depend on the type and host properties.  For the pure SN~Ia sample under initial consideration
+the first term is 1.
 
-The posterior is given by Bayes theorem
+A frequentist analysis can proceed with this likelihood.
+Alternatively, the Bayesian posterior is specified by parameter priors as follows:
 \begin{align*}
-p(M, M_i | O_o) & \propto \mathcal{L}(O_o| M, M_I) p(M_I|M) p(M)\\
- & \propto \mathcal{L}(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
+p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M)\\
+ & \propto \mathcal{L}_S(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
 \end{align*}
+Once all the $p$'s and $P$'s are specified we are ready to analyze!
 
 \section{Probabilities}
-Section~\ref{likelihood:sec} introduced a series of probabilities based on the structure of our model, which are now described
-in further detail.
+Section~\ref{likelihood:sec} introduces the probabilities and probability distribution functions that describe the likelihood
+and posterior for the model shown in Figure~\ref{pgm:fig}.
+In this Section we present one possible model by specifying those probabilities; the assembly of other models should
+resemble this example.
 
@@
 \item
-The Flat $w$CDM cosmological parameters are drawn from top-hat distributions
+Flat $w$CDM cosmological parameters are drawn from top-hat distributions
 \begin{align}
@@
 \item
-Photometric zeropoints for each band.
+Photometric zeropoints for each band are given by
 \begin{equation}
-Z \sim \mathcal{N}\left(0,\frac{\ln{10}}{2.5}0.01\right)
+Z \sim \mathcal{N}\left(0,\left(\frac{\ln{10}}{2.5}\right)^2 C_{cal}\right),
 \end{equation}
+where $C_{cal}$ is the covariance of the calibration zeropoints that Chris Lidman is calculating.
 
@@
 
-The luminosity of SNe~Ia as a function of time and wavelength are described by the SALT2 model and
-its relation to absolute magnitude.
-The mean luminosity and intrinsic magnitude dispersion for SNe~Ia are
+The luminosity of SNe~Ia as a function of time and wavelength is described by the SALT2 model and
+the linear relation of its parameters with absolute magnitude.
+There are parameters for the global supernova parameters.  For SNe~Ia
 \begin{align}
-\ln{L}_{SNIa} & \sim \mathcal{N}(\ln{1}, 0.02) \\
-\sigma_{SNIa} & \sim \ln{\mathcal{N}}(\ln{0.1},0.1).
+M_{SNIa} & \sim \mathcal{U}(-20, -18) \\
+\arctan{\alpha_{SNIa}} & \sim \mathcal{U}(-0.2, 0.3) \\
+\arctan{\beta_{SNIa}} & \sim \mathcal{U}(-1.4, 1.4),
 \end{align}
-and for the moment we don't worry about priors on $\alpha$ and $\beta$ or the distributions of $x_1$ and $c$.
+where $M_{SNIa} = -2.5\log_{10}({L_{SNIa}})$
+and for ``hyperparameters'' that describe the distributions of the per-supernova parameters
+\begin{align}
+\log_{10}{R^{x_0}_{SNIa}} & \sim \mathcal{U}({-2}, {-0.5})\\
+\log_{10}{R^{x_1}_{SNIa}} & \sim \mathcal{U}({-0.5}, {0.5})\\
+\log_{10}{R^{c}_{SNIa}} & \sim \mathcal{U}({-1.5}, {-0.5})\\
+x_{1,SNIa}^\star& \sim \text{Cauchy}(0,1)\\
+c^\star_{SNIa} & \sim \text{Cauchy}(0,0.3).
+\end{align}
 
-For the cosmology analysis only the relative SN~Ia rate between different transients are important.
+The following aspects of the model are irrelevant for the cosmology analysis of the pure SN~Ia sample
+but are presented for completeness.
+
 We take the rates relative to the SN~Ia rate
@@
 \end{equation}
-This is sufficient for the SN~Ia sample analysis.
-
-When considering the photometric sample the model needs to consider non-Ia populations.
-The mean luminosity and intrinsic dispersion for non-Ia in a simplified toy model
-\begin{align}
-\ln{L}_{SNII} & \sim \mathcal{N}(\ln{0.5}, 0.02) \\
-\sigma_{SNII} & \sim \ln{\mathcal{N}}(\ln{0.4},0.1),
-\end{align}
-and make the non-Ia rate a stochastic parameter
+If we were interested in the absolute rates this term could be replaced with your favorite parameterized
+SN rate model.
+The relative non-Ia rate would be a stochastic parameter, with a prior
 \begin{equation}
@@
 \end{equation}
+In a better model this relative rate would be redshift-dependent.
 
+When considering the photometric sample the model needs to consider non-Ia populations.
+We defer discussion on what model is used for non-Ia's for the moment.
+
+\item 
+In the photometric-sample analysis a prior must be specified for the redshift.  The redshift is drawn from a flat distribution
+\begin{equation}
+z\sim \mathcal{U}(0,\infty).
+\end{equation}
 \end{itemize}
@@
 \begin{itemize}
-\item We only have to worry about the type probabilities in the photometric-typing analysis.
+\item Type probabilities are relevant in the photometric analysis.
 The probability an object is SN~Ia is
@@
 
-For each SN~Ia the SALT2 parameters are drawn
-\begin{equation}
-x_1, c | \text{distribution hyperparameters }  \sim TBD.
-\end{equation}
-
-\item
-The peak luminosity is drawn from
-\begin{equation}
-\ln{L} | \tau, \ln{L}_{\tau},\sigma_{\tau} \sim \mathcal{N}\left( \ln{L}_{\tau},\frac{\ln{10}}{2.5}\sigma_{\tau}\right).
-\end{equation}
+Now we have the parameters that specify the sub-type.
+\begin{align}
+t_0 & \sim \mathcal{U}\\
+x_{0, SNIa} & \sim \mathcal{N}(0,R^{x_0}_{SNX})\\
+x_{1,SNIa} & \sim \mathcal{N}(x_{1,SNIa}^\star,R^{x_1}_{SNIa})\\
+c_{SNIa} & \sim \text{SkewNormal}(c^\star_{SNIa},R^{c}_{SNIa} )
+\end{align}
 
@@
 \begin{equation}
-T | \tau ~ \delta (T-\tau).
+T | \tau \sim \delta (T-\tau).
 \end{equation}
 
-\item The spectroscopic redshift is also assumed to be perfect
+\item The spectroscopic redshift of a spectroscopically typed transient assumed to be perfect
 \begin{equation}
@@
 \end{equation}
+
+When the transient is not spectroscopically typed, there is ambiguity in the correct assignment of the host galaxy.
+Then the measured redshift is a sum of delta functions
+\begin{equation}
+p(z_o|z) = \sum_i p_i \delta(z-z_i),
+\end{equation}
+where galaxy $i$ at
+redshift $z_i$ has probability $p_i$ of being
+the host.
+
 \item  The luminosity distance comes from the physics-theory function
@@
 \end{equation}
-the flux is something like
+the flux for an observation at date $t$ is
 \begin{equation}
-f = 10^{-\frac{\alpha x_1 + \beta c}{2.5}}\frac{L}{4\pi d_L^2} SALT2(t-t_0; x_1, c),
+f = 10^{-0.25 \left(M - \alpha x_1 + \beta c \right)}\frac{1}{4\pi d_L^2} F_{SALT2}((t-t_0)/(1+z); x_0,  x_1, c, \lambda/(1+z)),
 \end{equation}
@@
 \subsection{$\epsilon(O, M, M_I)$: Sample Inclusion Efficiency}
+The probability that a transient becomes a part of the analysis sample is given by $\epsilon(O, M, M_I)$ and is dependent
+on the properties of the experiment and choices made by the experimenter.
 
-There are selection criteria $S_c$ and $S_T$ for the inclusion of an object in the
-sample and for a successul typing.
-In what follows we assume that the typed objects are drawn from the same underlying
-distribution as the sample $S_c=S_T$, and that the selection criterion depends on a  brightness
-threshold $S_c = c_o \ge c_T$.  (The spectroscopic sample $S_T$ 
-eventually should depend on the
-photometry and the quality of the spectroscopic data.)
+\begin{itemize}
+\item The probability the transient is discovered is given by
+$P(discovered |c)$.  To first order, the discovery probability within a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
+an image with noise $N$ and observable counts the efficiency is then $\epsilon(c/N)$.  The probability of discovery in at least one image is
+thus
+\begin{equation}
+P(discovered |c)= 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
+ \end{equation}
+\item Typically not all discoveries, but only transients that satisfy sample selection criteria are included in the cosmology
+analysis.  For example, the object must have pass minimum signal-to-noise requirements in several bands and phases.
+The probability that a discovered transient is included in the sample is $P(sample| discovered, c)$.
+We could do a simulation to determine this function.
+\item Alternatively we could replace  $P(sample| discovered, c)P(discovered |c)$ with
+\begin{equation}
+P(discovered | sample, c)P(sample |c).
+\end{equation}
+If the sample selection criteria are selected such that $P(discovered | sample, c)=1$ (validated through simulation) then
+$P(sample |c)$ becomes the sum of  configurations of products of step functions and Gaussians.  This has significant
+computational benefits.
+\item The probability of obtaining a spectroscopic observation of a transient is given as $P(spec | discovered, c)$.  Each follow-up telescope
+has its own criteria: OzDES is magnitude limited, VLT prefers transients on faint hosts, others may try to get a random sample uniform in
+redshift.
+\item The probability that the spectroscopic observation yields a type is $P(typed | spec, s, M, M_I)$.  This is a function we have to
+figure out!
+\item This term appears for the typed SN~Ia sample, where it is trivial $P(SNIa|  typed, \tau)=1$.
+\end{itemize}
 
-For this preliminary discussion the top-level parameters are denoted by $X$,
-the latent parameters are $L$ and $T$ (for luminosity and type).
+%\subsection{Case Typed}
+%For Case Typed the likelihood is
+%\begin{align}
+%\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(T_o,z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
+%\end{align}
+%We assume that the observations give a perfect measurement of type and redshift,
+%so that
+%\begin{align}
+%p(T_o|T) & =\delta(T_o-T)\\
+%p(z_o|z) & =\delta(z_o-z);
+%\end{align}
+%$T$ and $z$ are no longer probabilistic but are fixed.
+%Then
+%\begin{align}
+%\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL\, p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
+%&= \int dL\, \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
+%\frac{p(L, S_c, S_T | T=T_o, z=z_o, X)}{P(S_c, S_T| T=T_o,  z=z_o, X)}
+%\frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
+%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.%\\
+%%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
+%\end{align}
+%
+%The photometric and spectroscopic sample selections can be treated as efficiencies $\epsilon_c$, $\epsilon_T$ as
+%\begin{align}
+%p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) & = \int ds\, p(c_o, s, S_c, S_T | T=T_o, L, z=z_o, X) \\
+% & =p(c_o, S_c | T=T_o, L, z=z_o, X)  \int ds\,  p(s, S_T | T=T_o, L, z=z_o, X) \\
+% & = \epsilon_c(c_o) p(c_o| T=T_o, L, z=z_o, X)  \int ds\, \epsilon_T(c_o, s)  p(s | T=T_o, L, z=z_o, X) \\
+% & \approx \epsilon_c(c_o)   \epsilon_T(c_o, s_\star) p(c_o| T=T_o, L, z=z_o, X).
+%\end{align}
+%where the supernova signal in the spectrum $s$ is not extracted and we make the approximation $ p(s | T=T_o, L, z=z_o, X)  = \delta(s-s_\star)$
+%and $s_\star(t_T; T=T_o, L, z=z_o, X)$ is the expected signal.
+%
+%
+%The  normalization term sums/integrates over all possible observed types/counts 
+%\begin{align}
+%P(S_c, S_T| z=z_o, X) & =\sum_i \int dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
+%% & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
+%%& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
+%& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right] \\
+%& \approx  \sum_i   P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[ \int dc_o\, \epsilon_c(c_o)  \epsilon_T(c_o, s_\star)  p(c_o| T=T_o, L, z=z_o, X)  \right] .
+%\end{align}
+%
+%\subsection{Case Not Typed}
+%
+%
+%For Case Not Typed the likelihood is
+%\begin{align}
+%\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int ds \,\mathcal{L}(z_o,c_o, s | S_c, S_T, z, X) \\
+%& =  \int ds \, \int dL \, \sum_i p(z_o,c_o, s | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
+%\end{align}
+%
+%For the redshift measurement, we take a set of potential host galaxies, where galaxy $j$
+%are redshift $z_j$ has probability $p_j$ of being the host
+%\begin{equation}
+%p(z_o|S_c, S_T, z, X) = \sum_j   p_j\delta(z_j-z).
+%\end{equation}
+%Then
+%\begin{align}
+%\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(s | T_i, L, S_c, S_T, z=z_j, X) \\
+%& \times p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
+%& \propto  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o, S_c, S_T | T_i, L, z=z_j, X) p(s, S_c, S_T | T_i, L, S_c, S_T, z=z_j, X) \\
+%& \times p(L,, S_c, S_T |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
+%&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c, S_T | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)}\\
+%&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)},
+%\end{align}
+%where we use the fact that the counts do not directly depend on type.
+%
+%The normalization term is
+%\begin{align}
+%P(S_c| z=z_o, X) & = \int dc_o \, \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
+%& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL \, p(L|T_i, z=z_j, X) 
+%\left[ \int dc_o \, p(c_o, S_c| T=T_i, L, z=z_j, X) \right].
+%\end{align}
 
-It is useful to split the data into two cases depending on whether or not
-the transient has a type measurement $T_o$.
+%\subsection{Sample Selection}
+%Both the Typed and Not Typed cases have in their normalizations the term
+%\begin{equation}
+%\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X),
+%\end{equation}
+%which is the fraction of objects whose counts satisfy the sample selection criteria.
+%
+%
+%In the case of a single count measurement per transient and neglecting the typing selection criteria,
+%\begin{align}
+% p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
+% &= \begin{cases}
+%   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
+%   0 & \text{if }  c_o < c_T
+% \end{cases}
+%\end{align}
+%so that the integral is
+%\begin{equation}
+%\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
+%\end{equation}
+%Our count measurements come from a Normal distribution, so that integral is simply related to the error function.
+%
+%For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
+%noise of a single observation is expected to be dominated by its own photon noise making this
+%supposition a good approximation.} and $k$ of those must be above threshold in order to be
+%included in the sample.
+%Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
+%of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
+%Then
+%\begin{align}
+%p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
+% &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
+%   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } (c_{o,\alpha} \ge c_T) = g_\alpha\\
+%   0 & \text{if }  (c_{o,\alpha} \ge c_T)  \ne g_\alpha
+% \end{cases}
+%\end{align}
+%and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
+%of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
+%If the count measurement uncertainties are independent, which may not be exactly but is approximately the case when we have high
+%signal-to-noise references, the integral because a sum of products of normalized error functions.
 
-\subsection{Case Typed}
-For Case Typed the likelihood is
-\begin{align}
-\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(T_o,z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
-\end{align}
-We assume that the observations give a perfect measurement of type and redshift,
-so that
-\begin{align}
-p(T_o|T) & =\delta(T_o-T)\\
-p(z_o|z) & =\delta(z_o-z);
-\end{align}
-$T$ and $z$ are no longer probabilistic but are fixed.
-Then
-\begin{align}
-\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL\, p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
-&= \int dL\, \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
-\frac{p(L, S_c, S_T | T=T_o, z=z_o, X)}{P(S_c, S_T| T=T_o,  z=z_o, X)}
-\frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
-&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.%\\
-%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
-\end{align}
-
-The photometric and spectroscopic sample selections can be treated as efficiencies $\epsilon_c$, $\epsilon_T$ as
-\begin{align}
-p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) & = \int ds\, p(c_o, s, S_c, S_T | T=T_o, L, z=z_o, X) \\
- & =p(c_o, S_c | T=T_o, L, z=z_o, X)  \int ds\,  p(s, S_T | T=T_o, L, z=z_o, X) \\
- & = \epsilon_c(c_o) p(c_o| T=T_o, L, z=z_o, X)  \int ds\, \epsilon_T(c_o, s)  p(s | T=T_o, L, z=z_o, X) \\
- & \approx \epsilon_c(c_o)   \epsilon_T(c_o, s_\star) p(c_o| T=T_o, L, z=z_o, X).
-\end{align}
-where the supernova signal in the spectrum $s$ is not extracted and we make the approximation $ p(s | T=T_o, L, z=z_o, X)  = \delta(s-s_\star)$
-and $s_\star(t_T; T=T_o, L, z=z_o, X)$ is the expected signal.
-
-The  normalization term sums/integrates over all possible observed types/counts 
-\begin{align}
-P(S_c, S_T| z=z_o, X) & =\sum_i \int dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
-% & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
-%& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right] \\
-& \approx  \sum_i   P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[ \int dc_o\, \epsilon_c(c_o)  \epsilon_T(c_o, s_\star)  p(c_o| T=T_o, L, z=z_o, X)  \right] .
-\end{align}
-
-\subsection{Case Not Typed}
-
-For Case Not Typed the likelihood is
-\begin{align}
-\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int ds \,\mathcal{L}(z_o,c_o, s | S_c, S_T, z, X) \\
-& =  \int ds \, \int dL \, \sum_i p(z_o,c_o, s | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
-\end{align}
-
-For the redshift measurement, we take a set of potential host galaxies, where galaxy $j$
-are redshift $z_j$ has probability $p_j$ of being the host
-\begin{equation}
-p(z_o|S_c, S_T, z, X) = \sum_j   p_j\delta(z_j-z).
-\end{equation}
-Then
-\begin{align}
-\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(s | T_i, L, S_c, S_T, z=z_j, X) \\
-& \times p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
-& \propto  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o, S_c, S_T | T_i, L, z=z_j, X) p(s, S_c, S_T | T_i, L, S_c, S_T, z=z_j, X) \\
-& \times p(L,, S_c, S_T |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
-&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c, S_T | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)}\\
-&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)},
-\end{align}
-where we use the fact that the counts do not directly depend on type.
-
-The normalization term is
-\begin{align}
-P(S_c| z=z_o, X) & = \int dc_o \, \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL \, p(L|T_i, z=z_j, X) 
-\left[ \int dc_o \, p(c_o, S_c| T=T_i, L, z=z_j, X) \right].
-\end{align}
-
-\subsection{Sample Selection}
-Both the Typed and Not Typed cases have in their normalizations the term
-\begin{equation}
-\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X),
-\end{equation}
-which is the fraction of objects whose counts satisfy the sample selection criteria.
-
-In the case of a single count measurement per transient and neglecting the typing selection criteria,
-\begin{align}
- p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
- &= \begin{cases}
-   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
-   0 & \text{if }  c_o < c_T
- \end{cases}
-\end{align}
-so that the integral is
-\begin{equation}
-\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
-\end{equation}
-Our count measurements come from a Normal distribution, so that integral is simply related to the error function.
-
-For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
-noise of a single observation is expected to be dominated by its own photon noise making this
-supposition a good approximation.} and $k$ of those must be above threshold in order to be
-included in the sample.
-Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
-of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
-Then
-\begin{align}
-p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
- &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
-   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } (c_{o,\alpha} \ge c_T) = g_\alpha\\
-   0 & \text{if }  (c_{o,\alpha} \ge c_T)  \ne g_\alpha
- \end{cases}
-\end{align}
-and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
-of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
-If the count measurement uncertainties are independent, which may not be exactly but is approximately the case when we have high
-signal-to-noise references, the integral because a sum of products of normalized error functions.
-
-\section{Model}
-\label{model:sec}
-
-We now go into more detail into the model and the form of the pdf's that enter
-into the likelihoods described in \S\ref{likelihood:sec}.
-
-A description of the nodes follow.  Numbers are illustrative.
-
-\subsection{SNIa~Probability}
-\subsection{Observed Type}
-The observed type is assumed to be perfect.
-\begin{equation}
-T_o = T.
-\end{equation}
-
-The corresponding pdf appears in the likelihood.
-
-\subsection{Redshift}
-The redshift is drawn from a flat distribution
-\begin{equation}
-z\sim \mathcal{U}(0,\infty).
-\end{equation}
-%
-%\subsection{Neighbor Galaxy Redshifts}
-%Each neighbor redshift is drawn from a flat distribution.  The redshift for neighbor
-%$i$ is  
-%\begin{equation}
-%z_{N,i} \sim \mathcal{U}(0,\infty).
-%\end{equation}
-
-\subsection{Observed Redshift}
-In our model the measured redshift is a sum of delta functions
-\begin{equation}
-p(z_o|z) = \sum_i p_i \delta(z-z_i),
-\end{equation}
-where galaxy $i$ at
-redshift $z_i$ has probability $p_i$ of being
-the host.
-
-\subsection{Counts}
-
-\subsection{Implementation Notes}
+\section{Implementation Notes}
 Anticipating the use of available statistical packages, the likelihoods have been constructed
@@
 \begin{equation}
-\int dL p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
+\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
 \end{equation}
@@
  \begin{equation}
-\sigma_X \int dn p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
+\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
 \end{equation}
@@
 
-\section{Adding Complexity}
-\subsection{SALT2 Model for Transient Luminosity}
-The SALT2 model for supernova luminosity as a function of SN-frame time $t$  and wavelength $\lambda$ is
-\begin{equation}
-L(t-t_0, \lambda; M-x_0-\alpha x_1 + \beta c, x_1, c),
-\label{SALTL:eqn}
-\end{equation}
-where $\alpha$ and $\beta$ are global parameters and $t_0$, $x_0$, $x_1$ and $c$ are per-supernova parameters.
-
-Incorporating the SALT2 model leads to changes to the PGM in Figure~\ref{pgm:fig}, specifically in the following Nodes.
-UNITY notation and priors are taken.
-\subsubsection{SNX Luminosity}
-There are parameters for the global supernova parameters.  For SNe~Ia
-\begin{align}
-M_{SNIa} & \sim \mathcal{U}(-20, -18) \\
-\arctan{\alpha_{SNIa}} & \sim \mathcal{U}(-0.2, 0.3) \\
-\arctan{\beta_{SNIa}} & \sim \mathcal{U}(-1.4, 1.4)
-\end{align}
-and for ``hyperparameters'' that describe the distributions of the per-supernova parameters
-\begin{align}
-\log_{10}{R^{x_0}_{SNIa}} & \sim \mathcal{U}({-2}, {-0.5})\\
-\log_{10}{R^{x_1}_{SNIa}} & \sim \mathcal{U}({-0.5}, {0.5})\\
-\log_{10}{R^{c}_{SNIa}} & \sim \mathcal{U}({-1.5}, {-0.5})\\
-x_{1,SNIa}^\star& \sim \text{Cauchy}(0,1)\\
-c^\star_{SNIa} & \sim \text{Cauchy}(0,0.3).
-\end{align}
-
-\subsubsection{Type}
-The type of the object is still obtained as before.  Now we have the parameters that specify the sub-type.
-\begin{align}
-t_0 & \sim \mathcal{U}\\
-x_{0, SNIa} & \sim \mathcal{N}(0,R^{x_0}_{SNX})\\
-x_{1,SNIa} & \sim \mathcal{N}(x_{1,SNIa}^\star,R^{x_1}_{SNIa})\\
-c_{SNIa} & \sim \text{SkewNormal}(c^\star_{SNIa},R^{c}_{SNIa} )
-\end{align}
-
-\subsubsection{Luminosity}
-The luminosity is now fixed by Eqn.\ \ref{SALTL:eqn}.  It may be made probabilistic if an error snake is considered.
-
-\subsubsection{Likelihood}
-The likelihood for typed objects is
-\begin{multline}
-\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X)
- \propto \int dt_0 \, dx_{0,T_o} \, dx_{1,T_o} dc_{T_o} 
- p(t_0, x_{0,T_o}, x_{1,T_o}, c_{T_o} |T_o, z=z_o, X) \\ P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L_{T_o}(t_0, x_{0,T_o}, x_{1,T_o}, c_{T_o}), z=z_o, X)
-\end{multline}
-and for non-typed
-\begin{multline}
-\mathcal{L}(z_o,c_o | S_c, S_T, z, X)
- \propto \sum_j p_j \sum_i P(T_i|z=z_j, X)   \int dt_0 \, dx_{0,T_i} \, dx_{1,T_i} dc_{T_i} 
- p(t_0, x_{0,T_i}, x_{1,T_i}, c_{T_i} |T_i, z=z_j, X) \\  p(c_o, S_c, S_T | T=T_o, L_{T_i}(t_0, x_{0,T_i}, x_{1,T_i}, c_{T_i}), z=z_j, X),
- \end{multline}
-where both likelihoods must be normalized.
-
 \end{document}

### 2016-05-18 dessn/examples/discrete_observed/__init__.py
Adding sampler changes

@@ -6,7 +6,7 @@ the balls, we fumble. On a fumble, we have a 30% chance of writing down the
 wrong colour.
 
 We represent this by giving colour as a discrete parameter, where 80% of the time
-there is only one colour option, but 10% of the time we give both colours as options,
+there is only one colour option, but 20% of the time we give both colours as options,
 quantifying the probability that we switched the colours or not.
 
 """

### 2016-05-18 dessn/examples/discrete_observed/discrete_observed.py

@@ -1,11 +1,11 @@
 import logging
 import os
 import sys
-
 import numpy as np
 from dessn.framework.edge import Edge
 from dessn.framework.model import Model
 from dessn.framework.parameter import ParameterObserved, ParameterUnderlying, ParameterDiscrete
+from dessn.framework.samplers.ensemble import EnsembleSampler
 
 
 def get_data(n=400):
@@ -130,10 +130,10 @@ class DiscreteModel2(Model):
     Running this file in python first generates a PGM of the framework, and then runs ``emcee``
     and creates a corner plot:
 
-    .. figure::     ../plots/discretePGM2.png
+    .. figure::     ../dessn/examples/discrete_observed/output/pgm.png
         :align:     center
 
-    .. figure::     ../plots/discrete2.png
+    .. figure::     ../dessn/examples/discrete_observed/output/surfaces.png
         :align:     center
 
     """
@@ -159,17 +159,17 @@ if __name__ == "__main__":
     else:
         logging.basicConfig(level=logging.DEBUG)
     dir_name = os.path.dirname(__file__)
-    temp_dir = os.path.abspath(dir_name + "/../../../temp/discrete2")
-    plot_file = os.path.abspath(dir_name + "/../../../plots/discrete2.png")
+    temp_dir = os.path.abspath(dir_name + "/output/data")
+    plot_file = os.path.abspath(dir_name + "/output/surfaces.png")
 
     if not only_data:
-        pgm_file = os.path.abspath(dir_name + "/../../../plots/discretePGM2.png")
-        # model.get_pgm(pgm_file)
+        pgm_file = os.path.abspath(dir_name + "/output/pgm.png")
+        model.get_pgm(pgm_file)
 
     logging.info("Starting fit")
-    model.fit_model(num_steps=3000, num_burn=300, temp_dir=temp_dir, save_interval=20)
+    sampler = EnsembleSampler(num_steps=3000, num_burn=500, temp_dir=temp_dir, save_interval=20)
+    chain_consumer = model.fit(sampler)
 
     if not only_data:
-        chain_consumer = model.get_consumer()
         print(chain_consumer.get_summary())
-        chain_consumer.plot(filename=plot_file, display=False, truth=[0.6])
+        chain_consumer.plot(filename=plot_file, truth=[0.6])

### 2016-05-18 dessn/examples/simple/modelbased/__init__.py

@@ -1,12 +1,0 @@
-""" I have placed the class based example for implementing the simplified model
-into its own module, so that the documentation generating for the ``simple`` module
-does not get cluttered with all the small classes this module will have.
-
-The primary class to look at in code is the :class:`.ExampleModel` class.
-
-I should finally note that in order to demonstrate parameter transformations, I
-have modified the model used in the previous two examples (:class:`.ExampleIntegral`
-and :class:`.ExampleLatent`) to also include a luminosity transformation, where I
-simply halve the luminosity before converting it to flux. Physically, this could
-represent a perfect 50% mirror absorption on the primary telescope mirror.
-"""

### 2016-05-18 dessn/framework/samplers/polychord.py

@@ -1,0 +1,76 @@
+from dessn.framework.samplers.scaffold import GenericSampler
+import logging
+import os
+import numpy as np
+
+
+class PolyChord(GenericSampler):  # pragma: no cover
+    # No test coverage until I can figure out how to properly install PolyChord on travis
+    def __init__(self, temp_dir, boost=0.0, num_repeats=None):
+        """ Uses `PolyChord
+        <https://ccpforge.cse.rl.ac.uk/gf/project/polychord/>`_ to fit the supplied
+        model.
+
+        Parameters
+        ----------
+        temp_dir : str
+            Specifies the directory in which to save results
+        boost : float, optional
+            How much to boost the number of posterior samples by. Cannot exceed
+            ``num_repeats``, which is by default ``5*num_dims``
+        num_repeats : int, optional
+            The number of slice slice-sampling steps to generate a new point.
+        """
+
+        self.logger = logging.getLogger(__name__)
+        if not self.can_use():
+            self.logger.warn("Cannot load the PyPolyChord package")
+            import PyPolyChord
+        self.chain = None
+        self.pool = None
+        self.master = True
+        self.temp_dir = os.path.abspath(temp_dir)
+        self.boost = boost
+        self.num_repeats = num_repeats
+
+    def can_use(self):  # pragma: no cover
+        try:
+            import PyPolyChord
+            return True
+        except ImportError as e:
+            self.logger.warn(str(e))
+            return False
+
+    def fit(self, model):
+        """ Runs the sampler over the model and returns the flat chain of results
+
+        Returns
+        -------
+        ndarray
+            The final flattened chain of dimensions
+            ``(num_dimensions, num_walkers * (num_steps - num_burn))``
+        """
+        import PyPolyChord
+
+        num_dim = len(model._theta_names)
+        if self.num_repeats is None:
+            self.num_repeats = 2 * num_dim
+        self.logger.debug("Fitting framework with %d dimensions" % num_dim)
+
+        self.logger.info("Using PolyChord")
+        if not os.path.exists(self.temp_dir):
+            os.mkdir(self.temp_dir)
+        chain_dir = self.temp_dir + os.sep + "clusters"
+        if not os.path.exists(chain_dir):
+            os.mkdir(chain_dir)
+        print("Do it")
+        PyPolyChord.run_nested_sampling(model.get_log_posterior_polychord, num_dim, 0,
+                                        base_dir=self.temp_dir,
+                                        prior=model.get_hypercube_convert,
[…]

### 2016-05-18 dessn/framework/samplers/tempered.py

@@ -1,0 +1,113 @@
+from dessn.framework.samplers.scaffold import GenericSampler
+from dessn.utility.ptemcee import PTWrapper
+import logging
+import sys
+
+
+class ParallelTemperedSampler(GenericSampler):
+
+    def __init__(self, num_walkers=None, num_steps=5000, num_burn=2000, num_temps=5,
+                 temp_dir=None, save_interval=300):
+        """ Uses ``emcee`` and the `PTSampler
+        <http://dan.iel.fm/emcee/current/api/#emcee.PTSampler>`_ to fit the supplied
+        model.
+
+        This method sets an emcee run using the ``PTSampler`` and manual
+        chain management to allow for low to medium dimensional models. MPI running
+        is detected automatically for less hassle, and chain progress is serialised
+        to disk automatically for convenience.
+
+
+        Parameters
+        ----------
+        num_walkers : int, optional
+            The number of walkers to run. If not supplied, it defaults to eight times the
+            framework dimensionality
+        num_temps : int, optional
+            The number of temperatures to use.
+        num_steps : int, optional
+            The number of steps to run
+        num_burn : int, optional
+            The number of steps to discard for burn in
+        temp_dir : str
+            If set, specifies a directory in which to save temporary results, like the emcee chain
+        save_interval : float
+            The amount of seconds between saving the chain to file. Setting to ``None``
+            disables serialisation.
+        """
+
+        self.logger = logging.getLogger(__name__)
+        if not self.can_use():
+            self.logger.warn("Cannot load the emcee package.")
+            import emcee
+        self.chain = None
+        self.pool = None
+        self.master = True
+        self.num_temps = num_temps
+        self.num_steps = num_steps
+        self.num_burn = num_burn
+        self.temp_dir = temp_dir
+        self.save_interval = save_interval
+        self.num_walkers = num_walkers
+
+    def can_use(self):  # pragma: no cover
+        try:
+            import emcee
+            return True
+        except ImportError as e:
+            self.logger.warn(str(e))
+            return False
+
+    def fit(self, model):
+        """ Runs the sampler over the model and returns the flat chain of results
+
+        Returns
+        -------
+        ndarray
+            The final flattened chain of dimensions
+            ``(num_dimensions, num_walkers * (num_steps - num_burn))``
+        """
+        from emcee.utils import MPIPool
+        import emcee
+        try:  # pragma: no cover
+            self.pool = MPIPool()
+            if not self.pool.is_master():
+                self.logger.info("Slave waiting")
[…]

### 2016-05-18 dessn/utility/ptemcee.py

@@ -3,14 +3,14 @@ import logging
 import numpy as np
 
 
-class EmceeWrapper(object):
+class PTWrapper(object):
     def __init__(self, sampler):
         self.sampler = sampler
         self.logger = logging.getLogger(__name__)
         self.chain = None
 
-    def run_chain(self, num_steps, num_burn, num_walkers, num_dim, start=None, save_interval=300,
-                  save_dim=None, temp_dir=None):
+    def run_chain(self, num_temps, num_steps, num_burn, num_walkers, num_dim, start=None,
+                  save_interval=300, save_dim=None, temp_dir=None):
         assert num_steps > num_burn, "num_steps has to be larger than num_burn"
         if save_dim is not None:
             assert save_dim <= num_dim, "You cannot save more dimensions than you actually have"
@@ -34,14 +34,13 @@ class EmceeWrapper(object):
             raise ValueError("You need to have either a starting function or existing chains")
 
         if pos is None:
-            pos = start(num_walkers)
-
+            pos = np.array([start(num_walkers) for i in range(num_temps)])
         step = 0
-        self.chain = np.zeros((num_walkers, num_steps, save_dim))
+        self.chain = np.zeros((num_temps, num_walkers, num_steps, save_dim))
         if past_chain is not None:
-            step = min(past_chain.shape[1], num_steps)
+            step = min(past_chain.shape[2], num_steps)
             num = num_steps - step
-            self.chain[:, :step, :] = past_chain[:, :step, :]
+            self.chain[:, :, :step, :] = past_chain[:, :, :step, :]
             self.logger.debug("A further %d steps are required" % num)
         else:
             num = num_steps
@@ -57,7 +56,7 @@ class EmceeWrapper(object):
                               % (temp_dir, save_interval))
 
         for result in self.sampler.sample(pos, iterations=num, storechain=False):
-            self.chain[:, step, :] = result[0][:, :save_dim]
+            self.chain[:, :, step, :] = result[0][:, :, :save_dim]
             step += 1
             if step == 1 or temp_dir is not None and save_interval is not None:
                 t2 = time()
@@ -66,9 +65,9 @@ class EmceeWrapper(object):
                     t = t2
                     position = result[0]
                     np.save(position_file, position)
-                    np.save(chain_file, self.chain[:, :step, :])
+                    np.save(chain_file, self.chain[:, :, :step, :])
[…]

### 2016-05-18 tex/toy.tex

@@
 
-The model has three types of observables $O$ that
+The model has the following types of observables $O$ that
 may be observed for each supernova:
@@
 
-Not all transients that occur in nature will be detected or pass our sample cuts, and the parameter pdf's
-depend on the sample under consideration: the $S$ subscript specifies a pdf for the analysis sample,
-whereas the $p$'s with no subscript specifies the pdf of the underlying sample.
+The likelihood is related to the pdf of observables given the model for the population of transients that satisfy sample
+selection criteria that is different from the population of all transients: we thus label the likelihood with
+an $S$ subscript to emphasize that its population is distinct from the underlying population.
 The likelihood for observations $O_o$ is expressed as
-\begin{align*}
+\begin{align}
 \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
-\end{align*}
-$p(O|M, M_I)$ is the pdf of the underlying sample, and $\epsilon(O, M, M_I) $
+\label{likelihood:eqn}
+\end{align}
+where $p(O|M, M_I)$ is the pdf of the underlying sample, and $\epsilon(O, M, M_I) $
 is the probability those observables will be included in the analysis sample.
-As the observable uncertainties
-are expected to be independent between different transients we can decompose the likelihood into terms
-for each.
+As the observable uncertainties across different transients
+are expected to be independent the likelihood can be split into contributions
+from each,
 \begin{align*}
@@
 \end{align*}
-Note that we work with counts: for fluxes this splitting would not be possible due to correlated calibration uncertainties.
+Note that we work with counts:  correlated calibration uncertainties would prevent this splitting for fluxes.
 
-We now consider the terms of the likelihood contributed by each transient and omit the $\alpha$ superscript. 
-The dependencies of the parameters are charted through the PGM to give for transients with a type measurement
+We now consider the terms in the likelihood contributed by each transient and omit the $\alpha$ superscript. 
+The dependencies of the parameters are charted through the PGM to give for transients with a  spectroscopic type measurement
 \begin{align*}
-p(O| M, M_I) & = P(T | \tau) p(z | M) p(c | L, M) p(s|L, M)
+p(O| M, M_I) & = P(T |\tau) p(z | M) p(c | L, M) p (s|L, M),
 \end{align*}
@@
 \end{align*}
-where the observations are taken to have measurement uncertainties that are independent.
+where the photometric and spectroscopic observations have independent measurement uncertainties.
 
 The efficiency $\epsilon(O, M, M_I) $ depends on the properties of the experiment and the sample selection criteria:
-the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, the typing
-is successful, and that the type is SN~Ia:
+the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, has successful typing, and is SN~Ia:
 \begin{align*}
 \epsilon(O, M, M_I)  & = P(SNIa, typed, spec, sample, discovered |O, M, L, \tau)\\
-& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | discovered, c) \\
-& \times P(sample| discovered, c)P(discovered |c).
+& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | discovered, c) P(sample, discovered| c).
 \end{align*}
@@
 of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
-on the signal $s$, but may also depend on the type and host properties.  For the pure SN~Ia sample under initial consideration
-the first term is 1.
+on the signal $s$, but may also depend on the type and host properties.
 
@@
 
-\section{Probabilities}
+\section{Probabilities: Minding Ones P's and p's}
 Section~\ref{likelihood:sec} introduces the probabilities and probability distribution functions that describe the likelihood
@@
 \end{equation}
-where $C_{cal}$ is the covariance of the calibration zeropoints that Chris Lidman is calculating.
+where $C_{cal}$ is the covariance of the calibration zeropoints calculated by Chris Lidman.
+(The PGM anticipates a more complicated calibration approach.)
 
@@
 the linear relation of its parameters with absolute magnitude.
-There are parameters for the global supernova parameters.  For SNe~Ia
+The following priors are taken from the Rubin et al.\ UNITY paper:
 \begin{align}
@@
 \end{align}
-where $M_{SNIa} = -2.5\log_{10}({L_{SNIa}})$
 and for ``hyperparameters'' that describe the distributions of the per-supernova parameters
@@
 
-The following aspects of the model are irrelevant for the cosmology analysis of the pure SN~Ia sample
+The following are irrelevant for the cosmology analysis of the pure SN~Ia sample
 but are presented for completeness.
-
 We take the rates relative to the SN~Ia rate
@@
 In a better model this relative rate would be redshift-dependent.
-
-When considering the photometric sample the model needs to consider non-Ia populations.
-We defer discussion on what model is used for non-Ia's for the moment.
+The description of non-Ia models is deferred.
 
@@
 
-Now we have the parameters that specify the sub-type.
+There are parameters that specify the sub-type.
 \begin{align}
@@
 \begin{equation}
-f = 10^{-0.25 \left(M - \alpha x_1 + \beta c \right)}\frac{1}{4\pi d_L^2} F_{SALT2}((t-t_0)/(1+z); x_0,  x_1, c, \lambda/(1+z)),
+f = 10^{-0.25 \left(M_{SNIa} - \alpha x_1 + \beta c \right)}\frac{1}{4\pi d_L^2} F_{SALT2}((t-t_0)/(1+z); x_0,  x_1, c, \lambda/(1+z)),
 \end{equation}
@@
 \end{equation}
-The distribution of counts for all transients is 
-\begin{equation}
-c | c_\star, \sigma_c \sim \mathcal{N}(c_\star, \sigma_c),
-\end{equation}
-where $\sigma_c$ is the measurement noise.
+The distribution of counts is 
+\begin{align*}
+c | c_\star &\sim \mathcal{N}(c_\star, C_c)\\
+s | c_\star &\sim \mathcal{N}(c_\star, C_s),
+\end{align*}
+where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.
 
@@
 on the properties of the experiment and choices made by the experimenter.
+\begin{itemize}
+\item The probability the transient is discovered and included in the sample is given by
+$P(sample, discovered |c)$.  To first order, the discovery probability within a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
+an image with noise $N$ the efficiency for discovery is $\epsilon(c/N)$, and the probability of discovery in at least one image
+\begin{equation}
+\epsilon_D(c) = 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
+ \end{equation}
 
-\begin{itemize}
-\item The probability the transient is discovered is given by
-$P(discovered |c)$.  To first order, the discovery probability within a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
-an image with noise $N$ and observable counts the efficiency is then $\epsilon(c/N)$.  The probability of discovery in at least one image is
-thus
-\begin{equation}
-P(discovered |c)= 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
- \end{equation}
-\item Typically not all discoveries, but only transients that satisfy sample selection criteria are included in the cosmology
-analysis.  For example, the object must have pass minimum signal-to-noise requirements in several bands and phases.
-The probability that a discovered transient is included in the sample is $P(sample| discovered, c)$.
-We could do a simulation to determine this function.
-\item Alternatively we could replace  $P(sample| discovered, c)P(discovered |c)$ with
-\begin{equation}
-P(discovered | sample, c)P(sample |c).
-\end{equation}
-If the sample selection criteria are selected such that $P(discovered | sample, c)=1$ (validated through simulation) then
-$P(sample |c)$ becomes the sum of  configurations of products of step functions and Gaussians.  This has significant
-computational benefits.
+As an illustrative example we write out the probability of a sample selection requiring two photometric points above
+a signal-to-noise threshold $STON_T$:
+\begin{align*}
+\epsilon_S(c) = &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c)\\
+&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \mathcal{N}(c_\star-c_\alpha, C_c)\prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c),
+ \end{align*}
+ where $\mathcal{H}$ is the Heaviside step function (other sample selection cases are similar but with different combinatorics).
+Then 
+$P(sample, discovered |c) = \epsilon_D(c) \epsilon_S(c)$.
+
+We could write
+\begin{align*}
+P(sample, discovered |c) & =P(sample| discovered, c)P(discovered |c)\\
+&= \epsilon_S(c)
+\end{align*}
+if the sample selection criteria are selected such that $P(discovered | sample, c)=1$.
+This would have significant
+computational benefits in the evaluation of the denominator of Eqn.~\ref{likelihood:eqn}.
 \item The probability of obtaining a spectroscopic observation of a transient is given as $P(spec | discovered, c)$.  Each follow-up telescope
 has its own criteria: OzDES is magnitude limited, VLT prefers transients on faint hosts, others may try to get a random sample uniform in
-redshift.
+redshift.  Mat Smith and Chris D'Andrea should be able to provide this term.
 \item The probability that the spectroscopic observation yields a type is $P(typed | spec, s, M, M_I)$.  This is a function we have to
@@
 
-%\subsection{Case Typed}
-%For Case Typed the likelihood is
-%\begin{align}
-%\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL \sum_i p(T_o,z_o,c_o | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
-%\end{align}
-%We assume that the observations give a perfect measurement of type and redshift,
-%so that
-%\begin{align}
-%p(T_o|T) & =\delta(T_o-T)\\
-%p(z_o|z) & =\delta(z_o-z);
-%\end{align}
-%$T$ and $z$ are no longer probabilistic but are fixed.
-%Then
-%\begin{align}
-%\mathcal{L}(T_o,z_o,c_o | S_c, S_T, z, X) & =  \int dL\, p(c_o | T=T_o, L, S_c, S_T, z=z_o, X) p(L| T=T_o, S_c, S_T, z=z_o, X)  P(T_o|S_c, S_T, z=z_o, X) \\
-%&= \int dL\, \frac{ p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) }{P(S_c, S_T, | T=T_o, L,  z=z_o, X) }
-%\frac{p(L, S_c, S_T | T=T_o, z=z_o, X)}{P(S_c, S_T| T=T_o,  z=z_o, X)}
-%\frac{p(T_o, S_c, S_T | z=z_o, X)}{P(S_c, S_T| z=z_o, X)} \\
-%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o, S_c, S_T | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.%\\
-%%&= \frac{\int dL\, p(L|T_o, z=z_o, X) P(T_o|z=z_o, X) p(c_o | T=T_o, L, z=z_o, X)}{P(S_c, S_T| z=z_o, X)}.
-%\end{align}
-%
-%The photometric and spectroscopic sample selections can be treated as efficiencies $\epsilon_c$, $\epsilon_T$ as
-%\begin{align}
-%p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) & = \int ds\, p(c_o, s, S_c, S_T | T=T_o, L, z=z_o, X) \\
-% & =p(c_o, S_c | T=T_o, L, z=z_o, X)  \int ds\,  p(s, S_T | T=T_o, L, z=z_o, X) \\
-% & = \epsilon_c(c_o) p(c_o| T=T_o, L, z=z_o, X)  \int ds\, \epsilon_T(c_o, s)  p(s | T=T_o, L, z=z_o, X) \\
-% & \approx \epsilon_c(c_o)   \epsilon_T(c_o, s_\star) p(c_o| T=T_o, L, z=z_o, X).
-%\end{align}
-%where the supernova signal in the spectrum $s$ is not extracted and we make the approximation $ p(s | T=T_o, L, z=z_o, X)  = \delta(s-s_\star)$
-%and $s_\star(t_T; T=T_o, L, z=z_o, X)$ is the expected signal.
+%\section{Implementation Notes}
+%Anticipating the use of available statistical packages, the likelihoods have been constructed
+%from standard distribution functions.  The truncated distributions from sample selection
+%are accounted by the normalization 
+%terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
+%in the PGM, they are treated as nodes who contribute 
+%$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
+%factor.
 %
 %
-%The  normalization term sums/integrates over all possible observed types/counts 
-%\begin{align}
-%P(S_c, S_T| z=z_o, X) & =\sum_i \int dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
-%% & =\sum_i \int_{-\infty}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X) \\
-%%& = \sum_i \int_{c_T}^{\infty} dc_o \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-%& = \sum_i  P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[\int dc_o\,  p(c_o,S_c, S_T | T=T_i, L, z=z_o, X)\right] \\
-%& \approx  \sum_i   P(T_i|z=z_o, X)\int dL\, p(L|T_i, z=z_o, X)  \left[ \int dc_o\, \epsilon_c(c_o)  \epsilon_T(c_o, s_\star)  p(c_o| T=T_o, L, z=z_o, X)  \right] .
-%\end{align}
-%
-%\subsection{Case Not Typed}
-%
-%
-%For Case Not Typed the likelihood is
-%\begin{align}
-%\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \int ds \,\mathcal{L}(z_o,c_o, s | S_c, S_T, z, X) \\
-%& =  \int ds \, \int dL \, \sum_i p(z_o,c_o, s | T_i, L, S_c, S_T, z, X) p(L |  T_i,  S_c, S_T, z, X) P(T_i|S_c, S_T, z, X).
-%\end{align}
-%
-%For the redshift measurement, we take a set of potential host galaxies, where galaxy $j$
-%are redshift $z_j$ has probability $p_j$ of being the host
+%The following integral appears in the normalization
 %\begin{equation}
-%p(z_o|S_c, S_T, z, X) = \sum_j   p_j\delta(z_j-z).
+%\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
 %\end{equation}
-%Then
-%\begin{align}
-%\mathcal{L}(z_o,c_o | S_c, S_T, z, X) & =  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o | T_i, L, S_c, S_T, z=z_j, X) p(s | T_i, L, S_c, S_T, z=z_j, X) \\
-%& \times p(L |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
-%& \propto  \sum_j p_j \int ds \, \int dL \,\sum_{i} p(c_o, S_c, S_T | T_i, L, z=z_j, X) p(s, S_c, S_T | T_i, L, S_c, S_T, z=z_j, X) \\
-%& \times p(L,, S_c, S_T |  T_i,  S_c, S_T,  z=z_j, X) P(T_i|S_c, S_T, z=z_j, X) \\
-%&=  \frac{\sum_j p_j \int dL \, p(c_o, S_c, S_T | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)}\\
-%&=  \frac{\sum_j p_j \int dL \, p(c_o | L, z=z_j, X) \sum_{i}  p(L|T_i, z=z_j, X) P(T_i|z=z_j, X)   }{P(S_c, S_T| z=z_j, X)},
-%\end{align}
-%where we use the fact that the counts do not directly depend on type.
-%
-%The normalization term is
-%\begin{align}
-%P(S_c| z=z_o, X) & = \int dc_o \, \mathcal{L}(T_i,z_o,c_o | S_c, S_T, z, X)\\
-%& =  \sum_j p_j  \sum_{i} P(T_i|z=z_j, X)  \int dL \, p(L|T_i, z=z_j, X) 
-%\left[ \int dc_o \, p(c_o, S_c| T=T_i, L, z=z_j, X) \right].
-%\end{align}
-
-%\subsection{Sample Selection}
-%Both the Typed and Not Typed cases have in their normalizations the term
+%We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
 %\begin{equation}
-%\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X),
+%n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
 %\end{equation}
-%which is the fraction of objects whose counts satisfy the sample selection criteria.
-%
-%
-%In the case of a single count measurement per transient and neglecting the typing selection criteria,
-%\begin{align}
-% p(c_o, S_c, S_T | T=T_o, L, z=z_o, X) &=  p(c_o, c_o \ge  c_T | T=T_o, L, z=z_o, X) \\
-% &= \begin{cases}
-%   p(c_o | T=T_o, L, z=z_o, X) & \text{if } c_o \ge c_T \\
-%   0 & \text{if }  c_o < c_T
-% \end{cases}
-%\end{align}
-%so that the integral is
+%the integral becomes
+% \begin{equation}
+%\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
+%\end{equation}
+%The two probability distributions are normal so that this reduces to
+% \begin{equation}
+%\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
+%\end{equation}
+%where
 %\begin{equation}
-%\int dc_o \, p(c_o,S_c, S_T | T=T_i, L, z=z_o, X) = \int_{c_T}^\infty dc_o\,  p(c_o | T=T_i, L, z=z_o, X).
+%c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
 %\end{equation}
-%Our count measurements come from a Normal distribution, so that integral is simply related to the error function.
-%
-%For the more general case, suppose a transient has $N$ independent count measurements\footnote{The independence of count measurements  depends on the flux extraction method; regardless the
-%noise of a single observation is expected to be dominated by its own photon noise making this
-%supposition a good approximation.} and $k$ of those must be above threshold in order to be
-%included in the sample.
-%Consider arrays of length $N$ in which each element has a value of True or False: there are $2^N$ possible configurations for this array
-%of which $\sum_{i=k}^N {N \choose i}$ have $k$ or more True's.  Let the set $\mathcal{G}$ represent the latter possible configurations.
-%Then
-%\begin{align}
-%p(c_o, S_c| T=T_o, L, z=z_o, X) &= \sum_{g \in \mathcal{G}} \prod_{\alpha} p(c_{o, \alpha}, g | T=T_o, L, z=z_o, X)\\
-% &=  \sum_{g \in \mathcal{G}}  \prod_{\alpha} \begin{cases}
-%   p(c_{o,\alpha} | T=T_o, L, z=z_o, X) & \text{if } (c_{o,\alpha} \ge c_T) = g_\alpha\\
-%   0 & \text{if }  (c_{o,\alpha} \ge c_T)  \ne g_\alpha
-% \end{cases}
-%\end{align}
-%and the normalization integral is composed of the sum over all possible configurations of the integrated probability of the truncated distributions
-%of the measured counts.  This simplifies to the the single-count case considered earlier in the case of $N=1$, $k=1$.
-%If the count measurement uncertainties are independent, which may not be exactly but is approximately the case when we have high
-%signal-to-noise references, the integral because a sum of products of normalized error functions.
-
-\section{Implementation Notes}
-Anticipating the use of available statistical packages, the likelihoods have been constructed
-from standard distribution functions.  The truncated distributions from sample selection
-are accounted by the normalization 
-terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
-in the PGM, they are treated as nodes who contribute 
-$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
-factor.
-
-The following integral appears in the normalization
-\begin{equation}
-\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
-\end{equation}
-We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
-\begin{equation}
-n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
-\end{equation}
-the integral becomes
- \begin{equation}
-\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
-\end{equation}
-The two probability distributions are normal so that this reduces to
- \begin{equation}
-\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
-\end{equation}
-where
-\begin{equation}
-c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
-\end{equation}
 

### 2016-05-20 tex/toy.tex
redo efficiencies in more rigorous fashion

@@
 
-The likelihood is related to the pdf of observables given the model for the population of transients that satisfy sample
-selection criteria that is different from the population of all transients: we thus label the likelihood with
-an $S$ subscript to emphasize that its population is distinct from the underlying population.
-The likelihood for observations $O_o$ is expressed as
+The model describes transients and their observables.  We are particularly interested in the pdf of those observables
+given the model.  The pdf is dependent on the sample and we consider two such samples: an underlying sample
+that represents the distribution of all possible transients and observables accommodated by the model; and
+an analysis sample that represents the distribution of transients and observables that end up being analyzed.
+The analysis sample differs from the underlying sample in that the transients of the former 
+must be discovered, pass photometric selection criteria, and for the SN~Ia sample be selected for spectroscopic
+observation, have a successful type from the spectrum, and be typed as SN~Ia.
+The pdf of the underlying sample is denoted as $p(O|M,M_I)$ whereas the pdf of the analysis sample, the likelihood,
+is denoted $\mathcal{L}_S(O | M, M_I)$. 
+The two are related by
 \begin{align}
-\mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
+% \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
+\mathcal{L}(O_o | M, M_I) & = \frac{p_S(O_o|M, M_I)}{\int  p_S(O|M, M_I)dO},
 \label{likelihood:eqn}
 \end{align}
-where $p(O|M, M_I)$ is the pdf of the underlying sample, and $\epsilon(O, M, M_I) $
-is the probability those observables will be included in the analysis sample.
-As the observable uncertainties across different transients
-are expected to be independent the likelihood can be split into contributions
+where
+\begin{align}
+p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I).
+\end{align}
+
+As the sample-inclusion probabilities and the uncertainties in $O_o$ are expected to be uncorrelated between different transients
+ the likelihood can be split into contributions
 from each,
 \begin{align*}
-\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients} \epsilon(O_o^\alpha, M, M_I) p(O_o^\alpha|M, M_I)}{\int \prod_{\alpha \in transients}  \epsilon(O^\alpha, M, M_I) p(O^\alpha|M, M_I)dO^\alpha}.
+\mathcal{L}(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients} p_S(O_o^\alpha|M, M_I)}{\prod_{\alpha \in transients} \int  p_S(O^\alpha|M, M_I)dO^\alpha}.
 \end{align*}
@@
 
-The efficiency $\epsilon(O, M, M_I) $ depends on the properties of the experiment and the sample selection criteria:
-the transient is discovered, satisfies sample selection, is selected for spectroscopic typing, has successful typing, and is SN~Ia:
-\begin{align*}
-\epsilon(O, M, M_I)  & = P(SNIa, typed, spec, sample, discovered |O, M, L, \tau)\\
-& = P(SNIa|  typed, \tau)P(typed | spec, s, M, M_I)P(spec | discovered, c) P(sample, discovered| c).
-\end{align*}
-Transient discovery and sample selection are purely based on observed photometry $c$.
-The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
-of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
-on the signal $s$, but may also depend on the type and host properties.
+The probability density for an object to enter the analysis sample can be broken out
+\begin{align}
+p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I)\\
+&= P(T, T=SNIa| s, c, disc=True, anal=True, sobs=True, type=True, M, M_I)\\
+& \times  p(s, type=True | c, disc=True, anal=True, sobs=True, M, M_I)\\
+& \times  p(c, disc=True, anal=True, sobs=True| M, M_I)\\
+&= P(T, T=SNIa| M, M_I)  p(s, type=True | M, M_I) \\
+& \times  p(c, disc=True, anal=True, sobs=True| M, M_I).
+\end{align}
+
+%Transient discovery and sample selection are purely based on observed photometry $c$.
+%The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
+%of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
+%on the signal $s$, but may also depend on the type and host properties.
 
@@
 \begin{align*}
-p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M)\\
- & \propto \mathcal{L}_S(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
+p(M, M_i | O_o) & \propto \mathcal{L}(O_o| M, M_I) p(M_I|M) p(M)\\
+ & \propto \mathcal{L}(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
 \end{align*}
@@
 \begin{align}
-p_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
-p_{SNII}&=1-p_{SNIa}.
+P_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
+P_{SNII}&=1-P_{SNIa}.
 \label{prob:eqn}
@@
 T | \tau \sim \delta (T-\tau).
+\label{typepdf:eqn}
 \end{equation}
@@
 
-\subsection{$\epsilon(O, M, M_I)$: Sample Inclusion Efficiency}
-The probability that a transient becomes a part of the analysis sample is given by $\epsilon(O, M, M_I)$ and is dependent
-on the properties of the experiment and choices made by the experimenter.
+\subsection{$p_S(O|M, M_I)$: Sample Selection Probability}
+We have listed the probabilities for the underlying distribution.  We now describe the probabilities for making it into the analysis sample.
+There are three terms.
+
 \begin{itemize}
-\item The probability the transient is discovered and included in the sample is given by
-$P(sample, discovered |c)$.  To first order, the discovery probability within a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
-an image with noise $N$ the efficiency for discovery is $\epsilon(c/N)$, and the probability of discovery in at least one image
-\begin{equation}
-\epsilon_D(c) = 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
- \end{equation}
+\item $P(T, T=SNIa| M, M_I)$: the probability a type measurement $T$ yields an SN~Ia type.  It depends
+on rate parameters as passed through in Eqn.~\ref{typepdf:eqn}.
+\item $p(s, type=True | M, M_I)$: the probability of a spectral measurement $s$ that yields a successful typing.
+\item $p(c, disc=True, anal=True, sobs=True| M, M_I)$
+\end{itemize}
+%\subsection{$\epsilon(O, M, M_I)$: Sample Inclusion Efficiency}
+%The probability that a transient becomes a part of the analysis sample is given by  \begin{equation}
+%P(typed, spec, sample, discovered | s, c, M, M_I) = \epsilon_T(s, M, M_I) \epsilon_O(c) \epsilon_S(c) \epsilon_D(c)
+%\end{equation}
+%and is dependent
+%on the properties of the experiment and choices made by the experimenter.
+%The probability of typing depends on the spectroscopic data whereas the probability of spectroscopic observation,
+%sample selection, and discovery depend on photometry.
+%
+%\begin{itemize}
+%\item The probability the transient is discovered 
+%in a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
+%an image with noise $N$ the efficiency for discovery is $\epsilon(c/N)$, and the probability of discovery in at least one image
+%\begin{equation}
+%\epsilon_D(c) = 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
+% \end{equation}
+%\item
+%As an illustrative example we write out the probability of a sample selection requiring two photometric points above
+%a signal-to-noise threshold $STON_T$:
+%\begin{align*}
+%\epsilon_S(c) = &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c)\\
+%&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \mathcal{N}(c_\star-c_\alpha, C_c)\prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c),
+% \end{align*}
+% where $\mathcal{H}$ is the Heaviside step function (other sample selection cases are similar but with different combinatorics).
+%\item The probability of obtaining a spectroscopic observation of a transient is given as $\epsilon_O(c)$.  Each follow-up telescope
+%has its own criteria: OzDES is magnitude limited, VLT prefers transients on faint hosts, others may try to get a random sample uniform in
+%redshift.  Mat Smith and Chris D'Andrea should be able to provide this term.
+%\item The probability that the spectroscopic observation yields a type is $\epsilon_T( s, M, M_I)$.  This is a function we have to
+%figure out!
+%\end{itemize}
 
-As an illustrative example we write out the probability of a sample selection requiring two photometric points above
-a signal-to-noise threshold $STON_T$:
-\begin{align*}
-\epsilon_S(c) = &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c)\\
-&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \mathcal{N}(c_\star-c_\alpha, C_c)\prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c),
- \end{align*}
- where $\mathcal{H}$ is the Heaviside step function (other sample selection cases are similar but with different combinatorics).
-Then 
-$P(sample, discovered |c) = \epsilon_D(c) \epsilon_S(c)$.
-
+\subsection{$\int  p_S(O|M, M_I)dO$: Normalizing the Likelihood}
 We could write
@@
 computational benefits in the evaluation of the denominator of Eqn.~\ref{likelihood:eqn}.
-\item The probability of obtaining a spectroscopic observation of a transient is given as $P(spec | discovered, c)$.  Each follow-up telescope
-has its own criteria: OzDES is magnitude limited, VLT prefers transients on faint hosts, others may try to get a random sample uniform in
-redshift.  Mat Smith and Chris D'Andrea should be able to provide this term.
-\item The probability that the spectroscopic observation yields a type is $P(typed | spec, s, M, M_I)$.  This is a function we have to
-figure out!
-\item This term appears for the typed SN~Ia sample, where it is trivial $P(SNIa|  typed, \tau)=1$.
-\end{itemize}
 
-%\section{Implementation Notes}
-%Anticipating the use of available statistical packages, the likelihoods have been constructed
-%from standard distribution functions.  The truncated distributions from sample selection
-%are accounted by the normalization 
-%terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
-%in the PGM, they are treated as nodes who contribute 
-%$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
-%factor.
-%
-%
-%The following integral appears in the normalization
-%\begin{equation}
-%\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
-%\end{equation}
-%We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
-%\begin{equation}
-%n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
-%\end{equation}
-%the integral becomes
-% \begin{equation}
-%\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
-%\end{equation}
-%The two probability distributions are normal so that this reduces to
-% \begin{equation}
-%\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
-%\end{equation}
-%where
-%\begin{equation}
-%c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
-%\end{equation}
+\section{Implementation Notes}
+Anticipating the use of available statistical packages, the likelihoods have been constructed
+from standard distribution functions.  The truncated distributions from sample selection
+are accounted by the normalization 
+terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
+in the PGM, they are treated as nodes who contribute 
+$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
+factor.
+
+The following integral appears in the normalization
+\begin{equation}
+\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
+\end{equation}
+We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
+\begin{equation}
+n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
+\end{equation}
+the integral becomes
+ \begin{equation}
+\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
+\end{equation}
+The two probability distributions are normal so that this reduces to
+ \begin{equation}
+\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
+\end{equation}
+where
+\begin{equation}
+c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
+\end{equation}
 

### 2016-05-22 tex/toy.tex
A fixed up version that I think is rigorous

@@
 
-The ambition is to construct a model to determine the likelihood of DES-like
-supernova data. 
+The ambition is to construct a model that predicts the observations of
+transients.
 The conceptual structure of the model is shown in Figure~\ref{pgm:fig}. 
@@
 
+There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
+host-galaxy properties, and transient properties.
+There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $\tau$.
+
 The model has the following types of observables $O$ that
-may be observed for each supernova:
+may be observed for each transient:
 \begin{itemize}
@@
 
-There are top-level  sets of parameters $M$ that correspond to photometric calibration, cosmology,
-host-galaxy properties, and transient properties.
-There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $\tau$.
-
-The model describes transients and their observables.  We are particularly interested in the pdf of those observables
-given the model.  The pdf is dependent on the sample and we consider two such samples: an underlying sample
-that represents the distribution of all possible transients and observables accommodated by the model; and
-an analysis sample that represents the distribution of transients and observables that end up being analyzed.
-The analysis sample differs from the underlying sample in that the transients of the former 
+We are interested in what the model predicts for the pdf of those observables.  The pdf is dependent on the population and we consider two such populations.  The first is
+the underlying population
+of all possible transients and observables accommodated by the model: the pdf for this
+distribution is $p(O|M,M_I)$  and in this note unsubscripted $p$ and $P$'s are probabilities
+for this popiulation.
+The second is
+the analysis sample that represents the population of transients and observables that end up being analyzed.
+The analysis population differs from the underlying population in that the transients of the former 
 must be discovered, pass photometric selection criteria, and for the SN~Ia sample be selected for spectroscopic
 observation, have a successful type from the spectrum, and be typed as SN~Ia.
-The pdf of the underlying sample is denoted as $p(O|M,M_I)$ whereas the pdf of the analysis sample, the likelihood,
-is denoted $\mathcal{L}_S(O | M, M_I)$. 
-The two are related by
+The $p$ and $P$ probabilities for the analysis population are subscripted with $S$.
+For the SN~Ia sample the two are related by
 \begin{align}
 % \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
-\mathcal{L}(O_o | M, M_I) & = \frac{p_S(O_o|M, M_I)}{\int  p_S(O|M, M_I)dO},
+\mathcal{L}_S(O_o | M, M_I) & = \frac{p_S(O_o|M, M_I)}{\int  p_S(O|M, M_I)dO},
 \label{likelihood:eqn}
@@
 \begin{align}
-p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I).
+p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I)\\
+&= p(O, disc=True, anal=True, sobs=True, type=True| T=SNIa, M, M_I)\\
+& \times P(T=SNIa| M, M_I)\\
+&= p(T, z, s, type=True| T=SNIa ,M, M_I)\\
+& \times p(c, disc=True, anal=True, sobs=True| T=SNIa,  M, M_I)\\
+& \times P(T=SNIa| M, M_I) \\
+&= p(T, Œz, type=True|s,  T=SNIa, M, M_I) \\
+& \times P(disc=True, anal=True, sobs=True| c, T=SNIa, M, M_I) \\
+& \times p(s| T=SNIa, M, M_I) p(c| T=SNIa, M, M_I) P(T=SNIa| M, M_I).
 \end{align}
+We anticipate that the discovery, photometric selection, spectroscopic
+targeting, and photometric measurements are independent of the typing success, and the
+spectral, redshift and typing measurements.
 
@@
 
-We now consider the terms in the likelihood contributed by each transient and omit the $\alpha$ superscript. 
-The dependencies of the parameters are charted through the PGM to give for transients with a  spectroscopic type measurement
-\begin{align*}
-p(O| M, M_I) & = P(T |\tau) p(z | M) p(c | L, M) p (s|L, M),
-\end{align*}
-and for transients without a type measurement (not relevant for the SN~Ia-only analysis but included for completeness)
-\begin{align*}
-p(O| M, M_I) & = p(z | M) p(c | L, M),
-\end{align*}
-where the photometric and spectroscopic observations have independent measurement uncertainties.
-
-The probability density for an object to enter the analysis sample can be broken out
-\begin{align}
-p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I)\\
-&= P(T, T=SNIa| s, c, disc=True, anal=True, sobs=True, type=True, M, M_I)\\
-& \times  p(s, type=True | c, disc=True, anal=True, sobs=True, M, M_I)\\
-& \times  p(c, disc=True, anal=True, sobs=True| M, M_I)\\
-&= P(T, T=SNIa| M, M_I)  p(s, type=True | M, M_I) \\
-& \times  p(c, disc=True, anal=True, sobs=True| M, M_I).
-\end{align}
-
-%Transient discovery and sample selection are purely based on observed photometry $c$.
-%The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
-%of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
-%on the signal $s$, but may also depend on the type and host properties.
-
 A frequentist analysis can proceed with this likelihood.
@@
 
+%We now consider the terms in the likelihood contributed by each transient and omit the $\alpha$ superscript. 
+%The dependencies of the parameters are charted through the PGM to give for transients with a  spectroscopic type measurement
+%\begin{align*}
+%p(O| M, M_I) & = P(T |\tau) p(z | M) p(c | L, M) p (s|L, M),
+%\end{align*}
+%and for transients without a type measurement (not relevant for the SN~Ia-only analysis but included for completeness)
+%\begin{align*}
+%p(O| M, M_I) & = p(z | M) p(c | L, M),
+%\end{align*}
+%where the photometric and spectroscopic observations have independent measurement uncertainties.
+%
+
+%Transient discovery and sample selection are purely based on observed photometry $c$.
+%The decision for spectroscopic follow-up is also based on observed photometry,  though perhaps external redshift estimates
+%of the possible host galaxy may come into play.  The success of the typing measurement should primarily depend
+%on the signal $s$, but may also depend on the type and host properties.
+
 \section{Probabilities: Minding Ones P's and p's}
@@
 and posterior for the model shown in Figure~\ref{pgm:fig}.
-In this Section we present one possible model by specifying those probabilities; the assembly of other models should
+In this Section we present one possible model by specifying underlying the
+probabilities; the assembly of other models should
 resemble this example.
@@
 with the convention that $T=0$ corresponds to SN~Ia.
+This is a term of $p_S(O|M, M_I)$
 
@@
 \end{equation}
-The distribution of counts is 
+The distributions of counts are 
 \begin{align*}
 c | c_\star &\sim \mathcal{N}(c_\star, C_c)\\
-s | c_\star &\sim \mathcal{N}(c_\star, C_s),
+s | c_\star &\sim \mathcal{N}(c_\star, \sigma_s),
 \end{align*}
-where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.
+where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.  These are terms of $p_S(O|M, M_I)$.
 
@@
 \subsection{$p_S(O|M, M_I)$: Sample Selection Probability}
-We have listed the probabilities for the underlying distribution.  We now describe the probabilities for making it into the analysis sample.
-There are three terms.
+We have listed the probabilities for the underlying distribution.  We now describe the probabilities 
+of the two terms that involve sample selection.
 
 \begin{itemize}
-\item $P(T, T=SNIa| M, M_I)$: the probability a type measurement $T$ yields an SN~Ia type.  It depends
-on rate parameters as passed through in Eqn.~\ref{typepdf:eqn}.
-\item $p(s, type=True | M, M_I)$: the probability of a spectral measurement $s$ that yields a successful typing.
-\item $p(c, disc=True, anal=True, sobs=True| M, M_I)$
+\item Spectroscopic term: $p(T, z, type=True|s,  T=SNIa, M, M_I) $. 
+This term can be reduced as
+\begin{align*}
+p(T, z, type=True|  s,  T=SNIa, M, M_I) &= p(T|  s,  T=SNIa, M, M_I) \\
+&\times P(type=True| s,  T=SNIa, M, M_I)\\
+&\times p(z|s,  T=SNIa, M, M_I),
+ \end{align*}
+with the (reasonable) approximation that the three measurements are
+independent.
+
+Condition §$type=True$: The probability of successful spectroscopic typing needs to be determined.
+A simple possibility is that  it only depends on the signal-to-noise of the spectroscopic observation
+$\epsilon_T(ston)$.  Then given a  spectroscopic noise of $\sigma_s$
+\begin{equation}
+P(type=True| s,  T=SNIa, M, M_I) = \epsilon_T(s/\sigma_s).
+\end{equation}
+
+Assembling all the terms gives
+\begin{align*}
+p(T, z, type=True|  s,  T=SNIa, M, M_I) &= \delta_{T, SNIa} \delta(z-z_{host}) \epsilon_T(s/\sigma_s) .
+ \end{align*}
+
+\item Photometric term: $P(disc=True, anal=True, sobs=True| c, T=SNIa, M, M_I)$.
+\begin{align*}
+P(disc=True, anal  =True, sobs=True| c, T=SNIa, M, M_I) & =P(anal=True| c, T=SNIa, M, M_I) \\
+& \times P(sobs=True|  c, T=SNIa, M, M_I) \\
+& \times P(disc=True| c, T=SNIa, M, M_I),
+ \end{align*}
+taking that the random processes that determine discovery, sample selection, and
+spectroscopic observation are independent.
+
+Condition $disc=True$:
+The probability for discovery of a transient in a single image depends on the
+signal-to-noise $\epsilon_D(ston)$.  The probability of discovery given all
+the images is
+\begin{equation}
+P(disc=True | c, T=SNIa, M, M_I)
+=  1-\prod_{\beta \in images} \left(1-\epsilon_D(c_\beta/N_\beta)\right).
+\end{equation}
+
+Condition $anal=True$:
+As an illustrative example we write out the probability of a analysis-sample selection requiring two photometric points above
+a signal-to-noise threshold $STON_T$: other sample selection cases are similar but with different combinatorics
+\begin{align*}
+P(anal=True  | c, T=SNIa, M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\\
+&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
+ \end{align*}
+ where $\mathcal{H}$ is the Heaviside step function.
+
+Condition $sobs=True$:
+The probability a transient has a spectroscopic observation.
+Mat Smith and Chris D'Andrea should be able to provide this term.
+Until we get this, lets make the simplifying ansatz that
+\begin{align*}
+P(sobs=True | c, T=SNIa, M, M_I)= & const.
+ \end{align*}
+
 \end{itemize}
-%\subsection{$\epsilon(O, M, M_I)$: Sample Inclusion Efficiency}
-%The probability that a transient becomes a part of the analysis sample is given by  \begin{equation}
-%P(typed, spec, sample, discovered | s, c, M, M_I) = \epsilon_T(s, M, M_I) \epsilon_O(c) \epsilon_S(c) \epsilon_D(c)
-%\end{equation}
-%and is dependent
-%on the properties of the experiment and choices made by the experimenter.
-%The probability of typing depends on the spectroscopic data whereas the probability of spectroscopic observation,
-%sample selection, and discovery depend on photometry.
-%
-%\begin{itemize}
-%\item The probability the transient is discovered 
-%in a single image depends on the transient signal-to-noise $\epsilon(STON)$.  For
-%an image with noise $N$ the efficiency for discovery is $\epsilon(c/N)$, and the probability of discovery in at least one image
-%\begin{equation}
-%\epsilon_D(c) = 1-\prod_{\beta \in images} \left(1-\epsilon(c_\beta/N_\beta)\right).
-% \end{equation}
-%\item
-%As an illustrative example we write out the probability of a sample selection requiring two photometric points above
-%a signal-to-noise threshold $STON_T$:
-%\begin{align*}
-%\epsilon_S(c) = &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c)\\
-%&  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \mathcal{N}(c_\star-c_\alpha, C_c)\prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\mathcal{N}(c_\star-c_\beta, C_c),
-% \end{align*}
-% where $\mathcal{H}$ is the Heaviside step function (other sample selection cases are similar but with different combinatorics).
-%\item The probability of obtaining a spectroscopic observation of a transient is given as $\epsilon_O(c)$.  Each follow-up telescope
-%has its own criteria: OzDES is magnitude limited, VLT prefers transients on faint hosts, others may try to get a random sample uniform in
-%redshift.  Mat Smith and Chris D'Andrea should be able to provide this term.
-%\item The probability that the spectroscopic observation yields a type is $\epsilon_T( s, M, M_I)$.  This is a function we have to
-%figure out!
-%\end{itemize}
-
-\subsection{$\int  p_S(O|M, M_I)dO$: Normalizing the Likelihood}
-We could write
-\begin{align*}
-P(sample, discovered |c) & =P(sample| discovered, c)P(discovered |c)\\
-&= \epsilon_S(c)
-\end{align*}
-if the sample selection criteria are selected such that $P(discovered | sample, c)=1$.
-This would have significant
-computational benefits in the evaluation of the denominator of Eqn.~\ref{likelihood:eqn}.
 

### 2016-05-22 tex/toy.tex
recast the story in a cleaner way

@@
 \end{itemize}
-The measurements made of the observables are labeled  with a subscript ``o''.
+The actual measurements of the observables are labeled  with a subscript ``o''.
 
-We are interested in what the model predicts for the pdf of those observables.  The pdf is dependent on the population and we consider two such populations.  The first is
-the underlying population
-of all possible transients and observables accommodated by the model: the pdf for this
-distribution is $p(O|M,M_I)$  and in this note unsubscripted $p$ and $P$'s are probabilities
-for this popiulation.
-The second is
-the analysis sample that represents the population of transients and observables that end up being analyzed.
+We are interested in what the model predicts for the pdf of those observables
+for the population of transients that enter the analysis.  
+The probabilities for this ``analysis'' population are notated with the
+subscript $S$, for example the likelihood is expressed as $\mathcal{L}_S(O| M, M_I)$.
+
+The likelihood can be expressed in terms of probabilities of the observables
+for all possible transients and observables accommodated by the model.
+The probabilities for this ``underlying'' population are notated with no subscript,
+so that the pdf of observables for this population is $p(O|M, M_I$).
 The analysis population differs from the underlying population in that the transients of the former 
-must be discovered, pass photometric selection criteria, and for the SN~Ia sample be selected for spectroscopic
+must, for the typed SN~Ia analysis, be discovered, pass photometric selection criteria, be selected for spectroscopic
 observation, have a successful type from the spectrum, and be typed as SN~Ia.
-The $p$ and $P$ probabilities for the analysis population are subscripted with $S$.
-For the SN~Ia sample the two are related by
+
+The likelihood is proportional to the pdf of the underlying sample that
+enter the analysis sample, as expressed by
 \begin{align}
 % \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
-\mathcal{L}_S(O_o | M, M_I) & = \frac{p_S(O_o|M, M_I)}{\int  p_S(O|M, M_I)dO},
+\mathcal{L}_S(O | M, M_I) & \propto p(O, \text{in analysis}|M, M_I)\\
+&  \propto \prod_{\alpha \in transients} p(O^\alpha, \text{in analysis}|M, M_I),
+\end{align}
+where the second line holds because the observations and stochastic decisions
+for sample selection are expected to be independent from transient to transient.
+
+For a single supernova labeled by $\alpha$,   the probability of the analysis population is related
+to the underlying population by
+\begin{align*}
+p_S(O^\alpha|M, M_I) &\equiv p(O^\alpha, \text{in analysis}|M, M_I)\\
+& = P(disc=True, anal=True, sobs=True, type=True, T^\alpha=SNIa|O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
+& = P(disc=True, anal=True, sobs=True, type=True|T^\alpha=SNIa, O^\alpha, M, M_I)\\
+& \times P(T^\alpha=SNIa| O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
+&= P(disc=True|T^\alpha=SNIa,c^\alpha, M, M_I)
+P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I) \\
+&  \times P(sobs=True|T^\alpha=SNIa,c^\alpha, M, M_I)
+ P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I)\\
+&  \times
+P(T^\alpha=SNIa| M, M_I)p(c^\alpha|M, M_I)p(s^\alpha|M, M_I)p(z^\alpha|M, M_I).
+\end{align*}
+where we anticipate that the stochastic processes that enter each sample-selection step
+are independent of each other, and that the measurement uncertainties are independent.
+We find that $p_S$ is the pdf of the underlying
+distribution $p(O^\alpha|M, M_I)$ multiplied by the efficiency of a realized target to
+enter the analysis sample.
+
+The likelihood can be expressed as
+\begin{align}
+\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO}.
 \label{likelihood:eqn}
 \end{align}
-where
-\begin{align}
-p_S(O|M, M_I) &= p(O, disc=True, anal=True, sobs=True, type=True, T=SNIa |M, M_I)\\
-&= p(O, disc=True, anal=True, sobs=True, type=True| T=SNIa, M, M_I)\\
-& \times P(T=SNIa| M, M_I)\\
-&= p(T, z, s, type=True| T=SNIa ,M, M_I)\\
-& \times p(c, disc=True, anal=True, sobs=True| T=SNIa,  M, M_I)\\
-& \times P(T=SNIa| M, M_I) \\
-&= p(T, Œz, type=True|s,  T=SNIa, M, M_I) \\
-& \times P(disc=True, anal=True, sobs=True| c, T=SNIa, M, M_I) \\
-& \times p(s| T=SNIa, M, M_I) p(c| T=SNIa, M, M_I) P(T=SNIa| M, M_I).
-\end{align}
-We anticipate that the discovery, photometric selection, spectroscopic
-targeting, and photometric measurements are independent of the typing success, and the
-spectral, redshift and typing measurements.
-
-As the sample-inclusion probabilities and the uncertainties in $O_o$ are expected to be uncorrelated between different transients
- the likelihood can be split into contributions
-from each,
+A frequentist analysis can proceed with this likelihood.
+Alternatively, the Bayesian posterior is specified by parameter priors as 
 \begin{align*}
-\mathcal{L}(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients} p_S(O_o^\alpha|M, M_I)}{\prod_{\alpha \in transients} \int  p_S(O^\alpha|M, M_I)dO^\alpha}.
-\end{align*}
-Note that we work with counts:  correlated calibration uncertainties would prevent this splitting for fluxes.
-
-A frequentist analysis can proceed with this likelihood.
-Alternatively, the Bayesian posterior is specified by parameter priors as follows:
-\begin{align*}
-p(M, M_i | O_o) & \propto \mathcal{L}(O_o| M, M_I) p(M_I|M) p(M)\\
- & \propto \mathcal{L}(O_o| M, M_I) p(L| \tau, M) p(\tau|M) p(M).
+p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M)\\
+ & \propto \mathcal{L}_S(O_o| M, L, \tau) p(L| \tau, M) p(\tau|M) p(M).
 \end{align*}
@@
 and posterior for the model shown in Figure~\ref{pgm:fig}.
-In this Section we present one possible model by specifying underlying the
-probabilities; the assembly of other models should
+In this Section we present one possible model; the assembly of other models should
 resemble this example.
+We now drop the $\alpha$ label with the implicit understanding that the distributions
+now described are per transient.
 
@@
 with the convention that $T=0$ corresponds to SN~Ia.
-This is a term of $p_S(O|M, M_I)$
 
@@
 \end{align*}
-where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.  These are terms of $p_S(O|M, M_I)$.
+where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.
 
@@
 
-\subsection{$p_S(O|M, M_I)$: Sample Selection Probability}
-We have listed the probabilities for the underlying distribution.  We now describe the probabilities 
-of the two terms that involve sample selection.
+\subsection{Sample Selection Probability}
+We have listed the terms that contribute to the probability that a target enters
+the analysis sample.
 
 \begin{itemize}
-\item Spectroscopic term: $p(T, z, type=True|s,  T=SNIa, M, M_I) $. 
-This term can be reduced as
-\begin{align*}
-p(T, z, type=True|  s,  T=SNIa, M, M_I) &= p(T|  s,  T=SNIa, M, M_I) \\
-&\times P(type=True| s,  T=SNIa, M, M_I)\\
-&\times p(z|s,  T=SNIa, M, M_I),
- \end{align*}
-with the (reasonable) approximation that the three measurements are
-independent.
-
-Condition §$type=True$: The probability of successful spectroscopic typing needs to be determined.
+\item $P(T=SNIa| M, M_I) = \delta_{\tau, SNIa}$: The transient is
+measured to be SN~Ia if it is SN~Ia.
+\item $ P(type=True|T=SNIa, s, M, M_I)$: 
+The probability of successful spectroscopic typing needs to be determined.
 A simple possibility is that  it only depends on the signal-to-noise of the spectroscopic observation
@@
 \begin{equation}
-P(type=True| s,  T=SNIa, M, M_I) = \epsilon_T(s/\sigma_s).
+P(type=True|  T=SNIa, s, M, M_I) = \epsilon_T(s/\sigma_s).
 \end{equation}
 
-Assembling all the terms gives
-\begin{align*}
-p(T, z, type=True|  s,  T=SNIa, M, M_I) &= \delta_{T, SNIa} \delta(z-z_{host}) \epsilon_T(s/\sigma_s) .
- \end{align*}
-
-\item Photometric term: $P(disc=True, anal=True, sobs=True| c, T=SNIa, M, M_I)$.
-\begin{align*}
-P(disc=True, anal  =True, sobs=True| c, T=SNIa, M, M_I) & =P(anal=True| c, T=SNIa, M, M_I) \\
-& \times P(sobs=True|  c, T=SNIa, M, M_I) \\
-& \times P(disc=True| c, T=SNIa, M, M_I),
- \end{align*}
-taking that the random processes that determine discovery, sample selection, and
-spectroscopic observation are independent.
-
-Condition $disc=True$:
+\item $P(disc=True|T=SNIa,c, M, M_I)$:
 The probability for discovery of a transient in a single image depends on the
@@
 \begin{equation}
-P(disc=True | c, T=SNIa, M, M_I)
+P(disc=True | T=SNIa,c,  M, M_I)
 =  1-\prod_{\beta \in images} \left(1-\epsilon_D(c_\beta/N_\beta)\right).
@@
 
-Condition $anal=True$:
+\item $P(anal=True|T=SNIa,c, M, M_I)$:
 As an illustrative example we write out the probability of a analysis-sample selection requiring two photometric points above
@@
 \begin{align*}
-P(anal=True  | c, T=SNIa, M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\\
+P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\\
 &  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
@@
 
-Condition $sobs=True$:
+\item $P(sobs=True|T=SNIa,c, M, M_I)$:
 The probability a transient has a spectroscopic observation.
@@
 \begin{align*}
-P(sobs=True | c, T=SNIa, M, M_I)= & const.
+P(sobs=True | T=SNIa, c, M, M_I)= & const.
  \end{align*}

### 2016-05-22 tex/toy.tex
pad up with more words to clarify.

@@
 host-galaxy properties, and transient properties.
-There are two sets of intermediary parameters $M_I$ that correspond to luminosity $L$ and type $\tau$.
+There are two sets of intermediary parameters  $M_I=\{L, \tau\}$. that correspond to luminosity and type respectively.
 
@@
 for the population of transients that enter the analysis.  
-The probabilities for this ``analysis'' population are notated with the
-subscript $S$, for example the likelihood is expressed as $\mathcal{L}_S(O| M, M_I)$.
-
 The likelihood can be expressed in terms of probabilities of the observables
 for all possible transients and observables accommodated by the model.
-The probabilities for this ``underlying'' population are notated with no subscript,
-so that the pdf of observables for this population is $p(O|M, M_I$).
 The analysis population differs from the underlying population in that the transients of the former 
@@
 observation, have a successful type from the spectrum, and be typed as SN~Ia.
-
-The likelihood is proportional to the pdf of the underlying sample that
-enter the analysis sample, as expressed by
-\begin{align}
+The likelihood of the observations $O_o$ is
+\begin{align*}
 % \mathcal{L}_S(O_o | M, M_I) & = \frac{\epsilon(O_o, M, M_I) p(O_o|M, M_I)}{\int  \epsilon(O, M, M_I) p(O|M, M_I)dO},
-\mathcal{L}_S(O | M, M_I) & \propto p(O, \text{in analysis}|M, M_I)\\
-&  \propto \prod_{\alpha \in transients} p(O^\alpha, \text{in analysis}|M, M_I),
-\end{align}
-where the second line holds because the observations and stochastic decisions
+\mathcal{L}_S(O_o | M, M_I) & = p(O_o |\text{in analysis}, M, M_I)\\
+& = \frac{ p(O_o, \text{in analysis}|M, M_I)}{p(\text{in analysis}|M, M_I)}\\
+&  = \frac{\prod_{\alpha \in transients} p(O_o^\alpha, \text{in analysis}|M, M_I)}{p(\text{in analysis}|M, M_I)},
+\end{align*}
+where the last line holds because the observations and stochastic decisions
 for sample selection are expected to be independent from transient to transient.
@@
 to the underlying population by
-\begin{align*}
-p_S(O^\alpha|M, M_I) &\equiv p(O^\alpha, \text{in analysis}|M, M_I)\\
-& = P(disc=True, anal=True, sobs=True, type=True, T^\alpha=SNIa|O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
-& = P(disc=True, anal=True, sobs=True, type=True|T^\alpha=SNIa, O^\alpha, M, M_I)\\
-& \times P(T^\alpha=SNIa| O^\alpha, M, M_I)p(O^\alpha|M, M_I)\\
+\begin{align}
+p_S(O^\alpha|M, M_I) &\equiv p(O^\alpha, \text{in analysis}|M, M_I) \nonumber \\
+& = P(disc=True, anal=True, sobs=True, type=True, T^\alpha=SNIa|O^\alpha, M, M_I)p(O^\alpha|M, M_I) \nonumber \\
+& = P(disc=True, anal=True, sobs=True, type=True|T^\alpha=SNIa, O^\alpha, M, M_I) \nonumber \\
+& \times P(T^\alpha=SNIa| O^\alpha, M, M_I)p(O^\alpha|M, M_I) \nonumber \\
 &= P(disc=True|T^\alpha=SNIa,c^\alpha, M, M_I)
-P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I) \\
+P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I)  \nonumber \\
 &  \times P(sobs=True|T^\alpha=SNIa,c^\alpha, M, M_I)
- P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I)\\
+ P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I) \nonumber \\
 &  \times
 P(T^\alpha=SNIa| M, M_I)p(c^\alpha|M, M_I)p(s^\alpha|M, M_I)p(z^\alpha|M, M_I).
-\end{align*}
-where we anticipate that the stochastic processes that enter each sample-selection step
-are independent of each other, and that the measurement uncertainties are independent.
-We find that $p_S$ is the pdf of the underlying
+\label{p_S:eqn}
+\end{align}
+where we anticipate that the stochastic processes that operate on each sample-selection decision
+are independent of each other, and that the measurement uncertainties of type, counts,
+spectrum, and redshift are independent.
+$p_S$ is equal to the pdf of the underlying
 distribution $p(O^\alpha|M, M_I)$ multiplied by the efficiency of a realized target to
@@
 
-The likelihood can be expressed as
+The likelihood is then  re-expressed as
 \begin{align}
-\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO}.
+\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO},
 \label{likelihood:eqn}
 \end{align}
+where  the denominator $p(\text{in analysis}|M, M_I)$ is set
+by the condition that the integral over the likelihood over all possible
+observations must equal to one.
 A frequentist analysis can proceed with this likelihood.
 Alternatively, the Bayesian posterior is specified by parameter priors as 
-\begin{align*}
-p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M)\\
+\begin{align}
+p(M, M_i | O_o) & \propto \mathcal{L}_S(O_o| M, M_I) p(M_I|M) p(M) \nonumber \\
  & \propto \mathcal{L}_S(O_o| M, L, \tau) p(L| \tau, M) p(\tau|M) p(M).
-\end{align*}
-Once all the $p$'s and $P$'s are specified we are ready to analyze!
+\label{posterior:eqn}
+\end{align}
+Once all the $p$'s and $P$'s introduced in
+Eqns.~\ref{p_S:eqn}, \ref{likelihood:eqn}, and \ref{posterior:eqn}
+are specified we are ready to analyze!
 
@@
 \begin{itemize}
-\item Type probabilities are relevant in the photometric analysis.
+\item $p(\tau| M, M_I) $: Type probabilities are relevant in the photometric analysis.
 The probability an object is SN~Ia is
 \begin{align}
-P_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}} \nonumber \\
+P_{SNIa} &= \frac{r_{SNIa}}{r_{SNIa}+r_{SNII}}  \\
 P_{SNII}&=1-P_{SNIa}.
@@
 
-There are parameters that specify the sub-type.
+\item $P(L| \tau, M, M_I) $: 
+For the SALT2 model several parameters
+must be specified to define the subtype,
+including the parameter $x_{0, SNIa}$ that corresponds to $L$.
 \begin{align}
@@
 \end{align}
-
 \end{itemize}
@@
 \begin{itemize}
-\item The spectroscopic typing is assumed to be perfect
+\item $P(T=SNIa| M, M_I) $: The spectroscopic typing is assumed to be perfect
 \begin{equation}
@@
 
-\item The spectroscopic redshift of a spectroscopically typed transient assumed to be perfect
+\item $P(z| M, M_I) $: The spectroscopic redshift of a spectroscopically typed transient assumed to be perfect
 \begin{equation}
@@
 
-\item  The luminosity distance comes from the physics-theory function
+\item  $P(c| M, M_I)$, $P(s| M, M_I)$:
+The luminosity distance comes from the physics-theory function
 \begin{equation}
@@
 The distributions of counts are 
-\begin{align*}
+\begin{align}
 c | c_\star &\sim \mathcal{N}(c_\star, C_c)\\
 s | c_\star &\sim \mathcal{N}(c_\star, \sigma_s),
-\end{align*}
+\end{align}
 where $C_c$ and $C_s$ are the measurement covariances for the photometric and spectroscopic observations respectively.
@@
 \begin{itemize}
-\item $P(T=SNIa| M, M_I) = \delta_{\tau, SNIa}$: The transient is
-measured to be SN~Ia if it is SN~Ia.
 \item $ P(type=True|T=SNIa, s, M, M_I)$: 
@@
 \begin{equation}
-P(disc=True | T=SNIa,c,  M, M_I)
+P(disc=True | T=SNIa,c,  M, M_Ii)
 =  1-\prod_{\beta \in images} \left(1-\epsilon_D(c_\beta/N_\beta)\right).
@@
 a signal-to-noise threshold $STON_T$: other sample selection cases are similar but with different combinatorics
-\begin{align*}
-P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T))\\
+\begin{align}
+P(anal=True  | T=SNIa, c,  M, M_I)= &1 - \prod_{\beta \in images} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)) \nonumber \\
 &  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
- \end{align*}
+ \end{align}
  where $\mathcal{H}$ is the Heaviside step function.
@@
 Until we get this, lets make the simplifying ansatz that
-\begin{align*}
+\begin{align}
 P(sobs=True | T=SNIa, c, M, M_I)= & const.
- \end{align*}
+ \end{align}
 
 \end{itemize}
+
+\subsection{$p(\text{in analysis}|M, M_I)$: The Normalization}
 

### 2016-05-23 tex/toy.tex
add discussion on evaluating the integral for nice cases of sample selection.

@@
 \begin{align}
-\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\int  \prod_{\alpha \in transients} p_S(O^\alpha |M, M_I)dO},
+\mathcal{L}_S(O_o | M, M_I) & = \frac{\prod_{\alpha \in transients}p_S(O_o^\alpha|M, M_I)}{\prod_{\alpha \in transients} \int  p_S(O^\alpha |M, M_I)dO^\alpha},
 \label{likelihood:eqn}
@@
 &  - \sum_{\alpha \in images}\mathcal{H}(c_\alpha/N_\alpha; STON_T) \prod_{\beta \in \{images-\alpha\}} (1-\mathcal{H}(c_\beta/N_\beta; STON_T)),
+\label{anal:eqn}
  \end{align}
@@
 \begin{align}
-P(sobs=True | T=SNIa, c, M, M_I)= & const.
+P(sobs=True | T=SNIa, c, M, M_I)= & 1.
+\label{sobs:eqn}
  \end{align}
@@
 \subsection{$p(\text{in analysis}|M, M_I)$: The Normalization}
-
-\section{Implementation Notes}
-Anticipating the use of available statistical packages, the likelihoods have been constructed
-from standard distribution functions.  The truncated distributions from sample selection
-are accounted by the normalization 
-terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
-in the PGM, they are treated as nodes who contribute 
-$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
-factor.
-
-The following integral appears in the normalization
+The contribution to the normalization from transient $\alpha$ can be rewritten as
+\begin{align}
+\int p_S(O^\alpha|M, M_I)dO^\alpha &= I_c I_s P(T^\alpha=SNIa| M, M_I),
+\end{align}
+where the independence of the photometric and spectroscopic observations allow us to separate the contributions of each
+\begin{multline}
+I_c = \int  P(disc=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) \\ P(sobs=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha)
+p(c^\alpha|M, M_I,z_{host}=z^\alpha) dc\\
+\end{multline}
+and
 \begin{equation}
-\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
-\end{equation}
-We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
-\begin{equation}
-n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
-\end{equation}
-the integral becomes
- \begin{equation}
-\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
-\end{equation}
-The two probability distributions are normal so that this reduces to
- \begin{equation}
-\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
-\end{equation}
-where
-\begin{equation}
-c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
+I_s = \int P(type=True|T^\alpha=SNIa, s^\alpha, M, M_I,z_{host}=z^\alpha)
+ p(s^\alpha|M, M_I,z_{host}=z^\alpha)ds.
 \end{equation}
 
+Note that in the pure SN~Ia sample analysis, $P(T^\alpha=SNIa| M, M_I)$ divides out in the calculation of the likelihood,  so that the cosmological results
+are independent of rates.  This would not be the case for the photometric-sample analysis.
+
+The integrals $I_c$ and $I_s$ can be simple to evaluate in certain circumstances.  Suppose the photometric selection requirements
+are set so that $P(disc=True | T=SNIa,c,  M, M_I) \sim 1$ and the photometric uncertainties are independent (approximatively true when
+each point's noise is dominated by its image noise).  Then for the case of the sample selection (spectroscopic observation)
+criteria in Eqn.~\ref{anal:eqn} (\ref{sobs:eqn}), 
+\begin{align}
+I_c & = \int  P(anal=True|T^\alpha=SNIa,c^\alpha, M, M_I,z_{host}=z^\alpha) p(c^\alpha|M, M_I,z_{host}=z^\alpha) dc\\
+& = 1 - \prod_{\beta \in images} \text{CDF}\left(N_\beta STON_T; \mathcal{N}(c_\star,\sigma_\beta)\right)   \nonumber \\
+& - \sum_{\alpha \in images}  \text{CCDF}\left(N_\alpha STON_T;\mathcal{N}(c_\star,\sigma_\alpha)\right)   \prod_{\beta \in \{images-\alpha\}} \text{CDF}\left(N_\beta STON_T;\mathcal{N}(c_\star,\sigma_\beta)\right),
+\end{align}
+where CDF and CCDF are the cumulative and the complimentary cumulative distribution functions respectively of the normal distribution
+that are readily accessible in programming language libraries either on their own or as error fuctions.
+%
+%\section{Implementation Notes}
+%Anticipating the use of available statistical packages, the likelihoods have been constructed
+%from standard distribution functions.  The truncated distributions from sample selection
+%are accounted by the normalization 
+%terms $P(S_c, S_T| z=z_o, X)$, $P(S_c| z=z_o, X)$.  Although they do not appear
+%in the PGM, they are treated as nodes who contribute 
+%$-\ln{P(S_c, S_T| z=z_o, X)}$ to the log-likelihood, where the negative sign reflects that this is a normalization
+%factor.
+%
+%
+%The following integral appears in the normalization
+%\begin{equation}
+%\int dL\, p(L|T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o  p(c_o | T=T_i, L, z=z_o, X)\right].
+%\end{equation}
+%We actually work with $\ln{L_X}$ and $\sigma_X$.  Setting
+%\begin{equation}
+%n=\frac{\ln{L}-\ln{L_X}}{\sigma_X}
+%\end{equation}
+%the integral becomes
+% \begin{equation}
+%\sigma_X \int dn\, p(n\sigma_X + \ln{L_X} |T_i, z=z_o, X)  \left[\int_{c_T}^{\infty} dc_o\,  p(c_o | T=T_i, L_Xe^{n\sigma_X}, z=z_o, X)\right].
+%\end{equation}
+%The two probability distributions are normal so that this reduces to
+% \begin{equation}
+%\frac{1}{2\sqrt{2\pi}} \int dn\, e^{-0.5n^2} \left(1-erf\left(\frac{(c_T - c_\star)}{\sqrt{2}\sigma_c}\right)\right),
+%\end{equation}
+%where
+%\begin{equation}
+%c_\star = \frac{e^{\ln{L_X}+n\sigma_X}}{4\pi d_L^2}10^{Z/2.5}.
+%\end{equation}
+%
 \end{document}

### 2016-08-02 dessn/investigations/gaussianity/paper/gaussianity.tex
Adding doco outline

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Supernova Gaussianity]{Investigating the Gaussianity of Supernova SALT2 Summary Statistics}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. R. Hinton et al.]{
Samuel R. Hinton,$^{1}$\thanks{E-mail: \href{samuelreay@gmail.com}}
Alex G. Kim,$^{3}$
Tamara M. Davis,$^{1,2}$
Edward Macaulay$^{1,2}$
\\
% List of institutions
$^{1}$School of Mathematics and Physics, The University of Queensland, Brisbane, QLD 4072, Australia\\
$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
$^{3}$Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2016}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Write an abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
%\begin{keywords}
%keyword1 -- keyword2 -- keyword3
%\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Few sentences on cosmology.

Use of supernova in cosmology.

SALT2.

Previous studies using summary statistics and their data sets.

Assumed gaussianity.

DES supernova (segue to precision)

Precision measurements, systematics more important than ever.

Thus motivation

\section{Next Section}
Think of section title

Supernova models, summary statistics, why they are used (instead of light curve data), and the fitting methodologies that have been used in the past to get summary statistics.

\section{Next section}

to investigate, go to lowest level and simulate light curves

to start with, realise only canonical supernova from abs mag with some scatter, using WMAP9 cosmology. Do this for shallow and deep fields, and fit the light curves using different methods.

 compare the different methods

section detailing the bias as a function of redshift (and ston if possible).

%\begin{figure}
%	% To include a figure from a file named example.*
%	% Allowable file formats are eps or ps if compiling using latex
%	% or pdf, png, jpg if compiling using pdflatex
%	\includegraphics[width=\columnwidth]{example}
%    \caption{This is an example figure. Captions appear below each figure.
%	Give enough detail for the reader to understand what they're looking at,
%	but leave detailed discussion to the main body of the text.}
%    \label{fig:example_figure}
%\end{figure}
%
%% Example table
%\begin{table}
%	\centering
%	\caption{This is an example table. Captions appear above each table.
%	Remember to define the quantities, symbols and units used.}
%	\label{tab:example_table}
%	\begin{tabular}{lccr} % four columns, alignment for each
%		\hline
%		A & B & C & D\\
%		\hline
%		1 & 2 & 3 & 4\\
%		2 & 4 & 6 & 8\\
%		3 & 5 & 7 & 9\\
%		\hline
%	\end{tabular}
%\end{table}

\section{Conclusions}

The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.

\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{bibliography} % if your bibtex file is called example.bib

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2016-08-02 dessn/investigations/gaussianity/__init__.py

@@ -189,6 +189,32 @@ above :math:`z=0.5`.
 Cosmology Impact
 ----------------
 
+
+To test the impact on cosmology fits this bias might have, I implement a basic cosmology
+simulation.
+
+We start with the basics, and assume that supernova are drawn from some underlying
+distribution of absolute magnitude :math:`M_i \sim \mathcal{N}(M,\sigma_{\rm int})`. We
+set our simulation cosmology to ``astropy.cosmology.WMAP9``, and (as done before), set
+our Philip's relation variables :math:`\alpha` and :math:`\beta` to  :math:`0.14` and
+:math:`3.15` respectively, (following fiducial values from [1]_), and fix the values there.
+Supernova colours are drawn such that :math:`c_i \sim \mathcal{N}(0, 0.1)` and colours are drawn
+as :math:`x_{1i} \sim \mathcal{N}(0, 1)`. The redshift distribution is calculated from the
+comoving volume of the DES deep-field area of 30 square degrees using the ``sncosmo``
+function ``zdist``.
+
+We assume perfectly realised zero points, which at the time of analysis were given
+as :math:`34.24,\ 34.85,\ 34.94,\ 35.42` for the *g, r, i* and *z* bands. Observational cadence
+is set to 5 days, and weather conditions and moon cycle are simulated in light curve generation.
+
+Given this, we can compute the weight as a function of the two cosmological parameters free:
+:math:`\Omega_m` and :math:`w`. We generate 500 supernova having set :math:`\sigma_{\rm int} = 0.1`
+
+
+
+----------
+
+
 To test the impact on cosmology fits this bias might have, I use a very simplistic
 cosmology fitter and do not correct for selection bias. Thus we expect the final
 cosmology to be biased, however both input data (one using skewed ``minuit`` output

### 2016-08-02 dessn/investigations/gaussianity/des_sky.py

@@ -14,11 +14,20 @@ from dessn.investigations.gaussianity.model import PerfectRedshift
 from dessn.framework.samplers.ensemble import EnsembleSampler
 from astropy.cosmology import WMAP9
 
-def realise_light_curve(temp_dir, zeros, seed, scatter=0.3):
+
+def realise_light_curve(temp_dir, zeros, seed, scatter=0.3, cosmology=None, mabs=-19.3):
+    simulation = cosmology is None
+    if cosmology is None:
+        cosmology = WMAP9
     np.random.seed(seed)
     t0 = 1000
     num_obs = 20
-    z = np.random.uniform(0.1, 0.9)
+    if simulation:
+        zs = list(sncosmo.zdist(0.05, 1.2, area=30, time=10000))
+        z_ind = np.random.randint(0, len(zs))
+        z = zs[z_ind]
+    else:
+        z = np.random.uniform(0.1, 0.9)
     deltat = -35
 
     newmoon = 1000 - np.random.uniform(0, 29.5)
@@ -53,18 +62,19 @@ def realise_light_curve(temp_dir, zeros, seed, scatter=0.3):
 
     model = sncosmo.Model(source='salt2-extended')
     model.set(z=z)
-    mabs = np.random.normal(-19.3, scatter)
-    model.set_source_peakabsmag(mabs, 'bessellb', 'ab')
+    mabs = np.random.normal(mabs, scatter)
+    model.set_source_peakabsmag(mabs, 'bessellb', 'ab', cosmo=cosmology)
     x0 = model.get('x0')
-    x1 = 0
-    c = 0
+    if not simulation:
+        x1 = 0
+        c = 0
+    else:
+        x1 = np.random.normal(0, 1)
+        c = np.random.normal(0, 0.1)
     p = {'z': z, 't0': t0, 'x0': x0, 'x1': x1, 'c': c}
 
     lc = sncosmo.realize_lcs(obs, model, [p])[0]
-    # fig = sncosmo.plot_lc(lc)
-    # fig.savefig(temp_dir + os.sep + "lc_%d.png" % seed, bbox_inches="tight", dpi=300)
     ston = (lc["flux"] / lc["fluxerr"]).max()
-    # print(z, t0, x0, x1, c, ston)
     return z, t0, x0, x1, c, ston, lc
 
 
@@ -84,9 +94,7 @@ def get_gaussian_fit(z, t0, x0, x1, c, lc, seed, temp_dir, interped, type="iminu
     elif type == "mcmc":
[…]

### 2016-08-02 dessn/investigations/gaussianity/simple_cosmology_fitter.py

@@ -71,6 +71,19 @@ class W(ParameterUnderlying):
         return 1
 
 
+class BiasCorrection(Edge):
+    def __init__(self, interpolator):
+        super().__init__("z", ["omega_m", "w", "M"])
+        self.interpolator = interpolator
+
+    def get_log_likelihood(self, data):
+        omega_m, w, mabs = data["omega_m"], data["w"], data["mabs"]
+        biases = self.interpolator([omega_m, w, mabs])
+        ps = -np.log(biases)
+        ps[ps == np.inf] = -np.inf
+        return np.ones(data["z"].shape) * ps
+
+
 class Likelihood(Edge):
     def __init__(self):
         self.H0 = 68
@@ -87,7 +100,7 @@ class Likelihood(Edge):
 
 
 class SimpleCosmologyFitter(Model):
-    def __init__(self, name, zs, mus, mues):
+    def __init__(self, name, zs, mus, mues, interpolator):
         super().__init__(name)
         self.add(Redshifts(zs))
         self.add(Mus(mus))
@@ -96,4 +109,5 @@ class SimpleCosmologyFitter(Model):
         self.add(W())
         self.add(M())
         self.add(Likelihood())
+        self.add(BiasCorrection(interpolator))
         self.finalise()

### 2016-08-02 dessn/investigations/gaussianity/simple_cosmology_test.py

@@ -7,68 +7,108 @@ summary statistics fitting software cause any shift in output
 cosmological parameters.
 
 """
+from scipy.interpolate import RegularGridInterpolator
 
 from dessn.investigations.gaussianity.simple_cosmology_fitter import SimpleCosmologyFitter
-from dessn.investigations.gaussianity.des_sky import get_result
+from dessn.investigations.gaussianity.des_sky import get_result, realise_light_curve
 from dessn.framework.samplers.ensemble import EnsembleSampler
 import os
 from joblib import Parallel, delayed
 import numpy as np
 import logging
-from astropy.cosmology import WMAP9
+from astropy.cosmology import WMAP9, FlatwCDM
 
 
-if __name__ == "__main__":
-    temp_dir = os.path.dirname(__file__) + "/output/example"
-    temp_dir2 = os.path.dirname(__file__) + "/output/cosmology"
+def get_zp_and_name(shallow):
+    if shallow:
+        return [32.46, 32.28, 32.55, 33.12], "shallow"
+    else:
+        return [34.24, 34.85, 34.94, 35.42], "deep"
+
+
+def get_supernova_data(n=500, ston_thresh=5, shallow=True):
+    zp, name = get_zp_and_name(shallow)
+    temp_dir = os.path.dirname(__file__) + "/output/supernova_data_%s" % name
     if not os.path.exists(temp_dir):
         os.makedirs(temp_dir)
+    ress = Parallel(n_jobs=4, max_nbytes="20M", batch_size=5)(delayed(get_result)(
+        temp_dir, zp, i, 0.1, False) for i in range(n))
+    res = np.array([r for r in ress if r is not None])
+    ston = res[:, 6]
+    res = res[ston < ston_thresh, :]
+    # [seed, z, t0, x0, x1, c, ston] + mus + stds)
+    #            zs, mu_mcmc, mu_minuit, std_mcmc, std_minuit
+    return res[:, 1], res[:, 7], res[:, 8], res[:, 9], res[:, 10]
+
+
+def get_bias(omega_m, w0, mabs, n=800, ston_thresh=5, shallow=True):
+    zp, name = get_zp_and_name(shallow)
+    cosmology = FlatwCDM(70, omega_m, w0=w0)
[…]

### 2016-08-03 dessn/investigations/gaussianity/paper/gaussianity.tex
Unfortunately having some difficulty fitting

@@
 
-SALT2.
-
-Previous studies using summary statistics and their data sets.
-
-Assumed gaussianity.
-
 DES supernova (segue to precision)
@@
 
-\section{Next Section}
-Think of section title
+\section{Motivation}
 
-Supernova models, summary statistics, why they are used (instead of light curve data), and the fitting methodologies that have been used in the past to get summary statistics.
+SALT2 and other models
 
-\section{Next section}
+why summary stats are used (instead of light curve data)
+
+fitting methodologies
+
+highlight Previous studies using summary statistics and their data sets.
+
+Assumed gaussianity.
+
+\section{Methodology}
 
@@
 
- compare the different methods
+state skewness introduced as z increases and ston decreases
+
+which creates difference when between mean and max likelihood
 
 section detailing the bias as a function of redshift (and ston if possible).
+
+\section{Cosmological Impact}
+
+Fit cosmology against simulated SN
+
+Do for des shallow and deep, using survey area to produce z dist plus extra low-z sample (0.05<z<0.2). simulate from light curves, apply selection effect of ston>5to cull bad fits. have <x> shallow.
+
+toy cosmology model.
+adopt $\alpha$ and $\beta$ values for Phillips correction from Betoule2014, assume known gaussian intrinsic scatter, have $\Omega_m$, $w$ and $M_B$ as cosmology parameters - marginalise over $M_B$.
+
+Show cosmology results.
 
@@
 \bibliographystyle{mnras}
-\bibliography{bibliography} % if your bibtex file is called example.bib
+\bibliography{bibliography}
 
@@
 
-\appendix
-
-\section{Some extra material}
-
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
+%\appendix
+%\section{Some extra material}
+%If you want to present additional material which would interrupt the flow of the main paper, it can be placed in an Appendix which appears after the list of references.
 

### 2016-08-03 dessn/investigations/gaussianity/des_sky.py

@@ -23,9 +23,12 @@ def realise_light_curve(temp_dir, zeros, seed, scatter=0.3, cosmology=None, mabs
     t0 = 1000
     num_obs = 20
     if simulation:
-        zs = list(sncosmo.zdist(0.05, 1.2, area=30, time=10000))
-        z_ind = np.random.randint(0, len(zs))
-        z = zs[z_ind]
+        if np.random.random() < 0.05:
+            z = np.random.uniform(0.05, 0.2)
+        else:
+            zs = list(sncosmo.zdist(0.05, 1.2, area=30, time=10000))
+            z_ind = np.random.randint(0, len(zs))
+            z = zs[z_ind]
     else:
         z = np.random.uniform(0.1, 0.9)
     deltat = -35
@@ -92,8 +95,8 @@ def get_gaussian_fit(z, t0, x0, x1, c, lc, seed, temp_dir, interped, type="iminu
         fig = sncosmo.plot_lc(lc, model=[fitted_model, correct_model], errors=res.errors)
         fig.savefig(temp_dir + os.sep + "lc_%d.png" % seed, bbox_inches="tight", dpi=300)
     elif type == "mcmc":
-        res, fitted_model = sncosmo.mcmc_lc(lc, model, ['t0', 'x0', 'x1', 'c'], nburn=500, nwalkers=20,
-                                            nsamples=1500, guess_amplitude=False, guess_t0=False)
+        res, fitted_model = sncosmo.mcmc_lc(lc, model, ['t0', 'x0', 'x1', 'c'], nburn=300, nwalkers=10,
+                                            nsamples=1300, guess_amplitude=False, guess_t0=False)
         mmu = get_mu(interped, res.parameters[2], res.parameters[3], res.parameters[4])
         chain = np.random.multivariate_normal(res.parameters[1:], res.covariance, size=int(1e5))
     elif type == "nestle":
@@ -186,6 +189,7 @@ def get_result(temp_dir, zps, seed, scatter, special=False, full=True):
         z, t0, x0, x1, c, ston, lc = realise_light_curve(temp_dir, zps, seed, scatter=scatter)
 
         if ston < 4:
+            print("Failed ", z)
             np.save(save_file, np.array([]))
             return None
 
@@ -206,10 +210,10 @@ def get_result(temp_dir, zps, seed, scatter, special=False, full=True):
                 else:
                     chainf3, _, _, _ = get_gaussian_fit(z, t0, x0, x1, c, lc, seed, temp_dir, interp, type="nestle")
[…]

### 2016-08-03 dessn/investigations/gaussianity/simple_cosmology_fitter.py

@@ -28,11 +28,11 @@ class M(ParameterUnderlying):
         return -19.3
 
     def get_suggestion_sigma(self, data):
-        return 1
+        return 0.5
 
     def get_log_prior(self, data):
         om = data["M"]
-        if om < -25 or om > -15:
+        if om < -20 or om > -19:
             return -np.inf
         return 1
 
@@ -49,7 +49,7 @@ class OmegaM(ParameterUnderlying):
 
     def get_log_prior(self, data):
         om = data["omega_m"]
-        if om < 0.05 or om > 0.8:
+        if om < 0.1 or om > 0.6:
             return -np.inf
         return 1
 
@@ -66,7 +66,7 @@ class W(ParameterUnderlying):
 
     def get_log_prior(self, data):
         om = data["w"]
-        if om < -4 or om > 0:
+        if om < -2 or om > -0.2:
             return -np.inf
         return 1
 
@@ -77,23 +77,23 @@ class BiasCorrection(Edge):
         self.interpolator = interpolator
 
     def get_log_likelihood(self, data):
-        omega_m, w, mabs = data["omega_m"], data["w"], data["mabs"]
+        omega_m, w, mabs = data["omega_m"], data["w"], data["M"]
         biases = self.interpolator([omega_m, w, mabs])
         ps = -np.log(biases)
         ps[ps == np.inf] = -np.inf
-        return np.ones(data["z"].shape) * ps
+        return np.ones(data["z"].shape)# * ps
 
 
 class Likelihood(Edge):
     def __init__(self):
-        self.H0 = 68
+        self.H0 = 70
         super().__init__(["z", "mu", "mue"], ["omega_m", "w", "M"])
         # super().__init__(["z", "mu", "mue"], ["omega_m", "w"])
 
     def get_log_likelihood(self, data):
         cosmology = FlatwCDM(H0=self.H0, Om0=data["omega_m"], w0=data["w"])
         distmod = cosmology.distmod(data["z"]).value
-        error = np.sqrt(data["mue"]**2 + 0.01*0.01)
+        error = np.sqrt(data["mue"]**2 + 0.05*0.05)
         diff = (distmod - data["mu"] + data["M"]) / error
         # diff = (distmod - data["mu"]) / data["mue"]
         return -0.5 * (diff * diff)

### 2016-08-03 dessn/investigations/gaussianity/simple_cosmology_test.py

@@ -26,16 +26,17 @@ def get_zp_and_name(shallow):
         return [34.24, 34.85, 34.94, 35.42], "deep"
 
 
-def get_supernova_data(n=500, ston_thresh=5, shallow=True):
+def get_supernova_data(n=800, ston_thresh=5, shallow=True):
     zp, name = get_zp_and_name(shallow)
     temp_dir = os.path.dirname(__file__) + "/output/supernova_data_%s" % name
     if not os.path.exists(temp_dir):
         os.makedirs(temp_dir)
     ress = Parallel(n_jobs=4, max_nbytes="20M", batch_size=5)(delayed(get_result)(
-        temp_dir, zp, i, 0.1, False) for i in range(n))
+        temp_dir, zp, i, 0.05, False, False) for i in range(n))
+    print(ress)
     res = np.array([r for r in ress if r is not None])
     ston = res[:, 6]
-    res = res[ston < ston_thresh, :]
+    res = res[ston > ston_thresh, :]
     # [seed, z, t0, x0, x1, c, ston] + mus + stds)
     #            zs, mu_mcmc, mu_minuit, std_mcmc, std_minuit
     return res[:, 1], res[:, 7], res[:, 8], res[:, 9], res[:, 10]
@@ -45,7 +46,7 @@ def get_bias(omega_m, w0, mabs, n=800, ston_thresh=5, shallow=True):
     zp, name = get_zp_and_name(shallow)
     cosmology = FlatwCDM(70, omega_m, w0=w0)
     ress = Parallel(n_jobs=4, max_nbytes="20M", batch_size=100)(delayed(realise_light_curve)(
-        None, zp, i, 0.1, cosmology, mabs) for i in range(n))
+        None, zp, i, 0.05, cosmology, mabs) for i in range(n))
     res = np.array([r for r in ress if r is not None])
     ston = res[:, 5]
     mask = (ston > ston_thresh).sum()
@@ -79,12 +80,18 @@ def get_bias_matrix(filename, shallow=True):
 
 
 def plot_cosmology(zs, mu_mcmc, mu_minuit, std_mcmc, std_minuit):
-    distmod = WMAP9.distmod(zs).value - 19.3
     import matplotlib.pyplot as plt
     fig, ax = plt.subplots(1, 1, figsize=(5, 4))
[…]

### 2016-08-09 dessn/investigations/gaussianity/paper/gaussianity.tex
Adding a ton of citations from betoule for potential use

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
+\newcommand{\blue}{\color{blue}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+In the years since the discovery of the acceleration of our universe \citep{RiessFilippenko1998, SchmidtSuntzeff1998, PerlmutterAldering1999}, supernova cosmology continued to grow as an important cosmological probe.
+Recent supernova studies offer far greater statistical power than ever before, systematics are better understood, and the very phenomenon of supernova are described with improved models \citep{GuyAstier2007, ConleyGuy2011, BetouleKessler2014, RubinAldering2015}.
+
+\vspace{10mm}
+
+SNLS \citep{AstierGuy2006, SullivanGuy2011}
+
+ESSENCE \citep{WoodVaseyMiknaitis2007}
+
+Pan-STARRS \citep{TonryStubbs2012, ScolnicRest2014, RestScolnic2014}
+
+SDSS-II \citep{FriemanBassett2008, KesslerBecker2009, SollermanMortsell2009, LampeitlNichol2010, CampbellDAndrea2013}
+
+Harvard-Smithsonian Center for Astrophysics survey \citep[CfA,][]{HickenChallis2009}
+
+Carnegie Supernova Project \citep[CSP,][]{ContrerasHamuy2010, FolatelliPhillips2010, StritzingerPhillips2011}
+
+Lick Observatory Supernova Search \citep[LOSS,][]{GaneshalingamLi2013}
+
+Nearby Supernova Factory \cite[SNF,][]{AlderingAdam2002}
+
+HST used in \citep{RiessStrolger2007, SuzukiRubin2012}
+
+Modern results systematic limited \citep{ConleyGuy2011, SuzukiRubin2012, ScolnicRest2014}
+
+Efforts overcome \citep{IvezicSmith2007, RegnaultConley2009, TonryStubbs2012}
+
+DES
+
+Compilations:
+
+Suzuki 2012: Union 2.1
+For citing SALT2, cite Guy2007, Guy2010, Mosher2014
+
+{\blue
 Few sentences on cosmology.
@@
 
-Thus motivation
+Thus motivation}
 

### 2016-08-11 dessn/investigations/gaussianity/paper/gaussianity.tex
Update all

--- 2016-08-03 dessn/investigations/gaussianity/paper/gaussianity.tex
@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
+\newcommand{\blue}{\color{blue}}
+\newcommand{\red}{\color{red}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Few sentences on cosmology.
+In the years since the discovery of the acceleration of our universe \citep{RiessFilippenko1998, SchmidtSuntzeff1998, PerlmutterAldering1999}, supernova cosmology continued to grow as an important cosmological probe.
+Recent supernova studies offer far greater statistical power than ever before, systematics are better understood, and the very phenomenon of supernova are described with improved models \citep{GuyAstier2007, ConleyGuy2011, BetouleKessler2014, RubinAldering2015}, and with increased statistical power and model fidelity, systematics are now the limiting constraint in supernova studies \citep{ConleyGuy2011, SuzukiRubin2012, ScolnicRest2014} {\red UPDATE citations, used from Betoule}.
+In this paper we investigate a potential source of systematics unaddressed in prior studies -- the gaussianity of supernova summary statistics.
+We perform this investigation within the context of the Dark Energy Survey, which is introduced in Section~\ref{sec:des}.
+In Section~\ref{sec:motivation} we summarise recent supernova models and analyses, and in Section~\ref{sec:methodology} and Section~\ref{sec:cosmology} we investigate the effect of the assumption of gaussianity on both the supernova distributions obtained and the follow on impact on cosmology.
 
-Use of supernova in cosmology.
+\section{The Dark Energy Survey}
+\label{sec:des}
 
-DES supernova (segue to precision)
-
-Precision measurements, systematics more important than ever.
-
-Thus motivation
+The Dark Energy Survey \citep[DES,][]{des2005, des2016} is a 5000 sq.
+deg.
+photometric redshift survey.
+The survey will image approximately 300 million galaxies in 5 broadband filters (\textit{grizY}) using the Dark Energy Camera \citep[DECam,][]{FlaugherDiehl2015}.
+Forty sq.
+deg.
+of the sky will be repeatedly images in the \textit{griz} bands for the supernova luminosity distance probe, and is expected to yield observations of approximately 1900 Type Ia supernova up to a redshift of $z=1.2$.
+{\red No idea where to look for more details on this.
+On cadence, updated estimates, cadence, difference between shallow and deep fields.}
 
 \section{Motivation}
+\label{sec:motivation}
 
-SALT2 and other models
+In order to perform a supernova analysis, an underlying Type Ia supernova model must be adopted.
+The most used model in recent years has been that of the SALT2 model \citep{GuyAstier2007, GuySullivan2010, MosherGuy2014} and in this analysis we restrict our investigations to this model.
+The SALT2 model characterises the Type Ia spectral energy distribution (SED) as a function of magnitude $x_0$, stretch $x_1$, colour $c$, phase $p$ and wavelength $\lambda$ via
+\begin{align}
+F(p,\lambda) &= x_0 \left[ M_0(p,\lambda) + x_1 M_1(p,\lambda)\right] \exp\left[c CL(\lambda)\right].
+\end{align} 
+Thus characterising an individual supernova's magnitude, colour and stretch (and associated covariance) allows recovery of the modelled SED.
+These summary statistics are used instead of the observed light curves in supernova analyses for computational reasons across many different fitting methodologies.
+In the Bayesian hierarchical analysis performed on the Unity 2.1 dataset \citep{SuzukiRubin2012} in \citet{RubinAldering2015}, each supernova is described by a latent colour and stretch, which means an analysis with $N$ supernova grows in dimensionality as $2N + x$, where $x$ represents underlying (non-latent) parameters (such as $\Omega_m$).
+If each supernova has $M$ observations, a full hierarchical model would increase in dimensionality to $(2+M)N + x$, providing vastly increased computational complexity in resolving the underlying posterior surface.
+In frequentist analyses like that of the Joint Lightcurve Analysis \citep{BetouleKessler2014}, the correlation of all light curve points via the calibration zero points instead leads to computational difficulties in that a covariance matrix of approximate size $MN$ is introduced into the likelihood calculation, which would be prohibitively expensive to invert.
+Therefore, instead of utilising light curve data directly in cosmological models, light curves are fit to the supernova data to produce summary statistics, and these are then used in cosmology analyses.
 
-why summary stats are used (instead of light curve data)
-
-fitting methodologies
-
-highlight Previous studies using summary statistics and their data sets.
-
-Assumed gaussianity.
+One key assumption when doing this is that these summary statistics are in fact gaussian in nature.
+Often this is explicitly assumed in the analysis \citep{MarchTrotta2011, RubinAldering2015}, however it is often implicitly assumed \citep{SullivanGuy2011, SuzukiRubin2012, CampbellDAndrea2013, BetouleKessler2014}.
+Investigation into asymmetries in the summary statistics are generally limited to asymmetries in the parent colour and stretch distribution \citep{ScolnicRest2014}.
+{\red More citations.} In our paper we therefore do not investigate biases or asymmetries found in underlying population distributions, but asymmetries and biases in the SALT2 parameter probability surface, and the impact of assuming the Gaussianity of said surface.
 
 \section{Methodology}
+\label{sec:methodology}
 
+We begin our investigation by generating supernova light curves for mock supernova in both the DES deep and shallow fields.
+As prior studies have investigated the effects of parent colour and stretch distributions, we generate all canonical supernova ($x_1 = c = 0$).
+The supernova are simulated in the \textit{griz} bands with an observational cadence of 5 days, starting from a time $t = -35$, giving 7 observations before the peak to remove the biases introduced by having little data before the luminosity peak {\red CITE}.
+Weather conditions and moon phase are simulated based on observed DES seeing {\red Cite excel file?}.
+Simulated light curves undergo a simple selection cut of having at least one observation in any band above a signal to noise of $5$.
+
+We investigate the potential asymmetries in SALT2 summary statistics by fitting the light curves and obtaining summary statistics using different algorithms in the software package \verb|sncosmo| \citep{Barbary2014}.
+Specifically, we obtain a full posterior surface using our own MCMC fitting methodology, and compare against summary statistics generated using a different MCMC approach (the \verb|mcmc_lc| method in \verb|sncosmo|), nestled sampling (the \verb|nest_lc| method) and minuit (via the \verb|fit_lc| method).
+The MCMC based algorithm utilises the package \verb|emcee| \citep{ForemanMackey2013}, whilst the nestled sampling is based off \verb|nestle| \citep{Skilling2004, Barbary2015}
+
+{\blue
 to investigate, go to lowest level and simulate light curves
@@
 
-section detailing the bias as a function of redshift (and ston if possible).
+section detailing the bias as a function of redshift (and ston if possible). }
 
 \section{Cosmological Impact}
-
+\label{sec:cosmology}
 Fit cosmology against simulated SN
@@
 % End of mnras_template.tex
+
+%SNLS \citep{AstierGuy2006, SullivanGuy2011}
+%
+%ESSENCE \citep{WoodVaseyMiknaitis2007}
+%
+%Pan-STARRS \citep{TonryStubbs2012, ScolnicRest2014, RestScolnic2014}
+%
+%SDSS-II \citep{FriemanBassett2008, KesslerBecker2009, SollermanMortsell2009, LampeitlNichol2010, CampbellDAndrea2013}
+%
+%Harvard-Smithsonian Center for Astrophysics survey \citep[CfA,][]{HickenChallis2009}
+%
+%Carnegie Supernova Project \citep[CSP,][]{ContrerasHamuy2010, FolatelliPhillips2010, StritzingerPhillips2011}
+%
+%Lick Observatory Supernova Search \citep[LOSS,][]{GaneshalingamLi2013}
+%	
+%Nearby Supernova Factory \cite[SNF,][]{AlderingAdam2002}
+%
+%HST used in \citep{RiessStrolger2007, SuzukiRubin2012}
+%
+%Modern results systematic limited \citep{ConleyGuy2011, SuzukiRubin2012, ScolnicRest2014}
+%
+%Efforts overcome \citep{IvezicSmith2007, RegnaultConley2009, TonryStubbs2012}

### 2016-08-11 dessn/investigations/gaussianity/paper/gaussianity.tex
A few more sentences

@@
 % out the following line
-\usepackage{newtxtext,newtxmath}
+%\usepackage{newtxtext,newtxmath}
 % Depending on your LaTeX fonts installation, you might get better results with one of these:
@@
 Specifically, we obtain a full posterior surface using our own MCMC fitting methodology, and compare against summary statistics generated using a different MCMC approach (the \verb|mcmc_lc| method in \verb|sncosmo|), nestled sampling (the \verb|nest_lc| method) and minuit (via the \verb|fit_lc| method).
-The MCMC based algorithm utilises the package \verb|emcee| \citep{ForemanMackey2013}, whilst the nestled sampling is based off \verb|nestle| \citep{Skilling2004, Barbary2015}
-
-{\blue
-to investigate, go to lowest level and simulate light curves
-
-to start with, realise only canonical supernova from abs mag with some scatter, using WMAP9 cosmology. Do this for shallow and deep fields, and fit the light curves using different methods.
-
-state skewness introduced as z increases and ston decreases
-
-which creates difference when between mean and max likelihood
-
-section detailing the bias as a function of redshift (and ston if possible). }
+The MCMC based algorithm utilises the package \verb|emcee| \citep{ForemanMackey2013}, whilst the nestled sampling is based off \verb|nestle| \citep{Skilling2004, Barbary2015}, and both algorithms should map out the full posterior surface, giving summary statistics from the mean and covariance of that full surface.
+The third algorithm, invoked via \verb|fit_lc| uses the minuit algorithm \citep{James1975, minuit2015}, and as such is a maximum likelihood estimator and does not map out the full posterior surface.
+For a Gaussian surface, all three algorithms should agree, however for non-Gaussian surfaces the minuit algorithm will be expected to produce different output.
+As such, we test the SALT2 probability surfaces for skewness and fit with the four algorithms described previously.
+The summary statistics are converted into distance modulus via the Philips relation
+\begin{align}
+\mu = m_B - M + \alpha x_1 - \beta c,
+\end{align}
+where we adopt the fiducial $\alpha = 0.14$ and $\beta = 3.15$ values from \citet{BetouleKessler2014}.
+This allows us to directly see the effects of any skewness or non-Gaussianity in any parameter on the cosmologically important distance modulus.
+We can compare the distributions in $\mu$ for the full posterior surface and for summary statistics, and also verify the skewness of the full distribution of $\mu$.
+We investigate any shift in $\langle \mu \rangle$, and change in the standard deviation $\sigma_\mu$, binning by redshift, with results illustrated in Figure~\ref{fig:bias_des}.
+We find strong evidence that, with increasing redshift, the skewness of the SALT2 fit posterior increases, and causes the maximum likelihood estimator to not only produce biased output, but overestimate the uncertainty of the result.
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../output/bias_dessky.png}
+    \caption{The mean bias in {\red fill out caption when updated figure}}
+    \label{fig:bias_des}
+\end{figure*}
+The likelihood surface for one of the skewed supernova fits is shown as an example in Figure \ref{fig:example}, where the effect of the skewness is readily apparent.
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../output/bias_surface.png}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:example}
+\end{figure*}
+To confirm that the skewness is a result of redshift and not simply a result of the general trend of decreasing signal to noise, we plot the skewness as a function of both redshift and signal to noise in Figure \ref{fig:bias_skew}.
+Interestingly, doing this shows a negative skewness for low redshift, low signal-to-noise light curves, and a positive skewness for high redshift light curves, regardless of signal-to-noise.
+\begin{figure*}
+	\includegraphics[width=\textwidth]{../output/bias2.png}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:bias_skew}
+\end{figure*}
+Having confirmed that skewness becomes significant at higher redshift, we now seek to characterise this potential impact on cosmology.
 

### 2016-08-12 dessn/investigations/gaussianity/paper/gaussianity.tex
Adding some images. Need to ensure tight layout

@@
 We find strong evidence that, with increasing redshift, the skewness of the SALT2 fit posterior increases, and causes the maximum likelihood estimator to not only produce biased output, but overestimate the uncertainty of the result.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias_dessky.png}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/bias_dessky.pdf}
     \caption{The mean bias in {\red fill out caption when updated figure}}
     \label{fig:bias_des}
-\end{figure*}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/bias_dessky_deep.pdf}
+	\caption{The mean bias in {\red fill out caption when updated figure}}
+	\label{fig:bias_des_deep}
+\end{figure}
 The likelihood surface for one of the skewed supernova fits is shown as an example in Figure \ref{fig:example}, where the effect of the skewness is readily apparent.
@@
 Interestingly, doing this shows a negative skewness for low redshift, low signal-to-noise light curves, and a positive skewness for high redshift light curves, regardless of signal-to-noise.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias2.png}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/bias.pdf}
 	\caption{{\red Fill out caption in updated figure}}
 	\label{fig:bias_skew}
-\end{figure*}
+\end{figure}
 Having confirmed that skewness becomes significant at higher redshift, we now seek to characterise this potential impact on cosmology.
@@
 Show cosmology results.
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/obs_cosmology_shallow.pdf}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:obs_cosmology_shallow}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/obs_cosmology_deep.pdf}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:obs_cosmology_deep}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/comparison_shallow.pdf}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:comparison_shallow}
+\end{figure}
+\begin{figure}
+	\includegraphics[width=\columnwidth]{../output/comparison_deep.pdf}
+	\caption{{\red Fill out caption in updated figure}}
+	\label{fig:comparison_deep}
+\end{figure}
 

### 2016-08-15 dessn/investigations/gaussianity/paper/gaussianity.tex
Some more updates

@@
 
-One key assumption when doing this is that these summary statistics are in fact gaussian in nature.
+One key assumption when doing this is that these summary statistics are in fact Gaussian in nature.
 Often this is explicitly assumed in the analysis \citep{MarchTrotta2011, RubinAldering2015}, however it is often implicitly assumed \citep{SullivanGuy2011, SuzukiRubin2012, CampbellDAndrea2013, BetouleKessler2014}.
@@
 	\includegraphics[width=\columnwidth]{../output/bias_dessky.pdf}
-    \caption{The mean bias in {\red fill out caption when updated figure}}
+    \caption{The bias in distance modulus $\mu$ for observations generated as per the DES shallow field.
+Supernova results for all algorithms were binned into redshift groups (top two subplots) and the full posterior surface was analysed for skewness (bottom subplot).
+We find evidence for skewness at higher redshifts, leading to diverging results between mean and maximum likelihood algorithms.}
     \label{fig:bias_des}
@@
 	\includegraphics[width=\columnwidth]{../output/bias_dessky_deep.pdf}
-	\caption{The mean bias in {\red fill out caption when updated figure}}
+	\caption{The bias in distance modulus $\mu$ for observations generated as per the DES deep field.
+Supernova results for all algorithms were binned into redshift groups (top two subplots) and the full posterior surface was analysed for skewness (bottom subplot).
+We find evidence for skewness at higher redshifts, however the increased observational time (and subsequent increase in signal-to-noise) results in less biased results than the shallow field.}
 	\label{fig:bias_des_deep}
 \end{figure}
-The likelihood surface for one of the skewed supernova fits is shown as an example in Figure \ref{fig:example}, where the effect of the skewness is readily apparent.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias_surface.png}
-	\caption{{\red Fill out caption in updated figure}}
-	\label{fig:example}
-\end{figure*}
+%The likelihood surface for one of the skewed supernova fits is shown as an example in Figure \ref{fig:example}, where the effect of the skewness is readily apparent.
+
+%\begin{figure*}
+%	\includegraphics[width=\textwidth]{../output/bias_surface.png}
+%	\caption{The posterior surfaces for all algorithms in the SALT2 fit parameters $x_0$, $x_1$, $c$ and $t$, along with the derived value $\mu$.
+The skewness of the marginalised colour distribution has a significant effect on the skewness of the final $\mu$ distribution, which leads to disagreement between the algorithms used.}
+%	\label{fig:example}
+%\end{figure*}
 To confirm that the skewness is a result of redshift and not simply a result of the general trend of decreasing signal to noise, we plot the skewness as a function of both redshift and signal to noise in Figure \ref{fig:bias_skew}.
@@
 	\includegraphics[width=\columnwidth]{../output/bias.pdf}
-	\caption{{\red Fill out caption in updated figure}}
+	\caption{The deviation between the maximum likelihood value for $\mu$ generated by minuit and the mean of the full posterior surfaces.
+Each point on the plot represents a simulated supernova, with the background surface showing a simple second order polynomial fit to the surface.
+We can see that, whilst the bias does decrease as signal-to-noise increases, it is still present for the high redsfhit supernova.}
 	\label{fig:bias_skew}
@@
 \label{sec:cosmology}
-Fit cosmology against simulated SN
 
-Do for des shallow and deep, using survey area to produce z dist plus extra low-z sample (0.05<z<0.2). simulate from light curves, apply selection effect of ston>5to cull bad fits. have <x> shallow.
-
-toy cosmology model.
-adopt $\alpha$ and $\beta$ values for Phillips correction from Betoule2014, assume known gaussian intrinsic scatter, have $\Omega_m$, $w$ and $M_B$ as cosmology parameters - marginalise over $M_B$.
-
-Show cosmology results.
+\subsection{Uncorrected Biases}
+For our initial cosmology comparison we generate a thousand supernova for both the shallow and deep fields using simulation cosmology of $\Omega_m = 0.30$, $w = -1.0$, $M_B = -19.3$ and $\sigma_{\rm int} = 0.1$.
+10\% of the supernova are drawn from a linear redshift distribution ($0.05 < z < 0.2$) included emulate an external dataset used to constrain the low redshift range of the Hubble diagram, with the remaining 90\% of supernova drawn from a redshift distribution following DES survey volume and $z < 1.2$.
+The thousand supernova all have the selection effect of having at least one point above a signal-to-noise of 5.
+We take these thousand supernova and fit their light curves (generated with the same cadence, $t_0$ and weather effects as discussed in the previous section) using both MCMC methods and maximum likelihood methods.
+As the bias comes from a cut before fitting the light curves, both the set of summary statistics from MCMC and maximum likelihood fitting have the same bias, and any difference between fits to the different sets of summary statistics is thus a product of divergence between the two methodologies and not the cause of selection effects.
+We fix $\alpha$ and $\beta$ as done previously, set $\sigma_{\rm int} = 0.1$, and fit a simple model of $\Omega_m$, $w$ and $M_B$.
 
@@
 	\includegraphics[width=\columnwidth]{../output/obs_cosmology_shallow.pdf}
-	\caption{{\red Fill out caption in updated figure}}
+	\caption{A Hubble diagram sowing the underlying simulation cosmology and the resulting 1000 supernova distance moduli for observations in the DES shallow field.
+Each supernova is shown twice, once when using summary statistics based in the mean to calculate the distance modulus, and then when using maximum likelihood to generate summary statistics and calculate the distance modulus.
+Selection effects heavily prune high redshift supernova.}
 	\label{fig:obs_cosmology_shallow}
@@
 	\includegraphics[width=\columnwidth]{../output/obs_cosmology_deep.pdf}
-	\caption{{\red Fill out caption in updated figure}}
+	\caption{A Hubble diagram sowing the underlying simulation cosmology and the resulting 1000 supernova distance moduli for observations in the DES deep field.
+Each supernova is shown twice, once when using summary statistics based in the mean to calculate the distance modulus, and then when using maximum likelihood to generate summary statistics and calculate the distance modulus.
+The increased observation in the deep field gives us more high redshift supernova than the shallow field.}
 	\label{fig:obs_cosmology_deep}
@@
 
-%\begin{figure}
-%	% To include a figure from a file named example.*
-%	% Allowable file formats are eps or ps if compiling using latex
-%	% or pdf, png, jpg if compiling using pdflatex
-%	\includegraphics[width=\columnwidth]{example}
-%    \caption{This is an example figure. Captions appear below each figure.
-%	Give enough detail for the reader to understand what they're looking at,
-%	but leave detailed discussion to the main body of the text.}
-%    \label{fig:example_figure}
-%\end{figure}
-%
-%% Example table
-%\begin{table}
-%	\centering
-%	\caption{This is an example table. Captions appear above each table.
-%	Remember to define the quantities, symbols and units used.}
-%	\label{tab:example_table}
-%	\begin{tabular}{lccr} % four columns, alignment for each
-%		\hline
-%		A & B & C & D\\
-%		\hline
-%		1 & 2 & 3 & 4\\
-%		2 & 4 & 6 & 8\\
-%		3 & 5 & 7 & 9\\
-%		\hline
-%	\end{tabular}
-%\end{table}
+Figures~\ref{fig:obs_cosmology_shallow} and \ref{fig:obs_cosmology_deep} show the Hubble diagrams (and simulation cosmology) for the shallow and deep simulated supernova respectively.
+Visible in the diagrams is the systematic bias between the two methodologies at high redshift.
+Marginalising over the absolute magnitude $M_B$, the cosmological surfaces are presented in Figures~\ref{fig:comparison_shallow} and \ref{fig:comparison_deep}.
+The shallow field results show a highly significant bias, with even the deep field showing approximately a $1\sigma$ shift when using different methodologies for generating the summary statistics.
+Table \ref{tab:unbias_comp} shows the marginalised parameter summaries for comparison, and the shift between summary statistics is significant.
+
+\begin{table}
+	\centering
+	\caption{Comparison between the shallow and deep field cosmology for the mean and maximum likelihood summary statistics.}
+	\label{tab:unbias_comp}
+    \begin{tabular}{cccc}
+       	\hline
+       	Model & $\Omega_m$ & $w$ & $M$ \\ 
+       	\hline
+       	Deep MCMC & $0.317^{+0.027}_{-0.038}$ & $-1.08^{+0.12}_{-0.15}$ & $-19.290^{+0.016}_{-0.017}$ \\ 
+       	Deep Max. Like. & $0.362^{+0.024}_{-0.030}$ & $-1.29^{+0.17}_{-0.15}$ & $-19.299^{+0.014}_{-0.020}$ \\ 
+       	Shallow MCMC & $0.341^{+0.035}_{-0.051}$ & $-1.17^{+0.19}_{-0.17}$ & $-19.289^{+0.014}_{-0.017}$ \\ 
+       	Shallow Max. Like. & $0.426^{+0.018}_{-0.031}$ & $-1.59^{+0.23}_{-0.18}$ & $-19.306^{+0.014}_{-0.018}$ \\ 
+       	\hline
+    \end{tabular}
+\end{table}
+
+The current investigation is not sufficient grounds to claim that any previous cosmology analyses that use maximum likelihood statistics are biased, as this depends strongly on the method of bias correction utilised in prior analyses.
+For example, methods in which biases are computed from the underlying cosmology may be prone to biases, whilst methods rely on adjusting the data to take into account selection effects calibrated with simulations \citep{KesslerBecker2009, ConleyGuy2011} may be consistent, given simulation SALT2 parameter distributions are drawn from distributions also fit using the observed maximum likelihood summary statistics.
+Similarly, Bayesian methods, such as modelling selection effects as a Gaussian CDF \citep{RubinAldering2015} are algorithm independent and may correct both sets of summary statistics equally well.
+
+\subsection{Simulation Bias Correction}
+
+Try and do it the Rick Kessler way, verify everything is a-ok
+
+\subsection{Modelled Corrections}
+
+Try and do it the Rubin et al way
+
+\subsection{Cosmological Corrections}
+
+Try and do it the Alex Kim way.
 

### 2016-08-15 dessn/investigations/gaussianity/des_sky.py

@@ -337,7 +337,7 @@ def run_analysis(foldername, zeropoints, scatter, savefigure=False, n=400):
         from matplotlib import gridspec
         from matplotlib.ticker import MaxNLocator
         gs = gridspec.GridSpec(3, 1, height_ratios=[2, 2, 1], hspace=0.0, wspace=0.0)
-        fig = plt.figure(figsize=(4.5, 8))
+        fig = plt.figure(figsize=(5, 9))
         ax1 = fig.add_subplot(gs[0])
         ax2 = fig.add_subplot(gs[1], sharex=ax1)
         ax3 = fig.add_subplot(gs[2], sharex=ax1)

### 2016-08-15 dessn/investigations/gaussianity/simple_cosmology_test.py

@@ -28,7 +28,7 @@ def get_zp_and_name(shallow):
         return [34.24, 34.85, 34.94, 35.42], "deep"
 
 
-def get_supernova_data(n=1500, ston_thresh=5, shallow=True):
+def get_supernova_data(n=1500, nact=1000, ston_thresh=5, shallow=True):
     zp, name = get_zp_and_name(shallow)
     temp_dir = os.path.dirname(__file__) + "/output/supernova_data_%s" % name
     if not os.path.exists(temp_dir):
@@ -38,9 +38,8 @@ def get_supernova_data(n=1500, ston_thresh=5, shallow=True):
     res = np.array([r for r in ress if r is not None])
     ston = res[:, 6]
     res = res[ston > ston_thresh, :]
-
+    res = res[:nact, :]
     diff = np.abs(res[:, 7] - (WMAP9.distmod(res[:, 1]).value - 19.3))
-    print(diff)
     res = res[(diff < 4), :]
     # [seed, z, t0, x0, x1, c, ston] + mus + stds)
     #            zs, mu_mcmc, mu_minuit, std_mcmc, std_minuit
@@ -100,13 +99,13 @@ def plot_cosmology(zs, mu_mcmc, mu_minuit, std_mcmc, std_minuit, n):
     ms = 2
     alpha = 0.4
     axes[0].errorbar(zs, mu_minuit, yerr=np.sqrt(std_minuit*std_minuit + 0.1*0.1),
-                     ms=ms, fmt='o', label=r"minuit", color="r", alpha=alpha)
+                     ms=ms, fmt='o', label=r"Max. Likelihood", color="r", alpha=alpha)
     axes[0].errorbar(zs, mu_mcmc, yerr=np.sqrt(std_mcmc*std_mcmc + 0.1*0.1), fmt='o',
-                     ms=ms, label=r"mcmc", color="b", alpha=alpha)
+                     ms=ms, label=r"MCMC", color="b", alpha=alpha)
     axes[1].errorbar(zs, mu_minuit - distmod2, yerr=np.sqrt(std_minuit*std_minuit + 0.1*0.1),
-                     ms=ms, fmt='o', label=r"minuit", color="r", alpha=alpha)
+                     ms=ms, fmt='o', label=r"Max. Likelihood", color="r", alpha=alpha)
     axes[1].errorbar(zs, mu_mcmc - distmod2, yerr=np.sqrt(std_mcmc*std_mcmc + 0.1*0.1), fmt='o',
[…]

### 2016-08-15 dessn/investigations/gaussianity/ston_and_redshift_small.py

@@ -269,7 +269,9 @@ if __name__ == "__main__":
     ax.set_xlim(z.min(), z.max())
     div1 = make_axes_locatable(ax)
     cax1 = div1.append_axes("right", size="5%", pad=0.05)
-    plt.colorbar(h, cax=cax1)
+    cb = plt.colorbar(h, cax=cax1)
+    cb.set_label(r"$\Delta \mu$")
+    cb.ax.tick_params(labelsize=8)
     if False:
         for i, zp, sp in zip(seeds, z, s):
             ax.text(zp, sp, "%d" % i, alpha=0.3)

### 2017-05-17 papers/general_method/general.tex
Adding pdf notes I have to project

\documentclass[a4paper,fleqn,usenatbib]{mnras}
\pdfoutput=1
% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{epstopdf}
\usepackage{aas_macros}
\usepackage{braket}
\usepackage{verbatim}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{appendix}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage[normalem]{ulem}
\usepackage{array}
\bibliographystyle{mnras}

\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand{\tick}{\checkmark}
\newcommand{\gtick}{\color{ForestGreen} \tick }
\newcommand{\cross}{$\times$ }
\newcommand{\rcross}{\color{red} \cross }
\newcommand{\runz}{\textsc{Runz}}
\newcommand{\brac}[1]{\left( #1 \right)}
\newcommand*\mean[1]{\bar{#1}}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\newcommand{\green}{\color{green}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}
\newcommand{\orange}{\color{BurntOrange}}
\newcommand{\purple}{\color{Fuchsia}}
\newcommand{\hMpc}{h^{-1} {\rm Mpc}} % to be used in math mode
\newcommand{\camb}{\textsc{camb}}
\newcommand{\cosmomc}{\textsc{cosmomc}}

\newcommand{\kmsmpc}{km\,s$^{-1}$\,Mpc$^{-1}$}

\newcommand{\halofit}{\textsc{halofit}}
\newcommand{\specialcell}[2][c]{\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

%\shortauthors{}
\title[MC Corrections for Bayesian Methods]{Monte Carlo Corrections for Bayesian Methods}

\author[S. R. Hinton et al.]{Samuel R. Hinton,$^{1,2}$\thanks{E-mail: \href{samuelreay@gmail.com}}
\\
% List of institutions
$^{1}$School of Mathematics and Physics, The University of Queensland, Brisbane, QLD 4072, Australia\\
$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)
}
% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

\begin{document}

\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

\begin{abstract}
I present a rigorous treatment for truncated or biased datasets by combining Bayesian analysis methods with Monte Carlo simulations.
\end{abstract}

%\vspace{35mm}

\section{Introduction}

Find some treatment of biased or truncated data before. Maybe \citet{Gull1989bayesian}.

\section{The perfect world}

In a perfect world, data is neither biased nor truncated. The data is perfect. Uncertainties are well quantified and normally distributed around true values. Presumably everything is also spherical and in a vacuum. Let us create a mock model in this perfect world. Let us observe a series of iid events $\vec{x}$, which is known perfectly and drawn from a normal distribution such that 
\begin{align}
\vec{x} \sim \mathcal{N}(\mu,\sigma)
\end{align}
If, having collected our observations of $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface. Taking uniform priors on both parameters we simply wish to map the surface
\begin{align}
P(\mu,\sigma| \vec{x}) &\propto P(\vec{x} | \mu, \sigma) P(\mu, \sigma) \\
P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma).
\end{align}
Generating a hundred data points with $\mu=100,\ \sigma=10$, we can recover our input parameters easily, as shown in Figure \ref{fig:perfect}.
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{example/perfect.pdf}
	\end{center}
	\caption{A systematic test of our perfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points. Posterior surface mapped out using emcee.}
	\label{fig:perfect}
\end{figure}

\section{The Imperfect World}

In a slightly imperfect world we may have to deal with something like truncated data. For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$, assigning a value $\alpha=85$ for convenience. If we don't take this truncation into account, we will recovered biased parameter estimates, as shown in Figure \ref{fig:imperfect}. However, we can correct for this truncation. If restate our likelihood to take into account some selection effect $S$, our likelihood for a single event can be stated as
\begin{align}
\mathcal{L} &= P(x | \theta, S)\\
&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)} \\
&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S, D|\theta)\, dD} \\
&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD},
\end{align}
where $D$ is a potential observable. In our example, the selection efficiency is a step function, having observed $x$, $P(S|x,\theta) = 1$. To substitute in our normal model,
\begin{align}
\mathcal{L} &= \frac{ \mathcal{N}(x|\mu, \sigma)}{\int_{-\infty}^\infty \mathcal{H}(D - \alpha) \mathcal{N}(D|\mu, \sigma)\, dD} \\
&= \frac{ \mathcal{N}(x|\mu, \sigma)}{\int_{\alpha}^\infty \mathcal{N}(D|\mu, \sigma)\, dD} \\
&= \frac{ \mathcal{N}(x|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
\end{align}
if $\mu > \alpha$. We can add this correction to our model, and note that we now recover unbiased parameter estimates, also shown in Figure \ref{fig:imperfect}.
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{example/imperfect.pdf}
	\end{center}
	\caption{A systematic test of our imperfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.}
	\label{fig:imperfect}
\end{figure}

\section{The Real World}

Unfortunately it is a rare scenario when dealing with nature and all her faults for us to be able to have a perfect step function or for selection efficiency to be wholly encapsulated by a single parameter. A more realistic scenario involves a selection efficiency which cannot be conveniently described by an analytic function and instead is a high dimensional non-analytic function. And its probably stochastic too, just to throw another wrench in the works.

\section{Conclusion}
\label{sec:conclusion}

Still da bes

\section*{Acknowledgments}

We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.

\bibliography{bibliography}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2017-05-17 papers/general_method/example/imperfect.py

@@ -1,0 +1,49 @@
+import numpy as np
+import emcee
+from scipy.stats import norm
+from scipy.special import erfc
+
+mu, sigma, alpha = 100, 10, 85
+ndim, nwalkers = 2, 10
+s2 = np.sqrt(2)
+def lnprob(theta, data):
+    if theta[1] < 0:
+        return -np.inf
+    return np.sum(norm.logpdf(data, theta[0], theta[1]))
+
+def lnprob2(theta, data):
+    mu, sigma = theta
+    if sigma < 0:
+        return -np.inf
+    if mu < alpha:
+        return -np.inf
+    return np.sum(norm.logpdf(data, mu, sigma) - np.log(0.5 * erfc((alpha - mu)/(s2 * sigma))))
+
+all_samples, all_sampels_corrected = [], []
+for i in range(100):
+    np.random.seed(i)
+    
+    x = np.random.normal(loc=mu, scale=sigma, size=1000)
+    mask = x > alpha
+    x = x[mask][:100]
+    p0 = [[np.random.normal(100,10), np.random.normal(15, 2)] for i in range(nwalkers)]
+    
+    sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=[x])
+    sampler.run_mcmc(p0, 2000)
+    
+    sampler2 = emcee.EnsembleSampler(nwalkers, ndim, lnprob2, args=[x])
+    sampler2.run_mcmc(p0, 2000)
+    
+    all_samples.append(sampler.chain[:, 100:, :].reshape((-1, ndim)))
+    all_sampels_corrected.append(sampler2.chain[:, 100:, :].reshape((-1, ndim)))
+
+
+all_samples = np.vstack(all_samples)
+all_sampels_corrected = np.vstack(all_sampels_corrected)
+
+from chainconsumer import ChainConsumer
+c = ChainConsumer()
+c.add_chain(all_samples, parameters=[r"$\mu$", r"$\sigma$"], name="Biased")
+c.add_chain(all_sampels_corrected, parameters=[r"$\mu$", r"$\sigma$"], name="Corrected")
+c.configure(flip=False, sigmas=[0,1,2], colors=["#D32F2F", "#4CAF50"], linestyles=["-", "--"], shade_alpha=0.2)
+c.plot(filename="imperfect.pdf", figsize="column", truth=[100,10], extents=[[95, 105],[7,14]])

### 2017-05-17 papers/general_method/example/perfect.py

@@ -1,0 +1,32 @@
+import numpy as np
+import emcee
+from scipy.stats import norm
+
+
+mu, sigma = 100, 10
+ndim, nwalkers = 2, 10
+
+def lnprob(theta, data):
+    if theta[1] < 0:
+        return -np.inf
+    return np.sum(norm.logpdf(data, theta[0], theta[1]))
+
+all_samples = []
+for i in range(100):
+    np.random.seed(i)
+    x = np.random.normal(loc=mu, scale=sigma, size=100)
+    p0 = [[np.random.normal(100,10), np.random.normal(15, 2)] for i in range(nwalkers)]
+    
+    sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=[x])
+    sampler.run_mcmc(p0, 2000)
+    all_samples.append(sampler.chain[:, 100:, :].reshape((-1, ndim)))
+
+
+all_samples = np.vstack(all_samples)
+
+from chainconsumer import ChainConsumer
+c = ChainConsumer()
+c.add_chain(all_samples, parameters=[r"$\mu$", r"$\sigma$"])
+c.configure(flip=False, sigmas=[0,1,2])
+c.plot(filename="perfect.pdf", figsize="column", truth=[100,10])
+

### 2017-05-17 papers/general_method/general.tex
Real

@@
 
-Unfortunately it is a rare scenario when dealing with nature and all her faults for us to be able to have a perfect step function or for selection efficiency to be wholly encapsulated by a single parameter.
-A more realistic scenario involves a selection efficiency which cannot be conveniently described by an analytic function and instead is a high dimensional non-analytic function.
+Unfortunately it is a rare scenario when dealing with nature and all her faults for us to have an analytic selection function.
+Let alone a function encapsulated by a single parameter.
+A more realistic scenario involves a selection efficiency which cannot be conveniently described by an analytic function.
+Instead, we would have a high dimensional non-analytic function.
 And its probably stochastic too, just to throw another wrench in the works.

### 2017-05-17 papers/general_method/example/real.py

@@ -1,0 +1,84 @@
+import numpy as np
+import emcee
+from scipy.stats import norm
+from scipy.special import erfc
+
+mu, sigma, alpha = 100, 10, 82
+ndim, nwalkers = 2, 10
+num = 50
+
+s2 = np.sqrt(2)
+def lnprob(theta, data):
+    if theta[1] < 0:
+        return -np.inf
+    return np.sum(norm.logpdf(data, theta[0], theta[1]))
+
+def lnprob2(theta, data):
+    mu, sigma = theta
+    if sigma < 0:
+        return -np.inf
+    if mu < alpha:
+        return -np.inf
+    return np.sum(norm.logpdf(data, mu, sigma) - np.log(0.5 * erfc((alpha - mu)/(s2 * sigma))))
+
+all_samples, all_sampels_corrected = [], []
+
+def get_data(mu, sigma, n=1000):
+    x = np.random.normal(loc=mu, scale=sigma, size=n)
+    mask1 = x > alpha
+    mask2 = x > 85 #(mu - 2 * sigma)
+    mask = mask1 & mask2
+    return x, mask, mask1, mask2
+    
+def reweight(mu, sigma):
+    original_weight = 0.5 * erfc((alpha - mu)/(s2 * sigma))
+    x, mask, mask1, mask2 = get_data(mu, sigma, n=100000)
+    new_weight = mask.sum() * 1.0 / mask.size
+    diff = np.log(original_weight) - np.log(new_weight)    
+    return num * diff    
+
+val1 = 0.5 * erfc((alpha - mu)/(s2 * sigma))
+x, mask, mask1, mask2 = get_data(mu, sigma, n=100000)
+val2 = mask.sum() * 1.0 / mask.size
+print(val1, val2)
+import matplotlib.pyplot as plt
+plt.hist(x, 50, histtype='step', label="all")
+plt.hist(x[mask1], 50, histtype='step', label="mask1")
+plt.hist(x[mask2], 50, histtype='step', label="mask2")
+plt.hist(x[mask], 50, histtype='step', label="both")
+plt.legend()
+plt.show()
+        
+    
+    
+for i in range(100):
+    np.random.seed(i)
+    
+    x, mask, mask1, mask2 = get_data(mu, sigma)
+    x = x[mask][:100]
+    p0 = [[np.random.normal(100,10), np.random.normal(15, 2)] for i in range(nwalkers)]
+    
[…]

### 2017-05-18 papers/general_method/general.tex
Real world example

@@
 And its probably stochastic too, just to throw another wrench in the works.
+Provided a method of forward modelling or simulating observations, the solution is Monte Carlo integration combined with an analytic approximate correction.
+
+So let us extend our toy model.
+We observe not just $x \sim \mathcal{N}(\mu, \sigma)$, but also another independent variable, $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
+Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$.
+Our likelihood for such a toy model becomes
+\begin{align}
+\mathcal{L} &= \frac{ \mathcal{N}(x|\mu, \sigma) \mathcal{N}(y|\mu_y, \sigma_y)}
+{\iint_{-\infty}^\infty \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}
+\end{align}
+Assume that we cannot solve these integral analytically, and must resort to numeric solutions.
+These often clash with sampling methods, especially for high dimensional integrals.
+Inserting Monte Carlo integration into fitting algorithms can drastically slow them down, and such as Hamiltonian MCMC that require continuous surfaces can easily fail on surfaces that, via Monte Carlo integration, fluctuate.
+Even by fixing the samples used in MC integration (thereby giving a continuous surface), the complexity of the surface derivatives will pose almost insurmountable problems for any algorithms that utilise surface gradients.
+One solution is to find an approximate, analytic correction we can utilise in our fitting algorithm which seeks to shift the region of parameter space sampled by the sampler closer to the correct area.
+
+In our example, if $\beta \ll 1$, such that the majority of selection effect is encapsulated by $x$ and not $y$, our approximate correction can take the form found in the previous example.
+Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
+Assuming some prior, imperfect knowledge of $mu_y$  we estimate that the average contribution from $\beta y$ is around $4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
+\begin{align}
+w = \frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu - 4}{\sqrt{2}\sigma} \right]
+\end{align}
+Further, let us explicitly break our likelihood into two parts, $\mathcal{L} = \mathcal{L}_1 \mathcal{L}_2$, with the parts given by
+\begin{align}
+\mathcal{L}_1 &= \frac{ \mathcal{N}(x|\mu, \sigma) \mathcal{N}(y|\mu_y, \sigma_y)}{w} \\
+\mathcal{L}_2 &= \frac{ w }{\iint_{-\infty}^\infty \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}.
+\end{align}
+$\mathcal{L}_1$ can thus be fitted with a traditional sampler without numeric difficulty or slowdown, and $\mathcal{L}_2$ allows us to calculate the weight of each sample.
+We are effectively importance sampling our likelihood evaluations.
+The computational benefits of this should not be understated either - each sample in our chains can be reweighted independently, providing a task that is trivially parallelisable.
+Evaluating $\mathcal{L}_2$ using Monte Carlo integration of $n$ samples, we have
+\begin{align}
+\mathcal{L}_2 = \frac{ w n }{\sum_{i=1}^{n} \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X_i|\mu, \sigma) \mathcal{N}(Y_i|\mu_y, \sigma_y)}
+\end{align}
+Further tricks can be used to increase the efficiency with which the samples are reweighted.
+Firstly, the overarching analytic model often provides functions which can be drawn from efficiently.
+In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (ie traditional importance sampling) we need only evaluate the step function for our data points.
+This is demonstrated in Figure \ref{fig:real}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{example/real.pdf}
+	\end{center}
+	\caption{A systematic test of our more complicated model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.
+The likelihood $\mathcal{L}_1$ was evaluated with the fitting algorithm emcee, and reweighted using Monte Carlo integration of a hundred thousand possible events as per $\mathcal{L}_2$.
+The truncated data with no correction is shown as `Biased' in dotted red, the `Approximate` only correction ($\mathcal{L}_1$) shown in dashed green, and the final reweighted chain shown in solid black as `Corrected'.}
+	\label{fig:real}
+\end{figure}
+
+Alternatively, if evaluating the probability that an event is observed is numerically expensive, it is easy to pregenerate a set of events and reuse them for all weights - provided that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
+This method is however only efficient when prior knowledge of parameter values is known to allow a reasonable initial draw of events.
 

### 2017-05-18 papers/general_method/example/real.py

@@ -71,7 +71,7 @@ def get_stuff(i):
     weights = np.exp(weights)
     return (chain1, chain2, weights)
 
-res = Parallel(n_jobs=4, backend="threading")(delayed(get_stuff)(i) for i in range(32))
+res = Parallel(n_jobs=4, backend="threading")(delayed(get_stuff)(i) for i in range(100))
 all_samples = np.vstack([r[0] for r in res])
 all_sampels_corrected = np.vstack([r[1] for r in res])
 weights = np.array([r[2] for r in res]).flatten()
@@ -79,8 +79,8 @@ print(weights.max(), weights.min(), weights.mean())
 
 from chainconsumer import ChainConsumer
 c = ChainConsumer()
-c.add_chain(all_samples, parameters=[r"$\mu_x$", r"$\sigma_x$", r"$\mu_y$"], name="Biased")
+c.add_chain(all_samples, parameters=[r"$\mu$", r"$\sigma$", r"$\mu_y$"], name="Biased")
 c.add_chain(all_sampels_corrected, name="Approximate")
 c.add_chain(all_sampels_corrected, weights=weights, name="Corrected")
-c.configure(flip=False, sigmas=[0,1,2], colors=["#D32F2F", "#4CAF50", "#333333"], linestyles=["-", "--", ":"], shade_alpha=0.2, shade=True)
+c.configure(flip=False, sigmas=[0,1,2], colors=["#D32F2F", "#4CAF50", "#222222"], linestyles=[":", "--", "-"], shade_alpha=0.2, shade=True)
 c.plot(filename="real2.pdf", figsize="column", truth=[100, 10], extents=[[90, 110],[6,15]], parameters=2)

### 2017-05-18 dessn/framework/models/__init__.py
Why wont you work

@@ -118,7 +118,7 @@ shown under the Population header as :math:`\gamma`.
 
 Furthermore, in the interests of simplicity, let us examine only a single supernova for the time being. As I don't
 yet take calibration into account, the supernova are independent. Let us denote the unnormalised likelihood for a single
-supernova as :math:`P (D_i|\theta).
+supernova as :math:`P (D_i|\theta)`.
 
 .. math::
     :label: d

### 2017-05-18 papers/general_method/general.tex
Why wont you work
Adding intro

@@
 
-Find some treatment of biased or truncated data before. Maybe \citet{Gull1989bayesian}.
+Truncated data is a problem in many areas of scientific inquiry.
+It is one of the primary difficulties when performing supernovae cosmology analysis, as our telescopes have visual limits that truncate our observed supernovae distribution from the actual underlying distribution.
+This bias, termed Malmquist bias, is source of much investigation \citep{Butkevich2005}.
+It is considered during analysis by either modifying the observed data to remove the expected bias \citep{BetouleKessler2014, ConleyGuySullivan2011}, or by incorporating the expected bias into the underlying model \citep{Rubin2015}.
+Truncated data is also commonly encountered in biological fields, where data such as mortality rates are left-truncated \citep{JANE1898}.
+Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with different fitting algorithms \citep{Gelfand1992}.
+Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
+As such, this work provides a simple treatment of truncated data in a common Bayesian technique.
+Section \ref{sec:perfect} discusses the ever-elusive case of perfect data without truncation to provide a common basis for Sections \ref{sec:imperfect} and \ref{sec:real}, which respectively cover analytically correctable data truncation and analytically intractable models.
+Section \ref{sec:tricks} details numeric
 
 \section{The perfect world}
-
+\label{sec:perfect}
 In a perfect world, data is neither biased nor truncated.
@@
 \section{The Imperfect World}
-
+\label{sec:imperfect}
 In a slightly imperfect world we may have to deal with something like truncated data.
@@
 \section{The Real World}
-
+\label{sec:real}
 Unfortunately it is a rare scenario when dealing with nature and all her faults for us to have an analytic selection function.
@@
 
+\section{Numerical Tricks}
+\label{sec:tricks}
 Alternatively, if evaluating the probability that an event is observed is numerically expensive, it is easy to pregenerate a set of events and reuse them for all weights - provided that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.

### 2017-05-18 papers/general_method/general.tex
Adding moreeee

@@
 \begin{abstract}
-I present a rigorous treatment for truncated or biased datasets by combining Bayesian analysis methods with Monte Carlo simulations.
+Truncated datasets are a pervasive problem in modern science.
+In this paper I present a simple overview of a Bayesian consideration of truncation, giving a solution to analytically tractable models.
+Additionally, truncation which cannot be analytically corrected is shown to be correctable easily via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
+Toy models are included, along with numerical considerations for implementation.
 \end{abstract}
@@
 Section \ref{sec:perfect} discusses the ever-elusive case of perfect data without truncation to provide a common basis for Sections \ref{sec:imperfect} and \ref{sec:real}, which respectively cover analytically correctable data truncation and analytically intractable models.
-Section \ref{sec:tricks} details numeric
+Section \ref{sec:tricks} details numeric concerns and tricks to be aware of for effective implementation of Monte Carlo corrections applied to analytic approximations.
 
@@
 \end{align}
-Further tricks can be used to increase the efficiency with which the samples are reweighted.
-Firstly, the overarching analytic model often provides functions which can be drawn from efficiently.
-In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (ie traditional importance sampling) we need only evaluate the step function for our data points.
-This is demonstrated in Figure \ref{fig:real}.
 
@@
 \end{figure}
+Thus we end up with a corrected posterior surface as shown in Figure \ref{fig:real}.
 
@@
 \label{sec:tricks}
-Alternatively, if evaluating the probability that an event is observed is numerically expensive, it is easy to pregenerate a set of events and reuse them for all weights - provided that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
+Further tricks can be used to increase the efficiency with which the samples are reweighted.
+Firstly, the overarching analytic model often provides functions which can be drawn from efficiently.
+In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (ie traditional importance sampling) we need only evaluate the step function for our data points.
+
+If evaluating the probability that an event is observed is numerically expensive (i.e.
+not a step function), it is easy to pregenerate a set of events and reuse them for all weights - provided that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
 This method is however only efficient when prior knowledge of parameter values is known to allow a reasonable initial draw of events.
+Without this prior information, samples need to span the entire posterior volume, which is numerically intractable even for low dimensional models.
 
@@
 
-Still da bes
+Reword the abstract or similar
 

### 2017-05-19 papers/general_method/general.tex
feedback

@@
 \begin{abstract}
-Truncated datasets are a pervasive problem in modern science.
+{\red Add example.
+Make abstract less boring} Truncated datasets are a pervasive problem in modern science.
 In this paper I present a simple overview of a Bayesian consideration of truncation, giving a solution to analytically tractable models.
@@
 Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
-As such, this work provides a simple treatment of truncated data in a common Bayesian technique.
+This work provides a simple treatment of truncated data in a common Bayesian technique.
 Section \ref{sec:perfect} discusses the ever-elusive case of perfect data without truncation to provide a common basis for Sections \ref{sec:imperfect} and \ref{sec:real}, which respectively cover analytically correctable data truncation and analytically intractable models.
@@
 
-\section{The perfect world}
+\section{The perfect world: No Truncation}
 \label{sec:perfect}
@@
 Let us create a mock model in this perfect world.
-Let us observe a series of iid events $\vec{x}$, which is known perfectly and drawn from a normal distribution such that
+Let us observe a series of independent and identically distributed events $\vec{x}$, which are drawn from a normal distribution such that
 \begin{align}
-\vec{x} \sim \mathcal{N}(\mu,\sigma)
+\vec{x} \sim \mathcal{N}(\mu,\sigma).
 \end{align}
-If, having collected our observations of $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface.
+If, having collected our observations $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface.
 Taking uniform priors on both parameters we simply wish to map the surface
 \begin{align}
-P(\mu,\sigma| \vec{x}) &\propto P(\vec{x} | \mu, \sigma) P(\mu, \sigma) \\
+P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
+\end{align}
+where our model parameters $\theta = \lbrace \mu, \sigma \rbrace$ and our data is given by $\vec{x}$.
+\begin{align}
+P(\mu,\sigma| \vec{x}) &\propto P(\vec{x} | \mu, \sigma) P(\mu, \sigma)
+\end{align}
+With uniform priors, $P(\mu,\sigma) = {\rm constant}$, and can be absorbed into the constant of proportionality. Expanding our observation vector, the posterior surface is given by
+\begin{align}
 P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma).
@@
 	\end{center}
-	\caption{A systematic test of our perfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points. Posterior surface mapped out using emcee.}
+	\caption{A systematic test of our perfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points.
+Any systematic offset in our model would be revealed by a shift in the stacked results away from the true parameter values.}
 	\label{fig:perfect}
@@
 
-\section{The Imperfect World}
+\section{The Imperfect World: Analytic Truncation}
 \label{sec:imperfect}
@@
 For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$, assigning a value $\alpha=85$ for convenience.
-If we don't take this truncation into account, we will recovered biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
+If we do not take this truncation into account, we will recover biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
 However, we can correct for this truncation.
-If restate our likelihood to take into account some selection effect $S$, our likelihood for a single event can be stated as
+If we restate our likelihood to take into account some selection effect $S$, our likelihood can be colloquially stated as the probability of observation given our model parameters \textit{and} our selection effects.
+Mathematically for a single event this can be stated as
 \begin{align}
 \mathcal{L} &= P(x | \theta, S)\\
-&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)} \\
+&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)}
+\end{align}
+Introducing an integral over all possible data to make $P(S|\theta)$ physical,
+\begin{align}
 &= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S, D|\theta)\, dD} \\
-&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD},
+&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}.
 \end{align}
-where $D$ is a potential observable. In our example, the selection efficiency is a step function, having observed $x$, $P(S|x,\theta) = 1$. To substitute in our normal model,
+In our example, the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
+Having successfully observed $x$, it follows that $x > \alpha$ and so $P(S|x,\theta) = 1$.
+To substitute in our normal model,
 \begin{align}
@@
 \end{align}
-if $\mu > \alpha$. We can add this correction to our model, and note that we now recover unbiased parameter estimates, also shown in Figure \ref{fig:imperfect}.
+where in the last line I have evaluated the integral in the case $\mu > \alpha$.
+We can add this correction to our model, and note that we now recover unbiased parameter estimates, also shown in Figure \ref{fig:imperfect}.
 \begin{figure}
@@
 	\end{center}
-	\caption{A systematic test of our imperfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.}
+	\caption{A systematic test of our imperfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.
+The bias shown in the red `Biased' contour can be corrected to via the techniques shown in Section \ref{sec:imperfect} to recover unbiased surfaces.}
 	\label{fig:imperfect}
@@
 
-\section{The Real World}
+\section{The Real World: Intractable Truncation}
 \label{sec:real}
-Unfortunately it is a rare scenario when dealing with nature and all her faults for us to have an analytic selection function.
-Let alone a function encapsulated by a single parameter.
-A more realistic scenario involves a selection efficiency which cannot be conveniently described by an analytic function.
-Instead, we would have a high dimensional non-analytic function.
-And its probably stochastic too, just to throw another wrench in the works.
-Provided a method of forward modelling or simulating observations, the solution is Monte Carlo integration combined with an analytic approximate correction.
+Unfortunately it is a rare scenario when dealing with nature and all her faults for us to have an analytic selection function, let alone a function encapsulated by a single parameter.
+A more realistic scenario involves a selection efficiency instead would take the form of non-analytic function of many model parameters.
+And the function would probably be stochastic too, just to throw another wrench in the works.
+Provided a method of forward modelling or simulating observations, the solution is to combine an analytic approximate correction with Monte Carlo integration.
 
-So let us extend our toy model.
-We observe not just $x \sim \mathcal{N}(\mu, \sigma)$, but also another independent variable, $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
+So let us modify our imperfect toy model.
+Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
 Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$.
-Our likelihood for such a toy model becomes
+Our likelihood for such a toy model becomes now the combination of probabilities for observing both $x$ and $y$, with the denominator becoming an integral over all possible $X$ and $Y$ observations.
 \begin{align}
@@
 \end{align}
-Assume that we cannot solve these integral analytically, and must resort to numeric solutions.
+Assume that we cannot solve this integral analytically, and must resort to numeric solutions.
 These often clash with sampling methods, especially for high dimensional integrals.
-Inserting Monte Carlo integration into fitting algorithms can drastically slow them down, and such as Hamiltonian MCMC that require continuous surfaces can easily fail on surfaces that, via Monte Carlo integration, fluctuate.
+Inserting Monte Carlo integration into fitting algorithms can drastically slow them down, and algorithms such as Hamiltonian MCMC that require continuous surfaces can easily fail on surfaces that fluctuate from Monte Carlo integration.
 Even by fixing the samples used in MC integration (thereby giving a continuous surface), the complexity of the surface derivatives will pose almost insurmountable problems for any algorithms that utilise surface gradients.
@@
 Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
-Assuming some prior, imperfect knowledge of $mu_y$  we estimate that the average contribution from $\beta y$ is around $4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
+Assuming some prior, imperfect knowledge of $\mu_y$  we estimate that the average contribution from $\beta y$ is around $4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
 \begin{align}

### 2017-05-19 papers/general_method/general.tex
Adding moreeee

@@
 Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
-Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$.
-Our likelihood for such a toy model becomes now the combination of probabilities for observing both $x$ and $y$, with the denominator becoming an integral over all possible $X$ and $Y$ observations.
+Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$, giving $P(S|x,y,\theta) = \mathcal{H}(x + \beta y - \alpha)$.
+Our likelihood for such a toy model becomes now the combination of probabilities for observing both $x$ and $y$, with the denominator becoming an integral over all possible $X$ and $Y$ observations subject to our selection effects.
 \begin{align}
@@
 
-In our example, if $\beta \ll 1$, such that the majority of selection effect is encapsulated by $x$ and not $y$, our approximate correction can take the form found in the previous example.
+In our example, if $\beta \ll 1$, such that the majority of selection effect is encapsulated by $x$ and not $y$, our approximate correction can take the form found in the previous correction from Section \ref{sec:imperfect}.
 Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
-Assuming some prior, imperfect knowledge of $\mu_y$  we estimate that the average contribution from $\beta y$ is around $4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
+Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is around $20\beta = 4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
 \begin{align}
-w = \frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu - 4}{\sqrt{2}\sigma} \right]
+w = \frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu - 4}{\sqrt{2}\sigma} \right].
 \end{align}

### 2017-05-19 papers/general_method/example/perfect.py

@@ -27,6 +27,6 @@ all_samples = np.vstack(all_samples)
 from chainconsumer import ChainConsumer
 c = ChainConsumer()
 c.add_chain(all_samples, parameters=[r"$\mu$", r"$\sigma$"])
-c.configure(flip=False, sigmas=[0,1,2])
+c.configure(flip=False, sigmas=[0,1,2], summary=False)
 c.plot(filename="perfect.pdf", figsize="column", truth=[100,10])
 

### 2017-05-19 papers/general_method/general.tex
Clarifying imperfect model

@@
 \begin{abstract}
-{\red Add example.
-Make abstract less boring} Truncated datasets are a pervasive problem in modern science.
-In this paper I present a simple overview of a Bayesian consideration of truncation, giving a solution to analytically tractable models.
-Additionally, truncation which cannot be analytically corrected is shown to be correctable easily via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
-Toy models are included, along with numerical considerations for implementation.
+Lifetimes are strictly positive.
+Fainter objects are harder to see with a telescope than brighter ones.
+We cut our data to remove background noise.
+Truncated or biased datasets are inescapable in modern physics.
+In this paper, I present a simple overview of a Bayesian consideration of truncation, giving a solution to both analytically tractable and intractable models.
+This can be accomplished via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
+Toy models are included, along with numerical considerations and optimisations for implementation.
 \end{abstract}
@@
 In a slightly imperfect world we may have to deal with something like truncated data.
-For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$, assigning a value $\alpha=85$ for convenience.
+For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$.
+I assign a value $\alpha=85$ for convenience.
 If we do not take this truncation into account, we will recover biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
 However, we can correct for this truncation.
-If we restate our likelihood to take into account some selection effect $S$, our likelihood can be colloquially stated as the probability of observation given our model parameters \textit{and} our selection effects.
-Mathematically for a single event this can be stated as
+If we restate our likelihood as the probability of observation given our model parameters \textit{and} our selection effects $S$, we have
 \begin{align}
 \mathcal{L} &= P(x | \theta, S)\\
-&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)}
+&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)}.
 \end{align}

### 2017-05-20 papers/general_method/general.tex
Increasing explanations. Adding comments about conclusion

@@
 %\shortauthors{}
-\title[MC Corrections for Bayesian Methods]{Monte Carlo Corrections for Bayesian Methods}
+\title[Correcting for data truncation in Bayesian analyses]{Correcting for data truncation in Bayesian analyses}
 
 \author[S. R. Hinton et al.]{Samuel R. Hinton,$^{1,2}$\thanks{E-mail: \href{samuelreay@gmail.com}}
+	Alex Kim,$^{3}$
+	Tamara M. Davis,$^{1,2}$
 \\
@@
 $^{1}$School of Mathematics and Physics, The University of Queensland, Brisbane, QLD 4072, Australia\\
-$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)
+$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
+$^{3}$Lawrence Berkeley National Labs
 }
@@
 Truncated or biased datasets are inescapable in modern physics.
-In this paper, I present a simple overview of a Bayesian consideration of truncation, giving a solution to both analytically tractable and intractable models.
+In this paper, we present a simple overview of a Bayesian consideration of truncation, giving a solution to both analytically tractable and intractable models.
 This can be accomplished via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
@@
 \end{abstract}
-
-%\vspace{35mm}
 
@@
 Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
+
 This work provides a simple treatment of truncated data in a common Bayesian technique.
@@
 \begin{align}
-P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma).
+P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma). \label{eq:prod}
 \end{align}
@@
 For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$.
-I assign a value $\alpha=85$ for convenience.
+We assign a value $\alpha=85$ for convenience.
 If we do not take this truncation into account, we will recover biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
@@
 \begin{align}
-\mathcal{L} &= P(x | \theta, S)\\
-&= \frac{P(S|x,\theta) P(x|\theta)}{P(S|\theta)}.
+\mathcal{L} &= P({\rm data} | \theta, S)\\
+&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{P(S|\theta)}.
 \end{align}
@@
 \begin{align}
-&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S, D|\theta)\, dD} \\
-&= \frac{P(S|x,\theta) P(x|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}.
+&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
+&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
 \end{align}
-In our example, the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
+where the integral in the denominator has the same dimensionality is the experimental data.
+\textbf{Equation \ref{eq:main} is the generalised likelihood of experiments with sample selection.} In our example, our data is one dimensional (as we only observe one variable, $x$) and the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
 Having successfully observed $x$, it follows that $x > \alpha$ and so $P(S|x,\theta) = 1$.
@@
 \begin{align}
-\mathcal{L} &= \frac{ \mathcal{N}(x|\mu, \sigma)}{\int_{-\infty}^\infty \mathcal{H}(D - \alpha) \mathcal{N}(D|\mu, \sigma)\, dD} \\
-&= \frac{ \mathcal{N}(x|\mu, \sigma)}{\int_{\alpha}^\infty \mathcal{N}(D|\mu, \sigma)\, dD} \\
-&= \frac{ \mathcal{N}(x|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
+\mathcal{L}_i &= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\int_{-\infty}^\infty \mathcal{H}(D - \alpha) \mathcal{N}(D|\mu, \sigma)\, dD} \\
+&= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\int_{\alpha}^\infty \mathcal{N}(D|\mu, \sigma)\, dD} \\
+&= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
 \end{align}
-where in the last line I have evaluated the integral in the case $\mu > \alpha$.
+where in the last line we have evaluated the integral in the case $\mu > \alpha$.
+Note that this is for a single observation of $x$, and so for a set of independent observations we need to introduce the product found in equation \eqref{eq:prod}.
+\begin{align}
+\mathcal{L} &= \prod_{i=1}^N \frac{ \mathcal{N}(x|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
+\end{align}
+However, as our selection efficiency correction is observationally independent it is identical for all observations, allowing us to take it outside the product.
+\begin{align}
+\mathcal{L} &= 2 \left( {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right] \right)^{-N}\prod_{i=1}^N  \mathcal{N}(x|\mu, \sigma), 
+\end{align}
 We can add this correction to our model, and note that we now recover unbiased parameter estimates, also shown in Figure \ref{fig:imperfect}.
@@
 
-\section{The Real World: Intractable Truncation}
+\section{The Real World: Analytically Intractable Truncation}
 \label{sec:real}
@@
 \begin{align}
-\mathcal{L} &= \frac{ \mathcal{N}(x|\mu, \sigma) \mathcal{N}(y|\mu_y, \sigma_y)}
-{\iint_{-\infty}^\infty \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}
+\mathcal{L}_i &= \frac{ \mathcal{N}(x_i|\mu, \sigma) \mathcal{N}(y_i|\mu_y, \sigma_y)}
+{\iint_{-\infty}^\infty \mathcal{H}(X + \beta Y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}
 \end{align}
@@
 \begin{align}
-w = \frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu - 4}{\sqrt{2}\sigma} \right].
+w_{\rm approx} = \frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu - 4}{\sqrt{2}\sigma} \right].
 \end{align}
-Further, let us explicitly break our likelihood into two parts, $\mathcal{L} = \mathcal{L}_1 \mathcal{L}_2$, with the parts given by
+Further, let us explicitly break our likelihood into two parts, $\mathcal{L}_i = \mathcal{L}_{i1} \mathcal{L}_{i2}$, with the parts given by
 \begin{align}
-\mathcal{L}_1 &= \frac{ \mathcal{N}(x|\mu, \sigma) \mathcal{N}(y|\mu_y, \sigma_y)}{w} \\
-\mathcal{L}_2 &= \frac{ w }{\iint_{-\infty}^\infty \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}.
+\mathcal{L}_{i1} &= \frac{ \mathcal{N}(x_i|\mu, \sigma) \mathcal{N}(y_i|\mu_y, \sigma_y)}{w_{\rm approx} } \\
+\mathcal{L}_{i2} &= \frac{ w_{\rm approx}  }{\iint_{-\infty}^\infty \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}.
 \end{align}
@@
 \begin{align}
-\mathcal{L}_2 = \frac{ w n }{\sum_{i=1}^{n} \mathcal{H}(x + \beta y - \alpha) \mathcal{N}(X_i|\mu, \sigma) \mathcal{N}(Y_i|\mu_y, \sigma_y)}
+\mathcal{L}_{i2} = \frac{ w_{\rm approx}  n }{\sum_{j=1}^{n} \mathcal{H}(X_j + \beta Y_j - \alpha) \mathcal{N}(X_j|\mu, \sigma) \mathcal{N}(Y_j|\mu_y, \sigma_y)} \label{eq:mc}.
+\end{align}
+We can now easily move from a single observation to a set of $N$ observations.
+\begin{align}
+\mathcal{L}_1 &= w_{\rm approx}^{-N} \prod_{i=1}^N \mathcal{N}(x_i|\mu, \sigma) \mathcal{N}(y_i|\mu_y, \sigma_y) \\
+\mathcal{L}_{2} &= \left(  \frac{ w_{\rm approx}  n }{\sum_{j=1}^{n} \mathcal{H}(X_j + \beta Y_j - \alpha) \mathcal{N}(X_j|\mu, \sigma) \mathcal{N}(Y_j|\mu_y, \sigma_y)}\right)^N
 \end{align}
@@
 \label{sec:tricks}
+
+\subsection{Importance Sampling}
 Further tricks can be used to increase the efficiency with which the samples are reweighted.
@@
 
+\subsection{Precomputing selection}
 If evaluating the probability that an event is observed is numerically expensive (i.e.
@@
 
+Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution, but utilise the Monte Carlo integration technique from Section \ref{sec:real} and do not have an analytic approximation (i.e.
+we set $w_{\rm approx} =1$).
+We could estimate, given some prior knowledge, that variable $x \approx \mathcal{N}(\mu_{\rm guess}, \sigma_{\rm guess})$.
+We then drawn samples of $x$ from this distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
+That is, we assign $P(S|x,\theta) = P(S|x) = 1$ or $0$ given it passed cuts or not.
+We discard all events with $0$ weight (as they have $0$ weight), and only track those events which pass.
+Then, when calculating the sample reweighting after running chains, $\mathcal{L}_2$ becomes
+\begin{align}
+\mathcal{L}_2 \propto w_{\rm approx}  \left[\sum_{i=1}^{n} \frac{\mathcal{N}(X_i | \mu, \sigma)}{\mathcal{N}(X_i | \mu_{\rm guess}, \sigma_{\rm guess})} \right]^{-1},
+\end{align} 
+where you can see that we discard the constant $n$ from equation \eqref{eq:mc} as we only care about likelihood proportionality.
+Provided our parameter estimate is reasonably well informed, the computation benefit this precomputation provides is enormous for any nontrivial selection function.
+Not only do we now waste no time when calculating $\mathcal{L}_2$ determining $P(S|{\rm data}, \theta)$, and because we only save results that pass the cuts, we have no wasted evaluations of $\mathcal{N}(X | \mu, \sigma)$.
+
+The astute reader may have picked up on one assumption - that selection efficiency of an observation is independent of model parameters $\theta$.
+For most experimental cases this should hold, however if it does not this method cannot be used to increase efficiency.
+Gridding or interpolating the parameter space is strongly not recommended due the required accuracy of $\mathcal{L}_2$.
+Even a small error when raised to the power of $N$ can spiral out of control.
+
+\subsection{Log-space}
+
+Following from the previous section, as our reweighting $\mathcal{L}_2$ is raised to the power of the number of our observations, they should definitely be computed in log-space, which turns the power into a linear factor.
+As most probabilistic work is already computed in log-space, this subsection barely needs to be stated.
+However, whilst working in log-space an efficient way of increasing the accuracy of the approximate analytic correction is to fit the correction such that the spread of the distribution $\log\mathcal{L}_2$ is minimised.
+
+\pagebreak
+
 \section{Conclusion}
@@
 
-Reword the abstract or similar
+{\red I keep trying to write this but just end up restarting the abstract almost verbatim.
+Do you have some good tips (for this write up and for the future) on proper ways to write the conclusion?
+How it should differ from the abstract and how to avoid repeating yourself too much?}
 

### 2017-05-23 papers/general_method/general.tex
Updates

@@
 \begin{abstract}
+Fainter objects are harder to see with a telescope than brighter ones.
 Lifetimes are strictly positive.
-Fainter objects are harder to see with a telescope than brighter ones.
 We cut our data to remove background noise.
@@
 Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with different fitting algorithms \citep{Gelfand1992}.
-Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
+Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques in a similar fashion to this work \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
 
@@
 Section \ref{sec:tricks} details numeric concerns and tricks to be aware of for effective implementation of Monte Carlo corrections applied to analytic approximations.
+
+\section{Theory}
+
+When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
+\begin{align}
+P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta).
+\end{align}
+Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$.
+When dealing with experiments which have selection efficiency, our likelihood necessarily includes that efficiency, for we want to describe the probability that our observations were both drawn from the underlying theoretical model \textit{and} that those observations, given they happened, were subsequently successfully observed.
+To make this extra conditional explicit, we can write our likelihood as
+with selection effect $S$:
+\begin{align}
+\mathcal{L} &= P({\rm data} | \theta, S).
+\end{align}
+As we wish to describe our selection efficiency generally as a function dependent on both our data and our model, we can reformulate to
+\begin{align}
+\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{P(S|\theta)}.
+\end{align}
+Introducing an integral over all possible data to make $P(S|\theta)$ physical,
+\begin{align}
+\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
+&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
+\end{align}
+where the integral in the denominator has the same dimensionality is the experimental data. \textbf{Equation \ref{eq:main} is the generalised likelihood of experiments with sample selection.}
 
@@
 If, having collected our observations $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface.
-Taking uniform priors on both parameters we simply wish to map the surface
+Taking uniform priors on both parameters, we simply wish to map the surface
 \begin{align}
@@
 \label{sec:imperfect}
-In a slightly imperfect world we may have to deal with something like truncated data.
+In a slightly imperfect world we may have to deal with truncated data.
 For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$.
@@
 If we do not take this truncation into account, we will recover biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
-However, we can correct for this truncation.
-If we restate our likelihood as the probability of observation given our model parameters \textit{and} our selection effects $S$, we have
-\begin{align}
-\mathcal{L} &= P({\rm data} | \theta, S)\\
-&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{P(S|\theta)}.
-\end{align}
-Introducing an integral over all possible data to make $P(S|\theta)$ physical,
-\begin{align}
-&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
-\end{align}
-where the integral in the denominator has the same dimensionality is the experimental data.
-\textbf{Equation \ref{eq:main} is the generalised likelihood of experiments with sample selection.} In our example, our data is one dimensional (as we only observe one variable, $x$) and the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
-Having successfully observed $x$, it follows that $x > \alpha$ and so $P(S|x,\theta) = 1$.
+However, we can correct for this truncation using equation \eqref{eq:main}.
+
+In our example, our data is one dimensional (as we only observe one variable, $x$) and the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
+Having successfully observed $x_i$, it follows that $x_i > \alpha$ and so $P(S|x_i,\theta) = 1$.
 To substitute in our normal model,
 \begin{align}
-\mathcal{L}_i &= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\int_{-\infty}^\infty \mathcal{H}(D - \alpha) \mathcal{N}(D|\mu, \sigma)\, dD} \\
+\mathcal{L}_i &= \frac{P(S|x_i,\theta) P(x_i|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD} \\
+&= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\int_{-\infty}^\infty \mathcal{H}(D - \alpha) \mathcal{N}(D|\mu, \sigma)\, dD} \\
 &= \frac{ \mathcal{N}(x_i|\mu, \sigma)}{\int_{\alpha}^\infty \mathcal{N}(D|\mu, \sigma)\, dD} \\
@@
 where in the last line we have evaluated the integral in the case $\mu > \alpha$.
-Note that this is for a single observation of $x$, and so for a set of independent observations we need to introduce the product found in equation \eqref{eq:prod}.
+Note that this is for a single observation, and so for a set of independent observations we need to introduce the product found in equation \eqref{eq:prod}.
 \begin{align}
-\mathcal{L} &= \prod_{i=1}^N \frac{ \mathcal{N}(x|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
+\mathcal{L} &= \prod_{i=1}^N \mathcal{L}_i =  \prod_{i=1}^N \frac{ \mathcal{N}(x|\mu, \sigma)}{\frac{1}{2} {\rm erfc}\left[ \frac{\alpha - \mu}{\sqrt{2}\sigma} \right]}, 
 \end{align}
@@
 Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
-Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is around $20\beta = 4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood is
+Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is approximately $20\beta = 4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood can be
 \begin{align}
@@
 	\caption{A systematic test of our more complicated model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.
-The likelihood $\mathcal{L}_1$ was evaluated with the fitting algorithm emcee, and reweighted using Monte Carlo integration of a hundred thousand possible events as per $\mathcal{L}_2$.
+The likelihood $\mathcal{L}_1$ was evaluated, and reweighted using Monte Carlo integration of a hundred thousand possible events as per $\mathcal{L}_2$.
 The truncated data with no correction is shown as `Biased' in dotted red, the `Approximate` only correction ($\mathcal{L}_1$) shown in dashed green, and the final reweighted chain shown in solid black as `Corrected'.}
@@
 We discard all events with $0$ weight (as they have $0$ weight), and only track those events which pass.
-Then, when calculating the sample reweighting after running chains, $\mathcal{L}_2$ becomes
+Then, when calculating the sample reweighting after running chains, $\mathcal{L}_{i2}$ becomes
 \begin{align}
-\mathcal{L}_2 \propto w_{\rm approx}  \left[\sum_{i=1}^{n} \frac{\mathcal{N}(X_i | \mu, \sigma)}{\mathcal{N}(X_i | \mu_{\rm guess}, \sigma_{\rm guess})} \right]^{-1},
+\mathcal{L}_{i2} \propto w_{\rm approx}  \left[\sum_{j=1}^{n} \frac{\mathcal{N}(X_j | \mu, \sigma)}{\mathcal{N}(X_j | \mu_{\rm guess}, \sigma_{\rm guess})} \right]^{-1},
 \end{align} 
@@
 Provided our parameter estimate is reasonably well informed, the computation benefit this precomputation provides is enormous for any nontrivial selection function.
-Not only do we now waste no time when calculating $\mathcal{L}_2$ determining $P(S|{\rm data}, \theta)$, and because we only save results that pass the cuts, we have no wasted evaluations of $\mathcal{N}(X | \mu, \sigma)$.
+Not only do we now waste no time when calculating $\mathcal{L}_{i2}$ determining $P(S|{\rm data}, \theta)$, as we only save results that pass the cuts, we have no wasted evaluations of $\mathcal{N}(X_j | \mu, \sigma)$.
 
 The astute reader may have picked up on one assumption - that selection efficiency of an observation is independent of model parameters $\theta$.
-For most experimental cases this should hold, however if it does not this method cannot be used to increase efficiency.
-Gridding or interpolating the parameter space is strongly not recommended due the required accuracy of $\mathcal{L}_2$.
+For many experimental cases this should hold, however if it does not this method cannot be used to increase efficiency.
+Gridding or interpolating the parameter space is strongly not recommended due the required accuracy of $\mathcal{L}_{i2}$.
 Even a small error when raised to the power of $N$ can spiral out of control.
@@
 
-\pagebreak
-
 \section{Conclusion}
@@
 
-{\red I keep trying to write this but just end up restarting the abstract almost verbatim.
-Do you have some good tips (for this write up and for the future) on proper ways to write the conclusion?
-How it should differ from the abstract and how to avoid repeating yourself too much?}
+Data truncation is a pervasive issue in many scientific domains.
+For simple cases of data truncation which can be encapsulated with analytic functions, it is possible to analytically correct likelihood surfaces by introducing selection efficiency into the likelihood formulation.
+When analytic corrections fail to provide an adequate description of selection effects, Monte Carlo integration can be used on top of analytic approximations to further correct the likelihood surface, provided the experiment can be effectively simulated.
 
-\section*{Acknowledgments}
+%\section*{Acknowledgments}
 
-We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.
+%We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.
 

### 2017-05-25 papers/general_method/general.tex
Updating based on Alex's feedback

@@
 \begin{abstract}
-Fainter objects are harder to see with a telescope than brighter ones.
 Lifetimes are strictly positive.
 We cut our data to remove background noise.
-Truncated or biased datasets are inescapable in modern physics.
-In this paper, we present a simple overview of a Bayesian consideration of truncation, giving a solution to both analytically tractable and intractable models.
+Astronomical observations suffer from Malmquist bias - our samples are preferentially biased toward those sources that are more likely to be observationally detected.
+These sample selections should be accounted for.
+In this paper we present a simple overview of a Bayesian consideration of sample selection, giving a solution to both analytically tractable and intractable models.
 This can be accomplished via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
@@
 This work provides a simple treatment of truncated data in a common Bayesian technique.
-Section \ref{sec:perfect} discusses the ever-elusive case of perfect data without truncation to provide a common basis for Sections \ref{sec:imperfect} and \ref{sec:real}, which respectively cover analytically correctable data truncation and analytically intractable models.
+The general theory for considering selection effects is discussed in Section \ref{sec:theory}.
+Section \ref{sec:examples} provides three examples of increasing complexity with sample selection.
 Section \ref{sec:tricks} details numeric concerns and tricks to be aware of for effective implementation of Monte Carlo corrections applied to analytic approximations.
@@
 \section{Theory}
+\label{sec:theory}
 
@@
 \begin{align}
-\mathcal{L} &= P({\rm data} | \theta, S).
+\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:like}
 \end{align}
@@
 
-\section{The perfect world: No Truncation}
+\section{Sample Selection}
+\label{sec:examples}
+
+ In this section we present sample selections of increasing complexity.
+Each sample selection is accompanied by an illustrative example inspired by Type Ia supernova cosmology, where we characterize the properties of a standard or standardizable candle.
+
+\subsection{Complete Selection}
 \label{sec:perfect}
@@
 Presumably everything is also spherical and in a vacuum.
-Let us create a mock model in this perfect world.
-Let us observe a series of independent and identically distributed events $\vec{x}$, which are drawn from a normal distribution such that
+We thus begin by considering an ideal situation where the sample is complete.
+All events are included in the sample such that per object, $P(S | {\rm data}, \theta)=1$.
+Trivially this expression is independent of $\theta$, and our likelihood from equation \eqref{eq:main} reduces down to equation \eqref{eq:like}.
+As a concrete example of this case, let us consider a model for a population of objects whose brightnesses form a normal distribution with average $\mu$ and standard deviation $\sigma$.
+Let us also assume that our experiment produces data $x$, measurements of the brightnesses with negligible measurement uncertainty, which we can formally state as
 \begin{align}
@@
 
-\section{The Imperfect World: Analytic Truncation}
+\subsection{Analytic Truncation}
 \label{sec:imperfect}
-In a slightly imperfect world we may have to deal with truncated data.
-For an example, consider the previous model, but with an instrumentation deficiency such that we can only observe events above a certain threshold, such that $x > \alpha$.
+
+We now consider a case with selection bias and has an analytic expression for the likelihood.
+This illustrative case is useful for the reader as the influence of selection bias is simple and intuitive.
+The case we consider is identical to the one in the previous section, \textit{except} that only the subset of objects brighter than a threshold $\alpha$ can be observed.
+
+With this sample selection, all events satisfy $x > \alpha$, giving $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
 We assign a value $\alpha=85$ for convenience.
-If we do not take this truncation into account, we will recover biased parameter estimates, as shown in Figure \ref{fig:imperfect}.
-However, we can correct for this truncation using equation \eqref{eq:main}.
-
-In our example, our data is one dimensional (as we only observe one variable, $x$) and the selection efficiency is the step function $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
+If we do not take this truncation into account, we will recover biased parameter estimates.
+However, we can correct for this truncation using equation \eqref{eq:main}, as the integral in the denominator has an analytic solution.
 Having successfully observed $x_i$, it follows that $x_i > \alpha$ and so $P(S|x_i,\theta) = 1$.
@@
 \end{align}
-We can add this correction to our model, and note that we now recover unbiased parameter estimates, also shown in Figure \ref{fig:imperfect}.
+We can add this correction to our model, and note that we now recover unbiased parameter estimates.
+This is demonstrated in Figure \ref{fig:imperfect}, which shows the posterior surfaces for when you take sample selection into account and when you do not.
 \begin{figure}
@@
 
-\section{The Real World: Analytically Intractable Truncation}
+\subsection{Analytically Intractable Truncation}
 \label{sec:real}
@@
 
-So let us modify our imperfect toy model.
+So let us modify our imperfect toy model from the previous section.
 Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
@@
 
-%\section*{Acknowledgments}
+\section*{Acknowledgments}
 
-%We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.
+We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.
 

### 2017-05-26 papers/general_method/general.tex
Truncation -> Sample selection

@@
 %\shortauthors{}
-\title[Correcting for data truncation in Bayesian analyses]{Correcting for data truncation in Bayesian analyses}
+\title[Correcting for sample selection in Bayesian analyses]{Correcting for sample selection in Bayesian analyses}
 
@@
 
-Truncated data is a problem in many areas of scientific inquiry.
-It is one of the primary difficulties when performing supernovae cosmology analysis, as our telescopes have visual limits that truncate our observed supernovae distribution from the actual underlying distribution.
+Sample selection is a problem in many areas of scientific inquiry.
+It is one of the primary difficulties when performing supernovae cosmology analysis, as our telescopes have visual limits that modify our observed supernovae distribution from the actual underlying distribution.
 This bias, termed Malmquist bias, is source of much investigation \citep{Butkevich2005}.
@@
 Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with different fitting algorithms \citep{Gelfand1992}.
-Whilst generalised resources exist that provide a comprehensive overview of truncated data and analysis techniques in a similar fashion to this work \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
+Whilst generalised resources exist that provide a comprehensive overview of sample selection and analysis techniques in a similar fashion to this work \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
 
-This work provides a simple treatment of truncated data in a common Bayesian technique.
+This work provides a simple treatment of sample selection in a common Bayesian technique.
 The general theory for considering selection effects is discussed in Section \ref{sec:theory}.
@@
 
-\subsection{Analytic Truncation}
+\subsection{Analytic Sample Selection}
 \label{sec:imperfect}
@@
 
-\subsection{Analytically Intractable Truncation}
+\subsection{Analytically Intractable Sample Selection}
 \label{sec:real}
@@
 
-Data truncation is a pervasive issue in many scientific domains.
-For simple cases of data truncation which can be encapsulated with analytic functions, it is possible to analytically correct likelihood surfaces by introducing selection efficiency into the likelihood formulation.
+Sample selection is a pervasive issue in many scientific domains.
+For simple cases of sample selection which can be encapsulated with analytic functions, it is possible to analytically correct likelihood surfaces by introducing selection efficiency into the likelihood formulation.
 When analytic corrections fail to provide an adequate description of selection effects, Monte Carlo integration can be used on top of analytic approximations to further correct the likelihood surface, provided the experiment can be effectively simulated.

### 2017-05-26 papers/general_method/general.tex
More feedback from Alex

@@
 This is demonstrated in Figure \ref{fig:imperfect}, which shows the posterior surfaces for when you take sample selection into account and when you do not.
+The selection bias preferentially selects intrinsically brighter objects and, by cutting out some of the distribution, narrows the observed distribution, and so we note that the bias correction correctly increases the weight of low $\mu$ and high $\sigma$ parametrisations, as those models would be subject to the most sample bias.
+We also note that not only does the best fit location for each parameter shift, but the shape of the posterior surface itself is significantly modified.
 \begin{figure}
@@
 
-So let us modify our imperfect toy model from the previous section.
-Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$.
+In this subsection we introduce an example that presents computational challenges in the evaluation of the posterior, and which better reflects the model complexity in supernova cosmology analysis.
+The specific challenge is that the denominator in the equation for the likelihood equation \eqref{eq:main} is dependent on $\theta$ and does not have an analytic solution (i.e.
+has to be solved numerically).
+When using Monte Carlo sampling methods to probe the posterior with this kind of likelihood, the integral in the denominator must be calculated with each draw.
+We therefore present importance sampling as an approach that make these posteriors more computationally tractable.
+
+Let us modify our imperfect toy model from the previous section.
+Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$, and has no measurement uncertainty (like $x$).
 Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$, giving $P(S|x,y,\theta) = \mathcal{H}(x + \beta y - \alpha)$.
@@
 Even by fixing the samples used in MC integration (thereby giving a continuous surface), the complexity of the surface derivatives will pose almost insurmountable problems for any algorithms that utilise surface gradients.
-One solution is to find an approximate, analytic correction we can utilise in our fitting algorithm which seeks to shift the region of parameter space sampled by the sampler closer to the correct area.
+
+One solution is to find an approximate, analytic correction we can utilise in our fitting algorithm which seeks to shift the region of parameter space sampled by the sampler closer to the correct area, and then importance sample our MC chains to provide a fully corrected surface.
+Given an approximate analytic correction $w_{\rm approx}$, we explicitly break our likelihood into two parts, $\mathcal{L}_i = \mathcal{L}_{i1} \mathcal{L}_{i2}$, with the parts given by
+\begin{align}
+\mathcal{L}_{i1} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{w_{\rm approx}} \\
+\mathcal{L}_{i2} &= \frac{w_{\rm approx}}{\int P(S, D|\theta)\, dD}
+\end{align}
+Where $\mathcal{L}_{i1}$ represents an analytic function which can efficiently be sampled using traditional Monte Carlo sampling methods, and $\mathcal{L}_{i2}$ is a numerically calculated weight applied after Monte Carlo sampling in order to model the full likelihood surface.
 
@@
 Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
-Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is approximately $20\beta = 4$ (which is close to the correct value of $6$), and from this our analytic correction to our likelihood can be
+Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is approximately $20\beta = 4$ (which is close to the correct value of $6$), and from this we add in a small adjustment to the analytic correction from Section \ref{sec:imperfect}:
 \begin{align}
@@
 \end{align}
-Further, let us explicitly break our likelihood into two parts, $\mathcal{L}_i = \mathcal{L}_{i1} \mathcal{L}_{i2}$, with the parts given by
+This gives our likelihood parts as
 \begin{align}
@@
 \end{align}
-$\mathcal{L}_1$ can thus be fitted with a traditional sampler without numeric difficulty or slowdown, and $\mathcal{L}_2$ allows us to calculate the weight of each sample.
+As stated previously, $\mathcal{L}_1$ can thus be fitted with a traditional sampler without numeric difficulty or slowdown, and $\mathcal{L}_2$ allows us to calculate the weight of each sample.
 We are effectively importance sampling our likelihood evaluations.
@@
 \end{align}
-
+Thus we end up with a corrected posterior surface as shown in Figure \ref{fig:real}.
+It is important to note that the point of maximum likelihood is biased by roughly the same amount in the biased and approximate posterior surfaces in Figure \ref{fig:real}.
+However, as the approximately correctly posterior is broader than the biased surface, it has far more samples in the region of parameter space mapped by the fully corrected posterior.
+This is the entire purpose of the approximate correction - to maximise the number of samples in the correct region of parameter space, so that we can importance sample our chains efficiently.
 \begin{figure}
@@
 \end{figure}
-Thus we end up with a corrected posterior surface as shown in Figure \ref{fig:real}.
 
@@
 In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (ie traditional importance sampling) we need only evaluate the step function for our data points.
+That is, we replace
+\begin{align}
+\int f(X) \mathcal{N}(X|\mu,\sigma)\, dX = \frac{1}{N} \sum_{i=1}^N f(X_i) \mathcal{N}(X_i | \mu, \sigma),
+\end{align}
+where $X$ is drawn from a uniform distribution over all space, with 
+\begin{align}
+\int f(X) \mathcal{N}(X|\mu,\sigma)\, dX = \frac{1}{N} \sum_{i=1}^N f(X_i),
+\end{align}
+where $X$ drawn from $\mathcal{N}(X_i | \mu, \sigma)$.
 

### 2017-05-26 papers/general_method/general.tex
Elaborating in the techniques page

@@
 We cut our data to remove background noise.
-Astronomical observations suffer from Malmquist bias - our samples are preferentially biased toward those sources that are more likely to be observationally detected.
-These sample selections should be accounted for.
+Astronomical observations suffer from Malmquist bias - brighter objects are easier to see than faint ones.
+These sample selection effects should be accounted for.
 In this paper we present a simple overview of a Bayesian consideration of sample selection, giving a solution to both analytically tractable and intractable models.
@@
 Truncated data is also commonly encountered in biological fields, where data such as mortality rates are left-truncated \citep{JANE1898}.
-Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with different fitting algorithms \citep{Gelfand1992}.
+Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with application to specific fitting algorithms \citep{Gelfand1992}.
 Whilst generalised resources exist that provide a comprehensive overview of sample selection and analysis techniques in a similar fashion to this work \citep{klein2005survival}, these sources are often opaque due to volume and mathematical complexity.
@@
 Uncertainties are well quantified and normally distributed around true values.
-Presumably everything is also spherical and in a vacuum.
+Presumably everything is also spherical and in a vacuum
+
 We thus begin by considering an ideal situation where the sample is complete.
-All events are included in the sample such that per object, $P(S | {\rm data}, \theta)=1$.
-Trivially this expression is independent of $\theta$, and our likelihood from equation \eqref{eq:main} reduces down to equation \eqref{eq:like}.
+All events are included in the sample such that $P(S | {\rm data}, \theta)=1$.
+Trivially this expression is independent of $\theta$, and our likelihood from Equation \eqref{eq:main} reduces down to Equation \eqref{eq:like}.
 As a concrete example of this case, let us consider a model for a population of objects whose brightnesses form a normal distribution with average $\mu$ and standard deviation $\sigma$.
-Let us also assume that our experiment produces data $x$, measurements of the brightnesses with negligible measurement uncertainty, which we can formally state as
+Let us also assume that our experiment produces data $x$, which are independent measurements of unique objects' brightnesses (with negligible measurement uncertainty), which we can formally state as
 \begin{align}
@@
 If we do not take this truncation into account, we will recover biased parameter estimates.
-However, we can correct for this truncation using equation \eqref{eq:main}, as the integral in the denominator has an analytic solution.
+However, we can correct for this truncation using Equation \eqref{eq:main}, as the integral in the denominator has an analytic solution.
 Having successfully observed $x_i$, it follows that $x_i > \alpha$ and so $P(S|x_i,\theta) = 1$.
@@
 where in the last line we have evaluated the integral in the case $\mu > \alpha$.
-Note that this is for a single observation, and so for a set of independent observations we need to introduce the product found in equation \eqref{eq:prod}.
+Note that this is for a single observation, and so for a set of independent observations we need to introduce the product found in Equation \eqref{eq:prod}.
 \begin{align}
@@
 In this subsection we introduce an example that presents computational challenges in the evaluation of the posterior, and which better reflects the model complexity in supernova cosmology analysis.
-The specific challenge is that the denominator in the equation for the likelihood equation \eqref{eq:main} is dependent on $\theta$ and does not have an analytic solution (i.e.
+The specific challenge is that the denominator in Equation \eqref{eq:main} is dependent on $\theta$ and does not have an analytic solution (i.e.
 has to be solved numerically).
@@
 Instead of observing just one variable, $x$, we also observe a new independent variable, $y$, which is drawn from its own distribution $y \sim \mathcal{N}(\mu_y, \sigma_y)$, and has no measurement uncertainty (like $x$).
-Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$, giving $P(S|x,y,\theta) = \mathcal{H}(x + \beta y - \alpha)$.
+Our selection efficiency can now become a combination of $x$ and $y$, such that we only observe events that satisfy $x + \beta y > \alpha$, giving $P(S|x,y,\theta) = \mathcal{H}(x + \beta y - \alpha)$, giving a relationship reminiscent of the stretch and colour corrections for standardising supernovae.
 Our likelihood for such a toy model becomes now the combination of probabilities for observing both $x$ and $y$, with the denominator becoming an integral over all possible $X$ and $Y$ observations subject to our selection effects.
@@
 
-\section{Numerical Tricks}
+\section{Numerical Techniques}
 \label{sec:tricks}
@@
 Firstly, the overarching analytic model often provides functions which can be drawn from efficiently.
-In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (ie traditional importance sampling) we need only evaluate the step function for our data points.
+In the case of our example, by drawing the random numbers $X$ and $Y$ respectively from the normal distributions $\mathcal{N}(\mu,\sigma)$ and $\mathcal{N}(\mu_y,\sigma_y)$ (i.e.
+traditional importance sampling) we need only evaluate the step function for our data points.
 That is, we replace
@@
 \subsection{Precomputing selection}
-If evaluating the probability that an event is observed is numerically expensive (i.e.
-not a step function), it is easy to pregenerate a set of events and reuse them for all weights - provided that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
-This method is however only efficient when prior knowledge of parameter values is known to allow a reasonable initial draw of events.
-Without this prior information, samples need to span the entire posterior volume, which is numerically intractable even for low dimensional models.
 
-Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution, but utilise the Monte Carlo integration technique from Section \ref{sec:real} and do not have an analytic approximation (i.e.
+The integral in the denominator of Equation \eqref{eq:main} must be recalculated for each set of $\theta$ parameter-values that are considered.
+A significant cost savings is possible when Monte Carlo integration is done, if the same Monte Carlo realizations are used for all sets of $\theta$.
+This is especially true if evaluating the sample selection $P(S|{\rm data}, \theta)$ is numerically expensive, but can only be used when the sample selection is independent of $\theta$, so that $P(S|{\rm data}, \theta) = P(S|{\rm data})$, and when there is some prior knowledge of $\theta$ such that we can estimate a reasonable $\theta_{\rm approx}$.
+Considering the integral from Equation \eqref{eq:main}, we have
+\begin{align}
+&\int P(S|D, \theta) P(D|\theta) \, dD = \notag \\
+&\quad\quad\quad\int P(S|D) \frac{P(D|\theta)}{P(D|\theta_{\rm approx})} \left[ P(D|\theta_{\rm approx}) \,dD \right]
+\end{align}
+With the mathematics laid as so, it is easy to pregenerate a set of events $D_i$ drawn from distribution $ P(D|\theta_{\rm approx})$, evaluate $P(S|D_i)$, and reuse them for all samples.
+Formulated using Monte Carlo integration,
+\begin{align}
+&\int P(S|D, \theta) P(D|\theta) \, dD = \frac{1}{N} \sum_{i=1}^N \frac{P(S|D_i)}{P(D_i|\theta_{\rm approx})} P(D_i|\theta),
+\end{align}
+the only term that needs to be evaluated for each sample is $P(D_i|\theta)$, as all other terms have been pre-computed and stored.
+When using this technique, take care that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
+
+Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution as was done in Section \ref{sec:imperfect}, but utilise the Monte Carlo integration technique from Section \ref{sec:real}.
+We work without an analytic approximation (i.e.
 we set $w_{\rm approx} =1$).
-We could estimate, given some prior knowledge, that variable $x \approx \mathcal{N}(\mu_{\rm guess}, \sigma_{\rm guess})$.
+We could estimate, given some prior knowledge, that variable $x \approx \mathcal{N}(\mu_{\rm approx}, \sigma_{\rm approx})$.
+In this example, it would be better to overestimate $\sigma_{\rm approx}$ rather than underestimate it, so that we still effectively sample the correct area of parameter space.
 We then drawn samples of $x$ from this distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
@@
 \begin{align}
-\mathcal{L}_{i2} \propto w_{\rm approx}  \left[\sum_{j=1}^{n} \frac{\mathcal{N}(X_j | \mu, \sigma)}{\mathcal{N}(X_j | \mu_{\rm guess}, \sigma_{\rm guess})} \right]^{-1},
+\mathcal{L}_{i2} \propto w_{\rm approx}  \left[\sum_{j=1}^{n} \frac{\mathcal{N}(X_j | \mu, \sigma)}{\mathcal{N}(X_j | \mu_{\rm approx}, \sigma_{\rm approx})} \right]^{-1},
 \end{align} 
-where you can see that we discard the constant $n$ from equation \eqref{eq:mc} as we only care about likelihood proportionality.
+where you can see that we discard the constant $n$ from Equation \eqref{eq:mc} as we only care about likelihood proportionality.
 Provided our parameter estimate is reasonably well informed, the computation benefit this precomputation provides is enormous for any nontrivial selection function.
@@
 
-The astute reader may have picked up on one assumption - that selection efficiency of an observation is independent of model parameters $\theta$.
-For many experimental cases this should hold, however if it does not this method cannot be used to increase efficiency.
+As stated previously, this technique requires that selection efficiency of an observation is independent of model parameters $\theta$.
+For many experimental cases this may hold, however if it does not this method cannot be used to increase efficiency.
 Gridding or interpolating the parameter space is strongly not recommended due the required accuracy of $\mathcal{L}_{i2}$.

### 2017-06-02 papers/general_method/general.tex
Paper updates from Tam's feedback

@@
 \begin{abstract}
-Lifetimes are strictly positive.
-We cut our data to remove background noise.
-Astronomical observations suffer from Malmquist bias - brighter objects are easier to see than faint ones.
-These sample selection effects should be accounted for.
-In this paper we present a simple overview of a Bayesian consideration of sample selection, giving a solution to both analytically tractable and intractable models.
-This can be accomplished via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
-Toy models are included, along with numerical considerations and optimisations for implementation.
+Lifetimes are strictly positive; data are often cut to remove background noise; astronomical observations preferentially discover bright objects.
+These sample selection effects need to be accounted for in statistical analyses.
+In this paper we present a simple overview of a Bayesian consideration of sample selection, giving solutions to both analytically tractable and intractable models.
+This is accomplished via a combination of analytic approximations and Monte Carlo integration, in which dataset simulation is efficiently used to correct for issues in the observed dataset.
+Toy models are included for demonstration, along with discussions of numerical considerations and how to optimise for implementation.
 \end{abstract}
@@
 Sample selection is a problem in many areas of scientific inquiry.
-It is one of the primary difficulties when performing supernovae cosmology analysis, as our telescopes have visual limits that modify our observed supernovae distribution from the actual underlying distribution.
-This bias, termed Malmquist bias, is source of much investigation \citep{Butkevich2005}.
-It is considered during analysis by either modifying the observed data to remove the expected bias \citep{BetouleKessler2014, ConleyGuySullivan2011}, or by incorporating the expected bias into the underlying model \citep{Rubin2015}.
-Truncated data is also commonly encountered in biological fields, where data such as mortality rates are left-truncated \citep{JANE1898}.
+For example, it is one of the primary difficulties when performing supernova cosmology analysis, as telescopes more easily detect bright objects than faint ones, causing our observed supernova distribution to deviate from the actual underlying distribution.
+This  ``Malmquist bias'', has been a subject of much investigation \citep{Butkevich2005}.
+It is taken into account during analyses by either modifying the observed data to remove the expected bias \citep{BetouleKessler2014, ConleyGuySullivan2011}, or by incorporating the expected bias into the underlying model \citep{Rubin2015}.
+Truncated data are also commonly encountered in biological fields, where data such as mortality rates are left-truncated \citep{JANE1898}.
 Simplified and generalised examples have been investigated in numerous fashions \citep{woodroofe1985estimating, Gull1989bayesian, grogger1991models, o1995truncated} and with application to specific fitting algorithms \citep{Gelfand1992}.
@@
 
-This work provides a simple treatment of sample selection in a common Bayesian technique.
+This work provides a simple treatment of sample selection using a common Bayesian technique.
 The general theory for considering selection effects is discussed in Section \ref{sec:theory}.
@@
 \begin{align}
-P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta).
+P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
 \end{align}
+which gives the probability of the model parameter values ($\theta$) given the data.
+Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$.
 Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$.
-When dealing with experiments which have selection efficiency, our likelihood necessarily includes that efficiency, for we want to describe the probability that our observations were both drawn from the underlying theoretical model \textit{and} that those observations, given they happened, were subsequently successfully observed.
-To make this extra conditional explicit, we can write our likelihood as
-with selection effect $S$:
+When dealing with experiments which have imperfect selection efficiency, our likelihood must take that efficiency into account.
+We need to describe the probability that the events we observe were both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, were subsequently successfully observed.
+To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} selection efficiency, $S$, as:
 \begin{align}
@@
 \end{align}
-As we wish to describe our selection efficiency generally as a function dependent on both our data and our model, we can reformulate to
+If we seek to describe the chance of our observation passing the sample selection as a function of the observation itself and model parameters, we need to rearrange the above equation to include $P(S|{\rm data},\theta)$.
+Doing so gives us
 \begin{align}
@@
 \end{align}
-Introducing an integral over all possible data to make $P(S|\theta)$ physical,
+Introducing an integral over all possible data $D$, so we can evaluate $P(S|\theta)$, 
 \begin{align}
@@
 \end{align}
-where the integral in the denominator has the same dimensionality is the experimental data. \textbf{Equation \ref{eq:main} is the generalised likelihood of experiments with sample selection.}
+where the integral in the denominator has the same dimensionality as the experimental data. \textbf{Equation \ref{eq:main} is the generalised likelihood of experiments with sample selection.}
 
@@
 In a perfect world, data is neither biased nor truncated.
-The data is perfect.
+Data is perfect.
 Uncertainties are well quantified and normally distributed around true values.
-Presumably everything is also spherical and in a vacuum
+Presumably everything is also spherical and in a vacuum.
 
@@
 All events are included in the sample such that $P(S | {\rm data}, \theta)=1$.
-Trivially this expression is independent of $\theta$, and our likelihood from Equation \eqref{eq:main} reduces down to Equation \eqref{eq:like}.
+Trivially this expression is independent of our data and model parameters $\theta$, and our likelihood from Equation \eqref{eq:main} reduces down to Equation \eqref{eq:like}.
 As a concrete example of this case, let us consider a model for a population of objects whose brightnesses form a normal distribution with average $\mu$ and standard deviation $\sigma$.
-Let us also assume that our experiment produces data $x$, which are independent measurements of unique objects' brightnesses (with negligible measurement uncertainty), which we can formally state as
+Let us also assume that our experiment produces data $x_i$, which are independent measurements of unique objects' brightnesses (with negligible measurement uncertainty), which we can formally state as
 \begin{align}
@@
 \end{align}
-If, having collected our observations $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface.
-Taking uniform priors on both parameters, we simply wish to map the surface
+If, having collected our observations $\vec{x}$, we wanted to constrain $\mu$ and $\sigma$, this would be a simple task of modelling the posterior surface. We simply wish to map the surface
 \begin{align}
@@
 \end{align}
-With uniform priors, $P(\mu,\sigma) = {\rm constant}$, and can be absorbed into the constant of proportionality. Expanding our observation vector, the posterior surface is given by
+With uniform priors, $P(\mu,\sigma) = {\rm constant}$, and can be absorbed into the constant of proportionality.
+Expanding our observation vector to take into account all observations and not just one, the posterior surface is given by
 \begin{align}
@@
 
-We now consider a case with selection bias and has an analytic expression for the likelihood.
+We now consider a case with selection bias that allows us to form an analytic expression for the likelihood.
 This illustrative case is useful for the reader as the influence of selection bias is simple and intuitive.
@@
 
-With this sample selection, all events satisfy $x > \alpha$, giving $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
+With this sample selection, all events satisfy $x_i > \alpha$, giving $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
 We assign a value $\alpha=85$ for convenience.
@@
 \end{align}
-However, as our selection efficiency correction is observationally independent it is identical for all observations, allowing us to take it outside the product.
+However, as our selection efficiency correction is independent of our observations $\vec{x}$, we can take it outside the product.
 \begin{align}
@@
 This is demonstrated in Figure \ref{fig:imperfect}, which shows the posterior surfaces for when you take sample selection into account and when you do not.
-The selection bias preferentially selects intrinsically brighter objects and, by cutting out some of the distribution, narrows the observed distribution, and so we note that the bias correction correctly increases the weight of low $\mu$ and high $\sigma$ parametrisations, as those models would be subject to the most sample bias.
+The selection bias preferentially selects intrinsically brighter objects and, by cutting out some of the distribution, narrows the observed distribution.
+As such, we note that the bias correction correctly increases the weight of low $\mu$ and high $\sigma$ parametrisations, as those models would be subject to the most sample bias.
 We also note that not only does the best fit location for each parameter shift, but the shape of the posterior surface itself is significantly modified.
@@
 Unfortunately it is a rare scenario when dealing with nature and all her faults for us to have an analytic selection function, let alone a function encapsulated by a single parameter.
-A more realistic scenario involves a selection efficiency instead would take the form of non-analytic function of many model parameters.
-And the function would probably be stochastic too, just to throw another wrench in the works.
+A more realistic selection efficiency would take the form of non-analytic function of many model parameters.
+And the function would probably be stochastic too, just to make things more difficult.
 Provided a method of forward modelling or simulating observations, the solution is to combine an analytic approximate correction with Monte Carlo integration.
@@
 Our likelihood for such a toy model becomes now the combination of probabilities for observing both $x$ and $y$, with the denominator becoming an integral over all possible $X$ and $Y$ observations subject to our selection effects.
+\begin{align}
+\mathcal{L}_i &= \frac{P(S|x_i, y_i,\theta) P(x_i, y_i|\theta)}{\iint P(S | X, Y, \theta) P(X, Y|\theta)\, dX\,dY},
+\end{align}
+where $\theta = \lbrace \mu, \sigma, \mu_y, \sigma_y \rbrace$.
+As before, given a successful observation and a deterministic selection efficiency, $P(S|x_i, y_i,\theta) = 1$.
+We continue to use uniform priors to keep the example simple.
+Substituting in the step function selection efficiency for $P(S|X, Y,\theta)$ and the normal distributions for $P(x_i, y_i|\theta)$, we have
 \begin{align}
@@
 traditional importance sampling) we need only evaluate the step function for our data points.
-That is, we replace
+That is, for a simplified 1D example only in $X$, we replace
 \begin{align}
@@
 \begin{align}
-&\int P(S|D, \theta) P(D|\theta) \, dD = \frac{1}{N} \sum_{i=1}^N \frac{P(S|D_i)}{P(D_i|\theta_{\rm approx})} P(D_i|\theta),
+&\int P(S|D, \theta) P(D|\theta) \, dD = \frac{1}{N} \sum_{i=1}^N \frac{P(S|D_i)}{P(D_i|\theta_{\rm approx})} P(D_i|\theta).
 \end{align}
-the only term that needs to be evaluated for each sample is $P(D_i|\theta)$, as all other terms have been pre-computed and stored.
+The \textit{only} term that needs to be evaluated for each sample is $P(D_i|\theta)$, as all other terms can be pre-computed and stored.
 When using this technique, take care that the number of events used when calculating the weights is sufficient to make the statistical error of Monte Carlo integration insignificant when compared to the constraining power of your dataset.
 
-Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution as was done in Section \ref{sec:imperfect}, but utilise the Monte Carlo integration technique from Section \ref{sec:real}.
-We work without an analytic approximation (i.e.
+Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution as was done in Section \ref{sec:imperfect}, but utilise the Monte Carlo integration technique from Section \ref{sec:real}, and work without an analytic approximation (i.e.
 we set $w_{\rm approx} =1$).
@@
 In this example, it would be better to overestimate $\sigma_{\rm approx}$ rather than underestimate it, so that we still effectively sample the correct area of parameter space.
-We then drawn samples of $x$ from this distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
+We then drawn samples of $x$ from this approximate distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
 That is, we assign $P(S|x,\theta) = P(S|x) = 1$ or $0$ given it passed cuts or not.
@@
 As most probabilistic work is already computed in log-space, this subsection barely needs to be stated.
-However, whilst working in log-space an efficient way of increasing the accuracy of the approximate analytic correction is to fit the correction such that the spread of the distribution $\log\mathcal{L}_2$ is minimised.
+However, whilst working in log-space an efficient way of increasing the accuracy of the approximate analytic correction is to fit the correction such that the spread of the distribution $\log\left(\mathcal{L}_2\right)$ is minimised.
 
@@
 
-We gratefully acknowledge the input of the many researchers that were consulted during the creation of this paper.
+We gratefully acknowledge the input of the researchers that were consulted during the creation of this paper.
 

### 2017-06-02 papers/general_method/example/real.py

@@ -83,4 +83,4 @@ c.add_chain(all_samples, parameters=[r"$\mu$", r"$\sigma$", r"$\mu_y$"], name="B
 c.add_chain(all_sampels_corrected, name="Approximate")
 c.add_chain(all_sampels_corrected, weights=weights, name="Corrected")
 c.configure(flip=False, sigmas=[0,1,2], colors=["#D32F2F", "#4CAF50", "#222222"], linestyles=[":", "--", "-"], shade_alpha=0.2, shade=True)
-c.plot(filename="real2.pdf", figsize="column", truth=[100, 10], extents=[[90, 110],[6,15]], parameters=2)
+c.plot(filename="real2.pdf", figsize="column", truth=[100, 10], extents=[[90, 105],[6,15]], parameters=2)

### 2017-06-08 papers/general_method/general.tex
More feedback

@@
 $^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
-$^{3}$Lawrence Berkeley National Labs
+$^{3}$Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720,
+USA
 }
@@
 \begin{abstract}
-Lifetimes are strictly positive; data are often cut to remove background noise; astronomical observations preferentially discover bright objects.
+Samples of astronomical sources are drawn from underlying distributions.
+The samples we observe generally do not necessarily represent the full population of sources in the Universe, but are preferentially biased toward those sources that are more likely to be detected.
+Malmquist bias is a well-known example.
+Sample selection is not unique to astronomy - noisy data is pruned in almost all fields, and physical limits  (such as age only being positive) are inescapable.
 These sample selection effects need to be accounted for in statistical analyses.
@@
 \end{align}
-If we seek to describe the chance of our observation passing the sample selection as a function of the observation itself and model parameters, we need to rearrange the above equation to include $P(S|{\rm data},\theta)$.
-Doing so gives us
+If we seek to describe the chance of our observation passing the sample selection as a function of the observation itself and model parameters, we need to rearrange the above equation to include $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of selection given our observation and model}.
+This can be done in a few lines:
+\begin{align}
+P({\rm data}, S | \theta) &= P({\rm data}, S | \theta) \\
+P({\rm data} | \theta, S) P(S | \theta)  &= P(S | {\rm data}, \theta) P({\rm data} | \theta)  \\
+P({\rm data} | \theta, S) &= \frac{P(S | {\rm data}, \theta) P({\rm data} | \theta) }{P(S | \theta)},
+\end{align}
+thus allowing us to explicitly write our likelihood as
 \begin{align}
@@
 \mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
+\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
 \end{align}
@@
 \begin{align}
-P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma). \label{eq:prod}
+P(\mu,\sigma| \vec{x}) &\propto \prod_{i=1}^N \mathcal{N}(x_i | \mu, \sigma), \label{eq:prod}
 \end{align}
+where $\mathcal{N}(x_i|\mu,\sigma)$ represents a the probability of drawing sample $x_i$ from the Normal distribution with mean $\mu$ and standard deviation $\sigma$.
 Generating a hundred data points with $\mu=100,\ \sigma=10$, we can recover our input parameters easily, as shown in Figure \ref{fig:perfect}.
@@
 
-With this sample selection, all events satisfy $x_i > \alpha$, giving $P(S|x,\theta) = \mathcal{H}(x - \alpha)$.
-We assign a value $\alpha=85$ for convenience.
+With this sample selection, all events satisfy $x_i > \alpha$, giving $P(S|x,\theta) = \mathcal{H}(x - \alpha)$, where $\mathcal{H}$ is the Heaviside step function
+\begin{align}
+\mathcal{H}(y)\equiv \begin{cases}
+1 \quad {\rm if }\  y \ge 0 \\
+0 \quad {\rm otherwise.}
+\end{cases}
+\end{align} 
+We assign an arbitrary value $\alpha=85$ for convenience.
 If we do not take this truncation into account, we will recover biased parameter estimates.
@@
 Provided a method of forward modelling or simulating observations, the solution is to combine an analytic approximate correction with Monte Carlo integration.
+Whilst other integration algorithms can be used, Monte Carlo integration is an appropriate choice, as it is based on drawing random numbers from underlying distributions and is efficient for evaluating high dimensional integrals.
 
@@
 When using Monte Carlo sampling methods to probe the posterior with this kind of likelihood, the integral in the denominator must be calculated with each draw.
-We therefore present importance sampling as an approach that make these posteriors more computationally tractable.
+We therefore present importance sampling as an approach that makes these posteriors more computationally tractable.
 
@@
 \mathcal{L}_i &= \frac{ \mathcal{N}(x_i|\mu, \sigma) \mathcal{N}(y_i|\mu_y, \sigma_y)}
-{\iint_{-\infty}^\infty \mathcal{H}(X + \beta Y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY}
+{\iint_{-\infty}^\infty \mathcal{H}(X + \beta Y - \alpha) \mathcal{N}(X|\mu, \sigma) \mathcal{N}(Y|\mu_y, \sigma_y)\, dX dY} \label{eq:above}
 \end{align}
@@
 These often clash with sampling methods, especially for high dimensional integrals.
-Inserting Monte Carlo integration into fitting algorithms can drastically slow them down, and algorithms such as Hamiltonian MCMC that require continuous surfaces can easily fail on surfaces that fluctuate from Monte Carlo integration.
+Inserting Monte Carlo integration into fitting algorithms can drastically slow them down, and algorithms such as Hamiltonian MCMC that require continuous surfaces can easily fail on surfaces that fluctuate due to the stochastic nature of Monte Carlo integration.
 Even by fixing the samples used in MC integration (thereby giving a continuous surface), the complexity of the surface derivatives will pose almost insurmountable problems for any algorithms that utilise surface gradients.
 
-One solution is to find an approximate, analytic correction we can utilise in our fitting algorithm which seeks to shift the region of parameter space sampled by the sampler closer to the correct area, and then importance sample our MC chains to provide a fully corrected surface.
-Given an approximate analytic correction $w_{\rm approx}$, we explicitly break our likelihood into two parts, $\mathcal{L}_i = \mathcal{L}_{i1} \mathcal{L}_{i2}$, with the parts given by
+It is tempting to separate out the numerator and denominator of Equation \eqref{eq:above} and only sample from the numerator, adding corrections from the denominator after the fact.
+However, this strategy rarely works, as ignoring the denominator completely can lead to fits in completely wrong areas of parameter space.
+The solution to find an approximate, analytic correction we can utilise in our fitting algorithm which seeks to shift the region of parameter space sampled by the sampler closer to the correct area, and then importance sample our MC chains to provide a fully corrected surface.
+Given an approximate analytic correction $w_{\rm approx}$, we explicitly break our likelihood into two parts, $\mathcal{L}_i = \mathcal{L}_{i1} \mathcal{L}_{i2}$, with the parts given in general by
 \begin{align}
 \mathcal{L}_{i1} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{w_{\rm approx}} \\
-\mathcal{L}_{i2} &= \frac{w_{\rm approx}}{\int P(S, D|\theta)\, dD}
+\mathcal{L}_{i2} &= \frac{w_{\rm approx}}{\int P(S, D|\theta)\, dD}.
 \end{align}
-Where $\mathcal{L}_{i1}$ represents an analytic function which can efficiently be sampled using traditional Monte Carlo sampling methods, and $\mathcal{L}_{i2}$ is a numerically calculated weight applied after Monte Carlo sampling in order to model the full likelihood surface.
 
@@
 \end{align}
-As stated previously, $\mathcal{L}_1$ can thus be fitted with a traditional sampler without numeric difficulty or slowdown, and $\mathcal{L}_2$ allows us to calculate the weight of each sample.
-We are effectively importance sampling our likelihood evaluations.
-The computational benefits of this should not be understated either - each sample in our chains can be reweighted independently, providing a task that is trivially parallelisable.
-Evaluating $\mathcal{L}_2$ using Monte Carlo integration of $n$ samples, we have
+Evaluating $\mathcal{L}_{i2}$ using Monte Carlo integration of $n$ samples, we have for a single observation that
 \begin{align}
@@
 \end{align}
-We can now easily move from a single observation to a set of $N$ observations.
+We now wish to move to a set of observations. As $\mathcal{L}_{i2}$ is independent of the specific observation $i$, this is a simple step:
 \begin{align}
@@
 \end{align}
-Thus we end up with a corrected posterior surface as shown in Figure \ref{fig:real}.
+$\mathcal{L}_1$ can thus be fitted with a traditional sampler without numeric difficulty or slowdown, and $\mathcal{L}_2$ allows us to calculate the weight of each sample given by fitting $\mathcal{L}_{1}$.
+We are effectively importance sampling our likelihood evaluations.
+The computational benefits of this should not be understated either --- each sample in our chains can be reweighted independently, providing a task that is trivially parallelisable.
+
+Taking our concrete example stated previously, we end up with a corrected posterior surface as shown in Figure \ref{fig:real}.
 It is important to note that the point of maximum likelihood is biased by roughly the same amount in the biased and approximate posterior surfaces in Figure \ref{fig:real}.
 However, as the approximately correctly posterior is broader than the biased surface, it has far more samples in the region of parameter space mapped by the fully corrected posterior.
-This is the entire purpose of the approximate correction - to maximise the number of samples in the correct region of parameter space, so that we can importance sample our chains efficiently.
+This is the entire purpose of the approximate correction --- to maximise the number of samples in the correct region of parameter space, so that we can importance sample our chains efficiently.
 \begin{figure}
@@
 The likelihood $\mathcal{L}_1$ was evaluated, and reweighted using Monte Carlo integration of a hundred thousand possible events as per $\mathcal{L}_2$.
-The truncated data with no correction is shown as `Biased' in dotted red, the `Approximate` only correction ($\mathcal{L}_1$) shown in dashed green, and the final reweighted chain shown in solid black as `Corrected'.}
+The truncated data with no correction is shown as `Biased' in dotted red, the `Approximate` only correction ($\mathcal{L}_1$) shown in dashed green, and the final reweighted chain shown in solid black as `Corrected'.
+A large number of samples of $\mathcal{L}_1$ had to be generated to ensure sufficient sampling of our `Corrected' posterior.}
 	\label{fig:real}
@@
 The integral in the denominator of Equation \eqref{eq:main} must be recalculated for each set of $\theta$ parameter-values that are considered.
-A significant cost savings is possible when Monte Carlo integration is done, if the same Monte Carlo realizations are used for all sets of $\theta$.
+A significant cost saving is possible when Monte Carlo integration is done, if the same Monte Carlo realizations are used for all sets of $\theta$.
 This is especially true if evaluating the sample selection $P(S|{\rm data}, \theta)$ is numerically expensive, but can only be used when the sample selection is independent of $\theta$, so that $P(S|{\rm data}, \theta) = P(S|{\rm data})$, and when there is some prior knowledge of $\theta$ such that we can estimate a reasonable $\theta_{\rm approx}$.
@@
 
-Consider the imperfect example - where we observe $x$ drawn from an underlying normal distribution as was done in Section \ref{sec:imperfect}, but utilise the Monte Carlo integration technique from Section \ref{sec:real}, and work without an analytic approximation (i.e.
+Consider the imperfect example --- where we observe $x$ drawn from an underlying normal distribution as was done in Section \ref{sec:imperfect}, but utilise the Monte Carlo integration technique from Section \ref{sec:real}, and work without an analytic approximation (i.e.
 we set $w_{\rm approx} =1$).
@@
 In this example, it would be better to overestimate $\sigma_{\rm approx}$ rather than underestimate it, so that we still effectively sample the correct area of parameter space.
-We then drawn samples of $x$ from this approximate distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
+We then draw samples of $x$ from this approximate distribution, recording the probability of each draw and then calculating whether our potential observation of $x$ would be observed in the experiment or not.
 That is, we assign $P(S|x,\theta) = P(S|x) = 1$ or $0$ given it passed cuts or not.
@@
 
-We gratefully acknowledge the input of the researchers that were consulted during the creation of this paper.
+We gratefully acknowledge the input of the researchers that were consulted during the creation of this paper, especially the Dark Energy Survey Supernova Working Group.
+Parts of this research were conducted by the Australian Research Council Centre of Excellence for All-sky Astrophysics (CAASTRO), through project number CE110001020.
 

### 2017-06-08 papers/general_method/general.tex
Updating paper
Adding github links

@@
 Each sample selection is accompanied by an illustrative example inspired by Type Ia supernova cosmology, where we characterize the properties of a standard or standardizable candle.
+Python code implementations of all examples, including plot generation, can be found at the Github repository \verb|SampleSelection|\footnote{\url{https://github.com/samreay/SampleSelection}}.
 
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{example/perfect.pdf}
+		\includegraphics[width=\columnwidth]{fig_0_perfect.pdf}
 	\end{center}
 	\caption{A systematic test of our perfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points.
-Any systematic offset in our model would be revealed by a shift in the stacked results away from the true parameter values.}
+Any systematic offset in our model would be revealed by a shift in the stacked results away from the true parameter values.
+This plot can be generated by executing  \mbox{\texttt{0\_perfect.py}}.}
 	\label{fig:perfect}
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{example/imperfect.pdf}
+		\includegraphics[width=\columnwidth]{fig_1_imperfect.pdf}
 	\end{center}
 	\caption{A systematic test of our imperfect model, done by stacking the output chains from fitting 100 independent realisations of our 100 data points, subject to our thresholding.
-The bias shown in the red `Biased' contour can be corrected to via the techniques shown in Section \ref{sec:imperfect} to recover unbiased surfaces.}
+The bias shown in the red `Biased' contour can be corrected to via the techniques shown in Section \ref{sec:imperfect} to recover unbiased surfaces.
+This plot can be generated by executing  \mbox{\texttt{1\_imperfect.py}}.}
 	\label{fig:imperfect}
@@
 In our example, if $\beta \ll 1$, such that the majority of selection effect is encapsulated by $x$ and not $y$, our approximate correction can take the form found in the previous correction from Section \ref{sec:imperfect}.
-Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, and a known $\beta = 0.2$, we can give a concrete example.
+Having true values of $\mu = 100$, $\sigma = 10$, $\mu_y = 30$, $\sigma_y = 5$, $\alpha=92$ and a known $\beta = 0.2$, we can give a concrete example.
 Assuming some prior, imperfect knowledge of $\mu_y$ (perhaps we believe it is approximately $20$) we estimate that the average contribution from $\beta y$ is approximately $20\beta = 4$ (which is close to the correct value of $6$), and from this we add in a small adjustment to the analytic correction from Section \ref{sec:imperfect}:
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{example/real.pdf}
+		\includegraphics[width=\columnwidth]{fig_2_real.pdf}
 	\end{center}
@@
 The truncated data with no correction is shown as `Biased' in dotted red, the `Approximate` only correction ($\mathcal{L}_1$) shown in dashed green, and the final reweighted chain shown in solid black as `Corrected'.
-A large number of samples of $\mathcal{L}_1$ had to be generated to ensure sufficient sampling of our `Corrected' posterior.}
+A large number of samples of $\mathcal{L}_1$ had to be generated to ensure sufficient sampling of our `Corrected' posterior.
+This plot can be generated by executing \mbox{\texttt{2\_real.py}}.}
 	\label{fig:real}

### 2017-09-18 dessn/configurations/approximate_simple_test.py
Turning everything back on but fixing w

@@ -39,7 +39,7 @@ if __name__ == "__main__":
         c.add_chain(chain, weights=weight, posterior=posterior, name="Approx")
         c.configure(spacing=1.0)
 
-        parameters = [r"$\Omega_m$", r"$\alpha$", r"$\beta$", r"$\langle M_B \rangle$",
+        parameters = [r"$\Omega_m$", r"$w$", r"$\alpha$", r"$\beta$", r"$\langle M_B \rangle$",
                       r"$\delta(0)$", r"$\delta(\infty)/\delta(0)$"]
         print(c.analysis.get_latex_table(transpose=True))
         c.plotter.plot(filename=plot_filename, truth=truth, parameters=parameters)

### 2017-09-18 papers/methods/paper.tex
How should I structure the paper?

--- 2016-08-11 dessn/investigations/gaussianity/paper/gaussianity.tex
@@
-% mnras_template.tex
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
-
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % out the following line
-%\usepackage{newtxtext,newtxmath}
+\usepackage{newtxtext,newtxmath}
 % Depending on your LaTeX fonts installation, you might get better results with one of these:
@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\blue}{\color{blue}}
-\newcommand{\red}{\color{red}}
+\newcommand{\name}{SN-BHM}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Keep the title short and informative.
-\title[Supernova Gaussianity]{Investigating the Gaussianity of Supernova SALT2 Summary Statistics}
+\title[\name]{\name: A hierarchical Bayesian model for Supernova Cosmology}
 
@@
 \author[S. R. Hinton et al.]{
-Samuel R. Hinton,$^{1}$\thanks{E-mail: \href{samuelreay@gmail.com}}
+	Samuel R. Hinton,$^{1,2}$\thanks{E-mail: samuelreay@gmail.com}
 	Alex G. Kim,$^{3}$
-Tamara M. Davis,$^{1,2}$
-Edward Macaulay$^{1,2}$
+	Tamara M. Davis$^{1,2}$
 	\\
@@
 % Enter the current year, for the copyright statements etc.
-\pubyear{2016}
+\pubyear{2017}
 
@@
 \begin{abstract}
-Write an abstract.
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
 \end{abstract}
@@
 % Don't make up new ones.
-%\begin{keywords}
-%keyword1 -- keyword2 -- keyword3
-%\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{keywords}
+keyword1 -- keyword2 -- keyword3
+\end{keywords}
 
@@
 
-In the years since the discovery of the acceleration of our universe \citep{RiessFilippenko1998, SchmidtSuntzeff1998, PerlmutterAldering1999}, supernova cosmology continued to grow as an important cosmological probe.
-Recent supernova studies offer far greater statistical power than ever before, systematics are better understood, and the very phenomenon of supernova are described with improved models \citep{GuyAstier2007, ConleyGuy2011, BetouleKessler2014, RubinAldering2015}, and with increased statistical power and model fidelity, systematics are now the limiting constraint in supernova studies \citep{ConleyGuy2011, SuzukiRubin2012, ScolnicRest2014} {\red UPDATE citations, used from Betoule}.
-In this paper we investigate a potential source of systematics unaddressed in prior studies -- the gaussianity of supernova summary statistics.
-We perform this investigation within the context of the Dark Energy Survey, which is introduced in Section~\ref{sec:des}.
-In Section~\ref{sec:motivation} we summarise recent supernova models and analyses, and in Section~\ref{sec:methodology} and Section~\ref{sec:cosmology} we investigate the effect of the assumption of gaussianity on both the supernova distributions obtained and the follow on impact on cosmology.
+Introduce significance of SN historically, and the quest to remain competitive with BAO and other probes, we have to reduce systematic uncert whilst larger surveys reduce stat.
 
-\section{The Dark Energy Survey}
-\label{sec:des}
+Review old methods
 
-The Dark Energy Survey \citep[DES,][]{des2005, des2016} is a 5000 sq.
-deg.
-photometric redshift survey.
-The survey will image approximately 300 million galaxies in 5 broadband filters (\textit{grizY}) using the Dark Energy Camera \citep[DECam,][]{FlaugherDiehl2015}.
-Forty sq.
-deg.
-of the sky will be repeatedly images in the \textit{griz} bands for the supernova luminosity distance probe, and is expected to yield observations of approximately 1900 Type Ia supernova up to a redshift of $z=1.2$.
-{\red No idea where to look for more details on this.
-On cadence, updated estimates, cadence, difference between shallow and deep fields.}
+Review current methods
 
-\section{Motivation}
-\label{sec:motivation}
+Potential places for improvement
 
-In order to perform a supernova analysis, an underlying Type Ia supernova model must be adopted.
-The most used model in recent years has been that of the SALT2 model \citep{GuyAstier2007, GuySullivan2010, MosherGuy2014} and in this analysis we restrict our investigations to this model.
-The SALT2 model characterises the Type Ia spectral energy distribution (SED) as a function of magnitude $x_0$, stretch $x_1$, colour $c$, phase $p$ and wavelength $\lambda$ via
-\begin{align}
-F(p,\lambda) &= x_0 \left[ M_0(p,\lambda) + x_1 M_1(p,\lambda)\right] \exp\left[c CL(\lambda)\right].
-\end{align} 
-Thus characterising an individual supernova's magnitude, colour and stretch (and associated covariance) allows recovery of the modelled SED.
-These summary statistics are used instead of the observed light curves in supernova analyses for computational reasons across many different fitting methodologies.
-In the Bayesian hierarchical analysis performed on the Unity 2.1 dataset \citep{SuzukiRubin2012} in \citet{RubinAldering2015}, each supernova is described by a latent colour and stretch, which means an analysis with $N$ supernova grows in dimensionality as $2N + x$, where $x$ represents underlying (non-latent) parameters (such as $\Omega_m$).
-If each supernova has $M$ observations, a full hierarchical model would increase in dimensionality to $(2+M)N + x$, providing vastly increased computational complexity in resolving the underlying posterior surface.
-In frequentist analyses like that of the Joint Lightcurve Analysis \citep{BetouleKessler2014}, the correlation of all light curve points via the calibration zero points instead leads to computational difficulties in that a covariance matrix of approximate size $MN$ is introduced into the likelihood calculation, which would be prohibitively expensive to invert.
-Therefore, instead of utilising light curve data directly in cosmological models, light curves are fit to the supernova data to produce summary statistics, and these are then used in cosmology analyses.
+\section{Our Method}
 
-One key assumption when doing this is that these summary statistics are in fact gaussian in nature.
-Often this is explicitly assumed in the analysis \citep{MarchTrotta2011, RubinAldering2015}, however it is often implicitly assumed \citep{SullivanGuy2011, SuzukiRubin2012, CampbellDAndrea2013, BetouleKessler2014}.
-Investigation into asymmetries in the summary statistics are generally limited to asymmetries in the parent colour and stretch distribution \citep{ScolnicRest2014}.
-{\red More citations.} In our paper we therefore do not investigate biases or asymmetries found in underlying population distributions, but asymmetries and biases in the SALT2 parameter probability surface, and the impact of assuming the Gaussianity of said surface.
+General comments about the method (BHM), Stan
 
-\section{Methodology}
-\label{sec:methodology}
+Mapping population of observables on a population of underlying SN, where the map function encodes
+cosmology. Difficulty is creating an underlying SN population that is flexible enough to not introduce bias whilst still being physically motivated.
 
-We begin our investigation by generating supernova light curves for mock supernova in both the DES deep and shallow fields.
-As prior studies have investigated the effects of parent colour and stretch distributions, we generate all canonical supernova ($x_1 = c = 0$).
-The supernova are simulated in the \textit{griz} bands with an observational cadence of 5 days, starting from a time $t = -35$, giving 7 observations before the peak to remove the biases introduced by having little data before the luminosity peak {\red CITE}.
-Weather conditions and moon phase are simulated based on observed DES seeing {\red Cite excel file?}.
-Simulated light curves undergo a simple selection cut of having at least one observation in any band above a signal to noise of $5$.
+Observables -> Transformation function (latent, mass, cosmology, systematics) -> Underlying pop (and outlier)
 
-We investigate the potential asymmetries in SALT2 summary statistics by fitting the light curves and obtaining summary statistics using different algorithms in the software package \verb|sncosmo| \citep{Barbary2014}.
-Specifically, we obtain a full posterior surface using our own MCMC fitting methodology, and compare against summary statistics generated using a different MCMC approach (the \verb|mcmc_lc| method in \verb|sncosmo|), nestled sampling (the \verb|nest_lc| method) and minuit (via the \verb|fit_lc| method).
-The MCMC based algorithm utilises the package \verb|emcee| \citep{ForemanMackey2013}, whilst the nestled sampling is based off \verb|nestle| \citep{Skilling2004, Barbary2015}, and both algorithms should map out the full posterior surface, giving summary statistics from the mean and covariance of that full surface.
-The third algorithm, invoked via \verb|fit_lc| uses the minuit algorithm \citep{James1975, minuit2015}, and as such is a maximum likelihood estimator and does not map out the full posterior surface.
-For a Gaussian surface, all three algorithms should agree, however for non-Gaussian surfaces the minuit algorithm will be expected to produce different output.
-As such, we test the SALT2 probability surfaces for skewness and fit with the four algorithms described previously.
-The summary statistics are converted into distance modulus via the Philips relation
-\begin{align}
-\mu = m_B - M + \alpha x_1 - \beta c,
-\end{align}
-where we adopt the fiducial $\alpha = 0.14$ and $\beta = 3.15$ values from \citet{BetouleKessler2014}.
-This allows us to directly see the effects of any skewness or non-Gaussianity in any parameter on the cosmologically important distance modulus.
-We can compare the distributions in $\mu$ for the full posterior surface and for summary statistics, and also verify the skewness of the full distribution of $\mu$.
-We investigate any shift in $\langle \mu \rangle$, and change in the standard deviation $\sigma_\mu$, binning by redshift, with results illustrated in Figure~\ref{fig:bias_des}.
-We find strong evidence that, with increasing redshift, the skewness of the SALT2 fit posterior increases, and causes the maximum likelihood estimator to not only produce biased output, but overestimate the uncertainty of the result.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias_dessky.png}
-    \caption{The mean bias in {\red fill out caption when updated figure}}
-    \label{fig:bias_des}
-\end{figure*}
-The likelihood surface for one of the skewed supernova fits is shown as an example in Figure \ref{fig:example}, where the effect of the skewness is readily apparent.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias_surface.png}
-	\caption{{\red Fill out caption in updated figure}}
-	\label{fig:example}
-\end{figure*}
-To confirm that the skewness is a result of redshift and not simply a result of the general trend of decreasing signal to noise, we plot the skewness as a function of both redshift and signal to noise in Figure \ref{fig:bias_skew}.
-Interestingly, doing this shows a negative skewness for low redshift, low signal-to-noise light curves, and a positive skewness for high redshift light curves, regardless of signal-to-noise.
-\begin{figure*}
-	\includegraphics[width=\textwidth]{../output/bias2.png}
-	\caption{{\red Fill out caption in updated figure}}
-	\label{fig:bias_skew}
-\end{figure*}
-Having confirmed that skewness becomes significant at higher redshift, we now seek to characterise this potential impact on cosmology.
+\section{Simple Model Validation}
 
-\section{Cosmological Impact}
-\label{sec:cosmology}
-Fit cosmology against simulated SN
+Simple
 
-Do for des shallow and deep, using survey area to produce z dist plus extra low-z sample (0.05<z<0.2). simulate from light curves, apply selection effect of ston>5to cull bad fits. have <x> shallow.
+\section{SNANA Validation}
 
-toy cosmology model.
-adopt $\alpha$ and $\beta$ values for Phillips correction from Betoule2014, assume known gaussian intrinsic scatter, have $\Omega_m$, $w$ and $M_B$ as cosmology parameters - marginalise over $M_B$.
+Multisim and bulk
 
-Show cosmology results.
+\section{Systematic Strength Test}
 
-%\begin{figure}
-%	% To include a figure from a file named example.*
-%	% Allowable file formats are eps or ps if compiling using latex
-%	% or pdf, png, jpg if compiling using pdflatex
-%	\includegraphics[width=\columnwidth]{example}
-%    \caption{This is an example figure. Captions appear below each figure.
-%	Give enough detail for the reader to understand what they're looking at,
-%	but leave detailed discussion to the main body of the text.}
-%    \label{fig:example_figure}
-%\end{figure}
-%
-%% Example table
-%\begin{table}
-%	\centering
-%	\caption{This is an example table. Captions appear above each table.
-%	Remember to define the quantities, symbols and units used.}
-%	\label{tab:example_table}
-%	\begin{tabular}{lccr} % four columns, alignment for each
-%		\hline
-%		A & B & C & D\\
-%		\hline
-%		1 & 2 & 3 & 4\\
-%		2 & 4 & 6 & 8\\
-%		3 & 5 & 7 & 9\\
-%		\hline
-%	\end{tabular}
-%\end{table}
+Checking that random systematic permutations are less significant than stat
+
+\section{Extension for Photometric SN}
+
+Better outlier pop
 
@@
 
-The last numbered section should briefly summarise what has been done, and describe
-the final conclusions which the authors draw from their work.
-
 \section*{Acknowledgements}
-
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \bibliographystyle{mnras}
-\bibliography{bibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{bib.bib}
 
@@
 
-%\appendix
-%\section{Some extra material}
-%If you want to present additional material which would interrupt the flow of the main paper, it can be placed in an Appendix which appears after the list of references.
+\appendix
+
+\section{Some extra material}
 
@@
 % End of mnras_template.tex
-
-%SNLS \citep{AstierGuy2006, SullivanGuy2011}
-%
-%ESSENCE \citep{WoodVaseyMiknaitis2007}
-%
-%Pan-STARRS \citep{TonryStubbs2012, ScolnicRest2014, RestScolnic2014}
-%
-%SDSS-II \citep{FriemanBassett2008, KesslerBecker2009, SollermanMortsell2009, LampeitlNichol2010, CampbellDAndrea2013}
-%
-%Harvard-Smithsonian Center for Astrophysics survey \citep[CfA,][]{HickenChallis2009}
-%
-%Carnegie Supernova Project \citep[CSP,][]{ContrerasHamuy2010, FolatelliPhillips2010, StritzingerPhillips2011}
-%
-%Lick Observatory Supernova Search \citep[LOSS,][]{GaneshalingamLi2013}
-%	
-%Nearby Supernova Factory \cite[SNF,][]{AlderingAdam2002}
-%
-%HST used in \citep{RiessStrolger2007, SuzukiRubin2012}
-%
-%Modern results systematic limited \citep{ConleyGuy2011, SuzukiRubin2012, ScolnicRest2014}
-%
-%Efforts overcome \citep{IvezicSmith2007, RegnaultConley2009, TonryStubbs2012}

### 2017-09-20 papers/methods/paper.tex
Tex update
Adding bib and starting paper intro

@@
 
-Introduce significance of SN historically, and the quest to remain competitive with BAO and other probes, we have to reduce systematic uncert whilst larger surveys reduce stat.
+Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
+Since that time, of the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
+Cosmological analysis of these supernovae samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}, and yet, despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
+
+In attempts to tease out the nature of dark energy, currently running and planned surveys are once again ramping up their statistical power.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric confirmation.
+The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae.
+Such increased statistical power demands a similarly increased fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
+
+As such, staggering effort is being put into developing more sophisticated supernovae analyses.
+\citet{Scolnic2016} and \citet{Kessler2017} explore sophisticated simulation corrections to traditional analyses.
+Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with only the cost of increased computational time \citep{Weyant2013, Jennings2016}.
+Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014a, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
+\rule{\linewidth}{1pt}
+
+Introduce significance of SN historically, and the quest to remain
+competitive with BAO and other probes, we have to reduce systematic 
+uncert whilst larger surveys reduce stat.
 
@@
 
+\subsection{General Description}
+
 Mapping population of observables on a population of underlying SN, where the map function encodes
@@
 
-\section{Simple Model Validation}
+\subsection{Applied to Spectroscopic Sample}
 
-Simple
+Minimal outliers
 
-\section{SNANA Validation}
+\section{Application to DES}
 
-Multisim and bulk
+\subsection{Simulating DES SN data}
 
-\section{Systematic Strength Test}
+\subsection{Results on simulated data (ie projections)}
 
-Checking that random systematic permutations are less significant than stat
+\subsubsection{Spectroscopic Sample}
 
-\section{Extension for Photometric SN}
+\subsubsection{Photometric sample}
 
-Better outlier pop
+\subsection{Comparison with bells and whistles fixed}
+
+\subsection{Model validation}
+
+appoximate\_simple\_test.py
+
+multisim
+
+bulk
+
+\section{Systematics Strength Test}
+
+systematics test
+
+\section{Interesting Implementation Details}
+
+Anything interesting.
+
+Also talk about non-analytic correction factors (and their failure - mc integration, GP, NNGP)
 
@@
 \bibliographystyle{mnras}
-\bibliography{bib.bib}
+\bibliography{bib}
 
@@
 
-\section{Some extra material}
+\section{Papers}
 

### 2017-09-20 papers/methods/paper.tex
Paper updates

@@
 Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014a, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
+
+In this paper, we lay out a new hierarchical model that extends on past work.
+In Section \ref{sec:method} we outline our methodology and apply it to simulated datasets.
+Forecasts for the impending DES three year spectroscopic supernova survey are contain in Section \ref{sec:des}.
+Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.
+
 \rule{\linewidth}{1pt}
@@
 \section{Our Method}
+\label{sec:method}
 
@@
 \section{Application to DES}
+\label{sec:des}
 
 \subsection{Simulating DES SN data}
+
+\subsection{Model validation}
+
+appoximate\_simple\_test.py
+
+multisim
+
+bulk
 
@@
 
-\subsection{Model validation}
-
-appoximate\_simple\_test.py
-
-multisim
-
-bulk
-
 \section{Systematics Strength Test}
+\label{sec:sys}
 
@@
 \section{Interesting Implementation Details}
-
+\label{sec:details}
 Anything interesting.

### 2017-09-21 papers/methods/paper.tex
Paper update

@@
 Since that time, of the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
-Cosmological analysis of these supernovae samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}, and yet, despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
+Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}, and yet, despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
 
@@
 In this paper, we lay out a new hierarchical model that extends on past work.
+Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
 In Section \ref{sec:method} we outline our methodology and apply it to simulated datasets.
@@
 
+\section{Review}
+\label{sec:review}
+
+Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements of apparent magnitude and colour.
+Instead, most techniques fit these observations of magnitude (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
+This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}.
+The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
+As such, the product at the end is a (redshift dependent) population of $m_B$, $x_1$ and $c$.
+
+The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology.
+However, given some modelled underlying population that lives in the redshift dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
+Specifically, for any given supernova our functional map may take the traditional form:
+\begin{equation}
+M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections},
+\end{equation}
+where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
+The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010,Sullivan2010,Rigault2013,Uddin2017}.
+The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.
+
 \rule{\linewidth}{1pt}
-
-Introduce significance of SN historically, and the quest to remain
-competitive with BAO and other probes, we have to reduce systematic 
-uncert whilst larger surveys reduce stat.
-
 Review old methods

### 2017-09-21 papers/methods/paper.tex
Paper update

@@
 \newcommand{\name}{SN-BHM}
+\newcommand{\myemail}{samuelreay@gmail.com}
+\newcommand\abs[1]{\left|#1\right|}
+\newcommand {\etal} {\emph{~et~al.} }
+
+\newcommand{\green}{\color{green}}
+\newcommand{\blue}{\color{blue}}
+\newcommand{\red}{\color{red}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}.
-The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
+The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
 As such, the product at the end is a (redshift dependent) population of $m_B$, $x_1$ and $c$.
@@
 
+{\red Do I need to put more here?}
+
+\
+
 \rule{\linewidth}{1pt}
-Review old methods
+Review old methods {\green done}
 
-Review current methods
+Review current methods {\red not done}
 
-Potential places for improvement
+Potential places for improvement {\red not done}
 

### 2017-09-21 papers/methods/paper.tex
Paper updates

@@
 where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
-The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010,Sullivan2010,Rigault2013,Uddin2017}.
+The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, Rigault2013, Uddin2017}.
 The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.
+
+For traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
+\begin{align}
+\mu_C &= 5 \log\left[ \frac{(1+z)\chi}{10} \right]\\
+&=\\
+\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B
+\end{align}
 

### 2017-09-22 papers/methods/paper.tex
Raising the question as to why things are not constrained!

@@
 \begin{align}
-\mu_C &= 5 \log\left[ \frac{(1+z)\chi}{10} \right]\\
-&=\\
+\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
+r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \\
 \mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B
 \end{align}
-
-{\red Do I need to put more here?}
+The minimising function is then given as
+\begin{equation}
+\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C)
+\end{equation}
+where $C^{-1}$ is an uncertainty matrix which combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
 

### 2017-09-22 dessn/configurations/approximate_simple_test_ol.py

@@ -19,14 +19,14 @@ if __name__ == "__main__":
 
     num_nodes = 4
     model = ApproximateModelOl(num_nodes=num_nodes, global_calibration=1)
-    simulation = [SimpleSimulation(700, alpha_c=0, mass=True, dscale=0.08, num_nodes=num_nodes),
-                  SimpleSimulation(300, alpha_c=0, mass=True, dscale=0.08, num_nodes=num_nodes, lowz=True)]
+    simulation = [SimpleSimulation(300, alpha_c=0, mass=True, dscale=0.08, num_nodes=num_nodes),
+                  SimpleSimulation(200, alpha_c=0, mass=True, dscale=0.08, num_nodes=num_nodes, lowz=True)]
     fitter = Fitter(dir_name)
     fitter.set_models(model)
     fitter.set_simulations(simulation)
     fitter.set_num_cosmologies(50)
     fitter.set_num_walkers(1)
-    fitter.set_max_steps(5000)
+    fitter.set_max_steps(2000)
 
     h = socket.gethostname()
     if h != "smp-hk5pn72":  # The hostname of my laptop. Only will work for me, ha!

### 2017-09-22 papers/methods/paper.tex
Turning off calibration, numerator weight, and tightening priors

@@
 where $C^{-1}$ is an uncertainty matrix which combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
+The benefit this analysis methodology provides is speed - for samples of hundreds of supernova, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
+The speed comes with two costs.
+Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity.
+Secondly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
+As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
 

### 2017-09-28 papers/methods/paper.tex
Updating paper

@@
 \newcommand {\etal} {\emph{~et~al.} }
-
+\newcommand{\cov}{\mathcal{C}^{-1}}
 \newcommand{\green}{\color{green}}
@@
 
-For traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
+\subsection{Traditional Analyses}
+
+Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
 \begin{align}
@@
 
-\
+Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data.
+Simulations following survey observational strategies and geometry and used to calculate the expected bias in distance modulus, which is then added onto the observational data.
 
-\rule{\linewidth}{1pt}
-Review old methods {\green done}
+\subsection{Approximate Bayesian Computation}
 
-Review current methods {\red not done}
+To try and escape the limitations of the traditional analysis methodology, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
+\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and Flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}.
+In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.
 
-Potential places for improvement {\red not done}
+By building the systematic uncertainties and selection effects into the simulation package, there is vastly more freedom in how to treat and model those effects.
+Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms which simply cannot be expressed in a tractable likelihood.
+This freedom comes with a cost -- computation.
+The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion.
+For ABC methods, we must instead simulate an entire supernova population in its entirety -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts.
+This is an intensive process.
+Luckily, efficient sampling algorithms that have walkers which rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits, such as used in the BAMBIS framework {\red CITE RACHEL}.
+
+One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
+\citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, however evaluation of systematic uncertainty was only performed using the Tripp metric.
+
+\subsection{Hierarchical Bayesian Models}
+
+Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies Hierarchical Bayesian Models.
+With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
+\citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model which they apply to the light curve fitting for supernova.
+\citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters provide a model which offers increased flexibility with reduced uncertainty over the traditional method.
+The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
+\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant difference in $\alpha$ and $\beta$ values from the original as well.
+Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
+The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application \citep{Roberts2017}, as zBEAMS.
+
+The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying populations, rates, redshift distributions, mass corrections and redshift evolution, each of will will be discussed further in the outline of our model below.
 
@@
 
-General comments about the method (BHM), Stan
+We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects.
+As this is closest to the UNITY method from \citet{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
 
-\subsection{General Description}
+At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function.
+The difficulty arises in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
 
+\subsection{Observed and Latent Population}
+
+Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Additionally, each supernova has an observed redshift $\hat{z}$ which we assume is precisely known as we are focused on the spectroscopically confirmed DES supernovae for this iteration of the method.
+Finally, each supernova also has a host galaxy mass associated with it $\hat{m}$.
+
+The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$.
+This can be thought of as the `true' values for the supernova, and is normally distributed around the observed values such that
+\begin{equation}
+P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov)
+\end{equation}
+
+\subsection{Underlying Population}
+
+\subsection{Cosmological Map}
+
+\vspace{20mm}
 Mapping population of observables on a population of underlying SN, where the map function encodes

### 2017-09-29 papers/methods/paper.tex
Updating paper

@@
 
+\newcommand{\kmsmpc}{{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
+
+\defcitealias{Guy2010}{G10}
+\defcitealias{Chotard2011}{C11}
+\defcitealias{Rubin2015}{R15}	 
+\newcommand{\gten}{\citetalias{Guy2010}}
+\newcommand{\celeven}{\citetalias{Chotard2011}}
+\newcommand{\rubin}{\citetalias{Rubin2015}}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with only the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014a, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
+Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
 
@@
 where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
-The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, Rigault2013, Uddin2017}.
+The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
 The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.
@@
 Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application \citep{Roberts2017}, as zBEAMS.
+The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017}.
 
@@
 We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects.
-As this is closest to the UNITY method from \citet{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
+As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
 
@@
 Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
-Additionally, each supernova has an observed redshift $\hat{z}$ which we assume is precisely known as we are focused on the spectroscopically confirmed DES supernovae for this iteration of the method.
-Finally, each supernova also has a host galaxy mass associated with it $\hat{m}$.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$.
+Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.
 
@@
 \begin{equation}
-P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov)
+P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov).
 \end{equation}
+As we are focused on the spectroscopically confirmed DES supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$, and the similarly that the observed mass is the true mass $m$.
+Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.
 
@@
 
-\subsection{Cosmological Map}
+The underlying supernova population is often treated with two components - a population distribution in colour and stretch, and intrinsic dispersion.
+Analytic models often treat the colour and stretch populations with a skew normal and normal, respectively, and have them as independent.
+Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution.
+Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the population possesses similar ability to change as a function of redshift.
+
+We model the underlying population and intrinsic dispersion together as a redshift dependent multivariate skew normal for each survey.
+Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes.
+Both the colour and stretch means are modelled with normal priors.
+The correlation between the magnitude, stretch and colour populations is represented in a correlation matrix $\rho$, which is treated with an LKJ prior.
+Skewness is fixed to 0 for the absolute magnitude population, but left free for colour and stretch, with normal priors centered at 0 to draw the skewness to zero if not well constrained by the data.
+The decision to include skewness on the stretch distribution as well as the colour was motivated by \citet{Scolnic2016}, which found highly asymmetric underlying populations for both colour and stretch, regardless of survey or scatter model adopted.
+
+The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.
+
+As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift dependent means, width and correlations fit individual for each survey.
+
+On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} (and therefore \citet{Kunz2007}) as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples which will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population.
+Proof of concept simulation fits show that an acceptable parameterisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+For the purposes of the DES spectroscopic sample which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
+We assume that supernova fall into either population as determined by their observed classification probability $\hat{p}$.
+
+\subsection{Population Map}
+
+\subsubsection{Cosmology}
+
+We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM and standard $\Lambda$CDM.
+$\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$.
+For calculating the distance modulus, we fix $H_0 = 70 \kmsmpc $.
+
+\subsubsection{Standardisation Parameters}
+
+With increasingly large datasets and more nuanced analyses, the choice of how to handle $\alpha$ and $\beta$ becomes an important consideration when constructing a model.
+{\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0).
+\citet{Shariff2016} instead of employing a colour-dependent $\beta$, model $\beta$ as redshift dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
+
+We tested two models against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$.
+See Section \ref{sec:simdes} for details on simulation generation.
+We found for both models that non-zero values for $\beta$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
+This degeneracy resulted in a significant bias in cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+
+\subsubsection{Host Galaxy Environment}
+
+It is now well known that host galaxy environment has a significant effect on supernova properties.
+The latest sample of over 1300 spectroscopically confirmed Type Ia supernova show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
+The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses.
+The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
+In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
+\begin{equation}
+\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right]
+\end{equation}
+We also take flat priors on the parameterisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
+
+\subsubsection{Selection Effects}
+
+\subsubsection{Systematics}
 
@@
 \subsection{Simulating DES SN data}
+\label{sec:simdes}
+{\green BUT THE BELOW IN THE TESTING AREA}
+
+Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require more a more sophisticated approach.
+In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
+The first is the model from \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
 

### 2017-09-29 papers/methods/paper.tex
Systematics

@@
 See Section \ref{sec:simdes} for details on simulation generation.
-We found for both models that non-zero values for $\beta$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
+We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
 This degeneracy resulted in a significant bias in cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
@@
 
+\subsubsection{Systematics}
+
+The chief difficulty with including systematics in supernova analyses is that the generally occur during the observational pipeline, and have difficult to model effects on the output observations.
+As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial Z_i}$, $\frac{\partial \hat{x_1}}{\partial Z_i}$, $\frac{\partial \hat{c}}{\partial Z_i}$, where $Z_i$ represents the $i$th systematic.
+
+Assuming that the gradients can be linearly extrapolated -- which is a reasonable approximation for modern surveys with high quality control of systematics -- we can incorporate into our model a deviation from the observed original values by constructing a $(3 \times N_{\rm sys})$ matrix containing the numerical derivatives for the $N_{\rm sys}$ systematics and multiplying it with the row vector containing the offset for each systematic.
+But scaling the gradient matrix to represent the shift over $1\sigma$ of systematic uncertainty, we can simply enforce a unit normal prior on the systematic row vector to increase computational efficiency.
+
+This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses, however it is normally constrained to band systematics.
+That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty.
+We include these in our approach, in addition to including HST Calspec calibration uncertainty, 10 SALT2 model systematic uncertainties, three dust systematics (scale, dust model, and transfer law) and also the systematic peculiar velocity uncertainty.
+This gives fourteen global systematics shared by all surveys, plus two systematics per band in each survey.
+
 \subsubsection{Selection Effects}
 
-\subsubsection{Systematics}
+Error function and skew normal
 

### 2017-10-03 papers/methods/paper.tex
Selection effects in paper

@@
 Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
-In Section \ref{sec:method} we outline our methodology and apply it to simulated datasets.
-Forecasts for the impending DES three year spectroscopic supernova survey are contain in Section \ref{sec:des}.
+In Section \ref{sec:method} we outline our methodology.
+Model verification on simulated datasets is given in Section \ref{sec:verification}.
+Forecasts for the impending DES three year spectroscopic supernova survey are contain in Section \ref{sec:forecast}.
 Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.
@@
 The chief difficulty with including systematics in supernova analyses is that the generally occur during the observational pipeline, and have difficult to model effects on the output observations.
-As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial Z_i}$, $\frac{\partial \hat{x_1}}{\partial Z_i}$, $\frac{\partial \hat{c}}{\partial Z_i}$, where $Z_i$ represents the $i$th systematic.
+As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial Z_i}$, $\frac{\partial \hat{x_1}}{\partial Z_i}$, $\frac{\partial \hat{c}}{\partial Z_i}$, where $Z_i$ represents the $i$\textsuperscript{th} systematic.
 
@@
 
-Error function and skew normal
+Our treatment of selection effects is to incorporate selection efficiency into our model, rather than relying on simulation-driven data corrections.
+As such, we need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
+To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
+\begin{align}
+\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:like}
+\end{align}
+As the model so far described in previous sections describes $P(D|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
+\begin{align}
+\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
+\end{align}
+where the denominator represents an integral over all potential data.
+Full derivation of this can be found in Appendix \ref{app:selection}.
+As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
+Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC.
+We therefore simplify the integral and approximate the selection effects in apparent magnitude and redshift space independently, such that the denominator, denoted now $w$ for simplicity, is given as
+\begin{align}
+w &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
+\end{align}
+We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sample the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
+The population $P(m_B | z, \theta)$ thus becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
+\begin{align}
+m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
+\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2 \notag\\ 
+&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c})
+\end{align}
+What then remains is determining the functional form of $P(S|m_B)$.
+For the treatment of the DES, we find that the error function which smoothly transitions from some constant efficiency down to zero is sufficient.
+This similar result has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014} {\red THESE CITATIONS EXACTLY MIRROR RUBIN2015, NOT SURE IF I SHOULD CHANGE THEM}.
+For our treatment of the low redshift surveys included {\red Introduce which lowz surveys earlier on} we simplify the model fitting and group all low redshift surveys together.
+This requires a more complicated selection function than the error function used for DES, however is reasonably well fit with a skew normal.
+The selection functions were fit to apparent magnitude efficiency ratios calculated from SNANA simulations.
+The DES survey is thus characterised by $\mu_{\rm DES} = 22.3$, $\sigma_{\rm DES} = 0.7$, with the combined low redshift surveys characterised by $\mu_{\rm LowZ} = 14.1$, $\sigma_{\rm LowZ} = 1.36$, $\alpha_{\rm LowZ} = 3.8$.
 
-\vspace{20mm}
-Mapping population of observables on a population of underlying SN, where the map function encodes
-cosmology. Difficulty is creating an underlying SN population that is flexible enough to not introduce bias whilst still being physically motivated.
+With the well sampled approximation as specified previously, we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
+For the error function (DES) and skew normal selection (LowZ) functions respectively, this correction becomes
+\begin{align}
+w_{\rm DES} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm DES}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm DES}^2}}  \right) \\
+w_{\rm LowZ} &= \mathcal{N}\left( \frac{m_B^* - \mu_{\rm LowZ}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm LowZ}^2}}\right) \times \notag\\
+&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm LowZ})(m_B^* - \mu_{\rm LowZ})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}{\sigma^2_{\rm LowZ}} \sqrt{\frac{\sigma^2_{\rm LowZ}}{\alpha^2_{\rm LowZ}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm LowZ}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}} }\right),
+\end{align}
+where $\Phi$ represents the normal CDF function, and $\Phi^c$ the complimentary CDF.
 
-Observables -> Transformation function (latent, mass, cosmology, systematics) -> Underlying pop (and outlier)
+\section{Model Verification}
+\label{sec:verification}
 
-\subsection{Applied to Spectroscopic Sample}
-
-Minimal outliers
-
-\section{Application to DES}
-\label{sec:des}
+\subsection{Applied to Toy Spectroscopic Data}
+\label{sec:toy}
 
@@
 
-\subsection{Model validation}
+\subsection{DES SN data validation}
+\label{sec:des}
 
@@
 
-\subsection{Results on simulated data (ie projections)}
+\section{Forecasts}
+\label{sec:forecasts}
 
@@
 
-\section{Papers}
+\section{Selection Effect Derivation}
+\label{app:selection}
 

### 2017-10-03 papers/methods/paper.tex
More fit results

@@
 \newcommand{\rubin}{\citetalias{Rubin2015}}
+
+\graphicspath{ {./figures/} }
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+In order to verify our model we run it through several tests.
+First, we validate on toy models, verifying that there is not significant cosmological bias in data sets more constraining than the DES spectroscopic sample.
+We then also generate
+
 \subsection{Applied to Toy Spectroscopic Data}
@@
 
-\subsection{Simulating DES SN data}
+We generate simple toy data to validate the basic premise of the model.
+For both DES and LowZ data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
+Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
+Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
+The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae.
+We draw from each survey simulation until we have 300 LowZ supernovae and 700 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
+
+The above represents one realisation of data.
+We test three models (Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM) with 300 realisations and find recovery of underlying cosmology without significant bias.
+Combined posterior surfaces of 22 realisations are shown in Figures \ref{fig:simple_om}, \ref{fig:simple_w}, \ref{fig:simple_ol}, with the statistical distribution of realisations for the Flat $\Lambda$CDM cosmology shown in Figure \ref{fig:simple_om_summary}.
+{\red Need a think about which plots to show or good way to merge them}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_simple_test_cosmo.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
+	\label{fig:simple_om}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_simple_test_w_cosmo.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $w$CDM model.}
+	\label{fig:simple_w}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_simple_test_ol_cosmo.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}.
+	\label{fig:simple_ol}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_simple_test_summary.pdf}
+	\end{center}
+	\caption{Parameter summaries for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}.
+	\label{fig:simple_om_summary}
+\end{figure}
+
+\subsection{DES SN data validation}
 \label{sec:simdes}
@@
 Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require more a more sophisticated approach.
-In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
-The first is the model from \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The second model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+In our development of this model and tests of intrinsic dispersion, we analyse the effects of three different scatter models.
+The first is simplest model - coherent magnitude scatter.
+The second model is the \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The third model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
 
-\subsection{DES SN data validation}
-\label{sec:des}
+Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys.
+In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and now also include our full treatment of systematics.
+Additionally, we also simulate two different underlying population -- a Gaussian distribution in colour and stretch, and a highly skewed colour and stretch population using population values from \citet{Scolnic2016}.
+Both populations have been centered on 0, with the bifurcated Gaussian used for the skewed colour distribution having width $\sigma_c^+ = 0.101$, $\sigma_c^- = 0.043$ and the skewed stretch distribution having $\sigma_{x_1}^+ = 0.222$, $\sigma_{x_1}^- = 1.472$.
 
-appoximate\_simple\_test.py
+Each realisation of cosmology fitted contains 300 LowZ supernovae, and 250 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+Combined posterior surfaces for {\red 20} data realisations are shown in Figure \ref{fig:bulk_summary}.
+{\green Talk about the biases introduced}
 
-multisim
+{\red UPDATE THIS WHEN THE SIMS FINISH AND I CAN RERUN WITH MORE REALISATIONS}
 
-bulk
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_bulk_load_summary.pdf}
+	\end{center}
+	\caption{Parameter summaries for 20 realisations of supernova data with the Flat $\Lambda$CDM model.}.
+	\label{fig:bulk_summary}
+\end{figure}
 

### 2017-10-03 dessn/configurations/approximate_bulk_load.py

@@ -55,4 +55,4 @@ if __name__ == "__main__":
     c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=params, truth=truth, figsize=1.5)
     with open(pfn + ".txt", 'w') as f:
         f.write(c.analysis.get_latex_table(parameters=params))
-    c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=params, errorbar=True, truth=truth, extra_parameter_spacing=1.0)
+    c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=params, errorbar=True, truth=truth)

### 2017-10-03 dessn/configurations/approximate_bulk_c11_skew_test.py
Swapping mag smear with combined g10 and c11 models

@@ -25,8 +25,8 @@ if __name__ == "__main__":
     fitter = Fitter(dir_name)
     fitter.set_models(model)
     fitter.set_simulations(simulation)
-    fitter.set_num_cosmologies(20)
-    fitter.set_num_walkers(10)
+    fitter.set_num_cosmologies(100)
+    fitter.set_num_walkers(3)
     fitter.set_max_steps(3000)
 
     h = socket.gethostname()

### 2017-10-03 dessn/configurations/approximate_bulk_combined_gauss_test.py

@@ -3,7 +3,7 @@ import logging
 import socket
 from dessn.framework.fitter import Fitter
 from dessn.framework.models.approx_model import ApproximateModel
-from dessn.framework.simulations.snana_bulk import SNANABulkSimulation
+from dessn.framework.simulations.snana_bulk import SNANACombinedBulk
 
 
 if __name__ == "__main__":
@@ -20,14 +20,16 @@ if __name__ == "__main__":
 
     model = ApproximateModel()
     # Turn off mass and skewness for easy test
-    simulation = [SNANABulkSimulation(300, sim="SHINTON_LOWZ_MATRIX_SMEAR_SKEWC_SKEWX1", manual_selection=[13.70+0.5, 1.363, 3.8, 0.2], num_calib=58),
-                  SNANABulkSimulation(250, sim="SHINTON_DES_MATRIX_SMEAR_SKEWC_SKEWX1", manual_selection=[22.3, 0.7, None, 1.0], num_calib=22)]
+    simulation = [SNANACombinedBulk(300, ["SHINTON_LOWZ_MATRIX_G10_SYMC_SYMX1", "SHINTON_LOWZ_MATRIX_C11_SYMC_SYMX1"],
+                                    "CombinedLowZ", manual_selection=[13.70+0.5, 1.363, 3.8, 0.2], num_calib=58),
+                  SNANACombinedBulk(250, ["SHINTON_DES_MATRIX_G10_SYMC_SYMX1", "SHINTON_DES_MATRIX_C11_SYMC_SYMX1"],
+                                    "CombinedDES", manual_selection=[22.3, 0.7, None, 1.0], num_calib=22)]
 
     fitter = Fitter(dir_name)
     fitter.set_models(model)
     fitter.set_simulations(simulation)
-    fitter.set_num_cosmologies(20)
-    fitter.set_num_walkers(10)
+    fitter.set_num_cosmologies(100)
+    fitter.set_num_walkers(3)
     fitter.set_max_steps(3000)
 
     h = socket.gethostname()
@@ -48,3 +50,4 @@ if __name__ == "__main__":
         c.add_chain(chain, weights=weight, posterior=posterior, name="Approx")
         c.configure(label_font_size=10, tick_font_size=10, diagonal_tick_labels=False)
         c.plotter.plot_distributions(filename=pfn + "_dist.png", truth=truth, col_wrap=8)
+

### 2017-10-03 dessn/configurations/approximate_bulk_combined_skew_test.py

@@ -3,7 +3,7 @@ import logging
 import socket
 from dessn.framework.fitter import Fitter
 from dessn.framework.models.approx_model import ApproximateModel
-from dessn.framework.simulations.snana_bulk import SNANABulkSimulation
+from dessn.framework.simulations.snana_bulk import SNANACombinedBulk
 
 
 if __name__ == "__main__":
@@ -20,14 +20,16 @@ if __name__ == "__main__":
 
     model = ApproximateModel()
     # Turn off mass and skewness for easy test
-    simulation = [SNANABulkSimulation(300, sim="SHINTON_LOWZ_MATRIX_SMEAR_SYMC_SYMX1", manual_selection=[13.70+0.5, 1.363, 3.8, 0.2], num_calib=58),
-                  SNANABulkSimulation(250, sim="SHINTON_DES_MATRIX_SMEAR_SYMC_SYMX1", manual_selection=[22.3, 0.7, None, 1.0], num_calib=22)]
+    simulation = [SNANACombinedBulk(300, ["SHINTON_LOWZ_MATRIX_G10_SKEWC_SKEWX1", "SHINTON_LOWZ_MATRIX_C11_SKEWC_SKEWX1"],
+                                    "CombinedLowZ", manual_selection=[13.70+0.5, 1.363, 3.8, 0.2], num_calib=58),
+                  SNANACombinedBulk(250, ["SHINTON_DES_MATRIX_G10_SKEWC_SKEWX1", "SHINTON_DES_MATRIX_C11_SKEWC_SKEWX1"],
+                                    "CombinedDES", manual_selection=[22.3, 0.7, None, 1.0], num_calib=22)]
 
     fitter = Fitter(dir_name)
     fitter.set_models(model)
     fitter.set_simulations(simulation)
-    fitter.set_num_cosmologies(20)
-    fitter.set_num_walkers(10)
+    fitter.set_num_cosmologies(100)
+    fitter.set_num_walkers(3)
     fitter.set_max_steps(3000)
 
     h = socket.gethostname()

### 2017-10-03 dessn/configurations/approximate_bulk_g10_gauss_test.py

@@ -26,8 +26,8 @@ if __name__ == "__main__":
     fitter = Fitter(dir_name)
     fitter.set_models(model)
     fitter.set_simulations(simulation)
-    fitter.set_num_cosmologies(20)
-    fitter.set_num_walkers(10)
+    fitter.set_num_cosmologies(100)
+    fitter.set_num_walkers(3)
     fitter.set_max_steps(3000)
 
     h = socket.gethostname()

### 2017-10-03 dessn/framework/simulations/snana_bulk.py
Addendum to prev commit

@@ -41,7 +41,7 @@ class SNANABulkSimulation(Simulation):
             ("intrinsic_correlation", np.identity(3), r"$\rho$"),
             ("calibration", np.zeros(self.num_calib), r"$\delta \mathcal{Z}_%d$")
         ]
-    
+
     def get_name(self):
         return self.sim
 
@@ -110,3 +110,45 @@ class SNANABulkSimulation(Simulation):
             "passed": np.ones(n_sne, dtype=np.bool)
         }
         return result
+
+
+class SNANACombinedBulk(Simulation):
+    def __init__(self, n_sne, sims, name, manual_selection=None, use_sim=False, num_nodes=4, num_calib=0):
+        super().__init__()
+        self.name = name
+        n_sim = int(n_sne / len(sims))
+        self.sim_names = sims
+        self.manual_selection = manual_selection
+        self.num_supernova = n_sne
+        self.use_sim = use_sim
+        self.num_nodes = num_nodes
+        self.num_calib = num_calib
+        self.sims = [SNANABulkSimulation(n_sim, s, manual_selection=manual_selection,
+                                         use_sim=use_sim, num_nodes=num_nodes, num_calib=num_calib) for s in sims]
+
+    def get_approximate_correction(self, plot=False, manual=None):
+        if self.manual_selection is not None:
+            return self.manual_selection
+        else:
+            raise ValueError("Bulk combined must specify manual selection")
+
+    def get_truth_values(self):
+        return self.sims[0].get_truth_values()
+
+    def get_name(self):
+        return self.name
+
+    def get_all_supernova(self, n_sne, cosmology_index=0):
+        n_sim = int(n_sne / len(self.sims))
+        results = [s.get_all_supernova(n_sim, cosmology_index=cosmology_index) for s in self.sims]
+
+        # Now merge together
+        n_tot = np.sum([r["n_sne"] for r in results])
+        final = results[0]
+        final["n_sne"] = int(n_tot)
+        for r in results[1:]:
+            for key in final.keys():
+                if key in ["n_sne"]:
+                    continue
+                final[key] = np.concatenate((final[key], r[key]))
+        return final

### 2017-10-03 papers/methods/paper.tex
Updating paper

@@
 w_{\rm DES} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm DES}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm DES}^2}}  \right) \\
-w_{\rm LowZ} &= \mathcal{N}\left( \frac{m_B^* - \mu_{\rm LowZ}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm LowZ}^2}}\right) \times \notag\\
+w_{\rm LowZ} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm LowZ}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm LowZ}^2}}\right) \times \notag\\
 &\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm LowZ})(m_B^* - \mu_{\rm LowZ})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}{\sigma^2_{\rm LowZ}} \sqrt{\frac{\sigma^2_{\rm LowZ}}{\alpha^2_{\rm LowZ}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm LowZ}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}} }\right),
@@
 	\end{center}
-	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}.
+	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}
 	\label{fig:simple_ol}
@@
 	\end{center}
-	\caption{Parameter summaries for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}.
+	\caption{Parameter summaries for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
 	\label{fig:simple_om_summary}
@@
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of three different scatter models.
-The first is simplest model - coherent magnitude scatter.
+The first is simplest model - coherent magnitude scatter, denoted `Smear' in the figures.
 The second model is the \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
@@
 	\end{center}
-	\caption{Parameter summaries for 20 realisations of supernova data with the Flat $\Lambda$CDM model.}.
+	\caption{Parameter summaries for 20 realisations of supernova data with the Flat $\Lambda$CDM model. {\red Discuss any biases}}
 	\label{fig:bulk_summary}
 \end{figure}
+
+{\red More discussion here when I get more results}
 
@@
 
-\subsubsection{Spectroscopic Sample}
-
-\subsubsection{Photometric sample}
-
-\subsection{Comparison with bells and whistles fixed}
+{\red Put in multisim forecasts with systamtics both on and off}
 
@@
 
-\section{Interesting Implementation Details}
-\label{sec:details}
-Anything interesting.
-
-Also talk about non-analytic correction factors (and their failure - mc integration, GP, NNGP)
-
 \section{Conclusions}
@@
 \section*{Acknowledgements}
+
+Plots of posterior surfaces and parameter summaries were created with \verb|ChainConsumer| \citep{Hinton2016}.
 

### 2017-10-09 dessn/configurations/combined.py
Adding combined forecast sim

@@ -2,7 +2,7 @@ import os
 import logging
 import socket
 from dessn.framework.fitter import Fitter
-from dessn.framework.models.approx_model import ApproximateModel
+from dessn.framework.models.approx_model import ApproximateModelW, ApproximateModel, ApproximateModelOl
 from dessn.framework.simulations.snana_bulk import SNANACombinedBulk
 from dessn.framework.simulations.selection_effects import lowz_sel, des_sel
 
@@ -18,7 +18,10 @@ if __name__ == "__main__":
     if not os.path.exists(dir_name):
         os.makedirs(dir_name)
 
-    model = ApproximateModel()
+    models = ApproximateModelW(), ApproximateModelW(systematics_scale=0.001), \
+             ApproximateModel(), ApproximateModel(systematics_scale=0.001), \
+             ApproximateModelOl(), ApproximateModelOl(systematics_scale=0.001)
+
     # Turn off mass and skewness for easy test
     simulation = [SNANACombinedBulk(300, ["SHINTON_LOWZ_MATRIX_G10_SKEWC_SKEWX1", "SHINTON_LOWZ_MATRIX_C11_SKEWC_SKEWX1"],
                                     "CombinedLowZ", manual_selection=lowz_sel(), num_calib=58),
@@ -26,10 +29,10 @@ if __name__ == "__main__":
                                     "CombinedDES", manual_selection=des_sel(), num_calib=22)]
 
     fitter = Fitter(dir_name)
-    fitter.set_models(model)
+    fitter.set_models(*models)
     fitter.set_simulations(simulation)
     fitter.set_num_cosmologies(100)
-    fitter.set_num_walkers(3)
+    fitter.set_num_walkers(1)
     fitter.set_max_steps(3000)
 
     h = socket.gethostname()

### 2017-10-09 dessn/configurations/approximate_bulk_load.py
Plotting updates

@@ -25,7 +25,7 @@ if __name__ == "__main__":
 
     expression = re.compile("approximate_bulk_(.*)_test.py")
     matches = [re.match(expression, f) for f in sorted(os.listdir(dir_name))]
-    names = [m.group(1).replace("_", " ").title() for m in matches if m is not None]
+    names = [m.group(1).replace("_", " ").title().replace("Gauss", "Gaussian").replace("Skew", "SK16") for m in matches if m is not None]
     filenames = [m.string[:-3] for m in matches if m is not None]
 
     dir_names = [os.path.dirname(os.path.abspath(__file__)) + "/plots/%s/output" % f for f in filenames]
@@ -49,11 +49,10 @@ if __name__ == "__main__":
     ls = ["-", ":"] * 3
     colors = ['b', 'b', 'r', 'r', 'g', 'g']
     alphas = 0.1
-    c.configure(label_font_size=10, tick_font_size=10, diagonal_tick_labels=False, linestyles=ls,
-                colors=colors, shade_alpha=alphas, shade=True)
-    params = ['$\\Omega_m$', '$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$']
+    c.configure(diagonal_tick_labels=False, linestyles=ls, colors=colors, shade_alpha=alphas, shade=True)
+    params = ['$\\Omega_m$', '$\\alpha$', '$\\beta$']
     c.plotter.plot_distributions(filename=pfn + "_dist.png", truth=truth, col_wrap=8)
-    c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=params, truth=truth, figsize=1.5)
+    c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=params, truth=truth, figsize="column")
     with open(pfn + ".txt", 'w') as f:
         f.write(c.analysis.get_latex_table(parameters=params))
-    c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=params, errorbar=True, truth=truth)
+    c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=params, errorbar=True, truth=truth, extra_parameter_spacing=1.0)

### 2017-10-09 dessn/configurations/approximate_systematic_load.py

@@ -24,8 +24,8 @@ if __name__ == "__main__":
              "10\\% Gauss error in biasCor flux errors", "idem, but with incorrect reported fluxErr",
              "MWEBV scale from 20\\% Gauss error", "MW RV shift from 0.2 Gauss error"]
     names = ["Stat only", "ZP offset", r"Filter $\lambda$ shift",
-             "biasCor flux errors", "idem + fluxErr",
-             "MW EBV scale error", "MW RV shift error"]
+             "Bias corrected flux error", "idem + uncertainty error",
+             "MW $E(B-V)$ scale error", "MW $R_V$ shift"]
     dir_names = [os.path.dirname(os.path.abspath(__file__)) + "/plots/%s/output/" % f for f in filenames]
 
     c = ChainConsumer()
@@ -50,5 +50,5 @@ if __name__ == "__main__":
     c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=params)
     print("Plotting summary")
     c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=['$\\Omega_m$'],
-                           extents={'$\\Omega_m$': [0.1, 0.45]}, truth="Stat only", extra_parameter_spacing=1.0)
+                           extents={'$\\Omega_m$': [0.1, 0.45]}, truth="Stat only", extra_parameter_spacing=1.5)
 

### 2017-10-09 dessn/configurations/approximate_systematic_load_w.py

@@ -24,8 +24,8 @@ if __name__ == "__main__":
              "10\\% Gauss error in biasCor flux errors", "idem, but with incorrect reported fluxErr",
              "MWEBV scale from 20\\% Gauss error", "MW RV shift from 0.2 Gauss error"]
     names = ["Stat only", "ZP offset", r"Filter $\lambda$ shift",
-             "biasCor flux errors", "idem + fluxErr",
-             "MW EBV scale error", "MW RV shift error"]
+             "Bias corrected flux error", "idem + uncertainty error",
+             "MW $E(B-V)$ scale error", "MW $R_V$ shift"]
     dir_names = [os.path.dirname(os.path.abspath(__file__)) + "/plots/%s/output/" % f for f in filenames]
 
     c = ChainConsumer()
@@ -50,5 +50,5 @@ if __name__ == "__main__":
     # c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=params)
     print("Plotting summary")
     c.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], parameters=["$w$"],
-                           extents={"$w$": [-1.5, -0.5]}, truth="Stat only", extra_parameter_spacing=1.0)
+                           extents={"$w$": [-1.5, -0.5]}, truth="Stat only", extra_parameter_spacing=1.5)
 

### 2017-10-09 papers/methods/paper.tex
Paper updates

@@
 \defcitealias{Rubin2015}{R15}	
+\defcitealias{Scolnic2016}{SK16}	 
 \newcommand{\gten}{\citetalias{Guy2010}}
@@
 \newcommand{\rubin}{\citetalias{Rubin2015}}
+\newcommand{\sk}{\citetalias{Scolnic2016}}
+\newcommand{\angstrom}{\mbox{\normalfont\AA}}
 
@@
 \subsubsection{Systematics}
-
+\label{sec:systreat}
 The chief difficulty with including systematics in supernova analyses is that the generally occur during the observational pipeline, and have difficult to model effects on the output observations.
@@
 We test three models (Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM) with 300 realisations and find recovery of underlying cosmology without significant bias.
-Combined posterior surfaces of 22 realisations are shown in Figures \ref{fig:simple_om}, \ref{fig:simple_w}, \ref{fig:simple_ol}, with the statistical distribution of realisations for the Flat $\Lambda$CDM cosmology shown in Figure \ref{fig:simple_om_summary}.
-{\red Need a think about which plots to show or good way to merge them}
+Combined posterior surfaces of 22 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_simple_test_cosmo.pdf}
-	\end{center}
-	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
-	\label{fig:simple_om}
-\end{figure}
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[width=\columnwidth]{approximate_simple_test_cosmo.pdf}
+%	\end{center}
+%	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
+%	\label{fig:simple_om}
+%\end{figure}
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_simple_test_ol_cosmo.pdf}
-	\end{center}
-	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}
-	\label{fig:simple_ol}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_simple_test_summary.pdf}
-	\end{center}
-	\caption{Parameter summaries for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
-	\label{fig:simple_om_summary}
-\end{figure}
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[width=\columnwidth]{approximate_simple_test_ol_cosmo.pdf}
+%	\end{center}
+%	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}
+%	\label{fig:simple_ol}
+%\end{figure}
 
@@
 \label{sec:simdes}
-{\green BUT THE BELOW IN THE TESTING AREA}
 
 Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require more a more sophisticated approach.
-In our development of this model and tests of intrinsic dispersion, we analyse the effects of three different scatter models.
-The first is simplest model - coherent magnitude scatter, denoted `Smear' in the figures.
-The second model is the \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The third model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
+The first model is the \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+We also test a `Combined' smear model, which is simply to draw equal samples from the {\celeven} and {\gten} models.
+It is against this combined sample that we wish to rigorously validate our model.
 
@@
 In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and now also include our full treatment of systematics.
-Additionally, we also simulate two different underlying population -- a Gaussian distribution in colour and stretch, and a highly skewed colour and stretch population using population values from \citet{Scolnic2016}.
-Both populations have been centered on 0, with the bifurcated Gaussian used for the skewed colour distribution having width $\sigma_c^+ = 0.101$, $\sigma_c^- = 0.043$ and the skewed stretch distribution having $\sigma_{x_1}^+ = 0.222$, $\sigma_{x_1}^- = 1.472$.
+Additionally, we also simulate two different underlying population -- a Gaussian distribution in colour and stretch, and skewed colour and stretch populations using population values from \citet[][hereafter {\sk}]{Scolnic2016}.
+
+\begin{table}
+	\centering
+	\caption{Tested population distributions, where the {\sk} LowZ distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
+	\label{tab:dist}
+	\begin{tabular}{l|llll}
+		\hline
+		Model & $\langle x_1 \rangle$ & $\sigma_{x_1}$ &  $\langle c \rangle$ & $\sigma_c$  \\
+		\hline
+		Gaussian LowZ &  $0.0$ & $1.0$ & $0.0$ & $0.1$ \\
+		Gaussian DES  &  $0.0$ & $1.0$ & $0.0$ & $0.1$  \\
+		{\sk} LowZ    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
+		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
+		\hline
+	\end{tabular}
+\end{table}
 
 Each realisation of cosmology fitted contains 300 LowZ supernovae, and 250 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-Combined posterior surfaces for {\red 20} data realisations are shown in Figure \ref{fig:bulk_summary}.
-{\green Talk about the biases introduced}
+Combined posterior surfaces for 150 data realisations are shown in Figure \ref{fig:bulk_summary}.
+Note that we have combined the posteriors for 150 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{150} \approx 12$ times less than a single realisation.
+Plot summaries for the fits are shown in Figure \ref{fig:bulk_summary}, and the parameter bounds are listed in Table \ref{tab:bulk_summary}.
+The results indicate changes in the underlying population can effect significant changes in the recovered standardisation parameters $\alpha$ and $\beta$, however do not appear to have a significant effect on recovered cosmology.
+Whilst the combined sample (on which the selection function was fit) shows no significant bias (as expected), the {\celeven} and {\gten} only models show opposite biases in $\Omega_m$.
+Cosmological bias is detected with $\approx 3.0\sigma$ significance for the {\celeven} model and $\approx 1.7\sigma$ significance for the {\gten} model, indicating that the population and selection effect treatment cannot capture all necessary information to encapsulate the magnitude and chromatic smearing of the supernovae population.
+However, for the sample size of the DES and LowZ supernova samples (of order 600 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $<0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
 
-{\red UPDATE THIS WHEN THE SIMS FINISH AND I CAN RERUN WITH MORE REALISATIONS}
+\begin{table}
+	\centering
+	\caption{Parameter summaries shown for 150 realisations of supernovae data, each comprising of 300 LowZ supernovae and 250 DES-like supernovae.
+Maximum likelihood statistics, with errors quoted to $1\sigma$ (68\% confidence) are reported.}
+	\label{tab:bulk_summary}
+	\begin{tabular}{ccccc}
+		\hline
+		Model & $\Omega_m$ & $\alpha$ & $\beta$  \\ 
+		\hline
+		C11 Gaussian & $0.316^{+0.062}_{-0.064}$ & $0.143^{+0.028}_{-0.027}$ & $2.97^{+0.20}_{-0.19}$  \\ 
+		C11 {\sk} & $0.316^{+0.063}_{-0.062}$ & $0.155\pm 0.024$ & $2.77^{+0.26}_{-0.27}$  \\ 
+		Combined Gaussian & $0.299^{+0.067}_{-0.060}$ & $0.144^{+0.030}_{-0.028}$ & $3.11^{+0.21}_{-0.22}$  \\ 
+		Combined {\sk} & $0.297^{+0.075}_{-0.064}$ & $0.150^{+0.026}_{-0.025}$ & $3.00^{+0.29}_{-0.31}$  \\ 
+		G10 Gaussian & $0.290^{+0.070}_{-0.060}$ & $0.145\pm 0.030$ & $3.27\pm 0.23$ \\ 
+		G10 {\sk} & $0.290^{+0.079}_{-0.073}$ & $0.147^{+0.026}_{-0.028}$ & $3.26^{+0.34}_{-0.33}$ \\ 
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 	\end{center}
-	\caption{Parameter summaries for 20 realisations of supernova data with the Flat $\Lambda$CDM model. {\red Discuss any biases}}
+	\caption{Graphical parameter summaries for 150 realisations of supernova data, following Table \ref{tab:bulk_summary}.}
 	\label{fig:bulk_summary}
@@
 
-{\red More discussion here when I get more results}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_bulk_load.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 150 realisations of supernova data.
+Neither population nor scatter model has a significant effect on posterior surface location or shape.
+Population skewness has the primary effect of changing the degeneracy direction for the $\alpha$ -- $\beta$ contour, whilst the scatter models and skewness have noticeable effects on the value and uncertainty of the recovered parameter $\beta$.}
+	\label{fig:bulk_posterior}
+\end{figure}
+
+\section{Systematics Strength Test}
+\label{sec:sys}
+
+\subsection{Singular Permutations}
+
+To test the impact of various systematics, we generate an array of simulations in which we simulate a dataset with one of the systematic variables permuted.
+We inspect the effects of systematic shifts using two cosmological models - Flat $\Lambda$CDM and Flat $w$CDM.
+To make the effects of the systematics more easily seen, we increase our simulated sample size to 500 LowZ and 500 DES-like supernovae, and enforce a prior on $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ on the Flat $w$CDM model.
+In these test we wish to inspect the shift in statistical uncertainty with simulated systematic effects, and so compare first to a series of simulations with no simulated systematic changes, and also disable our treatment of systematics as described in Section \ref{sec:systreat} for these tests.
+Our tests are broken down into groups, where systematics of different nature are treated separately.
+In each simulation, only one systemic is permuted.
+For a baseleine, our first set has no systematic shift, and so the uncertainty in the parameter summaries is given by statistical uncertainty only.
+We then test several variations:
+\begin{enumerate}
+	\item 0.02 Gaussian error on the zero point offsets in a randomly selected band.
+	\item $20 \angstrom $ Gaussian error on the filter wavelength.
+	\item 10\% Gaussian error in the bias corrected flux measurement (with correct flux uncertainty).
+	\item 10\% Gaussian error in the bias corrected flux measurement, but with incorrect flux error.
+	\item 20\% Gaussian error on the Milky Way E(B-V) dust correction scale.
+	\item 0.2 Gaussian error on the value of $R_V$ for calculating the extinction curve.
+\end{enumerate}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_systematic_load_summary.pdf}
+	\end{center}
+	\caption{Investigating the impact of $\Omega_m$ by systematic permutations under Flat $\Lambda$CDM cosmology, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
+	\label{fig:sys_om}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{approximate_systematic_load_w_summary.pdf}
+	\end{center}
+	\caption{Investigating the impact of $w$ by systematic permutations under $w$CDM cosmology with the prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
+	\label{fig:sys_w}
+\end{figure}
+
+We find that the most significant systematic are the zero points, as the 0.02 Gaussian error given to them resulted in significant changes to the shape of the marginalised posterior surface for both $\Omega_m$ and $w$, shifting them both to lower values and increasing uncertainty.
+
+{\red What are the actual size of the uncertainties of the systematics? Ie are zps really that badly known? Once I know, I can talk about if any of these are significant.}
+
+\subsection{Full Permutations}
+
+{\red When I have the final sims from Rick I can flesh this out}
 
@@
 {\red Put in multisim forecasts with systamtics both on and off}
-
-\section{Systematics Strength Test}
-\label{sec:sys}
-
-systematics test
 

### 2017-10-10 dessn/configurations/combined.py
15k samples for combined

@@ -33,7 +33,7 @@ if __name__ == "__main__":
     fitter.set_simulations(simulation)
     fitter.set_num_cosmologies(100)
     fitter.set_num_walkers(1)
-    fitter.set_max_steps(5000)
+    fitter.set_max_steps(15000)
 
     h = socket.gethostname()
     if h != "smp-hk5pn72":  # The hostname of my laptop. Only will work for me, ha!

### 2017-10-10 papers/methods/paper.tex
Paper updates

@@
 Cosmological bias is detected with $\approx 3.0\sigma$ significance for the {\celeven} model and $\approx 1.7\sigma$ significance for the {\gten} model, indicating that the population and selection effect treatment cannot capture all necessary information to encapsulate the magnitude and chromatic smearing of the supernovae population.
-However, for the sample size of the DES and LowZ supernova samples (of order 600 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $<0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
+However, for the sample size of the DES and LowZ supernova samples (of order 600 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
 
@@
 
-\section{Forecasts}
+\section{DES Forecasts}
 \label{sec:forecasts}
 
-{\red Put in multisim forecasts with systamtics both on and off}
+To forecast the results for the 3 year DES spectroscopic supernovae sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
+We fit these through the three cosmological models - Flat $\Lambda$CDM, $\Lambda$CDM and $w$CDM, however focus mostly on the $w$CDM model as it is the primary scientific concern of the Dark Energy Survey.
+We run all three cosmological models with systematics enabled and disabled to explicitly show the uncertainty included from systematic contribution.
+
+{\red How to include flat lambdacdm - the plot is boring, just quote the results, or just quote uncertainty, should I round the uncertainty to make it symmetric?
+Cant do that for wCDM or LambdaCDM because theyre definitely not gaussian}
+
+{\red Talk about how these arent even competitive with JLA and we really need to combine data. Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{combined_ApproximateModelW.pdf}
+	\end{center}
+	\caption{w.}
+	\label{fig:combined_w}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{combined_ApproximateModelOl.pdf}
+	\end{center}
+	\caption{ole.}
+	\label{fig:combined_ol}
+\end{figure}
 

### 2017-10-11 papers/methods/paper.tex
Selection effects appendix

@@
 \subsubsection{Selection Effects}
-
+\label{sec:selection}
 Our treatment of selection effects is to incorporate selection efficiency into our model, rather than relying on simulation-driven data corrections.
@@
 \begin{align}
-w_{\rm DES} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm DES}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm DES}^2}}  \right) \\
+w_{\rm DES} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm DES}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm DES}^2}}  \right) \label{eq:seldes}\\
 w_{\rm LowZ} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm LowZ}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm LowZ}^2}}\right) \times \notag\\
-&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm LowZ})(m_B^* - \mu_{\rm LowZ})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}{\sigma^2_{\rm LowZ}} \sqrt{\frac{\sigma^2_{\rm LowZ}}{\alpha^2_{\rm LowZ}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm LowZ}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}} }\right),
+&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm LowZ})(m_B^* - \mu_{\rm LowZ})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}{\sigma^2_{\rm LowZ}} \sqrt{\frac{\sigma^2_{\rm LowZ}}{\alpha^2_{\rm LowZ}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm LowZ}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}} }\right), \label{eq:sellowz}
 \end{align}
@@
 
+\subsection{General Selection Effects}
+
+When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
+\begin{align}
+P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
+\end{align}
+which gives the probability of the model parameter values ($\theta$) given the data.
+Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$.
+Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$.
+When dealing with experiments which have imperfect selection efficiency, our likelihood must take that efficiency into account.
+We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
+To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
+\begin{align}
+\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:app_like}
+\end{align}
+A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.
+That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}.
+We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
+\begin{align}
+P({\rm data} | S , \theta) P(S, \theta) &= P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
+P({\rm data} | S , \theta) &= \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
+&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
+&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
+\end{align}
+which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
+\begin{align}
+\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
+\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
+\end{align}
+where we define the denominator as $w$ for simplicity in future derivations.
+
+\subsection{Supernova Selection Effects}
+
+We assume that our selection effects can be reasonably well encapsulated by independent functions of (actual) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
+Our denominator then becomes
+\begin{align}
+w &= \idotsint \, d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta),
+\end{align}
+where for simplicity we have not written out all the integrals which do not interact with the selection effects explicitly.
+Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
+$P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$.
+Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.
+We can thus state
+\begin{align}
+w = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
+\end{align}
+By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernovae redshifts, we can approximate the integral over redshift by evaluating
+\begin{align}
+\int P(S|m_B) P(m_B | z, \theta)\, dm_B
+\end{align}
+for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.
+
+As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
+\begin{align}
+m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
+\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2 \notag\\ 
+&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c}).
+\end{align}
+Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-10-11 dessn/configurations/combined.py

@@ -42,6 +42,7 @@ if __name__ == "__main__":
     else:
         from chainconsumer import ChainConsumer
         results = fitter.load()
+        print("Data loaded")
 
         classes = list(set([r[0].__class__.__name__ for r in results]))
         for cls in ["ApproximateModelW"]:  # classes:

### 2017-12-12 papers/methods/paper.tex
Updating paper. Removing statistics

@@
 This requires a more complicated selection function than the error function used for DES, however is reasonably well fit with a skew normal.
-The selection functions were fit to apparent magnitude efficiency ratios calculated from SNANA simulations.
-The DES survey is thus characterised by $\mu_{\rm DES} = 22.3$, $\sigma_{\rm DES} = 0.7$, with the combined low redshift surveys characterised by $\mu_{\rm LowZ} = 14.1$, $\sigma_{\rm LowZ} = 1.36$, $\alpha_{\rm LowZ} = 3.8$.
+
+The selection functions were fit to apparent magnitude efficiency ratios calculated from SNANA simulations by calculating an efficiency ratio as a function of apparent magnitude.
+Uncertainty of the Malmquist bias (entering both through statistical uncertainty from finite sized simulations in the efficiency ratio and the discrepancy between the analytic approximation and non-analytic simulation results) is incorporated into the fitting for the analytic approximation.
+Uncertainty is uniformly added to the efficiency ratio until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits.
+
+The DES survey is thus characterised by mean fit values $\mu_{\rm DES} = 22.3$, $\sigma_{\rm DES} = 0.7$, with the combined low redshift surveys characterised by $\mu_{\rm LowZ} = 14.1$, $\sigma_{\rm LowZ} = 1.36$, $\alpha_{\rm LowZ} = 3.8$.
 
@@
 
-\section{Systematics Strength Test}
-\label{sec:sys}
-
-\subsection{Singular Permutations}
-
-To test the impact of various systematics, we generate an array of simulations in which we simulate a dataset with one of the systematic variables permuted.
+%
+%\section{Systematics Strength Test}
+%\label{sec:sys}
+%
+%\subsection{Singular Permutations}
+%
+%To test the impact of various systematics, we generate an array of simulations in which we simulate a dataset with one of the systematic variables permuted.
 We inspect the effects of systematic shifts using two cosmological models - Flat $\Lambda$CDM and Flat $w$CDM.
@@
 We then test several variations:
-\begin{enumerate}
-	\item 0.02 Gaussian error on the zero point offsets in a randomly selected band.
-	\item $20 \angstrom $ Gaussian error on the filter wavelength.
-	\item 10\% Gaussian error in the bias corrected flux measurement (with correct flux uncertainty).
-	\item 10\% Gaussian error in the bias corrected flux measurement, but with incorrect flux error.
-	\item 20\% Gaussian error on the Milky Way E(B-V) dust correction scale.
-	\item 0.2 Gaussian error on the value of $R_V$ for calculating the extinction curve.
-\end{enumerate}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_systematic_load_summary.pdf}
-	\end{center}
-	\caption{Investigating the impact of $\Omega_m$ by systematic permutations under Flat $\Lambda$CDM cosmology, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
-	\label{fig:sys_om}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_systematic_load_w_summary.pdf}
-	\end{center}
-	\caption{Investigating the impact of $w$ by systematic permutations under $w$CDM cosmology with the prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
-	\label{fig:sys_w}
-\end{figure}
-
-We find that the most significant systematic are the zero points, as the 0.02 Gaussian error given to them resulted in significant changes to the shape of the marginalised posterior surface for both $\Omega_m$ and $w$, shifting them both to lower values and increasing uncertainty.
-
-{\red What are the actual size of the uncertainties of the systematics? Ie are zps really that badly known? Once I know, I can talk about if any of these are significant.}
-
-\subsection{Full Permutations}
-
-{\red When I have the final sims from Rick I can flesh this out}
+%\begin{enumerate}
+%	\item 0.02 Gaussian error on the zero point offsets in a randomly selected band.
+%	\item $20 \angstrom $ Gaussian error on the filter wavelength.
+%	\item 10\% Gaussian error in the bias corrected flux measurement (with correct flux uncertainty).
+%	\item 10\% Gaussian error in the bias corrected flux measurement, but with incorrect flux error.
+%	\item 20\% Gaussian error on the Milky Way E(B-V) dust correction scale.
+%	\item 0.2 Gaussian error on the value of $R_V$ for calculating the extinction curve.
+%\end{enumerate}
+%
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[width=\columnwidth]{approximate_systematic_load_summary.pdf}
+%	\end{center}
+%	\caption{Investigating the impact of $\Omega_m$ by systematic permutations under Flat $\Lambda$CDM cosmology, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
+%	\label{fig:sys_om}
+%\end{figure}
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[width=\columnwidth]{approximate_systematic_load_w_summary.pdf}
+%	\end{center}
+%	\caption{Investigating the impact of $w$ by systematic permutations under $w$CDM cosmology with the prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
+%	\label{fig:sys_w}
+%\end{figure}
+%
+%We find that the most significant systematic are the zero points, as the 0.02 Gaussian error given to them resulted in significant changes to the shape of the marginalised posterior surface for both $\Omega_m$ and $w$, shifting them both to lower values and increasing uncertainty.
+%
+%{\red What are the actual size of the uncertainties of the systematics? Ie are zps really that badly known? Once I know, I can talk about if any of these are significant.}
+%
+%\subsection{Full Permutations}
+%
+%{\red When I have the final sims from Rick I can flesh this out}
 
@@
 \section{Conclusions}
+
+Conclude.
 

### 2018-01-15 papers/methods/paper.tex
Updating paper

@@
 \usepackage{amssymb}	% Extra maths symbols
-
+\usepackage{tikz}
+\usetikzlibrary{bayesnet}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In this paper, we lay out a new hierarchical model that extends on past work.
+In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015}.
 Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
@@
 Model verification on simulated datasets is given in Section \ref{sec:verification}.
-Forecasts for the impending DES three year spectroscopic supernova survey are contain in Section \ref{sec:forecast}.
-Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.
+Forecasts for the impending DES three year spectroscopic supernova survey are contained in Section \ref{sec:forecasts}.
+{\red Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.
+}
 
@@
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
-As such, the product at the end is a (redshift dependent) population of $m_B$, $x_1$ and $c$.
+As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$.
 
 The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology.
-However, given some modelled underlying population that lives in the redshift dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
+However, given some modelled underlying population that lives in the redshift-dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
 Specifically, for any given supernova our functional map may take the traditional form:
@@
 \end{equation}
-where $C^{-1}$ is an uncertainty matrix which combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
-The benefit this analysis methodology provides is speed - for samples of hundreds of supernova, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
+where $C^{-1}$ is an uncertainty matrix that combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
+The benefit this analysis methodology provides is speed - for samples of hundreds of supernovae, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
 The speed comes with two costs.
@@
 Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data.
-Simulations following survey observational strategies and geometry and used to calculate the expected bias in distance modulus, which is then added onto the observational data.
+Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
+As these effects are not built into the likelihood, their influence on the error budget is not captured fully in the $\chi^2$ distribution.
 
@@
 By building the systematic uncertainties and selection effects into the simulation package, there is vastly more freedom in how to treat and model those effects.
-Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms which simply cannot be expressed in a tractable likelihood.
+Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood.
 This freedom comes with a cost -- computation.
@@
 This is an intensive process.
-Luckily, efficient sampling algorithms that have walkers which rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits, such as used in the BAMBIS framework {\red CITE RACHEL}.
+Luckily, efficient sampling algorithms that have walkers that rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits and can help ease the computational burden.
 
@@
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model which they apply to the light curve fitting for supernova.
-\citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters provide a model which offers increased flexibility with reduced uncertainty over the traditional method.
+\citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model that they apply to the light curve fitting for supernova.
+\citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method.
 The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
-\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant difference in $\alpha$ and $\beta$ values from the original as well.
+\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well.
 Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017}.
+The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
 
@@
 
-We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects.
+We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift-dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects.
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
@@
 At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function.
-The difficulty arises in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
+The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
 
-\subsection{Observed and Latent Population}
+\subsection{Observables Nature Provides: Populations}
 
@@
 Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.
+Our set of observables is therefore given as $\lbrace \hat{m_B}, \hat{x_1}, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 \end{equation}
-As we are focused on the spectroscopically confirmed DES supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$, and the similarly that the observed mass is the true mass $m$.
-Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.
+As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
+Similarly, we assume that the observed mass is the true mass $m$, such that  Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.
+
+\begin{figure}
+\begin{tikzpicture}
+% Obs nodes
+\node[obs]  (obssum) {$\hat{m_B}, \hat{x_1}, \hat{c}, \mathcal{C}^{-1}$};
+\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
+\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
+\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
+
+% Latent nodes
+\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
+\node[det, above=of zobs] (z) {$z$};
+\node[det, above=of mobs] (m) {$m$};
+\node[det, above=of z] (mu) {$\mu$};
+\node[latent, above=of m] (MBobs) {$M_B$};
+% Survey nodes
+\node[latent, above=of sum, yshift=1cm,text width=2.3cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,$\\$ \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha_c$};
+\node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
+% Global nodes
+\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
+\node[latent, above=1.7cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta,$\\$\delta(0), \delta(\infty)$};
+
+% Connect the nodes
+\edge {sum} {obssum};
+\edge {z} {zobs};
+\edge {m} {mobs};
+\edge {z,cosmo} {mu};
+\edge {mu, sum, m, pop1, ab, pobs, sf} {MBobs};
+
+% Plates
+\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(m)(mu)} {$\forall\ \rm{SN}$} ;
+\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
+\end{tikzpicture}
+\caption{Probabilistic graphical model for our likelihood.
+Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.}
+\label{fig:pgm}
+\end{figure}
 
@@
 
-We model the underlying population and intrinsic dispersion together as a redshift dependent multivariate skew normal for each survey.
+We model the underlying population and intrinsic dispersion together as a redshift-dependent multivariate skew normal for each survey.
 Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes.
+These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
 Both the colour and stretch means are modelled with normal priors.
 The correlation between the magnitude, stretch and colour populations is represented in a correlation matrix $\rho$, which is treated with an LKJ prior.
-Skewness is fixed to 0 for the absolute magnitude population, but left free for colour and stretch, with normal priors centered at 0 to draw the skewness to zero if not well constrained by the data.
-The decision to include skewness on the stretch distribution as well as the colour was motivated by \citet{Scolnic2016}, which found highly asymmetric underlying populations for both colour and stretch, regardless of survey or scatter model adopted.
+Skewness is fixed to 0 for the absolute magnitude and stretch, but left free for colour.
+We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$.
 
@@
 
-As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift dependent means, width and correlations fit individual for each survey.
+As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means, width and correlations fit individually for each survey.
 
-On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} (and therefore \citet{Kunz2007}) as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples which will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population.
-Proof of concept simulation fits show that an acceptable parameterisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-For the purposes of the DES spectroscopic sample which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
-We assume that supernova fall into either population as determined by their observed classification probability $\hat{p}$.
+On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population.
+Proof of concept simulation fits show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+For the purposes of the DES spectroscopic sample, which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
+We assume that supernovae fall into either population as determined by their observed classification probability $\hat{p}$.
 
@@
 {\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0).
-\citet{Shariff2016} instead of employing a colour-dependent $\beta$, model $\beta$ as redshift dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
+\citet{Shariff2016} instead of employ a colour-dependent $\beta$, model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
 
@@
 
-{\red Talk about how these arent even competitive with JLA and we really need to combine data. Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
+{\red Talk about how these arent even competitive with JLA and we really need to combine data? Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
 

### 2018-01-15 papers/methods/paper.tex
Paper updates

@@
 We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
-This degeneracy resulted in a significant bias in cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+This degeneracy resulted in a significant bias in recovered cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
 
@@
 It is now well known that host galaxy environment has a significant effect on supernova properties.
-The latest sample of over 1300 spectroscopically confirmed Type Ia supernova show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
+The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
 The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses.
@@
 That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty.
-We include these in our approach, in addition to including HST Calspec calibration uncertainty, 10 SALT2 model systematic uncertainties, three dust systematics (scale, dust model, and transfer law) and also the systematic peculiar velocity uncertainty.
-This gives fourteen global systematics shared by all surveys, plus two systematics per band in each survey.
+We include these in our approach, in addition to including HST Calspec calibration uncertainty, 10 SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty.
+This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
 

### 2018-01-16 papers/methods/paper.tex
Updating paper

@@
 \usepackage{tikz}
+\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
 \usetikzlibrary{bayesnet}
@@
 \newcommand {\etal} {\emph{~et~al.} }
-\newcommand{\cov}{\mathcal{C}^{-1}}
 \newcommand{\green}{\color{green}}
@@
 \newcommand{\kmsmpc}{{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
+\newcommand{\cov}{\mathcal{C}}
+\newcommand{\Z}{\mathcal{Z}}
 
@@
 \begin{equation}
-P(m_B, x_1, c|\hat{m_B},\hat{x_1},\hat{c},\cov) = \mathcal{N}(\lbrace m_B, x_1, c \rbrace|\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace, \cov).
+P(\hat{m_B},\hat{x_1},\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
 \end{equation}
@@
 The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
-The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses.
+The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses.
 The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
@@
 \end{equation}
-We also take flat priors on the parameterisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
+We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
 
-\subsubsection{Systematics}
+\subsubsection{Uncertainty Propagation}
 \label{sec:systreat}
-The chief difficulty with including systematics in supernova analyses is that the generally occur during the observational pipeline, and have difficult to model effects on the output observations.
-As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial Z_i}$, $\frac{\partial \hat{x_1}}{\partial Z_i}$, $\frac{\partial \hat{c}}{\partial Z_i}$, where $Z_i$ represents the $i$\textsuperscript{th} systematic.
+The chief difficulty with including systematic uncertainties in supernova analyses is that they generally occur during the observational pipeline, and have difficult-to-model effects on the output observations.
+As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial \Z_i}$, $\frac{\partial \hat{x_1}}{\partial \Z_i}$, $\frac{\partial \hat{c}}{\partial \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
 
 Assuming that the gradients can be linearly extrapolated -- which is a reasonable approximation for modern surveys with high quality control of systematics -- we can incorporate into our model a deviation from the observed original values by constructing a $(3 \times N_{\rm sys})$ matrix containing the numerical derivatives for the $N_{\rm sys}$ systematics and multiplying it with the row vector containing the offset for each systematic.
-But scaling the gradient matrix to represent the shift over $1\sigma$ of systematic uncertainty, we can simply enforce a unit normal prior on the systematic row vector to increase computational efficiency.
+By scaling the gradient matrix to represent the shift over $1\sigma$ of systematic uncertainty, we can simply enforce a unit normal prior on the systematic row vector to increase computational efficiency.
 
@@
 This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
+Denoting $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
+\begin{align}
+P(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C) = \mathcal{N}(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov). \label{eq:pop2}
+\end{align}
 
@@
 
+\subsection{Approximate Selection Effects}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=\textwidth]{simple_w_super_dist.pdf}
+	\end{center}
+	\caption{Marginalised distributions for 200 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
+The dashed green surfaces represent a fit to an underlying Gaussian colour population with the unshifted model.
+The blue solid surface represents fits to a skewed colour population with the unshifted model, and the purple dotted surface represents a fit to a skewed colour population with the shifted model.
+The superscript $0$ and $1$ denote the two different surveys (high-z and low-z respectively), and similarly the first four $\langle c^i \rangle$ parameters represent the four redshift nodes in the high-z survey, and the last four represent the nodes for the low-z survey.
+We can see that the shifted model is far better able to recover skewed input populations than the unshifted, performing better in terms of recovering skewness $\alpha_c$, mean colour $\langle c \rangle$ and width of the colour distribution $\sigma_c$.
+The unshifted model recovers the correct colour mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the colour population mean.
+Whilst the performance of the shifted model on skewed data populations is an improvement over the unshifted model, it comes at a cost - significant bias appears in the fits to $\alpha$, $\beta$, $\sigma_{m_B}$ and $w$.
+These biases in cosmological parameters are not found in the unshifted model, and similarly the bias when moving from a Gaussian to skewed colour population is negligible.
+For these reasons, we adopt the unshifted model.}
+	\label{fig:simple_w_super}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-01-16 dessn/configurations/simple_w_super.py

@@ -50,8 +50,9 @@ if __name__ == "__main__":
         res = fitter.load(squeeze=False)
 
         ls = []
-        cs = ["#086ed3", "lb", "r", "p", "p", "brown", "brown", "c", "c","o", "o", "e", "e", "g", "g", ]
+        cs = ["#086ed3", "lg", "p", "p", "p", "brown", "brown", "c", "c","o", "o", "e", "e", "g", "g", ]
         for i, (m, s, ci, chain, truth, weight, old_weight, posterior) in enumerate(res):
+            params = list(chain.keys())
             name_skew = "Gaussian" if s[0].alpha_c == 0 else "Skewed"
             name_approx = "unshifted" if m.frac_shift == 0 else "shifted"
             name = "%s population, %s  normal colour approximation" % (name_skew, name_approx)
@@ -60,10 +61,15 @@ if __name__ == "__main__":
             if name_skew == "Gaussian":
                 ls.append("--")
             else:
-                ls.append("-")
+                if name_approx == "shifted":
+                    ls.append(":")
+                else:
+                    ls.append("-")
             c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
-
-        c1.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, linestyles=ls, colors=cs)
+        print(params)
+        red_params = [p for p in params if "langle x_1" not in p and "Delta" not in p and "delta" not in p and "Omega" not in p]
+        c1.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, linestyles=ls,
+                     colors=cs, max_ticks=4, linewidths=1.1)
 
         # print("Adding individual realisations")
         # res = fitter.load(split_cosmo=True)
@@ -90,22 +96,28 @@ if __name__ == "__main__":
         # with open(pfn + "_cosmo_params.txt", 'w') as f:
         #     f.write(c1.analysis.get_latex_table(parameters=parameters))
         #
-        print("Plotting cosmology")
-        c1.plotter.plot(filename=[pfn + "_cosmo.png", pfn + "_cosmo.pdf"], truth=truth, parameters=parameters,
-                        figsize="column")
-
-        print("Plotting summaries")
-        c1.plotter.plot_summary(filename=[pfn + "_summary.png", pfn + "_summary.pdf"], truth=truth, parameters=parameters, errorbar=True,
[…]

### 2018-01-17 papers/methods/paper.tex
Paper updates

@@
 \mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
-r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \\
-\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B
+r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \label{eq:mu0}\\
+\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B \label{eq:mu1}
 \end{align}
@@
 Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^10$ solar masses.
 Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.
@@
 We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$.
-
 The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.
@@
 As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means, width and correlations fit individually for each survey.
+The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
+\begin{align}
+P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \rho, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
+\mathcal{N}^{\rm{skew}}(\lbrace M_B, x_1, c \rbrace | \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle \rbrace, \rho \cdot (\vec{\sigma} \vec{\sigma}^T), 
+\vec{\alpha}) \label{eq:l1},
+\end{align}
+where $\vec{\alpha} = \lbrace 0, 0, \alpha_c \rbrace$ and $\vec{\sigma} = \lbrace \sigma_{m_B}, \sigma_{x_1}, \sigma_c \rbrace$.
 
@@
 This degeneracy resulted in a significant bias in recovered cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+As such, our calculation of distance modulus $\mu$ mirrors that found in Equations \eqref{eq:mu0} and \eqref{eq:mu1}.
 
@@
 \end{equation}
-We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
+We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes
+\begin{align}
+M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot m. \label{eq:l2}
+\end{align}
 
@@
 \begin{align}
-P(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C) = \mathcal{N}(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov). \label{eq:pop2}
+P(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C) = \mathcal{N}(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov). \label{eq:l3}
 \end{align}
@@
 \end{align}
-As the model so far described in previous sections describes $P(D|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
+As the model so far described in previous sections describe components of a basic likelihood $P(D|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
 \begin{align}
@@
 As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC.
+Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC (a brief dismissal of months of struggle).
 We therefore simplify the integral and approximate the selection effects in apparent magnitude and redshift space independently, such that the denominator, denoted now $w$ for simplicity, is given as
@@
 \end{align}
-We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sample the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
+We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
+The latter assumption is investigated in Appendix \ref{app:approx} and verified to be true for cosmological purposes, if not for robust population determination.
 The population $P(m_B | z, \theta)$ thus becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
@@
 What then remains is determining the functional form of $P(S|m_B)$.
-For the treatment of the DES, we find that the error function which smoothly transitions from some constant efficiency down to zero is sufficient.
-This similar result has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014} {\red THESE CITATIONS EXACTLY MIRROR RUBIN2015, NOT SURE IF I SHOULD CHANGE THEM}.
-For our treatment of the low redshift surveys included {\red Introduce which lowz surveys earlier on} we simplify the model fitting and group all low redshift surveys together.
-This requires a more complicated selection function than the error function used for DES, however is reasonably well fit with a skew normal.
+For the treatment of most surveys, we find that the error function which smoothly transitions from some constant efficiency down to zero is sufficient.
+Formally, this gives
+\begin{align}
+P(S|m_B) = \Phi^C(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
+\end{align}
+where $\Phi^c$ the complimentary CDF and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
+The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
+However, for surveys which suffer from saturation and thus rejection of low-z supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a good analytic form, taking the form
+\begin{align}
+P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
+\end{align}
 
-The selection functions were fit to apparent magnitude efficiency ratios calculated from SNANA simulations by calculating an efficiency ratio as a function of apparent magnitude.
+The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude.
 Uncertainty of the Malmquist bias (entering both through statistical uncertainty from finite sized simulations in the efficiency ratio and the discrepancy between the analytic approximation and non-analytic simulation results) is incorporated into the fitting for the analytic approximation.
-Uncertainty is uniformly added to the efficiency ratio until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits.
-
-The DES survey is thus characterised by mean fit values $\mu_{\rm DES} = 22.3$, $\sigma_{\rm DES} = 0.7$, with the combined low redshift surveys characterised by $\mu_{\rm LowZ} = 14.1$, $\sigma_{\rm LowZ} = 1.36$, $\alpha_{\rm LowZ} = 3.8$.
+Uncertainty is uniformly added to the efficiency ratio until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
 
 With the well sampled approximation as specified previously, we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
-For the error function (DES) and skew normal selection (LowZ) functions respectively, this correction becomes
+For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), this correction becomes
 \begin{align}
-w_{\rm DES} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm DES}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm DES}^2}}  \right) \label{eq:seldes}\\
-w_{\rm LowZ} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm LowZ}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm LowZ}^2}}\right) \times \notag\\
-&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm LowZ})(m_B^* - \mu_{\rm LowZ})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}{\sigma^2_{\rm LowZ}} \sqrt{\frac{\sigma^2_{\rm LowZ}}{\alpha^2_{\rm LowZ}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm LowZ}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm LowZ}}} }\right), \label{eq:sellowz}
+w_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
+w_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
+&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right).
+\label{eq:sellowz}
 \end{align}
-where $\Phi$ represents the normal CDF function, and $\Phi^c$ the complimentary CDF.
 
@@
 In order to verify our model we run it through several tests.
-First, we validate on toy models, verifying that there is not significant cosmological bias in data sets more constraining than the DES spectroscopic sample.
-We then also generate
+First, we validate on toy models, verifying that there is not significant cosmological bias in highly constraining datasets.
+We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three year spectroscopic sample.
 
@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 200 realisations of supernova data with the $\Lambda$CDM model.
+Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
+	\label{fig:simple_ol}
+\end{figure}
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
+	\end{center}
+	\caption{Posterior surfaces for 200 realisations of supernova data with the Flat $w$CDM model.
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing incorrect statistics due to the highly non-Gaussian nature of the posterior surface.}
+	\label{fig:simple_w}
+\end{figure}
+
 We generate simple toy data to validate the basic premise of the model.
@@
 Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
+Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$.
+
 Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
 The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae.
-We draw from each survey simulation until we have 300 LowZ supernovae and 700 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
+We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
 
-The above represents one realisation of data.
-We test three models (Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM) with 300 realisations and find recovery of underlying cosmology without significant bias.
-Combined posterior surfaces of 22 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
+We test four models: Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM and Flat $w$CDM with a prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, with the latter included to allow sensitive tests on bias for $w$.
+To achieve statistical precision, we fit 200 realisations of supernovae datasets.
+Cosmological parameters are recovered without significant bias.
+Combined posterior surfaces of all 200 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
+By utilising the Stan framework, fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
 
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[width=\columnwidth]{approximate_simple_test_cosmo.pdf}
-%	\end{center}
-%	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $\Lambda$CDM model.}
-%	\label{fig:simple_om}
-%\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_simple_test_w_cosmo.pdf}
-	\end{center}
-	\caption{Posterior surfaces for 22 realisations of supernova data with the Flat $w$CDM model.}
-	\label{fig:simple_w}
-\end{figure}
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[width=\columnwidth]{approximate_simple_test_ol_cosmo.pdf}
-%	\end{center}
-%	\caption{Posterior surfaces for 22 realisations of supernova data with the $\Lambda$CDM model.}
-%	\label{fig:simple_ol}
-%\end{figure}
+To enforce investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$.
+This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated surfaces, and the results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
+
+\begin{table}
+	\centering
+	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 200 fits to independent realisations of toy supernova data.
+As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
+Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
+No significant bias is detected in either cosmological model.}
+	\label{tab:simple_model}
+	\begin{tabular}{ccc}
+		\hline
+		Model & $\Omega_m$ & $w$ \\ 
+		\hline
+		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
+		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 The second model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-We also test a `Combined' smear model, which is simply to draw equal samples from the {\celeven} and {\gten} models.
-It is against this combined sample that we wish to rigorously validate our model.
 
@@
 	\centering
-	\caption{Tested population distributions, where the {\sk} LowZ distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
+	\caption{Tested population distributions, where the {\sk} LowZ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
 	\label{tab:dist}
@@
 
-To forecast the results for the 3 year DES spectroscopic supernovae sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
+To forecast the results for the 3 year DES spectroscopic supernova sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
 We fit these through the three cosmological models - Flat $\Lambda$CDM, $\Lambda$CDM and $w$CDM, however focus mostly on the $w$CDM model as it is the primary scientific concern of the Dark Energy Survey.
@@
 \begin{align}
-w &= \idotsint \, d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta),
+w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta),
 \end{align}
@@
 \begin{align}
-m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
+m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle\label{eq:disc1} \\
 \sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2 \notag\\ 
-&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c}).
+&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c}). \label{eq:disc2}
 \end{align}
@@
 \subsection{Approximate Selection Effects}
+\label{app:approx}
+
+Equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our colour distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
+We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to match the actual mean and standard deviation of skew normal.
+With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
+\begin{align}
+\mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
+\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}}.
+\end{align}
+We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$.
+Unfortunately, this shift to the mean and standard deviation of the normal approximation did not produce stable posterior surfaces which Stan could fit.
+The surface instability was caused by severe degeneracy between three parameters: $\langle c \rangle$, $\sigma_c$ and $\alpha_c$.
+Shifting the mean colour was thus accomplished either by shifting the actual mean colour $\langle c \rangle$, by increasing skewness $\alpha_c$, or by modifying $\sigma_c$.
+Stable surfaces were only found when we fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
+Figure \ref{fig:shift} shows that even an different fixed value than the correct value still give a better approximation than the unshifted approximation.
+However, implementing the shifted approximation in our full model resulted in biased parameter recover for cosmological parameters, resulting in us ultimately discarding the shifted approximation.
+These biases are shown in Figure \ref{fig:simple_w_super}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{shift.pdf}
+	\end{center}
+	\caption{Testing the significance of fixing $\sigma_c$ when calculating the efficiency using a shifted normal approximation.
+The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$.
+The $x$-axis is analogous to $m_B$ is cosmological context.
+As expected, all efficiencies drop towards zero as shift increases (as objects get fainter).
+With a correct choice of $\sigma_c$, there is negligible error between the correct solution and the shifted normal approximation.
+For both under and over-estimating $\sigma_c$, the shifted normal approximation outperforms the unshifted normal approximation.
+Thus we implemented the shifted normal approximation into our model to test it in full: whilst $\sigma_c$ itself gets sampled.}
+	\label{fig:shift}
+\end{figure}
 
@@
 	\end{center}
-	\caption{Marginalised distributions for 200 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
+	\caption{Marginalised probability distributions for 200 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
 The dashed green surfaces represent a fit to an underlying Gaussian colour population with the unshifted model.

### 2018-01-17 dessn/configurations/simple.py

@@ -31,37 +31,28 @@ if __name__ == "__main__":
     if h != "smp-hk5pn72":  # The hostname of my laptop. Only will work for me, ha!
         fitter.fit(file)
     else:
-        parameters = [r"$\Omega_m$", "$w$", r"$\Omega_\Lambda$"]
+        parameters = [r"$\Omega_m$", "$w$"]
 
         from chainconsumer import ChainConsumer
 
         res = fitter.load(split_models=True, squeeze=False, split_cosmo=False)
 
         print("Adding data")
-        c1 = ChainConsumer()
-
+        cb = ChainConsumer()
+        names = [r"Flat $\Lambda$CDM", r"$\Lambda$CDM", "Flat $w$CDM", r"Flat $w$CDM + $\Omega_m$ prior"]
         for i, (m, s, ci, chain, truth, weight, old_weight, posterior) in enumerate(res):
-            c1.add_chain(chain, weights=weight, posterior=posterior)
-            print(truth["$w$"])
+            c = ChainConsumer()
+            c.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
+            if i in [0, 3]:
+                cb.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
+            c.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, sigmas=[0, 0.3, 1, 2], contour_labels="confidence")
+            ps = m.get_cosmo_params()
+            n = m.__class__.__name__
+            if m.prior:
+                n += "OmPrior"
+            c.plotter.plot(filename=[pfn + n + ".png", pfn + n + ".pdf"], parameters=ps, truth=truth, figsize="column")
 
-        c1.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False)
-
-        # print("Saving table")
-        # print(c1.analysis.get_latex_table(transpose=True))
-        # with open(pfn + "_cosmo_params.txt", 'w') as f:
-        #     f.write(c1.analysis.get_latex_table(parameters=parameters))
-
-        print("Plotting cosmology")
-        c1.plotter.plot(filename=[pfn + "_cosmo.png", pfn + "_cosmo.pdf"], parameters=parameters)
-        #
-        #
-        # print("Plotting distributions")
-        # c1.plotter.plot_distributions(filename=[pfn + "_dist.png", pfn + "_dist.pdf"], truth=truth, col_wrap=6)
-        #
-        # print("Saving Parameter values")
[…]

### 2018-01-17 dessn/framework/model.py

@@ -40,3 +40,7 @@ class Model(ABC):
         def init():
             return self.get_init(**kwargs)
         return init
+
+    @abstractmethod
+    def get_cosmo_params(self):
+        raise NotImplementedError()

### 2018-01-17 dessn/framework/models/approx_model.py

@@ -32,7 +32,7 @@ class ApproximateModel(Model):
         self.apply_efficiency = 1 if apply_efficiency else 0
         self.beta_contrib = beta_contrib
         self.kfactor = kfactor
-
+        self.prior = False
         self.fixed_sigma_c = fixed_sigma_c
 
     def get_parameters(self):
@@ -339,6 +339,9 @@ class ApproximateModel(Model):
         del dictionary["intrinsic_correlation"]
         return dictionary
 
+    def get_cosmo_params(self):
+        return [r"$\Omega_m$"]
+
 
 class ApproximateModelOl(ApproximateModel):
     def __init__(self, filename="approximate_ol.stan", num_nodes=4, systematics_scale=1.0, statonly=False,
@@ -346,6 +349,9 @@ class ApproximateModelOl(ApproximateModel):
         super().__init__(filename, num_nodes=num_nodes, systematics_scale=systematics_scale, statonly=statonly,
                          frac_alpha=frac_alpha, frac_shift=frac_shift, frac_shift2=frac_shift2,
                          fixed_sigma_c=fixed_sigma_c, beta_contrib=beta_contrib, kfactor=kfactor)
+
+    def get_cosmo_params(self):
+        return [r"$\Omega_m$", r"$\Omega_\Lambda$"]
 
 
 class ApproximateModelW(ApproximateModel):
@@ -356,5 +362,7 @@ class ApproximateModelW(ApproximateModel):
         super().__init__(filename, num_nodes=num_nodes, systematics_scale=systematics_scale, statonly=statonly,
                          frac_alpha=frac_alpha, frac_shift=frac_shift, frac_shift2=frac_shift2,
                          fixed_sigma_c=fixed_sigma_c, beta_contrib=beta_contrib, kfactor=kfactor)
+        self.prior = prior
 
-
+    def get_cosmo_params(self):
+        return [r"$\Omega_m$", r"$w$"]

### 2018-01-17 papers/methods/snippets/shift.py

@@ -1,0 +1,50 @@
+import numpy as np
+from scipy.stats import norm, skewnorm
+import matplotlib.pyplot as plt
+from scipy.integrate import simps
+
+sigma = 0.1
+xs = np.linspace(-5, 5, 1000)
+sf = 1 - norm.cdf(xs, 2, 0.5)
+mids = np.linspace(0.5, 3.5, 100)
+
+alpha = 2
+delta = alpha / np.sqrt(1 + alpha**2)
+sigma_cor = [0.05, 0.075, 0.1, 0.15]
+shift = np.sqrt(2 / np.pi) * delta * sigma
+sigma_ratio1 = np.sqrt(1 - 2 * delta**2 / np.pi)
+
+ef_a = np.zeros(mids.shape)
+ef_no = np.zeros(mids.shape)
+efs = np.zeros((mids.size, len(sigma_cor)))
+
+for i, mid in enumerate(mids):
+    pop_actual = skewnorm.pdf(xs, alpha, mid, sigma)    
+    pop_approx1 = norm.pdf(xs, mid, sigma)
+    
+    ef_a[i] = simps(pop_actual * sf, x=xs)
+    ef_no[i] = simps(pop_approx1 * sf, x=xs)
+    
+    for j, s in enumerate(sigma_cor):
+        pop = norm.pdf(xs, mid + (shift * s / sigma), sigma * sigma_ratio1)
+        efs[i, j] = simps(pop * sf, x=xs)
+
+fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(5,6), gridspec_kw={"hspace":0.0})
+axes[0].plot(mids, ef_a, label="Correct", c="k")
+axes[0].plot(mids, ef_no, label="Unshifted")
+
+
+axes[1].axhline(0, c="k")
+axes[1].plot(mids, ef_no - ef_a, label="Unshifted")
+for row, c in zip(efs.T, sigma_cor):
+    axes[0].plot(mids, row, label="Shifted, $\sigma_c = %0.2f$" % c, ls="--")
+    axes[1].plot(mids, row - ef_a, label="Shifted, $\sigma_c = %0.2f$" % c, ls="--")
+
+axes[1].set_xlabel("Shift")
+axes[0].set_ylabel("Efficiency")
+axes[1].set_ylabel("$\Delta$ Efficiency")
+axes[0].legend(frameon=False, markerfirst=False)
+
+fig.tight_layout()
[…]

### 2018-01-17 dessn/snana/convert_snana_data.py
Updating some des and lowz data

@@ -284,11 +284,11 @@ if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG, format="[%(funcName)20s()] %(message)s")
     # convert("DES3YR_LOWZ_COMBINED_FITS")
     # convert("DES3YR_DES_COMBINED_FITS")
-    # convert("DES3YR_DES_NOMINAL")
-    # convert("DES3YR_LOWZ_NOMINAL")
+    convert("DES3YR_DES_NOMINAL")
+    convert("DES3YR_LOWZ_NOMINAL")
     # convert("DES3YR_DES_BULK", skip=6)
     # convert("DES3YR_LOWZ_BULK", skip=6)
-    convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11)
+    # convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11)
     # convert("DES3YR_LOWZ_BHMEFF", load_dump=True, skip=11)
     # convert("DES3YR_LOWZ_VALIDATION", skip=6)
     # convert("DES3YR_DES_VALIDATION", skip=6)

### 2018-01-17 dessn/snana/systematic_names.py

@@ -18,6 +18,7 @@ def get_systematic_mapping():
         "Fitz99ColorLaw": r"{\rm MWE ColorLaw}",
         "OD94ColorLaw": r"{\rm MWE ColorLaw}",
         "REDSHIFT_BIAS": r"{\rm \delta z }",
+        "zshift": r"{\rm \delta z }",
         "DES_ZP_g_survcal": r"\Delta g",
         "DES_ZP_r_survcal": r"\Delta r",
         "DES_ZP_i_survcal": r"\Delta i",

### 2018-01-17 dessn/configurations/nominal.py
Adding population

@@ -51,7 +51,8 @@ if __name__ == "__main__":
         from chainconsumer import ChainConsumer
         res = fitter.load()
         c = ChainConsumer()
-        for m, s, chain, truth, weight, old_weight, posterior in res:
+
+        for m, s, ci, chain, truth, weight, old_weight, posterior in res:
             name = "Stat + Syst" if not m.statonly else "Stat"
             c.add_chain(chain, weights=weight, posterior=posterior, name=name)
 
@@ -67,7 +68,7 @@ if __name__ == "__main__":
               r"$\sigma_{\rm m_B}^{0}$", r"$\sigma_{\rm m_B}^{1}$",]
         for p in ps:
             mus, stds = [], []
-            for m, s, chain, truth, weight, old_weight, posterior in res:
+            for m, s, ci, chain, truth, weight, old_weight, posterior in res:
                 w = chain[p]
                 mus.append(np.mean(w))
                 stds.append(np.std(w))

### 2018-01-17 dessn/configurations/plot_pop_simple.py

@@ -1,0 +1,41 @@
+import os
+
+from dessn.framework.simulations.simple import SimpleSimulation
+from astropy.cosmology import FlatwCDM
+import matplotlib.pyplot as plt
+from matplotlib import rc
+
+if __name__ == "__main__":
+
+    plot_dir = os.path.dirname(os.path.abspath(__file__)) + "/plots/%s/" % os.path.basename(__file__)[:-3]
+    pfn = plot_dir + os.path.basename(__file__)[:-3]
+    if not os.path.exists(plot_dir):
+        os.makedirs(plot_dir)
+
+    sims = [SimpleSimulation(1000), SimpleSimulation(1000, lowz=True)]
+    names = ["High-redshift", "Low-redshift"]
+
+    cosmo = FlatwCDM(70, 0.3)
+    rc('text', usetex=True)
+    fig, axes = plt.subplots(nrows=len(sims), figsize=(4, 5))
+
+    for ax, sim, n in zip(axes, sims, names):
+        data = sim.get_all_supernova(1000)
+        obs_mBx1c = data["obs_mBx1c"]
+        mbs, x1s, cs = obs_mBx1c[:, 0], obs_mBx1c[:, 1], obs_mBx1c[:, 2]
+        mbs, x1s, cs = data["sim_apparents"], data["sim_stretches"], data["sim_colours"]
+        zs = data["redshifts"]
+        passed = data["passed"]
+        mus = cosmo.distmod(zs).value
+        MBs = mbs - mus
+        ax.scatter(zs[passed], MBs[passed], c=cs[passed], s=10, vmin=-0.2, vmax=0.4, alpha=1, cmap="plasma")
+
+        ax.set_xlabel("$z$", fontsize=14)
+        ax.set_ylabel(r"$m_B - \mu$", fontsize=14)
+        ax.text(0.98, 0.95, n, verticalalignment='top', horizontalalignment='right', transform=ax.transAxes)
+        ax.set_xlim(zs.min() - 0.001, zs.max() + 0.001)
+    plt.locator_params(numticks=5)
+    fig.tight_layout()
+
+    fig.savefig(pfn + ".png", bbox_inches="tight", transparent=True)
+    fig.savefig(pfn + ".pdf", bbox_inches="tight", transparent=True)

### 2018-01-17 papers/methods/paper.tex
Updating some des and lowz data
Adding population
Paper update: population caption

@@
 	\begin{center}
+		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
+	\end{center}
+	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys.
+Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.}
+	\label{fig:simple_pop}
+\end{figure}
+\begin{figure}
+	\begin{center}
 		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
@@
 We generate simple toy data to validate the basic premise of the model.
-For both DES and LowZ data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
+For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
 Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
 Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$.
+Redshifts are drawn from a power law to increase the number of events as redshift increases.
 
 Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
+The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
 The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae.
 We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
+Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, showing presence of strong selection effects in both toy surveys.
 
@@
 	\label{tab:simple_model}
-	\begin{tabular}{ccc}
+	\begin{tabular}{l|cc}
 		\hline
@@
 	\label{tab:dist}
-	\begin{tabular}{l|llll}
+	\begin{tabular}{l|cccc}
 		\hline
@@
 	\label{tab:bulk_summary}
-	\begin{tabular}{ccccc}
+	\begin{tabular}{lcccc}
 		\hline

### 2018-01-18 papers/methods/paper.tex
Adding selection functions to paper

@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{DES3YR_DES_BHMEFF_AMG10.pdf}
+		\includegraphics[width=\columnwidth]{DES3YR_LOWZ_BHMEFF_G10.pdf}
+	\end{center}
+	\caption{Fitting the selection function for both the DES 3YR spectroscopically confirmed supernova sample and the combined low-redshift sample.
+Blue errorbars represent the efficiency calculated by determining the ratio of discovered to generated supernovae in apparent magnitude bins for SNANA simulations.
+The black line represents the best fit analytic function for each sample, and the light grey lines surrounding the best fit value represent random realisations of analytic function taking into account uncertainty on the best fit value.}
+	\label{fig:sf_snana}
+\end{figure}
+
 Each realisation of cosmology fitted contains 300 LowZ supernovae, and 250 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.

### 2018-01-18 dessn/framework/simulations/selection_effects.py

@@ -1,6 +1,5 @@
 import numpy as np
 from scipy.ndimage import gaussian_filter1d
-import matplotlib.pyplot as plt
 from scipy.stats import norm, skewnorm
 from scipy.optimize import curve_fit
 import os
@@ -82,16 +81,15 @@ def get_selection_effects_cdf(dump_npy, plot=False, cut_mag=20):
         vals, cov, *_ = result
         chi2 = np.sum(((ratio - cdf(binc, *vals)) / ratio_error_adj) ** 2)
         red_chi2 = chi2 / (len(binc) - 3)
-        # print(red_chi2, adj)
-
-    print(vals)
-    print(cov)
-    print(np.sqrt(np.diag(cov)))
 
     if plot:
-        fig, ax = plt.subplots(1, 1)
-        ax.errorbar(binc, ratio, yerr=ratio_error_adj, ls="none", fmt='o', ms=3, label="Bin(err)")
-        ax.errorbar(binc, ratio_smooth, yerr=ratio_smooth_error, label="Bin(err) (smoothed)")
+        import matplotlib.pyplot as plt
+        from matplotlib import rc
+        rc('text', usetex=True)
+
+        fig, ax = plt.subplots(figsize=(4, 3))
+        ax.errorbar(binc, ratio, yerr=ratio_error_adj, ls="none", fmt='o', ms=3, label="Simulation")
+        #ax.errorbar(binc, ratio_smooth, yerr=ratio_smooth_error, label="Bin(err) (smoothed)")
 
         mbs = np.linspace(binc[0], binc[-1], 1000)
         cdf_eval = cdf(mbs, *vals)
@@ -100,16 +98,18 @@ def get_selection_effects_cdf(dump_npy, plot=False, cut_mag=20):
             rands = np.random.multivariate_normal([0, 0, 0, 0], cov=cov)
             vals2 = vals + rands
             cdf_eval2 = cdf(mbs, *vals2)
-            opts = {}
-            if i == 0:
-                opts["label"] = "uncert"
-            plt.plot(mbs, cdf_eval2, c='k', alpha=0.05, **opts)
-        plt.ylabel("Probability")
-        plt.xlabel(r"$m_B + \alpha x_1 + \beta c$")
-        plt.legend()
-        ax.plot(mbs, cdf_eval, c='k', label="Fitting efficiency")
-        plt.title(os.path.basename(dump_npy))
-        plt.show()
[…]

### 2018-01-18 papers/methods/snippets/shift.py

@@ -2,11 +2,13 @@ import numpy as np
 from scipy.stats import norm, skewnorm
 import matplotlib.pyplot as plt
 from scipy.integrate import simps
+from matplotlib import rc
+
 
 sigma = 0.1
 xs = np.linspace(-5, 5, 1000)
 sf = 1 - norm.cdf(xs, 2, 0.5)
-mids = np.linspace(0.5, 3.5, 100)
+mids = np.linspace(0.5, 4, 100)
 
 alpha = 2
 delta = alpha / np.sqrt(1 + alpha**2)
@@ -29,7 +31,8 @@ for i, mid in enumerate(mids):
         pop = norm.pdf(xs, mid + (shift * s / sigma), sigma * sigma_ratio1)
         efs[i, j] = simps(pop * sf, x=xs)
 
-fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(5,6), gridspec_kw={"hspace":0.0})
+rc('text', usetex=True)
+fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(4, 5), gridspec_kw={"hspace": 0.0})
 axes[0].plot(mids, ef_a, label="Correct", c="k")
 axes[0].plot(mids, ef_no, label="Unshifted")
 

### 2018-02-07 dessn/framework/simulations/simple.py
Better fix

@@ -124,13 +124,13 @@ class SimpleSimulation(Simulation):
                 mb = MB + mu - alpha * x1 + beta * c - mass_correction * p
                 vector = np.array([mb, x1, c])
                 # Add intrinsic scatter to the mix
-                diag = np.array([0.04**2, 0.2**2, 0.03**2 + (self.kappa0 + self.kappa1 * red)**2])
-                diag2 = np.array([0.04**2, 0.2**2, 0.03**2])
+                diag = np.array([0.04, 0.2, 0.03])**2
                 cov = np.diag(diag)
-                cov2 = np.diag(diag)
+                obs_mBx1c_cov.append(cov)
                 sim_mBx1c.append(vector)
-                vector += np.random.multivariate_normal([0, 0, 0], cov)
-                obs_mBx1c_cov.append(cov2)
+                cov2 = cov.copy()
+                cov2[2, 2] += (self.kappa0 + self.kappa1 * red)**2
+                vector += np.random.multivariate_normal([0, 0, 0], cov2)
                 obs_mBx1c.append(vector)
                 deta_dcalib.append(np.zeros((3, self.num_calib)))
             redshifts_all += redshifts.tolist()

### 2018-02-07 papers/methods/paper.tex
Slight paper update

@@
 
-In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015}.
+In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects whilst improving computational performance.
 Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
@@
 \end{equation}
-where $C^{-1}$ is an uncertainty matrix that combined the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
+where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
 The benefit this analysis methodology provides is speed - for samples of hundreds of supernovae, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
@@
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
-As these effects are not built into the likelihood, their influence on the error budget is not captured fully in the $\chi^2$ distribution.
+As these effects are not built into the likelihood, their influence on the error budget is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
 
@@
 The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
+However, whilst there are a large amount of hierarchical models available, none of them have undergone high-statistics simulation verification to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own.
 
-The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying populations, rates, redshift distributions, mass corrections and redshift evolution, each of will will be discussed further in the outline of our model below.
+The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae type rates and redshift distributions, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
 

### 2018-02-20 papers/methods/paper.tex
Paper update

@@
 The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$.
-This can be thought of as the `true' values for the supernova, and is normally distributed around the observed values such that
+This can be thought of as the `true' values for the supernova, and with correct summary statistics and uncertainty would be normally distributed around the observed values such that
 \begin{equation}
@@
 \end{equation}
+However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics report from light curve fitting are not Gaussian.
+We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation.
+As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence.
+As this may represent underlying physics more than the effects of Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
+We thus add $(\kappa_0 + \kappa_1 z)^2$ to our observed colour uncertainty, where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
+Each survey has its own set of $\kappa$ parameters.
+Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the bias in the mean colour measured as a function of redshift, and subtract this bias from our observed data.
+
 As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Similarly, we assume that the observed mass is the true mass $m$, such that  Whilst the latter is not physically motivated like the redshift measurement, the relatively small contribution of the host galaxy mass to the model makes this an acceptable approximation.
+Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable.
 
@@
 % Survey nodes
-\node[latent, above=of sum, yshift=1cm,text width=2.3cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,$\\$ \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha_c$};
+\node[latent, above=of sum, yshift=1cm,text width=2.3cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
 \node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
@@
 
-We model the underlying population and intrinsic dispersion together as a redshift-dependent multivariate skew normal for each survey.
+We follow prior work and model the colour population as an independent redshift-dependent skew normal for each survey.
+For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal.
 Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes.
@@
 Both the colour and stretch means are modelled with normal priors.
-The correlation between the magnitude, stretch and colour populations is represented in a correlation matrix $\rho$, which is treated with an LKJ prior.
-Skewness is fixed to 0 for the absolute magnitude and stretch, but left free for colour.
+Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
 We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$.
@@
 
-As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means, width and correlations fit individually for each survey.
+As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means and width fit individually for each survey.
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
@@
 P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \rho, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
-\mathcal{N}^{\rm{skew}}(\lbrace M_B, x_1, c \rbrace | \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle \rbrace, \rho \cdot (\vec{\sigma} \vec{\sigma}^T), 
-\vec{\alpha}) \label{eq:l1},
+\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
 \end{align}
-where $\vec{\alpha} = \lbrace 0, 0, \alpha_c \rbrace$ and $\vec{\sigma} = \lbrace \sigma_{m_B}, \sigma_{x_1}, \sigma_c \rbrace$.
 
@@
 \begin{align}
-P(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C) = \mathcal{N}(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov). \label{eq:l3}
+P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right).
+\label{eq:l3}
 \end{align}
@@
 We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
-The latter assumption is investigated in Appendix \ref{app:approx} and verified to be true for cosmological purposes, if not for robust population determination.
-The population $P(m_B | z, \theta)$ thus becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
+It was found that discarding the skewness entirely resulted in highly biased population recover, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal, with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
+This shifted Gaussian approximation removes the unintended bias when simply discarding skewness.
+More detail on this shift can be found in Appendix \ref{app:approx}.
+The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
 \begin{align}
 m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
-\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2 \notag\\ 
-&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c})
+\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2
 \end{align}
@@
 	\end{center}
-	\caption{Posterior surfaces for 200 realisations of supernova data with the $\Lambda$CDM model.
+	\caption{Posterior surfaces for 100 realisations of supernova data with the $\Lambda$CDM model.
 Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
@@
 	\end{center}
-	\caption{Posterior surfaces for 200 realisations of supernova data with the Flat $w$CDM model.
+	\caption{Posterior surfaces for 100 realisations of supernova data with the Flat $w$CDM model.
 The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing incorrect statistics due to the highly non-Gaussian nature of the posterior surface.}
@@
 Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
+Extra dispersion of $0.02(1+3z)$ is added in quadrature to the $c$ uncertainty to simulation unexplained colour dispersion from chromatic variation in supernovae.
 The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
@@
 We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
-Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, showing presence of strong selection effects in both toy surveys.
+Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys.
 
 We test four models: Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM and Flat $w$CDM with a prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, with the latter included to allow sensitive tests on bias for $w$.
-To achieve statistical precision, we fit 200 realisations of supernovae datasets.
+To achieve statistical precision, we fit 100 realisations of supernovae datasets.
 Cosmological parameters are recovered without significant bias.
-Combined posterior surfaces of all 200 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
-By utilising the Stan framework, fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
+Combined posterior surfaces of all 100 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
+By utilising the Stan framework and several efficient parametrisations (discussed further in Appendix \ref{app:optimisations}), fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
 
 To enforce investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$.
-This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated surfaces, and the results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
+This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces, and the results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
 
@@
 	\centering
-	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 200 fits to independent realisations of toy supernova data.
+	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 100 fits to independent realisations of toy supernova data.
 As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
 Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
+As our surfaces are stacked, a shift away from the simulated values of more than $\sigma/\sqrt{100} = 0.1\sigma$ would give a $1\sigma$ detection of bias.
 No significant bias is detected in either cosmological model.}
@@
 In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and now also include our full treatment of systematics.
-Additionally, we also simulate two different underlying population -- a Gaussian distribution in colour and stretch, and skewed colour and stretch populations using population values from \citet[][hereafter {\sk}]{Scolnic2016}.
+Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016}.
+The selection effects observed by comparing the generated supernovae to those that pass our simulated cuts and were successfully `observed' are shown in Figure~\ref{fig:sf_snana}, and it is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
+We note that there was no significant difference between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.
 
@@
 		\hline
-		Gaussian LowZ &  $0.0$ & $1.0$ & $0.0$ & $0.1$ \\
-		Gaussian DES  &  $0.0$ & $1.0$ & $0.0$ & $0.1$  \\
 		{\sk} LowZ    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
@@
 
-Each realisation of cosmology fitted contains 300 LowZ supernovae, and 250 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-Combined posterior surfaces for 150 data realisations are shown in Figure \ref{fig:bulk_summary}.
-Note that we have combined the posteriors for 150 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{150} \approx 12$ times less than a single realisation.
+Each realisation of cosmology fitted contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+{\red Focus on w}.
+
+ Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_summary}.
+Note that we have combined the posteriors for 100 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{100} = 10$ times less than a single realisation.
 Plot summaries for the fits are shown in Figure \ref{fig:bulk_summary}, and the parameter bounds are listed in Table \ref{tab:bulk_summary}.
@@
 Cosmological bias is detected with $\approx 3.0\sigma$ significance for the {\celeven} model and $\approx 1.7\sigma$ significance for the {\gten} model, indicating that the population and selection effect treatment cannot capture all necessary information to encapsulate the magnitude and chromatic smearing of the supernovae population.
-However, for the sample size of the DES and LowZ supernova samples (of order 600 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
-
-\begin{table}
-	\centering
-	\caption{Parameter summaries shown for 150 realisations of supernovae data, each comprising of 300 LowZ supernovae and 250 DES-like supernovae.
-Maximum likelihood statistics, with errors quoted to $1\sigma$ (68\% confidence) are reported.}
-	\label{tab:bulk_summary}
-	\begin{tabular}{lcccc}
-		\hline
-		Model & $\Omega_m$ & $\alpha$ & $\beta$  \\ 
-		\hline
-		C11 Gaussian & $0.316^{+0.062}_{-0.064}$ & $0.143^{+0.028}_{-0.027}$ & $2.97^{+0.20}_{-0.19}$  \\ 
-		C11 {\sk} & $0.316^{+0.063}_{-0.062}$ & $0.155\pm 0.024$ & $2.77^{+0.26}_{-0.27}$  \\ 
-		Combined Gaussian & $0.299^{+0.067}_{-0.060}$ & $0.144^{+0.030}_{-0.028}$ & $3.11^{+0.21}_{-0.22}$  \\ 
-		Combined {\sk} & $0.297^{+0.075}_{-0.064}$ & $0.150^{+0.026}_{-0.025}$ & $3.00^{+0.29}_{-0.31}$  \\ 
-		G10 Gaussian & $0.290^{+0.070}_{-0.060}$ & $0.145\pm 0.030$ & $3.27\pm 0.23$ \\ 
-		G10 {\sk} & $0.290^{+0.079}_{-0.073}$ & $0.147^{+0.026}_{-0.028}$ & $3.26^{+0.34}_{-0.33}$ \\ 
-		\hline
-	\end{tabular}
-\end{table}
+However, for the sample size of the DES and LowZ supernova samples (of order 350 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
 
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_bulk_load_summary.pdf}
+		\includegraphics[width=\columnwidth]{bulk2.pdf}
 	\end{center}
-	\caption{Graphical parameter summaries for 150 realisations of supernova data, following Table \ref{tab:bulk_summary}.}
+	\caption{Graphical parameter summaries for 100 realisations of supernova data.}
 	\label{fig:bulk_summary}
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{approximate_bulk_load.pdf}
+		\includegraphics[width=\columnwidth]{bulk_small2.pdf}
 	\end{center}
-	\caption{Posterior surfaces for 150 realisations of supernova data.
+	\caption{Posterior surfaces for 100 realisations of supernova data.
 Neither population nor scatter model has a significant effect on posterior surface location or shape.
@@
 
+%\section{DES Forecasts}
+%\label{sec:forecasts}
 %
-%\section{Systematics Strength Test}
-%\label{sec:sys}
+%To forecast the results for the 3 year DES spectroscopic supernova sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
+We fit these through the three cosmological models - Flat $\Lambda$CDM, $\Lambda$CDM and $w$CDM, however focus mostly on the $w$CDM model as it is the primary scientific concern of the Dark Energy Survey.
+We run all three cosmological models with systematics enabled and disabled to explicitly show the uncertainty included from systematic contribution.
 %
-%\subsection{Singular Permutations}
+%{\red How to include flat lambdacdm - the plot is boring, just quote the results, or just quote uncertainty, should I round the uncertainty to make it symmetric?
+Cant do that for wCDM or LambdaCDM because theyre definitely not gaussian}
 %
-%To test the impact of various systematics, we generate an array of simulations in which we simulate a dataset with one of the systematic variables permuted.
-We inspect the effects of systematic shifts using two cosmological models - Flat $\Lambda$CDM and Flat $w$CDM.
-To make the effects of the systematics more easily seen, we increase our simulated sample size to 500 LowZ and 500 DES-like supernovae, and enforce a prior on $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ on the Flat $w$CDM model.
-In these test we wish to inspect the shift in statistical uncertainty with simulated systematic effects, and so compare first to a series of simulations with no simulated systematic changes, and also disable our treatment of systematics as described in Section \ref{sec:systreat} for these tests.
-Our tests are broken down into groups, where systematics of different nature are treated separately.
-In each simulation, only one systemic is permuted.
-For a baseleine, our first set has no systematic shift, and so the uncertainty in the parameter summaries is given by statistical uncertainty only.
-We then test several variations:
-%\begin{enumerate}
-%	\item 0.02 Gaussian error on the zero point offsets in a randomly selected band.
-%	\item $20 \angstrom $ Gaussian error on the filter wavelength.
-%	\item 10\% Gaussian error in the bias corrected flux measurement (with correct flux uncertainty).
-%	\item 10\% Gaussian error in the bias corrected flux measurement, but with incorrect flux error.
-%	\item 20\% Gaussian error on the Milky Way E(B-V) dust correction scale.
-%	\item 0.2 Gaussian error on the value of $R_V$ for calculating the extinction curve.
-%\end{enumerate}
+%{\red Talk about how these arent even competitive with JLA and we really need to combine data? Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
 %
@@
 %	\begin{center}
-%		\includegraphics[width=\columnwidth]{approximate_systematic_load_summary.pdf}
+%		\includegraphics[width=\columnwidth]{combined_ApproximateModelW.pdf}
 %	\end{center}
-%	\caption{Investigating the impact of $\Omega_m$ by systematic permutations under Flat $\Lambda$CDM cosmology, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
-%	\label{fig:sys_om}
+%	\caption{w.}
+%	\label{fig:combined_w}
 %\end{figure}
@@
 %	\begin{center}
-%		\includegraphics[width=\columnwidth]{approximate_systematic_load_w_summary.pdf}
+%		\includegraphics[width=\columnwidth]{combined_ApproximateModelOl.pdf}
 %	\end{center}
-%	\caption{Investigating the impact of $w$ by systematic permutations under $w$CDM cosmology with the prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, validated using 200 realisations of data, with each dataset containing 500 LowZ supernovae and 500 DES-like supernovae.}
-%	\label{fig:sys_w}
+%	\caption{ole.}
+%	\label{fig:combined_ol}
 %\end{figure}
-%
-%We find that the most significant systematic are the zero points, as the 0.02 Gaussian error given to them resulted in significant changes to the shape of the marginalised posterior surface for both $\Omega_m$ and $w$, shifting them both to lower values and increasing uncertainty.
-%
-%{\red What are the actual size of the uncertainties of the systematics? Ie are zps really that badly known? Once I know, I can talk about if any of these are significant.}
-%
-%\subsection{Full Permutations}
-%
-%{\red When I have the final sims from Rick I can flesh this out}
-
-\section{DES Forecasts}
-\label{sec:forecasts}
-
-To forecast the results for the 3 year DES spectroscopic supernova sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
-We fit these through the three cosmological models - Flat $\Lambda$CDM, $\Lambda$CDM and $w$CDM, however focus mostly on the $w$CDM model as it is the primary scientific concern of the Dark Energy Survey.
-We run all three cosmological models with systematics enabled and disabled to explicitly show the uncertainty included from systematic contribution.
-
-{\red How to include flat lambdacdm - the plot is boring, just quote the results, or just quote uncertainty, should I round the uncertainty to make it symmetric?
-Cant do that for wCDM or LambdaCDM because theyre definitely not gaussian}
-
-{\red Talk about how these arent even competitive with JLA and we really need to combine data? Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{combined_ApproximateModelW.pdf}
-	\end{center}
-	\caption{w.}
-	\label{fig:combined_w}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{combined_ApproximateModelOl.pdf}
-	\end{center}
-	\caption{ole.}
-	\label{fig:combined_ol}
-\end{figure}
 
@@
 
+\section{Numerical Optimisations}
+\label{app:optimisations}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-02-20 dessn/configurations/bulk.py

@@ -19,22 +19,15 @@ if __name__ == "__main__":
         os.makedirs(dir_name)
 
     models = [
-        # ApproximateModelW(prior=True, statonly=False, frac_shift=1.0),
+        ApproximateModelW(prior=True, statonly=False, frac_shift=1.0),
         ApproximateModelW(prior=True, statonly=True, frac_shift=1.0)
     ]
 
     ndes = 204
     nlowz = 137
-    bcor = False
     simulations = [
-        # [SNANASimulation(ndes, "DES3YR_DES_SAMTEST_MAGSMEAR", use_sim=use_sim), SNANASimulation(nlowz, "DES3YR_LOWZ_SAMTEST_MAGSMEAR", use_sim=use_sim)],
-        # [SNANASimulation(ndes, "DES3YR_DES_BULK_G10_SKEW", use_sim=use_sim, bias_cor=False), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_G10_SKEW", use_sim=use_sim, bias_cor=False)],
-        # [SNANASimulation(ndes, "DES3YR_DES_BULK_C11_SKEW", use_sim=use_sim, bias_cor=False), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_C11_SKEW", use_sim=use_sim, bias_cor=False)],
-        # [SNANASimulation(ndes, "DES3YR_DES_BULK_G10_SKEW", bias_cor=bcor), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_G10_SKEW", bias_cor=bcor)],
-        # [SNANASimulation(ndes, "DES3YR_DES_BULK_G10_SKEW", bias_cor=bcor), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_G10_SKEW", bias_cor=bcor)],
-        [SNANASimulation(ndes, "DES3YR_DES_BULK_C11_SKEW", bias_cor=bcor, zlim=0.9), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_C11_SKEW", bias_cor=bcor, zlim=0.9)],
-        [SNANASimulation(ndes, "DES3YR_DES_BULK_C11_SKEW", bias_cor=bcor, zlim=0.8), SNANASimulation(nlowz, "DES3YR_LOWZ_BULK_C11_SKEW", bias_cor=bcor, zlim=0.8)],
[…]

### 2018-02-20 dessn/configurations/simple.py

@@ -23,7 +23,7 @@ if __name__ == "__main__":
     fitter = Fitter(dir_name)
     fitter.set_models(*models)
     fitter.set_simulations(simulation)
-    fitter.set_num_cosmologies(200)
+    fitter.set_num_cosmologies(100)
     fitter.set_num_walkers(1)
     fitter.set_max_steps(3000)
 

### 2018-02-22 dessn/configurations/simple_w_shift.py
More plots

@@ -68,19 +68,30 @@ if __name__ == "__main__":
             c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
         print(params)
         red_params = [p for p in params if "langle x_1" not in p and "Delta" not in p and "delta" not in p and "Omega" not in p]
+
+        red_params = ['$w$', '$\\sigma_{c}^{0}$', '$\\sigma_{c}^{1}$', '$\\alpha_c^{0}$',
+                      '$\\alpha_c^{1}$', '$\\langle c^{0} \\rangle$', '$\\langle c^{1} \\rangle$',
+                      '$\\langle c^{2} \\rangle$', '$\\langle c^{3} \\rangle$', '$\\langle c^{4} \\rangle$',
+                      '$\\langle c^{5} \\rangle$', '$\\langle c^{6} \\rangle$', '$\\langle c^{7} \\rangle$']
+        red_params2 = ['$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$', '$\\sigma_{\\rm m_B}^{0}$',
+                       '$\\sigma_{\\rm m_B}^{1}$', '$\\sigma_{x_1}^{0}$', '$\\sigma_{x_1}^{1}$',
+                       '$\\kappa_{c0}^{0}$', '$\\kappa_{c0}^{1}$', '$\\kappa_{c1}^{0}$',
+                       '$\\kappa_{c1}^{1}$', '$s_c^{0}$']
         c1.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, linestyles=ls,
                      colors=cs, max_ticks=4, linewidths=1.1)
 
         c1.plotter.plot(filename=pfn + "_small.png", truth=truth, parameters=2)
 
         print("Plotting distributions")
-        fig = c1.plotter.plot_distributions(truth=truth, col_wrap=5, parameters=red_params)
-        ax = fig.get_axes()
-        ax[10].set_ylim(0, 10)
-        ax[11].set_ylim(0, 10)
-        filenames = [pfn + "_dist.png", pfn + "_dist.pdf"]
-        for f in filenames:
-            c1.plotter._save_fig(fig, f, 300)
+
+        for i, p in enumerate([red_params, red_params + red_params2]):
+            fig = c1.plotter.plot_distributions(truth=truth, col_wrap=5, parameters=p)
+            ax = fig.get_axes()
+            ax[3].set_ylim(0, 5)
+            ax[4].set_ylim(0, 5)
[…]

### 2018-02-22 papers/methods/paper.tex
Paper updates

@@
 As this may represent underlying physics more than the effects of Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $(\kappa_0 + \kappa_1 z)^2$ to our observed colour uncertainty, where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
 Each survey has its own set of $\kappa$ parameters.
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the bias in the mean colour measured as a function of redshift, and subtract this bias from our observed data.
+We allow the amount of bias to be parametrised to account for different intrinsic scatter models, where parameter $\kappa_2$ controlled the fraction of bias subtraction - 0 for no bias, up to 1 for the maximum bias found in simulations (which corresponded to the intrinsic scatter model from \citet{Chotard2011}).
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $\kappa_2$ represents a universal shift in supernovae biases.
+We place a uniform prior on $\kappa_2$.
 
@@
 Each realisation of cosmology fitted contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-{\red Focus on w}.
+As our primary focus is upon Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
 
- Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_summary}.
+ Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
 Note that we have combined the posteriors for 100 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{100} = 10$ times less than a single realisation.
-Plot summaries for the fits are shown in Figure \ref{fig:bulk_summary}, and the parameter bounds are listed in Table \ref{tab:bulk_summary}.
-The results indicate changes in the underlying population can effect significant changes in the recovered standardisation parameters $\alpha$ and $\beta$, however do not appear to have a significant effect on recovered cosmology.
-Whilst the combined sample (on which the selection function was fit) shows no significant bias (as expected), the {\celeven} and {\gten} only models show opposite biases in $\Omega_m$.
-Cosmological bias is detected with $\approx 3.0\sigma$ significance for the {\celeven} model and $\approx 1.7\sigma$ significance for the {\gten} model, indicating that the population and selection effect treatment cannot capture all necessary information to encapsulate the magnitude and chromatic smearing of the supernovae population.
-However, for the sample size of the DES and LowZ supernova samples (of order 350 supernova), these effects are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.25\sigma$ and, accounting for statistical fluctuations up to $2\sigma$, a minimum bias of $0.08\sigma$ for the {\celeven} model.
+The parameter bounds are listed in Table \ref{tab:bulk_summary}.
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{bulk2.pdf}
-	\end{center}
-	\caption{Graphical parameter summaries for 100 realisations of supernova data.}
-	\label{fig:bulk_summary}
-\end{figure}
+Cosmological bias is detected with $\approx 2.7\sigma$ significance for the {\celeven} model and $\approx 1.5\sigma$ significance for the {\gten} model when fitting both systematics and statistics.
+For the statistics only fits, the respective bias detection levels are $0.9\sigma$ and $5.5\sigma$.
+We investigate the source of the {\celeven} bias and find its source to be bias in the observed summary statistics, in addition to incorrect reported uncertainty on those summary statistics.
+By replacing fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data, and is difficult to do whilst accounting for correlations with population and scatter model.
+This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $\kappa_2$, however neither the {\gten} nor {\celeven} simulations constrained the fitting of this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections without having to precompute standardisation parameters and populations.
+
+For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.27\sigma$ for the full statistics and systematics model fit for the {\celeven} model, and as such we will leave more complicated treatment of them for future work.
 
@@
 	\caption{Posterior surfaces for 100 realisations of supernova data.
-Neither population nor scatter model has a significant effect on posterior surface location or shape.
-Population skewness has the primary effect of changing the degeneracy direction for the $\alpha$ -- $\beta$ contour, whilst the scatter models and skewness have noticeable effects on the value and uncertainty of the recovered parameter $\beta$.}
+Surfaces for $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$ are shown, with the other fit parameters being marginalised over.
+The intrinsic scatter model has significant impact on the recovered $\beta$, which in itself effects cosmology, resulting in small biases in $w$.}
 	\label{fig:bulk_posterior}
 \end{figure}
+
+\begin{table}
+	\centering
+	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
+Whilst the {\gten} models do not show significant evidence of bias, the {\celeven} model does.
+With the current statistical size of the supernovae dataset, the bias is sub-dominant to the size of the uncertainty.
+}
+	\label{tab:bulk_summary}
+	\begin{tabular}{l|cc}
+		\hline
+		Model & $w$ & Bias Detection Level \\
+		\hline
+		{\gten} Stat + Syst  &  $-1.02\pm 0.13$ & $1.5\sigma$  \\
+		{\celeven} Stat + Syst  &  $-0.97\pm 0.11$ & $2.7\sigma$  \\
+		{\gten}  Stat         &  $-1.01\pm 0.11$ & $0.9\sigma$  \\
+		{\celeven} Stat         &  $-0.95\pm 0.09$ & $5.5\sigma$  \\
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 \begin{align}
-m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle\label{eq:disc1} \\
-\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2 \notag\\ 
-&\quad\quad + 2(\beta \sigma_{M_B,c} - \alpha \sigma{m_B,x_1} - \alpha \beta \sigma_{x_1,c}). \label{eq:disc2}
+m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
+\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
 \end{align}

### 2018-02-23 papers/methods/paper.tex
Updating paper

@@
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the bias in the mean colour measured as a function of redshift, and subtract this bias from our observed data.
-We allow the amount of bias to be parametrised to account for different intrinsic scatter models, where parameter $\kappa_2$ controlled the fraction of bias subtraction - 0 for no bias, up to 1 for the maximum bias found in simulations (which corresponded to the intrinsic scatter model from \citet{Chotard2011}).
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $\kappa_2$ represents a universal shift in supernovae biases.
-We place a uniform prior on $\kappa_2$.
+We allow the amount of bias to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlled the fraction of bias subtraction - 0 for no bias, up to 1 for the maximum bias found in simulations (which corresponded to the intrinsic scatter model from \citet{Chotard2011}).
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae biases.
+We place a uniform prior on $s_c$.
 
@@
 % Survey nodes
-\node[latent, above=of sum, yshift=1cm,text width=2.3cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
+\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1, s_c$ };
 \node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
@@
 This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $\kappa_2$, however neither the {\gten} nor {\celeven} simulations constrained the fitting of this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulations constrained the fitting of this parameter, and thus the majority of bias in the {\celeven} fits remain.
 Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections without having to precompute standardisation parameters and populations.
@@
 We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$.
-Unfortunately, this shift to the mean and standard deviation of the normal approximation did not produce stable posterior surfaces which Stan could fit.
-The surface instability was caused by severe degeneracy between three parameters: $\langle c \rangle$, $\sigma_c$ and $\alpha_c$.
-Shifting the mean colour was thus accomplished either by shifting the actual mean colour $\langle c \rangle$, by increasing skewness $\alpha_c$, or by modifying $\sigma_c$.
-Stable surfaces were only found when we fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
-Figure \ref{fig:shift} shows that even an different fixed value than the correct value still give a better approximation than the unshifted approximation.
-However, implementing the shifted approximation in our full model resulted in biased parameter recover for cosmological parameters, resulting in us ultimately discarding the shifted approximation.
-These biases are shown in Figure \ref{fig:simple_w_super}.
+Unfortunately, this shift to the mean and standard deviation of the normal approximation did not produce stable posterior surfaces which Stan could fit when we included full covariance between $m_B$, $x_1$ and $c$ in the underlying population, and so we simplified the model to have independent parameters.
+Stable surfaces with underlying population covariance were found when we fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
+However, this resulted in several population parameters becoming biased, and so we do not fix $\sigma_c$.
+Comparing whether we shift our normal in the approximation or simply discard skewness, Figure \ref{fig:shift} shows that the calculated efficiency is significantly discrepant to the actual efficiency if the normal approximation is not shifted.
+The biases when using shifted or unshifted normal approximations when we fit our model on Gaussian and skewed underlying populations are shown in Figure \ref{fig:simple_w_super}, and only the shifted normal approximation correctly recovers underlying population parameters.
 
@@
 	\end{center}
-	\caption{Testing the significance of fixing $\sigma_c$ when calculating the efficiency using a shifted normal approximation.
-The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$.
+	\caption{Testing the correctness of our normal approximation to the skewed colour distribution.
+The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$, calculated numerically.
 The $x$-axis is analogous to $m_B$ is cosmological context.
 As expected, all efficiencies drop towards zero as shift increases (as objects get fainter).
-With a correct choice of $\sigma_c$, there is negligible error between the correct solution and the shifted normal approximation.
-For both under and over-estimating $\sigma_c$, the shifted normal approximation outperforms the unshifted normal approximation.
-Thus we implemented the shifted normal approximation into our model to test it in full: whilst $\sigma_c$ itself gets sampled.}
+The unshifted normal approximation shows significant discrepancy in the calculated efficiency as it transitions from 1 to 0, whilst the shifted normal approximation shows negligible error to the correct solution.
+From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) as unnecessary.}
 	\label{fig:shift}
@@
 	\begin{center}
-		\includegraphics[width=\textwidth]{simple_w_super_dist.pdf}
+		\includegraphics[width=\textwidth]{simple_w_shift_dist_0.pdf}
 	\end{center}
-	\caption{Marginalised probability distributions for 200 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
+	\caption{Marginalised probability distributions for 100 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
 The dashed green surfaces represent a fit to an underlying Gaussian colour population with the unshifted model.
@@
 The unshifted model recovers the correct colour mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the colour population mean.
-Whilst the performance of the shifted model on skewed data populations is an improvement over the unshifted model, it comes at a cost - significant bias appears in the fits to $\alpha$, $\beta$, $\sigma_{m_B}$ and $w$.
-These biases in cosmological parameters are not found in the unshifted model, and similarly the bias when moving from a Gaussian to skewed colour population is negligible.
-For these reasons, we adopt the unshifted model.}
+Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.}
 	\label{fig:simple_w_super}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Not many fitting methodologies and algorithms can handle the thousands of fit parameters our model requires.
+By using Stan, we are able to take advantage automatic differentiation and the NUTS sampler, which is a class of Hamiltonian Monte Carlo samplers.
+Even with these advantages, early implementations of our model still had excessive fit times, with our desired sub-hour running time far exceeded.
+
+The simplest and most commonly found optimisation we employed was to precompute as much as possible to reduce the complexity of the mathematical graph our model is translated into to compute the surface derivatives.
+For example, when computing the distance modulus, redshift is encountered to various powers.
+Instead of computing those powers in Stan, we simply pass in several arrays of redshift values already raised to the correct power.
+Small changes like this however only give small results.
+
+The primary numerical improvement we made on existing frameworks was to remove costly probability evaluations of multivariate normals.
+To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, you sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$.
+In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution.
+Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to twenty minutes.
+
+This parametrisation does come with one significant downside - inflexibility.
+For each step the algorithm takes, we do not recompute the cholesky decomposition of the covariance of the summary statistics - that happens once at the beginning of the model setup.
+If we had kept the full covariance matrix parametrisation we could modify the matrix easily - for example when incorporating intrinsic dispersion we could simple add on a secondary matrix to create an updated covariance.
+Using cholesky decompositions, ${\rm cholesky}(\Sigma_1 + \Sigma_2) \neq L_1 + L_2$, and so we would need to recompute the decomposition for each step, which discards most of the computational benefit just gained.
+
+Considering a $3\times3$ matrix with cholesky decomposition
+\begin{align}
+L = \begin{pmatrix}
+a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
+\end{pmatrix},
+\end{align}
+the original covariance matrix $\Sigma$ is given by
+\begin{align}
+\Sigma = \begin{pmatrix}
+a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
+\end{pmatrix}.
+\end{align}
+Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered colour parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$.
+We can now see that it is possible to add extra uncertainty to the colour observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix.
+To take our covariance and add on the diagonal uncertainty for colour an extra $\sigma_e$ term, we get
+\begin{align}
+C = \begin{pmatrix}
+\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
+\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
+\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
+\end{pmatrix}
+\end{align}.
+The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
+\begin{align}
+L = \begin{pmatrix}
+a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
+\end{pmatrix},
+\end{align}
+where $g = \sqrt{f^2 + \sigma_e^2} - f$.
+This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent colour uncertainty.
+For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.
 

### 2018-02-23 dessn/configurations/simple_w_shift.py

@@ -98,6 +98,6 @@ if __name__ == "__main__":
             f.write(c1.analysis.get_latex_table(transpose=True))
 
         # print("Plotting big triangle. This might take a while")
-        # c1.plotter.plot(filename=pfn + "_big.png", truth=truth, parameters=7)
+        c1.plotter.plot(filename=pfn + "_big.png", truth=truth, parameters=20)
         # c1.plotter.plot_walks(filename=pfn + "_walk.png", truth=truth, parameters=3)
         # c2.plotter.plot_summary(filename=pfn + "_summary_big.png", truth=truth)

### 2018-02-23 papers/methods/snippets/shift.py

@@ -12,7 +12,7 @@ mids = np.linspace(0.5, 4, 100)
 
 alpha = 2
 delta = alpha / np.sqrt(1 + alpha**2)
-sigma_cor = [0.05, 0.075, 0.1, 0.15]
+sigma_cor = [0.1]
 shift = np.sqrt(2 / np.pi) * delta * sigma
 sigma_ratio1 = np.sqrt(1 - 2 * delta**2 / np.pi)
 
@@ -32,7 +32,7 @@ for i, mid in enumerate(mids):
         efs[i, j] = simps(pop * sf, x=xs)
 
 rc('text', usetex=True)
-fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(4, 5), gridspec_kw={"hspace": 0.0})
+fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(4, 4), gridspec_kw={"hspace": 0.0})
 axes[0].plot(mids, ef_a, label="Correct", c="k")
 axes[0].plot(mids, ef_no, label="Unshifted")
 
@@ -40,8 +40,8 @@ axes[0].plot(mids, ef_no, label="Unshifted")
 axes[1].axhline(0, c="k")
 axes[1].plot(mids, ef_no - ef_a, label="Unshifted")
 for row, c in zip(efs.T, sigma_cor):
-    axes[0].plot(mids, row, label="Shifted, $\sigma_c = %0.2f$" % c, ls="--")
-    axes[1].plot(mids, row - ef_a, label="Shifted, $\sigma_c = %0.2f$" % c, ls="--")
+    axes[0].plot(mids, row, label="Shifted", ls="--")
+    axes[1].plot(mids, row - ef_a, label="Shifted", ls="--")
 
 axes[1].set_xlabel("Shift")
 axes[0].set_ylabel("Efficiency")

### 2018-02-23 papers/methods/paper.tex
Paper updates and digesting latest data

@@
 
-%\section{DES Forecasts}
-%\label{sec:forecasts}
-%
-%To forecast the results for the 3 year DES spectroscopic supernova sample, we generate 100 realisations of datasets, each comprising 300 low redshift supernovae and 250 DES-like supernovae.
-We fit these through the three cosmological models - Flat $\Lambda$CDM, $\Lambda$CDM and $w$CDM, however focus mostly on the $w$CDM model as it is the primary scientific concern of the Dark Energy Survey.
-We run all three cosmological models with systematics enabled and disabled to explicitly show the uncertainty included from systematic contribution.
-%
-%{\red How to include flat lambdacdm - the plot is boring, just quote the results, or just quote uncertainty, should I round the uncertainty to make it symmetric?
-Cant do that for wCDM or LambdaCDM because theyre definitely not gaussian}
-%
-%{\red Talk about how these arent even competitive with JLA and we really need to combine data? Talk about how the photometric supernova dataset should be way better? Talk about statistics limited.}
-%
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[width=\columnwidth]{combined_ApproximateModelW.pdf}
-%	\end{center}
-%	\caption{w.}
-%	\label{fig:combined_w}
-%\end{figure}
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[width=\columnwidth]{combined_ApproximateModelOl.pdf}
-%	\end{center}
-%	\caption{ole.}
-%	\label{fig:combined_ol}
-%\end{figure}
-
 \section{Conclusions}
 
-Conclude.
+Points to make:
+
+1. Created more efficient BHM model with treatment of selection effects, populations, unexplained dispersion, systematics.
+2. Main issue is still scatter model. Need to characterise this theoretically because adding full variance for them into the model results in weak cosmology.
+3.
+Bias in observed summary statistics is difficult to model analytically, and future work should incorporate bias corrections calculated from simulations.
+However, even numerically calculated bias corrections run into issues in that the type Ia dispersion is not well characterised.
+4. Model will be applied shortly to the DES 3yr spectroscopic sample.
 

### 2018-02-23 dessn/configurations/bulk.py

@@ -93,7 +93,7 @@ if __name__ == "__main__":
                      linestyles=ls, colors=cs, shade_alpha=shades)
         c2.plotter.plot_summary(filename=[pfn + "2.png", pfn + "2.pdf"], parameters=["$w$"], truth=[-1.0], figsize=1.5, errorbar=True)
         c2.plotter.plot(filename=[pfn + "_small.png", pfn + "_small.pdf"], parameters=2, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize="column")
-        c2.plotter.plot(filename=[pfn + "_small2.png", pfn + "_small2.pdf"], parameters=5, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize=1.0)
+        c2.plotter.plot(filename=[pfn + "_small2.png", pfn + "_small2.pdf"], parameters=4, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize=1.0)
         # c2.plotter.plot(filename=pfn + "_big.png", parameters=14, truth=truth)
         # c2.plotter.plot_distributions(filename=pfn + "_dist.png", truth=truth, col_wrap=7)
         with open(pfn + "_summary.txt", "w") as f:

### 2018-02-23 dessn/configurations/real_data.py

@@ -20,8 +20,8 @@ if __name__ == "__main__":
         os.makedirs(dir_name)
 
     models = [ApproximateModelW(), ApproximateModelW(statonly=True)]
-    simulation = [SNANASimulation(-1, "DES3YR_LOWZ_COMBINED_FITS"),
-                  SNANASimulation(-1, "DES3YR_DES_COMBINED_FITS")]
+    simulation = [SNANASimulation(-1, "DES3YR_LOWZ_COMBINED_TEXT"),
+                  SNANASimulation(-1, "DES3YR_DES_COMBINED_TEXT")]
 
     fitter = Fitter(dir_name)
 

### 2018-02-23 dessn/snana/convert_snana_data.py

@@ -185,9 +185,14 @@ def digest_simulation(sim_dir, systematics_scales, output_dir, systematic_labels
         x1e = row["x1ERR"]
         ce = row["cERR"]
 
-        sim_mb = row["SIM_mB"]
-        sim_x1 = row["SIM_x1"]
-        sim_c = row["SIM_c"]
+        if "SIM_mB" not in row:
+            sim_mb = 0
+            sim_x1 = 0
+            sim_c = 0
+        else:
+            sim_mb = row["SIM_mB"]
+            sim_x1 = row["SIM_x1"]
+            sim_c = row["SIM_c"]
 
         cov_x1_c = row["COV_x1_c"]
         cov_x0_c = row["COV_c_x0"]
@@ -292,16 +297,16 @@ def convert(base_folder, load_dump=False, override=False, skip=11):
 
 if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG, format="[%(funcName)20s()] %(message)s")
-    # convert("DES3YR_LOWZ_COMBINED_FITS")
-    # convert("DES3YR_DES_COMBINED_FITS")
+    convert("DES3YR_LOWZ_COMBINED_TEXT")
+    convert("DES3YR_DES_COMBINED_TEXT")
     # convert("DES3YR_DES_NOMINAL")
     # convert("DES3YR_LOWZ_NOMINAL")
     # convert("DES3YR_DES_BULK", skip=6)
     # convert("DES3YR_LOWZ_BULK", skip=6)
     # convert("DES3YR_DES_SAMTEST", skip=11)
     # convert("DES3YR_LOWZ_SAMTEST", skip=11)
-    convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11)
-    convert("DES3YR_LOWZ_BHMEFF", load_dump=True, skip=11)
+    # convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11)
+    # convert("DES3YR_LOWZ_BHMEFF", load_dump=True, skip=11)
     # convert("DES3YR_LOWZ_VALIDATION", skip=6)
     # convert("DES3YR_DES_VALIDATION", skip=6)
 

### 2018-02-23 papers/methods/paper.tex
Adding paper conclusion. Tongue in cheek ending. Lets see if it passes review (ha)

@@
 
-Points to make:
+In this paper we have outlined the creation of a Hierarchical Bayesian model for supernova cosmology.
+The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocity uncertainty, calibration uncertainty, dust uncertainty and SALT2 model uncertainty.
+The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to quadratic or cubic run times for models that rely on matrix inversion.
 
-1. Created more efficient BHM model with treatment of selection effects, populations, unexplained dispersion, systematics.
-2. Main issue is still scatter model. Need to characterise this theoretically because adding full variance for them into the model results in weak cosmology.
-3.
-Bias in observed summary statistics is difficult to model analytically, and future work should incorporate bias corrections calculated from simulations.
-However, even numerically calculated bias corrections run into issues in that the type Ia dispersion is not well characterised.
-4. Model will be applied shortly to the DES 3yr spectroscopic sample.
+The importance of validating models using high-precision statistics gained by performing fits to hundreds of data realisations cannot be overstated, however is lacking in many prior BHM models for supernova cosmology.
+We have validated this model against many realisations of simplistic simulations with well-known and well-defined statistics, and found no bias.
+When validating using SNANA simulations, we find evidence of bias which is traced back to light curve fits reporting biased observables and incorrect covariance.
+Allowing fully parametrised corrections on observed supernovae summary statistics is found to make cosmology fits too weak, and allowing simulation based corrections to vary in strength is found to give minor reductions in bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
+For the data size represented in the DES three-year spectroscopic survey, the determined biases should be sub-dominant to other sources of uncertainty, however this cannot be expected for future analyses with larger datasets.
+Stricter bias corrections calculated from simulations are required to reduce bias.
+Ideally, this would include further work on the physical intrinsic dispersion of the Type Ia supernovae population such that we can better characterise this bias.
+
+With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $250\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin.
+With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-ear dataset.
+But that, my friends, is another, much more important, paper.
 

### 2018-02-23 dessn/configurations/real_data.py

@@ -33,6 +33,7 @@ if __name__ == "__main__":
     fitter.set_num_cosmologies(1)
     fitter.set_max_steps(3000)
     fitter.set_num_walkers(200)
+    fitter.set_num_cpu(500)
 
     h = socket.gethostname()
     if h != "smp-hk5pn72":  # The hostname of my laptop. Only will work for me, ha!

### 2018-02-23 papers/methods/paper.tex
Abstract

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\name}{SN-BHM}
+\newcommand{\name}{Steve-o}
 \newcommand{\myemail}{samuelreay@gmail.com}
@@
 \begin{abstract}
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
-Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
+We present a hierarchical Bayesian model for use in supernova cosmology which takes into account full analysis systematics, Malmquist bias, underlying populations and standardisation coefficients.
+Our model is optimised for efficient fitting and has been validated on a combined dataset of more than $250\,000$ supernovae.
+Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ allow rigorous determination of model bias in $w$ when provided with information on matter density (such as might be found from CMB constraints).
+Bias is found to range from $-0.15\sigma$ to $+0.27\sigma$ across intrinsic dispersion models for datasets of comparable strength to the Dark Energy Survey three-year analysis.
 \end{abstract}
@@
 In Section \ref{sec:method} we outline our methodology.
-Model verification on simulated datasets is given in Section \ref{sec:verification}.
-Forecasts for the impending DES three year spectroscopic supernova survey are contained in Section \ref{sec:forecasts}.
-{\red Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies.
-}
+Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.
 
@@
 In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution.
-Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to twenty minutes.
+Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to thirty minutes.
 

### 2018-02-26 papers/methods/paper.tex
Paper updates. Still need to work on motivation

@@
 As such, staggering effort is being put into developing more sophisticated supernovae analyses.
-\citet{Scolnic2016} and \citet{Kessler2017} explore sophisticated simulation corrections to traditional analyses.
+First used in ESSENSE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, the role of simulations has become increasingly important in determining biases and validating models.
+\citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with only the cost of increased computational time \citep{Weyant2013, Jennings2016}.
@@
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009} and \citet{Mandel2011} construct a hierarchical model that they apply to the light curve fitting for supernova.
-\citet{March2011, March2014, Karpenka2015} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method.
-The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
-\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well.
-Notably, these methods rely on data that is bias corrected, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
+\citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to the light curve fitting for supernova.
+\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
+\citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations.
+The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
+\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well by correctly including in their likelihood the determinant of the global covariance matrix.
+Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
+The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision.
+Their model was validated on thirty realisations of statistics.
+Whilst there are a large amount of hierarchical models available, all sofar mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
+None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
+
 The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
-However, whilst there are a large amount of hierarchical models available, none of them have undergone high-statistics simulation verification to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own.
+For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.
 
@@
 We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
-It was found that discarding the skewness entirely resulted in highly biased population recover, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal, with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
+It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal, with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
 This shifted Gaussian approximation removes the unintended bias when simply discarding skewness.

### 2018-02-26 papers/methods/paper.tex
Paper proof

@@
 Our model is optimised for efficient fitting and has been validated on a combined dataset of more than $250\,000$ supernovae.
-Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ allow rigorous determination of model bias in $w$ when provided with information on matter density (such as might be found from CMB constraints).
-Bias is found to range from $-0.15\sigma$ to $+0.27\sigma$ across intrinsic dispersion models for datasets of comparable strength to the Dark Energy Survey three-year analysis.
+Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
+Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.
 \end{abstract}
@@
 Since that time, of the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
-Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}, and yet, despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
+Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}.
+Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
 
-In attempts to tease out the nature of dark energy, currently running and planned surveys are once again ramping up their statistical power.
+In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power.
 The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric confirmation.
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae.
-Such increased statistical power demands a similarly increased fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
+Such increased statistical power demands increased fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
 
@@
 \citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases.
-Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with only the cost of increased computational time \citep{Weyant2013, Jennings2016}.
+Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
 Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
@@
 
-Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements of apparent magnitude and colour.
-Instead, most techniques fit these observations of magnitude (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
+Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements.
+Instead, most techniques fit these observations (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
 This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}.
@@
 \begin{equation}
-M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections},
+M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections}, \label{eq:standard}
 \end{equation}
@@
 
-Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the cosmologically predicted values $\mu_C$ and the observed distance modulus $\mu_{\rm obs}$, shown respectively below:
+Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$; given by
 \begin{align}
@@
 r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \label{eq:mu0}\\
-\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B \label{eq:mu1}
+\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
 \end{align}
@@
 Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity.
-Secondly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
+Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus loses all information about uncertainty correlation.
+Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
@@
 
-Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies Hierarchical Bayesian Models.
+Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies hierarchical Bayesian Models.
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
@@
 
-We construct our Bayesian Hierarchical Model with several goals in mind: creation of a redshift-dependent correlated underlying supernova population, increased treatment of systematics, and analytic correction of selection effects.
+We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty.
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
@@
 Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^10$ solar masses.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
 Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.
@@
 
-The underlying supernova population is often treated with two components - a population distribution in colour and stretch, and intrinsic dispersion.
+The underlying supernova population is characterised by underlying distributions in colour, stretch and magnitude.
 Analytic models often treat the colour and stretch populations with a skew normal and normal, respectively, and have them as independent.
@@
 P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \rho, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
-\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
+\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
 \end{align}
@@
 {\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0).
-\citet{Shariff2016} instead of employ a colour-dependent $\beta$, model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
+\citet{Shariff2016} instead model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
 
@@
 Our treatment of selection effects is to incorporate selection efficiency into our model, rather than relying on simulation-driven data corrections.
-As such, we need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
+We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
@@
 As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC (a brief dismissal of months of struggle).
-We therefore simplify the integral and approximate the selection effects in apparent magnitude and redshift space independently, such that the denominator, denoted now $w$ for simplicity, is given as
+Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
+We therefore simplify the integral and approximate the selection effects in apparent magnitude and redshift space independently, such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
 \begin{align}
@@
 We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
-It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal, with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
+It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
 This shifted Gaussian approximation removes the unintended bias when simply discarding skewness.
@@
 The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
-However, for surveys which suffer from saturation and thus rejection of low-z supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a good analytic form, taking the form
+However, for surveys which suffer from saturation and thus rejection of low-z supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form
 \begin{align}
@@
 The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude.
-Uncertainty of the Malmquist bias (entering both through statistical uncertainty from finite sized simulations in the efficiency ratio and the discrepancy between the analytic approximation and non-analytic simulation results) is incorporated into the fitting for the analytic approximation.
-Uncertainty is uniformly added to the efficiency ratio until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
+Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit.
+Uncertainty was uniformly added to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
 
@@
 w_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
-&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right).
-\label{eq:sellowz}
+&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
 \end{align}
+and is incorporated into our likelihood.
 
@@
 For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
-Masses are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
+Mass probabilities are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
 Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$.
@@
 The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae.
-We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES spectroscopic analysis.
+We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES spectroscopic analysis.
 Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys.
@@
 Cosmological parameters are recovered without significant bias.
-Combined posterior surfaces of all 100 realisations for Flat $wCDM$ fits are shown in Figure \ref{fig:simple_w}.
+Combined posterior surfaces of all 100 realisations fits for $\Lambda$CDM are shown in Figure \ref{fig:simple_ol} and fits for Flat $wCDM$ are shown in Figure \ref{fig:simple_w}.
 By utilising the Stan framework and several efficient parametrisations (discussed further in Appendix \ref{app:optimisations}), fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
 
-To enforce investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$.
-This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces, and the results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
+To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$.
+This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces.
+The results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
 
@@
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
-The first model is the \citet[][hereafter denoted the {\gten} scatter model]{Guy2010}, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The second model, denoted the {\celeven} model is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+The first model is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second model, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
 
@@
 
-Each realisation of cosmology fitted contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+Each realisation of cosmology simulated contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
 As our primary focus is upon Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
@@
 This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulations constrained the fitting of this parameter, and thus the majority of bias in the {\celeven} fits remain.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections without having to precompute standardisation parameters and populations.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
 
-For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, representing at most a deviation of $0.27\sigma$ for the full statistics and systematics model fit for the {\celeven} model, and as such we will leave more complicated treatment of them for future work.
+For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+For our full systematics model, the bias represents a deviation from $0.27\sigma$ to $-0.15\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
 
@@
 
-In this paper we have outlined the creation of a Hierarchical Bayesian model for supernova cosmology.
+In this paper we have outlined the creation of a hierarchical Bayesian model for supernova cosmology.
 The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocity uncertainty, calibration uncertainty, dust uncertainty and SALT2 model uncertainty.
-The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to quadratic or cubic run times for models that rely on matrix inversion.
+The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion.
 
@@
 With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-ear dataset.
-But that, my friends, is another, much more important, paper.
+But that, my friends, is another paper.
 

### 2018-02-26 papers/methods/paper.tex
Appendix update

@@
 
-We assume that our selection effects can be reasonably well encapsulated by independent functions of (actual) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
+We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
 Our denominator then becomes
 \begin{align}
-w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta),
+w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta), \label{eq:sel}
 \end{align}
@@
 Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-$P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$.
+$P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$ (which is why they were not included in equation \eqref{eq:sel}).
 Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.
@@
 \begin{align}
-\int P(S|m_B) P(m_B | z, \theta)\, dm_B
+\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
 \end{align}
@@
 \end{align}
-Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.
+Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.
 
@@
 \rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
-\end{pmatrix}
-\end{align}.
+\end{pmatrix}.
+\end{align}
 The cholesky decomposition of this is, in terms of the original cholesky decomposition, is

### 2018-02-26 papers/methods/paper.tex
Fixing typo

@@
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Hierarchical Bayesian Models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
+Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
 

### 2018-02-27 papers/methods/paper.tex
Updating conclusion
Adding correlation section

@@
 
+Table \ref{tab:parameter_correlations} list the fit correlations to our model fit parameters (excluding the LowZ band systematics, which had negligible correlation), and Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
+As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion.
+Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology than $\beta$.
+Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
+The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
+
 For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
@@
 
+\begin{table}
+	\centering
+	\caption{Parameter Correlations with $w$ for the combined 100 simulation fits.
+Correlations for the LowZ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.}
+	\label{tab:parameter_correlations}
+	\begin{tabular}{c|cc}
+		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst  \\
+		\hline
+		$\Omega_m$ &   -0.19 &     -0.21     \\
+		$\alpha$ &   -0.17 &     -0.20     \\
+		$\beta$ &   -0.29 &     -0.23     \\
+		$\langle M_B \rangle$ &    0.68 &      0.66     \\
+		$\sigma_{\rm m_B}^{0}$ &    0.04 &      0.07     \\
+		$\sigma_{\rm m_B}^{1}$ &    0.23 &      0.18     \\
+		$\sigma_{x_1}^{0}$ &    0.04 &      0.03     \\
+		$\sigma_{x_1}^{1}$ &    0.05 &      0.01     \\
+		$\sigma_{c}^{0}$ &    0.01 &      0.11     \\
+		$\sigma_{c}^{1}$ &    0.08 &      0.04     \\
+		$\alpha_c^{0}$ &   -0.04 &      0.04     \\
+		$\alpha_c^{1}$ &    0.03 &      0.01     \\
+		$\kappa_{c0}^{0}$ &   -0.10 &     -0.05     \\
+		$\kappa_{c0}^{1}$ &   -0.20 &     -0.17     \\
+		$\kappa_{c1}^{0}$ &   -0.05 &     -0.01     \\
+		$\kappa_{c1}^{1}$ &   -0.01 &      0.01     \\
+		$s_c^{0}$ &    0.00 &      0.01     \\
+		$\delta(0)$ &    0.00 &      0.00     \\
+		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
+		$\langle x_1^{0} \rangle$ &   -0.01 &     -0.05     \\
+		$\langle x_1^{1} \rangle$ &   -0.02 &      0.02     \\
+		$\langle x_1^{2} \rangle$ &   -0.04 &     -0.04     \\
+		$\langle x_1^{3} \rangle$ &   -0.03 &     -0.06     \\
+		$\langle x_1^{4} \rangle$ &   -0.06 &     -0.06     \\
+		$\langle x_1^{5} \rangle$ &    0.04 &      0.02     \\
+		$\langle x_1^{6} \rangle$ &    0.04 &      0.04     \\
+		$\langle x_1^{7} \rangle$ &    0.08 &      0.03     \\
+		$\langle c^{0} \rangle$ &   -0.05 &     -0.12     \\
+		$\langle c^{1} \rangle$ &    0.11 &      0.03     \\
+		$\langle c^{2} \rangle$ &    0.11 &      0.06     \\
+		$\langle c^{3} \rangle$ &    0.14 &      0.04     \\
+		$\langle c^{4} \rangle$ &   -0.11 &     -0.11     \\
+		$\langle c^{5} \rangle$ &   -0.15 &     -0.08     \\
+		$\langle c^{6} \rangle$ &   -0.12 &     -0.13     \\
+		$\langle c^{7} \rangle$ &   -0.12 &     -0.06     \\
+		$\delta [ {\rm SALT}_0 ]$ &    0.05 &      0.05     \\
+		$\delta [ {\rm SALT}_1 ]$ &   -0.01 &      0.02     \\
+		$\delta [ {\rm SALT}_2 ]$ &   -0.10 &     -0.09     \\
+		$\delta [ {\rm SALT}_3 ]$ &   -0.03 &     -0.03     \\
+		$\delta [ {\rm SALT}_4 ]$ &    0.08 &      0.09     \\
+		$\delta [ {\rm SALT}_5 ]$ &    0.01 &      0.02     \\
+		$\delta [ {\rm SALT}_6 ]$ &    0.05 &      0.07     \\
+		$\delta [ {\rm SALT}_7 ]$ &   -0.11 &     -0.10     \\
+		$\delta [ {\rm SALT}_8 ]$ &    0.01 &      0.02     \\
+		$\delta [ {\rm SALT}_9 ]$ &    0.02 &      0.02     \\
+		$\delta [ {\rm MWE}_{B-V} ]$ &    0.03 &      0.02     \\
+		$\delta [ {\rm HST Calib} ]$ &   -0.07 &     -0.07     \\
+		$\delta [ v_{\rm pec} ]$ &    0.00 &     -0.01     \\
+		$\delta [ {\rm \delta z } ]$ &    0.01 &      0.00     \\
+		$\delta [ \Delta g ]$ &    0.05 &      0.11     \\
+		$\delta [ \Delta r ]$ &    0.16 &      0.10     \\
+		$\delta [ \Delta i ]$ &   -0.16 &     -0.18     \\
+		$\delta [ \Delta z ]$ &   -0.26 &     -0.26     \\
+		$\delta [ \Delta \lambda_g ]$ &    0.16 &      0.20     \\
+		$\delta [ \Delta \lambda_r ]$ &    0.05 &      0.06     \\
+		$\delta [ \Delta \lambda_i ]$ &    0.00 &     -0.01     \\
+		$\delta [ \Delta \lambda_z ]$ &    0.09 &      0.07     \\
+		\hline
+	\end{tabular}
+\end{table}
+%		$\delta [ \Delta t ]$ &   -0.01 &      0.00     \\
+%		$\delta [ \Delta y ]$ &   -0.03 &     -0.04     \\
+%		$\delta [ \Delta Z ]$ &    0.10 &      0.10     \\
+%		$\delta [ \Delta A ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta u ]$ &   -0.04 &     -0.03     \\
+%		$\delta [ \Delta v ]$ &    0.01 &      0.02     \\
+%		$\delta [ \Delta \lambda_{t} ]$ &    0.00 &      0.02     \\
+%		$\delta [ \Delta \lambda_{y} ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta \lambda_{z} ]$ &   -0.05 &     -0.04     \\
+%		$\delta [ \Delta \lambda_{A} ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{u} ]$ &   -0.02 &     -0.02     \\
+%		$\delta [ \Delta \lambda_{v} ]$ &   -0.01 &      0.01     \\
+%		$\delta [ \Delta gl ]$ &   -0.12 &     -0.13     \\
+%		$\delta [ \Delta hm ]$ &    0.02 &      0.01     \\
+%		$\delta [ \Delta in ]$ &    0.08 &      0.13     \\
+%		$\delta [ \Delta jo ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{gl} ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta \lambda_{hm} ]$ &   -0.01 &     -0.03     \\
+%		$\delta [ \Delta \lambda_{in} ]$ &   -0.10 &     -0.12     \\
+%		$\delta [ \Delta \lambda_{jo} ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta p ]$ &   -0.01 &     -0.01     \\
+%		$\delta [ \Delta q ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta R ]$ &    0.00 &      0.01     \\
+%		$\delta [ \Delta s ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{p} ]$ &    0.00 &      0.01     \\
+%		$\delta [ \Delta \lambda_{q} ]$ &    0.01 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{r} ]$ &   -0.01 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{s} ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta b ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta c ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta d ]$ &    0.00 &      0.01     \\
+%		$\delta [ \Delta e ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta \lambda_{b} ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{c} ]$ &    0.00 &      0.00     \\
+%		$\delta [ \Delta \lambda_{d} ]$ &    0.00 &     -0.01     \\
+%		$\delta [ \Delta \lambda_{e} ]$ &    0.00 &     -0.01     \\
+%		$\Delta_{0}$ &    0.00 &      0.01     \\
+%		$\Delta_{1}$ &    0.00 &     -0.01     \\
+%		$\Delta_{2}$ &    0.00 &      0.00     \\
+%		$\Delta_{3}$ &   -0.01 &     -0.01     \\
+%		$\Delta_{4}$ &    0.00 &     -0.01     \\
+%		$\Delta_{5}$ &    0.00 &     -0.01     \\
+%		$\Delta_{6}$ &    0.00 &      0.00     \\
+%		$\Delta_{7}$ &    0.00 &     -0.01     \\
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=\textwidth]{bulk_cor.pdf}
+	\end{center}
+	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
+We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
+Also visible in this figure are several other interesting relationships.
+$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
+This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
+As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
+And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
+The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
+These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
+As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
+Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
+Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
+}
+	\label{fig:bulk_cor}
+\end{figure*}
+
 \section{Conclusions}
@@
 The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocity uncertainty, calibration uncertainty, dust uncertainty and SALT2 model uncertainty.
+Furthermore, these uncertainties, being explicitly parametrised in our model, are captured will covariance intact, an improvement on many previous methods.
 The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion.

### 2018-02-27 dessn/configurations/bulk.py

@@ -2,6 +2,7 @@ import os
 import logging
 import socket
 
+
 from dessn.framework.fitter import Fitter
 from dessn.framework.models.approx_model import ApproximateModelW, ApproximateModel
 from dessn.framework.simulations.snana import SNANASimulation
@@ -82,77 +83,99 @@ if __name__ == "__main__":
             else:
                 print("C1")
                 c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
-        # for m, s, chain, truth, weight, old_weight, posterior in res2:
-        #     name = "All"
-        #     if isinstance(m, ApproximateModelW):
-        #         print("C2")
-        #         c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
-        #     else:
-        #         print("C1")
-        #         c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
-        #
+
         c2.configure(spacing=1.0, sigma2d=False, flip=False, shade=True,
                      linestyles=ls, colors=cs, shade_alpha=shades)
-        c2.plotter.plot_summary(filename=[pfn + "2.png", pfn + "2.pdf"], parameters=["$w$"], truth=[-1.0], figsize=1.5, errorbar=True)
-        c2.plotter.plot(filename=[pfn + "_small.png", pfn + "_small.pdf"], parameters=2, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize="column")
-        c2.plotter.plot(filename=[pfn + "_small2.png", pfn + "_small2.pdf"], parameters=4, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize=1.0)
+        # c2.plotter.plot_summary(filename=[pfn + "2.png", pfn + "2.pdf"], parameters=["$w$"], truth=[-1.0], figsize=1.5, errorbar=True)
+        # c2.plotter.plot(filename=[pfn + "_small.png", pfn + "_small.pdf"], parameters=2, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize="column")
+        # c2.plotter.plot(filename=[pfn + "_small2.png", pfn + "_small2.pdf"], parameters=4, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize=1.0)
         # c2.plotter.plot(filename=pfn + "_big.png", parameters=14, truth=truth)
         # c2.plotter.plot_distributions(filename=pfn + "_dist.png", truth=truth, col_wrap=7)
-        with open(pfn + "_summary.txt", "w") as f:
[…]

### 2018-02-28 papers/methods/paper.tex
Reorgansing
Three-ear to three-year

@@
 
+\begin{table}
+	\centering
+	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 100 fits to independent realisations of toy supernova data.
+As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
+Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
+As our surfaces are stacked, a shift away from the simulated values of more than $\sigma/\sqrt{100} = 0.1\sigma$ would give a $1\sigma$ detection of bias.
+No significant bias is detected in either cosmological model.}
+	\label{tab:simple_model}
+	\begin{tabular}{l|cc}
+		\hline
+		Model & $\Omega_m$ & $w$ \\ 
+		\hline
+		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
+		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
+		\hline
+	\end{tabular}
+\end{table}
+
 We generate simple toy data to validate the basic premise of the model.
@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
-
-\begin{table}
-	\centering
-	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 100 fits to independent realisations of toy supernova data.
-As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
-Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
-As our surfaces are stacked, a shift away from the simulated values of more than $\sigma/\sqrt{100} = 0.1\sigma$ would give a $1\sigma$ detection of bias.
-No significant bias is detected in either cosmological model.}
-	\label{tab:simple_model}
-	\begin{tabular}{l|cc}
-		\hline
-		Model & $\Omega_m$ & $w$ \\ 
-		\hline
-		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
-		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
-		\hline
-	\end{tabular}
-\end{table}
 
@@
 \end{figure}
-
-Each realisation of cosmology simulated contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-As our primary focus is upon Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
-
- Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-Note that we have combined the posteriors for 100 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{100} = 10$ times less than a single realisation.
-The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-
-Cosmological bias is detected with $\approx 2.7\sigma$ significance for the {\celeven} model and $\approx 1.5\sigma$ significance for the {\gten} model when fitting both systematics and statistics.
-For the statistics only fits, the respective bias detection levels are $0.9\sigma$ and $5.5\sigma$.
-We investigate the source of the {\celeven} bias and find its source to be bias in the observed summary statistics, in addition to incorrect reported uncertainty on those summary statistics.
-By replacing fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data, and is difficult to do whilst accounting for correlations with population and scatter model.
-This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
-
-Table \ref{tab:parameter_correlations} list the fit correlations to our model fit parameters (excluding the LowZ band systematics, which had negligible correlation), and Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion.
-Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology than $\beta$.
-Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
-The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
-
-For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-For our full systematics model, the bias represents a deviation from $0.27\sigma$ to $-0.15\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
 
@@
 \end{table}
+
+Each realisation of cosmology simulated contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+As our primary focus is upon Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
+
+ Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
+Note that we have combined the posteriors for 100 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{100} = 10$ times less than a single realisation.
+The parameter bounds are listed in Table \ref{tab:bulk_summary}.
+
+Cosmological bias is detected with $\approx 2.7\sigma$ significance for the {\celeven} model and $\approx 1.5\sigma$ significance for the {\gten} model when fitting both systematics and statistics.
+For the statistics only fits, the respective bias detection levels are $0.9\sigma$ and $5.5\sigma$.
+We investigate the source of the {\celeven} bias and find its source to be bias in the observed summary statistics, in addition to incorrect reported uncertainty on those summary statistics.
+By replacing fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data, and is difficult to do whilst accounting for correlations with population and scatter model.
+This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=\textwidth]{bulk_cor.pdf}
+	\end{center}
+	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
+We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
+Also visible in this figure are several other interesting relationships.
+$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
+This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
+As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
+And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
+The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
+These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
+As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
+Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
+Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
+}
+	\label{fig:bulk_cor}
+\end{figure*}
+
+Table \ref{tab:parameter_correlations} list the fit correlations to our model fit parameters (excluding the LowZ band systematics, which had negligible correlation), and Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
+As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion.
+Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology than $\beta$.
+Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
+The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
+
+For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+For our full systematics model, the bias represents a deviation from $0.27\sigma$ to $-0.15\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
+
 %		$\delta [ \Delta t ]$ &   -0.01 &      0.00     \\
@@
 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=\textwidth]{bulk_cor.pdf}
-	\end{center}
-	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
-We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
-Also visible in this figure are several other interesting relationships.
-$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
-This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
-As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
-And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
-The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
-These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
-As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
-Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
-Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
-}
-	\label{fig:bulk_cor}
-\end{figure*}
-
 \section{Conclusions}
@@
 With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $250\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin.
-With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-ear dataset.
+With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-year dataset.
 But that, my friends, is another paper.

### 2018-03-12 dessn/configurations/real_data.py
Plots for Anais

@@ -60,41 +60,62 @@ if __name__ == "__main__":
             c.add_chain(chain, weights=weight, posterior=posterior, name=name)
             c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
 
+        import numpy as np
+        # data = np.vstack(chain[x] for x in ['$\\Omega_m$', '$w$', '$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$', '$\\delta(0)$'])
+        # data = data.T
+        # np.save("samchain.npy", data.astype(np.float32))
+        # np.savetxt("samchain.txt", data)
+        # print(data.shape)
+        # exit()
+
         chain_planck, params, weights, likelihood = get_planck()
         if blind:
             chain_planck[:, 0] = blind_om(chain_planck[:, 0])
             chain_planck[:, 1] = blind_w(chain_planck[:, 1])
         c.add_chain(chain_planck, parameters=params, name="Planck")
 
-        import numpy as np
         subset = np.vstack((chain_full[r"$\Omega_m$"], chain_full["$w$"]))
-        print(subset.shape, chain_planck.shape)
         kde = gaussian_kde(chain_planck.T)
         print("Trained")
-        weights = kde.evaluate(subset)
-        print(weights.shape)
-        c.add_chain(subset.T, weights=weights, name="Combined", parameters=params)
-        c2.add_chain(subset.T, weights=weights, name="Combined", parameters=params)
+        weights = kde.evaluate(subset[:, ::1])
+        print("Eval")
+        c.add_chain(subset.T[::1, :], weights=weights, name="Combined", parameters=params)
+        c2.add_chain(subset.T[::1, :], weights=weights, name="Combined", parameters=params)
+
+        if False:
+            bonnie_file = plot_dir + "/output/bonnie.txt"
+            import numpy as np
+            bnames = ['$\\Omega_m$', '$w$', '$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$', '$\\delta(0)$']
+            bonnie_data = np.loadtxt(bonnie_file, delimiter=",")
+            bonnie_data = bonnie_data[10000:, :]
+            bdic = {bnames[i]: bonnie_data[:, i] for i in range(len(bnames))}
+            subset_bonnie = np.vstack((bdic[r"$\Omega_m$"], bdic["$w$"]))
[…]

### 2018-03-12 papers/methods/paper.tex
Updating caption
Initial draft feedback

@@
 \begin{abstract}
-We present a hierarchical Bayesian model for use in supernova cosmology which takes into account full analysis systematics, Malmquist bias, underlying populations and standardisation coefficients.
-Our model is optimised for efficient fitting and has been validated on a combined dataset of more than $250\,000$ supernovae.
+We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects.
+This advances the work of \citet{Rubin2015} by including additional sources of systematic uncertainty and improved treatment of Malmquist bias whilst increasing numerical efficiency.
+Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae.
 Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.
+SNANA simulations can recover the input cosmology to within $0.27\sigma$, while taking into account many more effects than standard analyses.
 \end{abstract}
+% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.
 
@@
 Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, of the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
-Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structure \citep{Alam2017} and the CMB \citep{Hinshaw2013, PlanckCollaboration2013}.
+Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
+Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
+For a recent review, see \citet{Huterer2018}.
 Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
@@
 As such, staggering effort is being put into developing more sophisticated supernovae analyses.
-First used in ESSENSE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, the role of simulations has become increasingly important in determining biases and validating models.
+First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, the role of simulations has become increasingly important in determining biases and validating models.
 \citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases.
@@
 The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology.
-However, given some modelled underlying population that lives in the redshift-dependent space $M_B$, $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
+However, given some modelled underlying population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
 Specifically, for any given supernova our functional map may take the traditional form:
@@
 \end{equation}
-where $\alpha$ is the stretch correction \citep{Phillips1993}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
+where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
 The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
@@
 
-Traditional $\chi^2$ analyses such as that found in \citet{Kowalski2008, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$; given by
+Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$.
+This function is given by
+\begin{equation}
+\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
+\end{equation}
+where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
+The cosmological $\mu_C$ is defined as
 \begin{align}
@@
 \end{align}
-The minimising function is then given as
-\begin{equation}
-\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C)
-\end{equation}
-where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
-The benefit this analysis methodology provides is speed - for samples of hundreds of supernovae, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
+
+The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
 The speed comes with two costs.
 Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity.
-Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus loses all information about uncertainty correlation.
+Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation.
 Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
@@
 Their model was validated on thirty realisations of statistics.
-Whilst there are a large amount of hierarchical models available, all sofar mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
+
+Whilst there are a large amount of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
 None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
@@
 
-We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty.
+We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
@@
 
-Like most of the BHM methods introduced previously, we work from the summary statistics as well, where each observed supernova has an apparent magnitude $\hat{m_B}$, stretch $\hat{x_1}$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has an apparent magnitude $\hat{m}_B$, stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-Given machine learning classifiers to type supernovae, we will also have a probability of being a Type Ia, $\hat{p}$.
-Our set of observables is therefore given as $\lbrace \hat{m_B}, \hat{x_1}, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+We will also have a probability of being a Type Ia, $\hat{p}$.
+Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 \begin{equation}
-P(\hat{m_B},\hat{x_1},\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m_B},\hat{x_1},\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
+P(\hat{m}_B,\hat{x}_1,\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m}_B,\hat{x}_1,\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
 \end{equation}
-However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics report from light curve fitting are not Gaussian.
+However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian.
 We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation.
 As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence.
-As this may represent underlying physics more than the effects of Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
+As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
 We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
+The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
 Each survey has its own set of $\kappa$ parameters.
-Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the bias in the mean colour measured as a function of redshift, and subtract this bias from our observed data.
-We allow the amount of bias to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlled the fraction of bias subtraction - 0 for no bias, up to 1 for the maximum bias found in simulations (which corresponded to the intrinsic scatter model from \citet{Chotard2011}).
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae biases.
+Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
+We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlled the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponded to the intrinsic scatter model from][]{Chotard2011}.
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour.
 We place a uniform prior on $s_c$.
@@
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
 
@@
 % Obs nodes
-\node[obs]  (obssum) {$\hat{m_B}, \hat{x_1}, \hat{c}, \mathcal{C}^{-1}$};
+\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C}^{-1}$};
 \node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
@@
 \caption{Probabilistic graphical model for our likelihood.
-Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.}
+Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.
+The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernovae population for the survey and the systematics associated with it.}
 \label{fig:pgm}
@@
 Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution.
-Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the population possesses similar ability to change as a function of redshift.
+Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the model population possesses similar ability to change as a function of redshift.
 
@@
 The chief difficulty with including systematic uncertainties in supernova analyses is that they generally occur during the observational pipeline, and have difficult-to-model effects on the output observations.
-As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m_B}}{\partial \Z_i}$, $\frac{\partial \hat{x_1}}{\partial \Z_i}$, $\frac{\partial \hat{c}}{\partial \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
+As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m}_B}{\partial \Z_i}$, $\frac{\partial \hat{x}_1}{\partial \Z_i}$, $\frac{\partial \hat{c}}{\partial \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
 
@@
 Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
-We therefore simplify the integral and approximate the selection effects in apparent magnitude and redshift space independently, such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
+We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently, such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
 \begin{align}
@@
 \end{align}
-We apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
+We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
 It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
@@
 m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
-\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2
+\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
 \end{align}
@@
 \begin{align}
-P(S|m_B) = \Phi^C(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
+P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
 \end{align}
@@
 The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
-However, for surveys which suffer from saturation and thus rejection of low-z supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form
+However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form
 \begin{align}
@@
 Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit.
-Uncertainty was uniformly added to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
+Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
 
-With the well sampled approximation as specified previously, we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
+With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
 For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), this correction becomes
@@
 	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys.
-Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.}
+Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.
+The colour of the data points is representative over the supernovae colour itself, blue dots represent blue supernovae, with warmer colours representing redder supernovae.}
 	\label{fig:simple_pop}
@@
 	\caption{Posterior surfaces for 100 realisations of supernova data with the Flat $w$CDM model.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing incorrect statistics due to the highly non-Gaussian nature of the posterior surface.}
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface, where the point attained from each parameters independent maximum likelihood lands well outside the $3\sigma$ contour in parameter space.}
 	\label{fig:simple_w}
@@
 Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
-Extra dispersion of $0.02(1+3z)$ is added in quadrature to the $c$ uncertainty to simulation unexplained colour dispersion from chromatic variation in supernovae.
+Extra dispersion of $0.02(1+3z)$ is added in quadrature to the $c$ uncertainty to simulate unexplained colour dispersion from chromatic variation in supernovae.
 The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
@@
 We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES spectroscopic analysis.
-Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys.
+Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
 
@@
 Cosmological parameters are recovered without significant bias.
-Combined posterior surfaces of all 100 realisations fits for $\Lambda$CDM are shown in Figure \ref{fig:simple_ol} and fits for Flat $wCDM$ are shown in Figure \ref{fig:simple_w}.
+Combined posterior surfaces of all 100 realisations fits for $\Lambda$CDM are shown in Figure \ref{fig:simple_ol} and fits for Flat $w$CDM are shown in Figure \ref{fig:simple_w}.
 By utilising the Stan framework and several efficient parametrisations (discussed further in Appendix \ref{app:optimisations}), fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
@@
 This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces.
+The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}.
+Without making such a slice, the variation in $w$ can appear to be large due to a shift along the degeneracy direction of the banana.
+By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them.
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
@@
 The selection effects observed by comparing the generated supernovae to those that pass our simulated cuts and were successfully `observed' are shown in Figure~\ref{fig:sf_snana}, and it is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
-We note that there was no significant difference between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.
+We note that there was no significant difference in the selection effect of Malmquist bias between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.
 
@@
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
+Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+
 Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data, and is difficult to do whilst accounting for correlations with population and scatter model.
@@
 As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
+Strong anti-correlation between the $\kappa_{c0}^0$ and $\kappa_{c0}^1$ (and positive fit values) shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
 Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
@@
 
-Table \ref{tab:parameter_correlations} list the fit correlations to our model fit parameters (excluding the LowZ band systematics, which had negligible correlation), and Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
+Table \ref{tab:parameter_correlations} lists the fit correlations to our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
+Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
 As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion.
-Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology than $\beta$.
+Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology as $\beta$.
 Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
@@
 For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-For our full systematics model, the bias represents a deviation from $0.27\sigma$ to $-0.15\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
+For our full systematics model, the bias represents a deviation from $-0.15\sigma$ to $0.27\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
 
@@
 
-The importance of validating models using high-precision statistics gained by performing fits to hundreds of data realisations cannot be overstated, however is lacking in many prior BHM models for supernova cosmology.
-We have validated this model against many realisations of simplistic simulations with well-known and well-defined statistics, and found no bias.
-When validating using SNANA simulations, we find evidence of bias which is traced back to light curve fits reporting biased observables and incorrect covariance.
-Allowing fully parametrised corrections on observed supernovae summary statistics is found to make cosmology fits too weak, and allowing simulation based corrections to vary in strength is found to give minor reductions in bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
+The importance of validating models using high-precision statistics gained by performing fits to hundreds of data realisations cannot be overstated, however this validation is lacking in many earlier BHM models for supernova cosmology.
+We have validated this model against many realisations of simplistic simulations with well-known and well-defined statistics, and found no cosmological bias.
+When validating using SNANA simulations, we find evidence of cosmological bias which is traced back to light curve fits reporting biased observables and incorrect covariance.
+Allowing fully parametrised corrections on observed supernovae summary statistics is found to make cosmology fits too weak, and allowing simulation based corrections to vary in strength is found to give minor reductions in $w$ bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
 For the data size represented in the DES three-year spectroscopic survey, the determined biases should be sub-dominant to other sources of uncertainty, however this cannot be expected for future analyses with larger datasets.
@@
 With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-year dataset.
-But that, my friends, is another paper.
 
@@
 \begin{align}
-w &= \int d\hat{z} \, d\hat{m_B} \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m_B}|m_B) P(z, m_B | \theta), \label{eq:sel}
+w &= \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
 \end{align}
@@
 Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-$P(\hat{m_B} | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m_B}$ (which is why they were not included in equation \eqref{eq:sel}).
+$P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$ (which is why they were not included in equation \eqref{eq:sel}).
 Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.

### 2018-03-12 papers/methods/paper.tex
Anais feedback 1

@@
 
+{\red Todo: get info on all used lowz surveys and introduce them properly. Help in terms of the right papers to cite would be appreciated.}
+
 Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
 Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
-Cosmological analysis of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
+Cosmological analyses of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
 For a recent review, see \citet{Huterer2018}.
@@
 In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric confirmation.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples.
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae.
-Such increased statistical power demands increased fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
+Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
 
 As such, staggering effort is being put into developing more sophisticated supernovae analyses.
-First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, the role of simulations has become increasingly important in determining biases and validating models.
+The role of simulations has become increasingly important in determining biases and validating models.
+First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
 \citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, however often face difficulties finding sufficient analytic approximations for complicated effects such as Malmquist bias.
+Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either also use simulation corrections or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
 
@@
 Instead, most techniques fit these observations (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-This model is trained separately before fitting the supernovae light curves for the cosmology selected supernova sample \citep{Guy2010, Mosher2014}.
+This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}.
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
@@
 
-The underlying actual supernova population is not as clear cut, and indeed accurately characterising this population, its evolution over redshift and effects from environment is one of the challenges of supernova cosmology.
+The underlying actual supernova population is not such a clear cut, and accurately characterising this population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
 However, given some modelled underlying population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
@@
 where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
-The $\text{corrections}$ term at the end often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
+The $\text{corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
 The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.
 
-\subsection{Traditional Analyses}
+\subsection{Traditional Cosmology Analyses}
 
@@
 \begin{align}
-\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right]\\
-r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}} \label{eq:mu0}\\
+\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
+r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}\\
 \mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
@@
 The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
-The speed comes with two costs.
-Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of uncertainty Gaussianity.
+The speed comes with several limitations.
+Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty.
 Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation.
@@
 
-To try and escape the limitations of the traditional analysis methodology, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
+To try and escape the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
 \citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and Flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}.
@@
 One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
-\citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, however evaluation of systematic uncertainty was only performed using the Tripp metric.
+\citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, although evaluation of systematic uncertainty was only performed using the Tripp metric.
 
@@
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to the light curve fitting for supernova.
-\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide a model that offers increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
+\citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
+\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
 \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations.
-The recent BAHAMAS model \citep{Shariff2016} builds off this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
-\citet{Ma2016} also performed a reanalysis of the JLA dataset, finding significant differences in $\alpha$ and $\beta$ values from the original as well by correctly including in their likelihood the determinant of the global covariance matrix.
+The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
+\citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}.
 Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
@@
 Whilst there are a large amount of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
-None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
+None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
 
@@
 
-The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae type rates and redshift distributions, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
+The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae rates and redshift distributions, and host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
 
@@
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
-We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlled the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponded to the intrinsic scatter model from][]{Chotard2011}.
+We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}.
 Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour.
@@
 The underlying supernova population is characterised by underlying distributions in colour, stretch and magnitude.
-Analytic models often treat the colour and stretch populations with a skew normal and normal, respectively, and have them as independent.
+Analytic models often treat the colour and stretch populations independently with a skew normal and normal, respectively.
 Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution.
@@
 
-On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination it will be required that extra degrees of freedom are afforded the outlier population.
-Proof of concept simulation fits show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-For the purposes of the DES spectroscopic sample, which will be dominated by confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
-We assume that supernovae fall into either population as determined by their observed classification probability $\hat{p}$.
+On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SN Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population.
+Proof of concept analysis on simple simulation data show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+For the purposes of the DES spectroscopic sample, which will be confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
+We assume that supernovae fall into the SN Ia population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.
 
@@
 \begin{equation}
-\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right]
+\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right].
 \end{equation}
@@
 That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty.
-We include these in our approach, in addition to including HST Calspec calibration uncertainty, 10 SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty.
+We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty.
 This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
@@
 \end{align}
-As the model so far described in previous sections describe components of a basic likelihood $P(D|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
+As our model so far described in previous sections describe components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
 \begin{align}

### 2018-03-12 papers/methods/paper.tex
Anais feedback 2

@@
 
-We test four models: Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM and Flat $w$CDM with a prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, with the latter included to allow sensitive tests on bias for $w$.
+We test four models: Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM, and Flat $w$CDM with a prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, with the latter included to allow sensitive tests on bias for $w$.
 To achieve statistical precision, we fit 100 realisations of supernovae datasets.
@@
 
-Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require more a more sophisticated approach.
+Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
-The first model is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The second model, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
 
 Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys.
-In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and now also include our full treatment of systematics.
+In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics.
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016}.
-The selection effects observed by comparing the generated supernovae to those that pass our simulated cuts and were successfully `observed' are shown in Figure~\ref{fig:sf_snana}, and it is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
+The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
+It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
 We note that there was no significant difference in the selection effect of Malmquist bias between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.
@@
 
-Each realisation of cosmology simulated contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-As our primary focus is upon Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
+Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
 
@@
 For the statistics only fits, the respective bias detection levels are $0.9\sigma$ and $5.5\sigma$.
-We investigate the source of the {\celeven} bias and find its source to be bias in the observed summary statistics, in addition to incorrect reported uncertainty on those summary statistics.
-By replacing fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+We investigate the {\celeven} bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+This is confirmed by replacing the fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
@@
 In this paper we have outlined the creation of a hierarchical Bayesian model for supernova cosmology.
-The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocity uncertainty, calibration uncertainty, dust uncertainty and SALT2 model uncertainty.
-Furthermore, these uncertainties, being explicitly parametrised in our model, are captured will covariance intact, an improvement on many previous methods.
-The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion.
+The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocities, survey calibration, HST calibration, dust, a potential global redshift offset, and SALT2 model uncertainty.
+Furthermore, our uncertainties in standardisation, population, mass-step and more, being explicitly parametrised in our model, are captured will covariance intact, an improvement on many previous methods.
+The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion of other methods.
 

### 2018-03-19 papers/methods/paper.tex
Paper update with lowz citations

@@
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009}, \citet{Mandel2011} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
+\citet{Mandel2009}, \citet{Mandel2011a} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
 \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
@@
 
-Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys.
+Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys, where the LowZ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+
  In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics.

### 2018-03-19 papers/methods/paper.tex
Adding Alex's suggestions for intrinsic dispersion.

@@
 
-The first layer of the hierarchy represents the latent population of $m_B$, $x_1$ and $c$.
-This can be thought of as the `true' values for the supernova, and with correct summary statistics and uncertainty would be normally distributed around the observed values such that
-\begin{equation}
-P(\hat{m}_B,\hat{x}_1,\hat{c}|m_B, x_1, c,\cov) = \mathcal{N}(\lbrace \hat{m}_B,\hat{x}_1,\hat{c} \rbrace|\lbrace m_B, x_1, c \rbrace, \cov). \label{eq:pop}
-\end{equation}
-However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian.
-We model the extra dispersion included in the observations from colour smearing by adding extra independent uncertainty on the colour observation.
-As shown in \citep{Kessler2013}, the extra dispersion shows heavy redshift dependence.
+The first layer of the hierarchy represents the parameters that describe each supernova.
+For the moment, let us consider a single supernova, and represent its parameters as $\theta$.
+The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
+As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$.
+The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$.
+For notational simplify, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
+The model prediction for our derived data is
+\begin{align}
+p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
+\end{align}
+To help confuse things, we often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
+\begin{align}
+p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
+\end{align}
+It is common in analytic supernova cosmology models to take that
+\begin{align}
+p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
+\end{align}
+where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
+However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian and equation \eqref{eq:pop} is violated.
+We address this problem with a similar method to previous analytic likelihoods -- including extra dispersion in the reported covariance matrix $\mathcal{C}$.
+
+We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement over just colour dispersion and was far more computationally inefficient.
+As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence.
 As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
@@
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
-We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controlls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}.
+We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}.
 Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour.

### 2018-03-22 papers/methods/paper.tex
Add line numbers and taking some of alex's feedback into account.

@@
 \newcommand{\angstrom}{\mbox{\normalfont\AA}}
+
+\usepackage{lineno}
+\linenumbers
 
@@
 Specifically, for any given supernova our functional map may take the traditional form:
+
 \begin{equation}
@@
 \end{equation}
+
 where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
@@
 This function is given by
+
 \begin{equation}
@@
 \end{equation}
+
 where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
 The cosmological $\mu_C$ is defined as
+
 \begin{align}
@@
 
-\subsection{Observables Nature Provides: Populations}
+\subsection{Observed Populations}
 
@@
 The model prediction for our derived data is
+
 \begin{align}
@@
 \end{align}
+
 To help confuse things, we often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
+
 \begin{align}
@@
 \end{align}
+
 It is common in analytic supernova cosmology models to take that
+
 \begin{align}
@@
 \end{align}
+
 where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
@@
 Each survey has its own set of $\kappa$ parameters.
+As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
+
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
@@
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.
-The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernovae population for the survey and the systematics associated with it.}
+The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.}
 \label{fig:pgm}
@@
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
+
 \begin{align}
-P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \rho, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
+P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
 \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
@@
 For calculating the distance modulus, we fix $H_0 = 70 \kmsmpc $.
+If the Hubble constant has a different value, the absolute magnitude is $M_B + 5\log(H_0/70 \kmsmpc )$ with the other cosmological parameters unaffected.
 
@@
 In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
+
 \begin{equation}
@@
 \end{equation}
+
 We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes
+
 \begin{align}
-M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot m. \label{eq:l2}
+M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times m. \label{eq:l2}
 \end{align}
@@
 This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
-Denoting $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
+With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
+
 \begin{align}
@@
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
+
 \begin{align}
@@
 \end{align}
+
 As our model so far described in previous sections describe components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
+
 \begin{align}
@@
 \end{align}
+
 where the denominator represents an integral over all potential data.
-Full derivation of this can be found in Appendix \ref{app:selection}.
 As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
-We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently, such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
+Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
+We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
+
 \begin{align}
@@
 \end{align}
-We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well sampled the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
+
+A full derivation of this can be found in Appendix \ref{app:selection}.
+We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
 It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
@@
 The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
+
 \begin{align}
@@
 \end{align}
+
 What then remains is determining the functional form of $P(S|m_B)$.
@@
 Formally, this gives
+
 \begin{align}
@@
 \end{align}
+
 where $\Phi^c$ the complimentary CDF and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
@@
 However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form
+
 \begin{align}
@@
 For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), this correction becomes
+
 \begin{align}
@@
 \end{align}
+
 and is incorporated into our likelihood.

### 2018-04-05 papers/methods/paper.tex
Updating likelihood math notation
Removing todo section
Updating paper from Bobs feedback

@@
 Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae.
-Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
+Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
 SNANA simulations can recover the input cosmology to within $0.27\sigma$, while taking into account many more effects than standard analyses.
@@
 
-{\red Todo: get info on all used lowz surveys and introduce them properly. Help in terms of the right papers to cite would be appreciated.}
-
 Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Balland2009, Amanullah2010}.
-Cosmological analyses of these supernova samples \citep{Kowalski2008, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
+Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
+Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
 For a recent review, see \citet{Huterer2018}.
@@
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
-As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$.
+As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
 
@@
 \mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
-r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}\\
-\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
-\end{align}
+r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
+%\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
+\end{align}
+where $r$ is the comoving distance for redshift $z$ given a specific cosmology and $H_0$ is the current value of Hubble's constant in $\kmsmpc$.
 
@@
 Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation.
+For example, correlation the underlying supernova colour population's mean and skewness is highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
 Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
@@
 Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data.
+Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift when supernovae increasingly fall below the detection threshold.
+An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}.
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{malmquist.pdf}
+	\end{center}
+	\caption{An example of the effects of Malmquist bias.
+Here are shown 1000 simulated supernovae redshifts and distance modulus given fiducial cosmology.
+The simulated survey is magnitude limited, and all supernovae brighter than magnitude 24 are successfully observed (shown as blue dots), and all dimmer than 24th magnitude are not successfully observed (shown as red dots).
+By binning the supernovae along redshift, and taking the mean distance modulus of the supernovae in each bin, we can see that at higher redshift where Malmquist bias kicks in, the population mean drops and becomes bias.
+This source of bias must either be corrected by adjusting the data (such as subtracting the found bias) for by incorporating Malmquist bias explicitly in the cosmological model.}
+	\label{fig:malmquist}
+\end{figure}
+
 \subsection{Approximate Bayesian Computation}
@@
 To try and escape the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
-\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and Flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}.
+\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}.
 In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.
 
-By building the systematic uncertainties and selection effects into the simulation package, there is vastly more freedom in how to treat and model those effects.
+By building the systematic uncertainties and selection effects into the simulation package, there is more freedom in how to treat and model those effects.
 Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood.
@@
 The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion.
-For ABC methods, we must instead simulate an entire supernova population in its entirety -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts.
+For ABC methods, we must instead simulate an entire supernova population -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts.
 This is an intensive process.
-Luckily, efficient sampling algorithms that have walkers that rely on the Markov properties of groups instead of individual walkers, such as Ensemble sampling \citep{Foreman-Mackey2013} allow easy parallelisation of parameter fits and can help ease the computational burden.
 
@@
 This degeneracy resulted in a significant bias in recovered cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
-As such, our calculation of distance modulus $\mu$ mirrors that found in Equations \eqref{eq:mu0} and \eqref{eq:mu1}.
+As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.
 
@@
 \begin{align}
-\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:like}
+\mathcal{L}(\theta; {\rm data}) &= P({\rm data} | \theta, S). \label{eq:like}
 \end{align}
@@
 \begin{align}
-\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
+\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
 \end{align}

### 2018-04-05 dessn/configurations/bulk.py

@@ -58,34 +58,34 @@ if __name__ == "__main__":
         shades = []
         cs = ['b', 'g', 'p', 'a']
         cs = ['b', 'g', 'lb', 'lg']
-        for m, s, ci, chain, truth, weight, old_weight, posterior in res:
-            sim_name = s[0].simulation_name
-            if "MAGSMEAR" in sim_name:
-                name = "Coherent"
-            elif "G10" in sim_name:
-                name = "G10"
-            elif "C11" in sim_name:
-                name = "C11"
-            else:
-                name = sim_name.replace("DES3YR_DES_", "").replace("_", " ").replace("SKEW", "SK16")
-            name = "%s %s" % (name, "Stat" if m.statonly else "Stat+Syst")
-            # if s[0].bias_cor:
-            #     name += " Biascor"
-            if m.statonly:
-                ls.append("--")
-                shades.append(0.0)
-            else:
-                ls.append("-")
-                shades.append(0.2)
-            if isinstance(m, ApproximateModelW):
-                print("C2")
-                c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
-            else:
-                print("C1")
-                c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
-
-        c2.configure(spacing=1.0, sigma2d=False, flip=False, shade=True,
-                     linestyles=ls, colors=cs, shade_alpha=shades)
+        # for m, s, ci, chain, truth, weight, old_weight, posterior in res:
+        #     sim_name = s[0].simulation_name
+        #     if "MAGSMEAR" in sim_name:
+        #         name = "Coherent"
+        #     elif "G10" in sim_name:
+        #         name = "G10"
+        #     elif "C11" in sim_name:
+        #         name = "C11"
+        #     else:
+        #         name = sim_name.replace("DES3YR_DES_", "").replace("_", " ").replace("SKEW", "SK16")
+        #     name = "%s %s" % (name, "Stat" if m.statonly else "Stat+Syst")
+        #     # if s[0].bias_cor:
+        #     #     name += " Biascor"
+        #     if m.statonly:
[…]

### 2018-04-05 papers/methods/snippets/malmquist.py

@@ -1,0 +1,54 @@
+import numpy as np
+from scipy.stats import norm, skewnorm
+import matplotlib.pyplot as plt
+from matplotlib import rc
+from astropy.cosmology import FlatwCDM
+
+cosmo = FlatwCDM(70, 0.3)
+
+zs = np.linspace(0.1, 1.0, 100)
+mus = cosmo.distmod(zs).value
+
+MB = -19.3
+alpha = 0.1
+beta = 3.1
+
+n = 1000
+np.random.seed(5)
+sn_zs = np.random.uniform(0.1, 1.0, size=n)
+sn_mus = cosmo.distmod(sn_zs).value
+sn_MBs = np.random.normal(MB, 0.2, size=n)
+sn_cs = np.random.normal(0, 0.04, size=n)
+sn_x1s = np.random.normal(0, 0.5, size=n)
+sn_mbs = sn_MBs - alpha * sn_x1s + beta * sn_cs + sn_mus
+
+sn_disp = sn_mbs + alpha * sn_x1s - beta * sn_cs - MB
+#mask = np.random.uniform(size=n) <= (1 - norm.cdf(sn_mbs, 24, 0.5))
+mask = sn_mbs < 24
+
+from scipy.stats import binned_statistic
+means, bins, _ = binned_statistic(sn_zs[mask], sn_disp[mask])
+err, bins, _ = binned_statistic(sn_zs[mask], sn_disp[mask], bins=bins, statistic=lambda x: np.std(x)/np.sqrt(len(x)))
+bin_center = 0.5 * (bins[1:] + bins[:-1])
+bias = means - cosmo.distmod(bin_center).value
+
+#rc('text', usetex=True)
+fig, ax = plt.subplots(nrows=1, sharex=True, figsize=(4.5, 3.5), gridspec_kw={"hspace": 0.0})
+
+ax.plot(zs, mus, c='k', ls='--')
+ax.scatter(sn_zs[mask], sn_disp[mask], s=2, c="#4286f4", alpha=0.9, label="Observed SN")
+ax.scatter(sn_zs[~mask], sn_disp[~mask], s=2, c="#f44141", alpha=0.4, label="Unobserved SN")
+
+
+cax = ax.scatter(bin_center, means, c=np.abs(bias), s=50, lw=1, edgecolor='k', label="SN means")
+cbar = fig.colorbar(cax)
+cbar.set_label(r"$\Delta \mu$", fontsize=14)
+
[…]

### 2018-04-09 papers/methods/paper.tex
Responding to more of Bobs feedback

@@
 \usepackage{tikz}
+\usepackage{enumerate}
 \DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
@@
 Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
+Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
 Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
@@
 One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
-\citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, although evaluation of systematic uncertainty was only performed using the Tripp metric.
+\citet{Jennings2016} presents both two metrics (which are used to measure the distance between the forward modelled population and observed population, which is minimised in fitting).
+The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics.
+However, we note that evaluation of systematic uncertainty was only performed using the Tripp metric.
 
@@
 
-Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies hierarchical Bayesian Models.
+Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM).
+Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and physical motivated conditional probabilities.
+For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
+We can thus easily incorporate different distributions, populations, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.
+
 With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009}, \citet{Mandel2011a} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
+\citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
 \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
@@
 We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
-As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.
+As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
+The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS) (a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.
 
@@
 The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
+In the following sections, we will describe these layers and mapping functions.
 
@@
 
+As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
+Potential sources of redshift error (such as peculiar velocities) is taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
+This is discussed further in Section \ref{sec:systreat}.
+Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
+
 The first layer of the hierarchy represents the parameters that describe each supernova.
-For the moment, let us consider a single supernova, and represent its parameters as $\theta$.
+That is, observed parameters are denoted with a hat, whilst the true value is denoted without a hat.
+For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.
+
+For the moment, let us consider a single supernova, and represent its true parameters as $\theta$.
 The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
@@
 where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
-However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian and equation \eqref{eq:pop} is violated.
-We address this problem with a similar method to previous analytic likelihoods -- including extra dispersion in the reported covariance matrix $\mathcal{C}$.
+Unfortunately, this assumption of Gaussianity is incorrect.
+The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion, wherein some supernovae are uniformly brighter or dimmer than the population mean, and in chromatic dispersion, whereby the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
+The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated.
+We address this problem with a similar method to previous analytic likelihoods: including extra dispersion in the reported covariance matrix $\mathcal{C}$.
+Whilst this doesn't solve the issue of non-Gaussianity, it does address the under-reporting of uncertainty, which is the primary source of bias.
 
@@
 We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
-The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
-Each survey has its own set of $\kappa$ parameters.
+The $\kappa$ parameters are tightly constrained as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$, and the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
+The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulted in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion  Each survey has its own set of $\kappa$ parameters.
 As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
@@
 Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
+With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
 We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}.
@@
 We place a uniform prior on $s_c$.
-
-As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
+It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
 
@@
 Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
-We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$.
+We parameterise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
 The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.
 
-As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means and width fit individually for each survey.
+As such, the only constant between survey populations is the mean absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
@@
 
-On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SN Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+On top of the SNIa populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SNIa population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
 With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population.
-Proof of concept analysis on simple simulation data show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-For the purposes of the DES spectroscopic sample, which will be confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$.
-We assume that supernovae fall into the SN Ia population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.
+This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification and incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification  \citep{Gupta2016}.
+Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$.
+We assume that supernovae fall into the SNIa population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.
 
@@
 The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
-The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses.
+The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step.
 The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
@@
 \begin{equation}
-\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right].
+\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
 \end{equation}
 
-We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes
-
-\begin{align}
-M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times m. \label{eq:l2}
+where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift.
+We take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
+With this correction, our calculation of absolute magnitude becomes
+
+\begin{align}
+M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times {\rm mass\_prob}. \label{eq:l2}
 \end{align}
@@
 
-where the denominator represents an integral over all potential data.
+where the denominator represents an integral over all potential data $D$.
 As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
-We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as
-
-\begin{align}
-w &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
+Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
+We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as
+
+\begin{align}
+d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
 \end{align}
@@
 
-where $\Phi^c$ the complimentary CDF and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
+where $\Phi^c$ the complimentary cumulative distribution function (CDF) and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
 The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
@@
 Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit.
-Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
+Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
 
 With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
-For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), this correction becomes
-
-\begin{align}
-w_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
-w_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
+For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SNIa} becomes
+
+\begin{align}
+d_{\rm CDF/SNIa} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
+d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
 &\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
@@
 
-and is incorporated into our likelihood.
+and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency} for aid in visualisation.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{efficiency.pdf}
+	\end{center}
+	\caption{The efficiency for supernova discovery at an arbitrary redshift.
+Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution.
+The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), but as generic functions of apparent magnitude.
+The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
+This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellowz}.}
+	\label{fig:efficiency}
+\end{figure}
 
@@
 
-We generate simple toy data to validate the basic premise of the model.
+We generate simple toy data to validate the basic premise of the model. Data generation algorithm is described below:
+
+\begin{enumerate}[1.]
+	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
+	\item Draw a true absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
+	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) = 0.5$ and equation \eqref{eq:mass}.
+	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$.
+Use this to determine the true apparent magnitude of the object $m_B$ using equation \eqref{eq:l2}.
+	\item Determine if the SNIa is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
+For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$.
+Repeat from step one until we have a supernova which passes.
+	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
+Add extra colour uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
+\end{enumerate}
+
 For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
@@
 \begin{align}
-w &= \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
+d &= \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
 \end{align}
@@
 \begin{align}
-w = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
+d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
 \end{align}

### 2018-04-09 dessn/configurations/plot_pop_simple.py

@@ -1,9 +1,11 @@
 import os
 
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
 from dessn.framework.simulations.simple import SimpleSimulation
 from astropy.cosmology import FlatwCDM
 import matplotlib.pyplot as plt
-from matplotlib import rc
+from matplotlib import rc, ticker
 
 if __name__ == "__main__":
 
@@ -17,9 +19,10 @@ if __name__ == "__main__":
 
     cosmo = FlatwCDM(70, 0.3)
     rc('text', usetex=True)
-    fig, axes = plt.subplots(nrows=len(sims), figsize=(4, 5))
+    fig, axes = plt.subplots(nrows=len(sims), figsize=(4.5, 5))
 
     for ax, sim, n in zip(axes, sims, names):
+        # ax, ax2 = axs
         data = sim.get_all_supernova(1000)
         obs_mBx1c = data["obs_mBx1c"]
         mbs, x1s, cs = obs_mBx1c[:, 0], obs_mBx1c[:, 1], obs_mBx1c[:, 2]
@@ -28,14 +31,23 @@ if __name__ == "__main__":
         passed = data["passed"]
         mus = cosmo.distmod(zs).value
         MBs = mbs - mus
-        ax.scatter(zs[passed], MBs[passed], c=cs[passed], s=10, vmin=-0.2, vmax=0.4, alpha=1, cmap="plasma")
+        h1 = ax.scatter(zs[passed], MBs[passed], c=cs[passed], s=10, vmin=-0.2, vmax=0.4, alpha=1, cmap="plasma")
+        divider = make_axes_locatable(ax)
+        cax = divider.append_axes('right', size='3%', pad=0.05)
+        fig.colorbar(h1, cax=cax, orientation='vertical', label="Colour")
+        # ax2.scatter(zs[passed], cs[passed], c=cs[passed], s=10, vmin=-0.2, vmax=0.4, alpha=1, cmap="plasma")
 
         ax.set_xlabel("$z$", fontsize=14)
         ax.set_ylabel(r"$m_B - \mu$", fontsize=14)
         ax.text(0.98, 0.95, n, verticalalignment='top', horizontalalignment='right', transform=ax.transAxes)
         ax.set_xlim(zs.min() - 0.001, zs.max() + 0.001)
-    plt.locator_params(numticks=5)
+        # ax2.set_xlabel("$z$", fontsize=14)
[…]

### 2018-04-09 papers/methods/snippets/efficiency.py

@@ -1,0 +1,52 @@
+import numpy as np
+from scipy.stats import norm, skewnorm
+import matplotlib.pyplot as plt
+from matplotlib import rc
+from astropy.cosmology import FlatwCDM
+
+mbs = np.linspace(19, 27, 100)
+
+pop = norm.pdf(mbs, 22, 1)
+pop /= pop.max()
+
+cdf_mu, cdf_sigma = 21.5, 0.5
+cdf = 1 - norm.cdf(mbs, cdf_mu, cdf_sigma)
+cdf_eff = pop * cdf
+
+skew_mu, skew_sigma, skew_alpha = 20, 1, 5
+skew = skewnorm.pdf(mbs, skew_alpha, skew_mu, skew_sigma)
+skew /= skew.max()
+skew_eff = pop * skew
+
+rc('text', usetex=True)
+fig, axes = plt.subplots(nrows=2, sharex=True, figsize=(4.5, 4.5),gridspec_kw={"hspace": 0.08})
+
+axes[0].plot(mbs, pop, label="SNIa pop: $P(m_B|\\theta)$", ls="--")
+axes[0].plot(mbs, cdf, label="CDF eff: $P(S| m_B)$", c="g", ls=":")
+axes[0].plot(mbs, cdf_eff, label="$P(S| m_B)P(m_B|\\theta)$", c="g")
+#axes[0].axhline(0, c='k')
+axes[0].fill_between(mbs, 0, cdf_eff, color='g', alpha=0.2)
+axes[0].text(20.6, 0.14, "$d_{\\rm CDF}$", fontsize=14)
+
+axes[1].plot(mbs, pop, label="SNIa pop: $P(m_B|\\theta)$", ls="--")
+axes[1].plot(mbs, skew, label="Skew eff: $P(S| m_B)$", c="r", ls=":")
+axes[1].plot(mbs, skew_eff, label="$P(S| m_B)P(m_B|\\theta)$", c="r")
+#axes[1].axhline(0, c='k')
+axes[1].fill_between(mbs, 0, skew_eff, color='r', alpha=0.2)
+axes[1].text(20.6, 0.13, "$d_{\\rm Skew}$", fontsize=14)
+
+axes[0].legend(loc=1, markerfirst=False, frameon=False)
+axes[1].legend(loc=1, markerfirst=False, frameon=False)
+axes[1].set_xlabel("$m_B$")
+axes[0].set_ylabel("Probability")
[…]

### 2018-04-09 papers/methods/paper.tex
Responding to more of Bobs feedback

@@
 
-For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$.
-Mass probabilities are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included.
-Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$.
-Redshifts are drawn from a power law to increase the number of events as redshift increases.
-
-Independent observational errors of $0.04$, $0.2$, $0.03$ on $m_B$, $x_1$ and $c$ (following the mean uncertainty for DES SNANA simulations) are added to create the observables.
-Extra dispersion of $0.02(1+3z)$ is added in quadrature to the $c$ uncertainty to simulate unexplained colour dispersion from chromatic variation in supernovae.
-The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
-The results are then passed through the selection effects, where each supernova is only selected based on $P(S|m_B)$, using a skew normal function for the LowZ supernovae and error function for the DES-like supernovae.
+For supernova independent uncertainty, the selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
 We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES spectroscopic analysis.

### 2018-04-10 papers/methods/paper.tex
Separating out bulk plots for Bob

@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{bulk_small2.pdf}
+		\includegraphics[width=\columnwidth]{bulk_small_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_small_c11.pdf}
 	\end{center}
-	\caption{Posterior surfaces for 100 realisations of supernova data.
+	\caption{Posterior surfaces for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel.
 Surfaces for $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$ are shown, with the other fit parameters being marginalised over.

### 2018-04-10 dessn/configurations/bulk.py

@@ -57,38 +57,36 @@ if __name__ == "__main__":
         ls = []
         shades = []
         cs = ['b', 'g', 'p', 'a']
-        cs = ['b', 'g', 'lb', 'lg']
-        # for m, s, ci, chain, truth, weight, old_weight, posterior in res:
-        #     sim_name = s[0].simulation_name
-        #     if "MAGSMEAR" in sim_name:
-        #         name = "Coherent"
-        #     elif "G10" in sim_name:
-        #         name = "G10"
-        #     elif "C11" in sim_name:
-        #         name = "C11"
-        #     else:
-        #         name = sim_name.replace("DES3YR_DES_", "").replace("_", " ").replace("SKEW", "SK16")
-        #     name = "%s %s" % (name, "Stat" if m.statonly else "Stat+Syst")
-        #     # if s[0].bias_cor:
-        #     #     name += " Biascor"
-        #     if m.statonly:
-        #         ls.append("--")
-        #         shades.append(0.0)
-        #     else:
-        #         ls.append("-")
-        #         shades.append(0.2)
-        #     if isinstance(m, ApproximateModelW):
-        #         print("C2")
-        #         c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
-        #     else:
-        #         print("C1")
-        #         c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
-        #
-        # c2.configure(spacing=1.0, sigma2d=False, flip=False, shade=True,
-        #              linestyles=ls, colors=cs, shade_alpha=shades)
+        cs = ['g', 'r', 'k', 'k']
+        for m, s, ci, chain, truth, weight, old_weight, posterior in res:
+            sim_name = s[0].simulation_name
+            if "MAGSMEAR" in sim_name:
+                name = "Coherent"
+            elif "G10" in sim_name:
+                name = "G10"
+            elif "C11" in sim_name:
+                name = "C11"
+            else:
+                name = sim_name.replace("DES3YR_DES_", "").replace("_", " ").replace("SKEW", "SK16")
+            name = "%s %s" % (name, "Stat" if m.statonly else "Stat+Syst")
+            # if s[0].bias_cor:
+            #     name += " Biascor"
+            if m.statonly:
+                ls.append("--")
[…]

### 2018-04-26 dessn/configurations/lock_simple.py
Slight update before I head off. Need to update the plots using the new features in ChainConsumer

@@ -64,7 +64,7 @@ if __name__ == "__main__":
             else:
                 name += "Stat+Syst"
             names.append(name)
-            c.add_chain(chain, weights=weight, posterior=posterior, name=name)
+            c.add_chain(chain, weights=weight, posterior=posterior, name=name, power=2.0)
 
         # c.plotter.plot(filename=[pfn + "_small_c11.png", pfn + "_small_c11.pdf"], parameters=4, truth=truth, figsize=1.0, chains=[n for n in names if "C11" in n])
         # c.plotter.plot(filename=[pfn + "_small_g10.png", pfn + "_small_g10.pdf"], parameters=4, truth=truth, figsize=1.0, chains=[n for n in names if "G10" in n])

### 2018-04-26 dessn/configurations/simple.py

@@ -4,6 +4,7 @@ import socket
 from dessn.framework.fitter import Fitter
 from dessn.framework.models.approx_model import ApproximateModel, ApproximateModelOl, ApproximateModelW
 from dessn.framework.simulations.simple import SimpleSimulation
+import matplotlib.pyplot as plt
 
 if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG)
@@ -36,23 +37,41 @@ if __name__ == "__main__":
         from chainconsumer import ChainConsumer
 
         res = fitter.load(split_models=True, squeeze=False, split_cosmo=False)
+        res2 = fitter.load(split_models=True, squeeze=False, split_cosmo=True)
 
         print("Adding data")
         cb = ChainConsumer()
         names = [r"Flat $\Lambda$CDM", r"$\Lambda$CDM", "Flat $w$CDM", r"Flat $w$CDM + $\Omega_m$ prior"]
         for i, (m, s, ci, chain, truth, weight, old_weight, posterior) in enumerate(res):
-            c = ChainConsumer()
-            c.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
-            if i in [0, 3]:
-                cb.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
-            c.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, sigmas=[0, 0.3, 1, 2], contour_labels="confidence")
+            # c = ChainConsumer()
+            # c.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
+            # if i in [0, 3]:
+            #     cb.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
+            #
             ps = m.get_cosmo_params()
             n = m.__class__.__name__
             if m.prior:
                 n += "OmPrior"
-            c.plotter.plot(filename=[pfn + n + ".png", pfn + n + ".pdf"], parameters=ps, truth=truth, figsize="column")
 
-        print("Saving table")
-        print(cb.analysis.get_latex_table(transpose=True))
-        with open(pfn + "_cosmo_params.txt", 'w') as f:
-            f.write(cb.analysis.get_latex_table(parameters=parameters))
+            if len(ps) == 2:
+                c2 = ChainConsumer()
+                i = 0
+                for m2, s2, ci, chain, _, weight, _, _ in res2:
+                    if m2 == m and s2 == s:
[…]

### 2018-07-04 papers/methods/paper.tex
Slight update before I head off. Need to update the plots using the new features in ChainConsumer
Updating from some of Dans feedback

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\name}{Steve-o}
+\newcommand{\name}{Steve}
 \newcommand{\myemail}{samuelreay@gmail.com}
@@
 We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects.
-This advances the work of \citet{Rubin2015} by including additional sources of systematic uncertainty and improved treatment of Malmquist bias whilst increasing numerical efficiency.
-Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae.
+This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improved treatment of Malmquist bias, whilst increasing numerical efficiency.
+Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery.
 Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-SNANA simulations can recover the input cosmology to within $0.27\sigma$, while taking into account many more effects than standard analyses.
+SNANA simulations can recover the input cosmology to within $0.27\sigma$, while taking into an increased number of systematics than prior analyses.
 \end{abstract}
@@
 Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
+Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
 Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
@@
 In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} will be observing thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples.
-The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will produce scores of thousands of photometrically classified supernovae.
-Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainty will prove to be the limiting factor in our analyses.
-
-As such, staggering effort is being put into developing more sophisticated supernovae analyses.
-The role of simulations has become increasingly important in determining biases and validating models.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples.
+The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
+Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainties will prove to be the limiting factor in our analyses \citep{Betoule2014}.
+
+As such, staggering effort is being put into developing more sophisticated supernova cosmology analyses.
+The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models.
 First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
-\citet{Betoule2014} make use of redshift dependent $\mu$-bias to correct their analyses, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation corrections as a function of redshift, stretch and colour to better remove fitting biases.
+\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation to quantify observational bias in as a function of many factors to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
@@
 
-In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects whilst improving computational performance.
+In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance.
 Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
@@
 
-Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements.
-Instead, most techniques fit these observations (along with redshift) to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
+Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly.
+Instead, most techniques fit and observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
 This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}.
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
-As such, the product at the end is a (redshift-dependent) population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
-
-The underlying actual supernova population is not such a clear cut, and accurately characterising this population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
-However, given some modelled underlying population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
+As all supernova are not identical, an ensemble of supernova form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
+
+The population previous discussed represents an observed population, which -- due to the presence of various selection effects -- may not represent the underlying supernova population.
+Accurately characterising this population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
+However, given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
 Specifically, for any given supernova our functional map may take the traditional form:
@@
 \begin{equation}
-M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{corrections}, \label{eq:standard}
+M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{other corrections}, \label{eq:standard}
 \end{equation}
 
-where $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader and bluer supernovae are brighter.
-The $\text{corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
-The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and an input redshift.
+where $M_B$ is the mean absolute magnitude for all SN Ia  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
+The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and a redshift.
+The $\text{other corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017} In traditional analyses, the other corections term often includes bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
 
@@
 Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty.
-Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about uncertainty correlation.
-For example, correlation the underlying supernova colour population's mean and skewness is highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
+Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
+For example, the underlying supernova colour population's mean and skewness is highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
 Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
+A common solution to this computational cost problem is to bin the supernova.
 
@@
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
-As these effects are not built into the likelihood, their influence on the error budget is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
+As these effects are not built into the likelihood and instead are formed by correcting data, their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
 
@@
 
-By building the systematic uncertainties and selection effects into the simulation package, there is more freedom in how to treat and model those effects.
-Data does not need to be corrected, analytic approximations do not need to be applied, we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood.
+Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects that plague supernovae surveys.
+By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood.
 This freedom comes with a cost -- computation.
@@
 The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision.
-Their model was validated on thirty realisations of statistics.
+Their model was validated to be free of significant biases using fits to thirty realisations of statistics.
 
@@
 We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
+We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
@@
 The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
-In the following sections, we will describe these layers and mapping functions.
+In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
 
@@
 The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$.
-For notational simplify, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
+For notational simplicity, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
 The model prediction for our derived data is
@@
 With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
-We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}.
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour.
+We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations.
+The scatter models, sourced from \citet{Guy2010} and \citet{Chotard2011} are discussed in detail in Section \ref{sec:simdes}.
+The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}.
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
 We place a uniform prior on $s_c$.
@@
 
-The underlying supernova population is characterised by underlying distributions in colour, stretch and magnitude.
+The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude.
 Analytic models often treat the colour and stretch populations independently with a skew normal and normal, respectively.
@@
 Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes.
+Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
@@
 This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification and incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification  \citep{Gupta2016}.
-Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination, both from non type Ia supernovae and from catastrophic failure, is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
 For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$.
@@
 
-It is now well known that host galaxy environment has a significant effect on supernova properties.
+It is now well known that host galaxy environment is significantly correlated with supernova properties.
 The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
@@
 
-\begin{equation}
+\begin{align}
 \Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
-\end{equation}
+\end{align}
 
@@
 
-This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses, however it is normally constrained to band systematics.
-That is, each band for each survey has two systematics associated with it - the calibration uncertainty and the filter wavelength uncertainty.
-We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic and also the systematic peculiar velocity uncertainty.
+This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses.
+For each survey and band, we have two systematics --  the calibration uncertainty and the filter wavelength uncertainty.
+We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
 This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
@@
 		\hline
-		Flat $\Lambda$CDM & $0.301\pm 0.016$ & -- \\ 
-		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.5^{+10.5}_{-9.8} \right) \times 10^{-3}$ & $-0.998^{+0.043}_{-0.047}$ \\ 
+		Flat $\Lambda$CDM & $0.302\pm 0.015$ & -- \\ 
+		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.9\pm7.4 \right) \times 10^{-3}$ & $-0.996\pm 0.038$ \\ 
 		\hline

### 2018-07-11 papers/methods/paper.tex
Updating plots, paper and data

@@
 \begin{keywords}
-keyword1 -- keyword2 -- keyword3
+keyword1 -- keyword2 -- keyword3\vspace{-5mm}
 \end{keywords}
@@
 
-\begin{equation}
+\begin{align}
 M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{other corrections}, \label{eq:standard}
-\end{equation}
+\end{align}
 
@@
 
-\begin{equation}
+\begin{align}
 \chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
-\end{equation}
+\end{align}
 
@@
 r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
-%\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
 \end{align}
@@
 	\end{center}
-	\caption{Posterior surfaces for 100 realisations of supernova data with the $\Lambda$CDM model.
+	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context.
 Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
@@
 	\end{center}
-	\caption{Posterior surfaces for 100 realisations of supernova data with the Flat $w$CDM model.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface, where the point attained from each parameters independent maximum likelihood lands well outside the $3\sigma$ contour in parameter space.}
+	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context.
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface.
+The recovered posteior maximums show the same degeneracy direction as the representative surface and scatter around the truth values input into the simulation, shown in dashed lines.}
 	\label{fig:simple_w}
@@
 	\centering
-	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 100 fits to independent realisations of toy supernova data.
+	\caption{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data.
 As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
+For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviations in the table.
+The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets.
 Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
-As our surfaces are stacked, a shift away from the simulated values of more than $\sigma/\sqrt{100} = 0.1\sigma$ would give a $1\sigma$ detection of bias.
-No significant bias is detected in either cosmological model.}
+As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted standard deviation: $\sqrt{100} = 10$.
+As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model (which was run with a prior on $\Omega_m$).}
 	\label{tab:simple_model}
@@
 		\hline
-		Model & $\Omega_m$ & $w$ \\ 
+		Model & $\Omega_m$(scatter) & $w$(scatter) \\ 
 		\hline
-		Flat $\Lambda$CDM & $0.302\pm 0.015$ & -- \\ 
-		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.9\pm7.4 \right) \times 10^{-3}$ & $-0.996\pm 0.038$ \\ 
+		Flat $\Lambda$CDM & $0.301\pm 0.015(0.012)$ & -- \\ 
+		Flat $w$CDM & -- & $-1.00\pm 0.042(0.030)$ \\ 
 		\hline
@@
 	\begin{center}
-		\includegraphics[width=\columnwidth]{bulk_small_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_small_c11.pdf}
+		\includegraphics[width=\columnwidth]{bulk_points_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_points_c11.pdf}
 	\end{center}

### 2018-07-11 dessn/configurations/bulk.py

@@ -31,10 +31,7 @@ if __name__ == "__main__":
     ]
     fitter = Fitter(dir_name)
 
-    # data = models[0].get_data(simulations[0], 0, plot=True)  # For testing
-    # data2 = models[0].get_data(simulations[1], 0, plot=True)  # For testing
-    # print(data["shift_deltas"])
-    # exit()
+    # data = models[0].get_data(simulations[0], 0, plot=False)  # For testing
 
     fitter.set_models(*models)
     fitter.set_simulations(*simulations)
@@ -49,7 +46,7 @@ if __name__ == "__main__":
         fitter.fit(file)
     else:
         from chainconsumer import ChainConsumer
-        res = fitter.load(split_models=True, split_sims=True, squeeze=False)
+        res = fitter.load(split_models=True, split_sims=True, split_cosmo=True, squeeze=False)
         # res2 = fitter.load(split_models=True, split_sims=False)
 
         c1, c2, c3 = ChainConsumer(), ChainConsumer(), ChainConsumer()
@@ -59,12 +56,17 @@ if __name__ == "__main__":
         cs = ['g', 'r', 'k', 'k']
         for m, s, ci, chain, truth, weight, old_weight, posterior in res:
             sim_name = s[0].simulation_name
+            col = "k"
+            ms = "+"
+            msize = 25
             if "MAGSMEAR" in sim_name:
                 name = "Coherent"
             elif "G10" in sim_name:
                 name = "G10"
+                col = "g"
             elif "C11" in sim_name:
                 name = "C11"
+                col = "r"
             else:
                 name = sim_name.replace("DES3YR_DES_", "").replace("_", " ").replace("SKEW", "SK16")
             name = "%s %s" % (name, "Stat" if m.statonly else "Stat+Syst")
@@ -73,19 +75,32 @@ if __name__ == "__main__":
             if m.statonly:
                 ls.append("--")
                 shades.append(0.0)
+                col = "k"
+                ms = "^"
+                msize = 5
             else:
                 ls.append("-")
                 shades.append(0.7)
             if isinstance(m, ApproximateModelW):
-                c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
+                if "C11" in name:
[…]

### 2018-07-11 dessn/configurations/simple.py

@@ -6,6 +6,8 @@ from dessn.framework.models.approx_model import ApproximateModel, ApproximateMod
 from dessn.framework.simulations.simple import SimpleSimulation
 import matplotlib.pyplot as plt
 
+from dessn.general.helper import weighted_avg_and_std
+
 if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG)
     plot_dir = os.path.dirname(os.path.abspath(__file__)) + "/plots/%s/" % os.path.basename(__file__)[:-3]
@@ -35,6 +37,9 @@ if __name__ == "__main__":
         parameters = [r"$\Omega_m$", "$w$"]
 
         from chainconsumer import ChainConsumer
+        import numpy as np
+        # print(ChainConsumer.__version__)
+        # exit()
 
         res = fitter.load(split_models=True, squeeze=False, split_cosmo=False)
         res2 = fitter.load(split_models=True, squeeze=False, split_cosmo=True)
@@ -42,6 +47,14 @@ if __name__ == "__main__":
         print("Adding data")
         cb = ChainConsumer()
         names = [r"Flat $\Lambda$CDM", r"$\Lambda$CDM", "Flat $w$CDM", r"Flat $w$CDM + $\Omega_m$ prior"]
+
+        ws = []
+        ws_std = []
+        oms = []
+        oms_std = []
+        oms2 = []
+        oms2_std = []
+
         for i, (m, s, ci, chain, truth, weight, old_weight, posterior) in enumerate(res):
             # c = ChainConsumer()
             # c.add_chain(chain, weights=weight, posterior=posterior, name=names[i])
@@ -53,7 +66,22 @@ if __name__ == "__main__":
             if m.prior:
                 n += "OmPrior"
 
-            if len(ps) == 2:
+            name = names[i]
+            for m2, s2, ci, chain, _, weight, _, pp in res2:
+                if m2 == m and s2 == s:
+                    if name == r"Flat $w$CDM + $\Omega_m$ prior":
+                        w_avg, w_std = weighted_avg_and_std(chain["$w$"], weight)
+                        ws.append(w_avg)
+                        ws_std.append(w_std)
+                        om2_avg, om2_std = weighted_avg_and_std(chain["$\Omega_m$"], weight)
+                        oms2.append(om2_avg)
+                        oms2_std.append(om2_std)
+                    elif name == r"Flat $\Lambda$CDM":
[…]

### 2018-07-11 dessn/framework/fitter.py

@@ -196,7 +196,11 @@ class Fitter(object):
                         else:
                             temp_list.append([mapping[p] % i, vals[:, i]])
                         if truth.get(mapping[p]) is not None:
-                            truth[mapping[p] % i] = truth[mapping[p]][i]
+                            if len(truth[mapping[p]]) <= i:
+                                self.logger.warning("Truth values don't line up for %s %d" % (p, i))
+                                truth[mapping[p] % i] = 0
+                            else:
+                                truth[mapping[p] % i] = truth[mapping[p]][i]
                     if truth.get(mapping[p]) is not None:
                         del truth[mapping[p]]
                 else:

### 2018-07-11 dessn/general/helper.py

@@ -1,0 +1,12 @@
+import numpy as np
+
+
+def weighted_avg_and_std(values, weights):
+    """
+    Return the weighted average and standard deviation.
+
+    values, weights -- Numpy ndarrays with the same shape.
+    """
+    average = np.average(values, weights=weights)
+    variance = np.average((values-average)**2, weights=weights)
+    return average, np.sqrt(variance)

### 2018-07-11 dessn/snana/convert_snana_data.py

@@ -333,14 +333,14 @@ if __name__ == "__main__":
     # convert("DES3YR_DES_COMBINED_TEXT")
     # convert("DES3YR_DES_NOMINAL")
     # convert("DES3YR_LOWZ_NOMINAL")
-    # convert("DES3YR_DES_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
+    convert("DES3YR_DES_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
     # convert("DES3YR_LOWZ_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
     # convert("DES3YR_DES_BULK", skip=6)
     # convert("DES3YR_LOWZ_BULK", skip=6)
     # convert("DES3YR_DES_SAMTEST", skip=11)
     # convert("DES3YR_LOWZ_SAMTEST", skip=11)
     # convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11, zipped=False)
-    convert("DES3YR_LOWZ_BHMEFF", load_dump=True, skip=11, zipped=True)
+    # convert("DES3YR_LOWZ_BHMEFF", load_dump=True, skip=11, zipped=True)
     # convert("DES3YR_LOWZ_VALIDATION", skip=6)
     # convert("DES3YR_DES_VALIDATION", skip=6)
 

### 2018-07-12 papers/methods/paper.tex
Adding lowz bulk sims

@@
 	\end{center}
-	\caption{Posterior surfaces for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel.
-Surfaces for $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$ are shown, with the other fit parameters being marginalised over.
+	\caption{Maximum posterior points for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel.
+Points are shown for parameters $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$, with the other fit parameters being marginalised over.
 The intrinsic scatter model has significant impact on the recovered $\beta$, which in itself effects cosmology, resulting in small biases in $w$.}
@@
 	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
-Whilst the {\gten} models do not show significant evidence of bias, the {\celeven} model does.
-With the current statistical size of the supernovae dataset, the bias is sub-dominant to the size of the uncertainty.
+With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$.
+However, their deviation from the truth value represents a shift of approximately $0.3\sigma$ when taking into account the uncertainty on fits to $w$.
+The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
 }
 	\label{tab:bulk_summary}
-	\begin{tabular}{l|cc}
+	\begin{tabular}{l|c}
 		\hline
-		Model & $w$ & Bias Detection Level \\
+		Model & $w$ (scatter) \\
 		\hline
-		{\gten} Stat + Syst  &  $-1.02\pm 0.13$ & $1.5\sigma$  \\
-		{\celeven} Stat + Syst  &  $-0.97\pm 0.11$ & $2.7\sigma$  \\
-		{\gten}  Stat         &  $-1.01\pm 0.11$ & $0.9\sigma$  \\
-		{\celeven} Stat         &  $-0.95\pm 0.09$ & $5.5\sigma$  \\
+		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$   \\
+		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$  \\
+		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$   \\
+		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$   \\
 		\hline

### 2018-07-12 dessn/configurations/bulk.py

@@ -2,10 +2,10 @@ import os
 import logging
 import socket
 
-
 from dessn.framework.fitter import Fitter
 from dessn.framework.models.approx_model import ApproximateModelW, ApproximateModel
 from dessn.framework.simulations.snana import SNANASimulation
+from dessn.general.helper import weighted_avg_and_std
 
 if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG, format="[%(funcName)20s()] %(message)s")
@@ -46,6 +46,8 @@ if __name__ == "__main__":
         fitter.fit(file)
     else:
         from chainconsumer import ChainConsumer
+        import numpy as np
+
         res = fitter.load(split_models=True, split_sims=True, split_cosmo=True, squeeze=False)
         # res2 = fitter.load(split_models=True, split_sims=False)
 
@@ -54,6 +56,9 @@ if __name__ == "__main__":
         shades = []
         cs = ['b', 'g', 'p', 'a']
         cs = ['g', 'r', 'k', 'k']
+
+        ws = {}
+        ws_std = {}
         for m, s, ci, chain, truth, weight, old_weight, posterior in res:
             sim_name = s[0].simulation_name
             col = "k"
@@ -86,21 +91,34 @@ if __name__ == "__main__":
                     cc = c3
                 else:
                     cc = c2
+                if ws.get(name) is None:
+                    ws[name] = []
+                if ws_std.get(name) is None:
+                    ws_std[name] = []
+                w_mean, w_std = weighted_avg_and_std(chain["$w$"], weight)
+                ws[name].append(w_mean)
+                ws_std[name].append(w_std)
                 cc.add_chain(chain, weights=weight, posterior=posterior, name=name, plot_contour=False, plot_point=True,
                            color=col, marker_style=ms, marker_size=msize)
             else:
                 c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
 
+        for k in ws.keys():
+            print("%s %5.3f %5.3f (%5.3f) : %4.2f" % (k, np.mean(ws[k]), np.std(ws[k]), np.mean(ws_std[k]), np.sqrt(100)*(-1-np.mean(ws[k]))/np.std(ws[k])))
[…]

### 2018-07-12 dessn/snana/convert_snana_data.py

@@ -333,10 +333,10 @@ if __name__ == "__main__":
     # convert("DES3YR_DES_COMBINED_TEXT")
     # convert("DES3YR_DES_NOMINAL")
     # convert("DES3YR_LOWZ_NOMINAL")
-    convert("DES3YR_DES_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
+    # convert("DES3YR_DES_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
     # convert("DES3YR_LOWZ_BULK", skip=6, biascor="SALT2mu_DES_BULK+LOWZ_BULK")
     # convert("DES3YR_DES_BULK", skip=6)
-    # convert("DES3YR_LOWZ_BULK", skip=6)
+    convert("DES3YR_LOWZ_BULK", skip=6)
     # convert("DES3YR_DES_SAMTEST", skip=11)
     # convert("DES3YR_LOWZ_SAMTEST", skip=11)
     # convert("DES3YR_DES_BHMEFF", load_dump=True, skip=11, zipped=False)

### 2018-07-12 papers/methods/paper.tex
Updating paper

@@
 Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-SNANA simulations can recover the input cosmology to within $0.27\sigma$, while taking into an increased number of systematics than prior analyses.
+SNANA simulations can recover the input cosmology to within $0.3\sigma$, while taking into an increased number of systematics than prior analyses.
 \end{abstract}
@@
 	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
+The quoted value for $w$ represents the average mean of the fits, with the uncertainty being the average standard deviation of the fits.
+In brackets, the simulation scatter (the standard deviation of the mean of 100 fits) is shown, and the bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.
 With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$.
@@
 	\label{tab:bulk_summary}
-	\begin{tabular}{l|c}
+	\begin{tabular}{l|cc}
 		\hline
-		Model & $w$ (scatter) \\
+		Model & $w$ (scatter) & Bias Significance \\
 		\hline
-		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$   \\
-		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$  \\
-		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$   \\
-		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$   \\
+		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $3.9\sigma$  \\
+		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $4.4\sigma$  \\
+		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $2.0\sigma$   \\
+		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $7.9\sigma$   \\
 		\hline
@@
 
- Combined posterior surfaces for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-Note that we have combined the posteriors for 100 realisations, and so we should expect the size of the uncertainty to be representative of one realisation, but the statistical spread of the final surface should be $\sqrt{100} = 10$ times less than a single realisation.
-The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-
-Cosmological bias is detected with $\approx 2.7\sigma$ significance for the {\celeven} model and $\approx 1.5\sigma$ significance for the {\gten} model when fitting both systematics and statistics.
-For the statistics only fits, the respective bias detection levels are $0.9\sigma$ and $5.5\sigma$.
-We investigate the {\celeven} bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds are listed in Table \ref{tab:bulk_summary}.
+
+Cosmological bias is detected with a confidence of more than $3\sigma$ for the {\gten} and {\celeven} models.
+We investigate the bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
 This is confirmed by replacing the fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.

### 2018-07-12 dessn/configurations/bulk.py

@@ -104,7 +104,7 @@ if __name__ == "__main__":
                 c1.add_chain(chain, weights=weight, posterior=posterior, name=name)
 
         for k in ws.keys():
-            print("%s %5.3f %5.3f (%5.3f) : %4.2f" % (k, np.mean(ws[k]), np.std(ws[k]), np.mean(ws_std[k]), np.sqrt(100)*(-1-np.mean(ws[k]))/np.std(ws[k])))
+            print("%s %5.3f %5.3f (%5.3f) : %4.2f" % (k, np.mean(ws[k]), np.mean(ws_std[k]), np.std(ws[k]), np.sqrt(100)*(-1-np.mean(ws[k]))/np.std(ws[k])))
         # c2.configure(spacing=1.0, sigma2d=False, flip=False, shade=True, linestyles=ls, colors=cs, shade_gradient=1.4, shade_alpha=shades, linewidths=1.2)
         # c2.plotter.plot_summary(filename=[pfn + "2.png", pfn + "2.pdf"], parameters=["$w$"], truth=[-1.0], figsize=1.5, errorbar=True)
         # c2.plotter.plot(filename=[pfn + "_small.png", pfn + "_small.pdf"], parameters=2, truth=truth, extents={"$w$": (-1.4, -0.7)}, figsize="column")

### 2018-07-12 papers/methods/paper.tex
Updating paper

@@
 Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-SNANA simulations can recover the input cosmology to within $0.3\sigma$, while taking into an increased number of systematics than prior analyses.
+SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into an increased number of systematics than prior analyses.
 \end{abstract}
@@
 \begin{keywords}
-keyword1 -- keyword2 -- keyword3\vspace{-5mm}
+keyword1 -- keyword2 -- keyword3
 \end{keywords}

### 2018-07-16 papers/methods/paper.tex
Adding a bit more clarification from dans emails

@@
 Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
+We can thus model a parametrised selection effect by including deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
 
@@
 \end{figure}
+
+Table \ref{tab:param_summary} provides a quick summary of the top-level parameters defined in our model and their use.
+
+\begin{table}
+	\centering
+	\caption{Top-level parameters defined in our model and a summary of their use.
+The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}
+	\label{tab:param_summary}
+	\begin{tabular}{ll}
+		\hline
+		Parameter & Description \\
+		\hline
+		$\Omega_m$  & Matter density  \\
+		$w$  &  Dark energy equation of state  \\
+		$\alpha$ & Stretch standardisation   \\
+		$\beta$         &  Colour standardisation   \\
+		$\delta(0)$ & Scale of the mass-magnitude correction\\
+		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
+		$\delta\mathcal{Z}_i$ & Systematics scale\\
+		$\delta S$ & Selection effect deviation \\
+		$\langle M_B \rangle$ & Mean absolute magnitude \\
+		$s$ & Colour bias\\
+		\hline
+		$\langle x_1^i \rangle$ & Mean stretch nodes\\
+		$\langle c^i \rangle$ & Mean colour nodes\\
+		$\sigma_{M_B}$ & Population magnitude scatter \\
+		$\sigma_{x_1}$ & Population stretch scatter \\
+		$\sigma_{c}$ & Population colour scatter \\
+		$\kappa_{0}$ & Extra colour dispersion\\
+		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
-}
+We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}
 	\label{tab:bulk_summary}
@@
 Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
+Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.
 

### 2018-07-17 papers/methods/paper.tex
Updating from some of Dans feedback

@@
 \newcommand {\etal} {\emph{~et~al.} }
-\newcommand{\green}{\color{green}}
+\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
+\newcommand{\green}{\color{forestgreen}}
 \newcommand{\blue}{\color{blue}}
@@
 	\centering
-	\caption{Top-level parameters defined in our model and a summary of their use.
-The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}
+	\caption{{\green Top-level parameters defined in our model and a summary of their use.
+The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}}
 	\label{tab:param_summary}
@@
 However, their deviation from the truth value represents a shift of approximately $0.3\sigma$ when taking into account the uncertainty on fits to $w$.
-The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
-We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}
+{\green The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
+We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}}
 	\label{tab:bulk_summary}
@@
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
-Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+{\green Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
 We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
-Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.
+Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.}
 
@@
 For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-For our full systematics model, the bias represents a deviation from $-0.15\sigma$ to $0.27\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
+{\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
 

### 2018-07-27 papers/methods/paper.tex
Adding a bit more clarification from dans emails

@@
 Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
-Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB.
+Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complimentary probes of large scale structureand the CMB.
 For a recent review, see \citet{Huterer2018}.
@@
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
-Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainties will prove to be the limiting factor in our analyses \citep{Betoule2014}.
+Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainties will prove to be the limiting factor in our analyses \citep{Betoule2014, Scolnic2017}.
 
@@
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A common solution to this computational cost problem is to bin the supernova.
+{\green A common solution to this computational cost problem is to bin the supernova, which then produces a matrix of arbitrary size.
+Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.}
 
@@
 
-Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data.
-With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
+Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
+{\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtract this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
+We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
 We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations.
@@
 The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}.
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
-We place a uniform prior on $s_c$.
 It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
@@
  In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics.
-Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016}.
-The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
+{\green Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
+Initial tests were also done with a second, Gaussian population, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.} The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
@@
 {\green The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
-We also note that the simulations do not vary cosmological parameters nor population, and as our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and in fact represents a minimum bound on permissible uncertainty values.}}
+We also note that the simulations do not vary cosmological parameters nor population.
+As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.}}
 	\label{tab:bulk_summary}

### 2018-07-28 papers/methods/paper.tex
More updates from feedback with Dan

@@
 
-At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function.
+At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function.
 The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
@@
 
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has an apparent magnitude $\hat{m}_B$, stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of being a Type Ia, $\hat{p}$.
+We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
 Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
@@
 As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) is taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
 This is discussed further in Section \ref{sec:systreat}.
-Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable.
+Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
 One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
@@
 
-To help confuse things, we often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
+We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
 
@@
 We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement over just colour dispersion and was far more computationally inefficient.
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
 As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence.
@@
 We can thus model a parametrised selection effect by including deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
+{\green Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.}
 

### 2018-07-30 papers/methods/paper.tex
Adding a bit more clarification from dans emails

@@
 
+\subsubsection{Observables}
 Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
@@
 One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
+
+\subsubsection{Latent Variables for Observables}
 
@@
 where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
-Unfortunately, this assumption of Gaussianity is incorrect.
-The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion, wherein some supernovae are uniformly brighter or dimmer than the population mean, and in chromatic dispersion, whereby the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
+
+\subsubsection{Dealing with incorrect observables}
+
+{\green Unfortunately, the assumption of Gaussian uncertainty is incorrect - the errors reported by the light curve fitting process on $m_B$, $x_1$ and $c$ are not accurate.
+The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion and in chromatic dispersion, where respectively some supernovae are uniformly brighter or dimmer than the population mean, and the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
 The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated.
-We address this problem with a similar method to previous analytic likelihoods: including extra dispersion in the reported covariance matrix $\mathcal{C}$.
-Whilst this doesn't solve the issue of non-Gaussianity, it does address the under-reporting of uncertainty, which is the primary source of bias.
+We address this in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.}
 
 We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
-We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
 As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence.
@@
 
-Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
-{\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtract this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
+
+Our second way of addressing the issues with uncertainty are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
+{\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
 Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
@@
 
+\subsubsection{Type Ia population} 
 The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude.
@@
 
-On top of the SNIa populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SNIa population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+\subsubsection{Outlier populations}
+
+On top of the SNIa populations, as described above, we also include a simplistic outlier population model.
+We follows {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
+The mean of the outlier population is fixed to the SNIa population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section.
+
 With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population.
@@
 
-\subsection{Population Map}
+\subsection{Mapping function}
 

### 2018-08-01 papers/methods/paper.tex
Some feedback from Tams email, more to come

@@
 Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery.
-Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into an increased number of systematics than prior analyses.
+Simplified statistical simulations for ideal supernova cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
+SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics than previous analyses.
 \end{abstract}
@@
 
-\subsection{Traditional Cosmology Analyses}
-
-Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \eqref{eq:standard}), and the cosmologically predicted values $\mu_C$.
+\subsection{Traditional Cosmology Analyses} \label{sec:traditional}
+
+Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \ref{eq:standard}), and the cosmologically predicted values $\mu_C$.
 This function is given by
@@
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-{\green A common solution to this computational cost problem is to bin the supernova, which then produces a matrix of arbitrary size.
+{\green A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size.
 Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.}
@@
 
-Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects that plague supernovae surveys.
-By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood.
+Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects that plague supernova surveys.
+By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}.
 This freedom comes with a cost -- computation.
@@
 One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
-\citet{Jennings2016} presents both two metrics (which are used to measure the distance between the forward modelled population and observed population, which is minimised in fitting).
+\citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and is minimised in fitting.
 The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics.
@@
 \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
-\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction.
+\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction.
 \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations.
@@
 Their model was validated to be free of significant biases using fits to thirty realisations of statistics.
+The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
+For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.
 
@@
 None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
-
-The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
-For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.
 
@@
 where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
+
+\begin{table}
+	\centering
+	\caption{{\green Top-level parameters defined in our model and a summary of their use.
+The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}}
+	\label{tab:param_summary}
+	\begin{tabular}{ll}
+		\hline
+		Parameter & Description \\
+		\hline
+		$\Omega_m$  & Matter density  \\
+		$w$  &  Dark energy equation of state  \\
+		$\alpha$ & Stretch standardisation   \\
+		$\beta$         &  Colour standardisation   \\
+		$\delta(0)$ & Scale of the mass-magnitude correction\\
+		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
+		$\delta\mathcal{Z}_i$ & Systematics scale\\
+		$\delta S$ & Selection effect deviation \\
+		$\langle M_B \rangle$ & Mean absolute magnitude \\
+		$s_c$ & Colour bias\\
+		\hline
+		$\langle x_1^i \rangle$ & Mean stretch nodes\\
+		$\langle c^i \rangle$ & Mean colour nodes\\
+		$\sigma_{M_B}$ & Population magnitude scatter \\
+		$\sigma_{x_1}$ & Population stretch scatter \\
+		$\sigma_{c}$ & Population colour scatter \\
+		$\kappa_{0}$ & Extra colour dispersion\\
+		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 {\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
+Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$.
 Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
@@
 Table \ref{tab:param_summary} provides a quick summary of the top-level parameters defined in our model and their use.
-
-\begin{table}
-	\centering
-	\caption{{\green Top-level parameters defined in our model and a summary of their use.
-The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}}
-	\label{tab:param_summary}
-	\begin{tabular}{ll}
-		\hline
-		Parameter & Description \\
-		\hline
-		$\Omega_m$  & Matter density  \\
-		$w$  &  Dark energy equation of state  \\
-		$\alpha$ & Stretch standardisation   \\
-		$\beta$         &  Colour standardisation   \\
-		$\delta(0)$ & Scale of the mass-magnitude correction\\
-		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
-		$\delta\mathcal{Z}_i$ & Systematics scale\\
-		$\delta S$ & Selection effect deviation \\
-		$\langle M_B \rangle$ & Mean absolute magnitude \\
-		$s$ & Colour bias\\
-		\hline
-		$\langle x_1^i \rangle$ & Mean stretch nodes\\
-		$\langle c^i \rangle$ & Mean colour nodes\\
-		$\sigma_{M_B}$ & Population magnitude scatter \\
-		$\sigma_{x_1}$ & Population stretch scatter \\
-		$\sigma_{c}$ & Population colour scatter \\
-		$\kappa_{0}$ & Extra colour dispersion\\
-		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
-		\hline
-	\end{tabular}
-\end{table}
 

### 2018-08-01 papers/methods/paper.tex
More changes from Tam before refactor

@@
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
-The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS) (a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.
+The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.
 
@@
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
 
@@
 As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$.
-The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$.
+The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1, \hat{c} \rbrace = f(y)$.
 For notational simplicity, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
@@
 The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion and in chromatic dispersion, where respectively some supernovae are uniformly brighter or dimmer than the population mean, and the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
-The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated.
-We address this in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.}
+The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics uncertainty reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated.
+We address this in our analysis in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.}
 
@@
 As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to sufficient extra uncertainty in our model.
 The $\kappa$ parameters are tightly constrained as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$, and the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
-The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulted in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion  Each survey has its own set of $\kappa$ parameters.
+The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulting in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion.
+Each survey has its own set of $\kappa$ parameters.
 As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
@@
 
-Our second way of addressing the issues with uncertainty are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
+Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
 {\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
@@
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.
-The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.}
+The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.
+Variables that appear outside both boxes represent top level model parameters.}
 	\label{fig:pgm}

### 2018-08-02 papers/methods/paper.tex
Adding to the table

@@
 In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
+Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.
 
 \subsection{Observed Populations}
-
-\subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
-Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
-
-As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
-This is discussed further in Section \ref{sec:systreat}.
-Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
-
-\subsubsection{Latent Variables for Observables}
-
-The first layer of the hierarchy represents the parameters that describe each supernova.
-That is, observed parameters are denoted with a hat, whilst the true value is denoted without a hat.
-For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.
-
-For the moment, let us consider a single supernova, and represent its true parameters as $\theta$.
-The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
-As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$.
-The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1, \hat{c} \rbrace = f(y)$.
-For notational simplicity, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-The model prediction for our derived data is
-
-\begin{align}
-p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
-\end{align}
-
-We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
-
-\begin{align}
-p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
-\end{align}
-
-It is common in analytic supernova cosmology models to take that
-
-\begin{align}
-p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
-\end{align}
-
-where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
 
@@
 	\caption{{\green Top-level parameters defined in our model and a summary of their use.
-The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, and the bottom section for parameters which are fit for each survey.}}
+The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are fit for each survey, the third section are parameters which are defined for each supernova.
+The bottom section shows the observables (not parameters) that are the input data to the model.}}
 	\label{tab:param_summary}
@@
 		\hline
+		$m_B$ & True flux\\
+		$x_1$ & True stretch \\
+		$c$ & True colour \\
+		\hline
+		$\hat{m}_B$ & Measured flux\\
+		$\hat{x}_1$ & Measured stretch \\
+		$\hat{c}$ & Measured colour \\
+		$C$ & Covariance on flux, stretch and colour \\
+		$\hat{z}$ & Determined redshift \\
+		$\hat{m}$ & Determined mass probability \\
+		\hline
 	\end{tabular}
 \end{table}
+
+\subsubsection{Observables}
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
+We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
+Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+
+As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
+This is discussed further in Section \ref{sec:systreat}.
+Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
+
+\subsubsection{Latent Variables for Observables}
+
+The first layer of the hierarchy represents the parameters that describe each supernova.
+That is, observed parameters are denoted with a hat, whilst the true value is denoted without a hat.
+For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.
+
+For the moment, let us consider a single supernova, and represent its true parameters as $\theta$.
+The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
+As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$.
+The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1, \hat{c} \rbrace = f(y)$.
+For notational simplicity, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
+The model prediction for our derived data is
+
+\begin{align}
+p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
+\end{align}
+
+We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
+
+\begin{align}
+p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
+\end{align}
+
+It is common in analytic supernova cosmology models to take that
+
+\begin{align}
+p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
+\end{align}
+
+where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
 

### 2018-08-10 papers/methods/paper.tex
Small changes before bigger refactor

@@
 We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects.
-This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improved treatment of Malmquist bias, whilst increasing numerical efficiency.
+This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency.
 Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery.
@@
 Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
-Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complimentary probes of large scale structureand the CMB.
+Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complimentary probes of large scale structure and the CMB.
 For a recent review, see \citet{Huterer2018}.
@@
 First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
-\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulation to quantify observational bias in as a function of many factors to better remove fitting biases.
+\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in cosmological parameters as a function of many factors to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
@@
 Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly.
-Instead, most techniques fit and observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
+Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
 This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}.
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
-As all supernova are not identical, an ensemble of supernova form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
+As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
 
 The population previous discussed represents an observed population, which -- due to the presence of various selection effects -- may not represent the underlying supernova population.
-Accurately characterising this population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
-However, given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map between those two populations, from apparent magnitude space to absolute magnitude.
+Accurately characterising this underlying population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
+However, given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map that translates underlying population in absolute magnitude space into the observed population living in apparent magnitude space.
 Specifically, for any given supernova our functional map may take the traditional form:
@@
 The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and a redshift.
-The $\text{other corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant effects on supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017} In traditional analyses, the other corections term often includes bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
+The $\text{other corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017} In traditional analyses, the other corections term often includes bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
 
@@
 The simulated survey is magnitude limited, and all supernovae brighter than magnitude 24 are successfully observed (shown as blue dots), and all dimmer than 24th magnitude are not successfully observed (shown as red dots).
-By binning the supernovae along redshift, and taking the mean distance modulus of the supernovae in each bin, we can see that at higher redshift where Malmquist bias kicks in, the population mean drops and becomes bias.
+By binning the supernovae along redshift, and taking the mean distance modulus of the supernovae in each bin, we can see that at higher redshift where Malmquist bias kicks in, the population mean drops and becomes biased.
 This source of bias must either be corrected by adjusting the data (such as subtracting the found bias) for by incorporating Malmquist bias explicitly in the cosmological model.}
@@
 The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision.
-Their model was validated to be free of significant biases using fits to thirty realisations of statistics.
+Their model was validated to be free of significant biases using fits to thirty realisations of supernova datasets.
 The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
@@
 
+\subsubsection{Observables}
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
+We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
+Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+
+As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
+This is discussed further in Section \ref{sec:systreat}.
+Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
+
 \begin{table}
@@
 	\caption{{\green Top-level parameters defined in our model and a summary of their use.
-The parameters are broken into two sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are fit for each survey, the third section are parameters which are defined for each supernova.
+The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova.
 The bottom section shows the observables (not parameters) that are the input data to the model.}}
@@
 
-\subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
-Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
-
-As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
-This is discussed further in Section \ref{sec:systreat}.
-Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
-
 \subsubsection{Latent Variables for Observables}
@@
 
-{\green Unfortunately, the assumption of Gaussian uncertainty is incorrect - the errors reported by the light curve fitting process on $m_B$, $x_1$ and $c$ are not accurate.
+{\green Unfortunately, the assumption of Gaussian uncertainty is incorrect - the errors reported by the light curve fitting process on $m_B$, $x_1$ and $c$ are not accurate and thus \eqref{eq:pop} does not necessarily hold.
 The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion and in chromatic dispersion, where respectively some supernovae are uniformly brighter or dimmer than the population mean, and the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
@@
 As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to sufficient extra uncertainty in our model.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to allow sufficient extra uncertainty in our model.
 The $\kappa$ parameters are tightly constrained as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$, and the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.

### 2018-08-14 papers/methods/paper.tex
Adding challenges section

@@
 In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance.
-Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology.
+Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
@@
 This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}.
-The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics.
+The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
 As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
@@
 
+\section{Challenges in Supernova Cosmology}
+\label{sec:challenges}
+
+{\red 
+The diverse amount of approaches and implementations applied to supernova cosmology is a direct result of the large number of challenges and complications faced by when performing supernova cosmology analyses.
+In this Section, we outline several of the most prominent challenges.
+
+Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion.
+Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the variability of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
+We utilise two dispersion models in this work.
+The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+Each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics which is not encapsulated in the reported covariance $\mathcal{C}$.
+Adopting either scatter model results in markedly different biases, however our uncertainty on the actual dispersion model means we have to support both scatter models.
+
+The underlying supernova population is further complicated by the presence of outliers.
+Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae.
+This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
+For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernovae spectra, this outlier population is negligible.
+However, for photometric surveys which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.
+
+Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics.
+Because of the rarity of SNIa, datasets are commonly formed from the discoveries of multiple surveys.
+However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
+Peculiar velocities, different host environments, dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
+}
+
 \section{Our Method}
@@
 We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations.
-The scatter models, sourced from \citet{Guy2010} and \citet{Chotard2011} are discussed in detail in Section \ref{sec:simdes}.
 The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}.
@@
 With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population.
-This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification and incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification  \citep{Gupta2016}.
 Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination, both from non type Ia supernovae and from catastrophic failure, is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
@@
 Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach.
-In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models.
-The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
 

### 2018-08-24 papers/methods/paper.tex
Tams edits

@@
 Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery.
-Simplified statistical simulations for ideal supernova cosmology show no bias, and sophisticated SNANA simulations of the combined low-redshift and DES 3-year spectroscopic survey sample quantify the bias for flat $w$CDM cosmology, with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ included to allow rigorous determination of model bias in $w$.
-SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics than previous analyses.
+We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample.
+SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and being more simulation independent than previous analyses.
 \end{abstract}
@@
 
-Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
+Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
 Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
-Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complimentary probes of large scale structure and the CMB.
+Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB.
 For a recent review, see \citet{Huterer2018}.
@@
 In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance.
-Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
+Section \ref{sec:review} is dedicated to a quick review of the supernova cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
 Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.
+We summarise our methodology in Section \ref{sec:conclusion}.
 
@@
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-This model is trained separately before fitting the supernovae light curves for cosmology \citep{Guy2010, Mosher2014}.
+This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Mosher2014}.
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
-As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $m_B$, $x_1$ and $c$, where the width of this distribution is called the intrinsic dispersion of the supernova population.
-
-The population previous discussed represents an observed population, which -- due to the presence of various selection effects -- may not represent the underlying supernova population.
+As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the width of this distribution is called the intrinsic dispersion of the supernova population and the hat denotes an observed variable.
+
+This represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
 Accurately characterising this underlying population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
-However, given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it is encoded in the functional map that translates underlying population in absolute magnitude space into the observed population living in apparent magnitude space.
-Specifically, for any given supernova our functional map may take the traditional form:
+Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population living in apparent magnitude space.
+Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:
 
@@
 where $M_B$ is the mean absolute magnitude for all SN Ia  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
-The cosmological term, $\mu(z)$ represents the distance modulus, and is precisely known given cosmological parameters and a redshift.
-The $\text{other corrections}$ term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017} In traditional analyses, the other corections term often includes bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
+The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by rearranging equation \eqref{eq:standard} to isolate $\mu(z)$.
+The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
+In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
 
@@
 
-Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus $\mu_{\rm obs}$ (attained from rearranging equation \ref{eq:standard}), and the cosmologically predicted values $\mu_C$.
+Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from rearranging equation \ref{eq:standard}, and the cosmologically predicted values.
+To distinguish them, we denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$.
 This function is given by
@@
 \end{align}
-where $r$ is the comoving distance for redshift $z$ given a specific cosmology and $H_0$ is the current value of Hubble's constant in $\kmsmpc$.
+where $r$ is the comoving distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
 
@@
 Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
-For example, the underlying supernova colour population's mean and skewness is highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
+For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
 Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
@@
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
-As these effects are not built into the likelihood and instead are formed by correcting data, their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
+When using traditional fitting methods, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
 
@@
 One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
-\citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and is minimised in fitting.
+\citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and are minimised in fitting.
 The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics.
@@
 \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations.
-The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
+The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
 \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}.
@@
 
-Whilst there are a large amount of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
+Whilst there are a large number of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
 None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
 
-The flexibility afforded by a hierarchical model allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, supernovae rates and redshift distributions, and host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
+The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
+Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models and increasing simulation independence to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
 
@@
 
-{\red 
-The diverse amount of approaches and implementations applied to supernova cosmology is a direct result of the large number of challenges and complications faced by when performing supernova cosmology analyses.
+{\green 
+The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by when performing supernova cosmology analyses.
 In this Section, we outline several of the most prominent challenges.
@@
 Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion.
-Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the variability of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
+Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
 We utilise two dispersion models in this work.
-The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation.
 The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-Each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics which is not encapsulated in the reported covariance $\mathcal{C}$.
-Adopting either scatter model results in markedly different biases, however our uncertainty on the actual dispersion model means we have to support both scatter models.
+As the SALT2 model does not include intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population.
+We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.
 
@@
 Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae.
-This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
-For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernovae spectra, this outlier population is negligible.
-However, for photometric surveys which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.
+This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
+Depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
+For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
+However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.
 
@@
 However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
-Peculiar velocities, different host environments, dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
+Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
 }
@@
 We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
-As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
+As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
 The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.
@@
 \subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty $\cov$ on those values.
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
@@
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.
 
@@
 	\centering
-	\caption{{\green Top-level parameters defined in our model and a summary of their use.
+	\caption{{\green Parameters defined in our model and a summary of their use.
 The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova.
@@
 		$\langle c^i \rangle$ & Mean colour nodes\\
+		$\alpha_c$ & Skewness of colour population \\
 		$\sigma_{M_B}$ & Population magnitude scatter \\
@@
 		$c$ & True colour \\
+		$z$ & True redshift \\
+		$M_B$ & Derived absolute magnitude \\
+		$\mu$ & Derived distance modulus \\
 		\hline
@@
 		$C$ & Covariance on flux, stretch and colour \\
-		$\hat{z}$ & Determined redshift \\
+		$\hat{z}$ & Observed redshift \\
 		$\hat{m}$ & Determined mass probability \\
+		$\hat{p}$ & Determined outlier probability \\
 		\hline
@@
 The first layer of the hierarchy represents the parameters that describe each supernova.
-That is, observed parameters are denoted with a hat, whilst the true value is denoted without a hat.
+That is, observed parameters are denoted with a hat, whilst the true (latent) value is denoted without a hat.
 For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.
 
-For the moment, let us consider a single supernova, and represent its true parameters as $\theta$.
+For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$.
+For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
+A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves and uncertainty.
+However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$.
+When using this simplification, our latent variables are given by
+\begin{align}
+p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
+\end{align}
+As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including intrinsic dispersion in the supernova model.
+
+%$\theta$.
 The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
@@
 The model prediction for our derived data is
-
-\begin{align}
-p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
-\end{align}
-
-We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
-
-\begin{align}
-p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
-\end{align}
-
-It is common in analytic supernova cosmology models to take that
-
-\begin{align}
-p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
-\end{align}
-
-where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
-
-\subsubsection{Dealing with incorrect observables}
-
-{\green Unfortunately, the assumption of Gaussian uncertainty is incorrect - the errors reported by the light curve fitting process on $m_B$, $x_1$ and $c$ are not accurate and thus \eqref{eq:pop} does not necessarily hold.
-The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion and in chromatic dispersion, where respectively some supernovae are uniformly brighter or dimmer than the population mean, and the spectral energy density of the supernovae varies non-uniformly as a function of wavelength.
-The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics uncertainty reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated.
-We address this in our analysis in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.}
+%
+%\begin{align}
+%p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
+%\end{align}
+%
+%We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
+%
+%\begin{align}
+%p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
+%\end{align}
+%
+%It is common in analytic supernova cosmology models to take that
+%
+%\begin{align}
+%p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
+%\end{align}
+%
+%where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
+
+\subsubsection{Correcting biased summary statistics}
+
+With the report summary statistics being biased and their uncertainty under-reported, we face a significant challenge in supernova cosmology.
+We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
+We must also do this without assuming a specific dispersion model (either the {\gten} or {\celeven} model).
+We address this in our analysis in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.
 
 We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence.
+We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by $\sigma_{M_B}$ without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
+
+As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift.
 As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to allow sufficient extra uncertainty in our model.
-The $\kappa$ parameters are tightly constrained as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$, and the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature).
+The $\kappa$ parameters are tightly constrained with priors as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
+We subject them to strong Cauchy priors centered on zero and with width $0.002$, where the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
 The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulting in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion.
@@
 
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
 We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
@@
 Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
-{\green This redshift-dependent array is added to our data-set and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
-Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$.
+By comparing simulated true colour values to simulated observed colour values, we can quantify the mean bias in colour in redshift bins, giving us {\green a redshift-dependent array of biases.
+We can then add this into our data-set, and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
+Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$, where `bias$(z)$' is the binned mean colour bias determined with fiducial simulations.
 Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
 We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
-We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations.
-The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}.
+We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction: 0 for no shift, up to 1 for the maximum shift found in simulations.
+The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}, which has greater chromatic dispersion than the {\gten} model.
+
 It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
+Our fiducial cosmology is a Flat $\Lambda$CDM universe with $\Omega_m = 0.3$ and $H_0 = 70\kmsmpc$.
+
+{\red MAKE SOME ILLUSTRATIVE PLOT FOR THIS SECTION}
 
@@
 	\node[det, above=of zobs] (z) {$z$};
-	\node[det, above=of mobs] (m) {$m$};
+
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[latent, above=of m] (MBobs) {$M_B$};
+	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
 	% Survey nodes
-	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1, s_c$ };
+	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
 	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
@@
 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta,$\\$\delta(0), \delta(\infty)$};
+	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $s_c$, $\delta \mathcal{Z}_i$, $\delta S$};
 
@@
 	\edge {z} {zobs};
-	\edge {m} {mobs};
+
 	\edge {z,cosmo} {mu};
-	\edge {mu, sum, m, pop1, ab, pobs, sf} {MBobs};
+	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
 
 	% Plates
-	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(m)(mu)} {$\forall\ \rm{SN}$} ;
+	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
 	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
@@
 	\caption{Probabilistic graphical model for our likelihood.
-Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, and standard nodes represent fit variables.
+Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
 The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.
@@
 \subsubsection{Type Ia population}
+
 The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude.
-Analytic models often treat the colour and stretch populations independently with a skew normal and normal, respectively.
-Intrinsic dispersion is treated in a variety of manners in other models, from representing it simply as (Gaussian) scatter on the absolute magnitude of the supernova population, to correlated multivariate normal scatter on the combined magnitude, stretch and colour distribution.
-Additionally, redshift drift of populations' mean colour and stretch will introduce a cosmological bias in our fits unless the model population possesses similar ability to change as a function of redshift.
-
-We follow prior work and model the colour population as an independent redshift-dependent skew normal for each survey.
+We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey.
 For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal.
-Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes.
+Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift.
 Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey.
@@
 Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
-We parameterise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
+We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
 The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.
@@
 
-On top of the SNIa populations, as described above, we also include a simplistic outlier population model.
+For future use with photometrically classified surveys, we include a simplistic outlier population model.
 We follows {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
-The mean of the outlier population is fixed to the SNIa population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works.
+However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.
+We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
+We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
 The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section.
-
-With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population.
-Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination, both from non type Ia supernovae and from catastrophic failure, is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$.
-We assume that supernovae fall into the SNIa population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.
+For the spectroscopic survey, we set this to unity.
+For the photometric proof of concepts we provide an accurate probability estimate.
+Further investigation on the effect of inaccurate estimates we leave for future improvements during the analysis of the DES photometric sample.
 
@@
 {\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0).
-\citet{Shariff2016} instead model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and $\beta(z) = \beta_0 + \Delta \beta\left(0.5 + \arctan(100(z - z_t)) / \pi\right)$, where the later effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
-
-We tested two models against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$.
+\citet{Shariff2016} instead model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and a second model which effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
+
+We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$.
 See Section \ref{sec:simdes} for details on simulation generation.
 We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
-This degeneracy resulted in a significant bias in recovered cosmology, and so in our final model we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+This degeneracy resulted in a significant bias in recovered cosmology.
+As $\alpha$ and $\beta$ are constants in the SALT2 model, assuming varying $\alpha$ or $\beta$ is only self consistent if the same variations are used in the light curve fitting process.
+As this variation is not present in the DES spectroscopic sample's light curve fitting process, and because of the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
 As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.
@@
 where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift.
-We take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$.
+We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.
 With this correction, our calculation of absolute magnitude becomes
@@
 \begin{align}
-M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times {\rm mass\_prob}. \label{eq:l2}
+M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times \hat{m}. \label{eq:l2}
 \end{align}
@@
 \label{sec:selection}
-Our treatment of selection effects is to incorporate selection efficiency into our model, rather than relying on simulation-driven data corrections.
+One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
 We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
@@
 
-As our model so far described in previous sections describe components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
+As our model so far describes components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
 
@@
 \begin{align}
-d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
-\end{align}
-
+d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
+\end{align}
+
+where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$ and $c$ population to $m_B$ given cosmological parameters.
 A full derivation of this can be found in Appendix \ref{app:selection}.
-We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this integral.
-It was found that discarding the skewness entirely resulted in highly biased population recovery, and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
-This shifted Gaussian approximation removes the unintended bias when simply discarding skewness.
-More detail on this shift can be found in Appendix \ref{app:approx}.
+
+We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this evaluating $P(m_B | z, \theta)$.
+It was found that discarding the colour population skewness entirely resulted in highly biased population recovery (see Figure \ref{fig:simple_w_super} to see the populations), and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
+This shifted Gaussian approximation for colour completely removes the unintended bias when simply discarding skewness.
+This shift was not required for the stretch population, and so was left out for numerical reasons.
+The impact of this approximation on the calculated efficiency is shown in Figure \ref{fig:shift}, and more detail on this shift and resulting population recovery can be found in Appendix \ref{app:approx}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{shift.pdf}
+	\end{center}
+	\caption{Testing the correctness of our normal approximation to the skewed colour distribution.
+The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$, calculated numerically.
+The $x$-axis is analogous to $m_B$ is cosmological context.
+As expected, all efficiencies drop towards zero as shift increases (as objects get fainter).
+The unshifted normal approximation shows significant discrepancy in the calculated efficiency as it transitions from 1 to 0, whilst the shifted normal approximation shows negligible error to the correct solution.
+From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) as unnecessary.}
+	\label{fig:shift}
+\end{figure}
+
 The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
@@
 m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
-\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
+\sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
 \end{align}
@@
 
-where $\Phi^c$ the complimentary cumulative distribution function (CDF) and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
+where $\Phi^c$ the complimentary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
 The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
-However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form
+However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
 
@@
 The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude.
-Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit.
-Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
-We can thus model a parametrised selection effect by including deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
-{\green Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.}
+That is, we calculate the probability we would include a particular supernova in our sample, divided by the number of such supernovae in our simulated fields.
+To take into account the uncertainty introduced by the imperfection of our analytic fit to the efficiency ratio, uncertainty was uniformly added in quadrature to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
+This is mathematically identical to fitting the efficiency ratio with a second `intrinsic dispersion' parameter which adds uncertainty to the efficiency ratio data points.
+
+We can thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty.
+Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
+{\green Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.
+Tests with changing the selection efficiency model show that the uncertainty introduced is sub-dominant to the uncertainty introduced from assuming an analytic form.}
 
@@
 
-and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency} for aid in visualisation.
+and is incorporated into our likelihood.
+This is illustrated in Figure \ref{fig:efficiency}.
+Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency.
+Further details on this choice are given in Section \ref{sec:simdes}.
 
@@
 \end{figure}
-
-Table \ref{tab:param_summary} provides a quick summary of the top-level parameters defined in our model and their use.
 
@@
 \section{Conclusions}
+\label{sec:conclusion}
 
@@
 The biases when using shifted or unshifted normal approximations when we fit our model on Gaussian and skewed underlying populations are shown in Figure \ref{fig:simple_w_super}, and only the shifted normal approximation correctly recovers underlying population parameters.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{shift.pdf}
-	\end{center}
-	\caption{Testing the correctness of our normal approximation to the skewed colour distribution.
-The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$, calculated numerically.
-The $x$-axis is analogous to $m_B$ is cosmological context.
-As expected, all efficiencies drop towards zero as shift increases (as objects get fainter).
-The unshifted normal approximation shows significant discrepancy in the calculated efficiency as it transitions from 1 to 0, whilst the shifted normal approximation shows negligible error to the correct solution.
-From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) as unnecessary.}
-	\label{fig:shift}
-\end{figure}
 

### 2018-08-30 papers/methods/paper.tex
Updating more of the paper

@@
 \label{sec:selection}
+
 One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
@@
 \begin{align}
-d_{\rm CDF/SNIa} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
+d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
 d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
@@
 Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution.
-The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), but as generic functions of apparent magnitude.
+The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude.
 The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
@@
 
-In order to verify our model we run it through several tests.
-First, we validate on toy models, verifying that there is not significant cosmological bias in highly constraining datasets.
-We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three year spectroscopic sample.
+In order to verify our model we run it through stringent tests.
+First, we validate on toy models, verifying that we recover accurate cosmology in highly constraining datasets.
+We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample.
 
@@
 Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.
-The colour of the data points is representative over the supernovae colour itself, blue dots represent blue supernovae, with warmer colours representing redder supernovae.}
+The colour of the data points is representative over the supernovae colour itself, a negative colour value showing bluer supernovae, with positive colour values representing redder supernovae.}
 	\label{fig:simple_pop}
@@
 	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context.
-Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
+Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ or $\Omega_\Lambda$ separately due to the severe degeneracy between the parameters.}
 	\label{fig:simple_ol}
@@
 	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface.
-The recovered posteior maximums show the same degeneracy direction as the representative surface and scatter around the truth values input into the simulation, shown in dashed lines.}
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the non-Gaussian nature of the posterior surface.
+The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
 	\label{fig:simple_w}
@@
 As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted standard deviation: $\sqrt{100} = 10$.
-As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model (which was run with a prior on $\Omega_m$).}
+As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
+For the Flat $w$CDM model, the value of $w$ is reported with a prior on $\Omega_m$ of $\mathcal{N}(0.3, 0.01)$.}
 	\label{tab:simple_model}
@@
 
-We generate simple toy data to validate the basic premise of the model. Data generation algorithm is described below:
+We generate simple toy data to validate the basic premise of the model. The data generation algorithm is described below:
 
@@
 	\item Draw a true absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
-	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) = 0.5$ and equation \eqref{eq:mass}.
+	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
 	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$.
@@
 For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$.
-Repeat from step one until we have a supernova which passes.
+Repeat from step one until we have a supernova that passes.
 	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
@@
 
-For supernova independent uncertainty, the selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters.
-We draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES spectroscopic analysis.
+For supernova-independent uncertainty, the selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
+Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES three-year spectroscopic analysis.
 Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
@@
 By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them.
-The results of the analysis are detailed in Table \ref{tab:simple_model}, and do not reveal evidence of systematic bias in our model.
+The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
 
@@
 {\green Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
-Initial tests were also done with a second, Gaussian population, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.} The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
+Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.} The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
@@
 Points are shown for parameters $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$, with the other fit parameters being marginalised over.
-The intrinsic scatter model has significant impact on the recovered $\beta$, which in itself effects cosmology, resulting in small biases in $w$.}
+As we are unable to fully correct observed summary statistics, a step required by the lack of intrinsic scatter in the SALT2 model, we expect to see an offset in $\alpha$ and $\beta$.
+This in turn effects cosmology, resulting in small biases in $w$.}
 	\label{fig:bulk_posterior}
@@
 		\hline
-		Model & $w$ (scatter) & Bias Significance \\
-		\hline
-		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $3.9\sigma$  \\
-		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $4.4\sigma$  \\
-		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $2.0\sigma$   \\
-		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $7.9\sigma$   \\
+		Model & $w$ (scatter) & Bias \\
+		\hline
+		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $0.3\sigma$  \\
+		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $0.3\sigma$  \\
+		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $0.1\sigma$   \\
+		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $0.6\sigma$   \\
 		\hline
@@
 
-Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-
-Cosmological bias is detected with a confidence of more than $3\sigma$ for the {\gten} and {\celeven} models.
-We investigate the bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
-This is confirmed by replacing the fully simulated observed $m_B$, $x_1$ and $c$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
+The parameter bounds are listed in Table \ref{tab:bulk_summary}.
+Our recovered results are all within $0.5\sigma$ of the true values, however some bias does emerge.
+
+We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+This is confirmed by replacing the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.

### 2018-08-31 papers/methods/paper.tex
More changes

@@
 For supernova-independent uncertainty, the selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
-Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 250 supernovae for the DES three-year spectroscopic analysis.
+Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis.
 Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
@@
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty is incorrect, non-Gaussian and biased.
-{\green Our current model and techniques can quantify the effect of different scatter models, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
+{\green Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
 We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
@@
 
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data, and is difficult to do whilst accounting for correlations with population and scatter model.
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
+It is difficult to do whilst accounting for correlations with population and scatter model.
 This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strength of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
 Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
@@
 
-Table \ref{tab:parameter_correlations} lists the fit correlations to our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
+Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
 Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion.
+As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
 Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology as $\beta$.
@@
 
-For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models are sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
 {\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
@@
 When validating using SNANA simulations, we find evidence of cosmological bias which is traced back to light curve fits reporting biased observables and incorrect covariance.
-Allowing fully parametrised corrections on observed supernovae summary statistics is found to make cosmology fits too weak, and allowing simulation based corrections to vary in strength is found to give minor reductions in $w$ bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
+Allowing fully parametrised corrections on observed supernovae summary statistics introduces too many degrees of freedom and is found to make cosmology fits too weak.
+Allowing simulation based corrections to vary in strength is found to give minor reductions in $w$ bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
 For the data size represented in the DES three-year spectroscopic survey, the determined biases should be sub-dominant to other sources of uncertainty, however this cannot be expected for future analyses with larger datasets.
 Stricter bias corrections calculated from simulations are required to reduce bias.
-Ideally, this would include further work on the physical intrinsic dispersion of the Type Ia supernovae population such that we can better characterise this bias.
+Ideally, this would include further work on the calculation of intrinsic dispersion of the Type Ia supernovae population such that we can better characterise this bias.
 
@@
 Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$.
-When dealing with experiments which have imperfect selection efficiency, our likelihood must take that efficiency into account.
+When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.
 We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
@@
 \end{align}
-where we define the denominator as $w$ for simplicity in future derivations.
+where we define the denominator as $d$ for simplicity in future derivations.
 
@@
 
-Equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our colour distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
-We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to match the actual mean and standard deviation of skew normal.
+In this section, we investigate the effect of approximating the skew normal underlying colour distribution as a normal.
+Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our colour distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
+We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to more accurately describe the actual mean and standard deviation of a skew normal.
 With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
@@
 \mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
-\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}}.
-\end{align}
+\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
+\end{align}
+where we highlight that $\mu$ here represents the mean of the distribution, not distance modulus.
 We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$.
-Unfortunately, this shift to the mean and standard deviation of the normal approximation did not produce stable posterior surfaces which Stan could fit when we included full covariance between $m_B$, $x_1$ and $c$ in the underlying population, and so we simplified the model to have independent parameters.
-Stable surfaces with underlying population covariance were found when we fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
-However, this resulted in several population parameters becoming biased, and so we do not fix $\sigma_c$.
+Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$ and $c$ as a multivariate skew normal did not produce stable posterior surfaces.
+Due to this, we treat the underlying $m_B$, $x_1$ and $c$ populations as independent.
+
+We tested a fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
+This resulted in stable posterior surfaces, however this introduced recovery bias in several population parameters, and so we do not fix $\sigma_c$.
 Comparing whether we shift our normal in the approximation or simply discard skewness, Figure \ref{fig:shift} shows that the calculated efficiency is significantly discrepant to the actual efficiency if the normal approximation is not shifted.
@@
 The unshifted model recovers the correct colour mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the colour population mean.
-Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.}
+Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.
+Based on these results and the fantastic performance in correctly recovering underlying populations of the shifted normal approximation, we adopt that in our model.}
 	\label{fig:simple_w_super}
@@
 Instead of computing those powers in Stan, we simply pass in several arrays of redshift values already raised to the correct power.
-Small changes like this however only give small results.
+Small changes like this however only give small improvements.
 
@@
 If we had kept the full covariance matrix parametrisation we could modify the matrix easily - for example when incorporating intrinsic dispersion we could simple add on a secondary matrix to create an updated covariance.
-Using cholesky decompositions, ${\rm cholesky}(\Sigma_1 + \Sigma_2) \neq L_1 + L_2$, and so we would need to recompute the decomposition for each step, which discards most of the computational benefit just gained.
+However as the cholesky decomposition of a sum of matrices is not equal to the sum of the cholesky decomposition of each individual matrix, we would need to recompute the decomposition for each step, which discards most of the computational benefit just gained.
 

### 2018-08-31 dessn/configurations/nominal_uncert.py
Better uncert analysis

@@ -65,6 +65,7 @@ if __name__ == "__main__":
                   + ("lockdisp " if m.lock_disp else "") \
                   + ("lockpop " if m.lock_pop else "") \
                   + ("lockbase " if m.lock_base else "") \
+                  + ("lockdrift " if m.lock_drift else "") \
                   + ("noeff " if not m.apply_efficiency else "")
 
             if key not in ws:
@@ -76,7 +77,7 @@ if __name__ == "__main__":
             # print(key, vals[:, 0])
         for key in sorted(ws.keys()):
             vals = np.array(ws[key])
-            print("%55s %8.3f %8.3f %8.3f"
+            print("%65s %8.4f %8.4f %8.4f"
                   % (key, np.average(vals[:, 0], weights=1/(vals[:, 1]**2)),
                      np.std(vals[:, 0]), np.mean(vals[:, 1])))
             # from chainconsumer import ChainConsumer

### 2018-09-03 papers/methods/paper.tex
Paper updates

@@
 
-Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
-
-Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-Our recovered results are all within $0.5\sigma$ of the true values, however some bias does emerge.
-
-We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
-This is confirmed by replacing the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
-With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
-{\green Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
-We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
-Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.}
-
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
-It is difficult to do whilst accounting for correlations with population and scatter model.
-This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=\textwidth]{bulk_cor.pdf}
-	\end{center}
-	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
-We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
-Also visible in this figure are several other interesting relationships.
-$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
-This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
-As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
-And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
-The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
-These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
-As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
-Strong anti-correlation between the $\kappa_{c0}^0$ and $\kappa_{c0}^1$ (and positive fit values) shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
-Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
-Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
-}
-	\label{fig:bulk_cor}
-\end{figure*}
-
-Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
-Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
-Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology as $\beta$.
-Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
-The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
-
-For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-{\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
-
 %		$\delta [ \Delta t ]$ &   -0.01 &      0.00     \\
@@
 
+Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
+
+Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
+The parameter bounds are listed in Table \ref{tab:bulk_summary}.
+Our recovered results are all within $0.5\sigma$ of the true values, however some bias does emerge.
+
+We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+This is confirmed by replacing the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
+{\green Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
+Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.}
+
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
+It is difficult to do whilst accounting for correlations with population and scatter model.
+This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
+Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=\textwidth]{bulk_cor.pdf}
+	\end{center}
+	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
+We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
+Also visible in this figure are several other interesting relationships.
+$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
+This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
+As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
+And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
+The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
+These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
+As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
+Strong anti-correlation between the $\kappa_{c0}^0$ and $\kappa_{c0}^1$ (and positive fit values) shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
+Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
+Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
+}
+	\label{fig:bulk_cor}
+\end{figure*}
+
+Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
+Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
+As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
+Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology as $\beta$.
+Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
+The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
+
+For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+{\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
+
+\subsection{Uncertainty Analysis}
+\label{sec:ucnert}
+
+With the increased flexibility of Bayesian hierarchical models over traditional modes, we expect to find an increased uncertainty on parameter inference.
+To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to vary or locked to a fixed value.
+By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
+
+\begin{table}
+	\centering
+	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features.}
+	\label{tab:uncert}
+	\begin{tabular}{l|rr}
+		Feature & $\sigma_w$ & Cumulative  \\
+		\hline
+		Cosmology only                 & 0.051  &   0.051       \\
+		Standardisation parameters     & 0.046  &   0.068       \\
+		Intrinsic scatter              & 0.020  &   0.071       \\
+		Static survey populations      & 0.022  &   0.074       \\
+		Redshift-dependent populations & 0.030  &   0.080       \\
+		Systematics                    & 0.054  &   0.096       \\
+		\hline
+	\end{tabular}
+\end{table}
+
+The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is still dominated by statistical error, as the total statistical uncertainty is on $w$ is $0.08$.
+With the low number of supernovae in the DES three-year spectroscopic sample, this is expected.
+We note that the label `Systematics' in Table \ref{tab:uncert} represents all numerically computed systematics (as discussed in Section \ref{sec:systreat}) and systematic uncertainty on the selection function.
+
+\subsection{Methodology Comparison}
+\label{sec:comp}
+
+{\red Introduce BBC. Highlight its use on Panstars}
+
+{\red have plots for w-w comparison for g10 and c11}
+
+{\red conclude that correlation is good, yay, have extra uncertainty, but this can be explained from previous section}
+
 \section{Conclusions}

### 2018-09-03 dessn/configurations/bulk.py
Uncert analysis updates, bulk sim updates

@@ -118,6 +118,38 @@ if __name__ == "__main__":
             c3.plotter.plot(filename=[pfn + "_points_c11.png", pfn + "_points_c11.pdf"], parameters=4,
                             truth=truth, extents=ex, figsize=1.0)
 
+        if True:
+            import matplotlib.pyplot as plt
+            fig, axes = plt.subplots(nrows=2, figsize=(5, 9))
+            for t, ax in zip(["G10", "C11"], axes):
+                for key in ws.keys():
+                    if t not in key:
+                        continue
+                    means_steve = ws[key]
+                    std_steve = ws_std[key]
+
+                    means_bbc = means_steve + np.random.normal(loc=0, scale=0.04, size=len(means_steve))
+                    std_bbc = std_steve + np.random.normal(loc=0, scale=0.01, size=len(means_steve))
+
+                    minv = min(np.min(means_steve), np.min(means_bbc)) - 0.02
+                    maxv = max(np.max(means_steve), np.max(means_bbc)) + 0.02
+                    color = "k" if "Syst" not in key else ("g" if "G10" in key else "r")
+                    ecolor = "#888888" if "Syst" not in key else ("#abf4bd" if "G10" in key else "#f4abab")
+                    ax.errorbar(means_steve, means_bbc, yerr=std_steve, xerr=std_bbc, errorevery=2,
+                                fmt="o", capsize=2, markersize=5, c=color, ecolor=ecolor, elinewidth=0.5)
+
+                ax.plot([minv, maxv], [minv, maxv], c='k', linewidth=1)
+                ax.set_xlim([minv, maxv])
+                ax.set_ylim([minv, maxv])
+                ax.set_xlabel("BBC $w$")
+                ax.set_ylabel("Steve $w$")
+            # plt.subplots_adjust(wspace=0, hspace=0)
+
+            fig.tight_layout()
+            plt.show()
+
+
+
 
         # c2.plotter.plot(filename=pfn + "_big.png", parameters=14, truth=truth)
         # c2.plotter.plot_distributions(filename=pfn + "_dist.png", truth=truth, col_wrap=7)

### 2018-09-03 dessn/configurations/nominal_uncert.py

@@ -59,6 +59,7 @@ if __name__ == "__main__":
         res = fitter.load(split_cosmo=True, split_sims=True)
 
         ws = {}
+        ns = {}
         for m, s, ci, chain, truth, weight, old_weight, posterior in res:
             key = ("Statonly " if m.statonly else "Syst ") \
                   + ("locksyst " if m.lock_systematics else "") \
@@ -67,19 +68,42 @@ if __name__ == "__main__":
                   + ("lockbase " if m.lock_base else "") \
                   + ("lockdrift " if m.lock_drift else "") \
                   + ("noeff " if not m.apply_efficiency else "")
-
-            if key not in ws:
-                ws[key] = []
-            ws[key].append([chain["$w$"].mean(), np.std(chain["$w$"])])
+            key2 = (1 if m.statonly else 0) +  \
+                  + (1 if m.lock_disp else 0) \
+                  + (1 if m.lock_pop else 0) \
+                  + (1 if m.lock_base else 0) \
+                  + (1 if m.lock_drift else 0) \
+                  + (1 if not m.apply_efficiency else 0)
+            if key == "Statonly locksyst ":
+                continue
+            key2 = 6 - key2
+            if key2 not in ws:
+                ws[key2] = []
+            if key2 not in ns:
+                ns[key2] = key
+            ws[key2].append([chain["$w$"].mean(), np.std(chain["$w$"])])
 
         # for key in ws.keys():
         #     vals = np.array(ws[key])
             # print(key, vals[:, 0])
         for key in sorted(ws.keys()):
             vals = np.array(ws[key])
-            print("%65s %8.4f %8.4f %8.4f"
-                  % (key, np.average(vals[:, 0], weights=1/(vals[:, 1]**2)),
-                     np.std(vals[:, 0]), np.mean(vals[:, 1])))
+            key_text = ns[key]
+            mean = np.average(vals[:, 0], weights=1/(vals[:, 1]**2))
+            scatter = np.std(vals[:, 0])
+            std = np.mean(vals[:, 1])
+
+            if key != 0:
+                prev_key = key - 1
+                prev_std = np.mean(np.array(ws[prev_key])[:, 1])
+                prev_var = prev_std**2
+                diff = np.sqrt(std**2 - prev_std**2)
+            else:
+                diff = std
+            print("%65s %8.3f %8.3f" % (key_text, std, diff))
+
+
+
[…]

### 2018-09-18 papers/methods/paper.tex
Updating paper

@@
 		\hline
-		{\gten} Stat + Syst  &  $-1.03\pm 0.10\ (0.08)$ & $0.3\sigma$  \\
-		{\celeven} Stat + Syst  &  $-0.97\pm 0.09\ (0.06)$ & $0.3\sigma$  \\
-		{\gten}  Stat         &  $-1.01\pm 0.08\ (0.08)$ & $0.1\sigma$   \\
-		{\celeven} Stat         &  $-0.96\pm 0.07\ (0.05)$ & $0.6\sigma$   \\
+		{\gten} Stat + Syst  &  $-1.00\pm 0.10\ (0.08)$ & $0.0\sigma$  \\
+		{\celeven} Stat + Syst  &  $-0.95\pm 0.10\ (0.06)$ & $0.5\sigma$  \\
+		{\gten}  Stat         &  $-1.00\pm 0.08\ (0.08)$ & $0.0\sigma$   \\
+		{\celeven} Stat         &  $-0.95\pm 0.08\ (0.05)$ & $0.6\sigma$   \\
 		\hline
@@
 
+BHM Mean for G10: -0.998, BBC mean -1.042, average diff -0.044, difference scatter 0.057
+
+BHM Mean for C11: -0.945, BBC mean -0.978, average diff -0.033, difference scatter 0.062
+
+{\red For systematics paper, also find 60\% reduction in uncert using planck priors}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=\columnwidth]{bulk_comparison.pdf}
+	\end{center}
+	\caption{Whoa.}
+	\label{fig:bulk_comparison}
+\end{figure}
+
 \section{Conclusions}

### 2018-09-18 dessn/configurations/bulk.py

@@ -165,18 +165,23 @@ if __name__ == "__main__":
                         res[t + "_BBC"] = means_bbc
                         minv = min(np.min(means_steve), np.min(means_bbc)) - 0.02
                         maxv = max(np.max(means_steve), np.max(means_bbc)) + 0.02
-                        color = "k" if "Syst" not in key else ("#4FC3F7" if "G10" in key else "#673AB7")
-                        ecolor = "#888888" if "Syst" not in key else ("#8edcff" if "G10" in key else "#a785e2")
+                        # color = "k" if "Syst" not in key else ("#4FC3F7" if "G10" in key else "#673AB7")
+                        color = "k" if "Syst" not in key else ("#43A047" if "G10" in key else "#E53935")
+                        # ecolor = "#888888" if "Syst" not in key else ("#8edcff" if "G10" in key else "#a785e2")
+                        ecolor = "#888888" if "Syst" not in key else ("#a0e8a3" if "G10" in key else "#f9a19f")
                         ax.errorbar(means_steve, means_bbc, yerr=std_steve, xerr=std_bbc, errorevery=2,
                                     fmt="o", capsize=2, markersize=5, c=color, ecolor=ecolor, elinewidth=0.5, label=t)
-                    minv = -1.25
-                    maxv = -0.7
-                    ax.plot([minv, maxv], [minv, maxv], c='k', linewidth=1)
-                    ax.set_xlim([minv, maxv])
-                    ax.set_ylim([minv, maxv])
-                    ax.set_ylabel("BBC $w$")
-                    ax.set_xlabel("BHM $w$")
-                    ax.legend(frameon=False)
+
+                minv = -1.25
+                maxv = -0.7
+                ax.plot([minv, maxv], [minv, maxv], c='k', linewidth=1)
+                ax.set_xlim([minv, maxv])
+                ax.set_ylim([minv, maxv])
+                ax.set_ylabel("BBC $w$")
+                ax.set_xlabel("BHM $w$")
+                ax.legend(frameon=False)
+
+
                 # plt.subplots_adjust(wspace=0, hspace=0)
                 print(res.keys())
                 mid_bbc = 0.5 * (res["G10_BBC"] + res["C11_BBC"])
@@ -185,6 +190,7 @@ if __name__ == "__main__":
 
                 fig.tight_layout()
[…]

### 2018-09-18 dessn/configurations/real_data.py

@@ -61,12 +61,6 @@ if __name__ == "__main__":
             c2.add_chain(chain, weights=weight, posterior=posterior, name=name)
 
         import numpy as np
-        # data = np.vstack(chain[x] for x in ['$\\Omega_m$', '$w$', '$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$', '$\\delta(0)$'])
-        # data = data.T
-        # np.save("samchain.npy", data.astype(np.float32))
-        # np.savetxt("samchain.txt", data)
-        # print(data.shape)
-        # exit()
 
         chain_planck, params, weights, likelihood = get_planck()
         if blind:
@@ -88,7 +82,7 @@ if __name__ == "__main__":
         watermark = "Blinded" if blind else None
 
         c.configure(spacing=1.0, diagonal_tick_labels=False, sigma2d=False, plot_hists=False,
-                    sigmas=[0, 1, 2], linestyles=["-", "--", ':', '-', '--', '-'], kde=2.0,
+                    sigmas=[0, 1, 2], linestyles=["-", "--", ':', '-', '--', '-'], kde=1.0,
                     legend_kwargs={"loc": "center right"}, watermark_text_kwargs={"alpha": 0.2},
                     colors=["b", "k", 'a', 'g', 'r', 'lb'], shade_alpha=[0.5, 0.0, 0.2, 0.4, 0.8, 0.1, 0.1])
         c.plotter.plot(filename=[pfn + ".png", pfn + ".pdf"], parameters=parameters, watermark=watermark,

### 2018-09-19 papers/methods/paper.tex
Scaffolding out bbc comparison

@@
 		$\langle M_B \rangle$ & Mean absolute magnitude \\
-		$s_c$ & Colour bias\\
+		%$s_c$ & Colour bias\\
 		\hline
@@
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We must also do this without assuming a specific dispersion model (either the {\gten} or {\celeven} model).
-We address this in our analysis in two ways - increasing the reported uncertainty, and allowing for a global shift in observed colour.
+We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.
 
@@
 
-Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
+%Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
 By comparing simulated true colour values to simulated observed colour values, we can quantify the mean bias in colour in redshift bins, giving us {\green a redshift-dependent array of biases.
@@
 
-It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
+%It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
 Our fiducial cosmology is a Flat $\Lambda$CDM universe with $\Omega_m = 0.3$ and $H_0 = 70\kmsmpc$.
 
-{\red MAKE SOME ILLUSTRATIVE PLOT FOR THIS SECTION}
+%{\red MAKE SOME ILLUSTRATIVE PLOT FOR THIS SECTION}
 
@@
 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $s_c$, $\delta \mathcal{Z}_i$, $\delta S$};
+	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$, $\delta S$};
 
@@
 		$\kappa_{c1}^{1}$ &   -0.01 &      0.01     \\
-		$s_c^{0}$ &    0.00 &      0.01     \\
+		%$s_c^{0}$ &    0.00 &      0.01     \\
 		$\delta(0)$ &    0.00 &      0.00     \\
@@
 This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain.
 Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+% Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
 
@@
 As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
-Strong anti-correlation between the $\kappa_{c0}^0$ and $\kappa_{c0}^1$ (and positive fit values) shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
-Finally, we notice the blank cross in the middle of the figure, the result of four parameters.
-Firstly, the redshift-dependent extra colour dispersion for the LowZ simulation is negligible (due to the low redshift values), the bias correciton $s_c^0$ is unconstrained by our fits, and as the SNANA simulations do not include mass, we expect zero correlation across the mass step $\delta(0)$ and $\delta(\infty)/\delta(0)$ parameters.
+Strong anti-correlation between $\kappa_{c0}^0$ and $\kappa_{c0}^1$ with $\sigma_{m_B}$ reveals the strong population degeneracy, and -- for the C11 simulation results -- a constrained positive value shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
 }
@@
 Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-As expected, $w$ bias is introduced primarily through incorrectly determined $\beta$, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
-Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}, however none have as significant an impact on cosmology as $\beta$.
-Table \ref{tab:parameter_correlations} does show that $\langle M_B \rangle$ is strongly correlated with $w$, however it is recovered much more accurately than $\beta$.
+As expected, $w$ bias would be introduced primarily through incorrectly determined standardisation parameters and mean absolute magnitude, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
+Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}.
 The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
@@
 For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-{\green For our full systematics model, the bias represents a deviation between $-0.3\sigma$ to $0.3\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
+{\green For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
 
@@
 
+\citep{Scolnic2017}
+
 {\red Introduce BBC. Highlight its use on Panstars}
@@
 {\red For systematics paper, also find 60\% reduction in uncert using planck priors}
+
+\begin{table}
+	\centering
+	\caption{Good table right here. Yeah boi.}
+	\label{tab:bulk_summary_bbc}
+	\begin{tabular}{l|ccc}
+		\hline
+		Model & {\gten} & {\celeven} & Mean ({\gten} + {\celeven}) \\
+		\hline
+		Steve $\langle w \rangle$           & $-0.998$ & $-0.945$ & $-1.010$ \\
+		BBC $\langle w \rangle$             & $-1.044$ & $-0.978$ & $-0.972$ \\
+		(Steve - BBC)  $\langle w \rangle$  & $-0.044$ & $-0.033$ & $-0.039$ \\
+		\hline
+		(Steve - BBC)  $\sigma_w$           & $0.057$ & $0.062$ & $0.060$ \\
+		\hline
+	\end{tabular}
+\end{table}
 

### 2018-09-19 papers/methods/paper.tex
Adding comparison to BBC

@@
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-{\green A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size.
-Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.}
+A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size.
+Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.
 
@@
 
-{\green 
 The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by when performing supernova cosmology analyses.
@@
 Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
-}
 
@@
 	\centering
-	\caption{{\green Parameters defined in our model and a summary of their use.
+	\caption{ Parameters defined in our model and a summary of their use.
 The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova.
-The bottom section shows the observables (not parameters) that are the input data to the model.}}
+The bottom section shows the observables (not parameters) that are the input data to the model.}
 	\label{tab:param_summary}
@@
 Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
-{\green Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.
-Tests with changing the selection efficiency model show that the uncertainty introduced is sub-dominant to the uncertainty introduced from assuming an analytic form.}
+Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.
+Tests with changing the intrinsic scatter model used in the selection efficiency simulations show that the uncertainty introduced is negligible.
 
@@
 In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics.
-{\green Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
-Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.} The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
+Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
+Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
+The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
@@
 With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$.
-However, their deviation from the truth value represents a shift of approximately $0.3\sigma$ when taking into account the uncertainty on fits to $w$.
-{\green The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
+However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
+The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
 We also note that the simulations do not vary cosmological parameters nor population.
-As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.}}
+As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.}
 	\label{tab:bulk_summary}
@@
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+We then also test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
-{\green Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
-We note that the incorrect summary statistics do not only bias cosmology, but also cause significant bias with the recovery of the standardisation parameters $\alpha$ and $\beta$.
-Fixing $\alpha$ and $\beta$ values to simulation values when fitting aids in recovering correct cosmology, but does not remove all bias.}
+Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
 
@@
 For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-{\green For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.}
+For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
 
@@
 
-\citep{Scolnic2017}
-
-{\red Introduce BBC. Highlight its use on Panstars}
-
-{\red have plots for w-w comparison for g10 and c11}
-
-{\red conclude that correlation is good, yay, have extra uncertainty, but this can be explained from previous section}
-
-BHM Mean for G10: -0.998, BBC mean -1.042, average diff -0.044, difference scatter 0.057
-
-BHM Mean for C11: -0.945, BBC mean -0.978, average diff -0.033, difference scatter 0.062
-
-{\red For systematics paper, also find 60\% reduction in uncert using planck priors}
+We compare the results of our model against those of the BBC method \citep{Kessler2017}.
+BBC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017}, as a leading supernova cosmology method provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
+
+To this end, we take the results of the BBC method also run on our set of 200 validation simulations and compare the recovered $w$ values to those of our method.
+The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
+These show that the difference in recovered $w$ over different intrinsic scatter models is comparable, with our methodology having a small (but not statistically significant) reduction in the sensitivity to intrinsic scatter model.
+However, as the BBC method introduces a bias to lower $w$ for {\gten} and a bias towards higher $w$ for {\celeven}, taking the average result over both {\gten} and {\celeven} only results in a small bias of $0.01$ in $w$.
+As our method recovers $w$ without bias for {\gten} and finds a biased high $w$ for {\celeven}, the average bias over the two scatter models is $-0.028$.
+When comparing both the {\gten} and {\celeven} set of simulations independently, our model differs from BBC in its average prediction of $w$ by $-0.044$ and $-0.033$ respectively.
+For the {\gten} model this difference is a result of bias in the BBC results, however for the {\celeven} simulations this is a result of both bias from BBC, and a larger bias from our method.
+These results also allow us to state the expectation values for $w$ when run on the DES three year spectroscopic survey sample -- factoring in the calculated uncertainty from Planck priors (shrinking the scatter from $0.06$ to $0.04$), we expect our BHM method to, on average, recover $w_{\rm BHM} = w_{\rm BBC} -0.04\pm0.04$.
+
+Having established that existing methodologies exhibit similar shifts in their recovery of $w$ compared to our method, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
 
@@
 	\centering
-	\caption{Good table right here. Yeah boi.}
+	\caption{Characterising the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model.
+The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation.
+Finally, we also characterise the scatter between the methods in the final row.}
 	\label{tab:bulk_summary_bbc}
@@
 	\end{center}
-	\caption{Whoa.}
+	\caption{Recovered $w$ for the 200 validation simulations with full treatment of statistical and systematic errors.
+Uncertainty on the recovered $w$ value is shown for every second data point for visual clarity.}
 	\label{fig:bulk_comparison}

### 2018-09-20 papers/methods/paper.tex
Tams edits and help, finishing sections on comparison and uncert

@@
 
-With the increased flexibility of Bayesian hierarchical models over traditional modes, we expect to find an increased uncertainty on parameter inference.
-To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to vary or locked to a fixed value.
+With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
+To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to either vary or stay locked to a fixed value.
 By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
@@
 	\centering
-	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features.}
+	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features.
+We start from the top of the table, only varying cosmological parameters $\Omega_m$ and $w$, and then progressively unlock parameters and let them fit as we progress down the table.
+The cumulative uncertainty shows the total uncertainty on $w$ on the fit for all, where the $\sigma_w$ term is derived by taking the quadrature difference in cumulative uncertainty as we progress.}
 	\label{tab:uncert}
@@
 		Intrinsic scatter              & 0.020  &   0.071       \\
-		Static survey populations      & 0.022  &   0.074       \\
+		Redshift-independent populations      & 0.022  &   0.074       \\
 		Redshift-dependent populations & 0.030  &   0.080       \\
@@
 
-The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is still dominated by statistical error, as the total statistical uncertainty is on $w$ is $0.08$.
+The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is still dominated by statistical error, as the total statistical uncertainty is on $w$ is $\pm0.08$.
 With the low number of supernovae in the DES three-year spectroscopic sample, this is expected.
@@
 We compare the results of our model against those of the BBC method \citep{Kessler2017}.
-BBC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017}, as a leading supernova cosmology method provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
+BBC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017}.
+As a leading supernova cosmology method, it provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
 
@@
 The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
-These show that the difference in recovered $w$ over different intrinsic scatter models is comparable, with our methodology having a small (but not statistically significant) reduction in the sensitivity to intrinsic scatter model.
-However, as the BBC method introduces a bias to lower $w$ for {\gten} and a bias towards higher $w$ for {\celeven}, taking the average result over both {\gten} and {\celeven} only results in a small bias of $0.01$ in $w$.
-As our method recovers $w$ without bias for {\gten} and finds a biased high $w$ for {\celeven}, the average bias over the two scatter models is $-0.028$.
-When comparing both the {\gten} and {\celeven} set of simulations independently, our model differs from BBC in its average prediction of $w$ by $-0.044$ and $-0.033$ respectively.
+
+These results show that both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
+The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
+Our method shows a small (but not statistically significant) improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}.
+The average bias over the two scatter models is $+0.028$, representing a larger bias than the BBC method.
+
+When comparing both the {\gten} and {\celeven} set of simulations independently, our model differs from BBC in its average prediction of $w$ by $+0.044$ and $+0.033$ respectively.
 For the {\gten} model this difference is a result of bias in the BBC results, however for the {\celeven} simulations this is a result of both bias from BBC, and a larger bias from our method.
-These results also allow us to state the expectation values for $w$ when run on the DES three year spectroscopic survey sample -- factoring in the calculated uncertainty from Planck priors (shrinking the scatter from $0.06$ to $0.04$), we expect our BHM method to, on average, recover $w_{\rm BHM} = w_{\rm BBC} -0.04\pm0.04$.
-
-Having established that existing methodologies exhibit similar shifts in their recovery of $w$ compared to our method, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
+These results also allow us to state the expected values for $w$ when run on the DES three year spectroscopic survey sample.
+When using Planck priors our uncertainty on $w$ is reduced compared to using our simulation Gaussian prior on $\Omega_m$, shrinking the average scatter from $0.06$ to $0.04$.
+After factoring this into our uncertainty, we expect our BHM method to, on average, recover $w_{\rm BHM} = w_{\rm BBC} +0.04\pm0.04$.
+
+Having established that our method exhibits similar shifts in their recovery of $w$ compared to BBC, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
 
@@
 The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation.
-Finally, we also characterise the scatter between the methods in the final row.}
+Finally, we also characterise the scatter between the methods in the final row.
+Averages are computed giving each simulation sample the same weight.}
 	\label{tab:bulk_summary_bbc}
@@
 		\hline
-		Steve $\langle w \rangle$           & $-0.998$ & $-0.945$ & $-1.010$ \\
-		BBC $\langle w \rangle$             & $-1.044$ & $-0.978$ & $-0.972$ \\
-		(Steve - BBC)  $\langle w \rangle$  & $-0.044$ & $-0.033$ & $-0.039$ \\
+		Steve $\langle w \rangle$           & $-0.998$ & $-0.945$ & $-0.972$\\
+		BBC $\langle w \rangle$             & $-1.044$ & $-0.978$ & $-1.010$ \\
+		(Steve - BBC)  $\langle w \rangle$  & $+0.044$ & $+0.033$ & $+0.039$ \\
 		\hline

### 2018-09-20 dessn/configurations/bulk.py

@@ -140,7 +140,7 @@ if __name__ == "__main__":
                 truth[r"$\beta$"] = None
                 call.plotter.plot(filename=pfn + "_poster.png", truth=truth, parameters=4, extents=ex)
 
-            if False:
+            if True:
                 bbcs = [np.loadtxt(plot_dir + "dillon_g10.txt"), np.loadtxt(plot_dir + "dillon_c11.txt")]
 
                 import matplotlib.pyplot as plt
@@ -161,6 +161,11 @@ if __name__ == "__main__":
 
                         diff = means_bbc - means_steve
                         print("Diff ", t, np.mean(means_steve), np.mean(means_bbc), np.mean(diff), np.std(diff))
+
+                        bbc_mean, bbc_std = weighted_avg_and_std(means_bbc, 1/std_bbc**2)
+                        bhm_mean, bhm_std = weighted_avg_and_std(means_steve, 1/std_steve**2)
+                        print(t, bbc_mean, bbc_std, bhm_mean, bhm_std)
+
                         res[t + "_Steve"] = means_steve
                         res[t + "_BBC"] = means_bbc
                         minv = min(np.min(means_steve), np.min(means_bbc)) - 0.02
@@ -217,7 +222,7 @@ if __name__ == "__main__":
         # with open(pfn + "_summary.txt", "w") as f:
         #     f.write(c2.analysis.get_latex_table(transpose=True))
         #
-        if True:
+        if False:
             pps = ['$\\Omega_m$', '$w$', '$\\alpha$', '$\\beta$', '$\\langle M_B \\rangle$',
                '$\\sigma_{\\rm m_B}^{0}$', '$\\sigma_{\\rm m_B}^{1}$', '$\\sigma_{x_1}^{0}$',
                '$\\sigma_{x_1}^{1}$', '$\\sigma_{c}^{0}$', '$\\sigma_{c}^{1}$', '$\\alpha_c^{0}$',

### 2018-09-24 papers/methods/paper.tex
Updating with feedback from Dan

@@
 \newcommand {\etal} {\emph{~et~al.} }
-\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
+\definecolor{forestgreen}{rgb}{0.13, 0.75, 0.13}
 \newcommand{\green}{\color{forestgreen}}
@@
 We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample.
-SNANA simulations can recover the input cosmology to within one third of the typical uncertainty for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and being more simulation independent than previous analyses.
+SNANA simulations can recover the input cosmology {\green between $0$ and $0.5\sigma$ bias depending on the intrinsic scatter model} for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and increasing simulation independent over previous analyses.
+{\green Future development will focus on increasing model agnosticism towards intrinsic scatter model.} {\red Have left out explicit references for `previous analyses' as I dont want to bloat the abstract.
+Its talked about in the paper, so is this alright?}
 \end{abstract}
@@
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Mosher2014}.
-The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
-As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the width of this distribution is called the intrinsic dispersion of the supernova population and the hat denotes an observed variable.
+{\green This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.} {\red  Ive just swapped out references here as this isnt a critical point for my analysis to drill down on.} The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
+As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
+{\red Removed confusing statement on intrinsic dispersion.}
 
@@
 A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size.
-Whilst binning data does result in some loss of information, recent works tested against simulations show that this information is not cosmologically significant \citep{Scolnic2016, Kessler2017}.
+Whilst binning data does result in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
 
@@
 None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
+{\red Dan, you mention underselling the prize here.
+Any suggestions on an appropriate sell, so to speak?}
 
@@
 The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-As the SALT2 model does not include intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+As the SALT2 model does not include {\green full treatment of} intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
 These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population.
@@
 We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by $\sigma_{M_B}$ without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
+We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the {\green width of the underlying magnitude population (discussed in Section \ref{sec:underlying})} without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
+{\green Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.}
 
@@
 We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature).
-The $\kappa$ parameters are tightly constrained with priors as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
-We subject them to strong Cauchy priors centered on zero and with width $0.002$, where the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}.
-The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulting in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion.
-Each survey has its own set of $\kappa$ parameters.
-As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
-
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
+{\green We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
+We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.} As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
+
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, {\green by modelling the dispersion as a multivariate Gaussian}, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
 We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
+This extra dispersion is modelled with redshift independence.
+{\red I'm not too sure what you mean about telling a story here.
+I've tried to change the wording to put it back in line with the method, is that sufficient?
+Also, I did not try redshift dependent extra scatter on $m_B$.}
 
@@
 \subsubsection{Type Ia population} 
-
+\label{sec:underlying}
 The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude.
@@
 For future use with photometrically classified surveys, we include a simplistic outlier population model.
-We follows {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
+We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
 For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works.
@@
 This degeneracy resulted in a significant bias in recovered cosmology.
-As $\alpha$ and $\beta$ are constants in the SALT2 model, assuming varying $\alpha$ or $\beta$ is only self consistent if the same variations are used in the light curve fitting process.
-As this variation is not present in the DES spectroscopic sample's light curve fitting process, and because of the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+{\red Removed comments about varying alpha and beta.} {\green Due to the }recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
 As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.
@@
 
-Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population, but recent analyses require a more sophisticated approach.
+Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population {\green \citep{Conley2011, Betoule2014}}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
@@
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
-Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not significantly impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
+Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
 The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
-We note that there was no significant difference in the selection effect of Malmquist bias between using the {\gten} or {\celeven} scatter model compared to the fit uncertainty.
+{\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the selection effect of Malmquist bias between the two models.}
 
@@
 The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-Our recovered results are all within $0.5\sigma$ of the true values, however some bias does emerge.
+{\green Our recovered results are all within $0.5\sigma$ of the true values of $\Omega_m = 0.3$, $w=-1$.}
 
 We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
-This is confirmed by replacing the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+{\green To confirm this, we run two tests.
+The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-We then also test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$.
-From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
+Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
 Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
 
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
 It is difficult to do whilst accounting for correlations with population and scatter model.

### 2018-09-24 papers/methods/paper.tex
More feedback

@@
 
-These results show that both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
+{\green As shown in GET-BROUT-CITATION, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
+As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}.
+As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
+These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both} BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
 The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
-Our method shows a small (but not statistically significant) improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}.
+Our method shows a small improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}.
+This decrease in error is not statistically significant as we have statistical uncertainty of $\sim0.01$ for $100$ simulation realisations.
 The average bias over the two scatter models is $+0.028$, representing a larger bias than the BBC method.

### 2018-09-25 papers/methods/paper.tex
Updating paper for Dan

@@
 \usepackage{enumerate}
+%\usepackage{tabu}
+%\usepackage{tabularx}
 \DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
@@
 		$\delta\mathcal{Z}_i$ & Systematics scale\\
-		$\delta S$ & Selection effect deviation \\
 		$\langle M_B \rangle$ & Mean absolute magnitude \\
@@
 		\hline
+		$\delta S$ & Selection effect deviation \\
 		$\langle x_1^i \rangle$ & Mean stretch nodes\\
@@
 		$\hat{m}$ & Determined mass probability \\
-		$\hat{p}$ & Determined outlier probability \\
+		$\hat{p}$ & Determined {\green Ia} probability \\
 		\hline
@@
 	% Survey nodes
-	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\rho,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
-	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\Delta_i, \delta_i$};
+	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
+	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
 	% Global nodes
 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$, $\delta S$};
+	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
 
@@
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
-{\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the selection effect of Malmquist bias between the two models.}
+{\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.}
 
@@
 
+\begin{table*}
+	\centering
+	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
+We show the average parameter mean and standard deviation, with the simulation scatter shown in brackets.
+Truth values for the simulations are $\alpha=0.15$, $\langle M_B \rangle = -19.365$, $\beta_{\rm G10} = 3.1$, $\beta_{\rm C11} = 3.8$.
+The width of the intrinsic scatter does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.}
+	\label{tab:standardisation}
+	\begin{tabular}{l|ccccc}
+		\hline
+		Model & $\alpha$ & $\beta$ & $\langle M_B \rangle$ & $\sigma_{m_B}^0$ &  $\sigma_{m_B}^1$ \\
+		\hline
+		{\gten} Stat + Syst  & $0.152\pm0.009\ (0.008)$ & $ 3.44\pm0.19\ (0.18)$ & $-19.367\pm0.028\ (0.015)$ & $0.070\pm0.022\ (0.018)$ & $0.073\pm0.025\ (0.022)$\\
+		{\gten}  Stat         & $0.150\pm0.008\ (0.008)$ & $3.43\pm0.20\ (0.17)$ & $-19.374\pm0.016\ (0.013)$ & $0.069\pm0.023\ (0.019)$ & $0.072\pm0.026\ (0.023)$\\
+		{\celeven} Stat + Syst  &  $0.152\pm0.009\ (0.007)$ & $3.76\pm0.15\ (0.13)$ & $-19.351\pm0.030\ (0.018)$ & $0.024\pm0.016\ (0.011)$ & $0.029\pm0.020\ (0.014)$ \\
+		{\celeven} Stat         &  $0.150\pm0.008\ (0.007)$ & $3.75\pm0.16\ (0.13)$ & $-19.359\pm0.016\ (0.015)$ & $0.025\pm0.016\ (0.012)$ & $0.027\pm0.020\ (0.015)$\\
+		\hline
+	\end{tabular}
+\end{table*}
+
 \begin{figure}
@@
 	\caption{Maximum posterior points for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel.
-Points are shown for parameters $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$, with the other fit parameters being marginalised over.
+Points are shown for parameters $\Omega_m$, $w$, $\alpha$ and $\beta$, with the other fit parameters being marginalised over.
 As we are unable to fully correct observed summary statistics, a step required by the lack of intrinsic scatter in the SALT2 model, we expect to see an offset in $\alpha$ and $\beta$.
@@
 		\hline
-		Model & $w$ (scatter) & Bias \\
-		\hline
-		{\gten} Stat + Syst  &  $-1.00\pm 0.10\ (0.08)$ & $0.0\sigma$  \\
-		{\celeven} Stat + Syst  &  $-0.95\pm 0.10\ (0.06)$ & $0.5\sigma$  \\
-		{\gten}  Stat         &  $-1.00\pm 0.08\ (0.08)$ & $0.0\sigma$   \\
-		{\celeven} Stat         &  $-0.95\pm 0.08\ (0.05)$ & $0.6\sigma$   \\
+		Model & $w$ (scatter) & $w$-Bias \\
+		\hline
+		{\gten} Stat + Syst  &  $-0.998\pm 0.097\ (0.073)$ & $(0.02\pm0.07)\sigma$  \\
+		{\gten}  Stat         &  $-1.008\pm 0.080\ (0.068)$ & $(-0.10\pm0.08)\sigma$   \\
+		{\celeven} Stat + Syst  &  $-0.945\pm 0.098\ (0.077)$ & $(0.55\pm0.08)\sigma$  \\
+		{\celeven} Stat         &  $-0.948\pm 0.079\ (0.066)$ & $(0.65\pm0.08)\sigma$   \\
 		\hline
@@
 
-Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
-As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
-
-Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-The parameter bounds are listed in Table \ref{tab:bulk_summary}.
-{\green Our recovered results are all within $0.5\sigma$ of the true values of $\Omega_m = 0.3$, $w=-1$.}
-
-We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
-{\green To confirm this, we run two tests.
-The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
-With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
-Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
-
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model) and binning data.
-It is difficult to do whilst accounting for correlations with population and scatter model.
-This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit inherently.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
-
-% Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
-
 \begin{figure*}
@@
 
+Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
+
+Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
+The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  {\green and secondary parameters are shown in Table \ref{tab:standardisation}.
+Our recovered results for $w$ are all within $0.5\sigma$ of the true values of $w=-1$.
+Table \ref{tab:standardisation} shows a clear difference in both $\beta$ and $\sigma_{m_B}$ across the {\gten} and {\celeven} simulations.
+As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations.
+The extra smearing of the {\celeven} model does not result in a biased $\beta$ value, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
+Interestingly, $w$-bias is only found for the {\celeven} simulations.}
+
+We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+{\green To confirm this, we run two tests.
+The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
+Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
+Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+
+Unfortunately, adding extra fit parameters to allow for shifting observables {\green washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model.} This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically.
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+% Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
+
 Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
 Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-As expected, $w$ bias would be introduced primarily through incorrectly determined standardisation parameters and mean absolute magnitude, which itself is tied to the unexplained colour dispersion and biases in colour introduced by the unknown intrinsic scatter model.
+{\green As expected, $w$ bias would be introduced primarily through incorrectly determined standardisation parameters, mean absolute magnitude or incorrect dispersion parameters.
+As discussed previously, the culprit for our $w$-bias is the dispersion model, rather than standardisation parameters}.
 Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}.
@@
 
-With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
-To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to either vary or stay locked to a fixed value.
-By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
-
-\begin{table}
+\begin{table*}
 	\centering
@@
 	\label{tab:uncert}
-	\begin{tabular}{l|rr}
-		Feature & $\sigma_w$ & Cumulative  \\
-		\hline
-		Cosmology only                 & 0.051  &   0.051       \\
-		Standardisation parameters     & 0.046  &   0.068       \\
-		Intrinsic scatter              & 0.020  &   0.071       \\
-		Redshift-independent populations      & 0.022  &   0.074       \\
-		Redshift-dependent populations & 0.030  &   0.080       \\
-		Systematics                    & 0.054  &   0.096       \\
+	\begin{tabular}{l|l|rr}
+		Feature & {\green Parameters} & $\sigma_w$ & Cumulative  \\
+		\hline
+		Cosmology                      & $\Omega_m$, $w$ & 0.051  &   0.051       \\
+		Standardisation                & $\alpha$, $\beta$, $\langle M_B \rangle$, $\delta(0)$, $\delta(\infty)/\delta(0)$ & 0.046  &   0.068       \\
+		Intrinsic scatter              & $\kappa_0$, $\kappa_1$ & 0.020  &   0.071       \\
+		Redshift-independent populations             & $\sigma_{M_B}$, $\sigma_c$, $\sigma_{x_1}$, $\alpha_c$ & 0.022  &   0.074       \\
+		Redshift-dependent populations & $\langle c_i \rangle$, $\langle x_{1,i} \rangle$ & 0.030  &   0.080   \\
+		Systematics                    & $\delta \mathcal{Z}_i$, $\delta S$ & 0.054  &   0.096      \\ 
 		\hline
 	\end{tabular}
-\end{table}
+\end{table*}
+
+With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
+To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to either vary or stay locked to a fixed value.
+By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
 
@@
 
-{\green As shown in GET-BROUT-CITATION, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
+{\green As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
 As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}.

### 2018-09-25 dessn/configurations/bulk.py

@@ -61,6 +61,8 @@ if __name__ == "__main__":
             cs = ['b', 'g', 'p', 'a']
             cs = ['g', 'r', 'k', 'k']
 
+            pars = ["$w$", r"$\Omega_m$", r"$\alpha$", r"$\beta$", '$\\langle M_B \\rangle$',
+               '$\\sigma_{\\rm m_B}^{0}$', '$\\sigma_{\\rm m_B}^{1}$']
             ws = {}
             ws_std = {}
             for m, s, ci, chain, truth, weight, old_weight, posterior in res:
@@ -96,12 +98,17 @@ if __name__ == "__main__":
                     else:
                         cc = c2
                     if ws.get(name) is None:
-                        ws[name] = []
+                        ws[name] = {}
                     if ws_std.get(name) is None:
-                        ws_std[name] = []
-                    w_mean, w_std = weighted_avg_and_std(chain["$w$"], weight)
-                    ws[name].append(w_mean)
-                    ws_std[name].append(w_std)
+                        ws_std[name] = {}
+                    for p in pars:
+                        w_mean, w_std = weighted_avg_and_std(chain[p], weight)
+                        if ws[name].get(p) is None:
+                            ws[name][p] = []
+                        if ws_std[name].get(p) is None:
+                            ws_std[name][p] = []
+                        ws[name][p].append(w_mean)
+                        ws_std[name][p].append(w_std)
                     cc.add_chain(chain, weights=weight, posterior=posterior, name=name, plot_contour=False, plot_point=True, color=col, marker_style=ms, marker_size=msize)
                     if True or "Syst" in name:
                         ms = "+" if "G10" in name else "o"
@@ -113,11 +120,14 @@ if __name__ == "__main__":
 
             with open(pfn + "_res.txt", "w") as f:
                 for k in ws.keys():
-                    s = "%s %5.3f %5.3f (%5.3f) : %4.2f\n" % (k, np.mean(ws[k]), np.mean(ws_std[k]), np.std(ws[k]), np.sqrt(100)*(-1-np.mean(ws[k]))/np.std(ws[k]))
-                    # mls = [i["$w$"][1] for i in call.analysis.get_summary(parameters=["$w$"], chains=k)]
-                    print(s)
-                    f.write(s)
-                    # print(np.mean(mls))
+                    for p in ws[k].keys():
+                        means = ws[k][p]
[…]

### 2018-10-01 papers/methods/paper.tex
Updating paper

@@
 We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample.
-SNANA simulations can recover the input cosmology {\green between $0$ and $0.5\sigma$ bias depending on the intrinsic scatter model} for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and increasing simulation independent over previous analyses.
-{\green Future development will focus on increasing model agnosticism towards intrinsic scatter model.} {\red Have left out explicit references for `previous analyses' as I dont want to bloat the abstract.
-Its talked about in the paper, so is this alright?}
+SNANA simulations can recover the input cosmology between $0$ and half the statistical uncertainty from a DES 3-year spectroscopic sample.
+The method fits cosmology, population and systematics simultaneously while taking into account an increased number of systematics and minimising simulation dependence.
+Future development will focus on increasing model agnosticism towards intrinsic scatter model.
 \end{abstract}
@@
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-{\green This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.} {\red  Ive just swapped out references here as this isnt a critical point for my analysis to drill down on.} The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
+This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.
+The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
 As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
-{\red Removed confusing statement on intrinsic dispersion.}
 
@@
 
-Whilst there are a large number of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
-None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
-{\red Dan, you mention underselling the prize here.
-Any suggestions on an appropriate sell, so to speak?}
+Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias.
+Additionally, tests that have been done have generally been performed on either  $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology and not $w$CDM cosmology.
+However, quantifying the biases on $w$CDM cosmology with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
 
@@
 The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-As the SALT2 model does not include {\green full treatment of} intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
 These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population.
@@
 		$\hat{m}$ & Determined mass probability \\
-		$\hat{p}$ & Determined {\green Ia} probability \\
+		$\hat{p}$ & Determined Ia probability \\
 		\hline
@@
 We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the {\green width of the underlying magnitude population (discussed in Section \ref{sec:underlying})} without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
-{\green Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.}
+We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
+Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.
 
@@
 The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
-{\green We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
-We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.} As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
-
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, {\green by modelling the dispersion as a multivariate Gaussian}, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
+We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.
+As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
+
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
 We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
 This extra dispersion is modelled with redshift independence.
-{\red I'm not too sure what you mean about telling a story here.
-I've tried to change the wording to put it back in line with the method, is that sufficient?
-Also, I did not try redshift dependent extra scatter on $m_B$.}
 
@@
 This degeneracy resulted in a significant bias in recovered cosmology.
-{\red Removed comments about varying alpha and beta.} {\green Due to the }recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
+Due to the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
 As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.
@@
 
-Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population {\green \citep{Conley2011, Betoule2014}}, but recent analyses require a more sophisticated approach.
+Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
@@
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
-{\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.}
+We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.
 
@@
 Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  {\green and secondary parameters are shown in Table \ref{tab:standardisation}.
+The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  and secondary parameters are shown in Table \ref{tab:standardisation}.
 Our recovered results for $w$ are all within $0.5\sigma$ of the true values of $w=-1$.
@@
 The extra smearing of the {\celeven} model does not result in a biased $\beta$ value, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
-Interestingly, $w$-bias is only found for the {\celeven} simulations.}
+Interestingly, $w$-bias is only found for the {\celeven} simulations.
 
 We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
-{\green To confirm this, we run two tests.
+To confirm this, we run two tests.
 The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
+Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
 Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
 
-Unfortunately, adding extra fit parameters to allow for shifting observables {\green washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model.} This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically.
+Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model.
+This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically.
 Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
@@
 Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-{\green As expected, $w$ bias would be introduced primarily through incorrectly determined standardisation parameters, mean absolute magnitude or incorrect dispersion parameters.
-As discussed previously, the culprit for our $w$-bias is the dispersion model, rather than standardisation parameters}.
 Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}.
@@
 	\begin{tabular}{l|l|rr}
-		Feature & {\green Parameters} & $\sigma_w$ & Cumulative  \\
+		Feature &  Parameters & $\sigma_w$ & Cumulative  \\
 		\hline
@@
 
-{\green As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
+As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
 As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}.
 As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
-These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both} BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
+These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
 The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
@@
 Finally, we also characterise the scatter between the methods in the final row.
-Averages are computed giving each simulation sample the same weight.}
+Averages are computed giving each simulation sample the same weight.
+In the model, $\Delta$ represents Steve - BBC.
+The final row shows the scatter between Steve and BBC for the different simulations.}
 	\label{tab:bulk_summary_bbc}
@@
 		\hline
-		Model & {\gten} & {\celeven} & Mean ({\gten} + {\celeven}) \\
-		\hline
-		Steve $\langle w \rangle$           & $-0.998$ & $-0.945$ & $-0.972$\\
-		BBC $\langle w \rangle$             & $-1.044$ & $-0.978$ & $-1.010$ \\
-		(Steve - BBC)  $\langle w \rangle$  & $+0.044$ & $+0.033$ & $+0.039$ \\
-		\hline
-		(Steve - BBC)  $\sigma_w$           & $0.057$ & $0.062$ & $0.060$ \\
+		Model & {\gten} & {\celeven} & ({\gten} + {\celeven}) \\
+		\hline
+		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
+		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$ \\
+		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
+		\hline
+		$\Delta$  $\sigma_w$           & $0.057\pm0.004$ & $0.062\pm0.004$ & $0.060\pm0.003$ \\
 		\hline

### 2018-10-01 dessn/configurations/bulk.py

@@ -150,27 +150,31 @@ if __name__ == "__main__":
                 truth[r"$\beta$"] = None
                 call.plotter.plot(filename=pfn + "_poster.png", truth=truth, parameters=4, extents=ex)
 
-            if False:
+            if True:
                 bbcs = [np.loadtxt(plot_dir + "dillon_g10.txt"), np.loadtxt(plot_dir + "dillon_c11.txt")]
 
                 import matplotlib.pyplot as plt
                 fig, axes = plt.subplots(nrows=1, figsize=(5, 5))
                 axes = [axes, axes]
                 res = {}
+                m_s, m_b, m_d = [], [], []
                 for t, ax, bbc in zip(["G10", "C11"], axes, bbcs):
                     for key in ws.keys():
                         if t not in key:
                             continue
                         if "Syst" not in key:
                             continue
-                        means_steve = np.array(ws[key])
-                        std_steve = np.array(ws_std[key])
+                        means_steve = np.array(ws[key]["$w$"])
+                        std_steve = np.array(ws_std[key]["$w$"])
 
                         means_bbc = bbc[:, 0] - 1
                         std_bbc = bbc[:, 1]
-
                         diff = means_bbc - means_steve
-                        print("Diff ", t, np.mean(means_steve), np.mean(means_bbc), np.mean(diff), np.std(diff))
+
+                        m_d.append(diff)
+                        m_b.append(means_bbc)
+                        m_s.append(means_steve)
+                        print("Diff %s %7.4f %7.4f %7.4f %7.4f %7.4f %7.4f %7.4f" % (t, np.mean(means_steve), np.std(means_steve), np.mean(means_bbc), np.std(means_bbc), np.mean(diff), np.std(diff), np.std(diff)/np.sqrt(2*len(diff))))
 
                         bbc_mean, bbc_std = weighted_avg_and_std(means_bbc, 1/std_bbc**2)
                         bhm_mean, bhm_std = weighted_avg_and_std(means_steve, 1/std_steve**2)
@@ -186,6 +190,15 @@ if __name__ == "__main__":
                         ecolor = "#888888" if "Syst" not in key else ("#a0e8a3" if "G10" in key else "#f9a19f")
[…]

### 2018-10-02 papers/methods/paper.tex
Updating abstract

@@
 \newcommand{\gten}{\citetalias{Guy2010}}
+\newcommand{\steve}{\textit{Steve}}
 \newcommand{\celeven}{\citetalias{Chotard2011}}
@@
 \begin{abstract}
-We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects.
-This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency.
-Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery.
-We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample.
-SNANA simulations can recover the input cosmology between $0$ and half the statistical uncertainty from a DES 3-year spectroscopic sample.
-The method fits cosmology, population and systematics simultaneously while taking into account an increased number of systematics and minimising simulation dependence.
-Future development will focus on increasing model agnosticism towards intrinsic scatter model.
+We present a new Bayesian hierarchical model named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
+%
+This advances previous works with Bayesian hierarchical models by including 
+an improved treatment of Malmquist bias, 
+accounting for additional sources of systematic uncertainty, 
+and increasing numerical efficiency. 
+%
+Given light curve fit parameters, redshifts, and host-galaxy masses, {\steve} simultaneously fits for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
+Selection effects are characterised using Monte-Carlo simulations.
+
+We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
+%
+Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
+%
+These simulated datasets contain more than $250\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
+%
+This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model.
+This $w$-bias is less than $0.03$ on average, roughly half the uncertainty on $w$.
+%
+These simulation test results are a concern for upcoming large surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
 \end{abstract}
@@
 
+Funding for the DES Projects has been provided by the U.S. Department of Energy, the U.S. National Science Foundation, the Ministry of Science and Education of Spain, 
+the Science and Technology Facilities Council of the United Kingdom, the Higher Education Funding Council for England, the National Center for Supercomputing 
+Applications at the University of Illinois at Urbana-Champaign, the Kavli Institute of Cosmological Physics at the University of Chicago, 
+the Center for Cosmology and Astro-Particle Physics at the Ohio State University,
+the Mitchell Institute for Fundamental Physics and Astronomy at Texas A\&M University, Financiadora de Estudos e Projetos, 
+Funda{\c c}{\~a}o Carlos Chagas Filho de Amparo {\`a} Pesquisa do Estado do Rio de Janeiro, Conselho Nacional de Desenvolvimento Cient{\'i}fico e Tecnol{\'o}gico and 
+the Minist{\'e}rio da Ci{\^e}ncia, Tecnologia e Inova{\c c}{\~a}o, the Deutsche Forschungsgemeinschaft and the Collaborating Institutions in the Dark Energy Survey.
+
+The Collaborating Institutions are Argonne National Laboratory, the University of California at Santa Cruz, the University of Cambridge, Centro de Investigaciones Energ{\'e}ticas, 
+Medioambientales y Tecnol{\'o}gicas-Madrid, the University of Chicago, University College London, the DES-Brazil Consortium, the University of Edinburgh, 
+the Eidgen{\"o}ssische Technische Hochschule (ETH) Z{\"u}rich, 
+Fermi National Accelerator Laboratory, the University of Illinois at Urbana-Champaign, the Institut de Ci{\`e}ncies de l'Espai (IEEC/CSIC), 
+the Institut de F{\'i}sica d'Altes Energies, Lawrence Berkeley National Laboratory, the Ludwig-Maximilians Universit{\"a}t M{\"u}nchen and the associated Excellence Cluster Universe, 
+the University of Michigan, the National Optical Astronomy Observatory, the University of Nottingham, The Ohio State University, the University of Pennsylvania, the University of Portsmouth, 
+SLAC National Accelerator Laboratory, Stanford University, the University of Sussex, Texas A\&M University, and the OzDES Membership Consortium.
+
+Based in part on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory, which is operated by the Association of 
+Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation.
+
+The DES data management system is supported by the National Science Foundation under Grant Numbers AST-1138766 and AST-1536171.
+The DES participants from Spanish institutions are partially supported by MINECO under grants AYA2015-71825, ESP2015-66861, FPA2015-68048, SEV-2016-0588, SEV-2016-0597, and MDM-2015-0509, 
+some of which include ERDF funds from the European Union. IFAE is partially funded by the CERCA program of the Generalitat de Catalunya.
+Research leading to these results has received funding from the European Research
+Council under the European Union's Seventh Framework Program (FP7/2007-2013) including ERC grant agreements 240672, 291329, and 306478.
+We  acknowledge support from the Australian Research Council Centre of Excellence for All-sky Astrophysics (CAASTRO), through project number CE110001020, and the Brazilian Instituto Nacional de Ci\^encia
+e Tecnologia (INCT) e-Universe (CNPq grant 465376/2014-2).
+
+This manuscript has been authored by Fermi Research Alliance, LLC under Contract No.
+DE-AC02-07CH11359 with the U.S.
+Department of Energy, Office of Science, Office of High Energy Physics.
+The United States Government retains and the publisher, by accepting the article for publication, acknowledges that the United States Government retains a non-exclusive, paid-up, irrevocable, world-wide license to publish or reproduce the published form of this manuscript, or allow others to do so, for United States Government purposes.
+
 %%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

### 2018-10-04 papers/methods/paper.tex
Initial rick comments

@@
 Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}.
+Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, sako2018sdss}.
 Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB.
 For a recent review, see \citet{Huterer2018}.
-Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.
+{\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}
 
 In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric selected samples.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples.
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
-Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as systematic uncertainties will prove to be the limiting factor in our analyses \citep{Betoule2014, Scolnic2017}.
+{\green Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as we will required reduced systematic uncertainties to fully utilised increase statistics} \citep{Betoule2014, Scolnic2017}.
 
@@
 The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models.
-First used in ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
-\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in cosmological parameters as a function of many factors to better remove fitting biases.
+First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
+\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of many factors to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either also use simulation corrections or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
+Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
 
 In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance.
-Section \ref{sec:review} is dedicated to a quick review of the supernova cosmology, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
+Section \ref{sec:review} is dedicated to a quick review of the {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
@@
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.
+This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}.
 The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
@@
 
-This represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
-Accurately characterising this underlying population, its evolution over redshift, and effects from environment, is one of the challenges of supernova cosmology.
+This {\green ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$} represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
+Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology.
 Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population living in apparent magnitude space.
@@
 Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from rearranging equation \ref{eq:standard}, and the cosmologically predicted values.
-To distinguish them, we denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$.
-This function is given by
+We denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$.
+{\green The $\chi^2$ function minimised is}
 
@@
 
-where $C^{-1}$ is an uncertainty matrix that combines the uncertainty from the SALT2 fits, intrinsic dispersion, calibration, dust, peculiar velocity and many other factors (see \citet{Betoule2014} for a review).
-The cosmological $\mu_C$ is defined as
-
-\begin{align}
-\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
-r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
-\end{align}
-where $r$ is the comoving distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
+where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Betoule2014} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
+{\green 
+\begin{align}
+\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right],\\
+d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
+\end{align}
+}
+where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
 
@@
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A common solution to this computational cost problem is to bin the supernovae, which then produces a matrix of arbitrary size.
-Whilst binning data does result in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
+A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size.
+Whilst binning data results in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
 
@@
 An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}.
-Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then added onto the observational data.
+Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data.
 When using traditional fitting methods, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
@@
 
-To try and escape the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
-\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples, described in \citet{Kessler2015}.
+To avoid the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
+\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples.
 In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.
 
-Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects that plague supernova surveys.
+Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects associated with supernova surveys.
 By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}.
@@
 Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias.
-Additionally, tests that have been done have generally been performed on either  $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology and not $w$CDM cosmology.
-However, quantifying the biases on $w$CDM cosmology with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.
+{\green Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
+However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to $w$.}
 
@@
 
-The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by when performing supernova cosmology analyses.
-In this Section, we outline several of the most prominent challenges.
-
-Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion.
+{\green The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.} In this Section, we outline several of the most prominent challenges.
+
+{\green Forefront among these challenges is our ignorance of the underlying Type Ia population -- its intrinsic dispersion}.
 Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
 We utilise two dispersion models in this work.
-The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation.
-The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
-These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population.
+The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
+The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
+{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population.
 We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.
@@
 The underlying supernova population is further complicated by the presence of outliers.
-Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae.
-This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
-Depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
+Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae {\green and represent outliers from the standardised SN Ia sample.} This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
+Depending on the cuts {\green optimising the} ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
 For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
@@
 
-At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function.
-The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
+At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where {\green cosmological parameters are} encoded directly in the mapping function.
+The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both {\green the observed and underlying} populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
 In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
@@
 Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
 We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
@@
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.
+One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, {\green these misclassifications} can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.
 
@@
 That is, observed parameters are denoted with a hat, whilst the true (latent) value is denoted without a hat.
-For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.
+For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has measurement error, is different from $c$.
 
@@
 For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves and uncertainty.
+A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty.
 However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$.
@@
 
-With the report summary statistics being biased and their uncertainty under-reported, we face a significant challenge in supernova cosmology.
+With the {\green reported summary statistics $\hat{\eta}$ being biased} and their uncertainty under-reported, we face a significant challenge in supernova cosmology.
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.

### 2018-10-04 papers/methods/paper.tex
Initial rick comments

@@
 
-\subsubsection{Correcting biased summary statistics}
-
-With the {\green reported summary statistics $\hat{\eta}$ being biased} and their uncertainty under-reported, we face a significant challenge in supernova cosmology.
-We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.
-
-We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation.
-We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
-Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.
-
-As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift.
-As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature).
-The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
-We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
-We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.
-As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
-
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
-We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
-This extra dispersion is modelled with redshift independence.
-
 %Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
@@
 
+\subsection{Correcting biased summary statistics}
+
+With the {\green reported summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}.
+We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
+We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.
+
+We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation {\green $\hat{c}$}.
+We note that extra dispersion in magnitude {\green $\hat{m}_B$} (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
+Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.
+
+As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift.
+As the extra dispersion may arise from underlying supernova physics, rather than {\green any selection or observational effects}, we decide to incorporate redshift dependence in our extra uncertainty.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature).
+The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
+We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
+We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.
+As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
+
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
+This extra dispersion is modelled with redshift independence.
+
 \subsection{Mapping function}
@@
 In order to verify our model we run it through stringent tests.
-First, we validate on toy models, verifying that we recover accurate cosmology in highly constraining datasets.
+First, we validate on toy models, verifying that we recover accurate cosmology {\green when generating fake supernovae data constructed to satisfy the assumptions of the BHM construction}.
 We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample.
@@
 	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
-	\item Draw a true absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
 	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
-	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$.
-Use this to determine the true apparent magnitude of the object $m_B$ using equation \eqref{eq:l2}.
+	\item Draw an absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
+{\green Calculate the true absolute magnitude using Equation \eqref{eq:l2}.}
+	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology.
+Use this to determine the true apparent magnitude of the object $m_B$.
 	\item Determine if the SNIa is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
@@
 Repeat from step one until we have a supernova that passes.
+{\green We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.}
 	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
@@
 
-For supernova-independent uncertainty, the selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
+The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
 Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis.
@@
 
+{\red motivate this better}
+
 Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
@@
 Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys, where the LowZ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+Simulation specifics can be found in \citet{Kessler18}.
 
@@
 Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
-The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}.
+The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}.
 It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
@@
 	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
-We show the average parameter mean and standard deviation, with the simulation scatter shown in brackets.
+We show the average parameter mean and {\green average }standard deviation, with the simulation scatter shown in brackets.
 Truth values for the simulations are $\alpha=0.15$, $\langle M_B \rangle = -19.365$, $\beta_{\rm G10} = 3.1$, $\beta_{\rm C11} = 3.8$.
@@
 	\centering
-	\caption{Parameter Correlations with $w$ for the combined 100 simulation fits.
+	\caption{Reduced parameter correlations with $w$ for the combined 100 simulation fits.
 Correlations for the LowZ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.}
@@
 As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations.
-The extra smearing of the {\celeven} model does not result in a biased $\beta$ value, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
+The extra smearing of the {\celeven} {\green model does not result in a significantly biased $\beta$ value compared to the average uncertainty on $\beta$}, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
 Interestingly, $w$-bias is only found for the {\celeven} simulations.
@@
 BBC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017}.
-As a leading supernova cosmology method, it provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
-
-To this end, we take the results of the BBC method also run on our set of 200 validation simulations and compare the recovered $w$ values to those of our method.
+{\green The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected observables are fit using CosmoMC \citep{lewis2002cosmomc}.
+For shorthand, we refer to this combined process as the BBC method in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
+
+To this end, we take the results of the BBC method also run on {\green the} set of 200 validation simulations and compare the recovered $w$ values to those of our method.
 The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
@@
 	\caption{Characterising the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model.
+We also show the results when combining the {\gten} and {\celeven} models into a combined set of 200 simulations.
 The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation.

### 2018-10-05 papers/methods/paper.tex
Taking Dillon's comments on board

@@
 Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, sako2018sdss}.
+Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}.
 Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB.
@@
 
-In attempts to tease out the nature of dark energy, active and planned surveys are once again ramping up their statistical power.
+In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}.
 The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples.
@@
 
-As such, staggering effort is being put into developing more sophisticated supernova cosmology analyses.
+As such, {\green considerable resources are} being put into developing more sophisticated supernova cosmology analyses.
 The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models.
@@
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Hierarchical Bayesian models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
+Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
 
@@
 The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
-In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014}.
+In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Scolnic2017}.
 
@@
 \begin{align}
-\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right],\\
-d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
+\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
+d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}},
 \end{align}
@@
 A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size.
-Whilst binning data results in some loss of information, recent works tested against simulationsshow that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
+Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
 
@@
 Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM).
-Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and physical motivated conditional probabilities.
+Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and {\green physically} motivated conditional probabilities.
 For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
@@
 Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision.
-Their model was validated to be free of significant biases using fits to thirty realisations of supernova datasets.
-The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, mostly lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
+The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is determined perfectly.
+Their model was validated to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}.
+The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
 For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.
@@
 The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
-Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models and increasing simulation independence to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
+Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
+{\green We increase simulation independence by not utilising bias-corrections} to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
 
@@
 
-We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
+We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, {\green treatment of an increased number of systematics}, and analytic correction of selection effects, including systematic uncertainty on those corrections.
 We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
@@
 \end{align}
-As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including intrinsic dispersion in the supernova model.
+As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including {\green full treatment of} intrinsic dispersion in the supernova model.
 
@@
 Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift.
-Each set of nodes are anchored to equally spaced to span the minimum to maximum redshift of each survey.
+Each set of nodes are anchored to equally span the {\green redshift range }of each survey.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
@@
 Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
-We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
-The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.
+We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ {\green that} allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
+The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors.
+We sample $\delta_c$ in log space for efficiency in sampling close to zero.
 
@@
 We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-The probability of each supernovae falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section.
+The probability of each supernova falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section.
 For the spectroscopic survey, we set this to unity.
-For the photometric proof of concepts we provide an accurate probability estimate.
-Further investigation on the effect of inaccurate estimates we leave for future improvements during the analysis of the DES photometric sample.
+For the photometric proof-of-concept we provide an accurate probability estimate.
+Further investigation on the effect of inaccurate {\green estimates will be left for} future improvements during the analysis of the DES photometric sample.
 
@@
 
-With the {\green reported summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}.
+With the {\green fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}.
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
@@
 We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
-This extra dispersion is modelled with redshift independence.
+This magnitude dispersion is modelled with redshift independence.
 
@@
 \begin{align}
-M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times \hat{m}. \label{eq:l2}
+M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot \hat{m}. \label{eq:l2}
 \end{align}
@@
 We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
-This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
+{\green A comprehensive explanation of all systematics is given in \citet{Brout18SYS}.} This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
@@
 d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
-&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
+&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
 \end{align}
@@
 The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
-This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellowz}.}
+This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
 	\label{fig:efficiency}
@@
 The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
-Draw from each survey simulation until we have 1000 LowZ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis.
+Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis.
 Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
@@
 
-Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and LowZ surveys, where the LowZ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
 Simulation specifics can be found in \citet{Kessler18}.
 
-In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics.
+In addition to the improvements in the scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics.
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
@@
 The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}.
-It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based.
+It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based.
 We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.
@@
 	\centering
-	\caption{Tested population distributions, where the {\sk} LowZ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
+	\caption{Tested population distributions, where the {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
 	\label{tab:dist}
@@
 		\hline
-		{\sk} LowZ    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
+		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
 		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
@@
 		\hline
-		Model & $\alpha$ & $\beta$ & $\langle M_B \rangle$ & $\sigma_{m_B}^0$ &  $\sigma_{m_B}^1$ \\
+		Model & $\alpha$ & $\beta$ & $\langle M_B \rangle$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$ \\
 		\hline
@@
 In brackets, the simulation scatter (the standard deviation of the mean of 100 fits) is shown, and the bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.
-With systematics enabled, both the {\gten} and {\celeven} models show evidence of bias, scattering to either side of the simulated value of $w=-1$.
-However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
+{\green The {\gten} model is consistent with $w=-1$, whilst the {\celeven} model show evidence of bias on $w$, scattering high.} However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
 The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
@@
 	\caption{Reduced parameter correlations with $w$ for the combined 100 simulation fits.
-Correlations for the LowZ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.}
+Correlations for the low-$z$ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.
+Zero superscripts indicate the Dark Energy Survey, and a superscript one represents the low-$z$ survey.}
 	\label{tab:parameter_correlations}
@@
 Also visible in this figure are several other interesting relationships.
-$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and LowZ), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
+$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and low-$z$), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
 This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
@@
 
-Each realisation of simulated cosmology contains 137 LowZ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+Each realisation of simulated cosmology contains 137 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
 As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
@@
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+Our second test, to {\green test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through, we} test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
@@
 
-Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
+Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the low-$z$ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
 Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
@@
 
-For the sample size of the DES and LowZ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+For the sample size of the DES and low-$z$ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
 For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
@@
 Furthermore, our uncertainties in standardisation, population, mass-step and more, being explicitly parametrised in our model, are captured will covariance intact, an improvement on many previous methods.
-The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters and run in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion of other methods.
+The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters.
+It runs in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion of other methods.
 
@@
 \end{align}
-Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellowz}.
+Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.
 

### 2018-10-05 papers/methods/paper.tex
Initial Tam's comments

@@
 Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed Type Ia supernovae (SNIa) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}.
+Since that time, the number of observed type Ia supernovae (SN~Ia) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}.
 Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB.
@@
 In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of Type Ia supernova, attaining both spectroscopic and photometric identified samples.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of type Ia supernova, attaining both spectroscopic and photometric identified samples.
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
@@
 
-{\green Forefront among these challenges is our ignorance of the underlying Type Ia population -- its intrinsic dispersion}.
-Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
+{\green Forefront among these challenges is our ignorance of the underlying type Ia population -- its intrinsic dispersion}.
+Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
 We utilise two dispersion models in this work.
@@
 The underlying supernova population is further complicated by the presence of outliers.
-Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for Type Ia supernovae {\green and represent outliers from the standardised SN Ia sample.} This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
+Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
 Depending on the cuts {\green optimising the} ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
@@
 Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics.
-Because of the rarity of SNIa, datasets are commonly formed from the discoveries of multiple surveys.
+Because of the rarity of SN~Ia, datasets are commonly formed from the discoveries of multiple surveys.
 However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
@@
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of each supernovae being a Type Ia, $\hat{p}$.
+We will also have a probability of each supernovae being a type Ia, $\hat{p}$.
 Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
@@
 We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
-For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SNIa population in prior works.
+For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works.
 However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.
@@
 We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-The probability of each supernova falling into either population is determined by the observed Type Ia probability $\hat{p}$ discussed in the previous section.
+The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
 For the spectroscopic survey, we set this to unity.
@@
 It is now well known that host galaxy environment is significantly correlated with supernova properties.
-The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
+The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
 The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step.
@@
 With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
-For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SNIa} becomes
+For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
 
@@
 	\caption{The efficiency for supernova discovery at an arbitrary redshift.
-Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution.
+Shown in both panels in dashed blue is the SN~Ia population distribution, which takes the form of a normal distribution.
 The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude.
-The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
+The survey efficiency, given the SN~Ia population, is shown as a solid line in both panels, and the probability of observing a SN~Ia is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
 This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
@@
 Use this to determine the true apparent magnitude of the object $m_B$.
-	\item Determine if the SNIa is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
+	\item Determine if the SN~Ia is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
 For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$.
@@
 Stricter bias corrections calculated from simulations are required to reduce bias.
-Ideally, this would include further work on the calculation of intrinsic dispersion of the Type Ia supernovae population such that we can better characterise this bias.
+Ideally, this would include further work on the calculation of intrinsic dispersion of the type Ia supernovae population such that we can better characterise this bias.
 

### 2018-10-05 papers/methods/paper.tex
Tam's feedback

@@
 %
-These simulated datasets contain more than $250\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
+These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
 %
@@
 In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of type Ia supernova, attaining both spectroscopic and photometric identified samples.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
 The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
@@
 As such, {\green considerable resources are} being put into developing more sophisticated supernova cosmology analyses.
-The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific cosmological models.
-First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernovae cosmology.
-\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of many factors to better remove fitting biases.
+The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models.
+First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernova cosmology.
+\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of {\green multiple} factors to better remove fitting biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
@@
 
-Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly.
+Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands {\green and a redshift measurement of the supernova (or its assumed host)}, most analyses do not work from these measurements directly.
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
 This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}.
-The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
-As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
+The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
+As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
 
@@
 Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology.
-Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$ and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population living in apparent magnitude space.
+Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population in apparent magnitude space.
 Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:
@@
 
-where $M_B$ is the mean absolute magnitude for all SN Ia  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the colour correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
+where $M_B$ is the mean absolute magnitude for all SN Ia,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
 The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by rearranging equation \eqref{eq:standard} to isolate $\mu(z)$.
 The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
-In traditional analyses, the other corections term often includes Malmquist bias corrections, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Scolnic2017}.
+In traditional analyses, Malmquist bias corrections are also included, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Scolnic2017}.
 
@@
 }
-where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$ and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
+where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
 
@@
 Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
-For example, the underlying supernova colour population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
+For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
 Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
@@
 
-One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$).
+One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$, and $c$).
 \citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and are minimised in fitting.
@@
 
-The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, colour and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
+The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
 Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
@@
 The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
 These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population.
@@
 The underlying supernova population is further complicated by the presence of outliers.
-Non-Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
+Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN~Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
 Depending on the cuts {\green optimising the} ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
 For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
-However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit and correct for contaminants.
+However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit, and correct for contaminants.
 
@@
 \subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and colour $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of each supernovae being a type Ia, $\hat{p}$.
+We will also have a probability of each supernova being a type Ia, $\hat{p}$.
 Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
@@
 		$\alpha$ & Stretch standardisation   \\
-		$\beta$         &  Colour standardisation   \\
+		$\beta$         &  color standardisation   \\
 		$\delta(0)$ & Scale of the mass-magnitude correction\\
@@
 		$\langle M_B \rangle$ & Mean absolute magnitude \\
-		%$s_c$ & Colour bias\\
+		%$s_c$ & color bias\\
 		\hline
@@
 		$\langle x_1^i \rangle$ & Mean stretch nodes\\
-		$\langle c^i \rangle$ & Mean colour nodes\\
-		$\alpha_c$ & Skewness of colour population \\
+		$\langle c^i \rangle$ & Mean color nodes\\
+		$\alpha_c$ & Skewness of color population \\
 		$\sigma_{M_B}$ & Population magnitude scatter \\
 		$\sigma_{x_1}$ & Population stretch scatter \\
-		$\sigma_{c}$ & Population colour scatter \\
-		$\kappa_{0}$ & Extra colour dispersion\\
-		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
+		$\sigma_{c}$ & Population color scatter \\
+		$\kappa_{0}$ & Extra color dispersion\\
+		$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
 		\hline
@@
 		$x_1$ & True stretch \\
-		$c$ & True colour \\
+		$c$ & True color \\
 		$z$ & True redshift \\
@@
 		$\hat{x}_1$ & Measured stretch \\
-		$\hat{c}$ & Measured colour \\
-		$C$ & Covariance on flux, stretch and colour \\
+		$\hat{c}$ & Measured color \\
+		$C$ & Covariance on flux, stretch and color \\
 		$\hat{z}$ & Observed redshift \\
@@
 
-The first layer of the hierarchy represents the parameters that describe each supernova.
-That is, observed parameters are denoted with a hat, whilst the true (latent) value is denoted without a hat.
-For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has measurement error, is different from $c$.
-
-For the moment, let us consider a single supernova and it's classic summary statistics $m_B$, $x_1$, $c$.
+{\green The first layer of hierarchy are the true (latent) parameters that describe each supernova.
+In contrast to the observed parameters, the latent parameters are denoted without a hat.} For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the color we observe, which, as it has measurement error, is different from $c$.
+
+For the moment, let us consider a single supernova and its classic summary statistics $m_B$, $x_1$, $c$.
 For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
 A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty.
-However, this is computational prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$.
+However, this is computationally prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$.
 When using this simplification, our latent variables are given by
@@
 \end{align}
-As discussed in Section \ref{sec:challenges}, this approximation is known to fail due to the SALT2 model not including {\green full treatment of} intrinsic dispersion in the supernova model.
+As discussed in Section \ref{sec:challenges},{\green the SALT2 model does not include full treatment of intrinsic dispersion in the supernova model, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
 
@@
 
-%Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift.
-By comparing simulated true colour values to simulated observed colour values, we can quantify the mean bias in colour in redshift bins, giving us {\green a redshift-dependent array of biases.
+%Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean color measured as a function of redshift.
+By comparing simulated true color values to simulated observed color values, we can quantify the mean bias in color in redshift bins, giving us {\green a redshift-dependent array of biases.
 We can then add this into our data-set, and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
-Specifically, allow the observed colour to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$, where `bias$(z)$' is the binned mean colour bias determined with fiducial simulations.
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae colour across all surveys.
-We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing.
+Specifically, allow the observed color to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$, where `bias$(z)$' is the binned mean color bias determined with fiducial simulations.
+Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae color across all surveys.
+We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean color address the non-Gaussianity introduced by chromatic smearing.
 We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction: 0 for no shift, up to 1 for the maximum shift found in simulations.
@@
 \label{sec:underlying}
-The underlying supernova population is characterised by underlying distributions in colour, stretch and absolute magnitude.
-We follow the prior work of {\rubin} and model the colour population as an independent redshift-dependent skew normal for each survey.
+The underlying supernova population is characterised by underlying distributions in color, stretch and absolute magnitude.
+We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal for each survey.
 For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal.
-Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and colour for a given redshift.
-Each set of nodes are anchored to equally span the {\green redshift range }of each survey.
+Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
-Both the colour and stretch means are modelled with normal priors.
+Both the color and stretch means are modelled with normal priors.
 Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
@@
 
-As such, the only constant between survey populations is the mean absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
+The only constant between survey populations is the mean absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
@@
 For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works.
-However, with with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.
+However, with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.
 We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+{\green We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude.} We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
 The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
@@
 
-We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation {\green $\hat{c}$}.
-We note that extra dispersion in magnitude {\green $\hat{m}_B$} (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion.
-Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.
-
-As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift.
+We model the extra dispersion only in color, and do so by adding extra independent uncertainty on the color observation {\green $\hat{c}$}.
+We note that extra dispersion in magnitude {\green $\hat{m}_B$} (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the color term, hence the requirement for modelling additional color dispersion.
+Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than color, and causes negligible bias in cosmology.
+
+As shown in \citep{Kessler2013}, the extra color dispersion shows heavy redshift dependence, increasing with redshift.
 As the extra dispersion may arise from underlying supernova physics, rather than {\green any selection or observational effects}, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature).
-The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$.
+We thus add $\kappa_0 + \kappa_1 z$ to our observed color uncertainty (in quadrature).
+The $\kappa$ parameters are highly degenerate with the width of the intrinsic color population $\sigma_c$.
 We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
@@
 
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient.
+Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just color dispersion, and was far more computationally inefficient.
 We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
@@
 
-We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM and standard $\Lambda$CDM.
+We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM.
 $\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$.
@@
 
-\subsubsection{Standardisation Parameters}
+\subsubsection{Supernova Standardisation Parameters}
 
 With increasingly large datasets and more nuanced analyses, the choice of how to handle $\alpha$ and $\beta$ becomes an important consideration when constructing a model.
-{\rubin} employs a broken linear relationship for both colour and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are respectively positive or negative (although the cut could be placed at a location other than 0).
+{\rubin} employs a broken linear relationship for both color and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are positive or negative (although the cut could be placed at a location other than 0).
 \citet{Shariff2016} instead model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and a second model which effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
@@
 The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
-The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step.
-The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
+The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step.
+The scale of this step function varies from analysis to analysis, with the 0.08 valuesourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
 In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
@@
 \begin{align}
-M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \cdot \hat{m}. \label{eq:l2}
+M_B = m_B + \alpha x_1 - \beta c + \Delta M \cdot \hat{m}. \label{eq:l2}
 \end{align}
@@
 where the denominator represents an integral over all potential data $D$.
-As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
+This is derived in Appendix \ref{app:selection1} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
 Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
@@
 
-where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$ and $c$ population to $m_B$ given cosmological parameters.
-A full derivation of this can be found in Appendix \ref{app:selection}.
-
-We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernova reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey colour and stretch populations can be treated as Gaussian for the purposes of this evaluating $P(m_B | z, \theta)$.
-It was found that discarding the colour population skewness entirely resulted in highly biased population recovery (see Figure \ref{fig:simple_w_super} to see the populations), and so we instead characterise the skew normal colour distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
-This shifted Gaussian approximation for colour completely removes the unintended bias when simply discarding skewness.
-This shift was not required for the stretch population, and so was left out for numerical reasons.
+where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$, and $c$ population to $m_B$ given cosmological parameters.
+A full derivation of this can be found in Appendix \ref{app:selection2}.
+
+We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernovae reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey color and stretch populations can be treated as Gaussian for the purposes of evaluating $P(m_B | z, \theta)$.
+We found that discarding the color population skewness entirely resulted in highly biased population recovery (see Figure \ref{fig:simple_w_super} to see the populations), and so we instead characterise the skew normal color distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
+This shifted Gaussian approximation for color completely removes the unintended bias when simply discarding skewness.
+This shift was not required for the stretch population, and so was left out {\green for the stretch population} for numerical reasons.
 The impact of this approximation on the calculated efficiency is shown in Figure \ref{fig:shift}, and more detail on this shift and resulting population recovery can be found in Appendix \ref{app:approx}.
@@
 	\end{center}
-	\caption{Testing the correctness of our normal approximation to the skewed colour distribution.
+	\caption{Testing the correctness of our normal approximation to the skewed color distribution.
 The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$, calculated numerically.
@@
 What then remains is determining the functional form of $P(S|m_B)$.
-For the treatment of most surveys, we find that the error function which smoothly transitions from some constant efficiency down to zero is sufficient.
+For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient.
 Formally, this gives
@@
 
-where $\Phi^c$ the complimentary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
+where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
 The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
@@
 Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
-Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level - varying cosmology or spectroscopic efficiency.
+Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level --- varying cosmology or spectroscopic efficiency.
 Tests with changing the intrinsic scatter model used in the selection efficiency simulations show that the uncertainty introduced is negligible.
 
-With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
+With the {\green well-sampled redshift} approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
 For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
@@
 Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.
-The colour of the data points is representative over the supernovae colour itself, a negative colour value showing bluer supernovae, with positive colour values representing redder supernovae.}
+The color of the data points is representative of the supernova color itself, a negative color value showing bluer supernovae, with positive color values representing redder supernovae.}
 	\label{fig:simple_pop}
@@
 	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
-	\item Draw an absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
+	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
 {\green Calculate the true absolute magnitude using Equation \eqref{eq:l2}.}
@@
 	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
-Add extra colour uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
+Add extra color uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
 \end{enumerate}
@@
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
-Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
+Initial tests were also done with a second, Gaussian population with color and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
 The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}.
@@
 Truth values for the simulations are $\alpha=0.15$, $\langle M_B \rangle = -19.365$, $\beta_{\rm G10} = 3.1$, $\beta_{\rm C11} = 3.8$.
-The width of the intrinsic scatter does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.}
+The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.}
 	\label{tab:standardisation}
@@
 Also visible in this figure are several other interesting relationships.
-$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES like and low-$z$), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
-This relationship is indeed expected - as $\kappa^0_c$ grows larger (more unexplained dispersion on the colour observation), the width of the supernova population in apparent magnitude space increases.
-As the fit prefers it to conform to the observed width of the distribution, the extra width in colour causes the inherent magnitude smearing amount to decrease.
-And with extra freedom on the observed colour from $\kappa^0_c$, $\beta$ shifts in response.
-The other striking feature in the plot are the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
-These too are expected, for they show the relationship between the colour distributions mean value, its width and its skewness.
-As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean colour population.
-Strong anti-correlation between $\kappa_{c0}^0$ and $\kappa_{c0}^1$ with $\sigma_{m_B}$ reveals the strong population degeneracy, and -- for the C11 simulation results -- a constrained positive value shows that a finite non-zero extra colour dispersion is indeed preferred by our model.
+$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES-like and low-$z$), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
+This relationship is indeed expected --- as $\kappa^0_c$ grows larger (more unexplained dispersion on the color observation), the width of the supernova population in apparent magnitude space increases.
+As the fit prefers it to conform to the observed width of the distribution, the extra width in color causes the inherent magnitude smearing amount to decrease.
+And with extra freedom on the observed color from $\kappa^0_c$, $\beta$ shifts in response.
+The other striking feature in the plot is the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
+These too are expected, for they show the relationship between the color distribution's mean value, its width and its skewness.
+As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean color population.
+Strong anti-correlation between $\kappa_{c0}^0$ and $\kappa_{c0}^1$ with $\sigma_{m_B}$ reveals the strong population degeneracy, and -- for the C11 simulation results -- a constrained positive value shows that a finite non-zero extra color dispersion is indeed preferred by our model.
 }
@@
 
-Each realisation of simulated cosmology contains 137 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+Each realisation of simulated cosmology contains 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
 As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
@@
 
-We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics.
+We investigate the cosmological bias and find its source to be a bias in the observed summary statistics {\green(i.e.
+the $m_B$, $x_1$, $c$ output from SALT2 light curve fitting)}, in addition to incorrect reported uncertainty on the summary statistics.
 To confirm this, we run two tests.
@@
 This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
-
-% Colour bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
+Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies --- incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+% color bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
 
@@
 
-For the sample size of the DES and low-$z$ supernova samples (of order 350 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and as such we will leave more complicated treatment of them for future work.
+For the sample size of the DES $+$ low-$z$ supernova samples (332 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
+For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and given that they remain sub-dominant, we will leave more complicated treatment of them for future work.
 
@@
 {\green The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected observables are fit using CosmoMC \citep{lewis2002cosmomc}.
-For shorthand, we refer to this combined process as the BBC method in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency as to the current levels of accuracy in recovering cosmological parameters.
-
-To this end, we take the results of the BBC method also run on {\green the} set of 200 validation simulations and compare the recovered $w$ values to those of our method.
+For shorthand, we refer to this combined process as the BBC method in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency{\green check} as to the current levels of accuracy in recovering cosmological parameters.
+
+{\green To this end, we take the results of the BBC method which were also run on the} set of 200 validation simulations and compare the recovered $w$ values to those of our method.
 The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
@@
 As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
-These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
-The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
+{\green These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}.
+Both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
+The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$}.
 Our method shows a small improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}.
@@
 
-Having established that our method exhibits similar shifts in their recovery of $w$ compared to BBC, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
+Having established that our method exhibits similar shifts in the recovery of $w$ compared to BBC, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
 
@@
 In this paper we have outlined the creation of a hierarchical Bayesian model for supernova cosmology.
-The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter colour smearing and incorporates uncertainty from peculiar velocities, survey calibration, HST calibration, dust, a potential global redshift offset, and SALT2 model uncertainty.
-Furthermore, our uncertainties in standardisation, population, mass-step and more, being explicitly parametrised in our model, are captured will covariance intact, an improvement on many previous methods.
+The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter color smearing and incorporates uncertainty from peculiar velocities, survey calibration, HST calibration, dust, a potential global redshift offset, and SALT2 model uncertainty.
+Furthermore, our uncertainties in standardisation, population, mass-step and more, being explicitly parametrised in our model, are captured with covariance intact, an improvement on many previous methods.
 The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters.
@@
 Stricter bias corrections calculated from simulations are required to reduce bias.
-Ideally, this would include further work on the calculation of intrinsic dispersion of the type Ia supernovae population such that we can better characterise this bias.
-
-With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $250\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin.
+Ideally, this would include further work on the calculation of intrinsic dispersion of the type Ia supernova population such that we can better characterise this bias.
+
+With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $60\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin.
 With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-year dataset.
@@
 \subsection{General Selection Effects}
-
+\label{app:selection1}
 When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
@@
 \subsection{Supernova Selection Effects}
-
+\label{app:selection2}
 We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
@@
 \end{align}
-where for simplicity we have not written out all the integrals which do not interact with the selection effects explicitly.
+where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly.
 Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
@@
 \end{align}
-By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernovae redshifts, we can approximate the integral over redshift by evaluating
+By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
 \begin{align}
@@
 
-In this section, we investigate the effect of approximating the skew normal underlying colour distribution as a normal.
-Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our colour distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
+In this section, we investigate the effect of approximating the skew normal underlying color distribution as a normal.
+Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our color distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
 We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to more accurately describe the actual mean and standard deviation of a skew normal.
@@
 We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$.
-Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$ and $c$ as a multivariate skew normal did not produce stable posterior surfaces.
-Due to this, we treat the underlying $m_B$, $x_1$ and $c$ populations as independent.
+Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$, and $c$ as a multivariate skew normal did not produce stable posterior surfaces.
+Due to this, we treat the underlying $m_B$, $x_1$, and $c$ populations as independent.
 
@@
 	\end{center}
-	\caption{Marginalised probability distributions for 100 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-z and 1000 simulated low-z supernovae.
-The dashed green surfaces represent a fit to an underlying Gaussian colour population with the unshifted model.
-The blue solid surface represents fits to a skewed colour population with the unshifted model, and the purple dotted surface represents a fit to a skewed colour population with the shifted model.
-The superscript $0$ and $1$ denote the two different surveys (high-z and low-z respectively), and similarly the first four $\langle c^i \rangle$ parameters represent the four redshift nodes in the high-z survey, and the last four represent the nodes for the low-z survey.
-We can see that the shifted model is far better able to recover skewed input populations than the unshifted, performing better in terms of recovering skewness $\alpha_c$, mean colour $\langle c \rangle$ and width of the colour distribution $\sigma_c$.
-The unshifted model recovers the correct colour mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the colour population mean.
+	\caption{Marginalised probability distributions for 100 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-$z$ and 1000 simulated low-$z$ supernovae.
+The dashed green surfaces represent a fit to an underlying Gaussian color population with the unshifted model.
+The blue solid surface represents fits to a skewed color population with the unshifted model, and the purple dotted surface represents a fit to a skewed color population with the shifted model.
+The superscript $0$ and $1$ denote the two different surveys (high-$z$ and low-$z$ respectively), and similarly the first four $\langle c^i \rangle$ parameters represent the four redshift nodes in the high-$z$ survey, and the last four represent the nodes for the low-$z$ survey.
+We can see that the shifted model is far better able to recover skewed input populations than the unshifted, performing better in terms of recovering skewness $\alpha_c$, mean color $\langle c \rangle$ and width of the color distribution $\sigma_c$.
+The unshifted model recovers the correct color mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the color population mean.
 Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.
-Based on these results and the fantastic performance in correctly recovering underlying populations of the shifted normal approximation, we adopt that in our model.}
+Based on these results and the fantastic performance in correctly recovering underlying populations of the shifted normal approximation, we adopt the shifted normal approximation in our model.}
 	\label{fig:simple_w_super}
@@
 
-This parametrisation does come with one significant downside - inflexibility.
-For each step the algorithm takes, we do not recompute the cholesky decomposition of the covariance of the summary statistics - that happens once at the beginning of the model setup.
-If we had kept the full covariance matrix parametrisation we could modify the matrix easily - for example when incorporating intrinsic dispersion we could simple add on a secondary matrix to create an updated covariance.
+This parametrisation does come with one significant downside --- inflexibility.
+For each step the algorithm takes, we do not recompute the cholesky decomposition of the covariance of the summary statistics --- that happens once at the beginning of the model setup.
+If we had kept the full covariance matrix parametrisation we could modify the matrix easily --- for example when incorporating intrinsic dispersion we could simply add on a secondary matrix to create an updated covariance.
 However as the cholesky decomposition of a sum of matrices is not equal to the sum of the cholesky decomposition of each individual matrix, we would need to recompute the decomposition for each step, which discards most of the computational benefit just gained.
@@
 \end{align}
-Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered colour parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$.
-We can now see that it is possible to add extra uncertainty to the colour observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix.
-To take our covariance and add on the diagonal uncertainty for colour an extra $\sigma_e$ term, we get
+Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered color parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$.
+We can now see that it is possible to add extra uncertainty to the color observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix.
+To take our covariance and add on the diagonal uncertainty for color an extra $\sigma_e$ term, we get
 \begin{align}
@@
 where $g = \sqrt{f^2 + \sigma_e^2} - f$.
-This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent colour uncertainty.
+This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent color uncertainty.
 For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.

### 2018-10-09 papers/methods/paper.tex
Ricks additional feedback

@@
 %
-Given light curve fit parameters, redshifts, and host-galaxy masses, {\steve} simultaneously fits for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
+Given light curve fit parameters, redshifts, and host-galaxy masses, {\green we fit {\steve}} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
 Selection effects are characterised using Monte-Carlo simulations.
-
+%
 We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
@@
 This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model.
-This $w$-bias is less than $0.03$ on average, roughly half the uncertainty on $w$.
+This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
 %
-These simulation test results are a concern for upcoming large surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
+These simulation test results are a concern {\green for BHM applications on large upcoming surveys}, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
 \end{abstract}
@@
 First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernova cosmology.
-\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of {\green multiple} factors to better remove fitting biases.
+\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of {\green multiple} factors to better remove biases.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
@@
 
-In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects, whilst improving computational performance.
-Section \ref{sec:review} is dedicated to a quick review of the {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
+{\green In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}.
+We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty.
+We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.} Section \ref{sec:review} is dedicated to a quick review of {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
@@
 This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}.
-The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$.
+The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C}$).
 As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
@@
 Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology.
-Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population in apparent magnitude space.
+Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditional in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population in apparent magnitude space.
 Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:
@@
 \begin{align}
-M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{other corrections}, \label{eq:standard}
+\mu(z) = m_B + \alpha x_1 - \beta c - M_B + \text{other corrections}, \label{eq:standard}
 \end{align}
@@
 where $M_B$ is the mean absolute magnitude for all SN Ia,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
-The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by rearranging equation \eqref{eq:standard} to isolate $\mu(z)$.
+The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by {\green utilising equation \eqref{eq:standard} and supernova observations}.
 The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
@@
 
-Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from rearranging equation \ref{eq:standard}, and the cosmologically predicted values.
+Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted values.
 We denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$.
@@
 
-where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Betoule2014} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
+where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
 {\green 
@@
 For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
-Thirdly, the computational efficiency is dependent on inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
+{\green Whilst known correlations like this can be captured and incorporated into a covariance matrix if known, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.} Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size.
+A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size {\green and can allow fitting without matrix inversion on every step}.
 Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
@@
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data.
-When using traditional fitting methods, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
+When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
+{\green Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections.}
 
@@
 
-With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
+With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
 \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
@@
 Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is determined perfectly.
+The simplification made by the UNITY analysis is that the bias is {\green perfectly described} by an analytic function.
 Their model was validated to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}.
@@
 Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
-{\green We increase simulation independence by not utilising bias-corrections} to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
+{\green We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections} to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
 
@@
 Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
-We utilise two dispersion models in this work.
+We{\green test against }two dispersion models in this work.
 The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
@@
 
-At the most fundamental level, a supernova cosmology analysis is simply a mapping from an underlying population onto an observed population, where {\green cosmological parameters are} encoded directly in the mapping function.
+At the most fundamental level, a supernova cosmology analysis is a mapping from an underlying population onto an observed population, where {\green cosmological parameters are} encoded directly in the mapping function.
 The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both {\green the observed and underlying} populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
@@
 We will also have a probability of each supernova being a type Ia, $\hat{p}$.
-Our set of observables is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+Our set of observables {\green input into the BHM model} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal.
-Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
+{\green We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no cosmological benefit with the skew-normal distributions for stretch and magnitude.} Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
@@
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We must also do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model.
+We {\green should also} do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model, {\green or utilise the results of computing the bias from both models.}
 
@@
 Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just color dispersion, and was far more computationally inefficient.
-We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
+We note here that we model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
 This magnitude dispersion is modelled with redshift independence.
@@
 \label{sec:systreat}
-The chief difficulty with including systematic uncertainties in supernova analyses is that they generally occur during the observational pipeline, and have difficult-to-model effects on the output observations.
-As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{\partial \hat{m}_B}{\partial \Z_i}$, $\frac{\partial \hat{x}_1}{\partial \Z_i}$, $\frac{\partial \hat{c}}{\partial \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
+The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations.
+As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
 
@@
 We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
-{\green A comprehensive explanation of all systematics is given in \citet{Brout18SYS}.} This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
+{\green A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.} This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
@@
 One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
-We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
+We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are{\green observed and pass cuts}.
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
@@
 
-The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude.
+The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by {\green taking the ratio of supernovae that are observed and passed cuts over the total number of supernovae generated in that apparent magnitude bin}.
 That is, we calculate the probability we would include a particular supernova in our sample, divided by the number of such supernovae in our simulated fields.
@@
 
-We can thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty.
+We thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty.
 Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
@@
 In order to verify our model we run it through stringent tests.
-First, we validate on toy models, verifying that we recover accurate cosmology {\green when generating fake supernovae data constructed to satisfy the assumptions of the BHM construction}.
+First, we validate on toy models, verifying that we recover accurate cosmology {\green when generating toy supernovae data constructed to satisfy the assumptions of the BHM construction}.
 We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample.
@@
 As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
-For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviations in the table.
+For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviation in the table.
 The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets.
@@
 		\hline
-		Model & $\Omega_m$(scatter) & $w$(scatter) \\ 
-		\hline
-		Flat $\Lambda$CDM & $0.301\pm 0.015(0.012)$ & -- \\ 
-		Flat $w$CDM & -- & $-1.00\pm 0.042(0.030)$ \\ 
+		Model & $\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) & $w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) \\ 
+		\hline
+		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
+		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
 		\hline
@@
 
-{\red motivate this better}
-
-Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
-In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
-
-Simulations (using the SNANA package) follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+{\green Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model.
+This is a useful consistency check that the model implementation is correct, efficient and free of obvious pathologies.
+However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways.
+To this end, we test using simulations} (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
 Simulation specifics can be found in \citet{Kessler18}.
 
-In addition to the improvements in the scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics.
+Additionally, early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
+In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
+
+In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics.
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
@@
 	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
-We show the average parameter mean and {\green average }standard deviation, with the simulation scatter shown in brackets.
-Truth values for the simulations are $\alpha=0.15$, $\langle M_B \rangle = -19.365$, $\beta_{\rm G10} = 3.1$, $\beta_{\rm C11} = 3.8$.
+We show the average parameter mean and {\green average }standard deviation respectively, with the simulation scatter shown in brackets.
 The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.}
 	\label{tab:standardisation}
-	\begin{tabular}{l|ccccc}
-		\hline
-		Model & $\alpha$ & $\beta$ & $\langle M_B \rangle$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$ \\
-		\hline
-		{\gten} Stat + Syst  & $0.152\pm0.009\ (0.008)$ & $ 3.44\pm0.19\ (0.18)$ & $-19.367\pm0.028\ (0.015)$ & $0.070\pm0.022\ (0.018)$ & $0.073\pm0.025\ (0.022)$\\
-		{\gten}  Stat         & $0.150\pm0.008\ (0.008)$ & $3.43\pm0.20\ (0.17)$ & $-19.374\pm0.016\ (0.013)$ & $0.069\pm0.023\ (0.019)$ & $0.072\pm0.026\ (0.023)$\\
-		{\celeven} Stat + Syst  &  $0.152\pm0.009\ (0.007)$ & $3.76\pm0.15\ (0.13)$ & $-19.351\pm0.030\ (0.018)$ & $0.024\pm0.016\ (0.011)$ & $0.029\pm0.020\ (0.014)$ \\
-		{\celeven} Stat         &  $0.150\pm0.008\ (0.007)$ & $3.75\pm0.16\ (0.13)$ & $-19.359\pm0.016\ (0.015)$ & $0.025\pm0.016\ (0.012)$ & $0.027\pm0.020\ (0.015)$\\
+	\begin{tabular}{l|rrrrr}
+		\hline
+		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$ \\
+		\hline
+		{\gten} Stat + Syst  & $0.022, \ 0.009\ (0.008)$ & $ 0.34, \ 0.19\ (0.18)$ & $-0.002, \ 0.028\ (0.015)$ & $0.070, \ 0.022\ (0.018)$ & $0.073, \ 0.025\ (0.022)$\\
+		{\gten}  Stat         & $0.000, \ 0.008\ (0.008)$ & $0.33, \ 0.20\ (0.17)$ & $0.001, \ 0.016\ (0.013)$ & $0.069, \ 0.023\ (0.019)$ & $0.072, \ 0.026\ (0.023)$\\
+		{\celeven} Stat + Syst  &  $0.002, \ 0.009\ (0.007)$ & $-0.04, \ 0.15\ (0.13)$ & $-19.351, \ 0.030\ (0.018)$ & $0.024, \ 0.016\ (0.011)$ & $0.029, \ 0.020\ (0.014)$ \\
+		{\celeven} Stat         &  $0.000, \ 0.008\ (0.007)$ & $-0.05, \ 0.16\ (0.13)$ & $-19.359, \ 0.016\ (0.015)$ & $0.025, \ 0.016\ (0.012)$ & $0.027, \ 0.020\ (0.015)$\\
 		\hline
@@
 	\label{tab:bulk_summary}
-	\begin{tabular}{l|cc}
-		\hline
-		Model & $w$ (scatter) & $w$-Bias \\
-		\hline
-		{\gten} Stat + Syst  &  $-0.998\pm 0.097\ (0.073)$ & $(0.02\pm0.07)\sigma$  \\
-		{\gten}  Stat         &  $-1.008\pm 0.080\ (0.068)$ & $(-0.10\pm0.08)\sigma$   \\
-		{\celeven} Stat + Syst  &  $-0.945\pm 0.098\ (0.077)$ & $(0.55\pm0.08)\sigma$  \\
-		{\celeven} Stat         &  $-0.948\pm 0.079\ (0.066)$ & $(0.65\pm0.08)\sigma$   \\
+	\begin{tabular}{l|rr}
+		\hline
+		Model & $w\ \langle\mu\rangle,\ \langle\sigma_w\rangle$ (scatter) & $w$-Bias \\
+		\hline
+		{\gten} Stat + Syst  &  $-0.998, \ 0.097\ (0.073)$ & $(0.02\pm0.07)\sigma$  \\
+		{\gten}  Stat         &  $-1.008,\ 0.080\ (0.068)$ & $(-0.10\pm0.08)\sigma$   \\
+		{\celeven} Stat + Syst  &  $-0.945,\ 0.098\ (0.077)$ & $(0.55\pm0.08)\sigma$  \\
+		{\celeven} Stat         &  $-0.948,\ 0.079\ (0.066)$ & $(0.65\pm0.08)\sigma$   \\
 		\hline
@@
 The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation.
-Finally, we also characterise the scatter between the methods in the final row.
 Averages are computed giving each simulation sample the same weight.
@@
 		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
-		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$ \\
+		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$\footnotemark\\
 		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
@@
 \end{table}
-
+\footnotetext{This value is computed with each simulation having the same weight.
+It disagrees with the value provided in \citet[Table 10, row 3]{Brout18SYS} which uses inverse variance weighted averages.
+We do not utilise this weight because the variance is correlated with the value of $w$ due to the $\Omega_m$ prior applied in the fitting process.}
 \begin{figure}

### 2018-10-09 dessn/framework/simulations/selection_effects.py

@@ -40,7 +40,7 @@ def lowz_sel(cov_scale=1.0, shift=None, type="G10", kappa=0, zlim=None, version=
     return sn, mean, cov, kappa
 
 
-def get_data(base, zlim=None):
+def get_data(base, zlim=None, maxc=None, minc=None):
     file = os.path.abspath(inspect.stack()[0][1])
     dir_name = os.path.dirname(file)
     folder = dir_name + "/" + base
@@ -57,8 +57,8 @@ def get_data(base, zlim=None):
     return mags, passed
 
 
-def get_ratio(base_folder, cut_mag=19.75, delta=0, zlim=None):
-    mB_all, passed = get_data(base_folder, zlim=zlim)
+def get_ratio(base_folder, cut_mag=19.75, delta=0, zlim=None, maxc=None, minc=None):
+    mB_all, passed = get_data(base_folder, zlim=zlim, maxc=maxc, minc=minc)
     mB_passed = mB_all[passed]
 
     # Bin data and get ratio
@@ -135,8 +135,7 @@ def get_selection_effects_cdf(dump_npy, plot=False, cut_mag=19, kappa=0, zlim=No
         name = os.path.basename(dump_npy)
 
         ax.text(0.98, 0.95, "DES 3YR Spectroscopically Confirmed", verticalalignment='top', horizontalalignment='right', transform=ax.transAxes)
-        # fig.savefig("../../../papers/methods/figures/%s_%0.2f.png" % (name, delta), bbox_inches="tight", transparent=True)
-        # fig.savefig("../../../papers/methods/figures/%s.pdf" % name, bbox_inches="tight", transparent=True)
+        fig.savefig("../../../papers/methods/figures/%s.pdf" % name, bbox_inches="tight", transparent=True)
         plt.show()
 
     print(vals, cov, r2)
@@ -198,9 +197,9 @@ def get_selection_effects_skewnorm(dump_npy, plot=False, cut_mag=10, kappa=0, zl
         fig.tight_layout()
         name = os.path.basename(dump_npy)
         plt.show()
-        ax.text(0.98, 0.95, "Combined LowZ Sample", verticalalignment='top', horizontalalignment='right', transform=ax.transAxes)
-        # fig.savefig("../../../papers/methods/figures/%s.png" % name, bbox_inches="tight", transparent=True)
[…]

### 2018-10-10 papers/methods/paper.tex
Dan and Anais's feedback

@@
 In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has observed thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
-The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will observe scores of thousands of photometrically classified supernovae.
-{\green Such increased statistical power demands greater fidelity and flexibility in modelling the supernovae for cosmological purposes, as we will required reduced systematic uncertainties to fully utilised increase statistics} \citep{Betoule2014, Scolnic2017}.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has {\green discovered} thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
+The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will {\green discover} scores of thousands of photometrically classified supernovae.
+{\green Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics} \citep{Betoule2014, Scolnic2017}.
 
@@
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find sufficient analytic approximations for complicated effects such as Malmquist bias.
+Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find analytic approximations for complicated effects such as Malmquist bias.
 
@@
 
-Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands {\green and a redshift measurement of the supernova (or its assumed host)}, most analyses do not work from these measurements directly.
+Whilst supernova observations take the form of photometric time-series brightness measurements in many bands {\green and a redshift measurement of the supernova (or its assumed host)}, most analyses do not work from these measurements directly.
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
@@
 Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology.
-Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditional in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population in absolute magnitude space into the observed population in apparent magnitude space.
+Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditionally in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space.
 Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:
@@
 Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM).
-Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and {\green physically} motivated conditional probabilities.
+Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and {\green physically} motivated conditional probabilities.
 For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
@@
 \end{align}
-As discussed in Section \ref{sec:challenges},{\green the SALT2 model does not include full treatment of intrinsic dispersion in the supernova model, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
+As discussed in Section \ref{sec:challenges}, {\green the SALT2 model does not include full treatment of intrinsic dispersion in the supernova model, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
 
@@
 	\end{center}
-	\caption{Fitting the selection function for both the DES 3YR spectroscopically confirmed supernova sample and the combined low-redshift sample.
+	\caption{Fitting the selection function for both the DES 3YR spectroscopically confirmed supernova sample and the low-$z$ sample.
 Blue errorbars represent the efficiency calculated by determining the ratio of discovered to generated supernovae in apparent magnitude bins for SNANA simulations.
@@
 	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
-We show the average parameter mean and {\green average }standard deviation respectively, with the simulation scatter shown in brackets.
-The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.}
+We show the average parameter mean and {\green average }standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$.
+The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.
+A measure of the significance of the parameter bias can be calculated by comparing the bias to a tenth of the scatter (as our Monte-Carlo estimate uncertainty is $\sqrt{100}$ of the scatter).
+From this, we can see that most biases are detected with high statistical significance due to the large number of simulations tested against.}
 	\label{tab:standardisation}
@@
 		\hline
-		{\gten} Stat + Syst  & $0.022, \ 0.009\ (0.008)$ & $ 0.34, \ 0.19\ (0.18)$ & $-0.002, \ 0.028\ (0.015)$ & $0.070, \ 0.022\ (0.018)$ & $0.073, \ 0.025\ (0.022)$\\
-		{\gten}  Stat         & $0.000, \ 0.008\ (0.008)$ & $0.33, \ 0.20\ (0.17)$ & $0.001, \ 0.016\ (0.013)$ & $0.069, \ 0.023\ (0.019)$ & $0.072, \ 0.026\ (0.023)$\\
-		{\celeven} Stat + Syst  &  $0.002, \ 0.009\ (0.007)$ & $-0.04, \ 0.15\ (0.13)$ & $-19.351, \ 0.030\ (0.018)$ & $0.024, \ 0.016\ (0.011)$ & $0.029, \ 0.020\ (0.014)$ \\
-		{\celeven} Stat         &  $0.000, \ 0.008\ (0.007)$ & $-0.05, \ 0.16\ (0.13)$ & $-19.359, \ 0.016\ (0.015)$ & $0.025, \ 0.016\ (0.012)$ & $0.027, \ 0.020\ (0.015)$\\
+		{\gten} Stat + Syst  & $0.022 \ [0.009\ (0.008)]$ & $ 0.34 \ [0.19\ (0.18)]$ & $-0.002 \ [0.028\ (0.015)]$ & $0.070 \ [0.022\ (0.018)]$ & $0.073 \ [0.025\ (0.022)]$\\
+		{\gten}  Stat         & $0.000 \ [0.008\ (0.008)]$ & $0.33 \ [0.20\ (0.17)]$ & $0.001 \ [0.016\ (0.013)]$ & $0.069 \ [0.023\ (0.019)]$ & $0.072 \ [0.026\ (0.023)]$\\
+		{\celeven} Stat + Syst  &  $0.002 \ [0.009\ (0.007)]$ & $-0.04 \ [0.15\ (0.13)]$ & $0.014 \ [0.030\ (0.018)]$ & $0.024 \ [0.016\ (0.011)]$ & $0.029 \ [0.020\ (0.014)]$ \\
+		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$\\
 		\hline
@@
 	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
-The quoted value for $w$ represents the average mean of the fits, with the uncertainty being the average standard deviation of the fits.
-In brackets, the simulation scatter (the standard deviation of the mean of 100 fits) is shown, and the bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.
+The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets.
+The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.
 {\green The {\gten} model is consistent with $w=-1$, whilst the {\celeven} model show evidence of bias on $w$, scattering high.} However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
@@
 		\hline
-		Model & $w\ \langle\mu\rangle,\ \langle\sigma_w\rangle$ (scatter) & $w$-Bias \\
-		\hline
-		{\gten} Stat + Syst  &  $-0.998, \ 0.097\ (0.073)$ & $(0.02\pm0.07)\sigma$  \\
-		{\gten}  Stat         &  $-1.008,\ 0.080\ (0.068)$ & $(-0.10\pm0.08)\sigma$   \\
-		{\celeven} Stat + Syst  &  $-0.945,\ 0.098\ (0.077)$ & $(0.55\pm0.08)\sigma$  \\
-		{\celeven} Stat         &  $-0.948,\ 0.079\ (0.066)$ & $(0.65\pm0.08)\sigma$   \\
+		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias \\
+		\hline
+		{\gten} Stat + Syst  &  $-0.998 \ [0.097\ (0.073)]$ & $(0.02\pm0.07)\sigma$  \\
+		{\gten}  Stat         &  $-1.008\ [0.080\ (0.068)]$ & $(-0.10\pm0.08)\sigma$   \\
+		{\celeven} Stat + Syst  &  $-0.945\ [0.098\ (0.077)]$ & $(0.55\pm0.08)\sigma$  \\
+		{\celeven} Stat         &  $-0.948\ [0.079\ (0.066)]$ & $(0.65\pm0.08)\sigma$   \\
 		\hline
@@
 It disagrees with the value provided in \citet[Table 10, row 3]{Brout18SYS} which uses inverse variance weighted averages.
-We do not utilise this weight because the variance is correlated with the value of $w$ due to the $\Omega_m$ prior applied in the fitting process.}
+We do not utilise this weight because the variance is correlated with the value of $w$ due to the $\Omega_m$ prior applied in the fitting process.
+We note that if inverse variance weighting is applied to both datasets, they both shift by $\Delta w \approx 0.01$, and thus the predicted difference between the BBC method and {\steve} remains the same.}
 \begin{figure}

### 2018-10-10 papers/methods/paper.tex
Updating sims to test C11 pops

@@
 We do not utilise this weight because the variance is correlated with the value of $w$ due to the $\Omega_m$ prior applied in the fitting process.
-We note that if inverse variance weighting is applied to both datasets, they both shift by $\Delta w \approx 0.01$, and thus the predicted difference between the BBC method and {\steve} remains the same.}
+We note that if inverse variance weighting is applied to both datasets, they both shift by $\Delta w \approx 0.005$, and thus the predicted difference between the BBC method and {\steve} remains the same.}
 \begin{figure}

### 2018-10-10 dessn/framework/simulations/selection_effects.py

@@ -225,13 +225,13 @@ def test_colour_contribution():
     plt.show()
 
 if __name__ == "__main__":
-    get_selection_effects_skewnorm("snana_data/DES3YR_LOWZ_BHMEFF_G10", plot=True)
+    # get_selection_effects_skewnorm("snana_data/DES3YR_LOWZ_BHMEFF_G10", plot=True)
     # get_selection_effects_skewnorm("snana_data/DES3YR_LOWZ_BHMEFF_C11", plot=True)
     # test_colour_contribution()
     get_selection_effects_cdf("snana_data/DES3YR_DES_BHMEFF_AMG10", plot=True)
     print("---")
-    # get_selection_effects_cdf("snana_data/DES3YR_DES_BHMEFF_AMC11", plot=True)
-    # print("---")
+    get_selection_effects_cdf("snana_data/DES3YR_DES_BHMEFF_AMC11", plot=True)
+    print("---")
     # get_selection_effects_cdf("snana_data/DES3YR_DES_BHMEFF_CD", plot=True, cut_mag=19)
 
     # _, mean, cov = get_selection_effects_cdf("snana_data/DES3YR_DES_BHMEFF_AMG10")

### 2018-10-10 dessn/snana/convert_snana_data.py

@@ -339,7 +339,7 @@ def convert(base_folder, load_dump=False, override=False, skip=11, biascor=None,
 if __name__ == "__main__":
     logging.basicConfig(level=logging.DEBUG, format="[%(funcName)20s()] %(message)s")
     # convert("DES3YR_LOWZ_COMBINED_TEXT_v8")
-    convert("DES3YR_DES_COMBINED_TEXT_v8")
+    # convert("DES3YR_DES_COMBINED_TEXT_v8")
     # convert("DES3YR_DES_NOMINAL")
     # convert("DES3YR_LOWZ_NOMINAL")
     # convert("DES3YR_DES_BULK_v8", skip=6)
@@ -351,7 +351,7 @@ if __name__ == "__main__":
     # convert("DES3YR_LOWZ_BULK", skip=6)
     # convert("DES3YR_DES_SAMTEST", skip=11)
     # convert("DES3YR_LOWZ_SAMTEST", skip=11)
-    # convert("DES3YR_DES_BHMEFF_v8", load_dump=True, skip=11, zipped=True)
+    convert("DES3YR_DES_BHMEFF_v8", load_dump=True, skip=11, zipped=True)
     # convert("DES3YR_LOWZ_BHMEFF_v8", load_dump=True, skip=11, zipped=True)
     # convert("DES3YR_LOWZ_VALIDATION", skip=11)
     # convert("DES3YR_DES_VALIDATION", skip=11)

### 2018-10-11 papers/methods/paper.tex
More of ricks feedback

@@
 \newcommand{\kmsmpc}{{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
-\newcommand{\cov}{\mathcal{C}}
+\newcommand{\cov}{\mathcal{C}_\eta}
 \newcommand{\Z}{\mathcal{Z}}
@@
 For a recent review, see \citet{Huterer2018}.
-{\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}
+{\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}
 
@@
 The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has {\green discovered} thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
-The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will {\green discover} scores of thousands of photometrically classified supernovae.
+The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will {\green discover tens} of thousands of photometrically classified supernovae.
 {\green Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics} \citep{Betoule2014, Scolnic2017}.
 
-As such, {\green considerable resources are} being put into developing more sophisticated supernova cosmology analyses.
+As such, {\green considerable resources are aimed at} developing more sophisticated supernova cosmology analyses.
 The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models.
-First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved in \citet{Kessler2009}, simulations are now a fundamental component of modern supernova cosmology.
-\citet{Betoule2014} quantise and then correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN Ia distances as a function of {\green multiple} factors to better remove biases.
+First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for {\green SDSS \citep{Kessler2009},} simulations are a fundamental component of modern supernova cosmology.
+\citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of {\green multiple factors to improve bias correction}.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections} or attempt to find analytic approximations for complicated effects such as Malmquist bias.
+Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself}.
 
@@
 We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty.
-We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.} Section \ref{sec:review} is dedicated to a quick review of {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
+We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.}
+
+Section \ref{sec:review} is dedicated to a quick review of {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
@@
 This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}.
-The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C}$).
+The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C_\eta}$).
 As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
@@
 Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditionally in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space.
-Specifically, for any given supernova our map between absolute magnitude and apparent magnitude may take the traditional form:
-
-\begin{align}
-\mu(z) = m_B + \alpha x_1 - \beta c - M_B + \text{other corrections}, \label{eq:standard}
-\end{align}
-
-where $M_B$ is the mean absolute magnitude for all SN Ia,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
-The cosmological term, $\mu(z)$ represents the distance modulus, and can be precisely calculated given cosmological parameters and a redshift, or can be `observed' by {\green utilising equation \eqref{eq:standard} and supernova observations}.
-The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
-In traditional analyses, Malmquist bias corrections are also included, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Scolnic2017}.
+Specifically, for any given supernova our map between {\green absolute magnitude and apparent magnitude is given by the distance modulus:}
+
+\begin{align}
+\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \text{other corrections}, \label{eq:standard}
+\end{align}
+
+where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
+{\green The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift.} The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
+In traditional $\chi^2$ analyses, Malmquist bias corrections are also included, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.
 
@@
 
-Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted values.
-We denote the observed distance modulus $\mu_{\rm obs}$ and the cosmologically calculated one $\mu_C$.
+Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus.
 {\green The $\chi^2$ function minimised is}
@@
 \begin{align}
-\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
-\end{align}
-
-where $C^{-1}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties). The cosmological $\mu_C$ is defined as
+\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
+\end{align}
+
+where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis).
+The predicted $\mu_C$ is defined as
 {\green 
@@
 \mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
-d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}},
+d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
+E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
 \end{align}
@@
 The speed comes with several limitations.
-Firstly, formulating a $\chi^2$ likelihood requires a loss of model flexibility by building into the model assumptions of Gaussian uncertainty.
-Secondly, the method of creating a global covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
-For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is lost when determining population uncertainty using numerical derivatives of population permutations.
-{\green Whilst known correlations like this can be captured and incorporated into a covariance matrix if known, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.} Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
+Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty.
+Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
+For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations.
+{\green Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.} Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which then produces a matrix of arbitrary size {\green and can allow fitting without matrix inversion on every step}.
+A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which produces a matrix of manageable size {\green and can allow fitting without matrix inversion on every step}.
 Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
 
-Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting data.
-Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift when supernovae increasingly fall below the detection threshold.
-An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}.
+Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting the determined $\mu_{\rm obs}$ from the data, or equivalently, adding a distance bias to the $\mu_C$ prediction.
+Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift.
+An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}.
 Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data.
-When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data, which means their uncertainty is not captured fully in the $\chi^2$ distribution, and any subtle correlations between cosmological or population parameters and the bias is lost.
-{\green Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections.}
+When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data.
+This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost.
+{\green Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.}
 
@@
 For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
-We can thus easily incorporate different distributions, populations, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.
+We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.
 
 With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting.
+\citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting.
 \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction.
-\citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations.
+\citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations.
 The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
 \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}.
-Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The simplification made by the UNITY analysis is that the bias is {\green perfectly described} by an analytic function.
-Their model was validated to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}.
+Notably, these methods rely on data that is bias corrected or {\green the methods }ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
+The assumption made by the UNITY analysis is that the bias {\green in real data} is {\green perfectly described} by an analytic function.
+{\green They validate their } model to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}.
 The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
@@
 {\green Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
-However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to $w$.}
-
-The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections and redshift evolution, each of which will be discussed further in the outline of our model below.
+However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.}
+
+The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections, and redshift evolution, each of which will be discussed further in the outline of our model below.
 Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
-{\green We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections} to provide a valuable cross-check on analysis methodologies which are highly simulation-dependent.
+{\green We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort} to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.
 
@@
 
-{\green Forefront among these challenges is our ignorance of the underlying type Ia population -- its intrinsic dispersion}.
-Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements.
-We{\green test against }two dispersion models in this work.
+{\green Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion.
+Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae.
+However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements.
+The underlying dispersion model is not a solved problem, and we therefore test against }two dispersion models in this work.
 The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
 The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$.
+{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$.
 These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population.
@@
 The underlying supernova population is further complicated by the presence of outliers.
-Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN~Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts.
-Depending on the cuts {\green optimising the} ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification \citep{Gupta2016} and results in a broad population of outliers.
+Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN~Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts {\green being assigned to supernovae.
+Different optimizations to the host-galaxy algorithm can result in misidentification of the host at the 3\% to 9\% level \citep{Gupta2016}, resulting in a broad population of outliers}.
 For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
@@
 Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics.
-Because of the rarity of SN~Ia, datasets are commonly formed from the discoveries of multiple surveys.
-However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
+{\green Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset.} However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
 Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
@@
 
-We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, {\green treatment of an increased number of systematics}, and analytic correction of selection effects, including systematic uncertainty on those corrections.
-We also desire this to be more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
+We construct our hierarchical Bayesian model {\steve} with several goals in mind: creation of a redshift-dependent underlying supernova population, {\green treatment of an increased number of systematics}, and analytic correction of selection effects, including systematic uncertainty on those corrections.
+We also desire {\steve} more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
 As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
@@
 \subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a flux measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement takes the form of a probability of being above $10^{10}$ solar masses.
-We will also have a probability of each supernova being a type Ia, $\hat{p}$.
-Our set of observables {\green input into the BHM model} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses.
+We also have a probability of each supernova being a type Ia, $\hat{p}$.
+Our set of observables {\green input into the {\steve}} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
-This is discussed further in Section \ref{sec:systreat}.
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, {\green these misclassifications} can naturally be modelled and taken into account by introducing secondary populations that supernovae have a non-zero probability of belonging to.
+One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, {\green these misclassifications} can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.
 
@@
 
-{\green The first layer of hierarchy are the true (latent) parameters that describe each supernova.
-In contrast to the observed parameters, the latent parameters are denoted without a hat.} For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the color we observe, which, as it has measurement error, is different from $c$.
+{\green The first layer of hierarchy is the set of true (latent) parameters that describe each supernova.
+In contrast to the observed parameters, the latent parameters are denoted without a hat.} For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the observed color, which, as it has measurement error, is different from $c$.
 
@@
 For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainty.
-However, this is computationally prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C}$ describing the uncertainty on $\hat{\eta}$.
-When using this simplification, our latent variables are given by
-\begin{align}
-p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}). \label{eq:pop}
-\end{align}
-As discussed in Section \ref{sec:challenges}, {\green the SALT2 model does not include full treatment of intrinsic dispersion in the supernova model, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
+A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainties.
+However, this is computationally prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$.
+Using this simplification, our latent variables are given by
+\begin{align}
+p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
+\end{align}
+As discussed in Section \ref{sec:challenges}, {\green the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
 
@@
 	% Obs nodes
-	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C}^{-1}$};
+	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
@@
 \label{sec:underlying}
-The underlying supernova population is characterised by underlying distributions in color, stretch and absolute magnitude.
-We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal for each survey.
-For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal.
-{\green We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no cosmological benefit with the skew-normal distributions for stretch and magnitude.} Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
+The underlying supernova population is characterised by distributions in color, stretch and absolute magnitude.
+We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey.
+For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution.
+{\green We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude.} Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
 Both the color and stretch means are modelled with normal priors.
-Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment.
-We parametrise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ {\green that} allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
+Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment.
+We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ {\green that} allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
+We sample $\delta_c$ in log space for efficiency in sampling close to zero.
 The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors.
-We sample $\delta_c$ in log space for efficiency in sampling close to zero.
-
-The only constant between survey populations is the mean absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
+
+The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
@@
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We {\green should also} do this without assuming a specific dispersion model -- either the {\gten} or {\celeven} model, {\green or utilise the results of computing the bias from both models.}
-
-We model the extra dispersion only in color, and do so by adding extra independent uncertainty on the color observation {\green $\hat{c}$}.
+We {\green should also avoid} assuming a specific dispersion model -- either the {\gten} or {\celeven} model, {\green or utilise the results of computing the bias from both models.}
+
+We model the extra dispersion only in color, and do so by adding independent uncertainty on the color observation {\green $\hat{c}$}.
 We note that extra dispersion in magnitude {\green $\hat{m}_B$} (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the color term, hence the requirement for modelling additional color dispersion.
@@
 As shown in \citep{Kessler2013}, the extra color dispersion shows heavy redshift dependence, increasing with redshift.
-As the extra dispersion may arise from underlying supernova physics, rather than {\green any selection or observational effects}, we decide to incorporate redshift dependence in our extra uncertainty.
+{\green This is an artifact of different filters, however as we may be subject to similar effects in our observational data,} we decide to incorporate redshift dependence in our extra uncertainty.
 We thus add $\kappa_0 + \kappa_1 z$ to our observed color uncertainty (in quadrature).
@@
 We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.
-As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
+As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C_\eta} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
 
@@
 
-It is now well known that host galaxy environment is significantly correlated with supernova properties.
+{\green There are numerous results showing } that host galaxy environment is correlated with supernova properties.
 The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
-The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step.
-The scale of this step function varies from analysis to analysis, with the 0.08 valuesourced from \cite{Sullivan2010} and used in \citet{Betoule2014}.
+The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step.
+The scale of this step function varies from analysis to analysis, and is treated as a fit parameter.
 In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
@@
 \begin{align}
-M_B = m_B + \alpha x_1 - \beta c + \Delta M \cdot \hat{m}. \label{eq:l2}
+\mu_{\rm obs} = m_B + \alpha x_1 - \beta c + \Delta M \cdot \hat{m} - M_B. \label{eq:l2}
 \end{align}
@@
 The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations.
-As such, the normal treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
-
-Assuming that the gradients can be linearly extrapolated -- which is a reasonable approximation for modern surveys with high quality control of systematics -- we can incorporate into our model a deviation from the observed original values by constructing a $(3 \times N_{\rm sys})$ matrix containing the numerical derivatives for the $N_{\rm sys}$ systematics and multiplying it with the row vector containing the offset for each systematic.
+As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
+
+Assuming that the gradients can be linearly extrapolated -- which is a reasonable approximation for modern surveys with high quality control of systematics -- we can incorporate into our model a deviation from the observed values by constructing a $(3 \times N_{\rm sys})$ matrix containing the numerical derivatives for the $N_{\rm sys}$ systematics and multiplying it with the row vector containing the offset for each systematic.
 By scaling the gradient matrix to represent the shift over $1\sigma$ of systematic uncertainty, we can simply enforce a unit normal prior on the systematic row vector to increase computational efficiency.
 
-This method of adjusting the observed summary statistics is used throughout the traditional and BHM analyses.
-For each survey and band, we have two systematics --  the calibration uncertainty and the filter wavelength uncertainty.
+This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses.
+For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty.
 We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
@@
 First, we validate on toy models, verifying that we recover accurate cosmology {\green when generating toy supernovae data constructed to satisfy the assumptions of the BHM construction}.
-We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample.
+We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample, termed the DES-SN3YR sample
 
@@
 	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys.
-Selection effects are visible in both samples, where red supernovae are often cut as redshift increases.
+Selection effects are visible in both samples, where red supernovae are often cut as redshift increases, creating a skewed color population.
 The color of the data points is representative of the supernova color itself, a negative color value showing bluer supernovae, with positive color values representing redder supernovae.}
@@
 
-The selection functions (a skew normal for low-redshift and an error function for high-redshift) are given independent uncertainty of $0.01$ on all parameters (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
-Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the combined DES three-year and low-redshift spectroscopic analysis.
+The selection functions parameters (a skew normal for low-redshift and an error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
+Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample.
 Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
@@
 The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}.
-Without making such a slice, the variation in $w$ can appear to be large due to a shift along the degeneracy direction of the banana.
+Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'.
 By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them.
@@
 However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways.
-To this end, we test using simulations} (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is comprised of CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+To this end, we test using simulations} (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
 Simulation specifics can be found in \citet{Kessler18}.
 
-Additionally, early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
+Additionally, prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
 
-In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics.
+In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics as detailed in \citet{Brout18SYS}.
 Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
@@
 It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based.
-We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.
+We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models.
 
@@
 	\label{tab:parameter_correlations}
-	\begin{tabular}{c|cc}
+	\begin{tabular}{c|rr}
 		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst  \\
 		\hline
-		$\Omega_m$ &   -0.19 &     -0.21     \\
-		$\alpha$ &   -0.17 &     -0.20     \\
-		$\beta$ &   -0.29 &     -0.23     \\
+		$\Omega_m$ &   $-$0.19 &     $-$0.21     \\
+		$\alpha$ &   $-$0.17 &     $-$0.20     \\
+		$\beta$ &   $-$0.29 &     $-$0.23     \\
 		$\langle M_B \rangle$ &    0.68 &      0.66     \\
@@
 		$\sigma_{c}^{1}$ &    0.08 &      0.04     \\
-		$\alpha_c^{0}$ &   -0.04 &      0.04     \\
+		$\alpha_c^{0}$ &   $-$0.04 &      0.04     \\
 		$\alpha_c^{1}$ &    0.03 &      0.01     \\
-		$\kappa_{c0}^{0}$ &   -0.10 &     -0.05     \\
-		$\kappa_{c0}^{1}$ &   -0.20 &     -0.17     \\
-		$\kappa_{c1}^{0}$ &   -0.05 &     -0.01     \\
-		$\kappa_{c1}^{1}$ &   -0.01 &      0.01     \\
+		$\kappa_{c0}^{0}$ &   $-$0.10 &     $-$0.05     \\
+		$\kappa_{c0}^{1}$ &   $-$0.20 &     $-$0.17     \\
+		$\kappa_{c1}^{0}$ &   $-$0.05 &     $-$0.01     \\
+		$\kappa_{c1}^{1}$ &   $-$0.01 &      0.01     \\
 		%$s_c^{0}$ &    0.00 &      0.01     \\
@@
 		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
-		$\langle x_1^{0} \rangle$ &   -0.01 &     -0.05     \\
-		$\langle x_1^{1} \rangle$ &   -0.02 &      0.02     \\
-		$\langle x_1^{2} \rangle$ &   -0.04 &     -0.04     \\
-		$\langle x_1^{3} \rangle$ &   -0.03 &     -0.06     \\
-		$\langle x_1^{4} \rangle$ &   -0.06 &     -0.06     \\
+		$\langle x_1^{0} \rangle$ &   $-$0.01 &     $-$0.05     \\
+		$\langle x_1^{1} \rangle$ &   $-$0.02 &      0.02     \\
+		$\langle x_1^{2} \rangle$ &   $-$0.04 &     $-$0.04     \\
+		$\langle x_1^{3} \rangle$ &   $-$0.03 &     $-$0.06     \\
+		$\langle x_1^{4} \rangle$ &   $-$0.06 &     $-$0.06     \\
 		$\langle x_1^{5} \rangle$ &    0.04 &      0.02     \\
@@
 		$\langle x_1^{7} \rangle$ &    0.08 &      0.03     \\
-		$\langle c^{0} \rangle$ &   -0.05 &     -0.12     \\
+		$\langle c^{0} \rangle$ &   $-$0.05 &     $-$0.12     \\
 		$\langle c^{1} \rangle$ &    0.11 &      0.03     \\
@@
 		$\langle c^{3} \rangle$ &    0.14 &      0.04     \\
-		$\langle c^{4} \rangle$ &   -0.11 &     -0.11     \\
-		$\langle c^{5} \rangle$ &   -0.15 &     -0.08     \\
-		$\langle c^{6} \rangle$ &   -0.12 &     -0.13     \\
-		$\langle c^{7} \rangle$ &   -0.12 &     -0.06     \\
+		$\langle c^{4} \rangle$ &   $-$0.11 &     $-$0.11     \\
+		$\langle c^{5} \rangle$ &   $-$0.15 &     $-$0.08     \\
+		$\langle c^{6} \rangle$ &   $-$0.12 &     $-$0.13     \\
+		$\langle c^{7} \rangle$ &   $-$0.12 &     $-$0.06     \\
 		$\delta [ {\rm SALT}_0 ]$ &    0.05 &      0.05     \\
-		$\delta [ {\rm SALT}_1 ]$ &   -0.01 &      0.02     \\
-		$\delta [ {\rm SALT}_2 ]$ &   -0.10 &     -0.09     \\
-		$\delta [ {\rm SALT}_3 ]$ &   -0.03 &     -0.03     \\
+		$\delta [ {\rm SALT}_1 ]$ &   $-$0.01 &      0.02     \\
+		$\delta [ {\rm SALT}_2 ]$ &   $-$0.10 &     $-$0.09     \\
+		$\delta [ {\rm SALT}_3 ]$ &   $-$0.03 &     $-$0.03     \\
 		$\delta [ {\rm SALT}_4 ]$ &    0.08 &      0.09     \\
@@
 		$\delta [ {\rm SALT}_6 ]$ &    0.05 &      0.07     \\
-		$\delta [ {\rm SALT}_7 ]$ &   -0.11 &     -0.10     \\
+		$\delta [ {\rm SALT}_7 ]$ &   $-$0.11 &     $-$0.10     \\
 		$\delta [ {\rm SALT}_8 ]$ &    0.01 &      0.02     \\
@@
 		$\delta [ {\rm MWE}_{B-V} ]$ &    0.03 &      0.02     \\
-		$\delta [ {\rm HST Calib} ]$ &   -0.07 &     -0.07     \\
-		$\delta [ v_{\rm pec} ]$ &    0.00 &     -0.01     \\
+		$\delta [ {\rm HST Calib} ]$ &   $-$0.07 &     $-$0.07     \\
+		$\delta [ v_{\rm pec} ]$ &    0.00 &     $-$0.01     \\
 		$\delta [ {\rm \delta z } ]$ &    0.01 &      0.00     \\
@@
 		$\delta [ \Delta r ]$ &    0.16 &      0.10     \\
-		$\delta [ \Delta i ]$ &   -0.16 &     -0.18     \\
-		$\delta [ \Delta z ]$ &   -0.26 &     -0.26     \\
+		$\delta [ \Delta i ]$ &   $-$0.16 &     $-$0.18     \\
+		$\delta [ \Delta z ]$ &   $-$0.26 &     $-$0.26     \\
 		$\delta [ \Delta \lambda_g ]$ &    0.16 &      0.20     \\
 		$\delta [ \Delta \lambda_r ]$ &    0.05 &      0.06     \\
-		$\delta [ \Delta \lambda_i ]$ &    0.00 &     -0.01     \\
+		$\delta [ \Delta \lambda_i ]$ &    0.00 &     $-$0.01     \\
 		$\delta [ \Delta \lambda_z ]$ &    0.09 &      0.07     \\
@@
 
-Each realisation of simulated cosmology contains 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the final DES spectroscopic analysis.
+Each realisation of simulated SN~Ia light curves contains {\green the SALT2 light-curve fits and redshifts} to 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the DES-SN3YR sample.
 As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
@@
 We investigate the cosmological bias and find its source to be a bias in the observed summary statistics {\green(i.e.
-the $m_B$, $x_1$, $c$ output from SALT2 light curve fitting)}, in addition to incorrect reported uncertainty on the summary statistics.
+the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting)}, in addition to incorrect reported uncertainty on the summary statistics.
 To confirm this, we run two tests.
-The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
+The first of which, we replace the SALT2-fitted $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a Gaussian centered on the true SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test, to {\green test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through, we} test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+Our second test aims to {\green test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
+To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter,} and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
-Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.
+Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced by an unknown scatter model.
 
@@
 With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
-To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES three-year spectroscopic survey with various model parameters allowed to either vary or stay locked to a fixed value.
+To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
 By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
 
-The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is still dominated by statistical error, as the total statistical uncertainty is on $w$ is $\pm0.08$.
-With the low number of supernovae in the DES three-year spectroscopic sample, this is expected.
+The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is dominated by statistical error, as the total statistical uncertainty is on $w$ is $\pm0.08$.
+With the low number of supernovae in the DES-SN3YR sample, this is expected.
 We note that the label `Systematics' in Table \ref{tab:uncert} represents all numerically computed systematics (as discussed in Section \ref{sec:systreat}) and systematic uncertainty on the selection function.
@@
 
-We compare the results of our model against those of the BBC method \citep{Kessler2017}.
-BBC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017}.
-{\green The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected observables are fit using CosmoMC \citep{lewis2002cosmomc}.
-For shorthand, we refer to this combined process as the BBC method in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency{\green check} as to the current levels of accuracy in recovering cosmological parameters.
-
-{\green To this end, we take the results of the BBC method which were also run on the} set of 200 validation simulations and compare the recovered $w$ values to those of our method.
+We compare the results of our model against those of the BBC+CosmoMC method \citep{Kessler2017}.
+BBC+CosmoMC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017} and is being used in the primary analysis of the DES-SN3YR sample \citep{DESKEY}.
+{\green The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected distances are fit using CosmoMC \citep{lewis2002cosmomc}.
+For shorthand, we refer to this combined process as the BBC method hereafter in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency{\green check} as to the current levels of accuracy in recovering cosmological parameters.
+
+{\green To this end, we take the results of the BBC method which were also run on the} same set of 200 validation simulations and compare the recovered $w$ values to those of our method.
 The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
 
-As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
+As shown in \citet{Brout18SYS}, the BBC method recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
 As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}.
-As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
+As such, we expect the BBC method to have a $w$-bias in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
 {\green These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}.
-Both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model.
+Both BBC method and {\steve} are sensitive to the intrinsic scatter model, finding differences of $\sim0.066$ and $0.053$ respectively in $w$ when varying the scatter model.
 The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$}.
-Our method shows a small improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}.
+Our method shows a small improvement in the insensitivity to the intrinsic scatter model (having a decrease in difference in $w$ between the {\gten} and {\celeven} models), finding no bias for {\gten} but a $w$ biased high for {\celeven}.
 This decrease in error is not statistically significant as we have statistical uncertainty of $\sim0.01$ for $100$ simulation realisations.
@@
 For the {\gten} model this difference is a result of bias in the BBC results, however for the {\celeven} simulations this is a result of both bias from BBC, and a larger bias from our method.
-These results also allow us to state the expected values for $w$ when run on the DES three year spectroscopic survey sample.
-When using Planck priors our uncertainty on $w$ is reduced compared to using our simulation Gaussian prior on $\Omega_m$, shrinking the average scatter from $0.06$ to $0.04$.
+These results also allow us to state the expected values for $w$ when run on the DES-SN3YR sample.
+When using Planck priors our uncertainty on $w$ is reduced compared to using our simulation Gaussian prior on $\Omega_m$, shrinking the average $w$-difference from $0.06$ to $0.04$.
 After factoring this into our uncertainty, we expect our BHM method to, on average, recover $w_{\rm BHM} = w_{\rm BBC} +0.04\pm0.04$.

### 2018-10-11 papers/methods/snippets/malmquist.py

@@ -42,10 +42,10 @@ ax.scatter(sn_zs[~mask], sn_disp[~mask], s=2, c="#f44141", alpha=0.4, label="Uno
 
 cax = ax.scatter(bin_center, means, c=np.abs(bias), s=50, lw=1, edgecolor='k', label="SN means")
 cbar = fig.colorbar(cax)
-cbar.set_label(r"$\Delta \mu$", fontsize=14)
+cbar.set_label(r"$\langle\mu_{\rm obs} - \mu_C\rangle$", fontsize=14)
 
 ax.set_xlabel("$z$", fontsize=14)
-ax.set_ylabel(r"$\mu$", fontsize=14)
+ax.set_ylabel(r"$\mu_{\rm obs}$", fontsize=14)
 
 ax.legend(loc=4, frameon=False, markerfirst=False)
 

### 2018-10-11 papers/methods/paper.tex
More of ricks feedback

@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
+{\green With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.}
 
@@
 
+{\green Failure to provide an adequate parametrisation for realistic intrinsic dispersion and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters.
+Tests when fixing the population parameters to true does not resolve the cosmological bias observed in the {\celeven} simulations.
+As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.}
+
 % color bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain

### 2018-10-14 papers/methods/paper.tex
emulateapj

--- 2018-10-11 papers/methods/paper.tex
@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\documentclass[a4paper,fleqn,usenatbib]{emulateapj}
 
@@
 % Keep the title short and informative.
-\title[\name]{\name: A hierarchical Bayesian model for Supernova Cosmology}
-
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[S. R. Hinton et al.]{
-	Samuel R. Hinton,$^{1,2}$\thanks{E-mail: samuelreay@gmail.com}
-	Alex G. Kim,$^{3}$
-	Tamara M. Davis$^{1,2}$
-	\\
-	% List of institutions
-	$^{1}$School of Mathematics and Physics, The University of Queensland, Brisbane, QLD 4072, Australia\\
-	$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
-	$^{3}$Physics Division, Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA
-}
-
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2017}
-
+
+\shorttitle{\name}
 % Don't change these lines
 \begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
+\title{\name: A hierarchical Bayesian model for Supernova Cosmology}
+
+ The list of authors, and the short list which is used in the headers.
+ % If you need two or more lines of authors, add an extra line using \newauthor
+ \author{
+ 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
+ 	T.~M.~Davis\altaffilmark{1,2},
+ 	A.~G.~Kim\altaffilmark{3},
+ 	A.~M\"oller\altaffilmark{2,4},
+ 	M.~Sako\altaffilmark{5},
+ 	M.~Smith\altaffilmark{6},
+ 	\\}
+ \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
+ \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+ \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
+ \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+ \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+ \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
 
@@
 %
-Given light curve fit parameters, redshifts, and host-galaxy masses, {\green we fit {\steve}} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
+Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
 Selection effects are characterised using Monte-Carlo simulations.
@@
 %
-These simulation test results are a concern {\green for BHM applications on large upcoming surveys}, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
+These simulation test results are a concern for BHM applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
 \end{abstract}
-% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.
-
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-keyword1 -- keyword2 -- keyword3
-\end{keywords}
 
@@
 For a recent review, see \citet{Huterer2018}.
-{\green While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.}
-
-In attempts to tease out the nature of dark energy, active and planned surveys are {\green continually growing in size and scale}.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has {\green discovered} thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
-The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will {\green discover tens} of thousands of photometrically classified supernovae.
-{\green Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics} \citep{Betoule2014, Scolnic2017}.
-
-As such, {\green considerable resources are aimed at} developing more sophisticated supernova cosmology analyses.
+While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.
+
+In attempts to tease out the nature of dark energy, active and planned surveys are continually growing in size and scale.
+The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has discovered thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
+The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will discover tens of thousands of photometrically classified supernovae.
+Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics \citep{Betoule2014, Scolnic2017}.
+
+As such, considerable resources are aimed at developing more sophisticated supernova cosmology analyses.
 The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models.
-First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for {\green SDSS \citep{Kessler2009},} simulations are a fundamental component of modern supernova cosmology.
-\citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of {\green multiple factors to improve bias correction}.
+First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for SDSS \citep{Kessler2009}, simulations are a fundamental component of modern supernova cosmology.
+\citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of multiple factors to improve bias correction.
 Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, {\green and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself}.
-
-{\green In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}.
+Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself.
+
+In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}.
 We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty.
-We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.}
-
-Section \ref{sec:review} is dedicated to a quick review of {\green supernova cosmology analysis methods}, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
+We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.
+
+Section \ref{sec:review} is dedicated to a quick review of supernova cosmology analysis methods, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
 In Section \ref{sec:method} we outline our methodology.
@@
 
-Whilst supernova observations take the form of photometric time-series brightness measurements in many bands {\green and a redshift measurement of the supernova (or its assumed host)}, most analyses do not work from these measurements directly.
+Whilst supernova observations take the form of photometric time-series brightness measurements in many bands and a redshift measurement of the supernova (or its assumed host), most analyses do not work from these measurements directly.
 Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
@@
 
-This {\green ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$} represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
-Accurately characterising this underlying population, its evolution over redshift, and effects from {\green host-galaxy} environment, is one of the challenges of supernova cosmology.
-Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ {\green(absolute magnitude of the supernova, traditionally in the Bessell $B$ band)}, $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space.
-Specifically, for any given supernova our map between {\green absolute magnitude and apparent magnitude is given by the distance modulus:}
-
-\begin{align}
-\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \text{other corrections}, \label{eq:standard}
+This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
+Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology.
+Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space.
+Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:
+
+\begin{align}
+\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
 \end{align}
@@
 where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
-{\green The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift.} The `$\text{other corrections}$' term often includes corrections for host galaxy environment, as this has statistically significant correlations with supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013, Uddin2017}.
-In traditional $\chi^2$ analyses, Malmquist bias corrections are also included, which can take the form of a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.
+$\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}.
+The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift.
+The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses.
+Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.
 
@@
 Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus.
-{\green The $\chi^2$ function minimised is}
+The $\chi^2$ function minimised is
 
@@
 The predicted $\mu_C$ is defined as
-{\green 
+
 \begin{align}
@@
 \end{align}
-}
+
 where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
@@
 For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations.
-{\green Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.} Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
+Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.
+Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A recent solution to this computational cost problem is to bin the supernovae {\green in redshift bins}, which produces a matrix of manageable size {\green and can allow fitting without matrix inversion on every step}.
+A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion on every step.
 Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
@@
 An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}.
-Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then {\green subtracted from} the observational data.
+Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then subtracted from the observational data.
 When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data.
 This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost.
-{\green Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.}
+Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.
 
@@
 
-\subsection{Hierarchical Bayesian Models}
+\subsection{Bayesian Hierarchical Models}
 
 Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM).
-Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and {\green physically} motivated conditional probabilities.
+Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and physically motivated conditional probabilities.
 For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
 We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.
+Unlike ABC methods, which can model arbitrary probability distributions, BHM methods are generally constrained to representing probabilities using analytic forms.
 
@@
 \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}.
-Notably, these methods rely on data that is bias corrected or {\green the methods }ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The assumption made by the UNITY analysis is that the bias {\green in real data} is {\green perfectly described} by an analytic function.
-{\green They validate their } model to be free of significant biases using fits to {\green thirty realisations of supernova datasets that are constructed to mimic the UNITY framework}.
+Notably, these methods rely on data that is bias corrected or the methods ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
+The assumption made by the UNITY analysis is that the bias in real data is perfectly described by an analytic function.
+They validate their  model to be free of significant biases using fits to thirty realisations of supernova datasets that are constructed to mimic the UNITY framework.
 The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
@@
 Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias.
-{\green Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
-However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.}
+Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
+However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.
 
@@
 Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
-{\green We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort} to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.
+We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.
 
@@
 
-{\green The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.} In this Section, we outline several of the most prominent challenges.
-
-{\green Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion.
+ The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.
+In this Section, we outline several of the most prominent challenges.
+
+ Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion.
 Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae.
 However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements.
-The underlying dispersion model is not a solved problem, and we therefore test against }two dispersion models in this work.
-The {\green first is based on the} \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation and 30\% from chromatic variation.
-The second, {\green based on the} {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation.
-{\green These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.} As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$.
-These two scatter models are currently {\green assumed} to span the possible range of scatter in the underlying supernova population.
+The underlying dispersion model is not a solved problem, and we therefore test against two dispersion models in this work.
+The first is based on the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, the second is sourced from \citet[][hereafter denoted {\celeven}]{Chotard2011}.
+As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$.
+These two scatter models are currently assumed to span the possible range of scatter in the underlying supernova population.
 We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.
@@
 The underlying supernova population is further complicated by the presence of outliers.
-Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae {\green and represent outliers from the standardised SN~Ia sample.} This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts {\green being assigned to supernovae.
-Different optimizations to the host-galaxy algorithm can result in misidentification of the host at the 3\% to 9\% level \citep{Gupta2016}, resulting in a broad population of outliers}.
+Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae  and represent outliers from the standardised SN~Ia sample.
+This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts  being assigned to supernovae.
+Different optimizations to the host-galaxy algorithm can result in misidentification of the host at the 3\% to 9\% level \citep{Gupta2016}, resulting in a broad population of outliers.
 For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
@@
 Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics.
-{\green Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset.} However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
+Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset.
+However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
 Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
@@
 
-We construct our hierarchical Bayesian model {\steve} with several goals in mind: creation of a redshift-dependent underlying supernova population, {\green treatment of an increased number of systematics}, and analytic correction of selection effects, including systematic uncertainty on those corrections.
+We construct our hierarchical Bayesian model {\steve} with several goals in mind: creation of a redshift-dependent underlying supernova population,  treatment of an increased number of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
 We also desire {\steve} more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
@@
 
-At the most fundamental level, a supernova cosmology analysis is a mapping from an underlying population onto an observed population, where {\green cosmological parameters are} encoded directly in the mapping function.
-The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both {\green the observed and underlying} populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
+At the most fundamental level, a supernova cosmology analysis is a mapping from an underlying population onto an observed population, where cosmological parameters are encoded directly in the mapping function.
+The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both the observed and underlying populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
 In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
@@
 Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host galaxy {\green stellar }mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses.
+Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses.
 We also have a probability of each supernova being a type Ia, $\hat{p}$.
-Our set of observables {\green input into the {\steve}} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, {\green these misclassifications} can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.
+One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.
 
@@
 
-{\green The first layer of hierarchy is the set of true (latent) parameters that describe each supernova.
-In contrast to the observed parameters, the latent parameters are denoted without a hat.} For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the observed color, which, as it has measurement error, is different from $c$.
+The first layer of hierarchy is the set of true (latent) parameters that describe each supernova.
+In contrast to the observed parameters, the latent parameters are denoted without a hat.
+For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the observed color, which, as it has measurement error, is different from $c$.
 
@@
 For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves {\green fluxes} and uncertainties.
-However, this is computationally prohibitive, and as such we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$.
+A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves fluxes and uncertainties.
+However, this is computationally prohibitive  as it would require incorporating SALT2 light curve fitting inside our model fitting.
+Due to this computational expense, we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$.
 Using this simplification, our latent variables are given by
@@
 \end{align}
-As discussed in Section \ref{sec:challenges}, {\green the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.}
-
-%$\theta$.
-The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$.
-As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$.
-The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1, \hat{c} \rbrace = f(y)$.
-For notational simplicity, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-The model prediction for our derived data is
-%
-%\begin{align}
-%p(\hat{\eta} | \theta) = \int \delta(\hat{\eta} - f(y)) p(y | \theta) \, dy.
-%\end{align}
-%
-%We often consider models that are parametrised as $\theta = \lbrace m_B, x_1, c \rbrace = \eta$, so that the probability of derived data is expressed as
-%
-%\begin{align}
-%p(\hat{\eta}|\eta) = \int \delta(\hat{\eta} - f(y)) p(y|\eta) dy. \label{eq:intsmear}
-%\end{align}
-%
-%It is common in analytic supernova cosmology models to take that
-%
-%\begin{align}
-%p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C}), \label{eq:pop}
-%\end{align}
-%
-%where $\mathcal{C}$ is a covariance matrix reported from light curve fitting.
-
-%Our second way of addressing the issues with under-reported uncertainty and biased summary statistics are to use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean color measured as a function of redshift.
-By comparing simulated true color values to simulated observed color values, we can quantify the mean bias in color in redshift bins, giving us {\green a redshift-dependent array of biases.
-We can then add this into our data-set, and is included in our modelling by allowing a controlled subtraction of this shift from our observed data, where the amount of shift subtracted is dictated by parameter $s_c$.
-Specifically, allow the observed color to shift by $c_{\rm new} = c_{\rm biased} - s_c \times {\rm bias}(z)$, where `bias$(z)$' is the binned mean color bias determined with fiducial simulations.
-Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a global shift in supernovae color across all surveys.
-We place a uniform prior on $s_c$.} With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean color address the non-Gaussianity introduced by chromatic smearing.
-We allow the amount of shift to be parametrised to account for different intrinsic scatter models where parameter $s_c$ controls the fraction of bias subtraction: 0 for no shift, up to 1 for the maximum shift found in simulations.
-The maximum shift was found to using the intrinsic scatter model from \citet{Chotard2011}, which has greater chromatic dispersion than the {\gten} model.
-
-%It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.
-Our fiducial cosmology is a Flat $\Lambda$CDM universe with $\Omega_m = 0.3$ and $H_0 = 70\kmsmpc$.
-
-%{\red MAKE SOME ILLUSTRATIVE PLOT FOR THIS SECTION}
-
- \begin{figure}
-	\begin{tikzpicture}
-	% Obs nodes
-	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
-	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
-	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
-	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
-
-	% Latent nodes
-	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
-	\node[det, above=of zobs] (z) {$z$};
-
-	\node[det, above=of z] (mu) {$\mu$};
-	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
-	% Survey nodes
-	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
-	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
-	% Global nodes
-	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
-
-	% Connect the nodes
-	\edge {sum} {obssum};
-	\edge {z} {zobs};
-
-	\edge {z,cosmo} {mu};
-	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
-
-	% Plates
-	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
-	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
-	\end{tikzpicture}
-	\caption{Probabilistic graphical model for our likelihood.
+As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.
+
+% \begin{figure}
+% 	\begin{tikzpicture}
+% 	% Obs nodes
+% 	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
+% 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
+% 	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
+% 	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
+% 	% Latent nodes
+% 	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
+% 	\node[det, above=of zobs] (z) {$z$};
+% 	\node[det, above=of z] (mu) {$\mu$};
+% 	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
+% 	% Survey nodes
+% 	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
+% 	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
+% 	% Global nodes
+% 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
+% 	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
+% 	% Connect the nodes
+% 	\edge {sum} {obssum};
+% 	\edge {z} {zobs};
+% 	\edge {z,cosmo} {mu};
+% 	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
+% 	% Plates
+% 	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
+% 	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
+% 	\end{tikzpicture}
+% 	\caption{Probabilistic graphical model for our likelihood.
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
@@
 Variables that appear outside both boxes represent top level model parameters.}
-	\label{fig:pgm}
- \end{figure}
+% 	\label{fig:pgm}
+% \end{figure}
 
@@
 \label{sec:underlying}
-The underlying supernova population is characterised by distributions in color, stretch and absolute magnitude.
+
+Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space.
+To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
+
+In addition to absolute magnitude, the underlying supernova population is also characterised by distributions in color and stretch.
 We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey.
 For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution.
-{\green We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude.} Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
+We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude.
+Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
 These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
@@
 Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment.
-We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ {\green that} allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
+We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ that allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
 We sample $\delta_c$ in log space for efficiency in sampling close to zero.
@@
 
-The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey.
+The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$.
+We model the colour skewness and the redshift-dependent means and width of the colour and skew distributions individually for each survey.
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
@@
 We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-{\green We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude.} We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
+We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude.
+We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
 The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
@@
 For the photometric proof-of-concept we provide an accurate probability estimate.
-Further investigation on the effect of inaccurate {\green estimates will be left for} future improvements during the analysis of the DES photometric sample.
+Further investigation on the effect of inaccurate estimates will be left for future improvements during the analysis of the DES photometric sample.
 
@@
 
-With the {\green fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology}.
+With the fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology.
 We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We {\green should also avoid} assuming a specific dispersion model -- either the {\gten} or {\celeven} model, {\green or utilise the results of computing the bias from both models.}
-
-We model the extra dispersion only in color, and do so by adding independent uncertainty on the color observation {\green $\hat{c}$}.
-We note that extra dispersion in magnitude {\green $\hat{m}_B$} (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the color term, hence the requirement for modelling additional color dispersion.
+We should also avoid assuming a specific dispersion model -- either the {\gten} or {\celeven} model, or utilise the results of computing the bias from both models.
+
+We model the extra dispersion only in color, and do so by adding independent uncertainty on the color observation $\hat{c}$.
+We note that extra dispersion in magnitude $\hat{m}_B$ (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the color term, hence the requirement for modelling additional color dispersion.
 Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than color, and causes negligible bias in cosmology.
@@
 As shown in \citep{Kessler2013}, the extra color dispersion shows heavy redshift dependence, increasing with redshift.
-{\green This is an artifact of different filters, however as we may be subject to similar effects in our observational data,} we decide to incorporate redshift dependence in our extra uncertainty.
+This is an artifact of different filters, however as we may be subject to similar effects in our observational data, we decide to incorporate redshift dependence in our extra uncertainty.
 We thus add $\kappa_0 + \kappa_1 z$ to our observed color uncertainty (in quadrature).
@@
 See Section \ref{sec:simdes} for details on simulation generation.
-We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects.
+We found for both models that non-zero values for $\beta_1$ are preferred even with constant $\beta$ used in simulation, due to severe degeneracy with selection effects.
 This degeneracy resulted in a significant bias in recovered cosmology.
@@
 \subsubsection{Host Galaxy Environment}
-
-{\green There are numerous results showing } that host galaxy environment is correlated with supernova properties.
+\label{sec:hostgal}
+
+here are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}.
 The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
@@
 We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.
-With this correction, our calculation of absolute magnitude becomes
-
-\begin{align}
-\mu_{\rm obs} = m_B + \alpha x_1 - \beta c + \Delta M \cdot \hat{m} - M_B. \label{eq:l2}
-\end{align}
 
@@
 We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
-{\green A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.} This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
+A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.
+This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
@@
 One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
-We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are{\green observed and pass cuts}.
+We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
@@
 This shifted Gaussian approximation for color completely removes the unintended bias when simply discarding skewness.
-This shift was not required for the stretch population, and so was left out {\green for the stretch population} for numerical reasons.
+This shift was not required for the stretch population, and so was left out for the stretch population for numerical reasons.
 The impact of this approximation on the calculated efficiency is shown in Figure \ref{fig:shift}, and more detail on this shift and resulting population recovery can be found in Appendix \ref{app:approx}.
@@
 
-The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by {\green taking the ratio of supernovae that are observed and passed cuts over the total number of supernovae generated in that apparent magnitude bin}.
+The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by taking the ratio of supernovae that are observed and passed cuts over the total number of supernovae generated in that apparent magnitude bin.
 That is, we calculate the probability we would include a particular supernova in our sample, divided by the number of such supernovae in our simulated fields.
@@
 
-With the {\green well-sampled redshift} approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
+With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
 For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
@@
 In order to verify our model we run it through stringent tests.
-First, we validate on toy models, verifying that we recover accurate cosmology {\green when generating toy supernovae data constructed to satisfy the assumptions of the BHM construction}.
+First, we validate on toy models, verifying that we recover accurate cosmology when generating toy supernovae data constructed to satisfy the assumptions of the BHM construction.
 We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample, termed the DES-SN3YR sample
@@
 	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
-{\green Calculate the true absolute magnitude using Equation \eqref{eq:l2}.}
+Calculate the true absolute magnitude using Equation \eqref{eq:standard}.
 	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology.
@@
 Repeat from step one until we have a supernova that passes.
-{\green We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.}
+We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.
 	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
+With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.
 
@@
 
-{\green Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model.
+Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model.
 This is a useful consistency check that the model implementation is correct, efficient and free of obvious pathologies.
 However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways.
-To this end, we test using simulations} (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
+To this end, we test using simulations (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
 Simulation specifics can be found in \citet{Kessler18}.
@@
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
+The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation.
+These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.
 
@@
 	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
-We show the average parameter mean and {\green average }standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$.
-The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model, with {\gten} having 70\% achromatic scatter and {\celeven} only having 25\% achromatic scatter.
-A measure of the significance of the parameter bias can be calculated by comparing the bias to a tenth of the scatter (as our Monte-Carlo estimate uncertainty is $\sqrt{100}$ of the scatter).
-From this, we can see that most biases are detected with high statistical significance due to the large number of simulations tested against.}
+We show the average parameter mean and average standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$.
+The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model.}
 	\label{tab:standardisation}
@@
 The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets.
-The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.
-{\green The {\gten} model is consistent with $w=-1$, whilst the {\celeven} model show evidence of bias on $w$, scattering high.} However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
-The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
-We also note that the simulations do not vary cosmological parameters nor population.
-As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.}
+The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.}
 	\label{tab:bulk_summary}
@@
 
-%		$\delta [ \Delta t ]$ &   -0.01 &      0.00     \\
-%		$\delta [ \Delta y ]$ &   -0.03 &     -0.04     \\
-%		$\delta [ \Delta Z ]$ &    0.10 &      0.10     \\
-%		$\delta [ \Delta A ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta u ]$ &   -0.04 &     -0.03     \\
-%		$\delta [ \Delta v ]$ &    0.01 &      0.02     \\
-%		$\delta [ \Delta \lambda_{t} ]$ &    0.00 &      0.02     \\
-%		$\delta [ \Delta \lambda_{y} ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta \lambda_{z} ]$ &   -0.05 &     -0.04     \\
-%		$\delta [ \Delta \lambda_{A} ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{u} ]$ &   -0.02 &     -0.02     \\
-%		$\delta [ \Delta \lambda_{v} ]$ &   -0.01 &      0.01     \\
-%		$\delta [ \Delta gl ]$ &   -0.12 &     -0.13     \\
-%		$\delta [ \Delta hm ]$ &    0.02 &      0.01     \\
-%		$\delta [ \Delta in ]$ &    0.08 &      0.13     \\
-%		$\delta [ \Delta jo ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{gl} ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta \lambda_{hm} ]$ &   -0.01 &     -0.03     \\
-%		$\delta [ \Delta \lambda_{in} ]$ &   -0.10 &     -0.12     \\
-%		$\delta [ \Delta \lambda_{jo} ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta p ]$ &   -0.01 &     -0.01     \\
-%		$\delta [ \Delta q ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta R ]$ &    0.00 &      0.01     \\
-%		$\delta [ \Delta s ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{p} ]$ &    0.00 &      0.01     \\
-%		$\delta [ \Delta \lambda_{q} ]$ &    0.01 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{r} ]$ &   -0.01 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{s} ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta b ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta c ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta d ]$ &    0.00 &      0.01     \\
-%		$\delta [ \Delta e ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta \lambda_{b} ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{c} ]$ &    0.00 &      0.00     \\
-%		$\delta [ \Delta \lambda_{d} ]$ &    0.00 &     -0.01     \\
-%		$\delta [ \Delta \lambda_{e} ]$ &    0.00 &     -0.01     \\
-%		$\Delta_{0}$ &    0.00 &      0.01     \\
-%		$\Delta_{1}$ &    0.00 &     -0.01     \\
-%		$\Delta_{2}$ &    0.00 &      0.00     \\
-%		$\Delta_{3}$ &   -0.01 &     -0.01     \\
-%		$\Delta_{4}$ &    0.00 &     -0.01     \\
-%		$\Delta_{5}$ &    0.00 &     -0.01     \\
-%		$\Delta_{6}$ &    0.00 &      0.00     \\
-%		$\Delta_{7}$ &    0.00 &     -0.01     \\
-
 \begin{figure*}
@@
 We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
+Whilst $\langle M_B \rangle$ is generally thought to be a nuisance parameter, we find cosmological correlation.
+We note that, by fixing $H_0$ in our distance modulus calculation, $\langle M_B \rangle$ absorbs any cosmological uncertainty on this term.
+Additionally $\langle M_B \rangle$ also effects the selection efficiency, which was computed from simulations with a fixed $M_B$ value, introducing a second plausible source of correlation.
 Also visible in this figure are several other interesting relationships.
@@
 
-Each realisation of simulated SN~Ia light curves contains {\green the SALT2 light-curve fits and redshifts} to 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the DES-SN3YR sample.
+Each realisation of simulated SN~Ia light curves contains the SALT2 light-curve fits and redshifts to 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the DES-SN3YR sample.
 As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
@@
 The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  and secondary parameters are shown in Table \ref{tab:standardisation}.
-Our recovered results for $w$ are all within $0.5\sigma$ of the true values of $w=-1$.
+
+Table \ref{tab:bulk_summary} shows that the {\gten} model is consistent with $w=-1$, whilst the {\celeven} model show evidence of bias on $w$, scattering high.
+However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
+The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
+We also note that the simulations do not vary cosmological parameters nor population.
+As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.
+
 Table \ref{tab:standardisation} shows a clear difference in both $\beta$ and $\sigma_{m_B}$ across the {\gten} and {\celeven} simulations.
 As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations.
-The extra smearing of the {\celeven} {\green model does not result in a significantly biased $\beta$ value compared to the average uncertainty on $\beta$}, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
+The extra smearing of the {\celeven}  model does not result in a significantly biased $\beta$ value compared to the average uncertainty on $\beta$, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
 Interestingly, $w$-bias is only found for the {\celeven} simulations.
-
-We investigate the cosmological bias and find its source to be a bias in the observed summary statistics {\green(i.e.
-the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting)}, in addition to incorrect reported uncertainty on the summary statistics.
+A measure of the significance of the parameter bias can be calculated by comparing the bias to a tenth of the scatter (as our Monte-Carlo estimate uncertainty is $\sqrt{100}$ of the scatter).
+From this, we can see that most biases are detected with high statistical significance due to the large number of simulations tested against.
+
+We investigate the cosmological bias and find its source to be a bias in the observed summary statistics (i.e.
+the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting), in addition to incorrect reported uncertainty on the summary statistics.
 To confirm this, we run two tests.
@@
 With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test aims to {\green test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
-To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter,} and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
+To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
@@
 Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies --- incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
+
+Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters.
+Tests when fixing the population parameters to true does not resolve the cosmological bias observed in the {\celeven} simulations.
+As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
 
@@
 BBC+CosmoMC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017} and is being used in the primary analysis of the DES-SN3YR sample \citep{DESKEY}.
-{\green The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected distances are fit using CosmoMC \citep{lewis2002cosmomc}.
-For shorthand, we refer to this combined process as the BBC method hereafter in this paper, as we are concerned with the results of cosmological parameter inference.} As a leading supernova cosmology method, it provides a good consistency{\green check} as to the current levels of accuracy in recovering cosmological parameters.
-
-{\green To this end, we take the results of the BBC method which were also run on the} same set of 200 validation simulations and compare the recovered $w$ values to those of our method.
+The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected distances are fit using CosmoMC \citep{lewis2002cosmomc}.
+For shorthand, we refer to this combined process as the BBC method hereafter in this paper, as we are concerned with the results of cosmological parameter inference.
+As a leading supernova cosmology method, it provides a good consistency check as to the current levels of accuracy in recovering cosmological parameters.
+
+To this end, we take the results of the BBC method which were also run on the same set of 200 validation simulations and compare the recovered $w$ values to those of our method.
 The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
@@
 As such, we expect the BBC method to have a $w$-bias in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
-{\green These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}.
+These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}.
 Both BBC method and {\steve} are sensitive to the intrinsic scatter model, finding differences of $\sim0.066$ and $0.053$ respectively in $w$ when varying the scatter model.
-The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$}.
+The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
 Our method shows a small improvement in the insensitivity to the intrinsic scatter model (having a decrease in difference in $w$ between the {\gten} and {\celeven} models), finding no bias for {\gten} but a $w$ biased high for {\celeven}.
@@
 
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
 \end{document}

### 2018-10-14 papers/methods/figures/tikz.tex
Sooo finally got emulatapj working, now I need to rejig all the tables

\documentclass{article}
% main document, called main.tex
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
	\node[det, above=of zobs] (z) {$z$};
	\node[det, above=of z] (mu) {$\mu$};
	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
	% Survey nodes
	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
	% Global nodes
	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
	% Connect the nodes
	\edge {sum} {obssum};
	\edge {z} {zobs};
	\edge {z,cosmo} {mu};
	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
	% Plates
	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

### 2018-10-14 papers/methods/paper.tex
Sooo finally got emulatapj working, now I need to rejig all the tables

@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{emulateapj}
+\documentclass[a4paper,fleqn,usenatbib,manuscript]{emulateapj}
 
@@
 \usepackage{amssymb}	% Extra maths symbols
-\usepackage{tikz}
 \usepackage{enumerate}
+\usepackage{xcolor}
+
 %\usepackage{tabu}
@@
 \DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
-\usetikzlibrary{bayesnet}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\shorttitle{\name}
+\shorttitle{{\steve}}
 % Don't change these lines
 \begin{document}
-\title{\name: A hierarchical Bayesian model for Supernova Cosmology}
-
- The list of authors, and the short list which is used in the headers.
+\title{{\steve}: A hierarchical Bayesian model for Supernova Cosmology}
+
  % If you need two or more lines of authors, add an extra line using \newauthor
@@
 
+\keywords{cosmology: supernovae}
+
 %%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
@@
 
-\begin{align}
+\begin{equation}
 \mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
 \chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{eqnarray}
 \mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
@@
 E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
-\end{align}
+\end{eqnarray}
 
@@
 Using this simplification, our latent variables are given by
-\begin{align}
+\begin{equation}
 p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
-\end{align}
+\end{equation}
 As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.
 
-% \begin{figure}
-% 	\begin{tikzpicture}
-% 	% Obs nodes
-% 	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
-% 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
-% 	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
-% 	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
-% 	% Latent nodes
-% 	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
-% 	\node[det, above=of zobs] (z) {$z$};
-% 	\node[det, above=of z] (mu) {$\mu$};
-% 	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
-% 	% Survey nodes
-% 	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
-% 	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
-% 	% Global nodes
-% 	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-% 	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
-% 	% Connect the nodes
-% 	\edge {sum} {obssum};
-% 	\edge {z} {zobs};
-% 	\edge {z,cosmo} {mu};
-% 	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
-% 	% Plates
-% 	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
-% 	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
-% 	\end{tikzpicture}
-% 	\caption{Probabilistic graphical model for our likelihood.
+ \begin{figure}
+ 	\begin{center}
+ 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
+ 	\end{center}
+ 	\caption{Probabilistic graphical model for our likelihood.
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
@@
 Variables that appear outside both boxes represent top level model parameters.}
-% 	\label{fig:pgm}
-% \end{figure}
+ 	\label{fig:pgm}
+ \end{figure}
 
@@
 
-\begin{align}
+\begin{equation}
 P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
 \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
 \Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
 P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right).
 \label{eq:l3}
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
-d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
-\end{align}
+\begin{equation}
+d = \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{eqnarray}
 m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
 \sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
-\end{align}
+\end{eqnarray}
 
@@
 
-\begin{align}
+\begin{equation}
 P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{equation}
 P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
-\end{align}
+\end{equation}
 
@@
 
-\begin{align}
+\begin{eqnarray}
 d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
@@
 &\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
-\end{align}
+\end{eqnarray}
 
@@
 When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
-\begin{align}
+\begin{equation}
 P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
-\end{align}
+\end{equation}
 which gives the probability of the model parameter values ($\theta$) given the data.
@@
 To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
-\begin{align}
-\mathcal{L} &= P({\rm data} | \theta, S). \label{eq:app_like}
-\end{align}
+\begin{equation}
+\mathcal{L} = P({\rm data} | \theta, S). \label{eq:app_like}
+\end{equation}
 A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.
@@
 We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
-\begin{align}
+\begin{eqnarray}
 P({\rm data} | S , \theta) P(S, \theta) &= P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
@@
 &= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
-\end{align}
+\end{eqnarray}
 which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
-\begin{align}
+\begin{eqnarray}
 \mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
 \mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
-\end{align}
+\end{eqnarray}
 where we define the denominator as $d$ for simplicity in future derivations.
@@
 Our denominator then becomes
-\begin{align}
-d &= \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
-\end{align}
+\begin{equation}
+d = \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
+\end{equation}
 where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly.
@@
 We can thus state
-\begin{align}
+\begin{equation}
 d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
-\end{align}
+\end{equation}
 By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
-\begin{align}
+\begin{equation}
 \int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
-\end{align}
+\end{equation}
 for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.
@@
 As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
-\begin{align}
+\begin{eqnarray}
 m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
 \sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
-\end{align}
+\end{eqnarray}
 Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.
@@
 With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
-\begin{align}
+\begin{eqnarray}
 \mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
 \sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
-\end{align}
+\end{eqnarray}
 where we highlight that $\mu$ here represents the mean of the distribution, not distance modulus.
@@
 Considering a $3\times3$ matrix with cholesky decomposition
-\begin{align}
+\begin{equation}
 L = \begin{pmatrix}
@@
 \end{pmatrix},
-\end{align}
+\end{equation}
 the original covariance matrix $\Sigma$ is given by
-\begin{align}
+\begin{equation}
 \Sigma = \begin{pmatrix}
@@
 \end{pmatrix}.
-\end{align}
+\end{equation}
 Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered color parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$.
@@
 To take our covariance and add on the diagonal uncertainty for color an extra $\sigma_e$ term, we get
-\begin{align}
+\begin{equation}
 C = \begin{pmatrix}
@@
 \end{pmatrix}.
-\end{align}
+\end{equation}
 The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
-\begin{align}
+\begin{equation}
 L = \begin{pmatrix}
@@
 \end{pmatrix},
-\end{align}
+\end{equation}
 where $g = \sqrt{f^2 + \sigma_e^2} - f$.

### 2018-10-15 papers/methods/paper.tex
emulateapj tables and abstract comments

@@
 \linenumbers
-
+\renewcommand{\arraystretch}{1.2}
 \graphicspath{ {./figures/} }
@@
 
- % If you need two or more lines of authors, add an extra line using \newauthor
 \author{
- 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
- 	T.~M.~Davis\altaffilmark{1,2},
+	S.~R.~Hinton\altaffilmark{1},
+	T.~M.~Davis\altaffilmark{1},
 	A.~G.~Kim\altaffilmark{3},
- 	A.~M\"oller\altaffilmark{2,4},
- 	M.~Sako\altaffilmark{5},
- 	M.~Smith\altaffilmark{6},
- 	\\}
+	R.~Kessler\altaffilmark{2},
+	C.~Lidman\altaffilmark{4},
+	A.~M\"oller\altaffilmark{5,4},
+	M.~Sako\altaffilmark{6},
+	D.~Scolnic\altaffilmark{2},
+	M.~Smith\altaffilmark{7} 
+	\\ \vspace{0.2cm} (DES Collaboration) \\	
+}
+
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
- \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affil{$^{2}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
 \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
 \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
- \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
- \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\affil{$^{5}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affil{$^{6}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\affil{$^{7}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
 
@@
 \begin{abstract}
-We present a new Bayesian hierarchical model named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
+We present a new Bayesian hierarchical model (BHM) named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
 %
-This advances previous works with Bayesian hierarchical models by including 
+This advances previous works by including 
 an improved treatment of Malmquist bias, 
@@
 %
-This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model.
+This is the most rigorous test of a BHM method applied to SN~Ia cosmology fitting, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model.
 This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
 %
-These simulation test results are a concern for BHM applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
+These simulation test results are a concern for BHM cosmology fitting applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
 \end{abstract}
@@
 \begin{eqnarray}
-\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
-d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
-E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
+\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
+d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
+E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
 \end{eqnarray}
@@
 
-\begin{table}
-	\centering
-	\caption{ Parameters defined in our model and a summary of their use.
-The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova.
-The bottom section shows the observables (not parameters) that are the input data to the model.}
-	\label{tab:param_summary}
-	\begin{tabular}{ll}
-		\hline
-		Parameter & Description \\
-		\hline
+\begin{deluxetable}{ll}
+	\tablecolumns{2}
+	\tablewidth{0.99\columnwidth}
+	\tablecaption{Model Parameters}
+	\tablehead { \colhead{Parameter} & \colhead{Description} }
+	\startdata
+
+	&\textbf{Global Parameters} \\
+
 	$\Omega_m$  & Matter density  \\
@@
 	$\langle M_B \rangle$ & Mean absolute magnitude \\
-		%$s_c$ & color bias\\
-		\hline
+	\\
+	& \textbf{Survey Parameters} \\
+
 	$\delta S$ & Selection effect deviation \\
@@
 	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
-		\hline
+	\\
+	& \textbf{Supernova Parameters} \\
 	$m_B$ & True flux\\
@@
 	$\mu$ & Derived distance modulus \\
-		\hline
+	\\
+	& \textbf{Input Data}\textsuperscript{a} \\
+
 	$\hat{m}_B$ & Measured flux\\
@@
 	$\hat{p}$ & Determined Ia probability \\
-		\hline
-	\end{tabular}
-\end{table}
+	\enddata
+	\tablenotetext{a}{Not model parameters but shown for completeness.}
+\label{tab:param_summary}
+\end{deluxetable}
 
@@
 \begin{eqnarray}
-m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
-\sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
+m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
+\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
 \end{eqnarray}
@@
 \begin{eqnarray}
-d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
-d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
-&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
+d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
+d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
+&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
 \end{eqnarray}
@@
 
-\begin{table}
-	\centering
-	\caption{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data.
+\begin{deluxetable}{lcc}
+	\tablecolumns{3}
+	\tablecaption{Cosmological Parameters for toy supernova data}
+	\tablehead {
+		\colhead {Model}              &
+		\colhead {$\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}  &
+		\colhead{$w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}
+	}
+
+	\startdata
+		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
+		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
+	\enddata
+	\label{tab:simple_model}
+	\tablecomments{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data.
 As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
@@
 The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets.
-Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.
-As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted standard deviation: $\sqrt{100} = 10$.
-As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
-For the Flat $w$CDM model, the value of $w$ is reported with a prior on $\Omega_m$ of $\mathcal{N}(0.3, 0.01)$.}
-	\label{tab:simple_model}
-	\begin{tabular}{l|cc}
-		\hline
-		Model & $\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) & $w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) \\ 
-		\hline
-		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
-		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
-		\hline
-	\end{tabular}
-\end{table}
+Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.}
+\end{deluxetable}
 
@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
+As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted scatter.
+The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
 With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.
@@
 
-\begin{table}
-	\centering
-	\caption{Tested population distributions, where the {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
+\begin{deluxetable}{lcccc}
+\tablewidth{\columnwidth}
+
+\tablecolumns{5}
+\tablecaption{Supernova population distributions}
+\tablehead {
+	\colhead {Model} &
+	\colhead { $\langle x_1 \rangle$  } &
+	\colhead { $\sigma_{x_1}$ } &
+	\colhead { $\langle c \rangle$ } &
+	\colhead { $\sigma_c$ } 
+}
+	\startdata
+		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\ \\
+		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$  \\
+	\enddata
 	\label{tab:dist}
-	\begin{tabular}{l|cccc}
-		\hline
-		Model & $\langle x_1 \rangle$ & $\sigma_{x_1}$ &  $\langle c \rangle$ & $\sigma_c$  \\
-		\hline
-		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
-		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
-		\hline
-	\end{tabular}
-\end{table}
+	\tablecomments{The {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
+\end{deluxetable}
 
@@
 
-\begin{table*}
-	\centering
-	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
+\begin{deluxetable*}{lrrrrr}
+	\tablewidth{\textwidth}
+	\tablecolumns{6}
+	\tablecaption{Realistic simulation standardisation parameters}
+	\tablecomments{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
 We show the average parameter mean and average standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$.
 The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model.}
-	\label{tab:standardisation}
-	\begin{tabular}{l|rrrrr}
-		\hline
-		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$ \\
-		\hline
+	\tablehead{
+		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$	
+	}
+	\startdata
 		{\gten} Stat + Syst  & $0.022 \ [0.009\ (0.008)]$ & $ 0.34 \ [0.19\ (0.18)]$ & $-0.002 \ [0.028\ (0.015)]$ & $0.070 \ [0.022\ (0.018)]$ & $0.073 \ [0.025\ (0.022)]$\\
@@
 		{\celeven} Stat + Syst  &  $0.002 \ [0.009\ (0.007)]$ & $-0.04 \ [0.15\ (0.13)]$ & $0.014 \ [0.030\ (0.018)]$ & $0.024 \ [0.016\ (0.011)]$ & $0.029 \ [0.020\ (0.014)]$ \\
-		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$\\
-		\hline
-	\end{tabular}
-\end{table*}
+		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$ \\
+
+	\enddata
+	\label{tab:standardisation}
+\end{deluxetable*}
 
@@
 
-\begin{table}
-	\centering
-	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
+\begin{deluxetable}{lrr}
+	\tablecaption{Realistic simulation determination of $w$}
+	\tablewidth{\columnwidth}
+	\tablecomments{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
 The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets.
 The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.}
-	\label{tab:bulk_summary}
-	\begin{tabular}{l|rr}
-		\hline
-		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias \\
-		\hline
+	\tablehead{		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias		}
+	\startdata
 		{\gten} Stat + Syst  &  $-0.998 \ [0.097\ (0.073)]$ & $(0.02\pm0.07)\sigma$  \\
@@
 		{\celeven} Stat         &  $-0.948\ [0.079\ (0.066)]$ & $(0.65\pm0.08)\sigma$   \\
-		\hline
-	\end{tabular}
-\end{table}
-
-\begin{table}
-	\centering
-	\caption{Reduced parameter correlations with $w$ for the combined 100 simulation fits.
+	\enddata
+		\label{tab:bulk_summary}
+\end{deluxetable}
+
+\begin{deluxetable}{crr}
+
+	\tablecaption{Reduced parameter correlations with $w$.}
+	\tablecomments{Correlations determined from the combined 100 simulation fits.
 Correlations for the low-$z$ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.
 Zero superscripts indicate the Dark Energy Survey, and a superscript one represents the low-$z$ survey.}
-	\label{tab:parameter_correlations}
-	\begin{tabular}{c|rr}
-		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst  \\
-		\hline
+	\tablehead{		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst }
+	\startdata
 		$\Omega_m$ &   $-$0.19 &     $-$0.21     \\
@@
 		$\langle M_B \rangle$ &    0.68 &      0.66     \\
+		$\delta(0)$ &    0.00 &      0.00     \\
+		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
+		\\
 		$\sigma_{\rm m_B}^{0}$ &    0.04 &      0.07     \\
@@
 		$\kappa_{c1}^{1}$ &   $-$0.01 &      0.01     \\
-		%$s_c^{0}$ &    0.00 &      0.01     \\
-		$\delta(0)$ &    0.00 &      0.00     \\
-		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
 		$\langle x_1^{0} \rangle$ &   $-$0.01 &     $-$0.05     \\
@@
 		$\langle c^{7} \rangle$ &   $-$0.12 &     $-$0.06     \\
+		\\
 		$\delta [ {\rm SALT}_0 ]$ &    0.05 &      0.05     \\
@@
 		$\delta [ \Delta \lambda_z ]$ &    0.09 &      0.07     \\
-		\hline
-	\end{tabular}
-\end{table}
+	\enddata
+	\label{tab:parameter_correlations}
+\end{deluxetable}
 
@@
 
-\begin{table*}
-	\centering
-	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features.
+\begin{deluxetable*}{llrr}
+	\tablecaption{$w$ error budget}
+	\tablecomments{Error budget determined from analysing uncertainty on simulation data whilst progressively enabling model features.
 We start from the top of the table, only varying cosmological parameters $\Omega_m$ and $w$, and then progressively unlock parameters and let them fit as we progress down the table.
 The cumulative uncertainty shows the total uncertainty on $w$ on the fit for all, where the $\sigma_w$ term is derived by taking the quadrature difference in cumulative uncertainty as we progress.}
-	\label{tab:uncert}
-	\begin{tabular}{l|l|rr}
-		Feature &  Parameters & $\sigma_w$ & Cumulative  \\
-		\hline
+	\tablehead{		Feature &  Parameters & $\sigma_w$ & Cumulative }
+	\startdata
 		Cosmology                      & $\Omega_m$, $w$ & 0.051  &   0.051       \\
@@
 		Systematics                    & $\delta \mathcal{Z}_i$, $\delta S$ & 0.054  &   0.096      \\ 
-		\hline
-	\end{tabular}
-\end{table*}
+	\enddata
+	\label{tab:uncert}
+\end{deluxetable*}
 
@@
 
-\begin{table}
-	\centering
-	\caption{Characterising the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model.
+\begin{deluxetable}{lccc}
+	\tablecaption{$w$ bias comparison}
+	\tablecomments{We characterise the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model.
 We also show the results when combining the {\gten} and {\celeven} models into a combined set of 200 simulations.
@@
 The final row shows the scatter between Steve and BBC for the different simulations.}
-	\label{tab:bulk_summary_bbc}
-	\begin{tabular}{l|ccc}
-		\hline
-		Model & {\gten} & {\celeven} & ({\gten} + {\celeven}) \\
-		\hline
+	\tablehead{		Model & {\gten} & {\celeven} & ({\gten} + {\celeven}) }
+	\startdata
+
 		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
@@
 		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
-		\hline
+
 		$\Delta$  $\sigma_w$           & $0.057\pm0.004$ & $0.062\pm0.004$ & $0.060\pm0.003$ \\
-		\hline
-	\end{tabular}
-\end{table}
+
+	\enddata
+	\label{tab:bulk_summary_bbc}
+\end{deluxetable}
 \footnotetext{This value is computed with each simulation having the same weight.
@@
 \begin{eqnarray}
-P({\rm data} | S , \theta) P(S, \theta) &= P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
-P({\rm data} | S , \theta) &= \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
-&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
-&= \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
+P({\rm data} | S , \theta) P(S, \theta) &=& P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
+P({\rm data} | S , \theta) &=& \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
+&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
+&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathcal{L}&= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-\mathcal{L} &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
+\mathcal{L}&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
+\mathcal{L} &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
 \end{eqnarray}
@@
 \begin{eqnarray}
-m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
-\sigma^*_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
+m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
+\sigma^*_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mu_1 &= \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
-\sigma_1 &= \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
+\mu_1 &=& \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
+\sigma_1 &=& \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
 \end{eqnarray}

### 2018-10-15 papers/methods/paper.tex
Some more updates from ChrisL

@@
 Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty.
-Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses all information about correlation between sources of uncertainty.
+Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses information about correlation between sources of uncertainty.
 For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations.
-Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods which can automatically capture correlated uncertainties are preferable.
+Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods that can automatically capture correlated uncertainties are preferable.
 Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
 As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion on every step.
-Whilst binning data results in some loss of information, recent works tested against simulations show that this loss of information does not create significant cosmological biases \citep{Scolnic2016, Kessler2017}.
+A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion at every step.
+Whilst binning data results in some loss of information, recent works tested against simulations show that this loss does not result in significant cosmological biases \citep{Scolnic2016, Kessler2017}.
 
@@
 Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects associated with supernova surveys.
-By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied; we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}.
+By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied, and we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}.
 This freedom comes with a cost -- computation.
@@
 \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting.
-\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias correction.
+\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias corrections.
 \citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations.
@@
 
-Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias.
+Whilst there are a large number of hierarchical models available, none of them have undergone comprehensive tests using realistic simulations to verify each models' respective bias.
 Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
@@
 Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset.
-However, each different survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
+However, each survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
 Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
@@
 
-here are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}.
+There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}.
 The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
@@
 \begin{equation}
-P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right).
+P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right).
 \label{eq:l3}
@@
 The unshifted normal approximation shows significant discrepancy in the calculated efficiency as it transitions from 1 to 0, whilst the shifted normal approximation shows negligible error to the correct solution.
-From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) as unnecessary.}
+From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) are unnecessary.}
 	\label{fig:shift}
@@
 
-The selection functions parameters (a skew normal for low-redshift and an error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
+The selection functions parameters (a skew normal for low-redshift and a complimentary error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
 Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample.
@@
 \begin{equation}
-\mathcal{L} = P({\rm data} | \theta, S). \label{eq:app_like}
+\mathcal{L}(\theta; {\rm data}) = P({\rm data} | \theta, S). \label{eq:app_like}
 \end{equation}
@@
 \begin{eqnarray}
-\mathcal{L}&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-\mathcal{L} &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
+\mathcal{L}(\theta; {\rm data})&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
+\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
 \end{eqnarray}
@@
 Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.
-Based on these results and the fantastic performance in correctly recovering underlying populations of the shifted normal approximation, we adopt the shifted normal approximation in our model.}
+Based on these results and the good performance in correctly recovering underlying populations of the shifted normal approximation, we adopt the shifted normal approximation in our model.}
 	\label{fig:simple_w_super}
@@
 
-The simplest and most commonly found optimisation we employed was to precompute as much as possible to reduce the complexity of the mathematical graph our model is translated into to compute the surface derivatives.
+The simplest and most commonly found optimisation we employed was to precompute as much as possible.
+This is in a bid to reduce the complexity of the mathematical graph our model is translated into by Stan, to simplify the computation of surface derivatives.
 For example, when computing the distance modulus, redshift is encountered to various powers.
@@
 The primary numerical improvement we made on existing frameworks was to remove costly probability evaluations of multivariate normals.
-To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, you sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$.
+To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, we sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$.
 In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution.

### 2018-10-16 papers/methods/paper.tex
Tablemark
Fixing the table
Updating authors

@@
 	T.~M.~Davis\altaffilmark{1},
-	A.~G.~Kim\altaffilmark{3},
-	R.~Kessler\altaffilmark{2},
-	C.~Lidman\altaffilmark{4},
-	A.~M\"oller\altaffilmark{5,4},
-	M.~Sako\altaffilmark{6},
-	D.~Scolnic\altaffilmark{2},
-	M.~Smith\altaffilmark{7} 
+	A.~G.~Kim\altaffilmark{2},
+	D.~Brout\altaffilmark{3},
+	R.~Kessler\altaffilmark{4},
+	C.~Lidman\altaffilmark{5},
+	A.~M\"oller\altaffilmark{6,5},
+	M.~Sako\altaffilmark{3},
+	D.~Scolnic\altaffilmark{4},
+	M.~Smith\altaffilmark{7},
+	R.~C.~Wolf\altaffilmark{8}
 	\\ \vspace{0.2cm} (DES Collaboration) \\
@@
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
-\affil{$^{2}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
-\affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
-\affil{$^{5}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
-\affil{$^{6}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\affil{$^{2}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
+\affil{$^{3}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\affil{$^{4}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\affil{$^{5}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\affil{$^{6}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
 \affil{$^{7}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\affil{$^{8}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA}
 
@@
 	\\
-	& \textbf{Input Data}\textsuperscript{a} \\
+	& \textbf{Input Data}\tablenotemark{a} \\
 
@@
 
-\begin{equation}
-P(M_B, x_1, c, z | \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c) = \notag \\
-\mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1}) \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1}.
-\end{equation}
+\begin{eqnarray}
+P(M_B, x_1, c, z | \theta) &=& \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \, \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1})\notag\\ &\quad\quad& \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1},
+\end{eqnarray}
+
+where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.
 
@@
 		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
-		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$\footnotemark\\
+		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$\tablenotemark{a}\\
 		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
@@
 	\enddata
-	\label{tab:bulk_summary_bbc}
-\end{deluxetable}
-\footnotetext{This value is computed with each simulation having the same weight.
+	\tablenotetext{a}{This value is computed with each simulation having the same weight.
 It disagrees with the value provided in \citet[Table 10, row 3]{Brout18SYS} which uses inverse variance weighted averages.
@@
 We note that if inverse variance weighting is applied to both datasets, they both shift by $\Delta w \approx 0.005$, and thus the predicted difference between the BBC method and {\steve} remains the same.}
+	\label{tab:bulk_summary_bbc}
+\end{deluxetable}
+
 \begin{figure}

### 2018-11-01 papers/methods/paper.tex
Updating author list

@@
 	D.~Brout\altaffilmark{3},
-	R.~Kessler\altaffilmark{4},
-	C.~Lidman\altaffilmark{5},
-	A.~M\"oller\altaffilmark{6,5},
+	R.~Kessler\altaffilmark{4,5},
+	J.~Lasker\altaffilmark{4,5},
+	C.~Lidman\altaffilmark{6},
+	A.~M\"oller\altaffilmark{7,6},
 	M.~Sako\altaffilmark{3},
-	D.~Scolnic\altaffilmark{4},
-	M.~Smith\altaffilmark{7},
-	R.~C.~Wolf\altaffilmark{8}
+	D.~Scolnic\altaffilmark{5},
+	M.~Smith\altaffilmark{8},
+	R.~C.~Wolf\altaffilmark{9},
+	Y.~Zhang\altaffilmark{10},
+	M.~March\altaffilmark{3},
+	S.~Allam\altaffilmark{10},
+	J.~Annis\altaffilmark{10},
+	S.~Avila\altaffilmark{11},
+	E.~Bertin\altaffilmark{12,13},
+	D.~Brooks\altaffilmark{14},
+	D.~L.~Burke\altaffilmark{15,16},
+	A.~Carnero~Rosell\altaffilmark{17,18},
+	M.~Carrasco~Kind\altaffilmark{19,20},
+	J.~Carretero\altaffilmark{21},
+	C.~E.~Cunha\altaffilmark{15},
+	L.~N.~da Costa\altaffilmark{18,22},
+	C.~Davis\altaffilmark{15},
+	J.~De~Vicente\altaffilmark{17},
+	D.~L.~DePoy\altaffilmark{23},
+	P.~Doel\altaffilmark{14},
+	T.~F.~Eifler\altaffilmark{24,25},
+	B.~Flaugher\altaffilmark{10},
+	P.~Fosalba\altaffilmark{26,27},
+	J.~Frieman\altaffilmark{10,5},
+	J.~Garc\'ia-Bellido\altaffilmark{28},
+	E.~Gaztanaga\altaffilmark{26,27},
+	D.~W.~Gerdes\altaffilmark{29,30},
+	R.~A.~Gruendl\altaffilmark{19,20},
+	J.~Gschwend\altaffilmark{18,22},
+	G.~Gutierrez\altaffilmark{10},
+	W.~G.~Hartley\altaffilmark{14,31},
+	D.~L.~Hollowood\altaffilmark{32},
+	K.~Honscheid\altaffilmark{33,34},
+	E.~Krause\altaffilmark{24},
+	K.~Kuehn\altaffilmark{35},
+	N.~Kuropatkin\altaffilmark{10},
+	O.~Lahav\altaffilmark{14},
+	M.~Lima\altaffilmark{36,18},
+	M.~A.~G.~Maia\altaffilmark{18,22},
+	J.~L.~Marshall\altaffilmark{23},
+	F.~Menanteau\altaffilmark{19,20},
+	R.~Miquel\altaffilmark{37,21},
+	R.~L.~C.~Ogando\altaffilmark{18,22},
+	A.~A.~Plazas\altaffilmark{25},
+	E.~Sanchez\altaffilmark{17},
+	V.~Scarpine\altaffilmark{10},
+	R.~Schindler\altaffilmark{16},
+	M.~Schubnell\altaffilmark{30},
+	S.~Serrano\altaffilmark{26,27},
+	I.~Sevilla-Noarbe\altaffilmark{17},
+	M.~Soares-Santos\altaffilmark{38},
+	F.~Sobreira\altaffilmark{39,18},
+	E.~Suchyta\altaffilmark{40},
+	G.~Tarle\altaffilmark{30},
+	D.~Thomas\altaffilmark{11},
+	and V.~Vikram\altaffilmark{41}
 	\\ \vspace{0.2cm} (DES Collaboration) \\
@@
 \affil{$^{3}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
-\affil{$^{4}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{5}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
-\affil{$^{6}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
-\affil{$^{7}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
-\affil{$^{8}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA}
+\affil{$^{4}$ Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
+\affil{$^{5}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\affil{$^{6}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\affil{$^{7}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affil{$^{8}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\affil{$^{9}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
+\affil{$^{10}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\affil{$^{11}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\affil{$^{12}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\affil{$^{13}$ Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\affil{$^{14}$ Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\affil{$^{15}$ Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\affil{$^{16}$ SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
+\affil{$^{17}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\affil{$^{18}$ Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affil{$^{19}$ Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affil{$^{20}$ National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\affil{$^{21}$ Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
+\affil{$^{22}$ Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affil{$^{23}$ George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
+\affil{$^{24}$ Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
+\affil{$^{25}$ Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
+\affil{$^{26}$ Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affil{$^{27}$ Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\affil{$^{28}$ Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
+\affil{$^{29}$ Department of Astronomy, University of Michigan, Ann Arbor, MI 48109, USA}
+\affil{$^{30}$ Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\affil{$^{31}$ Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 16, CH-8093 Zurich, Switzerland}
+\affil{$^{32}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
+\affil{$^{33}$ Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
+\affil{$^{34}$ Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
+\affil{$^{35}$ Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia}
+\affil{$^{36}$ Departamento de F\'isica Matem\'atica, Instituto de F\'isica, Universidade de S\~ao Paulo, CP 66318, S\~ao Paulo, SP, 05314-970, Brazil}
+\affil{$^{37}$ Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
+\affil{$^{38}$ Brandeis University, Physics Department, 415 South Street, Waltham MA 02453}
+\affil{$^{39}$ Instituto de F\'isica Gleb Wataghin, Universidade Estadual de Campinas, 13083-859, Campinas, SP, Brazil}
+\affil{$^{40}$ Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
+\affil{$^{41}$ Argonne National Laboratory, 9700 South Cass Avenue, Lemont, IL 60439, USA}
 

### 2018-11-05 papers/methods/paper.tex
Updating authorlist and removing line numbers

@@
 
-\usepackage{lineno}
-\linenumbers
+%\usepackage{lineno}
+%\linenumbers
 \renewcommand{\arraystretch}{1.2}
@@
 D.~Brout\altaffilmark{3},
+C.~B.~D'Andrea\altaffilmark{3},
 R.~Kessler\altaffilmark{4,5},
@@
 C.~Lidman\altaffilmark{6},
-	A.~M\"oller\altaffilmark{7,6},
+E.~Macaulay\altaffilmark{7},
+A.~M\"oller\altaffilmark{8,6},
 M.~Sako\altaffilmark{3},
 D.~Scolnic\altaffilmark{5},
-	M.~Smith\altaffilmark{8},
-	R.~C.~Wolf\altaffilmark{9},
-	Y.~Zhang\altaffilmark{10},
-	M.~March\altaffilmark{3},
-	S.~Allam\altaffilmark{10},
-	J.~Annis\altaffilmark{10},
-	S.~Avila\altaffilmark{11},
+M.~Smith\altaffilmark{9},
+R.~C.~Wolf\altaffilmark{10},
+S.~Allam\altaffilmark{11},
+J.~Annis\altaffilmark{11},
+S.~Avila\altaffilmark{7},
 E.~Bertin\altaffilmark{12,13},
@@
 J.~Carretero\altaffilmark{21},
+M.~Childress\altaffilmark{9},
 C.~E.~Cunha\altaffilmark{15},
@@
 T.~F.~Eifler\altaffilmark{24,25},
-	B.~Flaugher\altaffilmark{10},
+B.~Flaugher\altaffilmark{11},
 P.~Fosalba\altaffilmark{26,27},
-	J.~Frieman\altaffilmark{10,5},
+J.~Frieman\altaffilmark{11,5},
 J.~Garc\'ia-Bellido\altaffilmark{28},
@@
 J.~Gschwend\altaffilmark{18,22},
-	G.~Gutierrez\altaffilmark{10},
+G.~Gutierrez\altaffilmark{11},
 W.~G.~Hartley\altaffilmark{14,31},
@@
 K.~Kuehn\altaffilmark{35},
-	N.~Kuropatkin\altaffilmark{10},
+N.~Kuropatkin\altaffilmark{11},
 O.~Lahav\altaffilmark{14},
@@
 M.~A.~G.~Maia\altaffilmark{18,22},
+M.~March\altaffilmark{3},
 J.~L.~Marshall\altaffilmark{23},
@@
 R.~Miquel\altaffilmark{37,21},
+E.~Morganson\altaffilmark{20},
 R.~L.~C.~Ogando\altaffilmark{18,22},
@@
 E.~Sanchez\altaffilmark{17},
-	V.~Scarpine\altaffilmark{10},
+V.~Scarpine\altaffilmark{11},
 R.~Schindler\altaffilmark{16},
@@
 G.~Tarle\altaffilmark{30},
-	D.~Thomas\altaffilmark{11},
-	and V.~Vikram\altaffilmark{41}
+D.~Thomas\altaffilmark{7},
+V.~Vikram\altaffilmark{41},
+and Y.~Zhang\altaffilmark{11}
 \\ \vspace{0.2cm} (DES Collaboration) \\
@@
 \affil{$^{6}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
-\affil{$^{7}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
-\affil{$^{8}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
-\affil{$^{9}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
-\affil{$^{10}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
-\affil{$^{11}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\affil{$^{7}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\affil{$^{8}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affil{$^{9}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\affil{$^{10}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
+\affil{$^{11}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
 \affil{$^{12}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
@@
 Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:
-
 \begin{equation}
@@
 \end{equation}
-
 where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
@@
 The $\chi^2$ function minimised is
-
 \begin{equation}
@@
 \end{equation}
-
 where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis).
 The predicted $\mu_C$ is defined as
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
@@
 The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.
@@
 In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
-
 \begin{equation}
@@
 \end{equation}
-
 where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift.
@@
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
-
 \begin{equation}
@@
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
-
 \begin{align}
@@
 \end{align}
-
 As our model so far describes components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
-
 \begin{align}
@@
 \end{align}
-
 where the denominator represents an integral over all potential data $D$.
@@
 We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as
-
 \begin{equation}
@@
 \end{equation}
-
 where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$, and $c$ population to $m_B$ given cosmological parameters.
@@
 The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 What then remains is determining the functional form of $P(S|m_B)$.
@@
 Formally, this gives
-
 \begin{equation}
@@
 \end{equation}
-
 where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
@@
 However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
-
 \begin{equation}
@@
 For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 and is incorporated into our likelihood.
@@
 
-% color bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain
-
 Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the low-$z$ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
@@
 	\startdata
-
 		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
@@
 		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
-
 		$\Delta$  $\sigma_w$           & $0.057\pm0.004$ & $0.062\pm0.004$ & $0.060\pm0.003$ \\
-
 	\enddata

### 2018-12-19 papers/methods/figures/tikz.tex
Updating paper for referees comments

@@
 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
-	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
-	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
+	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
+	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
 	% Latent nodes
@@
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
+	\node[det, right=4.33cm of mu] (MBobs) {$M_B$};
 	% Survey nodes
@@
 	\edge {z,cosmo} {mu};
-	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
+	\edge {mu, sum, pop1, ab, sf, pobs, mobs} {MBobs};	
 	% Plates

### 2018-12-19 papers/methods/paper.tex
Updating for arxiv upload
Updating paper for referees comments

@@
 \renewcommand{\arraystretch}{1.2}
-\graphicspath{ {./figures/} }
+%\graphicspath{ {./figures/} }
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement {\blue as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus.
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
@@
 Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space.
-To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
+{\blue This is mathematically equivalent, however allows us to model the underlying population and intrinsic scatter distinctly.} To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
 
@@
 We sample $\delta_c$ in log space for efficiency in sampling close to zero.
-The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors.
+The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors with {\blue mean 0 and width 1}.
 
@@
 
-We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM.
+We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM {\blue (without a flatness prior)}.
 $\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$.

### 2018-12-19 papers/methods/paper.tex
Responding to referee comments

@@
 The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations.
-As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
+As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$ {\blue for each supernova light curve fit}, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
 
@@
 A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.
-This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
+This gives thirteen global systematics shared by {\blue all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey.
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
@@
 We thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty.
-Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
+Formally, we include deviations from the determined mean selection function parameters with parameter vector $\delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
 Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level --- varying cosmology or spectroscopic efficiency.
@@
 and is incorporated into our likelihood.
-This is illustrated in Figure \ref{fig:efficiency}.
+{\blue Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}.
 Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency.
@@
 	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the non-Gaussian nature of the posterior surface.
-The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ {\blue shown to recover input cosmology.
+For contours that are non-Gaussian due to the curved degeneracy between $\Omega_m$ and $w$, the marginalised distributions can provide misleading statistics where maximum marginalised distribution can disagree with the maximum likelihood in multiple dimensions.
+For our contours, the non-Gaussianity is small and the marginalised distributions still provide a valuable metric.} The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
 	\label{fig:simple_w}
@@
 \begin{enumerate}[1.]
-	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
+	\item Draw a redshift from a power law distribution.
+For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.33}$.
+{\blue For the low redshift survey, this is equivalent to sampling $y = z$ from $0.02$ to $\sqrt{0.1}$, for the high redshift survey, this is equivalent to sampling $y = z^2$ from $0.2$ to $1.0$.
+These distributions are arbitrary, and this test has been performed with various flat and power law distributions.}
 	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
 	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
-Calculate the true absolute magnitude using Equation \eqref{eq:standard}.
 	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology.
-Use this to determine the true apparent magnitude of the object $m_B$.
+Use this to determine the true apparent magnitude of the object $m_B$ using $m_B = \mu + M_B - \alpha x_1 + \beta c$.
 	\item Determine if the SN~Ia is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
-As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted scatter.
+As we are using 100 independent realisations, the precision of our determination of the mean simulation result is {\blue approximately a tenth of the quoted scatter (as a degree of non-Gaussianity of our fits will make this relationship inexact)}.
 The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
@@
 Simulation specifics can be found in \citet{Kessler18}.
-
-Additionally, prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
+{\blue The primary difference between the toy data of the previous section are the different underlying colour and stretch, inclusion of spectroscopic data cuts and light curve cuts, and inclusion of intrinsic dispersion models.}
+
+{\blue The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
@@
 We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models.
+{\blue Uncertainty on the analytic selection function is incorporated into our fits, mitigating the imperfection of our analytic form by allowing it to vary in our fits.}
 
@@
 Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
-To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+{\blue That is, the first test ascertained that biases in the summary statistics are the cause of cosmological bias.
+It is thus important to determine the source of those biases; whether they are from intrinsic scatter model or another aspect of the simulation.
+To} this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter.
+{\blue We find $w=-1.00$, showing} that the source of the biases in summary statistics is the underlying intrinsic scatter model.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
@@
 Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters.
-Tests when fixing the population parameters to true does not resolve the cosmological bias observed in the {\celeven} simulations.
-As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
+{\blue To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation.
+These fits did not show reduced cosmological bias in the {\celeven} simulations, and thus the determination of underlying population in our model may give biased distributions, however does not impact cosmology.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
 
@@
 With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
-To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
+{\blue This increased uncertainty is one of the strengths of hierarchical models as it represents a more thorough accounting of model uncertainty.} To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
 By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.

### 2018-12-20 papers/methods/misc/author_list.tex
Adding author list

--- 2018-11-05 papers/methods/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib,manuscript]{emulateapj}
+% Author list file generated with: mkauthlist.py 1.2.4 
+% mkauthlist.py -f --doc -j emulateapj --sort --sort-builder -a order.csv author_list.csv author_list.tex
 
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-\usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
+\documentclass[iop]{emulateapj}
 
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
-\usepackage[T1]{fontenc}
-\usepackage{ae,aecompl}
+\begin{document}
+\title{DES Publication Title}
 
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
-\usepackage{enumerate}
-\usepackage{xcolor}
-
-%\usepackage{tabu}
-%\usepackage{tabularx}
-\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\name}{Steve}
-\newcommand{\myemail}{samuelreay@gmail.com}
-\newcommand\abs[1]{\left|#1\right|}
-\newcommand {\etal} {\emph{~et~al.} }
-\definecolor{forestgreen}{rgb}{0.13, 0.75, 0.13}
-\newcommand{\green}{\color{forestgreen}}
-\newcommand{\blue}{\color{blue}}
-\newcommand{\red}{\color{red}}
-
-\newcommand{\kmsmpc}{{\rm km}\,{\rm s}^{-1}\,{\rm Mpc}^{-1}}
-\newcommand{\cov}{\mathcal{C}_\eta}
-\newcommand{\Z}{\mathcal{Z}}
-
-\defcitealias{Guy2010}{G10}
-\defcitealias{Chotard2011}{C11}
-\defcitealias{Rubin2015}{R15}	
-\defcitealias{Scolnic2016}{SK16}	 
-\newcommand{\gten}{\citetalias{Guy2010}}
-\newcommand{\steve}{\textit{Steve}}
-\newcommand{\celeven}{\citetalias{Chotard2011}}
-\newcommand{\rubin}{\citetalias{Rubin2015}}
-\newcommand{\sk}{\citetalias{Scolnic2016}}
-\newcommand{\angstrom}{\mbox{\normalfont\AA}}
-
-%\usepackage{lineno}
-%\linenumbers
-\renewcommand{\arraystretch}{1.2}
-\graphicspath{ {./figures/} }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-
-\shorttitle{{\steve}}
-% Don't change these lines
-\begin{document}
-\title{{\steve}: A hierarchical Bayesian model for Supernova Cosmology}
+\def\andname{}
 
@@
 
-% Abstract of the paper
 \begin{abstract}
-We present a new Bayesian hierarchical model (BHM) named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
-%
-This advances previous works by including 
-an improved treatment of Malmquist bias, 
-accounting for additional sources of systematic uncertainty, 
-and increasing numerical efficiency. 
-%
-Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties.
-Selection effects are characterised using Monte-Carlo simulations.
-%
-We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
-%
-Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
-%
-These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
-%
-This is the most rigorous test of a BHM method applied to SN~Ia cosmology fitting, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model.
-This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
-%
-These simulation test results are a concern for BHM cosmology fitting applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
+This is a sample document created by \texttt{mkauthlist.py v1.2.4}.
 \end{abstract}
-
-\keywords{cosmology: supernovae}
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-
-Two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}.
-Since that time, the number of observed type Ia supernovae (SN~Ia) has increased by more than an order of magnitude, with contributions from modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, chambers2016panstarrs, sako2018sdss}.
-Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014, Scolnic2017} have been combined with complementary probes of large scale structure and the CMB.
-For a recent review, see \citet{Huterer2018}.
-While these efforts have reduced the uncertainty on the equation-of-state of dark energy ($w$) by more than a factor of two, it is still consistent with a cosmological constant and the nature of dark energy remains an unsolved mystery.
-
-In attempts to tease out the nature of dark energy, active and planned surveys are continually growing in size and scale.
-The Dark Energy Survey \citep[DES,][]{Bernstein2012, Abbott2016} has discovered thousands of type Ia supernovae, attaining both spectroscopically and photometrically identified samples.
-The Large Synoptic Survey Telescope \citep[LSST,][]{Ivezic2008, LSSTScienceCollaboration2009} will discover tens of thousands of photometrically classified supernovae.
-Such increased statistical power demands greater fidelity and flexibility in modelling supernovae for cosmological purposes, as we will require reduced systematic uncertainties to fully utilise these increased statistics \citep{Betoule2014, Scolnic2017}.
-
-As such, considerable resources are aimed at developing more sophisticated supernova cosmology analyses.
-The role of simulations mimicking survey observations has become increasingly important in determining biases in cosmological constraints and validating specific supernova models.
-First used in SNLS \citep{Astier2006} and ESSENCE analyses \citep{Wood-Vasey2007}, and then refined and improved for SDSS \citep{Kessler2009}, simulations are a fundamental component of modern supernova cosmology.
-\citet{Betoule2014} quantise and correct observational bias using simulations, and more recently \citet{Scolnic2016} and \citet{Kessler2017} explore simulations to quantify observational bias in SN~Ia distances as a function of multiple factors to improve bias correction.
-Approximate Bayesian computation methods also make use of simulations, trading traditional likelihoods and analytic approximations for more robust models with the cost of increased computational time \citep{Weyant2013, Jennings2016}.
-Bayesian Hierarchical models abound \citep{Mandel2009, March2011, March2014, Rubin2015, Shariff2016, Roberts2017}, and either use simulation-determined distance-corrections to correct for biases, or attempt to find analytic approximations for effects such as Malmquist bias to model the biases inside the BHM itself.
-
-In this paper, we lay out a new hierarchical model that builds off the past work of \citet{Rubin2015}.
-We include additional sources of systematic uncertainty, including an analytic formulation of selection efficiency which incorporates parametric uncertainty.
-We also implement a different model of intrinsic dispersion to both incorporate redshift-dependent scatter and to increase numerical efficiency, allowing our model to perform rapid fits to supernovae datasets.
-
-Section \ref{sec:review} is dedicated to a quick review of supernova cosmology analysis methods, and Section \ref{sec:challenges} outlines some of the common challenges faced by analysis methods.
-In Section \ref{sec:method} we outline our methodology.
-Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.
-We summarise our methodology in Section \ref{sec:conclusion}.
-
-\section{Review}
-\label{sec:review}
-
-Whilst supernova observations take the form of photometric time-series brightness measurements in many bands and a redshift measurement of the supernova (or its assumed host), most analyses do not work from these measurements directly.
-Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}.
-This model is trained separately before fitting the supernova light curves for cosmology \citep{Guy2010, Betoule2014}.
-The resulting output from the model is, for each supernova, an amplitude $x_0$ (which can be converted into apparent magnitude, $m_B = -2.5\log(x_0)$), a stretch term $x_1$, and color term $c$, along with a covariance matrix describing the uncertainty on these summary statistics ($\mathcal{C_\eta}$).
-As all supernovae are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.
-
-This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population.
-Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology.
-Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space.
-Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:
-\begin{equation}
-\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
-\end{equation}
-where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter.
-$\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}.
-The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift.
-The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses.
-Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.
-
-\subsection{Traditional Cosmology Analyses} \label{sec:traditional}
-
-Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus.
-The $\chi^2$ function minimised is
-\begin{equation}
-\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
-\end{equation}
-where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis).
-The predicted $\mu_C$ is defined as
-\begin{eqnarray}
-\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
-d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
-E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
-\end{eqnarray}
-where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.
-
-The benefit this analysis methodology provides is speed -- for samples of hundreds of supernovae or less, efficient matrix inversion algorithms allow the likelihood to be evaluated quickly.
-The speed comes with several limitations.
-Firstly, formulating a $\chi^2$ likelihood results in a loss of model flexibility by assuming Gaussian uncertainty.
-Secondly, the method of creating a covariance matrix relies on computing partial derivatives and thus any uncertainty estimated from this method loses information about correlation between sources of uncertainty.
-For example, the underlying supernova color population's mean and skewness are highly correlated, however this correlation is ignored when determining population uncertainty using numerical derivatives of population permutations.
-Whilst correlations can be incorporated into a covariance matrix, it requires human awareness of the correlations and thus methods that can automatically capture correlated uncertainties are preferable.
-Thirdly, the computational efficiency is dependent on both creating the global covariance matrix, and then inverting a covariance matrix with dimensionality linearly proportional to the number of supernovae.
-As this number increases, the cost of inversion rises quickly, and is not viable for samples with thousands of supernovae.
-A recent solution to this computational cost problem is to bin the supernovae in redshift bins, which produces a matrix of manageable size and can allow fitting without matrix inversion at every step.
-Whilst binning data results in some loss of information, recent works tested against simulations show that this loss does not result in significant cosmological biases \citep{Scolnic2016, Kessler2017}.
-
-Selection efficiency, such as the well known Malmquist bias \citep{MalmquistK.G.1922} is accounted for by correcting the determined $\mu_{\rm obs}$ from the data, or equivalently, adding a distance bias to the $\mu_C$ prediction.
-Specifically, Malmquist bias is the result of losing the fainter tail of the supernova population at high redshift.
-An example of Malmquist bias is illustrated in Figure \ref{fig:malmquist}, which simulates supernovae according to equation \eqref{eq:standard}.
-Simulations following survey observational strategies and geometry are used to calculate the expected bias in distance modulus, which is then subtracted from the observational data.
-When using traditional fitting methods such as that found in \citet{Betoule2014}, these effects are not built into the likelihood and instead are formed by correcting data.
-This means that the bias uncertainty is not captured fully in the $\chi^2$ distribution, and subtle correlations between cosmological or population parameters and the bias correction is lost.
-Recent developments such as the BBC method \citep{Kessler2017} incorporate corrections dependent on $\alpha$ and $\beta$, improving their capture of uncertainty on bias corrections in the $\chi^2$ likelihood.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{malmquist.pdf}
-	\end{center}
-	\caption{An example of the effects of Malmquist bias.
-Here are shown 1000 simulated supernovae redshifts and distance modulus given fiducial cosmology.
-The simulated survey is magnitude limited, and all supernovae brighter than magnitude 24 are successfully observed (shown as blue dots), and all dimmer than 24th magnitude are not successfully observed (shown as red dots).
-By binning the supernovae along redshift, and taking the mean distance modulus of the supernovae in each bin, we can see that at higher redshift where Malmquist bias kicks in, the population mean drops and becomes biased.
-This source of bias must either be corrected by adjusting the data (such as subtracting the found bias) for by incorporating Malmquist bias explicitly in the cosmological model.}
-	\label{fig:malmquist}
-\end{figure}
-
-\subsection{Approximate Bayesian Computation}
-
-To avoid the limitations of the traditional approaches, several recent methods have adopted Approximate Bayesian Computation, where supernova samples are forward modelled in parameter space and compared to observed distributions.
-\citet{Weyant2013} provides an introduction into ABC methods for supernova cosmology in the context of the SDSS-II results \citep{Sako2014} and flat $\Lambda$CDM cosmology, whilst \citet{Jennings2016} demonstrates their \textit{superABC} method on simulated first season Dark Energy Survey samples.
-In both examples, the supernova simulation package SNANA \citep{Kessler2009a} is used to forward model the data at each point in parameter space.
-
-Simulations provide great flexibility and freedom in how to treat the systematic uncertainties and selection effects associated with supernova surveys.
-By using forward modelling directly from these simulations, data does not need to be corrected, analytic approximations do not need to be applied, and we are free to incorporate algorithms that simply cannot be expressed in a tractable likelihood such as those found in traditional analyses from Section \ref{sec:traditional}.
-This freedom comes with a cost -- computation.
-The classical $\chi^2$ method's most computationally expensive step in a fit is matrix inversion.
-For ABC methods, we must instead simulate an entire supernova population -- drawing from underlying supernova populations, modelling light curves, applying selection effects, fitting light curves and applying data cuts.
-This is an intensive process.
-
-One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$, and $c$).
-\citet{Jennings2016} presents two metrics, which are used to measure the distance between the forward modelled population and observed population, and are minimised in fitting.
-The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics.
-However, we note that evaluation of systematic uncertainty was only performed using the Tripp metric.
-
-\subsection{Bayesian Hierarchical Models}
-
-Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM).
-Hierarchical models utilise multiple layers of connected parameters, with the layers linked via well defined and physically motivated conditional probabilities.
-For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter.
-We can thus incorporate different population distributions, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.
-Unlike ABC methods, which can model arbitrary probability distributions, BHM methods are generally constrained to representing probabilities using analytic forms.
-
-With the introduction of multiple layers in our model, we can add more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood.
-\citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light-curve fitting.
-\citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but do not incorporate bias corrections.
-\citet{March2014} and \citet{Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS-like simulations.
-The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset \citep[using redshift dependent bias corrections from][]{Betoule2014}, whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$.
-\citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}.
-Notably, these methods rely on data that is bias corrected or the methods ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}.
-The assumption made by the UNITY analysis is that the bias in real data is perfectly described by an analytic function.
-They validate their  model to be free of significant biases using fits to thirty realisations of supernova datasets that are constructed to mimic the UNITY framework.
-The well known BEAMS (Bayesian estimation applied to multiple species) methodology from \citet{Kunz2007} has been extended and applied in several works \citep{Hlozek2012}, most lately to include redshift uncertainty for photometric redshift application as zBEAMS \citep{Roberts2017} and to include simulated bias corrections in \citet{Kessler2017}.
-For the latter case, by inferring biases using Bayesian models, sophisticated corrections can be calculated and then applied to more traditional $\chi^2$ models.
-
-Whilst there are a large number of hierarchical models available, none of them have undergone comprehensive tests using realistic simulations to verify each models' respective bias.
-Additionally, testing has generally been performed on supernovae simulations with either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology.
-However, quantifying the biases on $w$CDM cosmology simulations with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own, and focus shifts from determination of $\Omega_m$ to both $\Omega_m$ and $w$.
-
-The flexibility afforded by hierarchical models allows for investigations into different treatments of underlying supernova magnitude, color and stretch populations, host-galaxy corrections, and redshift evolution, each of which will be discussed further in the outline of our model below.
-Our model is designed to increase the numerical efficiency of prior works whilst incorporating the flexibility of hierarchical models.
-We reduce our dependence on an assumed scatter model in simulations by not utilising bias-corrections in an effort to provide a valuable cross-check on analysis methodologies which utilise scatter-model-dependant bias corrections.
-
-\section{Challenges in Supernova Cosmology}
-\label{sec:challenges}
-
- The diverse approaches and implementations applied to supernova cosmology are a response to the significant challenges and complications faced by analyses.
-In this Section, we outline several of the most prominent challenges.
-
- Forefront among these challenges is our ignorance of the underlying type Ia intrinsic dispersion.
-Ideally, analysis of the underlying dispersion would make use of an ensemble of time-series spectroscopy to characterise the diversity of type Ia supernovae.
-However this data is difficult to obtain, and recent efforts to quantify the dispersion draw inference from photometric measurements.
-The underlying dispersion model is not a solved problem, and we therefore test against two dispersion models in this work.
-The first is based on the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, the second is sourced from \citet[][hereafter denoted {\celeven}]{Chotard2011}.
-As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$, and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C_\eta}$.
-These two scatter models are currently assumed to span the possible range of scatter in the underlying supernova population.
-We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.
-
-The underlying supernova population is further complicated by the presence of outliers.
-Non-type Ia supernovae often trigger transient follow-up in surveys and can easily be mistaken for type Ia supernovae  and represent outliers from the standardised SN~Ia sample.
-This contamination is not just a result of non-SN~Ia being observed, but can also arise from host galaxy misidentification causing incorrect redshifts  being assigned to supernovae.
-Different optimizations to the host-galaxy algorithm can result in misidentification of the host at the 3\% to 9\% level \citep{Gupta2016}, resulting in a broad population of outliers.
-For spectroscopic surveys, where both supernova type and redshift can be confirmed through the supernova spectra, this outlier population is negligible.
-However, for photometric surveys, which do not have the spectroscopic confirmation, it is one of the largest challenges; how to model, fit, and correct for contaminants.
-
-Finally, one of the other persistent challenges facing supernova cosmology analyses are the high number of systematics.
-Because of the rarity of SN~Ia events, datasets are commonly formed from the SN~Ia discoveries of multiple surveys in order to increase the number of supernovae in a dataset.
-However, each survey introduces additional sources of systematic error, from sources within each survey such as band calibration, to systematics introduced by calibration across surveys.
-Peculiar velocities, different host environments, and dust extinction represent additional sources of systematic uncertainty which must all be modelled and accounted for.
-
-\section{Our Method}
-\label{sec:method}
-
-We construct our hierarchical Bayesian model {\steve} with several goals in mind: creation of a redshift-dependent underlying supernova population,  treatment of an increased number of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections.
-We also desire {\steve} more computationally efficient than prior works, such that cosmological results from thousands of supernovae are obtainable in the order of hours, rather than days.
-As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in the programming language Stan \citep{Carpenter2017, StanDevelopmentTeam2017}.
-The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS, a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.
-
-At the most fundamental level, a supernova cosmology analysis is a mapping from an underlying population onto an observed population, where cosmological parameters are encoded directly in the mapping function.
-The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both the observed and underlying populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
-In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
-Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.
-
-\subsection{Observed Populations}
-
-\subsubsection{Observables}
-Like most of the BHM methods introduced previously, we work from the summary statistics, where each observed supernova has a brightness measurement $\hat{m}_B$ (which is analogous to apparent magnitude), stretch $\hat{x}_1$ and color $\hat{c}$, with uncertainty on those values encoded in the covariance matrix $\cov$.
-Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses.
-We also have a probability of each supernova being a type Ia, $\hat{p}$.
-Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
-
-As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus.
-Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
-One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.
-
-\begin{deluxetable}{ll}
-	\tablecolumns{2}
-	\tablewidth{0.99\columnwidth}
-	\tablecaption{Model Parameters}
-	\tablehead { \colhead{Parameter} & \colhead{Description} }
-	\startdata
-
-	&\textbf{Global Parameters} \\
-
-	$\Omega_m$  & Matter density  \\
-	$w$  &  Dark energy equation of state  \\
-	$\alpha$ & Stretch standardisation   \\
-	$\beta$         &  color standardisation   \\
-	$\delta(0)$ & Scale of the mass-magnitude correction\\
-	$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
-	$\delta\mathcal{Z}_i$ & Systematics scale\\
-	$\langle M_B \rangle$ & Mean absolute magnitude \\
-	\\
-	& \textbf{Survey Parameters} \\
-
-	$\delta S$ & Selection effect deviation \\
-	$\langle x_1^i \rangle$ & Mean stretch nodes\\
-	$\langle c^i \rangle$ & Mean color nodes\\
-	$\alpha_c$ & Skewness of color population \\
-	$\sigma_{M_B}$ & Population magnitude scatter \\
-	$\sigma_{x_1}$ & Population stretch scatter \\
-	$\sigma_{c}$ & Population color scatter \\
-	$\kappa_{0}$ & Extra color dispersion\\
-	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
-	\\
-	& \textbf{Supernova Parameters} \\
-	$m_B$ & True flux\\
-	$x_1$ & True stretch \\
-	$c$ & True color \\
-	$z$ & True redshift \\
-	$M_B$ & Derived absolute magnitude \\
-	$\mu$ & Derived distance modulus \\
-	\\
-	& \textbf{Input Data}\tablenotemark{a} \\
-
-	$\hat{m}_B$ & Measured flux\\
-	$\hat{x}_1$ & Measured stretch \\
-	$\hat{c}$ & Measured color \\
-	$C$ & Covariance on flux, stretch and color \\
-	$\hat{z}$ & Observed redshift \\
-	$\hat{m}$ & Determined mass probability \\
-	$\hat{p}$ & Determined Ia probability \\
-	\enddata
-	\tablenotetext{a}{Not model parameters but shown for completeness.}
-\label{tab:param_summary}
-\end{deluxetable}
-
-\subsubsection{Latent Variables for Observables}
-
-The first layer of hierarchy is the set of true (latent) parameters that describe each supernova.
-In contrast to the observed parameters, the latent parameters are denoted without a hat.
-For example, $c$ is the true color of the supernova, whilst $\hat{c}$ is the observed color, which, as it has measurement error, is different from $c$.
-
-For the moment, let us consider a single supernova and its classic summary statistics $m_B$, $x_1$, $c$.
-For convenience, let us define $\eta \equiv \lbrace m_B, x_1, c \rbrace$.
-A full treatment of the summary statistics would involve determining $p(\hat{y}|\eta)$, where $\hat{y}$ represents the observed light curves fluxes and uncertainties.
-However, this is computationally prohibitive  as it would require incorporating SALT2 light curve fitting inside our model fitting.
-Due to this computational expense, we rely on initially fitting the light curve observations to produce a best fit $\hat{\eta}$ along with a $3\times3$ covariance matrix $\mathcal{C_\eta}$ describing the uncertainty on $\hat{\eta}$.
-Using this simplification, our latent variables are given by
-\begin{equation}
-p(\hat{\eta}|\eta) \sim \mathcal{N}(\hat{\eta} | \eta, \mathcal{C_\eta}). \label{eq:pop}
-\end{equation}
-As discussed in Section \ref{sec:challenges},  the SALT2 model does not include full treatment of intrinsic dispersion, and thus this approximation does not encapsulate the full uncertainty introduced from this dispersion.
-
- \begin{figure}
- 	\begin{center}
- 		\includegraphics[width=\columnwidth]{tikz-figure0.pdf}
- 	\end{center}
- 	\caption{Probabilistic graphical model for our likelihood.
-Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
-The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.
-Variables that appear outside both boxes represent top level model parameters.}
- 	\label{fig:pgm}
- \end{figure}
-
-\subsection{Underlying Population}
-
-\subsubsection{Type Ia population} 
-\label{sec:underlying}
-
-Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space.
-To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
-
-In addition to absolute magnitude, the underlying supernova population is also characterised by distributions in color and stretch.
-We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey.
-For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution.
-We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude.
-Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift.
-These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$.
-Both the color and stretch means are modelled with normal priors.
-Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment.
-We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ that allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}.
-We sample $\delta_c$ in log space for efficiency in sampling close to zero.
-The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors.
-
-The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$.
-We model the colour skewness and the redshift-dependent means and width of the colour and skew distributions individually for each survey.
-The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
-\begin{eqnarray}
-P(M_B, x_1, c, z | \theta) &=& \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \, \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1})\notag\\ &\quad\quad& \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1},
-\end{eqnarray}
-where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.
-
-\subsubsection{Outlier populations}
-
-For future use with photometrically classified surveys, we include a simplistic outlier population model.
-We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
-For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works.
-However, with the increased number of contaminants expected in the DES photometric sample, we seek a more physically motivated outlier population.
-We find that an acceptable parametrisation is to model the outlier population with a mean absolute magnitude of $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations.
-We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude.
-We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
-For the spectroscopic survey, we set this to unity.
-For the photometric proof-of-concept we provide an accurate probability estimate.
-Further investigation on the effect of inaccurate estimates will be left for future improvements during the analysis of the DES photometric sample.
-
-\subsection{Correcting biased summary statistics}
-
-With the fitted summary statistics $\hat{\eta}$ being biased and their uncertainty under-reported, we face a significant challenge utilising these statistics naively in supernova cosmology.
-We must either correct the observables to remove the biases introduced by the intrinsic dispersion of the underlying population, or incorporate this dispersion into our model.
-We should also avoid assuming a specific dispersion model -- either the {\gten} or {\celeven} model, or utilise the results of computing the bias from both models.
-
-We model the extra dispersion only in color, and do so by adding independent uncertainty on the color observation $\hat{c}$.
-We note that extra dispersion in magnitude $\hat{m}_B$ (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the color term, hence the requirement for modelling additional color dispersion.
-Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than color, and causes negligible bias in cosmology.
-
-As shown in \citep{Kessler2013}, the extra color dispersion shows heavy redshift dependence, increasing with redshift.
-This is an artifact of different filters, however as we may be subject to similar effects in our observational data, we decide to incorporate redshift dependence in our extra uncertainty.
-We thus add $\kappa_0 + \kappa_1 z$ to our observed color uncertainty (in quadrature).
-The $\kappa$ parameters are highly degenerate with the width of the intrinsic color population $\sigma_c$.
-We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$.
-We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.
-As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C_\eta} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.
-
-Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just color dispersion, and was far more computationally inefficient.
-We note here that we model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations.
-This magnitude dispersion is modelled with redshift independence.
-
-\subsection{Mapping function}
-
-\subsubsection{Cosmology}
-
-We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM.
-$\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$.
-For calculating the distance modulus, we fix $H_0 = 70 \kmsmpc $.
-If the Hubble constant has a different value, the absolute magnitude is $M_B + 5\log(H_0/70 \kmsmpc )$ with the other cosmological parameters unaffected.
-
-\subsubsection{Supernova Standardisation Parameters}
-
-With increasingly large datasets and more nuanced analyses, the choice of how to handle $\alpha$ and $\beta$ becomes an important consideration when constructing a model.
-{\rubin} employs a broken linear relationship for both color and stretch, where different values of $\alpha$ and $\beta$ are adopted depending on whether $x_1$ and $c$ are positive or negative (although the cut could be placed at a location other than 0).
-\citet{Shariff2016} instead model $\beta$ as redshift-dependent, testing two phenomenological models; $\beta(z) = \beta_0 + \beta_1 z$ and a second model which effects a rapid but smooth change in $\beta$ at a turnover redshift $z_t$.
-
-We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$.
-See Section \ref{sec:simdes} for details on simulation generation.
-We found for both models that non-zero values for $\beta_1$ are preferred even with constant $\beta$ used in simulation, due to severe degeneracy with selection effects.
-This degeneracy resulted in a significant bias in recovered cosmology.
-Due to the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses.
-As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.
-
-\subsubsection{Host Galaxy Environment}
-\label{sec:hostgal}
-
-There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}.
-The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}.
-The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step.
-The scale of this step function varies from analysis to analysis, and is treated as a fit parameter.
-In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
-\begin{equation}
-\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
-\end{equation}
-where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift.
-We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.
-
-\subsubsection{Uncertainty Propagation}
-\label{sec:systreat}
-The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations.
-As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
-
-Assuming that the gradients can be linearly extrapolated -- which is a reasonable approximation for modern surveys with high quality control of systematics -- we can incorporate into our model a deviation from the observed values by constructing a $(3 \times N_{\rm sys})$ matrix containing the numerical derivatives for the $N_{\rm sys}$ systematics and multiplying it with the row vector containing the offset for each systematic.
-By scaling the gradient matrix to represent the shift over $1\sigma$ of systematic uncertainty, we can simply enforce a unit normal prior on the systematic row vector to increase computational efficiency.
-
-This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses.
-For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty.
-We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty.
-A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.
-This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey.
-With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
-\begin{equation}
-P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right).
-\label{eq:l3}
-\end{equation}
-
-\subsubsection{Selection Effects}
-\label{sec:selection}
-
-One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
-We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
-To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
-\begin{align}
-\mathcal{L}(\theta; {\rm data}) &= P({\rm data} | \theta, S). \label{eq:like}
-\end{align}
-As our model so far describes components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
-\begin{align}
-\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
-\end{align}
-where the denominator represents an integral over all potential data $D$.
-This is derived in Appendix \ref{app:selection1} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
-We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as
-\begin{equation}
-d = \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
-\end{equation}
-where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$, and $c$ population to $m_B$ given cosmological parameters.
-A full derivation of this can be found in Appendix \ref{app:selection2}.
-
-We now apply two further approximations similar to those made in {\rubin} -- that the redshift distribution of the observed supernovae reasonably well samples the $P(S|z)P(z|\theta)$ distribution, and that the survey color and stretch populations can be treated as Gaussian for the purposes of evaluating $P(m_B | z, \theta)$.
-We found that discarding the color population skewness entirely resulted in highly biased population recovery (see Figure \ref{fig:simple_w_super} to see the populations), and so we instead characterise the skew normal color distribution with a Gaussian that follows the mean and variance of a skew normal; with mean given by $\langle c(z) \rangle + \sqrt{\frac{2}{\pi}} \sigma_c \delta_c$ and variance $\sigma_c^2(1 - 2\delta_c^2/\pi)$.
-This shifted Gaussian approximation for color completely removes the unintended bias when simply discarding skewness.
-This shift was not required for the stretch population, and so was left out for the stretch population for numerical reasons.
-The impact of this approximation on the calculated efficiency is shown in Figure \ref{fig:shift}, and more detail on this shift and resulting population recovery can be found in Appendix \ref{app:approx}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{shift.pdf}
-	\end{center}
-	\caption{Testing the correctness of our normal approximation to the skewed color distribution.
-The `correct' line (shown in black) represents the exact integral $w = \int P(S|x) P(x) dx$ where $P(S|x)$ is an error function (following our high-redshift surveys) and $P(x) = \mathcal{N}^{\rm Skew}(x, 0.1, 2)$, calculated numerically.
-The $x$-axis is analogous to $m_B$ is cosmological context.
-As expected, all efficiencies drop towards zero as shift increases (as objects get fainter).
-The unshifted normal approximation shows significant discrepancy in the calculated efficiency as it transitions from 1 to 0, whilst the shifted normal approximation shows negligible error to the correct solution.
-From these plots, further refinement of the normal approximation (such as including kurtosis or higher powers) are unnecessary.}
-	\label{fig:shift}
-\end{figure}
-
-The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
-\begin{eqnarray}
-m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
-\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
-\end{eqnarray}
-What then remains is determining the functional form of $P(S|m_B)$.
-For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient.
-Formally, this gives
-\begin{equation}
-P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
-\end{equation}
-where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function.
-The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}.
-However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
-\begin{equation}
-P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
-\end{equation}
-
-The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by taking the ratio of supernovae that are observed and passed cuts over the total number of supernovae generated in that apparent magnitude bin.
-That is, we calculate the probability we would include a particular supernova in our sample, divided by the number of such supernovae in our simulated fields.
-To take into account the uncertainty introduced by the imperfection of our analytic fit to the efficiency ratio, uncertainty was uniformly added in quadrature to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.
-This is mathematically identical to fitting the efficiency ratio with a second `intrinsic dispersion' parameter which adds uncertainty to the efficiency ratio data points.
-
-We thus include into our model parametrised selection effects by including the covariance matrix of selection effect uncertainty.
-Formally, we include deviations from the determined mean selection function parameters with parameter vector $\Delta S$, and apply a normal prior on this parameter as per the determined uncertainty covariance matrix.
-Whilst this uncertainty encapsulates the potential error from the simulations not matching the analytic approximations, it does not cover potential variations of the selection function at the top level --- varying cosmology or spectroscopic efficiency.
-Tests with changing the intrinsic scatter model used in the selection efficiency simulations show that the uncertainty introduced is negligible.
-
-With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova.
-For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
-\begin{eqnarray}
-d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
-d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
-&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
-\end{eqnarray}
-and is incorporated into our likelihood.
-This is illustrated in Figure \ref{fig:efficiency}.
-Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency.
-Further details on this choice are given in Section \ref{sec:simdes}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{efficiency.pdf}
-	\end{center}
-	\caption{The efficiency for supernova discovery at an arbitrary redshift.
-Shown in both panels in dashed blue is the SN~Ia population distribution, which takes the form of a normal distribution.
-The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), as functions of apparent magnitude.
-The survey efficiency, given the SN~Ia population, is shown as a solid line in both panels, and the probability of observing a SN~Ia is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
-This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
-	\label{fig:efficiency}
-\end{figure}
-
-\section{Model Verification}
-\label{sec:verification}
-
-In order to verify our model we run it through stringent tests.
-First, we validate on toy models, verifying that we recover accurate cosmology when generating toy supernovae data constructed to satisfy the assumptions of the BHM construction.
-We then validate our model on SNANA simulations based on a collection of low redshift surveys and the DES three-year spectroscopic sample, termed the DES-SN3YR sample
-
-\subsection{Applied to Toy Spectroscopic Data}
-\label{sec:toy}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
-	\end{center}
-	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys.
-Selection effects are visible in both samples, where red supernovae are often cut as redshift increases, creating a skewed color population.
-The color of the data points is representative of the supernova color itself, a negative color value showing bluer supernovae, with positive color values representing redder supernovae.}
-	\label{fig:simple_pop}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
-	\end{center}
-	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context.
-Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ or $\Omega_\Lambda$ separately due to the severe degeneracy between the parameters.}
-	\label{fig:simple_ol}
-\end{figure}
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
-	\end{center}
-	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the non-Gaussian nature of the posterior surface.
-The recovered posteior maximums show the same degeneracy direction as the representative surface, and scatter around the truth values input into the simulation, which are shown in dashed lines.}
-	\label{fig:simple_w}
-\end{figure}
-
-\begin{deluxetable}{lcc}
-	\tablecolumns{3}
-	\tablecaption{Cosmological Parameters for toy supernova data}
-	\tablehead {
-		\colhead {Model}              &
-		\colhead {$\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}  &
-		\colhead{$w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}
-	}
-
-	\startdata
-		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
-		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
-	\enddata
-	\label{tab:simple_model}
-	\tablecomments{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data.
-As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae.
-For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviation in the table.
-The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets.
-Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.}
-\end{deluxetable}
-
-We generate simple toy data to validate the basic premise of the model. The data generation algorithm is described below:
-
-\begin{enumerate}[1.]
-	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
-	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$, $\delta(0)/\delta(\infty) = 0.5$, and equation \eqref{eq:mass}.
-	\item Draw an absolute magnitude, stretch and color from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
-Calculate the true absolute magnitude using Equation \eqref{eq:standard}.
-	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$ under Flat $\Lambda$CDM cosmology.
-Use this to determine the true apparent magnitude of the object $m_B$.
-	\item Determine if the SN~Ia is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data).
-For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$.
-Repeat from step one until we have a supernova that passes.
-We use realistic values for the selection probability to ensure our model is numerically stable with highly skewed selection functions.
-	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations).
-Add extra color uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
-\end{enumerate}
-
-The selection functions parameters (a skew normal for low-redshift and a complimentary error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
-Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample.
-Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.
-
-We test four models: Flat $\Lambda$CDM, Flat $w$CDM, $\Lambda$CDM, and Flat $w$CDM with a prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, with the latter included to allow sensitive tests on bias for $w$.
-To achieve statistical precision, we fit 100 realisations of supernovae datasets.
-Cosmological parameters are recovered without significant bias.
-Combined posterior surfaces of all 100 realisations fits for $\Lambda$CDM are shown in Figure \ref{fig:simple_ol} and fits for Flat $w$CDM are shown in Figure \ref{fig:simple_w}.
-By utilising the Stan framework and several efficient parametrisations (discussed further in Appendix \ref{app:optimisations}), fits to these simulations of 2000 supernovae take only on order of a single CPU-hour to run.
-
-To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$.
-This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces.
-The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}.
-Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'.
-By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them.
-The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
-As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted scatter.
-The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
-With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.
-
-\subsection{DES SN data validation}
-\label{sec:simdes}
-
-Many BHM methods have previously been validated on data constructed explicitly to validate the assumptions of the model.
-This is a useful consistency check that the model implementation is correct, efficient and free of obvious pathologies.
-However, the real test of a model is its application to realistic datasets that mimic expected observational data in as many possible ways.
-To this end, we test using simulations (using the SNANA package) that follow the observational schedule and observing conditions for the DES and low-$z$ surveys, where the low-$z$ sample is based on observations from CfA3 \citep{Hicken2009, Hicken2009a}, CfA4 \citep{Hicken2012} and CSP \citep{Contreras2010,Folatelli2010,Stritzinger2011}.
-Simulation specifics can be found in \citet{Kessler18}.
-
-Additionally, prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
-In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
-The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation.
-These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.
-
-In addition to the improvements in testing multiple scatter models, we also include peculiar velocities for the low-$z$ sample, and our full treatment of systematics as detailed in \citet{Brout18SYS}.
-Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}.
-Initial tests were also done with a second, Gaussian population with color and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency.
-The selection effects were quantified by comparing all the generated supernovae to those that pass our cuts, as shown in Figure~\ref{fig:sf_snana}.
-It is from this simulation that our analytic determination of the selection functions for the low-$z$ and DES survey are based.
-We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models.
-
-\begin{deluxetable}{lcccc}
-\tablewidth{\columnwidth}
-
-\tablecolumns{5}
-\tablecaption{Supernova population distributions}
-\tablehead {
-	\colhead {Model} &
-	\colhead { $\langle x_1 \rangle$  } &
-	\colhead { $\sigma_{x_1}$ } &
-	\colhead { $\langle c \rangle$ } &
-	\colhead { $\sigma_c$ } 
-}
-	\startdata
-		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\ \\
-		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$  \\
-	\enddata
-	\label{tab:dist}
-	\tablecomments{The {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
-\end{deluxetable}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{DES3YR_DES_BHMEFF_AMG10.pdf}
-		\includegraphics[width=\columnwidth]{DES3YR_LOWZ_BHMEFF_G10.pdf}
-	\end{center}
-	\caption{Fitting the selection function for both the DES 3YR spectroscopically confirmed supernova sample and the low-$z$ sample.
-Blue errorbars represent the efficiency calculated by determining the ratio of discovered to generated supernovae in apparent magnitude bins for SNANA simulations.
-The black line represents the best fit analytic function for each sample, and the light grey lines surrounding the best fit value represent random realisations of analytic function taking into account uncertainty on the best fit value.}
-	\label{fig:sf_snana}
-\end{figure}
-
-\begin{deluxetable*}{lrrrrr}
-	\tablewidth{\textwidth}
-	\tablecolumns{6}
-	\tablecaption{Realistic simulation standardisation parameters}
-	\tablecomments{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations.
-We show the average parameter mean and average standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$.
-The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model.}
-	\tablehead{
-		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$	
-	}
-	\startdata
-		{\gten} Stat + Syst  & $0.022 \ [0.009\ (0.008)]$ & $ 0.34 \ [0.19\ (0.18)]$ & $-0.002 \ [0.028\ (0.015)]$ & $0.070 \ [0.022\ (0.018)]$ & $0.073 \ [0.025\ (0.022)]$\\
-		{\gten}  Stat         & $0.000 \ [0.008\ (0.008)]$ & $0.33 \ [0.20\ (0.17)]$ & $0.001 \ [0.016\ (0.013)]$ & $0.069 \ [0.023\ (0.019)]$ & $0.072 \ [0.026\ (0.023)]$\\
-		{\celeven} Stat + Syst  &  $0.002 \ [0.009\ (0.007)]$ & $-0.04 \ [0.15\ (0.13)]$ & $0.014 \ [0.030\ (0.018)]$ & $0.024 \ [0.016\ (0.011)]$ & $0.029 \ [0.020\ (0.014)]$ \\
-		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$ \\
-
-	\enddata
-	\label{tab:standardisation}
-\end{deluxetable*}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{bulk_points_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_points_c11.pdf}
-	\end{center}
-	\caption{Maximum posterior points for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel.
-Points are shown for parameters $\Omega_m$, $w$, $\alpha$ and $\beta$, with the other fit parameters being marginalised over.
-As we are unable to fully correct observed summary statistics, a step required by the lack of intrinsic scatter in the SALT2 model, we expect to see an offset in $\alpha$ and $\beta$.
-This in turn effects cosmology, resulting in small biases in $w$.}
-	\label{fig:bulk_posterior}
-\end{figure}
-
-\begin{deluxetable}{lrr}
-	\tablecaption{Realistic simulation determination of $w$}
-	\tablewidth{\columnwidth}
-	\tablecomments{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics.
-The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets.
-The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.}
-	\tablehead{		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias		}
-	\startdata
-		{\gten} Stat + Syst  &  $-0.998 \ [0.097\ (0.073)]$ & $(0.02\pm0.07)\sigma$  \\
-		{\gten}  Stat         &  $-1.008\ [0.080\ (0.068)]$ & $(-0.10\pm0.08)\sigma$   \\
-		{\celeven} Stat + Syst  &  $-0.945\ [0.098\ (0.077)]$ & $(0.55\pm0.08)\sigma$  \\
-		{\celeven} Stat         &  $-0.948\ [0.079\ (0.066)]$ & $(0.65\pm0.08)\sigma$   \\
-	\enddata
-		\label{tab:bulk_summary}
-\end{deluxetable}
-
-\begin{deluxetable}{crr}
-
-	\tablecaption{Reduced parameter correlations with $w$.}
-	\tablecomments{Correlations determined from the combined 100 simulation fits.
-Correlations for the low-$z$ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation.
-Zero superscripts indicate the Dark Energy Survey, and a superscript one represents the low-$z$ survey.}
-	\tablehead{		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst }
-	\startdata
-		$\Omega_m$ &   $-$0.19 &     $-$0.21     \\
-		$\alpha$ &   $-$0.17 &     $-$0.20     \\
-		$\beta$ &   $-$0.29 &     $-$0.23     \\
-		$\langle M_B \rangle$ &    0.68 &      0.66     \\
-		$\delta(0)$ &    0.00 &      0.00     \\
-		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
-		\\
-		$\sigma_{\rm m_B}^{0}$ &    0.04 &      0.07     \\
-		$\sigma_{\rm m_B}^{1}$ &    0.23 &      0.18     \\
-		$\sigma_{x_1}^{0}$ &    0.04 &      0.03     \\
-		$\sigma_{x_1}^{1}$ &    0.05 &      0.01     \\
-		$\sigma_{c}^{0}$ &    0.01 &      0.11     \\
-		$\sigma_{c}^{1}$ &    0.08 &      0.04     \\
-		$\alpha_c^{0}$ &   $-$0.04 &      0.04     \\
-		$\alpha_c^{1}$ &    0.03 &      0.01     \\
-		$\kappa_{c0}^{0}$ &   $-$0.10 &     $-$0.05     \\
-		$\kappa_{c0}^{1}$ &   $-$0.20 &     $-$0.17     \\
-		$\kappa_{c1}^{0}$ &   $-$0.05 &     $-$0.01     \\
-		$\kappa_{c1}^{1}$ &   $-$0.01 &      0.01     \\
-		$\langle x_1^{0} \rangle$ &   $-$0.01 &     $-$0.05     \\
-		$\langle x_1^{1} \rangle$ &   $-$0.02 &      0.02     \\
-		$\langle x_1^{2} \rangle$ &   $-$0.04 &     $-$0.04     \\
-		$\langle x_1^{3} \rangle$ &   $-$0.03 &     $-$0.06     \\
-		$\langle x_1^{4} \rangle$ &   $-$0.06 &     $-$0.06     \\
-		$\langle x_1^{5} \rangle$ &    0.04 &      0.02     \\
-		$\langle x_1^{6} \rangle$ &    0.04 &      0.04     \\
-		$\langle x_1^{7} \rangle$ &    0.08 &      0.03     \\
-		$\langle c^{0} \rangle$ &   $-$0.05 &     $-$0.12     \\
-		$\langle c^{1} \rangle$ &    0.11 &      0.03     \\
-		$\langle c^{2} \rangle$ &    0.11 &      0.06     \\
-		$\langle c^{3} \rangle$ &    0.14 &      0.04     \\
-		$\langle c^{4} \rangle$ &   $-$0.11 &     $-$0.11     \\
-		$\langle c^{5} \rangle$ &   $-$0.15 &     $-$0.08     \\
-		$\langle c^{6} \rangle$ &   $-$0.12 &     $-$0.13     \\
-		$\langle c^{7} \rangle$ &   $-$0.12 &     $-$0.06     \\
-		\\
-		$\delta [ {\rm SALT}_0 ]$ &    0.05 &      0.05     \\
-		$\delta [ {\rm SALT}_1 ]$ &   $-$0.01 &      0.02     \\
-		$\delta [ {\rm SALT}_2 ]$ &   $-$0.10 &     $-$0.09     \\
-		$\delta [ {\rm SALT}_3 ]$ &   $-$0.03 &     $-$0.03     \\
-		$\delta [ {\rm SALT}_4 ]$ &    0.08 &      0.09     \\
-		$\delta [ {\rm SALT}_5 ]$ &    0.01 &      0.02     \\
-		$\delta [ {\rm SALT}_6 ]$ &    0.05 &      0.07     \\
-		$\delta [ {\rm SALT}_7 ]$ &   $-$0.11 &     $-$0.10     \\
-		$\delta [ {\rm SALT}_8 ]$ &    0.01 &      0.02     \\
-		$\delta [ {\rm SALT}_9 ]$ &    0.02 &      0.02     \\
-		$\delta [ {\rm MWE}_{B-V} ]$ &    0.03 &      0.02     \\
-		$\delta [ {\rm HST Calib} ]$ &   $-$0.07 &     $-$0.07     \\
-		$\delta [ v_{\rm pec} ]$ &    0.00 &     $-$0.01     \\
-		$\delta [ {\rm \delta z } ]$ &    0.01 &      0.00     \\
-		$\delta [ \Delta g ]$ &    0.05 &      0.11     \\
-		$\delta [ \Delta r ]$ &    0.16 &      0.10     \\
-		$\delta [ \Delta i ]$ &   $-$0.16 &     $-$0.18     \\
-		$\delta [ \Delta z ]$ &   $-$0.26 &     $-$0.26     \\
-		$\delta [ \Delta \lambda_g ]$ &    0.16 &      0.20     \\
-		$\delta [ \Delta \lambda_r ]$ &    0.05 &      0.06     \\
-		$\delta [ \Delta \lambda_i ]$ &    0.00 &     $-$0.01     \\
-		$\delta [ \Delta \lambda_z ]$ &    0.09 &      0.07     \\
-	\enddata
-	\label{tab:parameter_correlations}
-\end{deluxetable}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=\textwidth]{bulk_cor.pdf}
-	\end{center}
-	\caption{Parameter correlations for the combined fits to the 100 {\gten} scatter model simulations.
-We see that the primary correlations with $w$ enter through $\alpha$, $\beta$ and $\langle M_B \rangle$, as shown in Table \ref{tab:parameter_correlations}.
-Whilst $\langle M_B \rangle$ is generally thought to be a nuisance parameter, we find cosmological correlation.
-We note that, by fixing $H_0$ in our distance modulus calculation, $\langle M_B \rangle$ absorbs any cosmological uncertainty on this term.
-Additionally $\langle M_B \rangle$ also effects the selection efficiency, which was computed from simulations with a fixed $M_B$ value, introducing a second plausible source of correlation.
-Also visible in this figure are several other interesting relationships.
-$\beta$ is strongly anti-correlated with intrinsic dispersion $\sigma_{m_B}$ for both surveys (DES-like and low-$z$), with $\sigma_{m_B}$ showing strong anti-correlation with $\kappa^0_c$.
-This relationship is indeed expected --- as $\kappa^0_c$ grows larger (more unexplained dispersion on the color observation), the width of the supernova population in apparent magnitude space increases.
-As the fit prefers it to conform to the observed width of the distribution, the extra width in color causes the inherent magnitude smearing amount to decrease.
-And with extra freedom on the observed color from $\kappa^0_c$, $\beta$ shifts in response.
-The other striking feature in the plot is the strong correlation blocks in the bottom right and the anti-correlation stripes on the edges.
-These too are expected, for they show the relationship between the color distribution's mean value, its width and its skewness.
-As skewness or population width increases, the effective mean of the population shifts (see Appendix \ref{app:approx} for details), creating anti-correlation between skewness and the (Gaussian) mean color population.
-Strong anti-correlation between $\kappa_{c0}^0$ and $\kappa_{c0}^1$ with $\sigma_{m_B}$ reveals the strong population degeneracy, and -- for the C11 simulation results -- a constrained positive value shows that a finite non-zero extra color dispersion is indeed preferred by our model.
-}
-	\label{fig:bulk_cor}
-\end{figure*}
-
-Each realisation of simulated SN~Ia light curves contains the SALT2 light-curve fits and redshifts to 128 low-$z$ supernovae, and 204 DES-like supernovae, such that the uncertainties found when combining chains is representative of the uncertainty in the DES-SN3YR sample.
-As our primary focus is Dark Energy, we now focus specifically on the Flat $w$CDM model with matter prior.
-
-Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}.
-The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  and secondary parameters are shown in Table \ref{tab:standardisation}.
-
-Table \ref{tab:bulk_summary} shows that the {\gten} model is consistent with $w=-1$, whilst the {\celeven} model show evidence of bias on $w$, scattering high.
-However, their deviation from the truth value represents a shift of approximately $0.5\sigma$ when taking into account the uncertainty on fits to $w$.
-The bias is sub-dominant to both the size of the uncertainty for each fit, and the scatter induced by statistical variance in the simulations.
-We also note that the simulations do not vary cosmological parameters nor population.
-As our model does include uncertainty on those values, the simulation scatter is expected to be less than the model uncertainty, and represents a minimum bound on permissible uncertainty values.
-
-Table \ref{tab:standardisation} shows a clear difference in both $\beta$ and $\sigma_{m_B}$ across the {\gten} and {\celeven} simulations.
-As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations.
-The extra smearing of the {\celeven}  model does not result in a significantly biased $\beta$ value compared to the average uncertainty on $\beta$, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$.
-Interestingly, $w$-bias is only found for the {\celeven} simulations.
-A measure of the significance of the parameter bias can be calculated by comparing the bias to a tenth of the scatter (as our Monte-Carlo estimate uncertainty is $\sqrt{100}$ of the scatter).
-From this, we can see that most biases are detected with high statistical significance due to the large number of simulations tested against.
-
-We investigate the cosmological bias and find its source to be a bias in the observed summary statistics (i.e.
-the $\hat{m}_B$, $\hat{x}_1$, $\hat{c}$ output from SALT2 light curve fitting), in addition to incorrect reported uncertainty on the summary statistics.
-To confirm this, we run two tests.
-The first of which, we replace the SALT2-fitted $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a Gaussian centered on the true SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits.
-With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly.
-Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
-To this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.
-From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
-Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced by an unknown scatter model.
-
-Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model.
-This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically.
-Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies --- incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.
-
-Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters.
-Tests when fixing the population parameters to true does not resolve the cosmological bias observed in the {\celeven} simulations.
-As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
-
-Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the low-$z$ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics.
-Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters.
-Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}.
-The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.
-
-For the sample size of the DES $+$ low-$z$ supernova samples (332 supernova), the bias from intrinsic scatter models is sub-dominant to the statistical uncertainty, as shown in Figure \ref{fig:bulk_posterior}.
-For our full systematics model, the bias represents a deviation between $0\sigma$ to $0.5\sigma$ depending on scatter model, and given that they remain sub-dominant, we will leave more complicated treatment of them for future work.
-
-\subsection{Uncertainty Analysis}
-\label{sec:ucnert}
-
-\begin{deluxetable*}{llrr}
-	\tablecaption{$w$ error budget}
-	\tablecomments{Error budget determined from analysing uncertainty on simulation data whilst progressively enabling model features.
-We start from the top of the table, only varying cosmological parameters $\Omega_m$ and $w$, and then progressively unlock parameters and let them fit as we progress down the table.
-The cumulative uncertainty shows the total uncertainty on $w$ on the fit for all, where the $\sigma_w$ term is derived by taking the quadrature difference in cumulative uncertainty as we progress.}
-	\tablehead{		Feature &  Parameters & $\sigma_w$ & Cumulative }
-	\startdata
-		Cosmology                      & $\Omega_m$, $w$ & 0.051  &   0.051       \\
-		Standardisation                & $\alpha$, $\beta$, $\langle M_B \rangle$, $\delta(0)$, $\delta(\infty)/\delta(0)$ & 0.046  &   0.068       \\
-		Intrinsic scatter              & $\kappa_0$, $\kappa_1$ & 0.020  &   0.071       \\
-		Redshift-independent populations             & $\sigma_{M_B}$, $\sigma_c$, $\sigma_{x_1}$, $\alpha_c$ & 0.022  &   0.074       \\
-		Redshift-dependent populations & $\langle c_i \rangle$, $\langle x_{1,i} \rangle$ & 0.030  &   0.080   \\
-		Systematics                    & $\delta \mathcal{Z}_i$, $\delta S$ & 0.054  &   0.096      \\ 
-	\enddata
-	\label{tab:uncert}
-\end{deluxetable*}
-
-With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
-To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
-By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
-
-The error budget detailed in Table \ref{tab:uncert} shows that our uncertainty is dominated by statistical error, as the total statistical uncertainty is on $w$ is $\pm0.08$.
-With the low number of supernovae in the DES-SN3YR sample, this is expected.
-We note that the label `Systematics' in Table \ref{tab:uncert} represents all numerically computed systematics (as discussed in Section \ref{sec:systreat}) and systematic uncertainty on the selection function.
-
-\subsection{Methodology Comparison}
-\label{sec:comp}
-
-We compare the results of our model against those of the BBC+CosmoMC method \citep{Kessler2017}.
-BBC+CosmoMC has been used in prior analyses, such as the Pantheon sample analysis of \citet{Scolnic2017} and is being used in the primary analysis of the DES-SN3YR sample \citep{DESKEY}.
-The BBC method is a two-part process, BBC computes bias corrections for observables, and then the corrected distances are fit using CosmoMC \citep{lewis2002cosmomc}.
-For shorthand, we refer to this combined process as the BBC method hereafter in this paper, as we are concerned with the results of cosmological parameter inference.
-As a leading supernova cosmology method, it provides a good consistency check as to the current levels of accuracy in recovering cosmological parameters.
-
-To this end, we take the results of the BBC method which were also run on the same set of 200 validation simulations and compare the recovered $w$ values to those of our method.
-The results are detailed in Table \ref{tab:bulk_summary_bbc}, and a scatter plot of the simulation results is presented in Figure \ref{fig:bulk_comparison}.
-
-As shown in \citet{Brout18SYS}, the BBC method recovers cosmological parameters without bias so long as the intrinsic scatter model is known.
-As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}.
-As such, we expect the BBC method to have a $w$-bias in one direction for {\gten} simulations and the other direction for {\celeven} simulations.
-These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}.
-Both BBC method and {\steve} are sensitive to the intrinsic scatter model, finding differences of $\sim0.066$ and $0.053$ respectively in $w$ when varying the scatter model.
-The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$.
-Our method shows a small improvement in the insensitivity to the intrinsic scatter model (having a decrease in difference in $w$ between the {\gten} and {\celeven} models), finding no bias for {\gten} but a $w$ biased high for {\celeven}.
-This decrease in error is not statistically significant as we have statistical uncertainty of $\sim0.01$ for $100$ simulation realisations.
-The average bias over the two scatter models is $+0.028$, representing a larger bias than the BBC method.
-
-When comparing both the {\gten} and {\celeven} set of simulations independently, our model differs from BBC in its average prediction of $w$ by $+0.044$ and $+0.033$ respectively.
-For the {\gten} model this difference is a result of bias in the BBC results, however for the {\celeven} simulations this is a result of both bias from BBC, and a larger bias from our method.
-These results also allow us to state the expected values for $w$ when run on the DES-SN3YR sample.
-When using Planck priors our uncertainty on $w$ is reduced compared to using our simulation Gaussian prior on $\Omega_m$, shrinking the average $w$-difference from $0.06$ to $0.04$.
-After factoring this into our uncertainty, we expect our BHM method to, on average, recover $w_{\rm BHM} = w_{\rm BBC} +0.04\pm0.04$.
-
-Having established that our method exhibits similar shifts in the recovery of $w$ compared to BBC, future work will focus on improving the parametrisation of intrinsic scatter model into our framework, with the goal of minimising the effect of the underlying scatter model on the recovery of cosmology.
-
-\begin{deluxetable}{lccc}
-	\tablecaption{$w$ bias comparison}
-	\tablecomments{We characterise the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model.
-We also show the results when combining the {\gten} and {\celeven} models into a combined set of 200 simulations.
-The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation.
-Averages are computed giving each simulation sample the same weight.
-In the model, $\Delta$ represents Steve - BBC.
-The final row shows the scatter between Steve and BBC for the different simulations.}
-	\tablehead{		Model & {\gten} & {\celeven} & ({\gten} + {\celeven}) }
-	\startdata
-		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
-		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$\tablenotemark{a}\\
-		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
-		$\Delta$  $\sigma_w$           & $0.057\pm0.004$ & $0.062\pm0.004$ & $0.060\pm0.003$ \\
-	\enddata
-	\tablenotetext{a}{This value is computed with each simulation having the same weight.
-It disagrees with the value provided in \citet[Table 10, row 3]{Brout18SYS} which uses inverse variance weighted averages.
-We do not utilise this weight because the variance is correlated with the value of $w$ due to the $\Omega_m$ prior applied in the fitting process.
-We note that if inverse variance weighting is applied to both datasets, they both shift by $\Delta w \approx 0.005$, and thus the predicted difference between the BBC method and {\steve} remains the same.}
-	\label{tab:bulk_summary_bbc}
-\end{deluxetable}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[width=\columnwidth]{bulk_comparison.pdf}
-	\end{center}
-	\caption{Recovered $w$ for the 200 validation simulations with full treatment of statistical and systematic errors.
-Uncertainty on the recovered $w$ value is shown for every second data point for visual clarity.}
-	\label{fig:bulk_comparison}
-\end{figure}
-
-\section{Conclusions}
-\label{sec:conclusion}
-
-In this paper we have outlined the creation of a hierarchical Bayesian model for supernova cosmology.
-The model takes into account selection effects and their uncertainty, fits underlying populations and standardisation parameters, incorporates unexplained dispersion from intrinsic scatter color smearing and incorporates uncertainty from peculiar velocities, survey calibration, HST calibration, dust, a potential global redshift offset, and SALT2 model uncertainty.
-Furthermore, our uncertainties in standardisation, population, mass-step and more, being explicitly parametrised in our model, are captured with covariance intact, an improvement on many previous methods.
-The model has been optimised to allow for hundreds of supernovae to be modelled fully with latent parameters.
-It runs in under an hour of CPU time and scales linearly with the number of supernovae, as opposed to polynomial complexity of matrix inversion of other methods.
-
-The importance of validating models using high-precision statistics gained by performing fits to hundreds of data realisations cannot be overstated, however this validation is lacking in many earlier BHM models for supernova cosmology.
-We have validated this model against many realisations of simplistic simulations with well-known and well-defined statistics, and found no cosmological bias.
-When validating using SNANA simulations, we find evidence of cosmological bias which is traced back to light curve fits reporting biased observables and incorrect covariance.
-Allowing fully parametrised corrections on observed supernovae summary statistics introduces too many degrees of freedom and is found to make cosmology fits too weak.
-Allowing simulation based corrections to vary in strength is found to give minor reductions in $w$ bias, however the uncertainty on the intrinsic scatter model itself limits the efficacy of the bias corrections.
-For the data size represented in the DES three-year spectroscopic survey, the determined biases should be sub-dominant to other sources of uncertainty, however this cannot be expected for future analyses with larger datasets.
-Stricter bias corrections calculated from simulations are required to reduce bias.
-Ideally, this would include further work on the calculation of intrinsic dispersion of the type Ia supernova population such that we can better characterise this bias.
-
-With our model being validated against hundreds of simulation realisations, representing a combined dataset over more than $60\,000$ simulated supernovae, we have been able to accurately determine biases in our model and trace their origin.
-With the current biases being sub-dominant to the total uncertainty, we now prepare to analyse the DES three-year dataset.
-
-\section*{Acknowledgements}
-
-Plots of posterior surfaces and parameter summaries were created with \verb|ChainConsumer| \citep{Hinton2016}.
-
-Funding for the DES Projects has been provided by the U.S. Department of Energy, the U.S. National Science Foundation, the Ministry of Science and Education of Spain, 
-the Science and Technology Facilities Council of the United Kingdom, the Higher Education Funding Council for England, the National Center for Supercomputing 
-Applications at the University of Illinois at Urbana-Champaign, the Kavli Institute of Cosmological Physics at the University of Chicago, 
-the Center for Cosmology and Astro-Particle Physics at the Ohio State University,
-the Mitchell Institute for Fundamental Physics and Astronomy at Texas A\&M University, Financiadora de Estudos e Projetos, 
-Funda{\c c}{\~a}o Carlos Chagas Filho de Amparo {\`a} Pesquisa do Estado do Rio de Janeiro, Conselho Nacional de Desenvolvimento Cient{\'i}fico e Tecnol{\'o}gico and 
-the Minist{\'e}rio da Ci{\^e}ncia, Tecnologia e Inova{\c c}{\~a}o, the Deutsche Forschungsgemeinschaft and the Collaborating Institutions in the Dark Energy Survey.
-
-The Collaborating Institutions are Argonne National Laboratory, the University of California at Santa Cruz, the University of Cambridge, Centro de Investigaciones Energ{\'e}ticas, 
-Medioambientales y Tecnol{\'o}gicas-Madrid, the University of Chicago, University College London, the DES-Brazil Consortium, the University of Edinburgh, 
-the Eidgen{\"o}ssische Technische Hochschule (ETH) Z{\"u}rich, 
-Fermi National Accelerator Laboratory, the University of Illinois at Urbana-Champaign, the Institut de Ci{\`e}ncies de l'Espai (IEEC/CSIC), 
-the Institut de F{\'i}sica d'Altes Energies, Lawrence Berkeley National Laboratory, the Ludwig-Maximilians Universit{\"a}t M{\"u}nchen and the associated Excellence Cluster Universe, 
-the University of Michigan, the National Optical Astronomy Observatory, the University of Nottingham, The Ohio State University, the University of Pennsylvania, the University of Portsmouth, 
-SLAC National Accelerator Laboratory, Stanford University, the University of Sussex, Texas A\&M University, and the OzDES Membership Consortium.
-
-Based in part on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory, which is operated by the Association of 
-Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation.
-
-The DES data management system is supported by the National Science Foundation under Grant Numbers AST-1138766 and AST-1536171.
-The DES participants from Spanish institutions are partially supported by MINECO under grants AYA2015-71825, ESP2015-66861, FPA2015-68048, SEV-2016-0588, SEV-2016-0597, and MDM-2015-0509, 
-some of which include ERDF funds from the European Union. IFAE is partially funded by the CERCA program of the Generalitat de Catalunya.
-Research leading to these results has received funding from the European Research
-Council under the European Union's Seventh Framework Program (FP7/2007-2013) including ERC grant agreements 240672, 291329, and 306478.
-We  acknowledge support from the Australian Research Council Centre of Excellence for All-sky Astrophysics (CAASTRO), through project number CE110001020, and the Brazilian Instituto Nacional de Ci\^encia
-e Tecnologia (INCT) e-Universe (CNPq grant 465376/2014-2).
-
-This manuscript has been authored by Fermi Research Alliance, LLC under Contract No.
-DE-AC02-07CH11359 with the U.S.
-Department of Energy, Office of Science, Office of High Energy Physics.
-The United States Government retains and the publisher, by accepting the article for publication, acknowledges that the United States Government retains a non-exclusive, paid-up, irrevocable, world-wide license to publish or reproduce the published form of this manuscript, or allow others to do so, for United States Government purposes.
-
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-% The best way to enter references is to use BibTeX:
-
-\bibliographystyle{mnras}
-\bibliography{bib}
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Selection Effect Derivation}
-\label{app:selection}
-
-\subsection{General Selection Effects}
-\label{app:selection1}
-When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
-\begin{equation}
-P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
-\end{equation}
-which gives the probability of the model parameter values ($\theta$) given the data.
-Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$.
-Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$.
-When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.
-We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.
-To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
-\begin{equation}
-\mathcal{L}(\theta; {\rm data}) = P({\rm data} | \theta, S). \label{eq:app_like}
-\end{equation}
-A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.
-That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}.
-We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
-\begin{eqnarray}
-P({\rm data} | S , \theta) P(S, \theta) &=& P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
-P({\rm data} | S , \theta) &=& \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
-&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
-&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
-\end{eqnarray}
-which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
-\begin{eqnarray}
-\mathcal{L}(\theta; {\rm data})&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
-\end{eqnarray}
-where we define the denominator as $d$ for simplicity in future derivations.
-
-\subsection{Supernova Selection Effects}
-\label{app:selection2}
-We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
-Our denominator then becomes
-\begin{equation}
-d = \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
-\end{equation}
-where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly.
-Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-$P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$ (which is why they were not included in equation \eqref{eq:sel}).
-Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.
-We can thus state
-\begin{equation}
-d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
-\end{equation}
-By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
-\begin{equation}
-\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
-\end{equation}
-for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.
-
-As stated in Section \ref{sec:selection}, the underlying population in apparent magnitude, when we discard skewness, can be represented as $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
-\begin{eqnarray}
-m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \left(\langle c(z) \rangle + \sqrt{\frac{2}{\pi}}\sigma_c \delta_c\right)\label{eq:disc1} \\
-\sigma^*_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 +  \left(\beta \sigma_c \sqrt{1 - \frac{2\delta_c^2}{\pi}}\right)^2. \label{eq:disc2}
-\end{eqnarray}
-Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.
-
-\subsection{Approximate Selection Effects}
-\label{app:approx}
-
-In this section, we investigate the effect of approximating the skew normal underlying color distribution as a normal.
-Specifically, equations \eqref{eq:disc1} and \eqref{eq:disc2} make the assumption that, for our color distribution, $\mathcal{N}^{\rm Skew}(\mu, \sigma, \alpha)$ is well approximated by $\mathcal{N}(\mu, \sigma)$.
-We sought to improve on this approximation by adjusting the mean and standard deviation of the approximated normal to more accurately describe the actual mean and standard deviation of a skew normal.
-With $\delta \equiv \alpha/\sqrt{1+\alpha^2}$, the correct mean and standard deviation are
-\begin{eqnarray}
-\mu_1 &=& \mu_0 + \sqrt{\frac{2}{\pi}} \delta \sigma_0 \\
-\sigma_1 &=& \sigma_0 \sqrt{1 - \frac{2 \delta^2}{\pi}},
-\end{eqnarray}
-where we highlight that $\mu$ here represents the mean of the distribution, not distance modulus.
-We can then test the approximation $\mathcal{N}^{\rm Skew}(\mu_0, \sigma_0, \alpha) \rightarrow \mathcal{N}(\mu_1, \sigma_1)$.
-Unfortunately, this shift to the mean and standard deviation of the normal approximation where we treat $m_B$, $x_1$, and $c$ as a multivariate skew normal did not produce stable posterior surfaces.
-Due to this, we treat the underlying $m_B$, $x_1$, and $c$ populations as independent.
-
-We tested a fixed $\sigma_c$ in the shift correction, such that $\mu_1 = \mu_0 + \sqrt{2/\pi}\delta k$, where we set $k=0.1$ to mirror the width of the input simulation population.
-This resulted in stable posterior surfaces, however this introduced recovery bias in several population parameters, and so we do not fix $\sigma_c$.
-Comparing whether we shift our normal in the approximation or simply discard skewness, Figure \ref{fig:shift} shows that the calculated efficiency is significantly discrepant to the actual efficiency if the normal approximation is not shifted.
-The biases when using shifted or unshifted normal approximations when we fit our model on Gaussian and skewed underlying populations are shown in Figure \ref{fig:simple_w_super}, and only the shifted normal approximation correctly recovers underlying population parameters.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=\textwidth]{simple_w_shift_dist_0.pdf}
-	\end{center}
-	\caption{Marginalised probability distributions for 100 realisations of cosmology, fit to Flat $w$CDM with prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$, each containing 1000 simulated high-$z$ and 1000 simulated low-$z$ supernovae.
-The dashed green surfaces represent a fit to an underlying Gaussian color population with the unshifted model.
-The blue solid surface represents fits to a skewed color population with the unshifted model, and the purple dotted surface represents a fit to a skewed color population with the shifted model.
-The superscript $0$ and $1$ denote the two different surveys (high-$z$ and low-$z$ respectively), and similarly the first four $\langle c^i \rangle$ parameters represent the four redshift nodes in the high-$z$ survey, and the last four represent the nodes for the low-$z$ survey.
-We can see that the shifted model is far better able to recover skewed input populations than the unshifted, performing better in terms of recovering skewness $\alpha_c$, mean color $\langle c \rangle$ and width of the color distribution $\sigma_c$.
-The unshifted model recovers the correct color mean and width if you approximate a skew normal as a normal: $\Delta\mu = \sqrt{2/\pi}\sigma_c\delta_c \approx 0.071$, which is approximately the deviation found in fits to the color population mean.
-Importantly, the unshifted model when run on skewed data (the solid blue) shows extreme bias in $\alpha_c$, where it fits strongly around zero regardless, showing it to be a poor approximation.
-Based on these results and the good performance in correctly recovering underlying populations of the shifted normal approximation, we adopt the shifted normal approximation in our model.}
-	\label{fig:simple_w_super}
-\end{figure*}
-
-\section{Numerical Optimisations}
-\label{app:optimisations}
-
-Not many fitting methodologies and algorithms can handle the thousands of fit parameters our model requires.
-By using Stan, we are able to take advantage automatic differentiation and the NUTS sampler, which is a class of Hamiltonian Monte Carlo samplers.
-Even with these advantages, early implementations of our model still had excessive fit times, with our desired sub-hour running time far exceeded.
-
-The simplest and most commonly found optimisation we employed was to precompute as much as possible.
-This is in a bid to reduce the complexity of the mathematical graph our model is translated into by Stan, to simplify the computation of surface derivatives.
-For example, when computing the distance modulus, redshift is encountered to various powers.
-Instead of computing those powers in Stan, we simply pass in several arrays of redshift values already raised to the correct power.
-Small changes like this however only give small improvements.
-
-The primary numerical improvement we made on existing frameworks was to remove costly probability evaluations of multivariate normals.
-To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, we sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$.
-In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution.
-Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to thirty minutes.
-
-This parametrisation does come with one significant downside --- inflexibility.
-For each step the algorithm takes, we do not recompute the cholesky decomposition of the covariance of the summary statistics --- that happens once at the beginning of the model setup.
-If we had kept the full covariance matrix parametrisation we could modify the matrix easily --- for example when incorporating intrinsic dispersion we could simply add on a secondary matrix to create an updated covariance.
-However as the cholesky decomposition of a sum of matrices is not equal to the sum of the cholesky decomposition of each individual matrix, we would need to recompute the decomposition for each step, which discards most of the computational benefit just gained.
-
-Considering a $3\times3$ matrix with cholesky decomposition
-\begin{equation}
-L = \begin{pmatrix}
-a & 0 & 0 \\ b & c & 0 \\ d & e & f \\
-\end{pmatrix},
-\end{equation}
-the original covariance matrix $\Sigma$ is given by
-\begin{equation}
-\Sigma = \begin{pmatrix}
-a^2 & ab & ad \\ ab & b^2 + c^2 & bd + ce \\ ad & bd + ce & d^2 + e^2 + f^2\\
-\end{pmatrix}.
-\end{equation}
-Now, the primary source of extra uncertainty in the intrinsic dispersion models comes from chromatic smearing, which primarily influences the recovered color parameter, which is placed as the last element in the observables vector $\lbrace m_B, x_1, c\rbrace$.
-We can now see that it is possible to add extra uncertainty to the color observation on the diagonal without having to recompute the cholesky decomposition - notice that $f$ is unique in that it is the only element of $L$ that appears in only one position in the covariance matrix.
-To take our covariance and add on the diagonal uncertainty for color an extra $\sigma_e$ term, we get
-\begin{equation}
-C = \begin{pmatrix}
-\sigma_{m_B}^2 & \rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \rho_{0,2} \sigma_{m_B} \sigma_c \\
-\rho_{0,1} \sigma_{m_B} \sigma_{x_1} & \sigma_{x_1}^2 & \rho_{1, 2} \sigma_{x_1} \sigma_c \\
-\rho_{0,2} \sigma_{m_B} \sigma_c & \rho_{1, 2} \sigma_{x_1} \sigma_c &  \sigma_c^2 + \sigma_e^2 \\
-\end{pmatrix}.
-\end{equation}
-The cholesky decomposition of this is, in terms of the original cholesky decomposition, is
-\begin{equation}
-L = \begin{pmatrix}
-a & 0 & 0 \\ b & c & 0 \\ d & e & f + g \\
-\end{pmatrix},
-\end{equation}
-where $g = \sqrt{f^2 + \sigma_e^2} - f$.
-This allows an easy update to the cholesky decomposition to add extra uncertainty to the independent color uncertainty.
-For both the {\gten} and {\celeven} models, we ran fits without the cholesky parametrisation to allow for extra correlated dispersion (instead of just dispersion on $c$), but find no decrease in bias or improved fit statistics, allowing us to use the more efficient cholesky parametrisation.
-
+\maketitle
 \end{document}
-
-% End of mnras_template.tex

### 2018-12-20 papers/methods/misc/author_list_arxiv.tex
Adding author list

% Author list file generated with: mkauthlist.py 1.2.4 
% mkauthlist.py -j arxiv --sort-nonbuilder -a order.csv author_list.csv author_list_arxiv.tex 
S. R. Hinton, T. M. Davis, A. G. Kim, D. Brout, C. B. D'Andrea, R. Kessler, J. Lasker, C. Lidman, E. Macaulay, A. M\"oller, M. Sako, D. Scolnic, M. Smith, R. C. Wolf, M. Childress, E. Morganson, S. Allam, J. Annis, S. Avila, E. Bertin, D. Brooks, D. L. Burke, A. Carnero Rosell, M. Carrasco Kind, J. Carretero, C. E. Cunha, L. N. da Costa, C. Davis, J. De Vicente, D. L. DePoy, P. Doel, T. F. Eifler, B. Flaugher, P. Fosalba, J. Frieman, J. Garc\'ia-Bellido, E. Gaztanaga, D. W. Gerdes, R. A. Gruendl, J. Gschwend, G. Gutierrez, W. G. Hartley, D. L. Hollowood, K. Honscheid, E. Krause, K. Kuehn, N. Kuropatkin, O. Lahav, M. Lima, M. A. G. Maia, M. March, J. L. Marshall, F. Menanteau, R. Miquel, R. L. C. Ogando, A. A. Plazas, E. Sanchez, V. Scarpine, R. Schindler, M. Schubnell, S. Serrano, I. Sevilla-Noarbe, M. Soares-Santos, F. Sobreira, E. Suchyta, G. Tarle, D. Thomas, V. Vikram, Y. Zhang (DES Collaboration)

### 2018-12-20 papers/methods/paper.tex

@@
 One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
-We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
+We describe the probability that the {\blue possible} events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
@@
 where the denominator represents an integral over all potential data $D$.
-This is derived in Appendix \ref{app:selection1} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
+This is derived in Appendix \ref{app:selection1}.
+{\blue To evaluate the effect of our selection effects, we need to evaluate this integral.
+The integral represents our global selection efficiency at a location in parameter space, rather than the probability of our data being selected into our sample.} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
 Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
@@
 &=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
-&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) }
+&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) } \label{eq:general1}
 \end{eqnarray}
@@
 \label{app:selection2}
-We assume that our selection effects can be reasonably well encapsulated by independent functions of (true) apparent magnitude and redshift, such that $P(S|{\rm data}, \theta) = P(S|z) P(S|m_B)$.
+{\blue
+To turn the generalised equations from the previous sections into selection effects relevant for this model, we first need to write out our latent parameters, as $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc).
+Whilst they can be inserted into the denominator of equation \eqref{eq:general1} to get $P(S|\theta) = \int P(S,L|\theta)\, dL$ and expanded out, we can simply write down our denominator with general latent parameter(s) $L$, as such:
+\begin{equation}
+d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL
+\end{equation}
+Here we can now move from the generic labels $D$ and $L$ to something specific to our model.
+In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ However, as we assume that our selection effects can be sufficiently encapsulated by true apparent magnitude $m_B$ and redshift $z$, we will not write out the other parameters for simplicity.
+If we did, we'd simply find them integrated to unity.
+Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
 Our denominator then becomes
 \begin{equation}
-d = \int d\hat{z} \, d\hat{m}_B \, dz \, dm_B P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta), \label{eq:sel}
-\end{equation}
-where for simplicity we have not written out all the integrals that do not interact with the selection effects explicitly.
+d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
+\end{equation}
 Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-$P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and $m_B$, $x_1$, $c$, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$ (which is why they were not included in equation \eqref{eq:sel}).
+$P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$.
 Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.

### 2018-12-20 papers/methods/figures/tikz.tex
Updating PGM

@@
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[det, right=4.33cm of mu] (MBobs) {$M_B$};
+	\node[det, right=4.33cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
 	% Survey nodes
-	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
-	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
+	\node[latent, above=of sum, yshift=1.6cm, xshift=1cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
+	\node[latent, right=1.2cm of pop1, align=center] (sf) {$\delta S$};
+	\node[latent, left=0.8cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
 	% Global nodes
-	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
+	\node[latent, above=2.3cm of mu]  (cosmo) {$\Omega_m, w$};
+	\node[latent, above=1.7cm of MBobs, text width=2.5cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$};
+	\node[latent, right=0.5cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
 	% Connect the nodes
@@
 	\edge {z,cosmo} {mu};
-	\edge {mu, sum, pop1, ab, sf, pobs, mobs} {MBobs};	
+	\edge {mu, pop1, ab, sf} {MBobs};	
+	\edge {MBobs} {pobs, mobs};
+	\edge {dz, kappa, MBobs} {sum};
 	% Plates
-	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
-	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
+	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)(MBobs)} {$\forall\ \rm{SN}$} ;
+	\plated[thick] {survey} {(sn)(pop1)(kappa)(sf)} {\\$\forall\ \rm{Survey}$} ;
 	\end{tikzpicture}

### 2018-12-20 papers/methods/misc/author_list.tex
Updating PGM

@@
 % Author list file generated with: mkauthlist.py 1.2.4 
-% mkauthlist.py -f --doc -j emulateapj --sort --sort-builder -a order.csv author_list.csv author_list.tex
+% mkauthlist.py -j apj --sort-nonbuilder -a order.csv author_list.csv author_list.tex
 
-\documentclass[iop]{emulateapj}
+\author{S.~R.~Hinton}
+\affiliation{School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
+\author{T.~M.~Davis}
+\affiliation{School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
+\author{A.~G.~Kim}
+\affiliation{Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
+\author{D.~Brout}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{C.~B.~D'Andrea}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{R.~Kessler}
+\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{J.~Lasker}
+\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{C.~Lidman}
+\affiliation{The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\author{E.~Macaulay}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{A.~M\"oller}
+\affiliation{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affiliation{The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\author{M.~Sako}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{D.~Scolnic}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{M.~Smith}
+\affiliation{School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\author{R.~C.~Wolf}
+\affiliation{Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
+\author{M.~Childress}
+\affiliation{School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\author{E.~Morganson}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{S.~Allam}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{J.~Annis}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{S.~Avila}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{E.~Bertin}
+\affiliation{CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\affiliation{Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\author{D.~Brooks}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{D.~L.~Burke}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\affiliation{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
+\author{A.~Carnero~Rosell}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~Carrasco~Kind}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{J.~Carretero}
+\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
+\author{C.~E.~Cunha}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\author{L.~N.~da Costa}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{C.~Davis}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\author{J.~De~Vicente}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{D.~L.~DePoy}
+\affiliation{George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
+\author{P.~Doel}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{T.~F.~Eifler}
+\affiliation{Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
+\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
+\author{B.~Flaugher}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{P.~Fosalba}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{J.~Frieman}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{J.~Garc\'ia-Bellido}
+\affiliation{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
+\author{E.~Gaztanaga}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{D.~W.~Gerdes}
+\affiliation{Department of Astronomy, University of Michigan, Ann Arbor, MI 48109, USA}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{R.~A.~Gruendl}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{J.~Gschwend}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{G.~Gutierrez}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{W.~G.~Hartley}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\affiliation{Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 16, CH-8093 Zurich, Switzerland}
+\author{D.~L.~Hollowood}
+\affiliation{Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
+\author{K.~Honscheid}
+\affiliation{Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
+\affiliation{Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
+\author{E.~Krause}
+\affiliation{Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
+\author{K.~Kuehn}
+\affiliation{Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia}
+\author{N.~Kuropatkin}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{O.~Lahav}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{M.~Lima}
+\affiliation{Departamento de F\'isica Matem\'atica, Instituto de F\'isica, Universidade de S\~ao Paulo, CP 66318, S\~ao Paulo, SP, 05314-970, Brazil}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~A.~G.~Maia}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~March}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{J.~L.~Marshall}
+\affiliation{George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
+\author{F.~Menanteau}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{R.~Miquel}
+\affiliation{Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
+\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
+\author{R.~L.~C.~Ogando}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{A.~A.~Plazas}
+\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
+\author{E.~Sanchez}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{V.~Scarpine}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{R.~Schindler}
+\affiliation{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
+\author{M.~Schubnell}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{S.~Serrano}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{I.~Sevilla-Noarbe}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{M.~Soares-Santos}
+\affiliation{Brandeis University, Physics Department, 415 South Street, Waltham MA 02453}
+\author{F.~Sobreira}
+\affiliation{Instituto de F\'isica Gleb Wataghin, Universidade Estadual de Campinas, 13083-859, Campinas, SP, Brazil}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{E.~Suchyta}
+\affiliation{Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
+\author{G.~Tarle}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{D.~Thomas}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{V.~Vikram}
+\affiliation{Argonne National Laboratory, 9700 South Cass Avenue, Lemont, IL 60439, USA}
+\author{Y.~Zhang}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
 
-\begin{document}
-\title{DES Publication Title}
-
-\def\andname{}
-
-\author{
-S.~R.~Hinton\altaffilmark{1},
-T.~M.~Davis\altaffilmark{1},
-A.~G.~Kim\altaffilmark{2},
-D.~Brout\altaffilmark{3},
-C.~B.~D'Andrea\altaffilmark{3},
-R.~Kessler\altaffilmark{4,5},
-J.~Lasker\altaffilmark{4,5},
-C.~Lidman\altaffilmark{6},
-E.~Macaulay\altaffilmark{7},
-A.~M\"oller\altaffilmark{8,6},
-M.~Sako\altaffilmark{3},
-D.~Scolnic\altaffilmark{5},
-M.~Smith\altaffilmark{9},
-R.~C.~Wolf\altaffilmark{10},
-S.~Allam\altaffilmark{11},
-J.~Annis\altaffilmark{11},
-S.~Avila\altaffilmark{7},
-E.~Bertin\altaffilmark{12,13},
-D.~Brooks\altaffilmark{14},
-D.~L.~Burke\altaffilmark{15,16},
-A.~Carnero~Rosell\altaffilmark{17,18},
-M.~Carrasco~Kind\altaffilmark{19,20},
-J.~Carretero\altaffilmark{21},
-M.~Childress\altaffilmark{9},
-C.~E.~Cunha\altaffilmark{15},
-L.~N.~da Costa\altaffilmark{18,22},
-C.~Davis\altaffilmark{15},
-J.~De~Vicente\altaffilmark{17},
-D.~L.~DePoy\altaffilmark{23},
-P.~Doel\altaffilmark{14},
-T.~F.~Eifler\altaffilmark{24,25},
-B.~Flaugher\altaffilmark{11},
-P.~Fosalba\altaffilmark{26,27},
-J.~Frieman\altaffilmark{11,5},
-J.~Garc\'ia-Bellido\altaffilmark{28},
-E.~Gaztanaga\altaffilmark{26,27},
-D.~W.~Gerdes\altaffilmark{29,30},
-R.~A.~Gruendl\altaffilmark{19,20},
-J.~Gschwend\altaffilmark{18,22},
-G.~Gutierrez\altaffilmark{11},
-W.~G.~Hartley\altaffilmark{14,31},
-D.~L.~Hollowood\altaffilmark{32},
-K.~Honscheid\altaffilmark{33,34},
-E.~Krause\altaffilmark{24},
-K.~Kuehn\altaffilmark{35},
-N.~Kuropatkin\altaffilmark{11},
-O.~Lahav\altaffilmark{14},
-M.~Lima\altaffilmark{36,18},
-M.~A.~G.~Maia\altaffilmark{18,22},
-M.~March\altaffilmark{3},
-J.~L.~Marshall\altaffilmark{23},
-F.~Menanteau\altaffilmark{19,20},
-R.~Miquel\altaffilmark{37,21},
-E.~Morganson\altaffilmark{20},
-R.~L.~C.~Ogando\altaffilmark{18,22},
-A.~A.~Plazas\altaffilmark{25},
-E.~Sanchez\altaffilmark{17},
-V.~Scarpine\altaffilmark{11},
-R.~Schindler\altaffilmark{16},
-M.~Schubnell\altaffilmark{30},
-S.~Serrano\altaffilmark{26,27},
-I.~Sevilla-Noarbe\altaffilmark{17},
-M.~Soares-Santos\altaffilmark{38},
-F.~Sobreira\altaffilmark{39,18},
-E.~Suchyta\altaffilmark{40},
-G.~Tarle\altaffilmark{30},
-D.~Thomas\altaffilmark{7},
-V.~Vikram\altaffilmark{41},
-and Y.~Zhang\altaffilmark{11}
-\\ \vspace{0.2cm} (DES Collaboration) \\
-}
-
-\affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
-\affil{$^{2}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
-\affil{$^{3}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
-\affil{$^{4}$ Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{5}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{6}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
-\affil{$^{7}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
-\affil{$^{8}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
-\affil{$^{9}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
-\affil{$^{10}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
-\affil{$^{11}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
-\affil{$^{12}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
-\affil{$^{13}$ Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
-\affil{$^{14}$ Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
-\affil{$^{15}$ Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
-\affil{$^{16}$ SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
-\affil{$^{17}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
-\affil{$^{18}$ Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
-\affil{$^{19}$ Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
-\affil{$^{20}$ National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
-\affil{$^{21}$ Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
-\affil{$^{22}$ Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
-\affil{$^{23}$ George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
-\affil{$^{24}$ Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
-\affil{$^{25}$ Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
-\affil{$^{26}$ Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
-\affil{$^{27}$ Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
-\affil{$^{28}$ Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
-\affil{$^{29}$ Department of Astronomy, University of Michigan, Ann Arbor, MI 48109, USA}
-\affil{$^{30}$ Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
-\affil{$^{31}$ Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 16, CH-8093 Zurich, Switzerland}
-\affil{$^{32}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
-\affil{$^{33}$ Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
-\affil{$^{34}$ Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
-\affil{$^{35}$ Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia}
-\affil{$^{36}$ Departamento de F\'isica Matem\'atica, Instituto de F\'isica, Universidade de S\~ao Paulo, CP 66318, S\~ao Paulo, SP, 05314-970, Brazil}
-\affil{$^{37}$ Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
-\affil{$^{38}$ Brandeis University, Physics Department, 415 South Street, Waltham MA 02453}
-\affil{$^{39}$ Instituto de F\'isica Gleb Wataghin, Universidade Estadual de Campinas, 13083-859, Campinas, SP, Brazil}
-\affil{$^{40}$ Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
-\affil{$^{41}$ Argonne National Laboratory, 9700 South Cass Avenue, Lemont, IL 60439, USA}
-
-\begin{abstract}
-This is a sample document created by \texttt{mkauthlist.py v1.2.4}.
-\end{abstract}
-\maketitle
-\end{document}
+\collaboration{(DES Collaboration)}

### 2018-12-20 papers/methods/paper.tex
Updating PGM

@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib,manuscript]{emulateapj}
+\documentclass[twocolumn,trackchanges,tighten]{aastex62}
 
@@
 %\linenumbers
-\renewcommand{\arraystretch}{1.2}
 %\graphicspath{ {./figures/} }
@@
 
-\author{
-S.~R.~Hinton\altaffilmark{1},
-T.~M.~Davis\altaffilmark{1},
-A.~G.~Kim\altaffilmark{2},
-D.~Brout\altaffilmark{3},
-C.~B.~D'Andrea\altaffilmark{3},
-R.~Kessler\altaffilmark{4,5},
-J.~Lasker\altaffilmark{4,5},
-C.~Lidman\altaffilmark{6},
-E.~Macaulay\altaffilmark{7},
-A.~M\"oller\altaffilmark{8,6},
-M.~Sako\altaffilmark{3},
-D.~Scolnic\altaffilmark{5},
-M.~Smith\altaffilmark{9},
-R.~C.~Wolf\altaffilmark{10},
-S.~Allam\altaffilmark{11},
-J.~Annis\altaffilmark{11},
-S.~Avila\altaffilmark{7},
-E.~Bertin\altaffilmark{12,13},
-D.~Brooks\altaffilmark{14},
-D.~L.~Burke\altaffilmark{15,16},
-A.~Carnero~Rosell\altaffilmark{17,18},
-M.~Carrasco~Kind\altaffilmark{19,20},
-J.~Carretero\altaffilmark{21},
-M.~Childress\altaffilmark{9},
-C.~E.~Cunha\altaffilmark{15},
-L.~N.~da Costa\altaffilmark{18,22},
-C.~Davis\altaffilmark{15},
-J.~De~Vicente\altaffilmark{17},
-D.~L.~DePoy\altaffilmark{23},
-P.~Doel\altaffilmark{14},
-T.~F.~Eifler\altaffilmark{24,25},
-B.~Flaugher\altaffilmark{11},
-P.~Fosalba\altaffilmark{26,27},
-J.~Frieman\altaffilmark{11,5},
-J.~Garc\'ia-Bellido\altaffilmark{28},
-E.~Gaztanaga\altaffilmark{26,27},
-D.~W.~Gerdes\altaffilmark{29,30},
-R.~A.~Gruendl\altaffilmark{19,20},
-J.~Gschwend\altaffilmark{18,22},
-G.~Gutierrez\altaffilmark{11},
-W.~G.~Hartley\altaffilmark{14,31},
-D.~L.~Hollowood\altaffilmark{32},
-K.~Honscheid\altaffilmark{33,34},
-E.~Krause\altaffilmark{24},
-K.~Kuehn\altaffilmark{35},
-N.~Kuropatkin\altaffilmark{11},
-O.~Lahav\altaffilmark{14},
-M.~Lima\altaffilmark{36,18},
-M.~A.~G.~Maia\altaffilmark{18,22},
-M.~March\altaffilmark{3},
-J.~L.~Marshall\altaffilmark{23},
-F.~Menanteau\altaffilmark{19,20},
-R.~Miquel\altaffilmark{37,21},
-E.~Morganson\altaffilmark{20},
-R.~L.~C.~Ogando\altaffilmark{18,22},
-A.~A.~Plazas\altaffilmark{25},
-E.~Sanchez\altaffilmark{17},
-V.~Scarpine\altaffilmark{11},
-R.~Schindler\altaffilmark{16},
-M.~Schubnell\altaffilmark{30},
-S.~Serrano\altaffilmark{26,27},
-I.~Sevilla-Noarbe\altaffilmark{17},
-M.~Soares-Santos\altaffilmark{38},
-F.~Sobreira\altaffilmark{39,18},
-E.~Suchyta\altaffilmark{40},
-G.~Tarle\altaffilmark{30},
-D.~Thomas\altaffilmark{7},
-V.~Vikram\altaffilmark{41},
-and Y.~Zhang\altaffilmark{11}
-\\ \vspace{0.2cm} (DES Collaboration) \\
-}
-
-\affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
-\affil{$^{2}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
-\affil{$^{3}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
-\affil{$^{4}$ Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{5}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
-\affil{$^{6}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
-\affil{$^{7}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
-\affil{$^{8}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
-\affil{$^{9}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
-\affil{$^{10}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
-\affil{$^{11}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
-\affil{$^{12}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
-\affil{$^{13}$ Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
-\affil{$^{14}$ Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
-\affil{$^{15}$ Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
-\affil{$^{16}$ SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
-\affil{$^{17}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
-\affil{$^{18}$ Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
-\affil{$^{19}$ Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
-\affil{$^{20}$ National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
-\affil{$^{21}$ Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
-\affil{$^{22}$ Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
-\affil{$^{23}$ George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
-\affil{$^{24}$ Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
-\affil{$^{25}$ Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
-\affil{$^{26}$ Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
-\affil{$^{27}$ Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
-\affil{$^{28}$ Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
-\affil{$^{29}$ Department of Astronomy, University of Michigan, Ann Arbor, MI 48109, USA}
-\affil{$^{30}$ Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
-\affil{$^{31}$ Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 16, CH-8093 Zurich, Switzerland}
-\affil{$^{32}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
-\affil{$^{33}$ Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
-\affil{$^{34}$ Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
-\affil{$^{35}$ Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia}
-\affil{$^{36}$ Departamento de F\'isica Matem\'atica, Instituto de F\'isica, Universidade de S\~ao Paulo, CP 66318, S\~ao Paulo, SP, 05314-970, Brazil}
-\affil{$^{37}$ Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
-\affil{$^{38}$ Brandeis University, Physics Department, 415 South Street, Waltham MA 02453}
-\affil{$^{39}$ Instituto de F\'isica Gleb Wataghin, Universidade Estadual de Campinas, 13083-859, Campinas, SP, Brazil}
-\affil{$^{40}$ Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
-\affil{$^{41}$ Argonne National Laboratory, 9700 South Cass Avenue, Lemont, IL 60439, USA}
+\author{S.~R.~Hinton}
+\affiliation{School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
+\author{T.~M.~Davis}
+\affiliation{School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
+\author{A.~G.~Kim}
+\affiliation{Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
+\author{D.~Brout}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{C.~B.~D'Andrea}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{R.~Kessler}
+\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{J.~Lasker}
+\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{C.~Lidman}
+\affiliation{The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\author{E.~Macaulay}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{A.~M\"oller}
+\affiliation{ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
+\affiliation{The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
+\author{M.~Sako}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{D.~Scolnic}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{M.~Smith}
+\affiliation{School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\author{R.~C.~Wolf}
+\affiliation{Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
+\author{M.~Childress}
+\affiliation{School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
+\author{E.~Morganson}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{S.~Allam}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{J.~Annis}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{S.~Avila}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{E.~Bertin}
+\affiliation{CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\affiliation{Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France}
+\author{D.~Brooks}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{D.~L.~Burke}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\affiliation{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
+\author{A.~Carnero~Rosell}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~Carrasco~Kind}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{J.~Carretero}
+\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
+\author{C.~E.~Cunha}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\author{L.~N.~da Costa}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{C.~Davis}
+\affiliation{Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA}
+\author{J.~De~Vicente}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{D.~L.~DePoy}
+\affiliation{George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
+\author{P.~Doel}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{T.~F.~Eifler}
+\affiliation{Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
+\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
+\author{B.~Flaugher}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{P.~Fosalba}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{J.~Frieman}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\affiliation{Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
+\author{J.~Garc\'ia-Bellido}
+\affiliation{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain}
+\author{E.~Gaztanaga}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{D.~W.~Gerdes}
+\affiliation{Department of Astronomy, University of Michigan, Ann Arbor, MI 48109, USA}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{R.~A.~Gruendl}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{J.~Gschwend}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{G.~Gutierrez}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{W.~G.~Hartley}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\affiliation{Department of Physics, ETH Zurich, Wolfgang-Pauli-Strasse 16, CH-8093 Zurich, Switzerland}
+\author{D.~L.~Hollowood}
+\affiliation{Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA}
+\author{K.~Honscheid}
+\affiliation{Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA}
+\affiliation{Department of Physics, The Ohio State University, Columbus, OH 43210, USA}
+\author{E.~Krause}
+\affiliation{Department of Astronomy/Steward Observatory, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA}
+\author{K.~Kuehn}
+\affiliation{Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia}
+\author{N.~Kuropatkin}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{O.~Lahav}
+\affiliation{Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK}
+\author{M.~Lima}
+\affiliation{Departamento de F\'isica Matem\'atica, Instituto de F\'isica, Universidade de S\~ao Paulo, CP 66318, S\~ao Paulo, SP, 05314-970, Brazil}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~A.~G.~Maia}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{M.~March}
+\affiliation{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
+\author{J.~L.~Marshall}
+\affiliation{George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA}
+\author{F.~Menanteau}
+\affiliation{Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA}
+\affiliation{National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA}
+\author{R.~Miquel}
+\affiliation{Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain}
+\affiliation{Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain}
+\author{R.~L.~C.~Ogando}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\affiliation{Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{A.~A.~Plazas}
+\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Dr., Pasadena, CA 91109, USA}
+\author{E.~Sanchez}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{V.~Scarpine}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
+\author{R.~Schindler}
+\affiliation{SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA}
+\author{M.~Schubnell}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{S.~Serrano}
+\affiliation{Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain}
+\affiliation{Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain}
+\author{I.~Sevilla-Noarbe}
+\affiliation{Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain}
+\author{M.~Soares-Santos}
+\affiliation{Brandeis University, Physics Department, 415 South Street, Waltham MA 02453}
+\author{F.~Sobreira}
+\affiliation{Instituto de F\'isica Gleb Wataghin, Universidade Estadual de Campinas, 13083-859, Campinas, SP, Brazil}
+\affiliation{Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil}
+\author{E.~Suchyta}
+\affiliation{Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831}
+\author{G.~Tarle}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+\author{D.~Thomas}
+\affiliation{Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
+\author{V.~Vikram}
+\affiliation{Argonne National Laboratory, 9700 South Cass Avenue, Lemont, IL 60439, USA}
+\author{Y.~Zhang}
+\affiliation{Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
 
@@
 As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement {\blue as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement \added{as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus.
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
@@
 \begin{deluxetable}{ll}
+	\renewcommand{\arraystretch}{0.8}
 	\tablecolumns{2}
@@
 	\startdata	
-
 	&\textbf{Global Parameters} \\
-
 	$\Omega_m$  & Matter density  \\
@@
 	& \textbf{Survey Parameters} \\
-
 	$\delta S$ & Selection effect deviation \\
@@
 	& \textbf{Input Data}\tablenotemark{a} \\
-
 	$\hat{m}_B$ & Measured flux\\
@@
 The SN box represents observed variables and latent variables for each individual supernova, whilst the survey box represents survey-specific variables, which in general describe the supernova population for the survey and the systematics associated with it.
-Variables that appear outside both boxes represent top level model parameters.}
+Variables that appear outside both boxes represent top level model parameters.
+\added{To solidify that the underlying supernova population lives in $\lbrace M_B, x_1, c\rbrace$ space, we write out the stretch and colour populations along with $M_B$, however note here that those $x_1$ and $c$ values (denoted by an asterix) are identically equal to the $x_1$ and $c$ values in the latent node they are conditioned on, and in the code implementation, they are the same variable.}}
  	\label{fig:pgm}
@@
 Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space.
-{\blue This is mathematically equivalent, however allows us to model the underlying population and intrinsic scatter distinctly.} To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
+\added{This is mathematically equivalent, however allows us to model the underlying population and intrinsic scatter distinctly.} To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.
 
@@
 We sample $\delta_c$ in log space for efficiency in sampling close to zero.
-The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors with {\blue mean 0 and width 1}.
+The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors \added{with mean 0 and width 1, following recommendations from the Stan user guide}.
 
@@
 
-We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM {\blue (without a flatness prior)}.
+We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM \added{ (without a flatness prior)}.
 $\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$.
@@
 The chief difficulty with including systematic uncertainties in supernova analyses is that they have difficult-to-model effects on the output observations.
-As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$ {\blue for each supernova light curve fit}, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
+As such, the traditional treatment for systematics is to compute their effect on the supernova summary statistics -- computing the numerical derivatives $\frac{{\rm d} \hat{m}_B}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{x}_1}{{\rm d} \Z_i}$, $\frac{{\rm d} \hat{c}}{{\rm d} \Z_i}$ \added{ for each supernova light curve fit}, where $\Z_i$ represents the $i$\textsuperscript{th} systematic.
 
@@
 A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details.
-This gives thirteen global systematics shared by {\blue all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey.
+This gives thirteen global systematics shared by \added{ all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey.
 With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
@@
 One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections.
-We describe the probability that the {\blue possible} events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
+We describe the probability that the \added{ possible} events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.
 To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
@@
 This is derived in Appendix \ref{app:selection1}.
-{\blue To evaluate the effect of our selection effects, we need to evaluate this integral.
+\added{ To evaluate the effect of our selection effects, we need to evaluate both the selection effect terms in the numerator and the integral in the denominator.
+The numerator represents the probability we caught the supernova and it was selected into the cosmology sample.
 The integral represents our global selection efficiency at a location in parameter space, rather than the probability of our data being selected into our sample.} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
@@
 and is incorporated into our likelihood.
-{\blue Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}.
+\added{ Note that the above efficiencies utilise the common form of the normal distribution rather than the conditional probability notation found previously in this work.} This is illustrated in Figure \ref{fig:efficiency}.
 Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency.
@@
 	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context.
-The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ {\blue shown to recover input cosmology.
+The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ \added{ shown to recover input cosmology.
 For contours that are non-Gaussian due to the curved degeneracy between $\Omega_m$ and $w$, the marginalised distributions can provide misleading statistics where maximum marginalised distribution can disagree with the maximum likelihood in multiple dimensions.
@@
 For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.33}$.
-{\blue For the low redshift survey, this is equivalent to sampling $y = z$ from $0.02$ to $\sqrt{0.1}$, for the high redshift survey, this is equivalent to sampling $y = z^2$ from $0.2$ to $1.0$.
+\added{ For the low redshift survey, this is equivalent to sampling $y = z$ from $0.02$ to $\sqrt{0.1}$, for the high redshift survey, this is equivalent to sampling $y = z^2$ from $0.2$ to $1.0$.
 These distributions are arbitrary, and this test has been performed with various flat and power law distributions.}
@@
 The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.
-As we are using 100 independent realisations, the precision of our determination of the mean simulation result is {\blue approximately a tenth of the quoted scatter (as a degree of non-Gaussianity of our fits will make this relationship inexact)}.
+As we are using 100 independent realisations, the precision of our determination of the mean simulation result is \added{ approximately a tenth of the quoted scatter (as a degree of non-Gaussianity of our fits will make this relationship inexact)}.
 The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model.
@@
 Simulation specifics can be found in \citet{Kessler18}.
-{\blue The primary difference between the toy data of the previous section are the different underlying colour and stretch, inclusion of spectroscopic data cuts and light curve cuts, and inclusion of intrinsic dispersion models.}
-
-{\blue The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
+\added{ The primary difference between the toy data of the previous section are the different underlying colour and stretch, inclusion of spectroscopic data cuts and light curve cuts, and inclusion of intrinsic dispersion models.}
+
+\added{ The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
@@
 We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the functional form of the Malmquist bias between the two models.
-{\blue Uncertainty on the analytic selection function is incorporated into our fits, mitigating the imperfection of our analytic form by allowing it to vary in our fits.}
+\added{ Uncertainty on the analytic selection function is incorporated into our fits, mitigating the imperfection of our analytic form by allowing it to vary in our fits.}
 
 \begin{deluxetable}{lcccc}
+	\renewcommand{\arraystretch}{1.2}
 \tablewidth{\columnwidth}
-
+	\label{tab:dist}
+	\tablecomments{The {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
 \tablecolumns{5}
@@
 	\startdata
-		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\ \\
+		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
 		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$  \\
 	\enddata
-	\label{tab:dist}
-	\tablecomments{The {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
 \end{deluxetable}
@@
 \begin{deluxetable*}{lrrrrr}
+	\label{tab:standardisation}
 	\tablewidth{\textwidth}
@@
 		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$ \\
-
 	\enddata
-	\label{tab:standardisation}
 \end{deluxetable*}
@@
 \begin{deluxetable}{lrr}
+	\label{tab:bulk_summary}
 	\tablecaption{Realistic simulation determination of $w$}
@@
 	\enddata
-		\label{tab:bulk_summary}
 \end{deluxetable}
@@
 \begin{deluxetable}{crr}
-
+	\label{tab:parameter_correlations}
+	\renewcommand{\arraystretch}{0.6}
 	\tablecaption{Reduced parameter correlations with $w$.}
@@
 Zero superscripts indicate the Dark Energy Survey, and a superscript one represents the low-$z$ survey.}
-	\tablehead{		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst }
+	\tablehead{	\\	Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst }
 	\startdata
@@
 	\enddata
-	\label{tab:parameter_correlations}
 \end{deluxetable}
@@
 Our second test aims to test our model whilst allowing biases in the summary statistics not caused from intrinsic scatter through.
-{\blue That is, the first test ascertained that biases in the summary statistics are the cause of cosmological bias.
+\added{ That is, the first test ascertained that biases in the summary statistics are the cause of cosmological bias.
 It is thus important to determine the source of those biases; whether they are from intrinsic scatter model or another aspect of the simulation.
 To} this end, we test a set of 100 simulations generated using an intrinsic dispersion model of only coherent magnitude scatter.
-{\blue We find $w=-1.00$, showing} that the source of the biases in summary statistics is the underlying intrinsic scatter model.
+\added{ We find $w=-1.00$, showing} that the source of the biases in summary statistics is the underlying intrinsic scatter model.
 From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased.
@@
 Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters.
-{\blue To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation.
+\added{ To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation.
 These fits did not show reduced cosmological bias in the {\celeven} simulations, and thus the determination of underlying population in our model may give biased distributions, however does not impact cosmology.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
@@
 \begin{deluxetable*}{llrr}
+	\label{tab:uncert}
 	\tablecaption{$w$ error budget}
@@
 	\enddata
-	\label{tab:uncert}
 \end{deluxetable*}
@@
 With the increased flexibility of Bayesian hierarchical models over traditional models, we expect to find an increased uncertainty on parameter inference.
-{\blue This increased uncertainty is one of the strengths of hierarchical models as it represents a more thorough accounting of model uncertainty.} To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
+\added{This increased uncertainty is one of the strengths of hierarchical models as it represents a more thorough accounting of model uncertainty.} To characterise the influence of the extra degrees of freedom in our model, we analyse the uncertainty on $w$ averaged across 10 nominal simulations of the DES-SN3YR sample with various model parameters allowed to either vary or stay locked to a fixed value.
 By taking the difference in uncertainty in quadrature, we can infer the relative contribution for each model feature to the uncertainty error budget.
@@
 \end{eqnarray}
-which is equal to the likelihood $\mathcal{L}$. Introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
+which is equal to the likelihood $\mathcal{L}$.
+\added{ At this point, our derivation depends on whether our selection effects are best modelled as a function of observables or as a function of latent parameters.
+In most cases, selection effects will be best modelled directly as a function of observables, and we would introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$,
 \begin{eqnarray}
@@
 where we define the denominator as $d$ for simplicity in future derivations.
+The second option, wherein selection effects can be more efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
+\begin{eqnarray}
+\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S| L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|D,L,\theta)P(D|L,\theta) P(L|\theta)\, dD\, dL},\label{eq:latent}
+\end{eqnarray}
+where $L$ represents our latent parameters that model the true underlying values of our observables, such that our data is conditioned directly on them.
+}
 
@@
 \label{app:selection2}
-{\blue
-To turn the generalised equations from the previous sections into selection effects relevant for this model, we first need to write out our latent parameters, as $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc).
-Whilst they can be inserted into the denominator of equation \eqref{eq:general1} to get $P(S|\theta) = \int P(S,L|\theta)\, dL$ and expanded out, we can simply write down our denominator with general latent parameter(s) $L$, as such:
+\added{
+To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters.
+We thus continue from equation \eqref{eq:latent} and write out denominator
 \begin{equation}
@@
 Here we can now move from the generic labels $D$ and $L$ to something specific to our model.
-In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ However, as we assume that our selection effects can be sufficiently encapsulated by true apparent magnitude $m_B$ and redshift $z$, we will not write out the other parameters for simplicity.
+In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ However, as we assume that our selection effects can be sufficiently encapsulated by latent parameters.
+That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$.
+Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity.
 If we did, we'd simply find them integrated to unity.
 Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
-Our denominator then becomes
+Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
 \begin{equation}

### 2018-12-29 papers/methods/figures/tikz.tex
Updating paper based on referee comments

@@
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[det, right=4.33cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
+	\node[latent, right=4.1cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
 	% Survey nodes
@@
 	\node[latent, above=2.3cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.7cm of MBobs, text width=2.5cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$};
-	\node[latent, right=0.5cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
+	\node[latent, above=2.3cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, \\$\delta(0)$, $\delta(\infty)$};
+	\node[latent, right=0.7cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
 	% Connect the nodes

### 2018-12-29 papers/methods/paper.tex
Updating paper based on referee comments

@@
 As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement \added{as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus.
+Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement \added{as varying redshifts introduce computational complexity in computing the distance modulus integral by reducing the amount of pre-computation that can be utilised}) but instead uncertainty on distance modulus.
 Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate.
@@
 This gives thirteen global systematics shared by \added{ all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey.
-With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
+\added{Systematics with known correlations are shifted together to produce covariant deviations, and we thus assume that the numerical derivatives input into our model represent independent systematics.
+Full details can be found in \citet{Brout18SYS}.} With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
 \begin{equation}
@@
 
-The selection functions parameters (a skew normal for low-redshift and a complimentary error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
+The selection functions parameters (a skew normal for low-redshift and a \replaced{complimentary}{complementary} error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function).
 Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample.
@@
 
-\added{ The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
+Prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach.
 In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.
@@
 \added{ To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation.
-These fits did not show reduced cosmological bias in the {\celeven} simulations, and thus the determination of underlying population in our model may give biased distributions, however does not impact cosmology.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
+These fits did not change the bias in the {\celeven} simulation.
+We conclude that the biased population recovery is not the cause of cosmological bias.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.
 
@@
 \mathcal{L}(\theta; {\rm data})&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
-\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
+\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}. \label{eq:app_main}
 \end{eqnarray}
-where we define the denominator as $d$ for simplicity in future derivations.
 The second option, wherein selection effects can be more efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
@@
 To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters.
-We thus continue from equation \eqref{eq:latent} and write out denominator
-\begin{equation}
-d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL
+We thus continue from equation \eqref{eq:latent} and write out denominator $d$:
+\begin{equation}
+d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL.
 \end{equation}
 Here we can now move from the generic labels $D$ and $L$ to something specific to our model.
-In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ However, as we assume that our selection effects can be sufficiently encapsulated by latent parameters.
+In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters.
 That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$.
 Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity.
-If we did, we'd simply find them integrated to unity.
+If we did, we would simply find them integrated to unity.
 Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}

### 2019-01-04 papers/methods/figures/tikz.tex
Updating tikz

@@
 	% Obs nodes
-	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
+	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \hat{\mathcal{C}}_\eta^{-1}$};
 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
@@
 	% Survey nodes
-	\node[latent, above=of sum, yshift=1.6cm, xshift=1cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
-	\node[latent, right=1.2cm of pop1, align=center] (sf) {$\delta S$};
+	\node[latent, above=of sum, yshift=1.6cm, xshift=1.4cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
+
 	\node[latent, left=0.8cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
@@
 	\node[latent, above=2.3cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, \\$\delta(0)$, $\delta(\infty)$};
-	\node[latent, right=0.7cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
+	\node[latent, right=0.7cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i, \delta S$};
 	% Connect the nodes
@@
 	\edge {z,cosmo} {mu};
-	\edge {mu, pop1, ab, sf} {MBobs};	
+	\edge {mu, pop1, ab} {MBobs};	
 	\edge {MBobs} {pobs, mobs};
@@
 	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)(MBobs)} {$\forall\ \rm{SN}$} ;
-	\plated[thick] {survey} {(sn)(pop1)(kappa)(sf)} {\\$\forall\ \rm{Survey}$} ;
+	\plated[thick] {survey} {(sn)(pop1)(kappa)} {\\$\forall\ \rm{Survey}$} ;
 	\end{tikzpicture}

### 2019-01-28 papers/methods/paper.tex
Updating tikz
Alex's suggestions

@@
 The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both the observed and underlying populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.
-In the following sections, we will describe these layers, mapping functions, and occurrences of these fatal pathologies.
-Summaries of observables and model parameters are shown in Table \ref{tab:param_summary} for easy reference.
+
+\added{In the analysis of this article, the underlying model universe maps to the observed universe as sketched in the BHM of Figure \ref{fig:pgm}.
+The dependencies that between the model and observations can be tracked following the arrows of the BHM, and a summary of all the conditional probabilities can be found in \ref{sec:model_summary}.
+
+In the following sections, we will describe the model parameters, the mapping functions that connect them to data, the effect of sample selection (in Equation \eqref{eq:main}), and the pathologies that can occur when evaluating the model.
+Summaries of observables and model parameters are shown in Table 1 for easy reference.}
 
@@
  	\end{center}
- 	\caption{Probabilistic graphical model for our likelihood.
+ 	\caption{Probabilistic graphical model for our likelihood \added{and selection effects}.
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
@@
 \end{figure}
+
+\added{
+\subsection{Model Summary}
+\label{sec:model_summary}
+
+Having laid out each individual aspect of the model, the relationships between variables and our treatment of uncertainty, here we summarise the relationships in our model likelihood mathematically.
+In this likelihood summary, we leave out sample selection for simplicity.
+Our likelihood is thus given as:
+
+\begin{align}
+&P({\rm data}|\theta) = \notag \\
+&\int\, d\mu\, d\eta \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
+& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta) \notag \\
+& P(\mu |z, \Omega_m, w) 
+\end{align}
+}
 
@@
 If we did, we would simply find them integrated to unity.
+Note that this assumption does not capture biases engendered by
+Poisson noise fluctuations.  
+We advocate that future analyses with
+higher statistical precision do use a precisely determined $P(S|D)$.
 Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}

### 2019-01-29 papers/methods/paper.tex
Elaborating on sec 4.5

@@
 \begin{equation}
-P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
+P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}), \label{eq:selcdf}
 \end{equation}
@@
 \begin{equation}
-P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
+P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}). \label{eq:selnorm}
 \end{equation}
@@
 d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
-&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
+&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-z}
 \end{eqnarray}
@@
 The survey efficiency, given the SN~Ia population, is shown as a solid line in both panels, and the probability of observing a SN~Ia is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated.
-This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.}
+This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellow-z}.}
 	\label{fig:efficiency}
@@
 
-Having laid out each individual aspect of the model, the relationships between variables and our treatment of uncertainty, here we summarise the relationships in our model likelihood mathematically.
-In this likelihood summary, we leave out sample selection for simplicity.
-Our likelihood is thus given as:
+Having laid out each individual aspect of the model, the relationships between variables and our treatment of uncertainty, here we summarise the relationships in our model mathematically.
+In this summary of $P({\rm data} | \theta)$, we leave out sample selection for simplicity.
+Referring to  equation \eqref{eq:main}, the relationships with $P({\rm data} | \theta)$ are as follows:
 
@@
 &P({\rm data}|\theta) = \notag \\
-&\int\, d\mu\, d\eta \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
+&\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
 & P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta) \notag \\
@@
 \end{align}
+
+The denominator of equation \eqref{eq:main} is then given by either equation \eqref{eq:seldes} or \eqref{eq:sellow-z} depending on the survey, and similarly $P(S|{\rm data}, \theta)$ is given by equation \eqref{eq:selnorm} or \eqref{eq:selcdf} respectively.
+Combining all of these gives us our full model likelihood with selection effects accounted for.
+
 }
@@
 \end{eqnarray}
-Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-$z$}.
+Then, modelling $P(S|m_B)$ as either a normal or a skew normal, we can analytically perform the integral in equation \eqref{eq:selint} and reach equations \eqref{eq:seldes} and \eqref{eq:sellow-z}.
 

### 2019-01-31 papers/methods/figures/tikz.tex
More figure changes

@@
 	% Latent nodes
-	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
+	\node[det, above=of obssum, xshift=0.5cm] (sum) {$m_B, x_1, c$};
 	\node[det, above=of zobs] (z) {$z$};
+	\node[det, above=of mobs] (m) {$m$};
+	\node[det, above=of pobs, yshift=0.449cm] (p) {$p$};	
 	\node[det, above=of z] (mu) {$\mu$};
@@
 	% Survey nodes
-	\node[latent, above=of sum, yshift=1.6cm, xshift=1.4cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
+	\node[latent, above=of sum, yshift=0.3cm, xshift=1.4cm, text width=1.85cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
 
-	\node[latent, left=0.8cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
+	\node[latent, left=1.3cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
 	% Global nodes
-	\node[latent, above=2.3cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=2.3cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, \\$\delta(0)$, $\delta(\infty)$};
-	\node[latent, right=0.7cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i, \delta S$};
+	\node[latent, above=2.1cm of mu]  (cosmo) {$\Omega_m, w$};
+	\node[latent, above=2.1cm of MBobs, xshift=0.8cm, align=center]  (ab) {$\langle M_B \rangle$};
+	\node[latent, right=0.1cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
+	\node[latent, right=0.1cm of dz, align=center] (dd) {$\alpha, \beta, \delta(0)$, \\$\delta(\infty)$};
 	% Connect the nodes
-	\edge {sum} {obssum};
+	\edge {sum, dz, kappa} {obssum};
 	\edge {z} {zobs};
+	\edge {m} {mobs};
+	\edge {p} {pobs};
 	\edge {z,cosmo} {mu};
-	\edge {mu, pop1, ab} {MBobs};	
-	\edge {MBobs} {pobs, mobs};
-	\edge {dz, kappa, MBobs} {sum};
+	\edge {pop1, ab} {MBobs};	
+	\edge {mu, MBobs, m, p, dd} {sum};
+
 	% Plates

### 2019-01-31 papers/methods/paper.tex
More figure changes

@@
  	\end{center}
- 	\caption{Probabilistic graphical model for our likelihood \added{and selection effects}.
+ 	\caption{Probabilistic graphical model for our likelihood \added{without selection effects}.
 Double-lined nodes represent observed variables, diamond nodes represent deterministic variables, single-lined ellipse nodes represent fit variables.
@@
 Variables that appear outside both boxes represent top level model parameters.
-\added{To solidify that the underlying supernova population lives in $\lbrace M_B, x_1, c\rbrace$ space, we write out the stretch and colour populations along with $M_B$, however note here that those $x_1$ and $c$ values (denoted by an asterix) are identically equal to the $x_1$ and $c$ values in the latent node they are conditioned on, and in the code implementation, they are the same variable.}}
+\added{We note that we have shown the model to have latent variables $\lbrace M_B,\,x_1\,c\rbrace$, which uniquely determines $\lbrace m_B,\,x_1\,c\rbrace$, given $\mu$ and other parameters.
+Thus, conceptually the two nodes make up a single layer in our hierarchy, not two layers.
+In the code implementation, $m_B$ is more efficient parametrise instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other.}}
  	\label{fig:pgm}

### 2019-02-01 papers/methods/figures/tikz.tex
More figure changes

@@
 	% Latent nodes
-	\node[det, above=of obssum, xshift=0.5cm] (sum) {$m_B, x_1, c$};
+	\node[det, above=of obssum, xshift=1.2cm] (mb) {$m_B$};
+	\node[latent, left=of mb, xshift=0.1cm] (x1c) {$x_1, c$};
 	\node[det, above=of zobs] (z) {$z$};
-	\node[det, above=of mobs] (m) {$m$};
-	\node[det, above=of pobs, yshift=0.449cm] (p) {$p$};	
+	\node[det, above=of mobs, yshift=-0.5cm] (m) {$m$};
+	\node[det, above=of pobs, yshift=0.5cm] (p) {$p$};	
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[latent, right=4.1cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
+	\node[latent, above=of p, yshift=-0.5cm] (MB) {$M_B$};
 	% Survey nodes
-	\node[latent, above=of sum, yshift=0.3cm, xshift=1.4cm, text width=1.85cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
+	\node[latent, above=of x1c, yshift=0.9cm, xshift=-0.65cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
 
-	\node[latent, left=1.3cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
+	\node[latent, right=0.2cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
 	% Global nodes
 	\node[latent, above=2.1cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=2.1cm of MBobs, xshift=0.8cm, align=center]  (ab) {$\langle M_B \rangle$};
-	\node[latent, right=0.1cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
-	\node[latent, right=0.1cm of dz, align=center] (dd) {$\alpha, \beta, \delta(0)$, \\$\delta(\infty)$};
+	\node[latent, above=1.8cm of MB, xshift=0cm, align=center]  (MBpop) {$\langle M_B \rangle,$ \\ $\sigma_{M_B}$};
+	\node[latent, right=1.43cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
+	\node[latent, right=0.1cm of dz, align=center] (ab) {$\alpha, \beta, \delta(0)$, \\$\delta(\infty)$};
 	% Connect the nodes
-	\edge {sum, dz, kappa} {obssum};
+	\edge {mb, x1c, dz, kappa} {obssum};
 	\edge {z} {zobs};
@@
 	\edge {z,cosmo} {mu};
-	\edge {pop1, ab} {MBobs};	
-	\edge {mu, MBobs, m, p, dd} {sum};
+	\edge {MBpop} {MB};	
+	\edge {MB, mu, ab, m, p, x1c} {mb}
+	\edge {pop1, z} {x1c};
 
 	% Plates
-	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)(MBobs)} {$\forall\ \rm{SN}$} ;
+	\plated[thick] {sn} {(mb)(obssum)(zobs)(mobs)(pobs)(x1c)(z)(mu)(MB)(m)(p)} {$\forall\ \rm{SN}$} ;
 	\plated[thick] {survey} {(sn)(pop1)(kappa)} {\\$\forall\ \rm{Survey}$} ;

### 2019-02-01 papers/methods/paper.tex
More figure changes

@@
 Variables that appear outside both boxes represent top level model parameters.
-\added{We note that we have shown the model to have latent variables $\lbrace M_B,\,x_1\,c\rbrace$, which uniquely determines $\lbrace m_B,\,x_1\,c\rbrace$, given $\mu$ and other parameters.
-Thus, conceptually the two nodes make up a single layer in our hierarchy, not two layers.
-In the code implementation, $m_B$ is more efficient parametrise instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other.}}
+\added{We note that we have shown the model to have latent variables $\lbrace M_B,\,x_1\,c\rbrace$, which uniquely determines $m_B$, given $\mu$ and other parameters.
+Thus, the two nodes $M_B$ and $m_B$, make up a single layer in our hierarchy, not two layers.
+In the code implementation, $m_B$ is more efficiently parametrised instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other.
+Finally, as we take redshift measurement $\hat{z}$, mass probability $\hat{m}$ and classification probability $\hat{p}$ as exact they are not conditioned on underlying distributions and are top level parameters.
+We add the deterministic $m$ and $p$ nodes in our diagram for clarity, however as they have a one-to-one mapping onto $\hat{m}$ and $\hat{p}$ respectively we simply use the observed values in the mathematics of the rest of the paper.}}
  	\label{fig:pgm}

### 2019-02-05 papers/methods/paper.tex
Updating figure again

@@
 In the code implementation, $m_B$ is more efficiently parametrised instead of $M_B$, however the mathematics remains constant regardless of if you parametrise $M_B$ or $m_B$ as one can determine the other.
-Finally, as we take redshift measurement $\hat{z}$, mass probability $\hat{m}$ and classification probability $\hat{p}$ as exact they are not conditioned on underlying distributions and are top level parameters.
-We add the deterministic $m$ and $p$ nodes in our diagram for clarity, however as they have a one-to-one mapping onto $\hat{m}$ and $\hat{p}$ respectively we simply use the observed values in the mathematics of the rest of the paper.}}
+We write out ${\rm mass}$ instead of $m$ to reduce possible confusion with magnitude in the diagram.
+Finally, as we take redshift measurement $\hat{z}$, mass probability $\hat{{\rm mass}}$ and classification probability $\hat{p}$ as exact they are not conditioned on underlying distributions and are top level parameters.
+We add the deterministic $m$ and $p$ nodes in our diagram for clarity, however as they have a one-to-one mapping onto $\hat{{\rm mass}}$ and $\hat{p}$ respectively we simply use the observed values in the mathematics of the rest of the paper.}}
  	\label{fig:pgm}
@@
 &\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
-& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta) \notag \\
+& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z) \notag \\
 & P(\mu |z, \Omega_m, w) 

### 2019-02-07 papers/methods/paper.tex
Final update hopefully
Ading more text to eqn 4.2.2 for alex

@@
 The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
-For the spectroscopic survey, we set this to unity.
+\added{Thus our likelihood adapts by having probability $\hat{p}$ to fall into the SN~Ia population as given by equation \eqref{eq:l1} and probability $1 - \hat{p}$ to fall into the outlier population, which is also Gaussian with same mean but unitary covariance.}For the spectroscopic survey, we set this to unity.
 For the photometric proof-of-concept we provide an accurate probability estimate.
@@
 \begin{align}
-&P({\rm data}|\theta) = \notag \\
+&P({\rm data}|\theta) \propto \notag \\
 &\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
-& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z) \notag \\
-& P(\mu |z, \Omega_m, w) 
+& P(\eta | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z, \hat{m}, \hat{p}) \notag \\
+& P(\mu |z, \Omega_m, w),
 \end{align}
+where $\hat{m}$ and $\hat{p}$ are found on the right-hand-side of the conditional probability because we take them as exact measurements and not dependent on a prior distribution.
 

### 2019-02-08 papers/methods/figures/tikz.tex
Updating figure again
Removing p

@@
 	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
-	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
-	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
+	\node[obs, right=0.9cm of obssum]  (mobs) {$\hat{{\rm mass}}$};
+
 	% Latent nodes
-	\node[det, above=of obssum, xshift=1.2cm] (mb) {$m_B$};
-	\node[latent, left=of mb, xshift=0.1cm] (x1c) {$x_1, c$};
+	\node[det, above=of obssum, xshift=1.5cm] (mb) {$m_B$};
+	\node[latent, left=of mb, xshift=-0.2cm] (x1c) {$x_1, c$};
 	\node[det, above=of zobs] (z) {$z$};
-	\node[det, above=of mobs, yshift=-0.5cm] (m) {$m$};
-	\node[det, above=of pobs, yshift=0.5cm] (p) {$p$};	
+	\node[det, above=of mobs, yshift=-0.5cm] (m) {${\rm mass}$};
+
 	\node[det, above=of z] (mu) {$\mu$};
-	\node[latent, above=of p, yshift=-0.5cm] (MB) {$M_B$};
+	\node[latent, above=of m, xshift=0.8cm, yshift=-0.5cm] (MB) {$M_B$};
 	% Survey nodes
@@
 	\node[latent, above=2.1cm of mu]  (cosmo) {$\Omega_m, w$};
-	\node[latent, above=1.8cm of MB, xshift=0cm, align=center]  (MBpop) {$\langle M_B \rangle,$ \\ $\sigma_{M_B}$};
+	\node[latent, above=2.5cm of MB, xshift=0cm, align=center]  (MBpop) {$\langle M_B \rangle,$ \\ $\sigma_{M_B}$};
 	\node[latent, right=1.43cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
-	\node[latent, right=0.1cm of dz, align=center] (ab) {$\alpha, \beta, \delta(0)$, \\$\delta(\infty)$};
+	\node[latent, right=0.1cm of dz, align=center] (ab) {$\alpha, \beta,$\\ $\delta(0),\delta(\infty)$};
 	% Connect the nodes
@@
 	\edge {m} {mobs};
-	\edge {p} {pobs};
 	\edge {z,cosmo} {mu};
 	\edge {MBpop} {MB};	
-	\edge {MB, mu, ab, m, p, x1c} {mb}
+	\edge {MB, mu, ab, m, x1c} {mb}
 	\edge {pop1, z} {x1c};
@@
 	% Plates
-	\plated[thick] {sn} {(mb)(obssum)(zobs)(mobs)(pobs)(x1c)(z)(mu)(MB)(m)(p)} {$\forall\ \rm{SN}$} ;
+	\plated[thick] {sn} {(mb)(obssum)(zobs)(mobs)(x1c)(z)(mu)(MB)(m)} {$\forall\ \rm{SN}$} ;
 	\plated[thick] {survey} {(sn)(pop1)(kappa)} {\\$\forall\ \rm{Survey}$} ;

### 2019-02-08 papers/methods/paper.tex
Removing p

--- 2019-02-05 papers/methods/paper.tex
@@
 Additionally, each supernova has an observed redshift $\hat{z}$ and a host-galaxy stellar mass associated with it, $\hat{m}$, where the mass measurement is converted into a probability of being above $10^{10}$ solar masses.
-We also have a probability of each supernova being a type Ia, $\hat{p}$.
-Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{p}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
+Our set of observables input into the {\steve} is therefore given as $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \cov \rbrace$, as shown in the probabilistic graphical model (PGM) in Figure \ref{fig:pgm}.
 
@@
 	$\hat{z}$ & Observed redshift \\
-	$\hat{m}$ & Determined mass probability \\
-	$\hat{p}$ & Determined Ia probability \\
+	$\hat{m}$ & Observed mass probability \\
 	\enddata
@@
 We write out ${\rm mass}$ instead of $m$ to reduce possible confusion with magnitude in the diagram.
-Finally, as we take redshift measurement $\hat{z}$, mass probability $\hat{{\rm mass}}$ and classification probability $\hat{p}$ as exact they are not conditioned on underlying distributions and are top level parameters.
-We add the deterministic $m$ and $p$ nodes in our diagram for clarity, however as they have a one-to-one mapping onto $\hat{{\rm mass}}$ and $\hat{p}$ respectively we simply use the observed values in the mathematics of the rest of the paper.}}
+Finally, as we take redshift measurement $\hat{z}$ and mass probability $\hat{{\rm mass}}$ as exact they are not conditioned on underlying distributions and are top level parameters.}}
  	\label{fig:pgm}
@@
 
-For future use with photometrically classified surveys, we include a simplistic outlier population model.
-We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model.
+\added{For the spectroscopic paper we do not consider outlier populations, however we ensure that our model has flexibility for such populations for future use with photometrically classified surveys.
+We thus include a simplistic outlier population model.} We follow {\rubin} \citep[and therefore ][]{Kunz2007} by implementing a Gaussian mixture model, where an additional observable of the SN~Ia probability would be needed in order to inform the weights of the mixture model.
 For surveys with low rates of contamination, it is not possible to fit a contamination population, and the mean of the outlier population has been fixed to the SN~Ia population in prior works.
@@
 We note that this represents the mean brightness of outliers, and so outliers could both be brighter and dimmer than the mean SN~Ia absolute magnitude.
-We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$.
-The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$ discussed in the previous section.
-For the spectroscopic survey, we set this to unity.
+We set the population width to $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$ in our tests.
+The probability of each supernova falling into either population is determined by the observed type Ia probability $\hat{p}$.
+For the spectroscopic survey, we set this to unity, and thus it is not included in Figure \ref{fig:pgm} or Table \ref{tab:param_summary}.
 For the photometric proof-of-concept we provide an accurate probability estimate.
@@
 We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.
+\added{Finally, we assume that the observed mass probability $\hat{m}$ supplied to the model is perfectly determined, and thus set $P(\hat{m}|m) = \delta(\hat{m} - m)$.}
 
@@
 \begin{align}
-&P({\rm data}|\theta) = \notag \\
-&\int\, d\mu\, d\eta\, dz \ P( \hat{\eta}, \hat{\mathcal{C}}_\eta^{-1} | \eta, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) P(\hat{z}|z) \notag \\
-& P(\eta, \hat{m}, \hat{p} | \langle M_B \rangle, \langle x_1^i \rangle, \langle c^i \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z) \notag \\
-& P(\mu |z, \Omega_m, w) 
+&P({\rm data}|\theta) \propto \notag \\
+&\int\, d\mu\, dm_B\, dx_1\, dc\, dz, dm \  P(\hat{z}|z) P(\hat{m}|m)  \notag \\
+& P(\hat{m}_B, \hat{x}_1, \hat{c}, \hat{\mathcal{C}}_\eta^{-1} | m_B, x_1, c, \delta \mathcal{Z}_i, \kappa_0, \kappa_1) \notag \\
+& P(x_1, c |  \langle x_1^i \rangle, \langle c^i \rangle, \sigma_{x_1}, \sigma_c, z) \notag\\
+& P(m_B | \langle M_B \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z, m, p) \notag \\
+& P(\mu |z, \Omega_m, w).
 \end{align}
@@
 Here we can now move from the generic labels $D$ and $L$ to something specific to our model.
-In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters.
+In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m} \rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters.
 That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$.

### 2019-02-08 papers/methods/paper.tex
Removing p

@@
 & P(x_1, c |  \langle x_1^i \rangle, \langle c^i \rangle, \sigma_{x_1}, \sigma_c, z) \notag\\
-& P(m_B | \langle M_B \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z, m, p) \notag \\
+& P(m_B | \langle M_B \rangle, \mu, \delta(0), \delta(\infty), \sigma_{M_B}, \sigma_{x_1}, \sigma_c, \alpha, \beta, z, m) \notag \\
 & P(\mu |z, \Omega_m, w).

### 2019-03-11 papers/methods/paper.tex
Adding 'computationally'

@@
 \begin{align}
-\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
+\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, 
 \end{align}
-where the denominator represents an integral over all potential data $D$.
+where the denominator represents an integral over all potential data $D$, \added{and $\theta$ represents top-level parameters.
+In the case that our selection effects are best charactersied by latent variables \textit{instead of} data, we can add them to our formulation and our likelihood becomes
+\begin{eqnarray}
+\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S|L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|L,\theta) P(L|\theta) dL},\label{eq:main}
+\end{eqnarray}
+where $L$ represents our latent parameters.}
 This is derived in Appendix \ref{app:selection1}.
@@
 The numerator represents the probability we caught the supernova and it was selected into the cosmology sample.
-The integral represents our global selection efficiency at a location in parameter space, rather than the probability of our data being selected into our sample.} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral.
-Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
+The integral represents our global selection efficiency at a location in parameter space, rather than the probability of our data being selected into our sample.} As $\theta$ represents the vector of all \added{top-level model parameters, and $L$ represents a vector of all latent parameters, this is not a trivial integral.} Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle).
 We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as
@@
 \end{eqnarray}
-The second option, wherein selection effects can be more efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
+The second option, wherein selection effects can be more computationally efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
 \begin{eqnarray}
-\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S| L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|D,L,\theta)P(D|L,\theta) P(L|\theta)\, dD\, dL},\label{eq:latent}
+\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S|{\rm data}, L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|D,L,\theta)P(D|L,\theta) P(L|\theta)\, dD\, dL},\label{eq:latent}
 \end{eqnarray}
 where $L$ represents our latent parameters that model the true underlying values of our observables, such that our data is conditioned directly on them.
+In this formulation, the selection effects can depend both on data and latent variables.
 }
@@
 \end{equation}
-Here we can now move from the generic labels $D$ and $L$ to something specific to our model.
-In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m} \rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters.
+In our formulation, we assume that our selection effects can be sufficiently encapsulated by latent parameters.
 That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$.
-Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity.
-If we did, we would simply find them integrated to unity.
+This allows us to isolate our integral $\int P(D|L,\theta)$ in the above equation, integrate it to unity, and come to
+\begin{equation}
+d = \int  P(S|L) P(L | \theta)\, dL.
+\end{equation}	
+Here we can now move from the generic label $L$ to something specific to our model.
+Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and so $L \rightarrow \lbrace m_B, z\rbrace$, or formally $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.
+We do not write out other latent parameters found in the model as they do not impact selection effects --- we would simply find them integrated to unity.
 Note that this assumption does not capture biases engendered by
@@
 higher statistical precision do use a precisely determined $P(S|D)$.
-Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
-Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
-\begin{equation}
-d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
-\end{equation}
-Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$.
-$P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$.
-Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.
-We can thus state
+Writing out the latent variables, our denominator becomes
+\begin{equation}
+d = \int P(S|z) P(S|m_B)P(z, m_B | \theta)\, d\hat{z} \, dz \, dm_B.  \label{eq:sel} 
+\end{equation}
+We can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates.
+We can thus state}
 \begin{equation}

### 2019-03-26 papers/methods/paper.tex
Fixing two typos

@@
 where the denominator represents an integral over all potential data $D$, \added{and $\theta$ represents top-level parameters.
-In the case that our selection effects are best charactersied by latent variables \textit{instead of} data, we can add them to our formulation and our likelihood becomes
+In the case that our selection effects are best characterised by latent variables \textit{instead of} data, we can add them to our formulation and our likelihood becomes
 \begin{eqnarray}
@@
 \begin{equation}
-d = \int P(S|z) P(S|m_B)P(z, m_B | \theta)\, d\hat{z} \, dz \, dm_B.  \label{eq:sel} 
+d = \int P(S|z) P(S|m_B)P(z, m_B | \theta)\, dz \, dm_B.  \label{eq:sel} 
 \end{equation}
